### 2020-04-25 HBCC_sgACC_FEATURECOUNTS_GRCh38_CQN.Rmd
model selection hbcc sgACC cohort

---
title: "HBCC - sgACC - FEATURECOUNTS - GRCh38 - CQN"
author: "Kelsey Montgomery, Gabriel Hoffman and Thanneer Perumal"
date: "`r date()`"
output: html_document
---

```{r knit2synapse, eval=FALSE}
library(synapser)
library(knit2synapse)

synLogin()

knit2synapse::createAndKnitToFolderEntity(file = "./cmc-rnaseq-de/HBCC_sgACC_FEATURECOUNTS_GRCh38_CQN.Rmd",
                                          parentId ="syn21435176",
                                          folderName = 'HBCC - sgACC - DREAM - FEATURECOUNTS - GRCh38 - CQN')
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## It is assumed your working directory is where this file

## Load required libraries
library(CovariateAnalysis) # get the package from devtools::install_github('th1vairam/CovariateAnalysis@dev')
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)

library(ggplot2)
library(limma)
library(psych)
library(edgeR)
library(cqn)
library(lme4)
library(biomaRt)
library(ComplexHeatmap)
library(statmod)

library(variancePartition)
library(synapser)
library(knitr)
library(githubr) # get the package from devtools::install_github('brian-bot/githubr')

synLogin()

library(doParallel)
library(foreach)
library(future)

cl = makeCluster((availableCores()-2)/2)

registerDoParallel(cl)

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
parentId = 'syn21435176';
activityName = 'Covariate adjustments';
activityDescription = 'Covariate analysis of GRCh38 feature counts with CQN normalisation (sgACC)';

thisFileName <- 'HBCC_sgACC_FEATURECOUNTS_GRCh38_CQN.Rmd'

# Github link
thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=thisFileName)
```
### Data download
#### Obtain effective count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
downloadFile <- function(id){
  fread(synGet(id)$path, data.table = F)
}
downloadFile_version <- function(id , version){
  fread(synGet(id, version = version)$path, data.table = F)
}

# Get RNASeq QCmetadata
METADATA_QC_ACC_ID = 'syn16816488' 
ALL_USED_IDs = METADATA_QC_ACC_ID
metadata = downloadFile_version(METADATA_QC_ACC_ID, version = 20) %>%
  dplyr::select(`Individual_ID`, Institution, `Reported_Gender`, Sex, Ethnicity, ageOfDeath, `PMI_(in_hours)`, Dx, 
                Sample_RNA_ID, one_of('Cohort',
                                      'rnaSeq_isolation:RIN',
                                      'rnaSeq_dissection:Brain_Region',
                                      'rnaSeq_report:Ribozero_Batch', 'rnaSeq_report:Library_Batch', 
                                      'rnaSeq_isolation:RNA_Isolation_Batch', 'rnaSeq_report:Exclude?',
                                      'rnaSeq_report:Mapped_Reads',
                                      'rnaSeq_report:Intragenic_Rate','rnaSeq_report:Intronic_Rate',
                                      'rnaSeq_report:Intergenic_Rate','rnaSeq_report:Genes_Detected',
                                      'rnaSeq_report:Expression_Profiling_Efficiency',
                                      'rnaSeq_report:rRNA_Rate',
                                      'rnaSeq_report:Total_Reads', 'rnaSeq_report:Percent_Aligned',
                                      'rnaSeq_report:Transcripts_Detected')) %>% 
  dplyr::filter(`rnaSeq_dissection:Brain_Region` %in% c("sgACC")) %>% 
  dplyr::filter(Cohort %in% c("NIMH-HBCC-5"))

## Download reprocessed counts (ACC)
counts = 'syn21867762'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = counts
COUNT = downloadFile(counts) %>% data.frame()
COUNT$Chr <- NULL
COUNT$Start <- NULL
COUNT$End <- NULL
COUNT$Strand <- NULL

# Parse gene lengths
gene_lengths = COUNT[,c("Geneid", "Length")]
COUNT$Length <- NULL
gene_lengths <- dplyr::rename(gene_lengths, gene_id = Geneid)

# Get ancestry vector calculated using gemtools
ANCESTRY_ID = 'syn9922992'
ALL_USED_IDs = c(ALL_USED_IDs, ANCESTRY_ID)
ANCESTRY = downloadFile(ANCESTRY_ID) %>% 
  plyr::rename(c(ID = 'SNP_report:Genotyping_Sample_ID'))

# Get genotype ids from synapse
GENOTYPE_ID = 'syn16816490'
ALL_USED_IDs = c(ALL_USED_IDs, GENOTYPE_ID)
GENOTYPE = downloadFile(GENOTYPE_ID) %>% 
  dplyr::select(Individual_ID, `SNP_report:Genotyping_Sample_ID`, `SNP_report:Exclude?`) %>% 
  dplyr::inner_join(ANCESTRY) %>%
  dplyr::filter(is.na(`SNP_report:Exclude?`)) #Remove excluded samples here because there are duplicate individuals
```

### Data preprocessing
```{r preprocess.data}
# Merge metadata
METADATA = metadata %>%
  dplyr::left_join(GENOTYPE) %>% 
  dplyr::rename(Region = `rnaSeq_dissection:Brain_Region`,
                PMI = `PMI_(in_hours)`,
                RIN = `rnaSeq_isolation:RIN`, 
                ReportExclude = `rnaSeq_report:Exclude?`,
                GenotypeExclude = `SNP_report:Exclude?`,
                SampleID = Sample_RNA_ID,
                LibraryBatch = `rnaSeq_report:Library_Batch`,
                IsolationBatch = `rnaSeq_isolation:RNA_Isolation_Batch`,
                RibozeroBatch = `rnaSeq_report:Ribozero_Batch`,
                MappedReads = `rnaSeq_report:Mapped_Reads`,
                IntragenicRate = `rnaSeq_report:Intragenic_Rate`, IntronicRate = `rnaSeq_report:Intronic_Rate`, 
                IntergenicRate = `rnaSeq_report:Intergenic_Rate`, GenesDetected = `rnaSeq_report:Genes_Detected`,
                ExpProfEfficiency = `rnaSeq_report:Expression_Profiling_Efficiency`, rRNARate = `rnaSeq_report:rRNA_Rate`,
                TotalReads = `rnaSeq_report:Total_Reads`, AlignmentRate = `rnaSeq_report:Percent_Aligned`, TranscriptsDetected = `rnaSeq_report:Transcripts_Detected`) %>%
  dplyr::select(SampleID, Individual_ID, Institution, Reported_Gender,Sex, Ethnicity, ageOfDeath, PMI, Dx, 
                RIN, ReportExclude, GenotypeExclude, EV.1, EV.2, EV.3, EV.4, EV.5,
                LibraryBatch, IsolationBatch, RibozeroBatch, MappedReads, TotalReads, GenesDetected, AlignmentRate,
                IntragenicRate, IntergenicRate, IntronicRate, ExpProfEfficiency, rRNARate, TranscriptsDetected) %>%
  dplyr::mutate(Dx = forcats::fct_recode(Dx, Other = "BP", Other = "MDD", Control = "Control", SCZ = "SCZ"))

# Limit HBCC ACC Counts
rownames(COUNT) = COUNT$Geneid
COUNT <- COUNT[,colnames(COUNT) %in% METADATA$SampleID]

ind = METADATA$SampleID[which(METADATA$ReportExclude == 1 | METADATA$GenotypeExclude)]
writeLines(paste('Following',length(ind),'samples are marked exclude'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$Ethnicity) | is.na(METADATA$Institution) | is.na(METADATA$Dx)]
writeLines(paste('Following', length(ind), 'counts are missing any metadata'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$PMI)]
writeLines(paste('Following', length(ind), 'counts are missing PMI'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$Reported_Gender)]
writeLines(paste('Following', length(ind), 'counts are missing gender'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$ageOfDeath)]
writeLines(paste('Following', length(ind), 'counts are missing age of death'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$RIN)]
writeLines(paste('Following', length(ind), 'counts are missing RIN'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$EV.1)]
writeLines(paste('Following', length(ind), 'counts are missing ancestry information'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))
```

```{r preprocess.data1, results='asis'}
# Match covariates to expression data
indToRetain = intersect(METADATA$SampleID, colnames(COUNT))
removedIDs = setdiff(colnames(COUNT), METADATA$SampleID)

COUNT = COUNT[,indToRetain]

rownames(METADATA) = METADATA$SampleID
METADATA = METADATA[indToRetain,]

METADATA %>% 
  group_by(Dx) %>% 
  dplyr::summarise(count = n()) %>% 
  kable()
```
Following sample are removed `r paste(removedIDs, collapse = ',')`

```{r gene.param, include = F}
## Get GC content from biomart
backgroundGenes = data.frame(gene_id = rownames(COUNT)) %>% 
  dplyr::mutate(id = gene_id) %>% 
  tidyr::separate(id, c('ensembl_gene_id','position'), sep = '\\.')

# Define biomart object
mart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL", 
                host = "uswest.ensembl.org", # Ensembl Release 99 (January 2020)
                dataset = "hsapiens_gene_ensembl")

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol", "percentage_gene_gc_content", "gene_biotype", "chromosome_name"),
                       filters = "ensembl_gene_id", values = backgroundGenes$ensembl_gene_id,
                       mart = mart)

GENE.GC.CONT = Ensemble2HGNC %>%
  dplyr::left_join(backgroundGenes) %>% 
  dplyr::select(gene_id, percentage_gene_gc_content, chromosome_name) %>%
  unique
rownames(GENE.GC.CONT) = GENE.GC.CONT$gene_id
```
### Sex Chromosome-specific Gene Expression Patterns

```{r Reported_Gender_plots}
COUNT$gene_id = rownames(COUNT) 
REPORTED.GENDER.COUNTS = GENE.GC.CONT %>% 
  left_join(COUNT) %>%
  dplyr::select(-one_of("percentage_gene_gc_content")) %>%
  filter(chromosome_name == "X" |chromosome_name == "Y") %>% 
  tidyr::gather(key = item, value = value, -c(gene_id, chromosome_name)) %>%
  dplyr::mutate(value = log(value)) %>%
  dplyr::rename(`counts(log)`= value) %>% 
  dplyr::rename(SampleID = item) %>%
  left_join(METADATA[,c("SampleID", "Reported_Gender")]) %>% 
  dplyr::rename(`Reported Gender` = Reported_Gender)

my.theme <- theme_bw() %+replace% theme(legend.position = 'top', axis.text.x = element_text(angle = 90, hjust = 1), plot.title=element_text(hjust=0.5))

p = list()
p[[1]] = ggplot(filter(REPORTED.GENDER.COUNTS, chromosome_name == "X"), aes(x = `Reported Gender`, y = `counts(log)`)) + geom_boxplot()
p[[1]] = p[[1]] + ggtitle('X') + my.theme
p[[2]] = ggplot(filter(REPORTED.GENDER.COUNTS, chromosome_name == "Y"), aes(x = `Reported Gender`, y = `counts(log)`)) + geom_boxplot()
p[[2]] = p[[2]] + ggtitle('Y') + my.theme
multiplot(plotlist = p, cols = 2)

##XIST and UTY expression 
#ENSG00000229807.11 and ENSG00000183878.15 
FILT <- REPORTED.GENDER.COUNTS %>% 
  filter(gene_id == "ENSG00000229807.11" | gene_id == "ENSG00000183878.15") %>% 
  dplyr::select(-one_of("chromosome_name")) %>% 
  tidyr::spread(key = gene_id, value = `counts(log)`) %>% 
  mutate(XIST = as.numeric(`ENSG00000229807.11`)) %>% 
  mutate(UTY = as.numeric(`ENSG00000183878.15`)) %>% 
  mutate(UTY = ifelse(UTY == -Inf, 0, UTY)) %>% 
  mutate(XIST = ifelse(XIST == -Inf, 0, XIST))

p = ggplot(FILT, aes (x= XIST, y = UTY)) 
p = p + geom_point(aes(color=`Reported Gender`)) + 
  ggtitle("Sex Check HBCC Cohort") + 
  theme(plot.title = element_text(hjust = 0.5, size = 15)) +
  labs(colour = "Reported Gender")
p
```

### Covariates clustering
Determine relationship between covariates. 
```{r covariates.clustering}
FactorCovariates <- c('Individual_ID', "Reported_Gender", "LibraryBatch", "Dx", "IsolationBatch")
ContCovariates <- c("ageOfDeath", "PMI", "RIN", "MappedReads", "IntragenicRate", "IntronicRate", "IntergenicRate",
                    "GenesDetected", "ExpProfEfficiency", "rRNARate", "TotalReads", "AlignmentRate",
                    "EV.1", "EV.2", "EV.3", "EV.4", "EV.5")

# Find inter relation between factor covariates
COVARIATES = METADATA[,c(FactorCovariates,ContCovariates),drop=F]
COVARIATES[,FactorCovariates] <- data.frame(lapply(COVARIATES[,FactorCovariates],function(x){
  x <- sapply(x,function(y){str_replace_all(as.character(y),'[^[:alnum:]]','_')})}))
rownames(COVARIATES) <- METADATA$SampleID

# Convert factor covariates to factors
COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], function(x){
  x = as.numeric(as.character(gsub('[\\,\\%]','',x)))
})

# Add in RIN^2 values
COVARIATES$RIN2 = COVARIATES$RIN^2
ContCovariates = c(ContCovariates, 'RIN2')
```
Correlation/association between covariates at an FDR <= 0.1
```{r covariates.correlation, fig.width=10, fig.height=8}
COVARIATES.CORRELATION = CovariateAnalysis::getAssociationStatistics(COVARIATES, PVAL = 0.05)
tmp = COVARIATES.CORRELATION$ESTIMATE
tmp[COVARIATES.CORRELATION$PVAL > 0.05] = 0
h = Heatmap(tmp, col = colorRamp2(c(-1,0,1), c('blue','white','red')), name = 'AssocEstimate')
draw(h, heatmap_legend_side = 'left')
```

## Explore metatdata
```{r data.explore, fig.width = 12, fig.height = 12}
my.theme <- theme_bw() %+replace% theme(legend.position = 'top', axis.text.x = element_text(angle = 90, hjust = 1), plot.title=element_text(hjust=0.5))

# RIN
p = list()
p[[1]] = ggplot(COVARIATES, aes(x = Dx, y = RIN)) + geom_boxplot()
p[[1]] = p[[1]] + ggtitle('RIN') + my.theme

# Age of Death
p[[2]] = ggplot(COVARIATES, aes(x = Dx, y = ageOfDeath)) + geom_boxplot()
p[[2]] = p[[2]] + ggtitle('AgeOfDeath') + my.theme

# PMI
p[[3]] = ggplot(COVARIATES, aes(x = Dx, y = PMI)) + geom_boxplot()
p[[3]] = p[[3]] + ggtitle('PMI (in hours)') + my.theme

# Intronic Rate
p[[4]] = ggplot(COVARIATES, aes(x = Dx, y = IntronicRate)) + geom_boxplot()
p[[4]] = p[[4]] + ggtitle('Intronic Rate') + my.theme

# IntergenicRate
p[[5]] = ggplot(COVARIATES, aes(x = Dx, y = IntergenicRate)) + geom_boxplot()
p[[5]] = p[[5]] + ggtitle('Intergenic Rate') + my.theme

# Transcripts Detected
p[[6]] = ggplot(COVARIATES, aes(x = Dx, y = IntragenicRate)) + geom_boxplot()
p[[6]] = p[[6]] + ggtitle('Intragenic Rate') + my.theme

# Mapped Reads
p[[7]] = ggplot(COVARIATES, aes(x = Dx, y = MappedReads)) + geom_boxplot()
p[[7]] = p[[7]] + ggtitle('Mapped Reads') + my.theme

# PercentAligned
p[[8]] = ggplot(COVARIATES, aes(x = Dx, y = TotalReads)) + geom_boxplot()
p[[8]] = p[[8]] + ggtitle('Total Reads') + my.theme

# rRNARate
p[[9]] = ggplot(COVARIATES, aes(x = Dx, y = rRNARate)) + geom_boxplot()
p[[9]] = p[[9]] + ggtitle('rRNARate') + my.theme

multiplot(plotlist = p, cols = 3)

# Institution
# vcd::mosaic(~ Institution + Dx, data = COVARIATES)

# Gender
# vcd::mosaic(~ Reported_Gender + Dx, data = COVARIATES)
```
### Filter genes
Remove genes that have less than 1 cpm counts in at least 50% of samples per Dx and per Dx.sex. Also remove genes with missing gene length and percentage GC content

```{r cpmnormalisation}
rownames(COUNT) = COUNT$gene_id
COUNT$gene_id = NULL

genesToAnalyze = dlply(METADATA, .(Dx, Reported_Gender), .fun = function(mtd, count){
  processed.counts = getGeneFilteredGeneExprMatrix(count[,mtd$SampleID],
                                                   MIN_GENE_CPM=1, 
                                                   MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0.5)
  processed.counts$filteredExprMatrix$genes
}, COUNT)

genesToAnalyze = unlist(genesToAnalyze) %>% 
  unique() %>% 
  intersect(GENE.GC.CONT$gene_id[!is.na(GENE.GC.CONT$percentage_gene_gc_content)]) %>% 
  intersect(gene_lengths$gene_id[!is.na(gene_lengths$Length)])

PROCESSED_COUNTS = getGeneFilteredGeneExprMatrix(COUNT[genesToAnalyze, ], MIN_GENE_CPM=0, MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0)
pct.pc = PROCESSED_COUNTS$filteredExprMatrix$genes %>%
  dplyr::rename(gene_id = genes) %>%
  left_join(backgroundGenes) %>%
  left_join(Ensemble2HGNC) %>%
  group_by(gene_biotype) %>%
  dplyr::summarise(fraction  = n()) %>%
  filter(fraction > 100) %>%
  dplyr::mutate(fraction = fraction/length(PROCESSED_COUNTS$filteredExprMatrix$genes[,1]))
```
`r dim(PROCESSED_COUNTS$filteredExprMatrix$genes)[1]` genes are used for the analysis.

Following is the distribution of biotypes
`r kable(pct.pc)`

### Outlier analysis
Detect outlier samples based on their expression (logCPM) pattern
```{r detect.outliers, fig.height=8, fig.width=8, results='asis', cache = FALSE}
# Find principal components of expression to plot
PC <- prcomp(voom(PROCESSED_COUNTS$filteredExprMatrix$counts)$E, scale.=T, center = T)

# Plot first 2 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])

# Percentage from each PC
eigen <- PC$sdev^2
pc1 <- eigen[1]/sum(eigen)
pc2 <- eigen[2]/sum(eigen)

# Identify outliers - samples 4SDs from the mean
outliers <- as.character(plotdata$SampleID[c(which(plotdata$PC1 < mean(plotdata$PC1) - 4*sd(plotdata$PC1)),
                              which(plotdata$PC1 > mean(plotdata$PC1) + 4*sd(plotdata$PC1))), drop = T])

outliers <- c(outliers, as.character(plotdata$SampleID[c(which(plotdata$PC2 < mean(plotdata$PC2) - 4*sd(plotdata$PC2)),
                                           which(plotdata$PC2 > mean(plotdata$PC2) + 4*sd(plotdata$PC2))), drop = T] ))

plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, "SampleID")) %>%
  dplyr::mutate(label = SampleID) %>% 
  dplyr::mutate(label = ifelse((label %in% outliers), label, NA)) %>% 
  dplyr::mutate(Tissue = "sgACC") %>% 
  dplyr::mutate(Institution = "NIMH-HBCC")

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(shape= Tissue, color = Institution, size = ageOfDeath))
p <- p + theme_bw() %+replace% theme(legend.position="right")
p <- p + geom_text(aes(label= label), size=4, hjust=0)
p
# Plot abberent distribution of logcpm counts
tmp1 = voom(PROCESSED_COUNTS$filteredExprMatrix$counts)$E %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  tidyr::gather(SampleID, logCPM, -ensembl_gene_id) %>%
  left_join(COVARIATES %>%
              rownameToFirstColumn('SampleID'))

indToRetain = setdiff(colnames(PROCESSED_COUNTS$filteredExprMatrix$counts), outliers)
PROCESSED_COUNTS$filteredExprMatrix$counts = PROCESSED_COUNTS$filteredExprMatrix$counts[,indToRetain]
COVARIATES = COVARIATES[indToRetain,]

tmp = COVARIATES %>%
  group_by(Dx) %>%
  dplyr::summarise(count = n())

NEW.COUNTS = PROCESSED_COUNTS$filteredExprMatrix$counts
```
Processing `r dim(PROCESSED_COUNTS$filteredExprMatrix)[1]` genes in `r dim(PROCESSED_COUNTS$filteredExprMatrix)[2]` samples from `r length(unique(COVARIATES$Individual_ID))` unique individuals

Based on the expression pattern following samples were tagged as outliers: `r paste(outliers, collapse = ', ')`

Distribution of samples are:
`r kable(tmp)`

### Library Normalisation
Initial library normalisation usign cqn
```{r cqn}
# Compute offset for gene length and gc content
gene_lengths = tibble::column_to_rownames(gene_lengths, var = 'gene_id') %>% as.data.frame()
CQN.GENE_EXPRESSION = cqn(NEW.COUNTS, 
                          x = GENE.GC.CONT[PROCESSED_COUNTS$filteredExprMatrix$genes$genes, 'percentage_gene_gc_content'],
                          lengths = gene_lengths[PROCESSED_COUNTS$filteredExprMatrix$genes$genes, 'Length'],
                          lengthMethod = "smooth", 
                          verbose = FALSE)
CQN.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$y + CQN.GENE_EXPRESSION$offset
```

Remove 0 variance gene: ENSG00000210082.2
```{r zero_var}
CQN.GENE_EXPRESSION$E <- subset(CQN.GENE_EXPRESSION$E, rownames(CQN.GENE_EXPRESSION$E) != "ENSG00000210082.2")
```

### Co-expression distribution
Coexpression of genes 
```{r coexp1, cache=FALSE, fig.height=5, fig.width=5}
cr = cor(t(CQN.GENE_EXPRESSION$E))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```

### Sample clustering
```{r decompse.normalise.data, fig.height=8, fig.width=8, results='asis', cache=FALSE}
# Find principal components of expression to plot
PC <- prcomp(CQN.GENE_EXPRESSION$E, scale.=T, center = T)

# Plot first 2 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])

plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>%
  dplyr::mutate(Tissue = "sgACC") %>% 
  dplyr::mutate(Institution = "NIMH-HBCC")

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(shape= Tissue, color = Institution, size = ageOfDeath))
p <- p + theme_bw() %+replace% theme(legend.position="top", plot.title = element_text(hjust=0.5, size = 15)) + ggtitle ("PCA HBCC Cohort") + 
  guides(colour = guide_legend(nrow = 2),
         shape = guide_legend(nrow = 2),
         size = guide_legend(nrow = 2))
p
# p <- p + geom_text(aes(label= SampleID), size=4, hjust=0)
```
Tree based clustering of samples
```{r decompse.normalise.data.1, fig.height=6, fig.width=10, results='asis'}
# Eucledian tree based analysis
COVARIATES.tmp = data.matrix(COVARIATES[,c("Reported_Gender", "LibraryBatch", "Dx")])
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0

tree = hclust(as.dist(t(CQN.GENE_EXPRESSION$E)))
cols = WGCNA::labels2colors(COVARIATES.tmp);

WGCNA::plotDendroAndColors(tree, 
                           colors = cols, 
                           dendroLabels = FALSE, 
                           abHeight = 0.80, 
                           main = "Sample dendrogram",
                           groupLabels = colnames(COVARIATES.tmp))
```

```{r temp, include = F}
dev.off()
gc()
```

### Distribution of samples (log cpm)
```{r lcpm.dist, cache=FALSE}
# Plot abberent distribution of logcpm counts
tmp1 = (CQN.GENE_EXPRESSION$E) %>%
  rownameToFirstColumn('gene_id') %>%
  tidyr::gather(SampleID, logCPM, -gene_id) %>%
  left_join(COVARIATES %>%
              rownameToFirstColumn('SampleID'))

p = ggplot(tmp1, aes(x = logCPM, color = SampleID)) + geom_density() 
p = p + theme_bw() %+replace% theme(legend.position = 'NONE') + facet_grid(.~Dx, scale = 'free')
p
```

### Significant Covariates
Correlation between pca of unadjusted mRNA expression and covariates is used to find significant covariates
```{r preAdjusted.covariates, cache=TRUE}
# Find correlation between PC's of gene expression with covariates
preAdjustedSigCovars = runPCAandPlotCorrelations(CQN.GENE_EXPRESSION$E, 
                                                 COVARIATES,
                                                 'NULL design(voom-normalized)', 
                                                 isKeyPlot=TRUE, 
                                                 MIN_PVE_PCT_PC = 1)
```

Significant covariates to adjust at FDR 0.1 are `r preAdjustedSigCovars$significantCovars`
```{r preAdjustedSigCovars.NULL, fig.width=20, fig.height=12}
preAdjustedSigCovars[["PC_res"]][[2]]$plotData
```

### Model Identification

#### Phase I (clinical, ancestry and sample specific technical variables)
```{r phase1}
#Fit a base model with the following covariates
base_covars <- "Dx"
random_effect <- "Individual_ID"

# Set base formula
base_formula <- glue::glue("~ ", glue::glue_collapse(base_covars, sep = " + ")) %>%
  glue::glue(" + (1|{random_effect})")

# Stepwise regression
phase1 <- mvBIC::mvForwardStepwise(exprObj = CQN.GENE_EXPRESSION$E,
                         baseFormula = base_formula,
                         data = COVARIATES,
                         variables = array(c("(1|Reported_Gender)", "scale(PMI)",
                                             "scale(ageOfDeath)", "scale(RIN)", "scale(RIN2)", "scale(EV.1)",
                                             "scale(EV.2)", "scale(EV.3)", "scale(EV.4)", "scale(EV.5)")))
phase1
```

#### Phase II (batch effects)
```{r phase2}
# Fit a base model with the following covariates
phase2 <- mvBIC::mvForwardStepwise(exprObj = CQN.GENE_EXPRESSION$E,
                         baseFormula = phase1$formula, 
                         data = COVARIATES,
                         variables = array(c("(1|LibraryBatch)", "(1|IsolationBatch)")))
phase2
```

#### Phase III (RNASeq alignment specific covariates)
```{r phase3}
# Fit a base model with the following covariates
phase3 <- mvBIC::mvForwardStepwise(exprObj = CQN.GENE_EXPRESSION$E,
                         baseFormula = phase2$formula, 
                         data = COVARIATES,
                         variables = array(c("scale(IntragenicRate)", "scale(IntronicRate)","scale(IntergenicRate)","scale(rRNARate)", 
                                             "scale(TotalReads)", "scale(GenesDetected)", "scale(MappedReads)")))
phase3
```

#### Variables included in the model 
Covariates were added as fixed and random effects iteratively in three phases if model improvement by BIC criteria was observed. Clinical, ancestry and sample-specific technical variables were tested for model improvement first. Covariates related to batch effects were next tested for model improvement and finally, RNASeq alignment-specific covariates.

The model for clinical, technical and batch variables selected after phase I, II and III is `r paste(phase3$formula)`

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Set annotations
all.annotations = list(
  dataType = 'geneExpression',
  dataSubtype = 'processed',
  assay	 = 'rnaSeq',

  tissue	= 'anterior cingulate cortex', 
  study = 'CMC',

  species = 'Human',
  consortium	= 'CMC',

  analysisType	= "data normalization",

  transcriptNormalizationMethod	= 'CQN',
  transcriptQuantificationMethod = 'featureCounts',
  referenceSet = 'GRCh38'
)

# Code
CODE <- Folder(name = "HBCC - sgACC - DREAM - FEATURECOUNTS - GRCh38 - CQN", parentId = parentId)
CODE <- synStore(CODE)

# Store covariates
COVARIATES = rownameToFirstColumn(COVARIATES, 'SampleID')
write.table(COVARIATES, file = 'CMC_HBCC_sgACC_Covariates.tsv', sep = '\t', row.names=F, quote=F)
COV_OBJ = synapser::File('CMC_HBCC_sgACC_Covariates.tsv', name = 'Covariates', parentId = CODE$properties$id)
COV_OBJ = synStore(COV_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                   executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'covariates'
synSetAnnotations(COV_OBJ, annotations = all.annotations)

# Store filtered counts
PROCESSED_COUNTS$filteredExprMatrix$counts %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'CMC_HBCC_sgACC_Counts.tsv', sep = '\t', row.names=F, quote=F)
COUNT_OBJ = File('CMC_HBCC_sgACC_Counts.tsv', name = 'Counts (filtered raw)', parentId = CODE$properties$id)
COUNT_OBJ = synStore(COUNT_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                     executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'filteredCounts'
synSetAnnotations(COUNT_OBJ, annotations = all.annotations)

# Store logCPM
CQN.GENE_EXPRESSION$E %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'CMC_HBCC_sgACC_logCPM.tsv', sep = '\t', row.names=F, quote=F)
LCOUNT_OBJ = File('CMC_HBCC_sgACC_logCPM.tsv', name = 'Counts (filtered logCPM)', parentId = CODE$properties$id)
LCOUNT_OBJ = synStore(LCOUNT_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'filteredLCPM'
synSetAnnotations(LCOUNT_OBJ, annotations = all.annotations)

# Store cqn offsets
CQN.GENE_EXPRESSION$offset %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'CMC_HBCC_sgACC_offset.tsv', sep = '\t', row.names=F, quote=F)
OFFSET_OBJ = File('CMC_HBCC_sgACC_offset.tsv', name = 'Gene length and GC content offset', parentId = CODE$properties$id)
OFFSET_OBJ = synStore(OFFSET_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'offset'
synSetAnnotations(OFFSET_OBJ, annotations = all.annotations)
```

|  *Results*                                  |  *SynapseID*                     |
|  -----------------------------------------  |   ---------                      |
|  Covariates                                 |  `r COV_OBJ$properties$id`       |
|  Counts (raw)                               |  `r COUNT_OBJ$properties$id`     |
|  Counts (lcpm)                              |  `r LCOUNT_OBJ$properties$id`    |
|  Offset (for gene length and gc content)    |  `r OFFSET_OBJ$properties$id`    |

### Source Rmd
[Source markdown](`r thisFile`)

### 2020-04-26 HBCC_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd
completed cohorts

--- 2020-04-25 HBCC_sgACC_FEATURECOUNTS_GRCh38_CQN.Rmd
@@
 ---
-title: "HBCC - sgACC - FEATURECOUNTS - GRCh38 - CQN"
-author: "Kelsey Montgomery, Gabriel Hoffman and Thanneer Perumal"
+title: "HBCC - DLPFC - FEATURECOUNTS - GRCh38 - CQN"
+author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
 date: "`r date()`"
@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "./cmc-rnaseq-de/HBCC_sgACC_FEATURECOUNTS_GRCh38_CQN.Rmd",
+knit2synapse::createAndKnitToFolderEntity(file = "HBCC_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd",
                                           parentId ="syn21435176",
-                                          folderName = 'HBCC - sgACC - DREAM - FEATURECOUNTS - GRCh38 - CQN')
+                                          folderName = 'HBCC - DLPFC - DREAM - FEATURECOUNTS - GRCh38 - CQN')
 ```
@@
 library(stringr)
+library(mvBIC)
 
@@
 activityName = 'Covariate adjustments';
-activityDescription = 'Covariate analysis of GRCh38 feature counts with CQN normalisation (sgACC)';
+activityDescription = 'Covariate analysis of GRCh38 feature counts with CQN normalisation (DLPFC)';
 
-thisFileName <- 'HBCC_sgACC_FEATURECOUNTS_GRCh38_CQN.Rmd'
+thisFileName <- 'HBCC_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd'
 
@@
 thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
-thisFile <- getPermlink(repository = thisRepo, repositoryPath=thisFileName)
+thisFile <- getPermlink(repository = thisRepo, repositoryPath = thisFileName)
 ```
@@
 # Get RNASeq QCmetadata
-METADATA_QC_ACC_ID = 'syn16816488' 
-ALL_USED_IDs = METADATA_QC_ACC_ID
-metadata = downloadFile_version(METADATA_QC_ACC_ID, version = 20) %>%
-  dplyr::select(`Individual_ID`, Institution, `Reported_Gender`, Sex, Ethnicity, ageOfDeath, `PMI_(in_hours)`, Dx, 
-                Sample_RNA_ID, one_of('Cohort',
-                                      'rnaSeq_isolation:RIN',
+METADATA_QC_DLPFC_ID = 'syn16816488' 
+ALL_USED_IDs = METADATA_QC_DLPFC_ID
+metadata = downloadFile_version(METADATA_QC_DLPFC_ID, version = 19) %>%
+  dplyr::select(`Individual_ID`, Institution, Cohort, `Reported_Gender`, Sex, Ethnicity, ageOfDeath, `PMI_(in_hours)`, Dx, 
+                Sample_RNA_ID, one_of('rnaSeq_isolation:RIN',
                                       'rnaSeq_dissection:Brain_Region',
                                       'rnaSeq_report:Ribozero_Batch', 'rnaSeq_report:Library_Batch', 
-                                      'rnaSeq_isolation:RNA_Isolation_Batch', 'rnaSeq_report:Exclude?',
+                                      'rnaSeq_report:Flowcell_Batch', 'rnaSeq_report:Exclude?',
                                       'rnaSeq_report:Mapped_Reads',
@@
                                       'rnaSeq_report:Transcripts_Detected')) %>% 
-  dplyr::filter(`rnaSeq_dissection:Brain_Region` %in% c("sgACC")) %>% 
-  dplyr::filter(Cohort %in% c("NIMH-HBCC-5"))
+  dplyr::filter(`rnaSeq_dissection:Brain_Region` %in% c("DLPFC"))
 
-## Download reprocessed counts (ACC)
-counts = 'syn21867762'
+## Download reprocessed counts (DLPFC)
+counts = 'syn21754352'
 ALL_USED_IDs[length(ALL_USED_IDs)+1] = counts
@@
   dplyr::inner_join(ANCESTRY) %>% 
-  dplyr::filter(is.na(`SNP_report:Exclude?`)) #Remove excluded samples here because there are duplicate individuals
+  dplyr::filter(is.na(`SNP_report:Exclude?`))
 ```
@@
                 LibraryBatch = `rnaSeq_report:Library_Batch`,
-                IsolationBatch = `rnaSeq_isolation:RNA_Isolation_Batch`,
+                FlowcellBatch = `rnaSeq_report:Flowcell_Batch`,
                 RibozeroBatch = `rnaSeq_report:Ribozero_Batch`,
@@
                 TotalReads = `rnaSeq_report:Total_Reads`, AlignmentRate = `rnaSeq_report:Percent_Aligned`, TranscriptsDetected = `rnaSeq_report:Transcripts_Detected`) %>%
-  dplyr::select(SampleID, Individual_ID, Institution, Reported_Gender,Sex, Ethnicity, ageOfDeath, PMI, Dx, 
+  dplyr::select(SampleID, Individual_ID, Institution, Cohort, Reported_Gender,Sex, Ethnicity, ageOfDeath, PMI, Dx, 
                 RIN, ReportExclude, GenotypeExclude, EV.1, EV.2, EV.3, EV.4, EV.5,
-                LibraryBatch, IsolationBatch, RibozeroBatch, MappedReads, TotalReads, GenesDetected, AlignmentRate,
+                LibraryBatch, FlowcellBatch, RibozeroBatch, MappedReads, TotalReads, GenesDetected, AlignmentRate,
                 IntragenicRate, IntergenicRate, IntronicRate, ExpProfEfficiency, rRNARate, TranscriptsDetected) %>%
-  dplyr::mutate(Dx = forcats::fct_recode(Dx, Other = "BP", Other = "MDD", Control = "Control", SCZ = "SCZ"))
-
-# Limit HBCC ACC Counts
-rownames(COUNT) = COUNT$Geneid
-COUNT <- COUNT[,colnames(COUNT) %in% METADATA$SampleID]
+  dplyr::filter(Cohort %in% "NIMH-HBCC") %>% 
+  dplyr::filter(SampleID %in% colnames(COUNT), !is.na(SampleID)) %>% 
+  dplyr::mutate(Dx = forcats::fct_recode(Dx, Other = "BP", Other = "undetermined", Control = "Control", SCZ = "SCZ"))
 
@@
 
-ind = METADATA$SampleID [is.na(METADATA$RIN)]
-writeLines(paste('Following', length(ind), 'counts are missing RIN'))
-writeLines(paste(ind, collapse = ', '))
-METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))
-
 ind = METADATA$SampleID [is.na(METADATA$EV.1)]
@@
 indToRetain = intersect(METADATA$SampleID, colnames(COUNT))
-removedIDs = setdiff(colnames(COUNT), METADATA$SampleID)
 
+rownames(COUNT) <- COUNT$Geneid
 COUNT = COUNT[,indToRetain]
@@
 ```
-Following sample are removed `r paste(removedIDs, collapse = ',')`
 
@@
 mart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL",
-                host = "uswest.ensembl.org", # Ensembl Release 99 (January 2020)
+                host = "uswest.ensembl.org", # Ensembl Release 99 (January 2020)",
                 dataset = "hsapiens_gene_ensembl")
@@
 ```{r covariates.clustering}
-FactorCovariates <- c('Individual_ID', "Reported_Gender", "LibraryBatch", "Dx", "IsolationBatch")
+FactorCovariates <- c('Individual_ID', "Reported_Gender", "LibraryBatch", "Dx", "FlowcellBatch", "RibozeroBatch")
 ContCovariates <- c("ageOfDeath", "PMI", "RIN", "MappedReads", "IntragenicRate", "IntronicRate", "IntergenicRate",
@@
 ```{r covariates.correlation, fig.width=10, fig.height=8}
-COVARIATES.CORRELATION = CovariateAnalysis::getAssociationStatistics(COVARIATES, PVAL = 0.05)
+COVARIATES.CORRELATION = getAssociationStatistics(COVARIATES, PVAL = 0.05)
 tmp = COVARIATES.CORRELATION$ESTIMATE
@@
 h = Heatmap(tmp, col = colorRamp2(c(-1,0,1), c('blue','white','red')), name = 'AssocEstimate')
-draw(h, heatmap_legend_side = 'left')
+ComplexHeatmap::draw(h, heatmap_legend_side = 'left')
 ```
@@
   dplyr::mutate(label = ifelse((label %in% outliers), label, NA)) %>% 
-  dplyr::mutate(Tissue = "sgACC") %>% 
-  dplyr::mutate(Institution = "NIMH-HBCC")
+  dplyr::mutate(Institution = "NIMH-HBCC") %>% 
+  dplyr::mutate(Tissue = "DLPFC")
 
 p <- ggplot(plotdata, aes(x=PC1, y=PC2))
-p <- p + geom_point(aes(shape= Tissue, color = Institution, size = ageOfDeath))
+p <- p + geom_point(aes(shape= Tissue, color = Institution, size=ageOfDeath))
 p <- p + theme_bw() %+replace% theme(legend.position="right")
@@
               rownameToFirstColumn('SampleID'))
+# p = ggplot(tmp1 %>%
+#              dplyr::filter(SampleID %in% indToRemove),
+#            aes(x = logCPM, color = SampleID)) + geom_density() 
+# p = p + theme_bw() %+replace% theme(legend.position = 'none') + facet_grid(Dx~., scale = 'free')
+# p
 
@@
 
-Remove 0 variance gene: ENSG00000210082.2
-```{r zero_var}
-CQN.GENE_EXPRESSION$E <- subset(CQN.GENE_EXPRESSION$E, rownames(CQN.GENE_EXPRESSION$E) != "ENSG00000210082.2")
-```
-
 ### Co-expression distribution
@@
 plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>% 
-  dplyr::mutate(Tissue = "sgACC") %>% 
-  dplyr::mutate(Institution = "NIMH-HBCC")
+  dplyr::mutate(Institution = "NIMH-HBCC") %>% 
+  dplyr::mutate(Tissue = "DLPFC")
 
 p <- ggplot(plotdata, aes(x=PC1, y=PC2))
-p <- p + geom_point(aes(shape= Tissue, color = Institution, size = ageOfDeath))
-p <- p + theme_bw() %+replace% theme(legend.position="top", plot.title = element_text(hjust=0.5, size = 15)) + ggtitle ("PCA HBCC Cohort") + 
-  guides(colour = guide_legend(nrow = 2),
-         shape = guide_legend(nrow = 2),
-         size = guide_legend(nrow = 2))
+p <- p + geom_point(aes(color=Institution, shape=Tissue, size=ageOfDeath))
+p <- p + theme_bw() + theme(legend.position="right") + facet_grid(Tissue~.)
+# p <- p + geom_text(aes(label= SampleID), size=4, hjust=0)
 p
-# p <- p + geom_text(aes(label= SampleID), size=4, hjust=0)
 ```
@@
                          data = COVARIATES,
-                         variables = array(c("(1|LibraryBatch)", "(1|IsolationBatch)")))
+                         variables = array(c("(1|LibraryBatch)", "(1|FlowcellBatch)")))
 phase2
@@
 
-  tissue	= 'anterior cingulate cortex', 
+  tissue	= 'dorsolateral prefrontal cortex, anterior cingulate cortex', 
   study = 'CMC',
@@
 # Code
-CODE <- Folder(name = "HBCC - sgACC - DREAM - FEATURECOUNTS - GRCh38 - CQN", parentId = parentId)
+CODE <- Folder(name = "HBCC - DLPFC - DREAM - FEATURECOUNTS - GRCh38 - CQN", parentId = parentId)
 CODE <- synStore(CODE)
@@
 COVARIATES = rownameToFirstColumn(COVARIATES, 'SampleID')
-write.table(COVARIATES, file = 'CMC_HBCC_sgACC_Covariates.tsv', sep = '\t', row.names=F, quote=F)
-COV_OBJ = synapser::File('CMC_HBCC_sgACC_Covariates.tsv', name = 'Covariates', parentId = CODE$properties$id)
+write.table(COVARIATES, file = 'CMC_HBCC_DLPFC_Covariates.tsv', sep = '\t', row.names=F, quote=F)
+COV_OBJ = synapser::File('CMC_HBCC_DLPFC_Covariates.tsv', name = 'Covariates', parentId = CODE$properties$id)
 COV_OBJ = synStore(COV_OBJ, used = ALL_USED_IDs, activityName = activityName, 
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_HBCC_sgACC_Counts.tsv', sep = '\t', row.names=F, quote=F)
-COUNT_OBJ = File('CMC_HBCC_sgACC_Counts.tsv', name = 'Counts (filtered raw)', parentId = CODE$properties$id)
+  write.table(file = 'CMC_HBCC_DLPFC_Counts.tsv', sep = '\t', row.names=F, quote=F)
+COUNT_OBJ = File('CMC_HBCC_DLPFC_Counts.tsv', name = 'Counts (filtered raw)', parentId = CODE$properties$id)
 COUNT_OBJ = synStore(COUNT_OBJ, used = ALL_USED_IDs, activityName = activityName, 
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_HBCC_sgACC_logCPM.tsv', sep = '\t', row.names=F, quote=F)
-LCOUNT_OBJ = File('CMC_HBCC_sgACC_logCPM.tsv', name = 'Counts (filtered logCPM)', parentId = CODE$properties$id)
+  write.table(file = 'CMC_HBCC_DLPFC_logCPM.tsv', sep = '\t', row.names=F, quote=F)
+LCOUNT_OBJ = File('CMC_HBCC_DLPFC_logCPM.tsv', name = 'Counts (filtered logCPM)', parentId = CODE$properties$id)
 LCOUNT_OBJ = synStore(LCOUNT_OBJ, used = ALL_USED_IDs, activityName = activityName, 
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_HBCC_sgACC_offset.tsv', sep = '\t', row.names=F, quote=F)
-OFFSET_OBJ = File('CMC_HBCC_sgACC_offset.tsv', name = 'Gene length and GC content offset', parentId = CODE$properties$id)
+  write.table(file = 'CMC_HBCC_DLPFC_offset.tsv', sep = '\t', row.names=F, quote=F)
+OFFSET_OBJ = File('CMC_HBCC_DLPFC_offset.tsv', name = 'Gene length and GC content offset', parentId = CODE$properties$id)
 OFFSET_OBJ = synStore(OFFSET_OBJ, used = ALL_USED_IDs, activityName = activityName, 

### 2020-04-26 MSSM-Penn-Pitt_ACC_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd
completed cohorts

---
title: "MSSM-Penn-Pitt - ACC and DLPFC - DREAM - FEATURECOUNTS - GRCh38 - CQN - Model Selection"
author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
date: "`r date()`"
output: html_document
---

```{r knit2synapse, eval=FALSE}
library(synapser)
library(knit2synapse)

synLogin()

knit2synapse::createAndKnitToFolderEntity(file = "MSSM-Penn-Pitt_ACC_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd",
                                 parentId ="syn21435176",
                                 folderName = "MSSM-Penn-Pitt - ACC and DLPFC - DREAM - FEATURECOUNTS - GRCh38 - CQN")
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## Load required libraries
library(CovariateAnalysis) # get the package from devtools::install_github('th1vairam/CovariateAnalysis@dev')
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)
library(R.utils)
library(ggplot2)
library(limma)
library(psych)
library(edgeR)
library(cqn)
library(lme4)
library(biomaRt)
library(variancePartition)
library(synapser)
library(knitr)
library(githubr) # get the package from devtools::install_github('brian-bot/githubr')

synLogin()

library(BiocParallel)
library(foreach)
library(future)

param = SnowParam((availableCores()-2), "SOCK", progressbar=TRUE)
register(param)

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE, cache=FALSE}
parentId = 'syn21435176';
activityName = 'Covariate adjustments';
activityDescription = 'Covariate analysis of GRCh38 feature counts with CQN normalisation (ACC and DLPFC)';

thisFileName <- 'MSSM-Penn-Pitt_ACC_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd'

# Github link
thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=thisFileName)
```
### Data download
#### Obtain count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
downloadFile <- function(id){
  fread(synGet(id)$path, data.table = F)
}
downloadFile_version <- function(id , version){
  fread(synGet(id, version = version)$path, data.table = F)
}
# Download reprocessed counts
counts = 'syn21867938'
ALL_USED_IDs = counts
counts = downloadFile(counts) %>% data.frame()
counts$Chr <- NULL
counts$Start <- NULL
counts$End <- NULL
counts$Strand <- NULL

# Parse gene lengths
gene_lengths = counts[,c("Geneid", "Length")]
counts$Length <- NULL
gene_lengths <- dplyr::rename(gene_lengths, gene_id = Geneid)

# Get ancestry vector calculated using gemtools
ANCESTRY_ID = 'syn2511399'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = ANCESTRY_ID
ANCESTRY = downloadFile(ANCESTRY_ID) %>% 
  plyr::rename(c('DNA_report..Genotyping.Sample_ID' = 'SNP_report:Genotyping_Sample_ID'))

# Get genotype ids from synapse
GENOTYPE_ID = 'syn16816490'
ALL_USED_IDs = c(ALL_USED_IDs, GENOTYPE_ID)
GENOTYPE = downloadFile(GENOTYPE_ID) %>% 
  dplyr::select(Individual_ID, `SNP_report:Genotyping_Sample_ID`) %>%
  dplyr::inner_join(ANCESTRY)

#Get Metadata
metadata = 'syn16816488' 
ALL_USED_IDs[length(ALL_USED_IDs)+1] = metadata
metadata = downloadFile_version(metadata, version = 17)  %>% 
  dplyr::select(Individual_ID, Institution, Reported_Gender, Sex, 
                Ethnicity, ageOfDeath, `PMI_(in_hours)`, Dx,
                Sample_RNA_ID, one_of('rnaSeq_isolation:RIN',
                                      'rnaSeq_report:Ribozero_Batch', 'rnaSeq_report:Library_Batch', 
                                      'rnaSeq_report:Flowcell_Batch','rnaSeq_dissection:Brain_Region',
                                      'rnaSeq_report:Exclude?',
                                      'rnaSeq_report:Mapped_Reads',
                                      'rnaSeq_report:Intragenic_Rate','rnaSeq_report:Intronic_Rate',
                                      'rnaSeq_report:Intergenic_Rate','rnaSeq_report:Genes_Detected',
                                      'rnaSeq_report:Expression_Profiling_Efficiency',
                                      'rnaSeq_report:rRNA_Rate',
                                      'rnaSeq_report:Total_Reads', 'rnaSeq_report:Percent_Aligned'))

METADATA = metadata %>%
  dplyr::left_join(GENOTYPE) %>%
  dplyr::rename(PMI = `PMI_(in_hours)`,
                ReportExclude = `rnaSeq_report:Exclude?`,
                SampleID = Sample_RNA_ID,
                LibraryBatch = `rnaSeq_report:Library_Batch`,
                FlowcellBatch = `rnaSeq_report:Flowcell_Batch`,
                RibozeroBatch = `rnaSeq_report:Ribozero_Batch`,
                Tissue = `rnaSeq_dissection:Brain_Region`,
                MappedReads = `rnaSeq_report:Mapped_Reads`,
                IntragenicRate = `rnaSeq_report:Intragenic_Rate`, 
                IntronicRate = `rnaSeq_report:Intronic_Rate`, 
                IntergenicRate = `rnaSeq_report:Intergenic_Rate`, 
                GenesDetected = `rnaSeq_report:Genes_Detected`,
                ExpProfEfficiency = `rnaSeq_report:Expression_Profiling_Efficiency`, 
                rRNARate = `rnaSeq_report:rRNA_Rate`,
                TotalReads = `rnaSeq_report:Total_Reads`, 
                AlignmentRate = `rnaSeq_report:Percent_Aligned`,
                RIN = `rnaSeq_isolation:RIN`) %>%
  dplyr::select(SampleID, Individual_ID, Institution, Reported_Gender,Sex, Ethnicity, ageOfDeath, PMI, Dx, 
                RIN, ReportExclude, EV.1, EV.2, EV.3, EV.4, EV.5,
                LibraryBatch, FlowcellBatch, RibozeroBatch, Tissue, MappedReads, TotalReads, GenesDetected, AlignmentRate,
                IntragenicRate, IntergenicRate, IntronicRate, ExpProfEfficiency, rRNARate) %>% 
  dplyr::filter(Institution %in% c("MSSM", "Penn", "Pitt")) %>% 
  dplyr::filter(SampleID %in% colnames(counts), !is.na(SampleID)) %>%
  dplyr::mutate(Dx = forcats::fct_recode(Dx,  Other = "AFF", Other = "BP", Control = "Control", SCZ = "SCZ"))
```

### Data preprocessing
```{r preprocess.data}
ind = METADATA$SampleID[which(METADATA$ReportExclude == 1)]
writeLines(paste('Following',length(ind),'samples are marked exclude'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$Ethnicity) | is.na(METADATA$Institution) | is.na(METADATA$Dx)]
writeLines(paste('Following', length(ind), 'counts are missing clinical metadata'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$PMI)]
writeLines(paste('Following', length(ind), 'counts are missing PMI'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$Reported_Gender)]
writeLines(paste('Following', length(ind), 'counts are missing gender'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$ageOfDeath)]
writeLines(paste('Following', length(ind), 'counts are missing age of death'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$EV.1)]
writeLines(paste('Following', length(ind), 'counts are missing ancestry information'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))
```

```{r preprocess.data1, results='asis'}
# Match covariates to expression data
indToRetain = intersect(METADATA$SampleID, colnames(counts))

rownames(counts) <- counts$Geneid
counts = counts[,indToRetain]

rownames(METADATA) = METADATA$SampleID
METADATA = METADATA[indToRetain,]

METADATA %>% 
  group_by(Dx, Tissue) %>% 
  dplyr::summarise(count = n()) %>% 
  spread(Tissue, count) %>%
  kable()

METADATA = METADATA %>%
  dplyr::mutate(Dx.Tissue = paste(Dx, Tissue, sep = '.'))
```

```{r gene.param}
## Get GC content from biomart
backgroundGenes = data.frame(gene_id = rownames(counts)) %>%
  dplyr::mutate(id = gene_id) %>%
  tidyr::separate(id, c('ensembl_gene_id','position'), sep = '\\.')

# Biomart down until April 20th maintenance
# Define biomart object
# mart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL", 
#                 host = "uswest.ensembl.org", # Ensembl Release 99 (January 2020)
#                 dataset = "hsapiens_gene_ensembl")
# 
# # Query biomart
# Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol", "percentage_gene_gc_content", "gene_biotype", "chromosome_name"),
#                        filters = "ensembl_gene_id", values = backgroundGenes$ensembl_gene_id,
#                        mart = mart)
Ensemble2HGNC = downloadFile("syn21907981")

GENE.GC.CONT = Ensemble2HGNC %>%
  dplyr::left_join(backgroundGenes) %>% 
  dplyr::select(gene_id, percentage_gene_gc_content, chromosome_name) %>%
  unique
rownames(GENE.GC.CONT) = GENE.GC.CONT$gene_id
```

### Sex Check
```{r Reported_Gender_plots}
counts$gene_id = rownames(counts) 
REPORTED.GENDER.COUNTS = GENE.GC.CONT %>% 
  left_join(counts) %>%
  dplyr::select(-one_of("percentage_gene_gc_content")) %>%
  filter(chromosome_name == "X" |chromosome_name == "Y") %>% 
  tidyr::gather(key = item, value = value, -c(gene_id, chromosome_name)) %>%
  dplyr::mutate(value = log(value)) %>%
  dplyr::rename(`counts(log)`= value) %>% 
  dplyr::rename(SampleID = item) %>%
  left_join(METADATA[,c("SampleID", "Reported_Gender")]) %>% 
  dplyr::rename(`Reported Gender` = Reported_Gender)

##XIST and UTY expression 
#ENSG00000229807.11 and ENSG00000183878.15 
FILT <- REPORTED.GENDER.COUNTS %>% 
  filter(gene_id == "ENSG00000229807.11" | gene_id == "ENSG00000183878.15") %>% 
  dplyr::select(-one_of("chromosome_name")) %>% 
  tidyr::spread(key = gene_id, value = `counts(log)`) %>% 
  mutate(XIST = as.numeric(`ENSG00000229807.11`)) %>% 
  mutate(UTY = as.numeric(`ENSG00000183878.15`))

p = ggplot(FILT, aes (x= XIST, y = UTY)) 
p = p + geom_point(aes(color=`Reported Gender`)) 
p
```
### Covariates clustering
Determine relationship between covariates. 
```{r covariates.clustering}
FactorCovariates <- c('Individual_ID', "Institution", "Reported_Gender", "LibraryBatch", "Dx.Tissue", "RibozeroBatch", "FlowcellBatch")
ContCovariates <- c("ageOfDeath", "PMI", "RIN", "MappedReads", "IntragenicRate", "IntronicRate", "IntergenicRate",
                    "GenesDetected", "ExpProfEfficiency", "rRNARate", "TotalReads","AlignmentRate", 
                     "EV.1", "EV.2", "EV.3", "EV.4", "EV.5")

# Find inter relation between factor covariates
COVARIATES = METADATA[,c(FactorCovariates,ContCovariates),drop=F]
COVARIATES[,FactorCovariates] <- data.frame(lapply(COVARIATES[,FactorCovariates],function(x){
  x <- sapply(x,function(y){str_replace_all(as.character(y),'[^[:alnum:]]','_')})}))
rownames(COVARIATES) <- METADATA$SampleID

# Convert factor covariates to factors
COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], function(x){
  x = as.numeric(as.character(gsub('[\\,\\%]','',x)))
})

# Add in RIN^2 values
COVARIATES$RIN2 = COVARIATES$RIN^2
ContCovariates = c(ContCovariates, 'RIN2')
```
Correlation/association between covariates at an FDR <= 0.1
```{r covariates.correlation, fig.width=10, fig.height=8}
COVARIATES.CORRELATION = getAssociationStatistics(COVARIATES, PVAL = 0.05)
tmp = COVARIATES.CORRELATION$ESTIMATE
tmp[COVARIATES.CORRELATION$PVAL > 0.05] = 0
h = Heatmap(tmp, col = colorRamp2(c(-1,0,1), c('blue','white','red')), name = 'AssocEstimate')
ComplexHeatmap::draw(h, heatmap_legend_side = 'left')
```

## Explore metatdata
```{r data.explore, fig.width = 12, fig.height = 12}
my.theme <- theme_bw() %+replace% theme(legend.position = 'top', axis.text.x = element_text(angle = 90, hjust = 1), plot.title=element_text(hjust=0.5))

# RIN
p = list()
p[[1]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = RIN)) + geom_boxplot()
p[[1]] = p[[1]] + ggtitle('RIN') + my.theme

# Age of Death
p[[2]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = ageOfDeath)) + geom_boxplot()
p[[2]] = p[[2]] + ggtitle('AgeOfDeath') + my.theme

# PMI
p[[3]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = PMI)) + geom_boxplot()
p[[3]] = p[[3]] + ggtitle('PMI (in hours)') + my.theme

# Intronic Rate
p[[4]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = IntronicRate)) + geom_boxplot()
p[[4]] = p[[4]] + ggtitle('Intronic Rate') + my.theme

# IntergenicRate
p[[5]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = IntergenicRate)) + geom_boxplot()
p[[5]] = p[[5]] + ggtitle('Intergenic Rate') + my.theme

# Transcripts Detected
p[[6]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = IntragenicRate)) + geom_boxplot()
p[[6]] = p[[6]] + ggtitle('Intragenic Rate') + my.theme

# Mapped Reads
p[[7]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = MappedReads)) + geom_boxplot()
p[[7]] = p[[7]] + ggtitle('Mapped Reads') + my.theme

# PercentAligned
p[[8]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = TotalReads)) + geom_boxplot()
p[[8]] = p[[8]] + ggtitle('Total Reads') + my.theme

# rRNARate
p[[9]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = rRNARate)) + geom_boxplot()
p[[9]] = p[[9]] + ggtitle('rRNARate') + my.theme

multiplot(plotlist = p, cols = 3)

# Institution
# vcd::mosaic(~ Institution + Dx.Tissue, data = COVARIATES)

# Gender
# vcd::mosaic(~ Gender + Dx.Tissue, data = COVARIATES)
```
### Filter genes
Remove genes that have less than 1 cpm counts in at least 50% of samples per Dx.Tissue. Also remove genes with missing gene length and percentage GC content
```{r cpmnormalisation}
counts$gene_id = NULL

genesToAnalyze = dlply(METADATA, .(Dx.Tissue, Reported_Gender), .fun = function(mtd, count){
  processed.counts = getGeneFilteredGeneExprMatrix(count[,mtd$SampleID],
                                                   MIN_GENE_CPM=1, 
                                                   MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0.5)
  processed.counts$filteredExprMatrix$genes
}, counts)

genesToAnalyze = unlist(genesToAnalyze) %>% 
  unique() %>%
  intersect(GENE.GC.CONT$gene_id[!is.na(GENE.GC.CONT$percentage_gene_gc_content)]) %>%
  intersect(gene_lengths$gene_id[!is.na(gene_lengths$Length)])

#Subset union of all genes and convert subset to logcpm values 
PROCESSED_COUNTS = getGeneFilteredGeneExprMatrix(counts[genesToAnalyze, ], MIN_GENE_CPM=0, MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0)
pct.pc = PROCESSED_COUNTS$filteredExprMatrix$genes %>%
  dplyr::rename(gene_id = genes) %>%
  left_join(backgroundGenes) %>%
  left_join(Ensemble2HGNC) %>%
  group_by(gene_biotype) %>%
  dplyr::summarise(fraction  = n()) %>%
  filter(fraction > 100) %>%
  dplyr::mutate(fraction = fraction/length(PROCESSED_COUNTS$filteredExprMatrix$genes[,1]))
```
`r dim(PROCESSED_COUNTS$filteredExprMatrix$genes)[1]` genes are used for the analysis.

Following is the distribution of biotypes
`r kable(pct.pc)`

### Outlier analysis
Detect outlier samples based on their expression (logCPM) pattern
```{r detect.outliers, fig.height=8, fig.width=8, results='asis', cache = FALSE}
# Find principal components of expression to plot
PC <- prcomp(voom(PROCESSED_COUNTS$filteredExprMatrix$counts)$E, scale.=T, center = T)

# Plot first 2 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])

# Percentage from each PC
eigen <- PC$sdev^2
pc1 <- eigen[1]/sum(eigen)
pc2 <- eigen[2]/sum(eigen)

# Identify outliers - samples 4SDs from the mean
outliers <- as.character(plotdata$SampleID[c(which(plotdata$PC1 < mean(plotdata$PC1) - 4*sd(plotdata$PC1)),
                              which(plotdata$PC1 > mean(plotdata$PC1) + 4*sd(plotdata$PC1))), drop = T])

outliers <- c(outliers, as.character(plotdata$SampleID[c(which(plotdata$PC2 < mean(plotdata$PC2) - 4*sd(plotdata$PC2)),
                                           which(plotdata$PC2 > mean(plotdata$PC2) + 4*sd(plotdata$PC2))), drop = T] ))

plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, "SampleID")) %>%
    tidyr::separate(Dx.Tissue, c('Dx','Tissue'), sep = '_') %>%
  dplyr::mutate(label = SampleID) %>% 
  dplyr::mutate(label = ifelse((label %in% outliers), label, NA))

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=Institution, shape=Tissue, size=ageOfDeath))
p <- p + theme_bw() + theme(legend.position="right") + facet_grid(Tissue~.)
p <- p + geom_text(aes(label= label), size=4, hjust=0)
p

# Plot abberent distribution of logcpm counts
tmp1 = voom(PROCESSED_COUNTS$filteredExprMatrix$counts)$E %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  tidyr::gather(SampleID, logCPM, -ensembl_gene_id) %>%
  left_join(COVARIATES %>%
              rownameToFirstColumn('SampleID') %>%
              tidyr::separate(Dx.Tissue, c('Dx', 'Tissue'), sep = '_'))

p = ggplot(tmp1 %>%
             dplyr::filter(SampleID %in% outliers),
           aes(x = logCPM, color = SampleID)) + geom_density() 
p = p + theme(legend.position = 'top') + facet_grid(Dx+.~Tissue, scale = 'free')
p

indToRetain = setdiff(colnames(PROCESSED_COUNTS$filteredExprMatrix$counts), outliers)
PROCESSED_COUNTS$filteredExprMatrix$counts = PROCESSED_COUNTS$filteredExprMatrix$counts[,indToRetain]
COVARIATES = COVARIATES[indToRetain,]

tmp = tidyr::separate(COVARIATES, Dx.Tissue, c('Dx', 'Tissue'), sep = '_') %>%
  group_by(Dx, Tissue) %>%
  dplyr::summarise(count = n()) %>%
  spread(Tissue, count)

NEW.COUNTS = PROCESSED_COUNTS$filteredExprMatrix$counts
```
Processing `r dim(PROCESSED_COUNTS$filteredExprMatrix)[1]` genes in `r dim(PROCESSED_COUNTS$filteredExprMatrix)[2]` samples from `r length(unique(COVARIATES$Individual_ID))` unique individuals

Based on the expression pattern following samples were tagged as outliers: `r paste(outliers, collapse = ', ')`

Distribution of samples are:
`r kable(tmp)`

### Library Normalisation
Initial library normalisation usign cqn
```{r cqn}
# Compute offset for gene length and gc content
gene_lengths = tibble::column_to_rownames(gene_lengths, var = 'gene_id') %>% as.data.frame()
CQN.GENE_EXPRESSION = cqn(NEW.COUNTS, 
                          x = GENE.GC.CONT[PROCESSED_COUNTS$filteredExprMatrix$genes$genes, 'percentage_gene_gc_content'],
                          lengths = gene_lengths[PROCESSED_COUNTS$filteredExprMatrix$genes$genes, 'Length'],
                          lengthMethod = "smooth", 
                          verbose = FALSE)
CQN.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$y + CQN.GENE_EXPRESSION$offset
```

### Co-expression distribution
Coexpression of genes 
```{r coexp1, cache=FALSE, fig.height=5, fig.width=5}
cr = cor(t(CQN.GENE_EXPRESSION$E))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```

### Sample clustering
```{r decompse.normalise.data, fig.height=8, fig.width=8, results='asis', cache=FALSE}
# Find principal components of expression to plot
PC <- prcomp(CQN.GENE_EXPRESSION$E, scale.=T, center = T)

# Plot first 2 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])

plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>%
  tidyr::separate(Dx.Tissue, c('Dx','Tissue'), sep = '_')

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=Institution, shape=Tissue, size=ageOfDeath))
p <- p + theme_bw() + theme(legend.position="right") + facet_grid(Tissue~.)
# p <- p + geom_text(aes(label= SampleID), size=4, hjust=0)
p
```
Tree based clustering of samples
```{r decompse.normalise.data.1, fig.height=6, fig.width=10, results='asis'}
# Eucledian tree based analysis
COVARIATES.tmp = data.matrix(COVARIATES[,c("Institution", "Reported_Gender", "LibraryBatch", "Dx.Tissue")])
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0

tree = hclust(as.dist(t(CQN.GENE_EXPRESSION$E)))
cols = WGCNA::labels2colors(COVARIATES.tmp);

WGCNA::plotDendroAndColors(tree, 
                           colors = cols, 
                           dendroLabels = FALSE, 
                           abHeight = 0.80, 
                           main = "Sample dendrogram",
                           groupLabels = colnames(COVARIATES.tmp))
```

```{r temp, include = F}
dev.off()
gc()
```

### Distribution of samples (log cpm)
```{r lcpm.dist, cache=FALSE}
# Plot abberent distribution of logcpm counts
tmp1 = (CQN.GENE_EXPRESSION$E) %>%
  rownameToFirstColumn('gene_id') %>%
  tidyr::gather(SampleID, logCPM, -gene_id) %>%
  left_join(COVARIATES %>%
              rownameToFirstColumn('SampleID'))

p = ggplot(tmp1, aes(x = logCPM, color = SampleID)) + geom_density() 
p = p + theme_bw() %+replace% theme(legend.position = 'NONE') + facet_grid(.~Dx.Tissue, scale = 'free')
p
```

### Significant Covariates
Correlation between pca of unadjusted mRNA expression and covariates is used to find significant covariates
```{r preAdjusted.covariates, cache=TRUE}
# Find correlation between PC's of gene expression with covariates
preAdjustedSigCovars = runPCAandPlotCorrelations(CQN.GENE_EXPRESSION$E, 
                                                 COVARIATES,
                                                 'NULL design(voom-normalized)', 
                                                 isKeyPlot=TRUE, 
                                                 MIN_PVE_PCT_PC = 1)
```

Significant covariates to adjust at FDR 0.1 are `r preAdjustedSigCovars$significantCovars`
```{r preAdjustedSigCovars.NULL, fig.width=20, fig.height=12}
preAdjustedSigCovars[["PC_res"]][[2]]$plotData
```

### Model Identification

#### Phase I (clinical, ancestry and sample specific technical variables)
```{r phase1}
# Fit a base model with the following covariates
base_covars <- "Dx.Tissue"
random_effect <- "Individual_ID"

# Set base formula
base_formula <- glue::glue("~ ", glue::glue_collapse(base_covars, sep = " + ")) %>%
  glue::glue(" + (1|{random_effect})")

# Stepwise regression
phase1 <- mvBIC::mvForwardStepwise(exprObj = CQN.GENE_EXPRESSION$E,
                         baseFormula = base_formula,
                         data = COVARIATES,
                         variables = array(c("(1|Institution)", "(1|Reported_Gender)", "scale(PMI)",
                                                             "scale(ageOfDeath)", "scale(RIN)", "scale(RIN2)", "scale(EV.1)",
                                                             "scale(EV.2)", "scale(EV.3)", "scale(EV.4)", "scale(EV.5)")))
phase1
```

#### Phase II (batch effects)
```{r phase2}
# Fit a base model with the following covariates
phase2 <- mvBIC::mvForwardStepwise(exprObj = CQN.GENE_EXPRESSION$E,
                         baseFormula = phase1$formula, 
                         data = COVARIATES,
                         variables = array(c("(1|LibraryBatch)", "(1|FlowcellBatch)")))
phase2
```

#### Phase III (RNASeq alignment specific covariates)
```{r phase3}
# Fit a base model with the following covariates
phase3 <- mvBIC::mvForwardStepwise(exprObj = CQN.GENE_EXPRESSION$E,
                         baseFormula = phase2$formula, 
                         data = COVARIATES,
                         variables = array(c("scale(IntragenicRate)", "scale(IntronicRate)","scale(IntergenicRate)","scale(rRNARate)", 
                                             "scale(TotalReads)", "scale(GenesDetected)", "scale(MappedReads)")))
phase3
```

#### Variables included in the model 
Covariates were added as fixed and random effects iteratively in three phases if model improvement by BIC criteria was observed. Clinical, ancestry and sample-specific technical variables were tested for model improvement first. Covariates related to batch effects were next tested for model improvement and finally, RNASeq alignment-specific covariates.

The model for clinical, technical and batch variables selected after phase I, II and III is `r phase3$formula`

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Set annotations
all.annotations = list(
  dataType = 'geneExpression',
  dataSubtype = 'processed',
  assay	 = 'rnaSeq',

  tissue	= 'dorsolateral prefrontal cortex, anterior cingulate cortex', 
  study = 'CMC',

  species = 'Human',
  consortium	= 'CMC',

  analysisType	= "data normalization",

  transcriptNormalizationMethod	= 'CQN',
  transcriptQuantificationMethod = 'featureCounts',
  referenceSet = 'GRCh38'
)

# Code
CODE <- Folder(name = "MSSM-Penn-Pitt - ACC and DLPFC - DREAM - FEATURECOUNTS - GRCh38 - CQN", parentId = parentId)
CODE <- synStore(CODE)

# Store covariates
COVARIATES = rownameToFirstColumn(COVARIATES, 'SampleID')
write.table(COVARIATES, file = 'CMC_ACC_DLPFC_Covariates.tsv', sep = '\t', row.names=F, quote=F)
COV_OBJ = synapser::File('CMC_ACC_DLPFC_Covariates.tsv', name = 'Covariates', parentId = CODE$properties$id)
COV_OBJ = synStore(COV_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
synSetAnnotations(COV_OBJ, annotations = all.annotations)

# Store filtered counts
PROCESSED_COUNTS$filteredExprMatrix$counts %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'CMC_ACC_DLPFC_Counts.tsv', sep = '\t', row.names=F, quote=F)
COUNT_OBJ = File('CMC_ACC_DLPFC_Counts.tsv', name = 'Counts (filtered raw)', parentId = CODE$properties$id)
COUNT_OBJ = synStore(COUNT_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
synSetAnnotations(COUNT_OBJ, annotations = all.annotations)

# Store logCPM
CQN.GENE_EXPRESSION$E %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'CMC_ACC_DLPFC_logCPM.tsv', sep = '\t', row.names=F, quote=F)
LCOUNT_OBJ = File('CMC_ACC_DLPFC_logCPM.tsv', name = 'Counts (filtered logCPM)', parentId = CODE$properties$id)
LCOUNT_OBJ = synStore(LCOUNT_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
synSetAnnotations(LCOUNT_OBJ, annotations = all.annotations)

# Store cqn offsets
CQN.GENE_EXPRESSION$offset %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'CMC_ACC_DLPFC_offset.tsv', sep = '\t', row.names=F, quote=F)
OFFSET_OBJ = File('CMC_ACC_DLPFC_offset.tsv', name = 'Gene length and GC content offset', parentId = CODE$properties$id)
OFFSET_OBJ = synStore(OFFSET_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
synSetAnnotations(OFFSET_OBJ, annotations = all.annotations)
```

|  *Results*                                  |  *SynapseID*                     |
|  -----------------------------------------  |   ---------                      |
|  Covariates                                 |  `r COV_OBJ$properties$id`       |
|  Counts (raw)                               |  `r COUNT_OBJ$properties$id`     |
|  Counts (lcpm)                              |  `r LCOUNT_OBJ$properties$id`    |
|  Offset (for gene length and gc content)    |  `r OFFSET_OBJ$properties$id`    |

### Source Rmd
[Source markdown](`r thisFile`)

### 2020-04-26 MSSM-Penn-Pitt_DE_DxTissue.Rmd
de with mvBIC model

---
title: "Differential Expression - MSSM-Penn-Pitt - ACC DLPFC - Dx.Tissue"
author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
date: "`r date()`"
output: html_document
---

```{r knit2synapse, eval=FALSE}
library(synapser)
library(knit2synapse)

synLogin()

knit2synapse::createAndKnitToFolderEntity(file = "MSSM-Penn-Pitt_DE_DxTissue.Rmd",
                                 parentId ="syn21824667",
                                 folderName = "Differential Expression - Dx.Tissue")
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## Load required libraries
library(CovariateAnalysis) # get the package from devtools::install_github('th1vairam/CovariateAnalysis@dev')
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)
library(R.utils)
library(ggplot2)
library(limma)
library(psych)
library(lme4)
library(biomaRt)
library(variancePartition) #1.17.9
library(edgeR)
library(synapser)
library(knitr)
library(githubr) # get the package from devtools::install_github('brian-bot/githubr')

synLogin()

library(foreach)
library(future)
library(BiocParallel)

# See dream documentation to customize parallel processing       http://bioconductor.org/packages/release/bioc/vignettes/variancePartition/inst/doc/dream.html#parallel-processing
param = SnowParam((availableCores()-2)/2, "SOCK", progressbar=TRUE)

register(param)

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
parentId = 'syn21824667';
activityName = 'Differential Expression';
activityDescription = 'Differential Expression - Dx.Tissue';

thisFileName <- 'MSSM-Penn-Pitt_DE_DxTissue.Rmd'

# Github link
thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath = thisFileName)
```

### Data download
#### Obtain filtered and CQN normalized counts
```{r download.data, cache=TRUE}
downloadFile <- function(id){
  fread(synGet(id)$path, data.table = F)
}

# Download CQN counts
CQN_ID = "syn21825509"
ALL_USED_IDs = c(CQN_ID)
CQN.GENE_EXPRESSION = downloadFile(CQN_ID) %>% 
  tibble::column_to_rownames(var = "ensembl_gene_id") %>% 
  data.matrix()

# Download filtered counts
COUNTS_ID = "syn21825508"
ALL_USED_IDs = c(ALL_USED_IDs, COUNTS_ID)
NEW.COUNTS = downloadFile(COUNTS_ID) %>% 
  tibble::column_to_rownames(var = "ensembl_gene_id")

# Get covariates
COVAR_ID = "syn21825507"
ALL_USED_IDs = c(ALL_USED_IDs, COVAR_ID)
COVARIATES = downloadFile(COVAR_ID) %>% 
  tibble::column_to_rownames(var = "SampleID")

# Set variable class
FactorCovariates <- c("Individual_ID", "Institution", "Reported_Gender", "LibraryBatch", "Dx.Tissue", "RibozeroBatch", "FlowcellBatch")
COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)

ContCovariates <- c("ageOfDeath", "PMI", "RIN", "MappedReads", "IntragenicRate", "IntronicRate", "IntergenicRate",
                    "GenesDetected", "ExpProfEfficiency", "rRNARate", "TotalReads","AlignmentRate", 
                    "EV.1", "EV.2", "EV.3", "EV.4", "EV.5")
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], function(x){
  x = as.numeric(as.character(gsub('[\\,\\%]','',x)))
})

# Set covariates to adjust for from model selection
adjust.covars <- c("Dx.Tissue", "RIN2", "(1|Institution)", "ageOfDeath", "EV.1", "PMI", "(1|Reported_Gender)", "EV.4","EV.2","RIN", "EV.3", "EV.5", "IntronicRate")
```

### Normalisation - Dx.Tissue
1. Dx.Tissue is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis and tissue)
2. Individual_ID is chosen as random effect

```{r iterative.normalisation, results='asis'}
writeLines(paste('Using following covariates in the model:',
                 paste(adjust.covars, collapse=', '),
                 'as fixed effects and Individual_ID is chosen as random effect'))

# Post adjusted formula
random_effect <- "Individual_ID"
formula <- glue::glue("~ ", glue::glue_collapse(adjust.covars, sep = " + ")) %>% 
  glue::glue(" + (1|{random_effect})")

# Estimate voom weights with DREAM
geneExpr = DGEList(NEW.COUNTS)
geneExpr = edgeR::calcNormFactors(geneExpr)

VOOM.GENE_EXPRESSION = variancePartition::voomWithDreamWeights(counts = geneExpr, 
                                                               formula = formula,
                                                               data = COVARIATES,
                                                               save.plot = TRUE)

# How to plot a voom object if save.plot=TRUE
# Set plot generic to voom result EList
#' @export
setMethod("plot", signature(x="EList", y="missing"),
  function(x,  y, ...){
    if( is.null(x$voom.xy) || is.null(x$voom.line)){
        stop("x does not contain the trend information.\nvoom() must be run with save.plot=TRUE")
    }
    # points
    df = data.frame(x = x$voom.xy$x,
                    y = x$voom.xy$y)
    # trend line
    df.line = data.frame(x = x$voom.line$x,
                        y = x$voom.line$y)
    ggplot(df, aes(x,y)) + geom_point(size=.1) + theme_bw(15) + 
      theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + 
      geom_line(data=df.line, aes(x,y), color="red") + xlab(bquote(log[2](count + 0.5))) + 
      ylab(expression( sqrt("standard deviation"))) + ggtitle("Voom: Mean-variance trend") + 
      ylim(0, max(df$y))
  }
)

plot(VOOM.GENE_EXPRESSION)

# Fit linear model using new weights and new design
VOOM.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION
ADJUSTED.FIT = variancePartition::dream(exprObj = VOOM.GENE_EXPRESSION,
                                        formula = formula,
                                        data = COVARIATES,
                                        computeResiduals = TRUE)

col_type <- dplyr::select(COVARIATES, c(adjust.covars, "Individual_ID")) %>%
  dplyr::summarise_all(class) %>% 
  tidyr::gather(variable, class)

# Model categorical variables as random to report variance fractions
formula_for_variance_fractions = sapply(1:length(col_type$class), function(i){
  switch(col_type$class[i], 
    "factor" = paste0('(1|', col_type$variable[i], ')'),
    "numeric" = col_type$variable[i])
  })

formula_for_variance_fractions <- glue::glue("~ ", glue::glue_collapse(formula_for_variance_fractions, sep = " + "))

violin_plot <- variancePartition::fitExtractVarPartModel(VOOM.GENE_EXPRESSION,
                                                         formula = formula_for_variance_fractions,
                                                         data = COVARIATES)

plotVarPart(sortCols(violin_plot))

# Residuals after normalisation
RESIDUAL.GENE_EXPRESSION = residuals(ADJUSTED.FIT, CQN.GENE_EXPRESSION)

# Find PC of residual gene expression and significant covariates that are highly correlated with PCs
residualSigCovars = runPCAandPlotCorrelations(RESIDUAL.GENE_EXPRESSION, COVARIATES,
                                              'adjusted design(voom-normalized)',
                                              isKeyPlot=TRUE)
```

### Sanity check
```{r residualSigCovars.manual, fig.width=12, fig.height=8}
residualSigCovars[["PC_res"]][[2]]$plotData
```

### Tissue Correlation
```{r tissue_correlation}
hasReps = rownameToFirstColumn(COVARIATES, "SampleID") %>%
  dplyr::select(Individual_ID) %>% 
  table

df_dlpfc = rownameToFirstColumn(COVARIATES, "SampleID") %>% 
  filter( Individual_ID %in% names(hasReps[hasReps==2])) %>%
  filter(grepl("DLPFC", Dx.Tissue)) %>% 
  dplyr::select(SampleID, Individual_ID, Dx.Tissue)
df_acc = rownameToFirstColumn(COVARIATES, "SampleID") %>%
  filter( Individual_ID %in% names(hasReps[hasReps==2])) %>%
  filter(grepl("ACC", Dx.Tissue)) %>% 
  dplyr::select(SampleID, Individual_ID, Dx.Tissue)

df_merge = merge(df_dlpfc, df_acc, by="Individual_ID")

res_std <- rownameToFirstColumn(RESIDUAL.GENE_EXPRESSION, "ensembl_gene_id")

corRes = sapply(res_std$ensembl_gene_id, function(ensGene){

  x = VOOM.GENE_EXPRESSION$E[ensGene,df_merge$SampleID.x,drop=FALSE]
  y = VOOM.GENE_EXPRESSION$E[ensGene,df_merge$SampleID.y,drop=FALSE]

  cor(t(x)[,1], t(y)[,1], use="pairwise.complete.obs")
})

hist(corRes, xlab="Correlation", main="Correlation between ACC and DLPFC sample from the sample individual for each gene from ")

pca = prcomp(t(VOOM.GENE_EXPRESSION$E))

df_pca = merge(pca$x, as.data.frame(rownameToFirstColumn(COVARIATES, "SampleID")), 
               by.x="row.names", 
               by.y="SampleID" ) %>% 
  tidyr::separate(Dx.Tissue, into = c("Dx", "Tissue"), sep = "_")

ggplot(df_pca, aes(PC1, PC2, color=Tissue, shape=Institution)) + geom_point() + theme_bw(17) + theme(aspect.ratio=1)
```

### Residual calculation
Calculate weighted residuals and add back "Dx.Tissue" to the residuals
```{r varsToAddBack}
# Add variable of interest back to the residuals
varsToAddIn = grep("Dx.Tissue", colnames(ADJUSTED.FIT$design), value = T)
RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION +
  ADJUSTED.FIT$coefficients[,varsToAddIn] %*% t(ADJUSTED.FIT$design[,varsToAddIn])
```

### Coexpression of residual expression 
```{r coexp2, cache=FALSE, fig.height=5, fig.width=5}
cr = cor(t(RESIDUAL.GENE_EXPRESSION))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```

### Clustering residual data
```{r decompse.normalise.data2, fig.height=8, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(RESIDUAL.GENE_EXPRESSION, scale.=T, center = T)

# Plot first 4 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])

plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>%
  tidyr::separate(Dx.Tissue, c('Dx','Tissue'), sep = '_')

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=Institution, shape=Tissue, size=ageOfDeath))
p <- p + theme_bw() + theme(legend.position="right") + facet_grid(Tissue~., scales = 'free_y')
p
```

```{r decompose.normalise.data2.1, fig.height=8, fig.width=12, results='asis'}
# Eucledian tree based analysis
COVARIATES.tmp = data.matrix(COVARIATES[,c('Dx.Tissue','Institution')])

tree = hclust(as.dist(t(RESIDUAL.GENE_EXPRESSION)))
cols = WGCNA::labels2colors(COVARIATES.tmp);

WGCNA::plotDendroAndColors(tree, 
                           colors = cols, 
                           dendroLabels = FALSE, 
                           abHeight = 0.80, 
                           main = "Sample dendrogram",
                           groupLabels = colnames(COVARIATES.tmp[,c('Dx.Tissue','Institution')]))
```

```{r temp1, include=F}
dev.off()
gc()
```

### Differential expression analysis (SCZ - Control - Other)
Genes that are differentially expressed at an FDR <= 0.05 are:
```{r set-formula}
formula <- glue::glue("~ 0 + ", glue::glue_collapse(adjust.covars, sep = " + ")) %>% 
  glue::glue(" + (1|{random_effect})")
```

`r formula`

```{r diffExp, fig.height=10, fig.width=15}
L1 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                   formula = formula, 
                                   data = COVARIATES, 
                                   coefficient = c("Dx.TissueOther_ACC", "Dx.TissueControl_ACC"))
L2 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                   formula = formula, 
                                   data = COVARIATES, 
                                   coefficient = c("Dx.TissueSCZ_ACC", "Dx.TissueControl_ACC"))
L3 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                   formula = formula, 
                                   data = COVARIATES, 
                                   coefficient = c("Dx.TissueSCZ_ACC", "Dx.TissueOther_ACC"))
L4 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                   formula = formula, 
                                   data = COVARIATES, 
                                   coefficient = c("Dx.TissueOther_DLPFC", "Dx.TissueControl_DLPFC"))
L5 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                   formula = formula, 
                                   data = COVARIATES, 
                                   coefficient = c("Dx.TissueSCZ_DLPFC", "Dx.TissueControl_DLPFC"))
L6 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                   formula = formula, 
                                   data = COVARIATES, 
                                   coefficient = c("Dx.TissueSCZ_DLPFC", "Dx.TissueOther_DLPFC"))

L = cbind(L1, L2, L3, L4, L5, L6)

contrast <- c(glue::glue_collapse(names(L1[L1 != 0]), "-"), 
              glue::glue_collapse(names(L2[L2 != 0]), "-"),
              glue::glue_collapse(names(L3[L3 != 0]), "-"),
              glue::glue_collapse(names(L4[L4 != 0]), "-"),
              glue::glue_collapse(names(L5[L5 != 0]), "-"),
              glue::glue_collapse(names(L6[L6 != 0]), "-")
)

# Visualize contrast matrix
plotContrasts(L)

# Fit contrast
FIT.CONTR = variancePartition::dream(exprObj = VOOM.GENE_EXPRESSION,
                                     formula = formula,
                                     data = COVARIATES,
                                     L = L)

# Get background genes 
backgroundGenes = data.frame(gene_id = rownames(NEW.COUNTS)) %>%
  dplyr::mutate(id = gene_id) %>%
  tidyr::separate(id, c('ensembl_gene_id','position'), sep = '\\.')

# Get gene lengths
counts = 'syn21867938'
ALL_USED_IDs = counts
counts = downloadFile(counts) %>% data.frame()
gene_lengths = counts[,c("Geneid", "Length")] %>% 
  dplyr::rename(gene_id = Geneid)
rm(counts)

# Define biomart object
mart <- biomaRt::useMart(biomart = "ENSEMBL_MART_ENSEMBL",
                host = "uswest.ensembl.org", # Ensembl Release 99 (January 2020)
                dataset = "hsapiens_gene_ensembl")

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol", 
                                      "percentage_gene_gc_content", "gene_biotype", 
                                      "chromosome_name"),
                       filters = "ensembl_gene_id", 
                       values = backgroundGenes$ensembl_gene_id,
                       mart = mart)

# Get differential expression
DE = lapply(glue::glue("L{1:6}"), function(i, FIT){
  topTable(FIT, coef=i, number = Inf) %>%
    rownameToFirstColumn('gene_id') %>% 
    left_join(backgroundGenes)
}, FIT.CONTR) 
names(DE) <- contrast

DE = DE %>% 
  rbindlist(idcol = 'Comparison') %>%
  dplyr::mutate(Comparison = gsub('Dx.Tissue','',Comparison),
                Direction = logFC/abs(logFC),
                Direction = factor(Direction, c(-1,1), c('-1' = 'DOWN', '1' = 'UP')),
                Direction = as.character(Direction)) %>%
  tidyr::separate(Comparison, into = c('ref.state','to.state'), sep = '-') %>%
  tidyr::separate(ref.state, into = c('Dx1','Tissue'), sep = '_') %>%
  tidyr::separate(to.state, into = c('Dx2','Tissue'), sep = '_') %>%
  tidyr::unite(Comparison, Dx1, Dx2, sep = '_vs_') %>%
  left_join(Ensemble2HGNC) %>%
  left_join(gene_lengths)
DE$Direction[DE$adj.P.Val > 0.05 | abs(DE$logFC) < log2(1.2)] = 'NONE'

tmp = DE %>%
  dplyr::filter(adj.P.Val <= 0.05) %>%
  dplyr::select(ensembl_gene_id, Comparison, Tissue) %>%
  group_by(Comparison, Tissue) %>%
  dplyr::summarise(FDR_0_05 = length(unique(ensembl_gene_id)))

tmp1 = DE %>%
  dplyr::filter(adj.P.Val <= 0.05, abs(logFC) >= log2(1.2)) %>%
  dplyr::select(ensembl_gene_id, Comparison, Tissue) %>%
  group_by(Comparison, Tissue) %>%
  dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(ensembl_gene_id)))

kable(full_join(tmp,tmp1))

p = ggplot(DE, aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1))
p = p + scale_color_manual(values = c('green','grey','red')) + theme_bw() %+replace% theme(legend.position = 'top')
p = p + facet_grid(Tissue+.~Comparison, scales = 'fixed')
p
```

### Associate differential expression results with gc content, gene length and average expression
```{r associate.de, fig.height=10, fig.width=15}
pl = list()
pl[[1]] = ggplot(DE %>% 
                   arrange(match(Direction, c("NONE", "UP", "DOWN"))), aes(x = log10(Length), y = logFC, color = Direction)) + geom_point() 
pl[[1]] = pl[[1]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = log10(Length), y = logFC))
pl[[1]] = pl[[1]] + facet_grid(Tissue~.) + scale_color_manual(values = c('green','grey','red'))
pl[[1]] = pl[[1]] + theme(legend.position = 'top')

pl[[2]] = ggplot(DE %>% 
                   arrange(match(Direction, c("NONE", "UP", "DOWN"))), aes(x = percentage_gene_gc_content, y = logFC, color = Direction)) + geom_point()
pl[[2]] = pl[[2]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = percentage_gene_gc_content, y = logFC))
pl[[2]] = pl[[2]] + facet_grid(Tissue~.) + scale_color_manual(values = c('green','grey','red'))
pl[[2]] = pl[[2]] + theme(legend.position = 'top')

pl[[3]] = ggplot(DE %>% 
                    arrange(match(Direction, c("NONE", "UP", "DOWN"))), aes(x = AveExpr, y = logFC, color = Direction)) + geom_point()
pl[[3]] = pl[[3]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = AveExpr, y = logFC))
pl[[3]] = pl[[3]] + facet_grid(Tissue~.) + scale_color_manual(values = c('green','grey','red'))
pl[[3]] = pl[[3]] + theme(legend.position = 'top')

multiplot(plotlist = pl, cols = 3)
```

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Set annotations
all.annotations = list(
  dataType = 'geneExpression',
  dataSubtype = 'processed',
  assay	 = 'rnaSeq',

  tissue	= 'dorsolateral prefrontal cortex, anterior cingulate cortex', 
  study = 'CMC',

  species = 'Human',
  consortium	= 'CMC',

  analysisType	= "data normalization",

  transcriptNormalizationMethod	= 'CQN',
  transcriptQuantificationMethod = 'featureCounts',
  referenceSet = 'GRCh38'
)

# Code
CODE <- Folder(name = "Differential Expression - Dx.Tissue", parentId = parentId)
CODE <- synStore(CODE)

# Store residual gene expression for network analysis
RESIDUAL.GENE_EXPRESSION %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'CMC_ACC_DLPFC_Dx.Tissue_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
nEXP_OBJ = File('CMC_ACC_DLPFC_Dx.Tissue_netResidualExpression.tsv', 
                name = 'Dx.Tissue Normalised, covariates removed residual expression (for network analysis)', 
                parentId = CODE$properties$id)
nEXP_OBJ = synStore(nEXP_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                    executed = thisFile, activityDescription = activityDescription)
synSetAnnotations(nEXP_OBJ, annotations = all.annotations)

# Store differential expression results
write.table(DE, file = 'CMC_ACC_DLPFC_Dx.Tissue_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
DEXP_OBJ = File('CMC_ACC_DLPFC_Dx.Tissue_DiffExpression.tsv', 
                name = 'Differential Expression Results (Dx.Tissue)', 
                parentId = CODE$properties$id)
DEXP_OBJ = synStore(DEXP_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                    executed = thisFile, activityDescription = activityDescription)
synSetAnnotations(DEXP_OBJ, annotations = all.annotations)

# Store adjusted fit linear mixed model object as RDS - Dx.Tissue
saveRDS(ADJUSTED.FIT, file = 'CMC_ACC_DLPFC_Dx.Tissue_dream-output.rds')
ADJ_OBJ = File('CMC_ACC_DLPFC_Dx.Tissue_dream-output.rds',
               name = 'Dream Output object - Dx.Tissue', 
               parentId = CODE$properties$id)
ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
                   executed = thisFile, activityDescription = activityDescription)
synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
```

### 2020-04-27 MSSM-Penn-Pitt_DE_DxTissue.Rmd
fix random effects syntax in formula

@@
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Dx.Tissue", "RIN2", "(1|Institution)", "ageOfDeath", "EV.1", "PMI", "(1|Reported_Gender)", "EV.4","EV.2","RIN", "EV.3", "EV.5", "IntronicRate")
+adjust.covars <- c("Dx.Tissue", "RIN2", "ageOfDeath", "EV.1", "PMI", "EV.4","EV.2","RIN", "EV.3", "EV.5", "IntronicRate")
+random_effect <- c("Individual_ID", "Institution", "Reported_Gender")
 ```
@@
 ```{r iterative.normalisation, results='asis'}
-writeLines(paste('Using following covariates in the model:',
-                 paste(adjust.covars, collapse=', '),
-                 'as fixed effects and Individual_ID is chosen as random effect'))
+writeLines(glue::glue('Using following covariates in the model:',
+                 glue::glue_collapse(adjust.covars, sep = ", ", last = " and "),
+                 ' as fixed effects and',
+                 glue::glue_collapse(random_effect, sep = ", ", last = " and "),' chosen as random effects'))
 
 # Post adjusted formula
-random_effect <- "Individual_ID"
-formula <- glue::glue("~ ", glue::glue_collapse(adjust.covars, sep = " + ")) %>% 
-  glue::glue(" + (1|{random_effect})")
+formula <- glue::glue("~ ", glue::glue_collapse(adjust.covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))
 
@@
 
-col_type <- dplyr::select(COVARIATES, c(adjust.covars, "Individual_ID")) %>%
+col_type <- dplyr::select(COVARIATES, c(adjust.covars, random_effect)) %>%
   dplyr::summarise_all(class) %>% 
@@
 ```{r set-formula}
-formula <- glue::glue("~ 0 + ", glue::glue_collapse(adjust.covars, sep = " + ")) %>% 
-  glue::glue(" + (1|{random_effect})")
+formula <- glue::glue("~ 0 + ", glue::glue_collapse(adjust.covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))
+
 ```

### 2020-04-27 HBCC_sgACC_FEATURECOUNTS_GRCh38_CQN.Rmd
fix base formula and drop leelvs for mvForwardStepwise

@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "./cmc-rnaseq-de/HBCC_sgACC_FEATURECOUNTS_GRCh38_CQN.Rmd",
+knit2synapse::createAndKnitToFolderEntity(file = "HBCC_sgACC_FEATURECOUNTS_GRCh38_CQN.Rmd",
                                           parentId ="syn21435176",
@@
 base_covars <- "Dx"
-random_effect <- "Individual_ID"
 
 # Set base formula
-base_formula <- glue::glue("~ ", glue::glue_collapse(base_covars, sep = " + ")) %>%
-  glue::glue(" + (1|{random_effect})")
+base_formula <- glue::glue("~ ", glue::glue_collapse(base_covars, sep = " + "))
 
+COVARIATES <- droplevels(COVARIATES)
 # Stepwise regression
@@
                          baseFormula = base_formula,
-                         data = COVARIATES,
+                         data = bar,
                          variables = array(c("(1|Reported_Gender)", "scale(PMI)",

### 2020-05-04 individual-cohorts/MSSM-Penn-Pitt_ACC_FEATURECOUNTS_GRCh38_CQN.Rmd
deconstruct model selection by analyzing each tissue and cohort separately

--- 2020-04-26 MSSM-Penn-Pitt_ACC_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd
@@
 ---
-title: "MSSM-Penn-Pitt - ACC and DLPFC - DREAM - FEATURECOUNTS - GRCh38 - CQN - Model Selection"
+title: "MSSM-Penn-Pitt - ACC - DREAM - FEATURECOUNTS - GRCh38 - CQN - Model Selection"
 author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "MSSM-Penn-Pitt_ACC_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd",
-                                 parentId ="syn21435176",
-                                 folderName = "MSSM-Penn-Pitt - ACC and DLPFC - DREAM - FEATURECOUNTS - GRCh38 - CQN")
+knit2synapse::createAndKnitToFolderEntity(file = "MSSM-Penn-Pitt_ACC_FEATURECOUNTS_GRCh38_CQN.Rmd",
+                                 parentId ="syn22012942",
+                                 folderName = "MSSM-Penn-Pitt - ACC - DREAM - FEATURECOUNTS - GRCh38 - CQN")
 ```
@@
 activityName = 'Covariate adjustments';
-activityDescription = 'Covariate analysis of GRCh38 feature counts with CQN normalisation (ACC and DLPFC)';
+activityDescription = 'Covariate analysis of GRCh38 feature counts with CQN normalisation (ACC)';
 
-thisFileName <- 'MSSM-Penn-Pitt_ACC_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd'
+thisFileName <- 'MSSM-Penn-Pitt_ACC_FEATURECOUNTS_GRCh38_CQN.Rmd'
 
@@
 thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
-thisFile <- getPermlink(repository = thisRepo, repositoryPath=thisFileName)
+thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/',thisFileName))
 ```
@@
 ALL_USED_IDs[length(ALL_USED_IDs)+1] = metadata
-metadata = downloadFile_version(metadata, version = 17)  %>% 
+metadata = downloadFile_version(metadata, version = 20)  %>% 
   dplyr::select(Individual_ID, Institution, Reported_Gender, Sex, 
@@
   dplyr::filter(Institution %in% c("MSSM", "Penn", "Pitt")) %>% 
+  dplyr::filter(Tissue %in% c("ACC")) %>% 
   dplyr::filter(SampleID %in% colnames(counts), !is.na(SampleID)) %>%
@@
 METADATA %>% 
-  group_by(Dx, Tissue) %>% 
+  group_by(Dx) %>% 
   dplyr::summarise(count = n()) %>% 
-  spread(Tissue, count) %>%
   kable()
-
-METADATA = METADATA %>%
-  dplyr::mutate(Dx.Tissue = paste(Dx, Tissue, sep = '.'))
 ```
@@
 ```{r covariates.clustering}
-FactorCovariates <- c('Individual_ID', "Institution", "Reported_Gender", "LibraryBatch", "Dx.Tissue", "RibozeroBatch", "FlowcellBatch")
+FactorCovariates <- c("Institution", "Reported_Gender", "LibraryBatch", "Dx", "RibozeroBatch", "FlowcellBatch")
 ContCovariates <- c("ageOfDeath", "PMI", "RIN", "MappedReads", "IntragenicRate", "IntronicRate", "IntergenicRate",
@@
 p = list()
-p[[1]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = RIN)) + geom_boxplot()
+p[[1]] = ggplot(COVARIATES, aes(x = Dx, y = RIN)) + geom_boxplot()
 p[[1]] = p[[1]] + ggtitle('RIN') + my.theme
@@
 # Age of Death
-p[[2]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = ageOfDeath)) + geom_boxplot()
+p[[2]] = ggplot(COVARIATES, aes(x = Dx, y = ageOfDeath)) + geom_boxplot()
 p[[2]] = p[[2]] + ggtitle('AgeOfDeath') + my.theme
@@
 # PMI
-p[[3]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = PMI)) + geom_boxplot()
+p[[3]] = ggplot(COVARIATES, aes(x = Dx, y = PMI)) + geom_boxplot()
 p[[3]] = p[[3]] + ggtitle('PMI (in hours)') + my.theme
@@
 # Intronic Rate
-p[[4]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = IntronicRate)) + geom_boxplot()
+p[[4]] = ggplot(COVARIATES, aes(x = Dx, y = IntronicRate)) + geom_boxplot()
 p[[4]] = p[[4]] + ggtitle('Intronic Rate') + my.theme
@@
 # IntergenicRate
-p[[5]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = IntergenicRate)) + geom_boxplot()
+p[[5]] = ggplot(COVARIATES, aes(x = Dx, y = IntergenicRate)) + geom_boxplot()
 p[[5]] = p[[5]] + ggtitle('Intergenic Rate') + my.theme
@@
 # Transcripts Detected
-p[[6]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = IntragenicRate)) + geom_boxplot()
+p[[6]] = ggplot(COVARIATES, aes(x = Dx, y = IntragenicRate)) + geom_boxplot()
 p[[6]] = p[[6]] + ggtitle('Intragenic Rate') + my.theme
@@
 # Mapped Reads
-p[[7]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = MappedReads)) + geom_boxplot()
+p[[7]] = ggplot(COVARIATES, aes(x = Dx, y = MappedReads)) + geom_boxplot()
 p[[7]] = p[[7]] + ggtitle('Mapped Reads') + my.theme
@@
 # PercentAligned
-p[[8]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = TotalReads)) + geom_boxplot()
+p[[8]] = ggplot(COVARIATES, aes(x = Dx, y = TotalReads)) + geom_boxplot()
 p[[8]] = p[[8]] + ggtitle('Total Reads') + my.theme
@@
 # rRNARate
-p[[9]] = ggplot(COVARIATES, aes(x = Dx.Tissue, y = rRNARate)) + geom_boxplot()
+p[[9]] = ggplot(COVARIATES, aes(x = Dx, y = rRNARate)) + geom_boxplot()
 p[[9]] = p[[9]] + ggtitle('rRNARate') + my.theme
@@
 multiplot(plotlist = p, cols = 3)
-
-# Institution
-# vcd::mosaic(~ Institution + Dx.Tissue, data = COVARIATES)
-
-# Gender
-# vcd::mosaic(~ Gender + Dx.Tissue, data = COVARIATES)
 ```
 ### Filter genes
-Remove genes that have less than 1 cpm counts in at least 50% of samples per Dx.Tissue. Also remove genes with missing gene length and percentage GC content
+Remove genes that have less than 1 cpm counts in at least 50% of samples per Dx. Also remove genes with missing gene length and percentage GC content
 ```{r cpmnormalisation}
@@
 
-genesToAnalyze = dlply(METADATA, .(Dx.Tissue, Reported_Gender), .fun = function(mtd, count){
+genesToAnalyze = dlply(METADATA, .(Dx, Reported_Gender), .fun = function(mtd, count){
   processed.counts = getGeneFilteredGeneExprMatrix(count[,mtd$SampleID],
@@
 plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, "SampleID")) %>%
-    tidyr::separate(Dx.Tissue, c('Dx','Tissue'), sep = '_') %>%
   dplyr::mutate(label = SampleID) %>% 
@@
   left_join(COVARIATES %>%
-              rownameToFirstColumn('SampleID') %>%
-              tidyr::separate(Dx.Tissue, c('Dx', 'Tissue'), sep = '_'))
+              rownameToFirstColumn('SampleID'))
 
@@
            aes(x = logCPM, color = SampleID)) + geom_density() 
-p = p + theme(legend.position = 'top') + facet_grid(Dx+.~Tissue, scale = 'free')
+p = p + theme(legend.position = 'top') + facet_grid(Dx, scale = 'free')
 p
@@
 
-tmp = tidyr::separate(COVARIATES, Dx.Tissue, c('Dx', 'Tissue'), sep = '_') %>%
-  group_by(Dx, Tissue) %>%
-  dplyr::summarise(count = n()) %>%
-  spread(Tissue, count)
+tmp = group_by(COVARIATES, Dx) %>%
+  dplyr::summarise(count = n())
 
@@
 plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>% 
-  tidyr::separate(Dx.Tissue, c('Dx','Tissue'), sep = '_')
+  dplyr::mutate(Tissue = "ACC")
 
@@
 # Eucledian tree based analysis
-COVARIATES.tmp = data.matrix(COVARIATES[,c("Institution", "Reported_Gender", "LibraryBatch", "Dx.Tissue")])
+COVARIATES.tmp = data.matrix(COVARIATES[,c("Institution", "Reported_Gender", "LibraryBatch", "Dx")])
 COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0
@@
 p = ggplot(tmp1, aes(x = logCPM, color = SampleID)) + geom_density() 
-p = p + theme_bw() %+replace% theme(legend.position = 'NONE') + facet_grid(.~Dx.Tissue, scale = 'free')
+p = p + theme_bw() %+replace% theme(legend.position = 'NONE') + facet_grid(.~Dx, scale = 'free')
 p
@@
 # Fit a base model with the following covariates
-base_covars <- "Dx.Tissue"
-random_effect <- "Individual_ID"
+base_covars <- "Dx"
 
 # Set base formula
-base_formula <- glue::glue("~ ", glue::glue_collapse(base_covars, sep = " + ")) %>%
-  glue::glue(" + (1|{random_effect})")
+base_formula <- glue::glue("~ ", glue::glue_collapse(base_covars, sep = " + "))
 
@@
 
-  tissue	= 'dorsolateral prefrontal cortex, anterior cingulate cortex', 
+  tissue	= 'anterior cingulate cortex', 
   study = 'CMC',
@@
 # Code
-CODE <- Folder(name = "MSSM-Penn-Pitt - ACC and DLPFC - DREAM - FEATURECOUNTS - GRCh38 - CQN", parentId = parentId)
+CODE <- Folder(name = "MSSM-Penn-Pitt - ACC - DREAM - FEATURECOUNTS - GRCh38 - CQN", parentId = parentId)
 CODE <- synStore(CODE)
@@
 COVARIATES = rownameToFirstColumn(COVARIATES, 'SampleID')
-write.table(COVARIATES, file = 'CMC_ACC_DLPFC_Covariates.tsv', sep = '\t', row.names=F, quote=F)
-COV_OBJ = synapser::File('CMC_ACC_DLPFC_Covariates.tsv', name = 'Covariates', parentId = CODE$properties$id)
+write.table(COVARIATES, file = 'CMC_ACC_Covariates.tsv', sep = '\t', row.names=F, quote=F)
+COV_OBJ = synapser::File('CMC_ACC_Covariates.tsv', name = 'Covariates', parentId = CODE$properties$id)
 COV_OBJ = synStore(COV_OBJ, used = ALL_USED_IDs, activityName = activityName, 
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_ACC_DLPFC_Counts.tsv', sep = '\t', row.names=F, quote=F)
-COUNT_OBJ = File('CMC_ACC_DLPFC_Counts.tsv', name = 'Counts (filtered raw)', parentId = CODE$properties$id)
+  write.table(file = 'CMC_ACC_Counts.tsv', sep = '\t', row.names=F, quote=F)
+COUNT_OBJ = File('CMC_ACC_Counts.tsv', name = 'Counts (filtered raw)', parentId = CODE$properties$id)
 COUNT_OBJ = synStore(COUNT_OBJ, used = ALL_USED_IDs, activityName = activityName, 
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_ACC_DLPFC_logCPM.tsv', sep = '\t', row.names=F, quote=F)
-LCOUNT_OBJ = File('CMC_ACC_DLPFC_logCPM.tsv', name = 'Counts (filtered logCPM)', parentId = CODE$properties$id)
+  write.table(file = 'CMC_ACC_logCPM.tsv', sep = '\t', row.names=F, quote=F)
+LCOUNT_OBJ = File('CMC_ACC_logCPM.tsv', name = 'Counts (filtered logCPM)', parentId = CODE$properties$id)
 LCOUNT_OBJ = synStore(LCOUNT_OBJ, used = ALL_USED_IDs, activityName = activityName, 
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_ACC_DLPFC_offset.tsv', sep = '\t', row.names=F, quote=F)
-OFFSET_OBJ = File('CMC_ACC_DLPFC_offset.tsv', name = 'Gene length and GC content offset', parentId = CODE$properties$id)
+  write.table(file = 'CMC_ACC_offset.tsv', sep = '\t', row.names=F, quote=F)
+OFFSET_OBJ = File('CMC_ACC_offset.tsv', name = 'Gene length and GC content offset', parentId = CODE$properties$id)
 OFFSET_OBJ = synStore(OFFSET_OBJ, used = ALL_USED_IDs, activityName = activityName, 

### 2020-05-04 individual-cohorts/MSSM-Penn-Pitt_ACC_FEATURECOUNTS_GRCh38_CQN.Rmd
fix parent id

@@
 ```{r synapse.parameters, include=FALSE, cache=FALSE}
-parentId = 'syn21435176';
+parentId = 'syn22012942';
 activityName = 'Covariate adjustments';

### 2020-05-04 individual-cohorts/MSSM-Penn-Pitt_ACC_FEATURECOUNTS_GRCh38_CQN.Rmd
fix knit2synapse

@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "MSSM-Penn-Pitt_ACC_FEATURECOUNTS_GRCh38_CQN.Rmd",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/MSSM-Penn-Pitt_ACC_FEATURECOUNTS_GRCh38_CQN.Rmd",
                                  parentId ="syn22012942",

### 2020-05-04 individual-cohorts/MSSM-Penn-Pitt_ACC_FEATURECOUNTS_GRCh38_CQN.Rmd
fix plots where Tissue called

@@
   dplyr::mutate(label = SampleID) %>% 
-  dplyr::mutate(label = ifelse((label %in% outliers), label, NA))
+  dplyr::mutate(label = ifelse((label %in% outliers), label, NA)) %>% 
+  dplyr::mutate(Tissue = "ACC")
 
@@
            aes(x = logCPM, color = SampleID)) + geom_density() 
-p = p + theme(legend.position = 'top') + facet_grid(Dx, scale = 'free')
+p = p + theme(legend.position = 'top') + facet_grid("Dx", scale = 'free')
 p
@@
                                                              "scale(EV.2)", "scale(EV.3)", "scale(EV.4)", "scale(EV.5)")))
-phase1
+kable(phase1)
 ```
@@
                          variables = array(c("(1|LibraryBatch)", "(1|FlowcellBatch)")))
-phase2
+kable(phase2)
 ```
@@
                                              "scale(TotalReads)", "scale(GenesDetected)", "scale(MappedReads)")))
-phase3
+kable(phase3)
 ```

### 2020-05-12 individual-cohorts/MSSM-Penn-Pitt_ACC_FEATURECOUNTS_GRCh38_CQN.Rmd
covar selection with mvIC

@@
 library(githubr) # get the package from devtools::install_github('brian-bot/githubr')
+library(mvIC) # devtools::install_github("GabrielHoffman/mvIC", repos=BiocManager::repositories())
 
@@
 # Stepwise regression
-phase1 <- mvBIC::mvForwardStepwise(exprObj = CQN.GENE_EXPRESSION$E,
+phase1 <- mvIC::mvForwardStepwise(exprObj = CQN.GENE_EXPRESSION$E,
                          baseFormula = base_formula,
@@
 # Fit a base model with the following covariates
-phase2 <- mvBIC::mvForwardStepwise(exprObj = CQN.GENE_EXPRESSION$E,
+phase2 <- mvIC::mvForwardStepwise(exprObj = CQN.GENE_EXPRESSION$E,
                          baseFormula = phase1$formula, 
@@
 # Fit a base model with the following covariates
-phase3 <- mvBIC::mvForwardStepwise(exprObj = CQN.GENE_EXPRESSION$E,
+phase3 <- mvIC::mvForwardStepwise(exprObj = CQN.GENE_EXPRESSION$E,
                          baseFormula = phase2$formula, 

### 2020-05-14 individual-cohorts/MSSM-Penn-Pitt_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd
test DLPFC cohort

--- 2020-05-12 individual-cohorts/MSSM-Penn-Pitt_ACC_FEATURECOUNTS_GRCh38_CQN.Rmd
@@
 ---
-title: "MSSM-Penn-Pitt - ACC - DREAM - FEATURECOUNTS - GRCh38 - CQN - Model Selection"
+title: "MSSM-Penn-Pitt - DLPFC - DREAM - FEATURECOUNTS - GRCh38 - CQN - Model Selection"
 author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/MSSM-Penn-Pitt_ACC_FEATURECOUNTS_GRCh38_CQN.Rmd",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/MSSM-Penn-Pitt_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd",
                                  parentId ="syn22012942",
-                                 folderName = "MSSM-Penn-Pitt - ACC - DREAM - FEATURECOUNTS - GRCh38 - CQN")
+                                 folderName = "MSSM-Penn-Pitt - DLPFC - DREAM - FEATURECOUNTS - GRCh38 - CQN")
 ```
@@
 activityName = 'Covariate adjustments';
-activityDescription = 'Covariate analysis of GRCh38 feature counts with CQN normalisation (ACC)';
+activityDescription = 'Covariate analysis of GRCh38 feature counts with CQN normalisation (DLPFC)';
 
-thisFileName <- 'MSSM-Penn-Pitt_ACC_FEATURECOUNTS_GRCh38_CQN.Rmd'
+thisFileName <- 'MSSM-Penn-Pitt_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd'
 
@@
   dplyr::filter(Institution %in% c("MSSM", "Penn", "Pitt")) %>% 
-  dplyr::filter(Tissue %in% c("ACC")) %>% 
+  dplyr::filter(Tissue %in% c("DLPFC")) %>% 
   dplyr::filter(SampleID %in% colnames(counts), !is.na(SampleID)) %>%
@@
   dplyr::mutate(label = ifelse((label %in% outliers), label, NA)) %>% 
-  dplyr::mutate(Tissue = "ACC")
+  dplyr::mutate(Tissue = "DLPFC")
 
@@
 plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>% 
-  dplyr::mutate(Tissue = "ACC")
+  dplyr::mutate(Tissue = "DLPFC")
 
@@
                                                              "scale(EV.2)", "scale(EV.3)", "scale(EV.4)", "scale(EV.5)")))
-kable(phase1)
+kable(phase1$settings)
+kable(phase1$trace)
+as.character(phase1$formula)[2]
 ```
@@
                          variables = array(c("(1|LibraryBatch)", "(1|FlowcellBatch)")))
-kable(phase2)
+kable(phase2$settings)
+kable(phase2$trace)
+as.character(phase2$formula)[2]
 ```
@@
                                              "scale(TotalReads)", "scale(GenesDetected)", "scale(MappedReads)")))
-kable(phase3)
+kable(phase3$settings)
+kable(phase3$trace)
+as.character(phase3$formula)[2]
 ```
@@
 
-The model for clinical, technical and batch variables selected after phase I, II and III is `r phase3$formula`
+The model for clinical, technical and batch variables selected after phase I, II and III is `r as.character(phase3$formula)[2]`
 
@@
 # Code
-CODE <- Folder(name = "MSSM-Penn-Pitt - ACC - DREAM - FEATURECOUNTS - GRCh38 - CQN", parentId = parentId)
+CODE <- Folder(name = "MSSM-Penn-Pitt - DLPFC - DREAM - FEATURECOUNTS - GRCh38 - CQN", parentId = parentId)
 CODE <- synStore(CODE)
@@
 COVARIATES = rownameToFirstColumn(COVARIATES, 'SampleID')
-write.table(COVARIATES, file = 'CMC_ACC_Covariates.tsv', sep = '\t', row.names=F, quote=F)
-COV_OBJ = synapser::File('CMC_ACC_Covariates.tsv', name = 'Covariates', parentId = CODE$properties$id)
+write.table(COVARIATES, file = 'CMC_DLPFC_Covariates.tsv', sep = '\t', row.names=F, quote=F)
+COV_OBJ = synapser::File('CMC_DLPFC_Covariates.tsv', name = 'Covariates', parentId = CODE$properties$id)
 COV_OBJ = synStore(COV_OBJ, used = ALL_USED_IDs, activityName = activityName, 
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_ACC_Counts.tsv', sep = '\t', row.names=F, quote=F)
-COUNT_OBJ = File('CMC_ACC_Counts.tsv', name = 'Counts (filtered raw)', parentId = CODE$properties$id)
+  write.table(file = 'CMC_DLPFC_Counts.tsv', sep = '\t', row.names=F, quote=F)
+COUNT_OBJ = File('CMC_DLPFC_Counts.tsv', name = 'Counts (filtered raw)', parentId = CODE$properties$id)
 COUNT_OBJ = synStore(COUNT_OBJ, used = ALL_USED_IDs, activityName = activityName, 
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_ACC_logCPM.tsv', sep = '\t', row.names=F, quote=F)
-LCOUNT_OBJ = File('CMC_ACC_logCPM.tsv', name = 'Counts (filtered logCPM)', parentId = CODE$properties$id)
+  write.table(file = 'CMC_DLPFC_logCPM.tsv', sep = '\t', row.names=F, quote=F)
+LCOUNT_OBJ = File('CMC_DLPFC_logCPM.tsv', name = 'Counts (filtered logCPM)', parentId = CODE$properties$id)
 LCOUNT_OBJ = synStore(LCOUNT_OBJ, used = ALL_USED_IDs, activityName = activityName, 
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_ACC_offset.tsv', sep = '\t', row.names=F, quote=F)
-OFFSET_OBJ = File('CMC_ACC_offset.tsv', name = 'Gene length and GC content offset', parentId = CODE$properties$id)
+  write.table(file = 'CMC_DLPFC_offset.tsv', sep = '\t', row.names=F, quote=F)
+OFFSET_OBJ = File('CMC_DLPFC_offset.tsv', name = 'Gene length and GC content offset', parentId = CODE$properties$id)
 OFFSET_OBJ = synStore(OFFSET_OBJ, used = ALL_USED_IDs, activityName = activityName, 

### 2020-05-14 individual-cohorts/HBCC_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd
add hbcc dlpfc cohort

--- 2020-04-26 HBCC_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd
@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "HBCC_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd",
-                                          parentId ="syn21435176",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/HBCC_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd",
+                                          parentId ="syn22012942",
                                           folderName = 'HBCC - DLPFC - DREAM - FEATURECOUNTS - GRCh38 - CQN')
@@
 library(stringr)
-library(mvBIC)
 
@@
 ```{r synapse.parameters, include=FALSE, cache=TRUE}
-parentId = 'syn21435176';
+parentId = 'syn22012942';
 activityName = 'Covariate adjustments';
@@
 thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
-thisFile <- getPermlink(repository = thisRepo, repositoryPath = thisFileName)
+thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/',thisFileName))
 ```
@@
                                              "scale(EV.2)", "scale(EV.3)", "scale(EV.4)", "scale(EV.5)")))
-phase1
+kable(phase1$settings)
+kable(phase1$trace)
+as.character(phase1$formula)[2]
 ```
@@
                          variables = array(c("(1|LibraryBatch)", "(1|FlowcellBatch)")))
-phase2
+kable(phase2$settings)
+kable(phase2$trace)
+as.character(phase2$formula)[2]
 ```
@@
                                              "scale(TotalReads)", "scale(GenesDetected)", "scale(MappedReads)")))
-phase3
+kable(phase3$settings)
+kable(phase3$trace)
+as.character(phase3$formula)[2]
 ```
@@
 
-The model for clinical, technical and batch variables selected after phase I, II and III is `r paste(phase3$formula)`
+The model for clinical, technical and batch variables selected after phase I, II and III is:
+```{r selection_results}
+as.character(phase3$formula)[2]
+```
 

### 2020-05-15 individual-cohorts/HBCC_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd
fails if ageOfDeath scaled

@@
                          variables = array(c("(1|Reported_Gender)", "scale(PMI)",
-                                             "scale(ageOfDeath)", "scale(RIN)", "scale(RIN2)", "scale(EV.1)",
+                                             "ageOfDeath", "scale(RIN)", "scale(RIN2)", "scale(EV.1)",
                                              "scale(EV.2)", "scale(EV.3)", "scale(EV.4)", "scale(EV.5)")))

### 2020-05-16 individual-cohorts/HBCC_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd
fails with random effects

@@
                          data = COVARIATES,
-                         variables = array(c("(1|LibraryBatch)", "(1|FlowcellBatch)")))
+                         variables = array(c("LibraryBatch", "FlowcellBatch")))
 kable(phase2$settings)

### 2020-05-18 individual-cohorts/MSSM-Penn-Pitt_ACC_DE.Rmd
test DE of MPP ACC

--- 2020-04-27 MSSM-Penn-Pitt_DE_DxTissue.Rmd
@@
 ---
-title: "Differential Expression - MSSM-Penn-Pitt - ACC DLPFC - Dx.Tissue"
+title: "Differential Expression - MSSM-Penn-Pitt - ACC - Dx"
 author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "MSSM-Penn-Pitt_DE_DxTissue.Rmd",
-                                 parentId ="syn21824667",
-                                 folderName = "Differential Expression - Dx.Tissue")
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/MSSM-Penn-Pitt_ACC_DE.Rmd",
+                                 parentId ="syn22012953",
+                                 folderName = "Differential Expression - Dx")
 ```
@@
 library(biomaRt)
-library(variancePartition) #1.17.9
+library(variancePartition)
 library(edgeR)
@@
 ```{r synapse.parameters, include=FALSE, cache=TRUE}
-parentId = 'syn21824667';
+parentId = 'syn22012953';
 activityName = 'Differential Expression';
-activityDescription = 'Differential Expression - Dx.Tissue';
+activityDescription = 'Differential Expression - Dx';
 
-thisFileName <- 'MSSM-Penn-Pitt_DE_DxTissue.Rmd'
+thisFileName <- 'MSSM-Penn-Pitt_ACC_DE_Dx.Rmd'
 
@@
 # Download CQN counts
-CQN_ID = "syn21825509"
+CQN_ID = "syn22013154"
 ALL_USED_IDs = c(CQN_ID)
@@
 # Download filtered counts
-COUNTS_ID = "syn21825508"
+COUNTS_ID = "syn22013153"
 ALL_USED_IDs = c(ALL_USED_IDs, COUNTS_ID)
@@
 # Get covariates
-COVAR_ID = "syn21825507"
+COVAR_ID = "syn22013150"
 ALL_USED_IDs = c(ALL_USED_IDs, COVAR_ID)
@@
 # Set variable class
-FactorCovariates <- c("Individual_ID", "Institution", "Reported_Gender", "LibraryBatch", "Dx.Tissue", "RibozeroBatch", "FlowcellBatch")
+FactorCovariates <- c("Individual_ID", "Institution", "Reported_Gender", "LibraryBatch", "Dx", "RibozeroBatch", "FlowcellBatch")
 COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
@@
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Dx.Tissue", "RIN2", "ageOfDeath", "EV.1", "PMI", "EV.4","EV.2","RIN", "EV.3", "EV.5", "IntronicRate")
-random_effect <- c("Individual_ID", "Institution", "Reported_Gender")
+adjust.covars <- c("Dx", "ageOfDeath", "RIN2", "RIN", "rRNARate", "IntronicRate", "IntragenicRate", "IntergenicRate")
+random_effect <- c("Institution", "Reported_Gender")
 ```
 
-### Normalisation - Dx.Tissue
-1. Dx.Tissue is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis and tissue)
-2. Individual_ID is chosen as random effect
+### Normalisation - Dx
+1. Dx is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis)
 
@@
 
-### Tissue Correlation
-```{r tissue_correlation}
-hasReps = rownameToFirstColumn(COVARIATES, "SampleID") %>%
-  dplyr::select(Individual_ID) %>% 
-  table
-
-df_dlpfc = rownameToFirstColumn(COVARIATES, "SampleID") %>% 
-  filter( Individual_ID %in% names(hasReps[hasReps==2])) %>%
-  filter(grepl("DLPFC", Dx.Tissue)) %>% 
-  dplyr::select(SampleID, Individual_ID, Dx.Tissue)
-df_acc = rownameToFirstColumn(COVARIATES, "SampleID") %>%
-  filter( Individual_ID %in% names(hasReps[hasReps==2])) %>%
-  filter(grepl("ACC", Dx.Tissue)) %>% 
-  dplyr::select(SampleID, Individual_ID, Dx.Tissue)
-
-df_merge = merge(df_dlpfc, df_acc, by="Individual_ID")
-
-res_std <- rownameToFirstColumn(RESIDUAL.GENE_EXPRESSION, "ensembl_gene_id")
-
-corRes = sapply(res_std$ensembl_gene_id, function(ensGene){
-
-  x = VOOM.GENE_EXPRESSION$E[ensGene,df_merge$SampleID.x,drop=FALSE]
-  y = VOOM.GENE_EXPRESSION$E[ensGene,df_merge$SampleID.y,drop=FALSE]
-
-  cor(t(x)[,1], t(y)[,1], use="pairwise.complete.obs")
-})
-
-hist(corRes, xlab="Correlation", main="Correlation between ACC and DLPFC sample from the sample individual for each gene from ")
-
-pca = prcomp(t(VOOM.GENE_EXPRESSION$E))
-
-df_pca = merge(pca$x, as.data.frame(rownameToFirstColumn(COVARIATES, "SampleID")), 
-               by.x="row.names", 
-               by.y="SampleID" ) %>% 
-  tidyr::separate(Dx.Tissue, into = c("Dx", "Tissue"), sep = "_")
-
-ggplot(df_pca, aes(PC1, PC2, color=Tissue, shape=Institution)) + geom_point() + theme_bw(17) + theme(aspect.ratio=1)
-```
-
 ### Residual calculation
-Calculate weighted residuals and add back "Dx.Tissue" to the residuals
+Calculate weighted residuals and add back "Dx" to the residuals
 ```{r varsToAddBack}
 # Add variable of interest back to the residuals
-varsToAddIn = grep("Dx.Tissue", colnames(ADJUSTED.FIT$design), value = T)
+varsToAddIn = grep("Dx", colnames(ADJUSTED.FIT$design), value = T)
 RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION +
@@
 # Eucledian tree based analysis
-COVARIATES.tmp = data.matrix(COVARIATES[,c('Dx.Tissue','Institution')])
+COVARIATES.tmp = data.matrix(COVARIATES[,c('Dx','Institution')])
 
@@
                            main = "Sample dendrogram",
-                           groupLabels = colnames(COVARIATES.tmp[,c('Dx.Tissue','Institution')]))
+                           groupLabels = colnames(COVARIATES.tmp[,c('Dx','Institution')]))
 ```
@@
 ```{r set-formula}
+# remove RIN2 as RIN and RIN2 are highly correlated
+adjust.covars <- c("Dx", "ageOfDeath", "EV.1", "PMI", "EV.4","EV.2","RIN", "EV.3", "EV.5", "IntronicRate")
 formula <- glue::glue("~ 0 + ", glue::glue_collapse(adjust.covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))    
-
 ```
-
+#drop redundancy - RIN2 or RIN
+#use canCorPairs() to test colinearity, don't want var of interest to be correlated
+#leave out random effects
 `r formula`
@@
                                    data = COVARIATES, 
-                                   coefficient = c("Dx.TissueOther_ACC", "Dx.TissueControl_ACC"))
+                                   coefficient = c("DxOther", "DxControl"))
 L2 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
@@
                                    data = COVARIATES, 
-                                   coefficient = c("Dx.TissueSCZ_ACC", "Dx.TissueControl_ACC"))
+                                   coefficient = c("DxSCZ", "DxControl"))
 L3 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
@@
                                    data = COVARIATES, 
-                                   coefficient = c("Dx.TissueSCZ_ACC", "Dx.TissueOther_ACC"))
-L4 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
-                                   formula = formula, 
-                                   data = COVARIATES, 
-                                   coefficient = c("Dx.TissueOther_DLPFC", "Dx.TissueControl_DLPFC"))
-L5 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
-                                   formula = formula, 
-                                   data = COVARIATES, 
-                                   coefficient = c("Dx.TissueSCZ_DLPFC", "Dx.TissueControl_DLPFC"))
-L6 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
-                                   formula = formula, 
-                                   data = COVARIATES, 
-                                   coefficient = c("Dx.TissueSCZ_DLPFC", "Dx.TissueOther_DLPFC"))
+                                   coefficient = c("DxSCZ", "DxOther"))
 
-L = cbind(L1, L2, L3, L4, L5, L6)
+L = cbind(L1, L2, L3)
 
@@
               glue::glue_collapse(names(L2[L2 != 0]), "-"),
-              glue::glue_collapse(names(L3[L3 != 0]), "-"),
-              glue::glue_collapse(names(L4[L4 != 0]), "-"),
-              glue::glue_collapse(names(L5[L5 != 0]), "-"),
-              glue::glue_collapse(names(L6[L6 != 0]), "-")
+              glue::glue_collapse(names(L3[L3 != 0]), "-")
 )
@@
                                      data = COVARIATES,
-                                     L = L)
+                                     L = L,
+                                     suppressWarnings = TRUE)
 
@@
 # Get differential expression
-DE = lapply(glue::glue("L{1:6}"), function(i, FIT){
+DE = lapply(glue::glue("L{1:3}"), function(i, FIT){
   topTable(FIT, coef=i, number = Inf) %>%
@@
   rbindlist(idcol = 'Comparison') %>%
-  dplyr::mutate(Comparison = gsub('Dx.Tissue','',Comparison),
+  dplyr::mutate(Comparison = gsub('Dx','',Comparison),
                 Direction = logFC/abs(logFC),
@@
   tidyr::separate(Comparison, into = c('ref.state','to.state'), sep = '-') %>%
-  tidyr::separate(ref.state, into = c('Dx1','Tissue'), sep = '_') %>%
-  tidyr::separate(to.state, into = c('Dx2','Tissue'), sep = '_') %>%
-  tidyr::unite(Comparison, Dx1, Dx2, sep = '_vs_') %>%
+  tidyr::unite(Comparison, `ref.state`, `to.state`, sep = '_vs_') %>%
   left_join(Ensemble2HGNC) %>%
@@
   dplyr::filter(adj.P.Val <= 0.05) %>%
-  dplyr::select(ensembl_gene_id, Comparison, Tissue) %>%
-  group_by(Comparison, Tissue) %>%
+  dplyr::select(ensembl_gene_id, Comparison) %>%
+  group_by(Comparison) %>%
   dplyr::summarise(FDR_0_05 = length(unique(ensembl_gene_id)))
@@
   dplyr::filter(adj.P.Val <= 0.05, abs(logFC) >= log2(1.2)) %>%
-  dplyr::select(ensembl_gene_id, Comparison, Tissue) %>%
-  group_by(Comparison, Tissue) %>%
+  dplyr::select(ensembl_gene_id, Comparison) %>%
+  group_by(Comparison) %>%
   dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(ensembl_gene_id)))
@@
 p = p + scale_color_manual(values = c('green','grey','red')) + theme_bw() %+replace% theme(legend.position = 'top')
-p = p + facet_grid(Tissue+.~Comparison, scales = 'fixed')
+p = p + facet_grid(.~Comparison, scales = 'fixed')
 p
@@
 # Code
-CODE <- Folder(name = "Differential Expression - Dx.Tissue", parentId = parentId)
+CODE <- Folder(name = "Differential Expression - Dx", parentId = parentId)
 CODE <- synStore(CODE)
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_ACC_DLPFC_Dx.Tissue_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
-nEXP_OBJ = File('CMC_ACC_DLPFC_Dx.Tissue_netResidualExpression.tsv', 
-                name = 'Dx.Tissue Normalised, covariates removed residual expression (for network analysis)', 
+  write.table(file = 'CMC_ACC_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
+nEXP_OBJ = File('CMC_ACC_Dx_netResidualExpression.tsv', 
+                name = 'Dx Normalised, covariates removed residual expression (for network analysis)', 
                 parentId = CODE$properties$id)
@@
 # Store differential expression results
-write.table(DE, file = 'CMC_ACC_DLPFC_Dx.Tissue_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
-DEXP_OBJ = File('CMC_ACC_DLPFC_Dx.Tissue_DiffExpression.tsv', 
-                name = 'Differential Expression Results (Dx.Tissue)', 
+write.table(DE, file = 'CMC_ACC_Dx_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
+DEXP_OBJ = File('CMC_ACC_Dx_DiffExpression.tsv', 
+                name = 'Differential Expression Results (Dx)', 
                 parentId = CODE$properties$id)
@@
 
-# Store adjusted fit linear mixed model object as RDS - Dx.Tissue
-saveRDS(ADJUSTED.FIT, file = 'CMC_ACC_DLPFC_Dx.Tissue_dream-output.rds')
-ADJ_OBJ = File('CMC_ACC_DLPFC_Dx.Tissue_dream-output.rds',
-               name = 'Dream Output object - Dx.Tissue', 
+# Store adjusted fit linear mixed model object as RDS - Dx
+saveRDS(ADJUSTED.FIT, file = 'CMC_ACC_Dx_dream-output.rds')
+ADJ_OBJ = File('CMC_ACC_Dx_dream-output.rds',
+               name = 'Dream Output object - Dx', 
                parentId = CODE$properties$id)

### 2020-05-18 individual-cohorts/MSSM-Penn-Pitt_ACC_DE.Rmd
rm tissue facet

@@
 pl[[1]] = pl[[1]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = log10(Length), y = logFC))
-pl[[1]] = pl[[1]] + facet_grid(Tissue~.) + scale_color_manual(values = c('green','grey','red'))
+pl[[1]] = pl[[1]] + scale_color_manual(values = c('green','grey','red'))
 pl[[1]] = pl[[1]] + theme(legend.position = 'top')
@@
 pl[[2]] = pl[[2]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = percentage_gene_gc_content, y = logFC))
-pl[[2]] = pl[[2]] + facet_grid(Tissue~.) + scale_color_manual(values = c('green','grey','red'))
+pl[[2]] = pl[[2]] + scale_color_manual(values = c('green','grey','red'))
 pl[[2]] = pl[[2]] + theme(legend.position = 'top')
@@
 pl[[3]] = pl[[3]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = AveExpr, y = logFC))
-pl[[3]] = pl[[3]] + facet_grid(Tissue~.) + scale_color_manual(values = c('green','grey','red'))
+pl[[3]] = pl[[3]] + scale_color_manual(values = c('green','grey','red'))
 pl[[3]] = pl[[3]] + theme(legend.position = 'top')

### 2020-05-18 individual-cohorts/MSSM-Penn-Pitt_DLPFC_DE.Rmd
test DE DLPFC MPP

--- 2020-05-18 individual-cohorts/MSSM-Penn-Pitt_ACC_DE.Rmd
@@
 ---
-title: "Differential Expression - MSSM-Penn-Pitt - ACC - Dx"
+title: "Differential Expression - MSSM-Penn-Pitt - DLPFC - Dx"
 author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/MSSM-Penn-Pitt_ACC_DE.Rmd",
-                                 parentId ="syn22012953",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/MSSM-Penn-Pitt_DLPFC_DE.Rmd",
+                                 parentId ="syn22043922",
                                  folderName = "Differential Expression - Dx")
@@
 ```{r synapse.parameters, include=FALSE, cache=TRUE}
-parentId = 'syn22012953';
+parentId = 'syn22043922';
 activityName = 'Differential Expression';
@@
 
-thisFileName <- 'MSSM-Penn-Pitt_ACC_DE_Dx.Rmd'
+thisFileName <- 'MSSM-Penn-Pitt_DLPFC_DE.Rmd'
 
@@
 # Download CQN counts
-CQN_ID = "syn22013154"
+CQN_ID = "syn22045741"
 ALL_USED_IDs = c(CQN_ID)
@@
 # Download filtered counts
-COUNTS_ID = "syn22013153"
+COUNTS_ID = "syn22045729"
 ALL_USED_IDs = c(ALL_USED_IDs, COUNTS_ID)
@@
 # Get covariates
-COVAR_ID = "syn22013150"
+COVAR_ID = "syn22045728"
 ALL_USED_IDs = c(ALL_USED_IDs, COVAR_ID)
-COVARIATES = downloadFile(COVAR_ID) %>% 
-  tibble::column_to_rownames(var = "SampleID")
+COVARIATES = downloadFile(COVAR_ID)
+
+# dropped individual id 
+MD_ID = "syn16816488"
+ALL_USED_IDs = c(ALL_USED_IDs, COUNTS_ID)
+METADATA = downloadFile(MD_ID) %>% 
+  dplyr::select(Individual_ID, Sample_RNA_ID)
+
+COVARIATES = right_join(METADATA, COVARIATES, by = c("Sample_RNA_ID" = "SampleID")) %>% 
+  tibble::column_to_rownames(var = "Sample_RNA_ID")
 
@@
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Dx", "ageOfDeath", "RIN2", "RIN", "rRNARate", "IntronicRate", "IntragenicRate", "IntergenicRate")
+adjust.covars <- c("Dx", "ageOfDeath", "RIN2", "RIN", "IntronicRate", "IntragenicRate")war
 random_effect <- c("Institution", "Reported_Gender")
@@
 ```
-#drop redundancy - RIN2 or RIN
-#use canCorPairs() to test colinearity, don't want var of interest to be correlated
-#leave out random effects
+
 `r formula`
@@
 
-  tissue	= 'anterior cingulate cortex', 
+  tissue	= 'dorsolateral prefrontal cortex', 
   study = 'CMC',
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_ACC_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
-nEXP_OBJ = File('CMC_ACC_Dx_netResidualExpression.tsv', 
+  write.table(file = 'CMC_DLPFC_Dx_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
+nEXP_OBJ = File('CMC_DLPFC_Dx_netResidualExpression.tsv', 
                 name = 'Dx Normalised, covariates removed residual expression (for network analysis)', 
@@
 # Store differential expression results
-write.table(DE, file = 'CMC_ACC_Dx_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
-DEXP_OBJ = File('CMC_ACC_Dx_DiffExpression.tsv', 
+write.table(DE, file = 'CMC_DLPFC_Dx_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
+DEXP_OBJ = File('CMC_DLPFC_Dx_DiffExpression.tsv', 
                 name = 'Differential Expression Results (Dx)', 
@@
 # Store adjusted fit linear mixed model object as RDS - Dx
-saveRDS(ADJUSTED.FIT, file = 'CMC_ACC_Dx_dream-output.rds')
-ADJ_OBJ = File('CMC_ACC_Dx_dream-output.rds',
+saveRDS(ADJUSTED.FIT, file = 'CMC_DLPFC_Dx_dream-output.rds')
+ADJ_OBJ = File('CMC_DLPFC_Dx_dream-output.rds',
                name = 'Dream Output object - Dx', 

### 2020-05-18 individual-cohorts/MSSM-Penn-Pitt_DLPFC_DE.Rmd
typo

@@
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Dx", "ageOfDeath", "RIN2", "RIN", "IntronicRate", "IntragenicRate")war
+adjust.covars <- c("Dx", "ageOfDeath", "RIN2", "RIN", "IntronicRate", "IntragenicRate")
 random_effect <- c("Institution", "Reported_Gender")

### 2020-05-19 individual-cohorts/HBCC_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd
rm individual id as random effect to debug

--- 2020-05-15 individual-cohorts/HBCC_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd
@@
 base_covars <- "Dx"
-random_effect <- "Individual_ID"
 
 # Set base formula
-base_formula <- glue::glue("~ ", glue::glue_collapse(base_covars, sep = " + ")) %>%
-  glue::glue(" + (1|{random_effect})")
+base_formula <- glue::glue("~ ", glue::glue_collapse(base_covars, sep = " + "))
 

### 2020-05-19 individual-cohorts/HBCC_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd
point counts to data freeze

@@
 ## Download reprocessed counts (DLPFC)
-counts = 'syn21754352'
+counts = 'syn21886235'
 ALL_USED_IDs[length(ALL_USED_IDs)+1] = counts

### 2020-05-19 individual-cohorts/HBCC_DLPFC_DE.Rmd
test HBCC DLPFC DE

--- 2020-05-18 individual-cohorts/MSSM-Penn-Pitt_DLPFC_DE.Rmd
@@
 ---
-title: "Differential Expression - MSSM-Penn-Pitt - DLPFC - Dx"
+title: "Differential Expression - HBCC - DLPFC - Dx"
 author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/MSSM-Penn-Pitt_DLPFC_DE.Rmd",
-                                 parentId ="syn22043922",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/HBCC_DLPFC_DE.Rmd",
+                                 parentId ="syn22050019",
                                  folderName = "Differential Expression - Dx")
@@
 ```{r synapse.parameters, include=FALSE, cache=TRUE}
-parentId = 'syn22043922';
+parentId = 'syn22050019';
 activityName = 'Differential Expression';
@@
 
-thisFileName <- 'MSSM-Penn-Pitt_DLPFC_DE.Rmd'
+thisFileName <- 'HBCC_DLPFC_DE.Rmd'
 
@@
 # Download CQN counts
-CQN_ID = "syn22045741"
+CQN_ID = "syn22082460"
 ALL_USED_IDs = c(CQN_ID)
@@
 # Download filtered counts
-COUNTS_ID = "syn22045729"
+COUNTS_ID = "syn22082459"
 ALL_USED_IDs = c(ALL_USED_IDs, COUNTS_ID)
@@
 # Get covariates
-COVAR_ID = "syn22045728"
+COVAR_ID = "syn22082458"
 ALL_USED_IDs = c(ALL_USED_IDs, COVAR_ID)
@@
 # Set variable class
-FactorCovariates <- c("Individual_ID", "Institution", "Reported_Gender", "LibraryBatch", "Dx", "RibozeroBatch", "FlowcellBatch")
+FactorCovariates <- c("Individual_ID", "Reported_Gender", "LibraryBatch", "Dx", "RibozeroBatch", "FlowcellBatch")
 COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
@@
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Dx", "ageOfDeath", "RIN2", "RIN", "IntronicRate", "IntragenicRate")
-random_effect <- c("Institution", "Reported_Gender")
+adjust.covars <- c("Dx", "ageOfDeath", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate")
+random_effect <- c("Reported_Gender")
 ```
@@
 p <- ggplot(plotdata, aes(x=PC1, y=PC2))
-p <- p + geom_point(aes(color=Institution, shape=Tissue, size=ageOfDeath))
+p <- p + geom_point(aes(shape=Tissue, size=ageOfDeath))
 p <- p + theme_bw() + theme(legend.position="right") + facet_grid(Tissue~., scales = 'free_y')
@@
 # Eucledian tree based analysis
-COVARIATES.tmp = data.matrix(COVARIATES[,c('Dx','Institution')])
+COVARIATES.tmp = data.matrix(COVARIATES[,c('Dx')])
 
@@
                            main = "Sample dendrogram",
-                           groupLabels = colnames(COVARIATES.tmp[,c('Dx','Institution')]))
+                           groupLabels = colnames(COVARIATES.tmp[,c('Dx')]))
 ```
@@
 ```{r set-formula}
-# remove RIN2 as RIN and RIN2 are highly correlated
-adjust.covars <- c("Dx", "ageOfDeath", "EV.1", "PMI", "EV.4","EV.2","RIN", "EV.3", "EV.5", "IntronicRate")
 formula <- glue::glue("~ 0 + ", glue::glue_collapse(adjust.covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))    
@@
 # Get gene lengths
-counts = 'syn21867938'
+counts = 'syn21886235'
 ALL_USED_IDs = counts

### 2020-05-19 individual-cohorts/MSSM-Penn-Pitt_ACC_DE.Rmd
fix adjust.covars in DE analysis

@@
 
-thisFileName <- 'MSSM-Penn-Pitt_ACC_DE_Dx.Rmd'
+thisFileName <- 'MSSM-Penn-Pitt_ACC_DE.Rmd'
 
@@
 ALL_USED_IDs = c(ALL_USED_IDs, COVAR_ID)
-COVARIATES = downloadFile(COVAR_ID) %>% 
-  tibble::column_to_rownames(var = "SampleID")
+COVARIATES = downloadFile(COVAR_ID)
+
+# dropped individual id 
+MD_ID = "syn16816488"
+ALL_USED_IDs = c(ALL_USED_IDs, COUNTS_ID)
+METADATA = downloadFile(MD_ID) %>% 
+  dplyr::select(Individual_ID, Sample_RNA_ID)
+
+COVARIATES = right_join(METADATA, COVARIATES, by = c("Sample_RNA_ID" = "SampleID")) %>% 
+  tibble::column_to_rownames(var = "Sample_RNA_ID")
 
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_ACC_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
+  write.table(file = 'CMC_ACC_Dx_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
 nEXP_OBJ = File('CMC_ACC_Dx_netResidualExpression.tsv', 

### 2020-05-19 individual-cohorts/HBCC_DLPFC_DE.Rmd
no vars dropped

@@
 COVARIATES = downloadFile(COVAR_ID)
-
-# dropped individual id 
-MD_ID = "syn16816488"
-ALL_USED_IDs = c(ALL_USED_IDs, COUNTS_ID)
-METADATA = downloadFile(MD_ID) %>% 
-  dplyr::select(Individual_ID, Sample_RNA_ID)
-
-COVARIATES = right_join(METADATA, COVARIATES, by = c("Sample_RNA_ID" = "SampleID")) %>% 
-  tibble::column_to_rownames(var = "Sample_RNA_ID")
 

### 2020-05-19 individual-cohorts/HBCC_DLPFC_DE.Rmd
covar sampleid as rowname

@@
 ALL_USED_IDs = c(ALL_USED_IDs, COVAR_ID)
-COVARIATES = downloadFile(COVAR_ID)
+COVARIATES = downloadFile(COVAR_ID) %>% 
+  column_to_rownames(var = "SampleID")
 
@@
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Dx", "ageOfDeath", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate")
+adjust.covars <- c("Dx", "scale(ageOfDeath)", "scale(RIN)", "scale(IntronicRate)", "scale(IntragenicRate)", "scale(IntergenicRate)")
 random_effect <- c("Reported_Gender")

### 2020-05-19 individual-cohorts/HBCC_DLPFC_DE.Rmd
fix scaling

--- 2020-05-19 individual-cohorts/HBCC_DLPFC_DE.Rmd
@@
 ALL_USED_IDs = c(ALL_USED_IDs, COVAR_ID)
-COVARIATES = downloadFile(COVAR_ID)
+COVARIATES = downloadFile(COVAR_ID) %>% 
+  tibble::column_to_rownames(var = "SampleID")
 
@@
                  glue::glue_collapse(random_effect, sep = ", ", last = " and "),' chosen as random effects'))
+
+col_type <- dplyr::select(COVARIATES, c(adjust.covars, random_effect)) %>%
+  dplyr::summarise_all(class) %>% 
+  tidyr::gather(variable, class)
+
+# Scale continuous variables
+adjust.covars = sapply(1:length(col_type$class), function(i){
+  switch(col_type$class[i], 
+    "factor" = paste0(col_type$variable[i]),
+    "numeric" = paste0('scale(', col_type$variable[i], ')')
+  )
+  })
 
@@
 
-col_type <- dplyr::select(COVARIATES, c(adjust.covars, random_effect)) %>%
-  dplyr::summarise_all(class) %>% 
-  tidyr::gather(variable, class)
-
 # Model categorical variables as random to report variance fractions
@@
     "factor" = paste0('(1|', col_type$variable[i], ')'),
-    "numeric" = col_type$variable[i])
+    "numeric" = paste0('scale(', col_type$variable[i], ')')
+  )
   })

### 2020-05-19 individual-cohorts/HBCC_DLPFC_DE.Rmd
don't drop levels

@@
 # Eucledian tree based analysis
-COVARIATES.tmp = data.matrix(COVARIATES[,c('Dx')])
+COVARIATES.tmp = data.matrix(COVARIATES[,c('Dx'), drop = FALSE])
 
@@
                            main = "Sample dendrogram",
-                           groupLabels = colnames(COVARIATES.tmp[,c('Dx')]))
+                           groupLabels = colnames(COVARIATES.tmp[,c('Dx'), drop = FALSE]))
 ```

### 2020-05-19 individual-cohorts/HBCC_sgACC_FEATURECOUNTS_GRCh38_CQN.Rmd
test HBCC sgACC cohort

--- 2020-04-27 HBCC_sgACC_FEATURECOUNTS_GRCh38_CQN.Rmd
@@
 knit2synapse::createAndKnitToFolderEntity(file = "HBCC_sgACC_FEATURECOUNTS_GRCh38_CQN.Rmd",
-                                          parentId ="syn21435176",
+                                          parentId ="syn22012942",
                                           folderName = 'HBCC - sgACC - DREAM - FEATURECOUNTS - GRCh38 - CQN')

### 2020-05-19 individual-cohorts/HBCC_sgACC_FEATURECOUNTS_GRCh38_CQN.Rmd
fix mvIC syntax

@@
                          variables = array(c("(1|LibraryBatch)", "(1|IsolationBatch)")))
-phase2
+kable(phase2$settings)
+kable(phase2$trace)
+as.character(phase2$formula)[2]
 ```
@@
 
-The model for clinical, technical and batch variables selected after phase I, II and III is `r paste(phase3$formula)`
+The model for clinical, technical and batch variables selected after phase I, II and III is `r as.character(phase3$formula)[2]`
 

### 2020-05-19 individual-cohorts/HBCC_sgACC_FEATURECOUNTS_GRCh38_CQN.Rmd
fix mvIC syntax

@@
 
-COVARIATES <- droplevels(COVARIATES)
 # Stepwise regression
@@
                          baseFormula = base_formula,
-                         data = bar,
+                         data = COVARIATES,
                          variables = array(c("(1|Reported_Gender)", "scale(PMI)",

### 2020-05-19 individual-cohorts/HBCC_dsACC_FEATURECOUNTS_GRCh38_CQN.Rmd
test HBCC dsACC

--- 2020-05-19 individual-cohorts/HBCC_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd
@@
 ---
-title: "HBCC - DLPFC - FEATURECOUNTS - GRCh38 - CQN"
+title: "HBCC - dsACC - FEATURECOUNTS - GRCh38 - CQN"
 author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/HBCC_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/HBCC_dsACC_FEATURECOUNTS_GRCh38_CQN.Rmd",
                                           parentId ="syn22012942",
-                                          folderName = 'HBCC - DLPFC - DREAM - FEATURECOUNTS - GRCh38 - CQN')
+                                          folderName = 'HBCC - dsACC - DREAM - FEATURECOUNTS - GRCh38 - CQN')
 ```
@@
 activityName = 'Covariate adjustments';
-activityDescription = 'Covariate analysis of GRCh38 feature counts with CQN normalisation (DLPFC)';
+activityDescription = 'Covariate analysis of GRCh38 feature counts with CQN normalisation (dsACC)';
 
-thisFileName <- 'HBCC_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd'
+thisFileName <- 'HBCC_dsACC_FEATURECOUNTS_GRCh38_CQN.Rmd'
 
@@
 downloadFile <- function(id){
-  fread(synGet(id)$path, data.table = F)
+  fread(synGet(id)$path, data.table = F, check.names = F)
 }
@@
 # Get RNASeq QCmetadata
-METADATA_QC_DLPFC_ID = 'syn16816488' 
-ALL_USED_IDs = METADATA_QC_DLPFC_ID
-metadata = downloadFile_version(METADATA_QC_DLPFC_ID, version = 19) %>%
+METADATA_QC_dsACC_ID = 'syn16816488' 
+ALL_USED_IDs = METADATA_QC_dsACC_ID
+metadata = downloadFile_version(METADATA_QC_dsACC_ID, version = 20) %>%
   dplyr::select(`Individual_ID`, Institution, Cohort, `Reported_Gender`, Sex, Ethnicity, ageOfDeath, `PMI_(in_hours)`, Dx, 
@@
                                       'rnaSeq_report:Transcripts_Detected')) %>% 
-  dplyr::filter(`rnaSeq_dissection:Brain_Region` %in% c("DLPFC"))
+  dplyr::filter(`rnaSeq_dissection:Brain_Region` %in% c("dsACC"))
 
-## Download reprocessed counts (DLPFC)
-counts = 'syn21886235'
+## Download reprocessed counts (dsACC)
+counts = 'syn21867866'
 ALL_USED_IDs[length(ALL_USED_IDs)+1] = counts
@@
 COUNT$Strand <- NULL
+colnames(COUNT) <- gsub("X6", "6", colnames(COUNT))
 
@@
                 IntragenicRate, IntergenicRate, IntronicRate, ExpProfEfficiency, rRNARate, TranscriptsDetected) %>%
-  dplyr::filter(Cohort %in% "NIMH-HBCC") %>% 
   dplyr::filter(SampleID %in% colnames(COUNT), !is.na(SampleID)) %>% 
-  dplyr::mutate(Dx = forcats::fct_recode(Dx, Other = "BP", Other = "undetermined", Control = "Control", SCZ = "SCZ"))
+  dplyr::mutate(Dx = forcats::fct_recode(Dx, Other = "BP", Other = "MDD", Control = "Control", SCZ = "SCZ"))
 
@@
 ```{r covariates.clustering}
-FactorCovariates <- c('Individual_ID', "Reported_Gender", "LibraryBatch", "Dx", "FlowcellBatch", "RibozeroBatch")
+FactorCovariates <- c('Individual_ID', "Reported_Gender", "IsolationBatch", "Dx")
 ContCovariates <- c("ageOfDeath", "PMI", "RIN", "MappedReads", "IntragenicRate", "IntronicRate", "IntergenicRate",
@@
   dplyr::mutate(Institution = "NIMH-HBCC") %>% 
-  dplyr::mutate(Tissue = "DLPFC")
+  dplyr::mutate(Tissue = "dsACC")
 
@@
 # Code
-CODE <- Folder(name = "HBCC - DLPFC - DREAM - FEATURECOUNTS - GRCh38 - CQN", parentId = parentId)
+CODE <- Folder(name = "HBCC - dsACC - DREAM - FEATURECOUNTS - GRCh38 - CQN", parentId = parentId)
 CODE <- synStore(CODE)
@@
 COVARIATES = rownameToFirstColumn(COVARIATES, 'SampleID')
-write.table(COVARIATES, file = 'CMC_HBCC_DLPFC_Covariates.tsv', sep = '\t', row.names=F, quote=F)
-COV_OBJ = synapser::File('CMC_HBCC_DLPFC_Covariates.tsv', name = 'Covariates', parentId = CODE$properties$id)
+write.table(COVARIATES, file = 'CMC_HBCC_dsACC_Covariates.tsv', sep = '\t', row.names=F, quote=F)
+COV_OBJ = synapser::File('CMC_HBCC_dsACC_Covariates.tsv', name = 'Covariates', parentId = CODE$properties$id)
 COV_OBJ = synStore(COV_OBJ, used = ALL_USED_IDs, activityName = activityName, 
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_HBCC_DLPFC_Counts.tsv', sep = '\t', row.names=F, quote=F)
-COUNT_OBJ = File('CMC_HBCC_DLPFC_Counts.tsv', name = 'Counts (filtered raw)', parentId = CODE$properties$id)
+  write.table(file = 'CMC_HBCC_dsACC_Counts.tsv', sep = '\t', row.names=F, quote=F)
+COUNT_OBJ = File('CMC_HBCC_dsACC_Counts.tsv', name = 'Counts (filtered raw)', parentId = CODE$properties$id)
 COUNT_OBJ = synStore(COUNT_OBJ, used = ALL_USED_IDs, activityName = activityName, 
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_HBCC_DLPFC_logCPM.tsv', sep = '\t', row.names=F, quote=F)
-LCOUNT_OBJ = File('CMC_HBCC_DLPFC_logCPM.tsv', name = 'Counts (filtered logCPM)', parentId = CODE$properties$id)
+  write.table(file = 'CMC_HBCC_dsACC_logCPM.tsv', sep = '\t', row.names=F, quote=F)
+LCOUNT_OBJ = File('CMC_HBCC_dsACC_logCPM.tsv', name = 'Counts (filtered logCPM)', parentId = CODE$properties$id)
 LCOUNT_OBJ = synStore(LCOUNT_OBJ, used = ALL_USED_IDs, activityName = activityName, 
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_HBCC_DLPFC_offset.tsv', sep = '\t', row.names=F, quote=F)
-OFFSET_OBJ = File('CMC_HBCC_DLPFC_offset.tsv', name = 'Gene length and GC content offset', parentId = CODE$properties$id)
+  write.table(file = 'CMC_HBCC_dsACC_offset.tsv', sep = '\t', row.names=F, quote=F)
+OFFSET_OBJ = File('CMC_HBCC_dsACC_offset.tsv', name = 'Gene length and GC content offset', parentId = CODE$properties$id)
 OFFSET_OBJ = synStore(OFFSET_OBJ, used = ALL_USED_IDs, activityName = activityName, 

### 2020-05-19 individual-cohorts/HBCC_dsACC_FEATURECOUNTS_GRCh38_CQN.Rmd
add phase II covars for this dataset

@@
                                       'rnaSeq_dissection:Brain_Region',
+                                      'rnaSeq_isolation:RNA_Isolation_Batch',
                                       'rnaSeq_report:Ribozero_Batch', 'rnaSeq_report:Library_Batch', 
@@
                 FlowcellBatch = `rnaSeq_report:Flowcell_Batch`,
-                RibozeroBatch = `rnaSeq_report:Ribozero_Batch`,
+                IsolationBatch = `rnaSeq_isolation:RNA_Isolation_Batch`, #isolation batch to replace ribozero batch
                 MappedReads = `rnaSeq_report:Mapped_Reads`,
@@
                          data = COVARIATES,
-                         variables = array(c("(1|LibraryBatch)", "(1|FlowcellBatch)")))
+                         variables = array(c("(1|IsolationBatch)")))
 kable(phase2$settings)

### 2020-05-19 individual-cohorts/HBCC_dsACC_FEATURECOUNTS_GRCh38_CQN.Rmd
rm LibraryBatch

@@
                 RIN, ReportExclude, GenotypeExclude, EV.1, EV.2, EV.3, EV.4, EV.5,
-                LibraryBatch, FlowcellBatch, RibozeroBatch, MappedReads, TotalReads, GenesDetected, AlignmentRate,
+                LibraryBatch, FlowcellBatch, IsolationBatch, MappedReads, TotalReads, GenesDetected, AlignmentRate,
                 IntragenicRate, IntergenicRate, IntronicRate, ExpProfEfficiency, rRNARate, TranscriptsDetected) %>%
@@
 # Eucledian tree based analysis
-COVARIATES.tmp = data.matrix(COVARIATES[,c("Reported_Gender", "LibraryBatch", "Dx")])
+COVARIATES.tmp = data.matrix(COVARIATES[,c("Reported_Gender", "Dx")])
 COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0

### 2020-05-20 individual-cohorts/HBCC_sgACC_DE.Rmd
test DE HBCC sgACC

--- 2020-05-19 individual-cohorts/HBCC_DLPFC_DE.Rmd
@@
 ---
-title: "Differential Expression - HBCC - DLPFC - Dx"
+title: "Differential Expression - HBCC - sgACC - Dx"
 author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/HBCC_DLPFC_DE.Rmd",
-                                 parentId ="syn22050019",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/HBCC_sgACC_DE.Rmd",
+                                 parentId ="syn22084216",
                                  folderName = "Differential Expression - Dx")
@@
 ```{r synapse.parameters, include=FALSE, cache=TRUE}
-parentId = 'syn22050019';
+parentId = 'syn22084216';
 activityName = 'Differential Expression';
@@
 
-thisFileName <- 'HBCC_DLPFC_DE.Rmd'
+thisFileName <- 'HBCC_sgACC_DE.Rmd'
 
@@
 # Download CQN counts
-CQN_ID = "syn22082460"
+CQN_ID = "syn22084946"
 ALL_USED_IDs = c(CQN_ID)
@@
 # Download filtered counts
-COUNTS_ID = "syn22082459"
+COUNTS_ID = "syn22084938"
 ALL_USED_IDs = c(ALL_USED_IDs, COUNTS_ID)
@@
 # Get covariates
-COVAR_ID = "syn22082458"
+COVAR_ID = "syn22084935"
 ALL_USED_IDs = c(ALL_USED_IDs, COVAR_ID)
@@
 # Set variable class
-FactorCovariates <- c("Individual_ID", "Reported_Gender", "LibraryBatch", "Dx", "RibozeroBatch", "FlowcellBatch")
+FactorCovariates <- c("Individual_ID", "Reported_Gender", "LibraryBatch", "Dx", "IsolationBatch")
 COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
@@
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Dx", "ageOfDeath", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate")
-random_effect <- c("Reported_Gender")
+adjust.covars <- c("Dx", "RIN2", "IntronicRate")
+random_effect <- c("Reported_Gender", "LibraryBatch")
 ```
@@
 plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>% 
-  dplyr::mutate(Tissue = "DLPFC")
+  dplyr::mutate(Tissue = "sgACC")
 
@@
 # Get gene lengths
-counts = 'syn21886235'
+counts = 'syn21867762'
 ALL_USED_IDs = counts
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_DLPFC_Dx_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
-nEXP_OBJ = File('CMC_DLPFC_Dx_netResidualExpression.tsv', 
+  write.table(file = 'CMC_sgACC_Dx_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
+nEXP_OBJ = File('CMC_sgACC_Dx_netResidualExpression.tsv', 
                 name = 'Dx Normalised, covariates removed residual expression (for network analysis)', 
@@
 # Store differential expression results
-write.table(DE, file = 'CMC_DLPFC_Dx_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
-DEXP_OBJ = File('CMC_DLPFC_Dx_DiffExpression.tsv', 
+write.table(DE, file = 'CMC_sgACC_Dx_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
+DEXP_OBJ = File('CMC_sgACC_Dx_DiffExpression.tsv', 
                 name = 'Differential Expression Results (Dx)', 
@@
 # Store adjusted fit linear mixed model object as RDS - Dx
-saveRDS(ADJUSTED.FIT, file = 'CMC_DLPFC_Dx_dream-output.rds')
-ADJ_OBJ = File('CMC_DLPFC_Dx_dream-output.rds',
+saveRDS(ADJUSTED.FIT, file = 'CMC_sgACC_Dx_dream-output.rds')
+ADJ_OBJ = File('CMC_sgACC_Dx_dream-output.rds',
                name = 'Dream Output object - Dx', 

### 2020-05-20 individual-cohorts/HBCC_sgACC_DE.Rmd
fix scaling

@@
                  glue::glue_collapse(adjust.covars, sep = ", ", last = " and "),
-                 ' as fixed effects and',
+                 ' as fixed effects and ',
                  glue::glue_collapse(random_effect, sep = ", ", last = " and "),' chosen as random effects'))
 
-col_type <- dplyr::select(COVARIATES, c(adjust.covars, random_effect)) %>%
+scaled_col_type <- dplyr::select(COVARIATES, c(adjust.covars)) %>%
   dplyr::summarise_all(class) %>% 
@@
 # Scale continuous variables
-adjust.covars = sapply(1:length(col_type$class), function(i){
-  switch(col_type$class[i], 
-    "factor" = paste0(col_type$variable[i]),
-    "numeric" = paste0('scale(', col_type$variable[i], ')')
+scaled_adjust_covars = sapply(1:length(scaled_col_type$class), function(i){
+  switch(scaled_col_type$class[i], 
+    "factor" = paste0(scaled_col_type$variable[i]),
+    "numeric" = paste0('scale(', scaled_col_type$variable[i], ')')
   )
@@
 # Post adjusted formula
-formula <- glue::glue("~ ", glue::glue_collapse(adjust.covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))
+formula <- glue::glue("~ ", glue::glue_collapse(scaled_adjust_covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))
 

### 2020-05-20 individual-cohorts/HBCC_sgACC_DE.Rmd
fix dream input? did not fail in other cohorts

@@
 VOOM.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION
-ADJUSTED.FIT = variancePartition::dream(exprObj = VOOM.GENE_EXPRESSION,
+ADJUSTED.FIT = variancePartition::dream(exprObj = VOOM.GENE_EXPRESSION$E,
                                         formula = formula,

### 2020-05-20 individual-cohorts/HBCC_sgACC_DE.Rmd
fix dream input? did not fail in other cohorts

@@
     "factor" = paste0('(1|', col_type$variable[i], ')'),
-    "numeric" = paste0('scale(', col_type$variable[i], ')')
+    "numeric" = paste0(col_type$variable[i])
   )
@@
 
-violin_plot <- variancePartition::fitExtractVarPartModel(VOOM.GENE_EXPRESSION,
+violin_plot <- variancePartition::fitExtractVarPartModel(VOOM.GENE_EXPRESSION$E,
                                                          formula = formula_for_variance_fractions,

### 2020-05-20 individual-cohorts/HBCC_STG_FEATURECOUNTS_GRCh38_CQN.Rmd
test STG model sel

--- 2020-05-19 individual-cohorts/HBCC_dsACC_FEATURECOUNTS_GRCh38_CQN.Rmd
@@
 ---
-title: "HBCC - dsACC - FEATURECOUNTS - GRCh38 - CQN"
+title: "HBCC - STG - FEATURECOUNTS - GRCh38 - CQN"
 author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/HBCC_dsACC_FEATURECOUNTS_GRCh38_CQN.Rmd",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/HBCC_STG_FEATURECOUNTS_GRCh38_CQN.Rmd",
                                           parentId ="syn22012942",
-                                          folderName = 'HBCC - dsACC - DREAM - FEATURECOUNTS - GRCh38 - CQN')
+                                          folderName = 'HBCC - STG - DREAM - FEATURECOUNTS - GRCh38 - CQN')
 ```
@@
 activityName = 'Covariate adjustments';
-activityDescription = 'Covariate analysis of GRCh38 feature counts with CQN normalisation (dsACC)';
+activityDescription = 'Covariate analysis of GRCh38 feature counts with CQN normalisation (STG)';
 
-thisFileName <- 'HBCC_dsACC_FEATURECOUNTS_GRCh38_CQN.Rmd'
+thisFileName <- 'HBCC_STG_FEATURECOUNTS_GRCh38_CQN.Rmd'
 
@@
 # Get RNASeq QCmetadata
-METADATA_QC_dsACC_ID = 'syn16816488' 
-ALL_USED_IDs = METADATA_QC_dsACC_ID
-metadata = downloadFile_version(METADATA_QC_dsACC_ID, version = 20) %>%
+METADATA_QC_STG_ID = 'syn16816488' 
+ALL_USED_IDs = METADATA_QC_STG_ID
+metadata = downloadFile_version(METADATA_QC_STG_ID, version = 20) %>%
   dplyr::select(`Individual_ID`, Institution, Cohort, `Reported_Gender`, Sex, Ethnicity, ageOfDeath, `PMI_(in_hours)`, Dx, 
@@
                                       'rnaSeq_report:Transcripts_Detected')) %>% 
-  dplyr::filter(`rnaSeq_dissection:Brain_Region` %in% c("dsACC"))
+  dplyr::filter(`rnaSeq_dissection:Brain_Region` %in% c("STG"))
 
-## Download reprocessed counts (dsACC)
+## Download reprocessed counts (STG)
 counts = 'syn21867866'
@@
   dplyr::mutate(Institution = "NIMH-HBCC") %>% 
-  dplyr::mutate(Tissue = "dsACC")
+  dplyr::mutate(Tissue = "STG")
 
@@
 # Code
-CODE <- Folder(name = "HBCC - dsACC - DREAM - FEATURECOUNTS - GRCh38 - CQN", parentId = parentId)
+CODE <- Folder(name = "HBCC - STG - DREAM - FEATURECOUNTS - GRCh38 - CQN", parentId = parentId)
 CODE <- synStore(CODE)
@@
 COVARIATES = rownameToFirstColumn(COVARIATES, 'SampleID')
-write.table(COVARIATES, file = 'CMC_HBCC_dsACC_Covariates.tsv', sep = '\t', row.names=F, quote=F)
-COV_OBJ = synapser::File('CMC_HBCC_dsACC_Covariates.tsv', name = 'Covariates', parentId = CODE$properties$id)
+write.table(COVARIATES, file = 'CMC_HBCC_STG_Covariates.tsv', sep = '\t', row.names=F, quote=F)
+COV_OBJ = synapser::File('CMC_HBCC_STG_Covariates.tsv', name = 'Covariates', parentId = CODE$properties$id)
 COV_OBJ = synStore(COV_OBJ, used = ALL_USED_IDs, activityName = activityName, 
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_HBCC_dsACC_Counts.tsv', sep = '\t', row.names=F, quote=F)
-COUNT_OBJ = File('CMC_HBCC_dsACC_Counts.tsv', name = 'Counts (filtered raw)', parentId = CODE$properties$id)
+  write.table(file = 'CMC_HBCC_STG_Counts.tsv', sep = '\t', row.names=F, quote=F)
+COUNT_OBJ = File('CMC_HBCC_STG_Counts.tsv', name = 'Counts (filtered raw)', parentId = CODE$properties$id)
 COUNT_OBJ = synStore(COUNT_OBJ, used = ALL_USED_IDs, activityName = activityName, 
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_HBCC_dsACC_logCPM.tsv', sep = '\t', row.names=F, quote=F)
-LCOUNT_OBJ = File('CMC_HBCC_dsACC_logCPM.tsv', name = 'Counts (filtered logCPM)', parentId = CODE$properties$id)
+  write.table(file = 'CMC_HBCC_STG_logCPM.tsv', sep = '\t', row.names=F, quote=F)
+LCOUNT_OBJ = File('CMC_HBCC_STG_logCPM.tsv', name = 'Counts (filtered logCPM)', parentId = CODE$properties$id)
 LCOUNT_OBJ = synStore(LCOUNT_OBJ, used = ALL_USED_IDs, activityName = activityName, 
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_HBCC_dsACC_offset.tsv', sep = '\t', row.names=F, quote=F)
-OFFSET_OBJ = File('CMC_HBCC_dsACC_offset.tsv', name = 'Gene length and GC content offset', parentId = CODE$properties$id)
+  write.table(file = 'CMC_HBCC_STG_offset.tsv', sep = '\t', row.names=F, quote=F)
+OFFSET_OBJ = File('CMC_HBCC_STG_offset.tsv', name = 'Gene length and GC content offset', parentId = CODE$properties$id)
 OFFSET_OBJ = synStore(OFFSET_OBJ, used = ALL_USED_IDs, activityName = activityName, 

### 2020-05-20 individual-cohorts/HBCC_sgACC_FEATURECOUNTS_GRCh38_CQN.Rmd
most updated config

@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "HBCC_sgACC_FEATURECOUNTS_GRCh38_CQN.Rmd",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/HBCC_sgACC_FEATURECOUNTS_GRCh38_CQN.Rmd",
                                           parentId ="syn22012942",

### 2020-05-20 individual-cohorts/HBCC_dsACC_DE.Rmd
test DE HBCC dsACC

--- 2020-05-20 individual-cohorts/HBCC_sgACC_DE.Rmd
@@
 ---
-title: "Differential Expression - HBCC - sgACC - Dx"
+title: "Differential Expression - HBCC - dsACC - Dx"
 author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/HBCC_sgACC_DE.Rmd",
-                                 parentId ="syn22084216",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/HBCC_dsACC_DE.Rmd",
+                                 parentId ="syn22084347",
                                  folderName = "Differential Expression - Dx")
@@
 ```{r synapse.parameters, include=FALSE, cache=TRUE}
-parentId = 'syn22084216';
+parentId = 'syn22084347';
 activityName = 'Differential Expression';
@@
 
-thisFileName <- 'HBCC_sgACC_DE.Rmd'
+thisFileName <- 'HBCC_dsACC_DE.Rmd'
 
@@
 # Download CQN counts
-CQN_ID = "syn22084946"
+CQN_ID = "syn22086076"
 ALL_USED_IDs = c(CQN_ID)
@@
 # Download filtered counts
-COUNTS_ID = "syn22084938"
+COUNTS_ID = "syn22086075"
 ALL_USED_IDs = c(ALL_USED_IDs, COUNTS_ID)
@@
 # Get covariates
-COVAR_ID = "syn22084935"
+COVAR_ID = "syn22086074"
 ALL_USED_IDs = c(ALL_USED_IDs, COVAR_ID)
@@
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Dx", "RIN2", "IntronicRate")
-random_effect <- c("Reported_Gender", "LibraryBatch")
+adjust.covars <- c("Dx", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate", "ageOfDeath")
+random_effect <- c("Reported_Gender", "IsolationBatch")
 ```
@@
 # Model categorical variables as random to report variance fractions
+col_type <- dplyr::select(COVARIATES, c(adjust.covars, random_effect)) %>%
+  dplyr::summarise_all(class) %>% 
+  tidyr::gather(variable, class)
+
 formula_for_variance_fractions = sapply(1:length(col_type$class), function(i){
@@
 plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>% 
-  dplyr::mutate(Tissue = "sgACC")
+  dplyr::mutate(Tissue = "dsACC")
 
@@
 # Fit contrast
-FIT.CONTR = variancePartition::dream(exprObj = VOOM.GENE_EXPRESSION,
+FIT.CONTR = variancePartition::dream(exprObj = VOOM.GENE_EXPRESSION$E,
                                      formula = formula,
@@
 # Get gene lengths
-counts = 'syn21867762'
+counts = 'syn21867866'
 ALL_USED_IDs = counts
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_sgACC_Dx_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
-nEXP_OBJ = File('CMC_sgACC_Dx_netResidualExpression.tsv', 
+  write.table(file = 'CMC_dsACC_Dx_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
+nEXP_OBJ = File('CMC_dsACC_Dx_netResidualExpression.tsv', 
                 name = 'Dx Normalised, covariates removed residual expression (for network analysis)', 
@@
 # Store differential expression results
-write.table(DE, file = 'CMC_sgACC_Dx_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
-DEXP_OBJ = File('CMC_sgACC_Dx_DiffExpression.tsv', 
+write.table(DE, file = 'CMC_dsACC_Dx_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
+DEXP_OBJ = File('CMC_dsACC_Dx_DiffExpression.tsv', 
                 name = 'Differential Expression Results (Dx)', 
@@
 # Store adjusted fit linear mixed model object as RDS - Dx
-saveRDS(ADJUSTED.FIT, file = 'CMC_sgACC_Dx_dream-output.rds')
-ADJ_OBJ = File('CMC_sgACC_Dx_dream-output.rds',
+saveRDS(ADJUSTED.FIT, file = 'CMC_dsACC_Dx_dream-output.rds')
+ADJ_OBJ = File('CMC_dsACC_Dx_dream-output.rds',
                name = 'Dream Output object - Dx', 

### 2020-05-20 individual-cohorts/HBCC_dsACC_DE.Rmd
force header

@@
 downloadFile <- function(id){
-  fread(synGet(id)$path, data.table = F)
+  fread(synGet(id)$path, data.table = F, header = T)
 }

### 2020-05-20 individual-cohorts/HBCC_dsACC_DE.Rmd
no library batch data on NIMH-HBCC-4

@@
 # Set variable class
-FactorCovariates <- c("Individual_ID", "Reported_Gender", "LibraryBatch", "Dx", "IsolationBatch")
+FactorCovariates <- c("Individual_ID", "Reported_Gender", "Dx", "IsolationBatch")
 COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)

### 2020-05-20 individual-cohorts/MSSM-Penn-Pitt_ACC_DE.Rmd
rm redundant code

@@
 ```{r set-formula}
+adjust.covars <- adjust.covars[!(adjust.covars %in% "RIN2")]
 formula <- glue::glue("~ 0 + ", glue::glue_collapse(adjust.covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))    

### 2020-05-21 individual-cohorts/graphics_t_statistics.Rmd
plot t-stats and cor heatmap

---
title: "Graphics"
author: "KM"
date: "5/18/2020"
output: html_document
---

```{r knit2synapse, eval=FALSE}
library(synapser)
library(knit2synapse)
synLogin()
knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/graphics_t_statistics.Rmd",
                                          parentId ="syn22012942",
                                          folderName = 't-statistics across cohorts')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(ComplexHeatmap)
library(dendsort)

synLogin()

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

## t-statistics from differential expression results

- separate by comparison

```{r data}
mpp_acc <- readr::read_tsv(synGet("syn22079117")$path) %>% 
  dplyr::select(Comparison, gene_id, t) %>% 
  dplyr::rename(`MSSM-Penn-Pitt ACC t-statistic` = t)
mpp_dlpfc <- readr::read_tsv(synGet("syn22079472")$path) %>% 
  dplyr::select(Comparison, gene_id, t) %>% 
  dplyr::rename(`MSSM-Penn-Pitt DLPFC t-statistic` = t)
hbcc_dsACC <- readr::read_tsv(synGet("syn22086829")$path) %>% 
  dplyr::select(Comparison, gene_id, t) %>% 
  dplyr::rename(`HBCC dsACC t-statistic` = t)
hbcc_stg <- readr::read_tsv(synGet("syn22087386")$path) %>% 
  dplyr::select(Comparison, gene_id, t) %>% 
  dplyr::rename(`HBCC STG t-statistic` = t)
hbcc_sgACC <- readr::read_tsv(synGet("syn22085862")$path) %>% 
  dplyr::select(Comparison, gene_id, t) %>% 
  dplyr::rename(`HBCC sgACC t-statistic` = t)
hbcc_dlpfc <- readr::read_tsv(synGet("syn22084205")$path) %>% 
  dplyr::select(Comparison, gene_id, t) %>% 
  dplyr::rename(`HBCC DLPFC t-statistic` = t)

df <- inner_join(mpp_acc, mpp_dlpfc,
                 by = c("Comparison", "gene_id"))
df <- inner_join(df, hbcc_dsACC, by = c("Comparison", "gene_id"))
df <- inner_join(df, hbcc_stg, by = c("Comparison", "gene_id"))
df <- inner_join(df, hbcc_sgACC, by = c("Comparison", "gene_id"))
df <- inner_join(df, hbcc_dlpfc, by = c("Comparison", "gene_id"))

conditions = list(c(2,3), c(2,4), c(2,5), c(2,6), c(2,7), c(3,4), c(3,5),c(3,6), c(3,7), c(4,5), c(4,6), c(4,7), c(5,6), c(5,7), c(6,7))
```

```{r functions}
plot_t <- function(df, conditions){
  p = list()
  for(i in seq_along(conditions)){
  positions = conditions[i]
  positions
  x = df[unlist(positions)[1]]
  y = df[unlist(positions)[2]]
  computed_cor = cor(x, y, method = "spearman")
  x_var <- names(x)
  y_var <- names(y)
  p[[i]] = ggplot(df, aes(x = .data[[x_var]], y = .data[[y_var]])) +
  geom_hex() +
  geom_abline() +
  theme_bw() +
  labs(
    caption = paste0("spearman correlation between the statistics is ", round(computed_cor, digits = 3))
  )
  }
multiplot(plotlist = p, cols = 2)
}
```

# Control vs Other
```{r plota, fig.width = 10, fig.height = 20}
df_one <- df[df$Comparison %in% "Control_vs_Other",] %>% 
  group_by(gene_id) %>% 
  dplyr::filter(row_number()==1) %>% 
  tibble::column_to_rownames(var = "gene_id")

plot_t(df_one, conditions)
```

```{r heatmapa, fig.width = 5, fig.height = 5 }
for_fig <- dplyr::select(df_one, -Comparison)
colnames(for_fig) <- gsub(" t-statistic", "", colnames(for_fig))

cor <- corr.test(for_fig, use = "pairwise.complete.obs", method = "spearman")

colors <- colorRamp2(c(0, 1), c("white", "red"))
dend <-  dendsort(hclust(dist(as.matrix(cor$r))))

Heatmap(as.matrix(cor$r), 
        name = "spearman cor", 
        rect_gp = gpar(col = "white", lwd = 2),
        column_names_rot = 45,
        col = colors,
        cluster_rows = dend,
        row_names_gp = gpar(fontsize = 10),
        column_names_gp = gpar(fontsize = 10)
        )
```

# Control vs SCZ
```{r plotb, fig.width = 10, fig.height = 20}
df_two <- df[df$Comparison %in% "Control_vs_SCZ",] %>% 
  group_by(gene_id) %>% 
  dplyr::filter(row_number()==1) %>% 
  tibble::column_to_rownames(var = "gene_id")

plot_t(df_two, conditions)
```

```{r heatmapb, fig.width = 5, fig.height = 5 }
for_fig <- dplyr::select(df_two, -Comparison)
colnames(for_fig) <- gsub(" t-statistic", "", colnames(for_fig))

cor <- corr.test(for_fig, use = "pairwise.complete.obs", method = "spearman")

colors <- colorRamp2(c(0, 1), c("white", "red"))
dend <-  dendsort(hclust(dist(as.matrix(cor$r))))

Heatmap(as.matrix(cor$r), 
        name = "spearman cor", 
        rect_gp = gpar(col = "white", lwd = 2),
        column_names_rot = 45,
        col = colors,
        cluster_rows = dend,
        row_names_gp = gpar(fontsize = 10),
        column_names_gp = gpar(fontsize = 10)
        )
```

# Other vs SCZ
```{r plotc, fig.width = 10, fig.height = 20}
df_three <- df[df$Comparison %in% "Other_vs_SCZ",] %>% 
  group_by(gene_id) %>% 
  dplyr::filter(row_number()==1) %>% 
  tibble::column_to_rownames(var = "gene_id")

plot_t(df_three, conditions)
```

```{r heatmapc, fig.width = 5, fig.height = 5 }
for_fig <- dplyr::select(df_three, -Comparison)
colnames(for_fig) <- gsub(" t-statistic", "", colnames(for_fig))

cor <- corr.test(for_fig, use = "pairwise.complete.obs", method = "spearman")

colors <- colorRamp2(c(0, 1), c("white", "red"))
dend <-  dendsort(hclust(dist(as.matrix(cor$r))))

Heatmap(as.matrix(cor$r), 
        name = "spearman cor", 
        rect_gp = gpar(col = "white", lwd = 2),
        column_names_rot = 45,
        col = colors,
        cluster_rows = dend,
        row_names_gp = gpar(fontsize = 10),
        column_names_gp = gpar(fontsize = 10)
        )
```

```{r store.files}
CODE <- Folder(name = "t-statistics across cohorts", parentId = "syn22012942")
CODE <- synStore(CODE)
```

### 2020-05-21 individual-cohorts/HBCC_STG_DE.Rmd

--- 2020-05-20 individual-cohorts/HBCC_dsACC_DE.Rmd
@@
 ---
-title: "Differential Expression - HBCC - dsACC - Dx"
+title: "Differential Expression - HBCC - STG - Dx"
 author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/HBCC_dsACC_DE.Rmd",
-                                 parentId ="syn22084347",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/HBCC_STG_DE.Rmd",
+                                 parentId ="syn22086820",
                                  folderName = "Differential Expression - Dx")
@@
 ```{r synapse.parameters, include=FALSE, cache=TRUE}
-parentId = 'syn22084347';
+parentId = 'syn22086820';
 activityName = 'Differential Expression';
@@
 
-thisFileName <- 'HBCC_dsACC_DE.Rmd'
+thisFileName <- 'HBCC_STG_DE.Rmd'
 
@@
 # Download CQN counts
-CQN_ID = "syn22086076"
+CQN_ID = "syn22087258"
 ALL_USED_IDs = c(CQN_ID)
@@
 # Download filtered counts
-COUNTS_ID = "syn22086075"
+COUNTS_ID = "syn22087256"
 ALL_USED_IDs = c(ALL_USED_IDs, COUNTS_ID)
@@
 # Get covariates
-COVAR_ID = "syn22086074"
+COVAR_ID = "syn22087255"
 ALL_USED_IDs = c(ALL_USED_IDs, COVAR_ID)
@@
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Dx", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate", "ageOfDeath")
+adjust.covars <- c("Dx", "RIN2", "IntronicRate", "IntragenicRate", "IntergenicRate", "ageOfDeath")
 random_effect <- c("Reported_Gender", "IsolationBatch")
@@
 plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>% 
-  dplyr::mutate(Tissue = "dsACC")
+  dplyr::mutate(Tissue = "STG")
 
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_dsACC_Dx_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
-nEXP_OBJ = File('CMC_dsACC_Dx_netResidualExpression.tsv', 
+  write.table(file = 'CMC_STG_Dx_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
+nEXP_OBJ = File('CMC_STG_Dx_netResidualExpression.tsv', 
                 name = 'Dx Normalised, covariates removed residual expression (for network analysis)', 
@@
 # Store differential expression results
-write.table(DE, file = 'CMC_dsACC_Dx_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
-DEXP_OBJ = File('CMC_dsACC_Dx_DiffExpression.tsv', 
+write.table(DE, file = 'CMC_STG_Dx_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
+DEXP_OBJ = File('CMC_STG_Dx_DiffExpression.tsv', 
                 name = 'Differential Expression Results (Dx)', 
@@
 # Store adjusted fit linear mixed model object as RDS - Dx
-saveRDS(ADJUSTED.FIT, file = 'CMC_dsACC_Dx_dream-output.rds')
-ADJ_OBJ = File('CMC_dsACC_Dx_dream-output.rds',
+saveRDS(ADJUSTED.FIT, file = 'CMC_STG_Dx_dream-output.rds')
+ADJ_OBJ = File('CMC_STG_Dx_dream-output.rds',
                name = 'Dream Output object - Dx', 

### 2020-05-21 MSSM-Penn-Pitt_DE_DxTissue.Rmd
supress warning and drop colinear vars

@@
 ```{r set-formula}
+# remove RIN2 as RIN and RIN2 are highly correlated
+adjust.covars <- c("Dx.Tissue", "ageOfDeath", "EV.1", "PMI", "EV.4","EV.2","RIN", "EV.3", "EV.5", "IntronicRate")
 formula <- glue::glue("~ 0 + ", glue::glue_collapse(adjust.covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))    
-
 ```
-
+#drop redundancy - RIN2 or RIN
+#use canCorPairs() to test colinearity, don't want var of interest to be correlated
+#leave out random effects
 `r formula`

### 2020-06-02 individual-cohorts/graphics_t_statistics.Rmd
add 3 external cohorts

@@
 library(knit2synapse)
-synLogin()
 knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/graphics_t_statistics.Rmd",
@@
 library(ComplexHeatmap)
-library(dendsort)
+library(readxl)
+library(gtools)
 
@@
   cache = TRUE)
+```
+
+```{r synapse.parameters, include=FALSE, cache=TRUE}
+thisFileName <- 'graphics_t_statistics.Rmd'
+
+# Github link
+thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
+thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/',thisFileName))
 ```
@@
   dplyr::rename(`HBCC DLPFC t-statistic` = t)
+# CMC Analysis 1
+CMC1 <- readr::read_tsv(synGet("syn5607603")$path) %>% 
+  dplyr::rename(gene_id = genes,
+                `CMC 1 t-statistic` = t) %>% 
+  dplyr::select(gene_id, `CMC 1 t-statistic`) %>% 
+  dplyr::mutate(Comparison = "Control_vs_SCZ")
+# HBCC-DLPFC microarray
+CMC_HBCC <- readr::read_tsv(synGet("syn4941553")$path) %>% 
+  dplyr::rename(gene_id = genes,
+                `HBCC microarray t-statistic` = t) %>% 
+  dplyr::select(gene_id, `HBCC microarray t-statistic`) %>%
+  dplyr::mutate(Comparison = "Control_vs_SCZ")
+# Gandal, Science 2018, Table S1
+# Use Tab:DGE
+# wget https://science.sciencemag.org/highwire/filestream/720110/field_highwire_adjunct_files/1/aat8127_Table_S1.xlsx
+gandal <- readxl::read_xlsx('../aat8127_Table_S1.xlsx', sheet = "DGE") %>% 
+  dplyr::rename(gene_id = ensembl_gene_id,
+                `Gandal, Science 2018 t-statistic` = `SCZ.t.value`) %>% 
+  dplyr::select(gene_id, `Gandal, Science 2018 t-statistic`) %>% 
+  dplyr::mutate(Comparison = "Control_vs_SCZ")
 
@@
 df <- inner_join(df, hbcc_dlpfc, by = c("Comparison", "gene_id"))
+#Remove transcript Ids
+df <- df %>% 
+  separate(gene_id, c("gene_id", "transcript")) %>% 
+  dplyr::select(-one_of("transcript"))
+df <- inner_join(df, CMC1, by = c("Comparison", "gene_id")) 
+df <- inner_join(df, CMC_HBCC, by = c("Comparison", "gene_id"))
+df <- inner_join(df, gandal, by = c("Comparison", "gene_id"))
 
-conditions = list(c(2,3), c(2,4), c(2,5), c(2,6), c(2,7), c(3,4), c(3,5),c(3,6), c(3,7), c(4,5), c(4,6), c(4,7), c(5,6), c(5,7), c(6,7))
+#conditions = list(c(2,3), c(2,4), c(2,5), c(2,6), c(2,7), c(3,4), c(3,5),c(3,6), c(3,7), c(4,5), c(4,6), c(4,7), c(5,6), c(5,7), c(6,7))
 ```
@@
   p = list()
-  for(i in seq_along(conditions)){
-  positions = conditions[i]
+  for(i in seq_along(conditions[,1])){
+  positions = conditions[i,]
   positions
-  x = df[unlist(positions)[1]]
-  y = df[unlist(positions)[2]]
+  x = df[positions[1]]
+  y = df[positions[2]]
   computed_cor = cor(x, y, method = "spearman")
@@
     caption = paste0("spearman correlation between the statistics is ", round(computed_cor, digits = 3))
-  )
+  ) +
+    theme(aspect.ratio=1)
   }
@@
 # Control vs Other
-```{r plota, fig.width = 10, fig.height = 20}
+```{r plota, fig.width = 20, fig.height = 40}
 df_one <- df[df$Comparison %in% "Control_vs_Other",] %>% 
@@
 
+conditions <- gtools::combinations(n = ncol(df_one), 2, seq_len(ncol(df_one)))
 plot_t(df_one, conditions)
@@
 colors <- colorRamp2(c(0, 1), c("white", "red"))
-dend <-  dendsort(hclust(dist(as.matrix(cor$r))))
 
@@
         col = colors,
-        cluster_rows = dend,
         row_names_gp = gpar(fontsize = 10),
@@
 # Control vs SCZ
-```{r plotb, fig.width = 10, fig.height = 20}
+```{r plotb, fig.width = 20, fig.height = 40}
 df_two <- df[df$Comparison %in% "Control_vs_SCZ",] %>% 
@@
 # Other vs SCZ
-```{r plotc, fig.width = 10, fig.height = 20}
+```{r plotc, fig.width = 20, fig.height = 40}
 df_three <- df[df$Comparison %in% "Other_vs_SCZ",] %>% 
@@
 
+Code: `r thisFile`
+
 ```{r store.files}

### 2020-06-02 individual-cohorts/graphics_t_statistics.Rmd
remove missing values for certain comparisons

@@
   dplyr::select(-one_of("transcript"))
-df <- inner_join(df, CMC1, by = c("Comparison", "gene_id")) 
-df <- inner_join(df, CMC_HBCC, by = c("Comparison", "gene_id"))
-df <- inner_join(df, gandal, by = c("Comparison", "gene_id"))
+df <- left_join(df, CMC1, by = c("Comparison", "gene_id")) 
+df <- left_join(df, CMC_HBCC, by = c("Comparison", "gene_id"))
+df <- left_join(df, gandal, by = c("Comparison", "gene_id"))
 
@@
   dplyr::filter(row_number()==1) %>% 
-  tibble::column_to_rownames(var = "gene_id")
+  tibble::column_to_rownames(var = "gene_id") %>% 
+  dplyr::select_if( ~ !all(is.na(.)))
 
@@
   dplyr::filter(row_number()==1) %>% 
-  tibble::column_to_rownames(var = "gene_id")
+  tibble::column_to_rownames(var = "gene_id") %>% 
+  dplyr::select_if( ~ !all(is.na(.)))
 
+conditions <- gtools::combinations(n = ncol(df_two), 2, seq_len(ncol(df_two)))
 plot_t(df_two, conditions)
@@
   dplyr::filter(row_number()==1) %>% 
-  tibble::column_to_rownames(var = "gene_id")
+  tibble::column_to_rownames(var = "gene_id") %>% 
+  dplyr::select_if( ~ !all(is.na(.)))
 
+conditions <- gtools::combinations(n = ncol(df_three), 2, seq_len(ncol(df_three)))
 plot_t(df_three, conditions)

### 2020-06-02 individual-cohorts/graphics_t_statistics.Rmd
fix df

@@
   positions = conditions[i,]
-  positions
   x = df[positions[1]]
@@
   tibble::column_to_rownames(var = "gene_id") %>% 
-  dplyr::select_if( ~ !all(is.na(.)))
+  dplyr::select_if( ~ !all(is.na(.))) %>% 
+  dplyr::select(-Comparison)
 

### 2020-06-02 individual-cohorts/graphics_t_statistics.Rmd
fix heatmap

@@
 author: "KM"
-date: "5/18/2020"
+date: "`r date()`"
 output: html_document
@@
 ```{r heatmapa, fig.width = 5, fig.height = 5 }
-for_fig <- dplyr::select(df_one, -Comparison)
-colnames(for_fig) <- gsub(" t-statistic", "", colnames(for_fig))
+colnames(df_one) <- gsub(" t-statistic", "", colnames(df_one))
 
-cor <- corr.test(for_fig, use = "pairwise.complete.obs", method = "spearman")
+cor <- corr.test(df_one, use = "pairwise.complete.obs", method = "spearman")
 
@@
 ```{r heatmapb, fig.width = 5, fig.height = 5 }
-for_fig <- dplyr::select(df_two, -Comparison)
-colnames(for_fig) <- gsub(" t-statistic", "", colnames(for_fig))
+colnames(df_two) <- gsub(" t-statistic", "", colnames(df_two))
 
-cor <- corr.test(for_fig, use = "pairwise.complete.obs", method = "spearman")
+cor <- corr.test(df_two, use = "pairwise.complete.obs", method = "spearman")
 
@@
 ```{r heatmapc, fig.width = 5, fig.height = 5 }
-for_fig <- dplyr::select(df_three, -Comparison)
-colnames(for_fig) <- gsub(" t-statistic", "", colnames(for_fig))
+colnames(df_three) <- gsub(" t-statistic", "", colnames(df_three))
 
-cor <- corr.test(for_fig, use = "pairwise.complete.obs", method = "spearman")
+cor <- corr.test(df_three, use = "pairwise.complete.obs", method = "spearman")
 

### 2020-06-03 individual-cohorts/graphics_t_statistics.Rmd
add other comparisons to Gandal cohorts

@@
                 `CMC 1 t-statistic` = t) %>% 
-  dplyr::select(gene_id, `CMC 1 t-statistic`) %>% 
-  dplyr::mutate(Comparison = "Control_vs_SCZ")
+  dplyr::select(gene_id, `CMC 1 t-statistic`)
 # HBCC-DLPFC microarray
@@
                 `HBCC microarray t-statistic` = t) %>% 
-  dplyr::select(gene_id, `HBCC microarray t-statistic`) %>%
-  dplyr::mutate(Comparison = "Control_vs_SCZ")
+  dplyr::select(gene_id, `HBCC microarray t-statistic`)
 # Gandal, Science 2018, Table S1
@@
   dplyr::rename(gene_id = ensembl_gene_id,
-                `Gandal, Science 2018 t-statistic` = `SCZ.t.value`) %>% 
-  dplyr::select(gene_id, `Gandal, Science 2018 t-statistic`) %>% 
-  dplyr::mutate(Comparison = "Control_vs_SCZ")
+                `Gandal, SCZ t-statistic` = `SCZ.t.value`,
+                `Gandal, BD t-statistic` = `BD.t.value`,
+                `Gandal, ASD t-statistic` = `ASD.t.value`) %>%
+  dplyr::select(gene_id, `Gandal, SCZ t-statistic`,
+                `Gandal, BD t-statistic`, `Gandal, ASD t-statistic`)
 
@@
 df <- inner_join(df, hbcc_dlpfc, by = c("Comparison", "gene_id"))
-#Remove transcript Ids
-df <- df %>% 
-  separate(gene_id, c("gene_id", "transcript")) %>% 
-  dplyr::select(-one_of("transcript"))
-df <- left_join(df, CMC1, by = c("Comparison", "gene_id")) 
-df <- left_join(df, CMC_HBCC, by = c("Comparison", "gene_id"))
-df <- left_join(df, gandal, by = c("Comparison", "gene_id"))
-
-#conditions = list(c(2,3), c(2,4), c(2,5), c(2,6), c(2,7), c(3,4), c(3,5),c(3,6), c(3,7), c(4,5), c(4,6), c(4,7), c(5,6), c(5,7), c(6,7))
 ```
@@
 ```{r plota, fig.width = 20, fig.height = 40}
-df_one <- df[df$Comparison %in% "Control_vs_Other",] %>% 
+df_one <- df[df$Comparison %in% "Control_vs_SCZ",] %>% 
+  dplyr::select_if( ~ !all(is.na(.))) %>% 
+  dplyr::select(-Comparison)
+
+#Remove transcript Ids
+df_one <- df_one %>% 
+  separate(gene_id, c("gene_id", "transcript")) %>% 
+  dplyr::select(-one_of("transcript"))
+df_one <- inner_join(df_one, dplyr::select(gandal, gene_id,  `Gandal, BD t-statistic`,`Gandal, ASD t-statistic`), by = c("gene_id")) %>% 
   group_by(gene_id) %>% 
   dplyr::filter(row_number()==1) %>% 
-  tibble::column_to_rownames(var = "gene_id") %>% 
-  dplyr::select_if( ~ !all(is.na(.))) %>% 
-  dplyr::select(-Comparison)
+  tibble::column_to_rownames(var = "gene_id")
 
@@
 # Control vs SCZ
-```{r plotb, fig.width = 20, fig.height = 40}
+```{r plotb, fig.width = 10, fig.height = 100}
 df_two <- df[df$Comparison %in% "Control_vs_SCZ",] %>% 
+  dplyr::select_if( ~ !all(is.na(.))) %>% 
+  dplyr::select(-Comparison)
+
+#Remove transcript Ids
+df_two <- df_two %>% 
+  separate(gene_id, c("gene_id", "transcript")) %>% 
+  dplyr::select(-one_of("transcript"))
+df_two <- inner_join(df_two, CMC1, by = c("gene_id")) 
+df_two <- inner_join(df_two, CMC_HBCC, by = c("gene_id"))
+df_two <- inner_join(df_two, dplyr::select(gandal, gene_id, `Gandal, SCZ t-statistic`), by = c("gene_id")) %>% 
   group_by(gene_id) %>% 
   dplyr::filter(row_number()==1) %>% 
-  tibble::column_to_rownames(var = "gene_id") %>% 
-  dplyr::select_if( ~ !all(is.na(.))) %>% 
-  dplyr::select(-Comparison)
+  tibble::column_to_rownames(var = "gene_id")
 
@@
 
-```{r heatmapb, fig.width = 5, fig.height = 5 }
+```{r heatmapb, fig.width = 8, fig.height = 8 }
 colnames(df_two) <- gsub(" t-statistic", "", colnames(df_two))

### 2020-06-03 individual-cohorts/graphics_t_statistics.Rmd
add cor values to heatmap

@@
 library(knit2synapse)
+synLogin()
 knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/graphics_t_statistics.Rmd",
@@
 # Control vs Other
-```{r plota, fig.width = 20, fig.height = 40}
-df_one <- df[df$Comparison %in% "Control_vs_SCZ",] %>% 
+```{r plota, fig.width = 10, fig.height = 100}
+df_one <- df[df$Comparison %in% "Control_vs_Other",] %>% 
   dplyr::select_if( ~ !all(is.na(.))) %>% 
@@
         row_names_gp = gpar(fontsize = 10),
-        column_names_gp = gpar(fontsize = 10)
+        column_names_gp = gpar(fontsize = 10),
+        cell_fun = function(j, i, x, y, width, height, fill) {
+        grid.text(sprintf("%.1f", as.matrix(cor$r)[i, j]), x, y, gp = gpar(fontsize = 8))
+          }
         )
@@
 # Other vs SCZ
-```{r plotc, fig.width = 20, fig.height = 40}
+```{r plotc, fig.width = 10, fig.height = 100}
 df_three <- df[df$Comparison %in% "Other_vs_SCZ",] %>% 
+  dplyr::select_if( ~ !all(is.na(.))) %>% 
+  dplyr::select(-Comparison)
+
+#Remove transcript Ids
+df_three <- df_three %>% 
+  separate(gene_id, c("gene_id", "transcript")) %>% 
+  dplyr::select(-one_of("transcript"))
+df_three <- inner_join(df_three, dplyr::select(gandal, gene_id,  `Gandal, BD t-statistic`,`Gandal, ASD t-statistic`), by = c("gene_id")) %>% 
   group_by(gene_id) %>% 
   dplyr::filter(row_number()==1) %>% 
-  tibble::column_to_rownames(var = "gene_id") %>% 
-  dplyr::select_if( ~ !all(is.na(.))) %>% 
-  dplyr::select(-Comparison)
+  tibble::column_to_rownames(var = "gene_id")
 

### 2020-06-03 individual-cohorts/graphics_t_statistics.Rmd
improve plot and text size

@@
   ) +
-    theme(aspect.ratio=1)
+    theme(aspect.ratio=1, 
+          axis.text=element_text(size=12),
+          plot.caption = element_text(face = "bold", size = "12"))
   }
@@
   dplyr::select(-one_of("transcript"))
-df_one <- inner_join(df_one, dplyr::select(gandal, gene_id,  `Gandal, BD t-statistic`,`Gandal, ASD t-statistic`), by = c("gene_id")) %>% 
+df_one <- inner_join(df_one, dplyr::select(gandal, gene_id,  `Gandal, SCZ t-statistic`,`Gandal, BD t-statistic`,`Gandal, ASD t-statistic`), by = c("gene_id")) %>% 
   group_by(gene_id) %>% 
@@
 
-```{r heatmapa, fig.width = 5, fig.height = 5 }
+```{r heatmapa, fig.width = 8, fig.height = 8 }
 colnames(df_one) <- gsub(" t-statistic", "", colnames(df_one))
@@
         col = colors,
-        row_names_gp = gpar(fontsize = 10),
-        column_names_gp = gpar(fontsize = 10),
+        row_names_gp = gpar(fontsize = 12),
+        column_names_gp = gpar(fontsize = 12),
         cell_fun = function(j, i, x, y, width, height, fill) {
-        grid.text(sprintf("%.1f", as.matrix(cor$r)[i, j]), x, y, gp = gpar(fontsize = 8))
+        grid.text(sprintf("%.1f", as.matrix(cor$r)[i, j]), x, y, gp = gpar(fontsize = 12))
           }
@@
 df_two <- inner_join(df_two, CMC_HBCC, by = c("gene_id"))
-df_two <- inner_join(df_two, dplyr::select(gandal, gene_id, `Gandal, SCZ t-statistic`), by = c("gene_id")) %>% 
+df_two <- inner_join(df_two, dplyr::select(gandal, gene_id, `Gandal, SCZ t-statistic`, 
+                                           `Gandal, ASD t-statistic`, `Gandal, BD t-statistic`), by = c("gene_id")) %>% 
   group_by(gene_id) %>% 
@@
 
-```{r heatmapb, fig.width = 8, fig.height = 8 }
+```{r heatmapb, fig.width = 5, fig.height = 5 }
 colnames(df_two) <- gsub(" t-statistic", "", colnames(df_two))
@@
   dplyr::select(-one_of("transcript"))
-df_three <- inner_join(df_three, dplyr::select(gandal, gene_id,  `Gandal, BD t-statistic`,`Gandal, ASD t-statistic`), by = c("gene_id")) %>% 
+df_three <- inner_join(df_three, dplyr::select(gandal, gene_id,  `Gandal, BD t-statistic`,
+                                               `Gandal, ASD t-statistic`, `Gandal, SCZ t-statistic`), by = c("gene_id")) %>% 
   group_by(gene_id) %>% 

### 2020-06-03 individual-cohorts/graphics_t_statistics.Rmd
improve plot and text size

@@
 
-```{r heatmapc, fig.width = 5, fig.height = 5 }
+```{r heatmapc, fig.width = 8, fig.height = 8 }
 colnames(df_three) <- gsub(" t-statistic", "", colnames(df_three))

### 2020-06-03 individual-cohorts/graphics_t_statistics.Rmd
replace heatmap with corrplot

@@
 
-colors <- colorRamp2(c(0, 1), c("white", "red"))
-
-Heatmap(as.matrix(cor$r), 
-        name = "spearman cor", 
-        rect_gp = gpar(col = "white", lwd = 2),
-        column_names_rot = 45,
-        col = colors,
-        row_names_gp = gpar(fontsize = 12),
-        column_names_gp = gpar(fontsize = 12),
-        cell_fun = function(j, i, x, y, width, height, fill) {
-        grid.text(sprintf("%.1f", as.matrix(cor$r)[i, j]), x, y, gp = gpar(fontsize = 12))
-          }
-        )
+col2 <- colorRampPalette(rev(c("#67001F", "#B2182B", "#D6604D", "#F4A582",
+                           "#FDDBC7", "#FFFFFF", "#D1E5F0", "#92C5DE",
+                           "#4393C3", "#2166AC", "#053061")))
+corrplot(cor$r, col = col2(200))
 ```
@@
 
-colors <- colorRamp2(c(0, 1), c("white", "red"))
-
-Heatmap(as.matrix(cor$r), 
-        name = "spearman cor", 
-        rect_gp = gpar(col = "white", lwd = 2),
-        column_names_rot = 45,
-        col = colors,
-        row_names_gp = gpar(fontsize = 12),
-        column_names_gp = gpar(fontsize = 12),
-        cell_fun = function(j, i, x, y, width, height, fill) {
-        grid.text(sprintf("%.1f", as.matrix(cor$r)[i, j]), x, y, gp = gpar(fontsize = 12))
-          }
-        )
+corrplot(cor$r, col = col2(200))
 ```

### 2020-06-03 individual-cohorts/graphics_t_statistics.Rmd
hierarchical clustering

@@
 
-cor <- corr.test(df_one, use = "pairwise.complete.obs", method = "spearman")
-
+cor <- cor(df_one)
+# clustering based on distance
+hcl = hclust(as.dist(1-cor))
+# reorder based on clustering
+reorder = cor[hcl$order, hcl$order]
+# plot clustering
+plot(hcl)
 col2 <- colorRampPalette(rev(c("#67001F", "#B2182B", "#D6604D", "#F4A582",
@@
                            "#4393C3", "#2166AC", "#053061")))
-corrplot(cor$r, col = col2(200))
+corrplot(reorder, col = col2(200))
 ```
@@
 
-cor <- corr.test(df_two, use = "pairwise.complete.obs", method = "spearman")
-
-corrplot(cor$r, col = col2(200))
+cor <- cor(df_two)
+# clustering based on distance
+hcl = hclust(as.dist(1-cor))
+# reorder based on clustering
+reorder = cor[hcl$order, hcl$order]
+# plot clustering
+plot(hcl)
+corrplot(reorder, col = col2(200))
 ```
@@
 
-cor <- corr.test(df_three, use = "pairwise.complete.obs", method = "spearman")
-
-corrplot(cor$r, col = col2(200))
+cor <- cor(df_three)
+# clustering based on distance
+hcl = hclust(as.dist(1-cor))
+# reorder based on clustering
+reorder = cor[hcl$order, hcl$order]
+# plot clustering
+plot(hcl)
+corrplot(reorder, col = col2(200))
 ```

### 2020-06-03 individual-cohorts/graphics_t_statistics.Rmd
hierarchical clustering

--- 2020-06-03 individual-cohorts/graphics_t_statistics.Rmd
@@
 
-cor <- corr.test(df_one, use = "pairwise.complete.obs", method = "spearman")
-
+cor <- cor(df_one)
+# clustering based on distance
+hcl = hclust(as.dist(1-cor))
+# reorder based on clustering
+reorder = cor[hcl$order, hcl$order]
+# plot clustering
+plot(hcl, hang=-1)
 col2 <- colorRampPalette(rev(c("#67001F", "#B2182B", "#D6604D", "#F4A582",
@@
 
-cor <- corr.test(df_two, use = "pairwise.complete.obs", method = "spearman")
-
-corrplot(cor$r, col = col2(200))
+cor <- cor(df_two)
+# clustering based on distance
+hcl = hclust(as.dist(1-cor))
+# reorder based on clustering
+reorder = cor[hcl$order, hcl$order]
+# plot clustering
+plot(hcl, hang=-1)
+corrplot(reorder, col = col2(200))
 ```
@@
 
-cor <- corr.test(df_three, use = "pairwise.complete.obs", method = "spearman")
-
-corrplot(cor$r, col = col2(200))
+cor <- cor(df_three)
+# clustering based on distance
+hcl = hclust(as.dist(1-cor))
+# reorder based on clustering
+reorder = cor[hcl$order, hcl$order]
+# plot clustering
+plot(hcl, hang=-1)
+corrplot(reorder, col = col2(200))
 ```

### 2020-06-19 individual-cohorts/graphics_t_statistics.Rmd
add AD outgroup and pvclust

--- 2020-06-03 individual-cohorts/graphics_t_statistics.Rmd
@@
 library(gtools)
+library(pvclust)
 
@@
 
+# AMP-AD RNASeq reprocessing GRCh37 - https://www.synapse.org/#!Synapse:syn9702085
+MSSM <- readr::read_tsv(synGet("syn10526259")$path) %>%
+  dplyr::filter(Comparison %in% "AD-CONTROL",
+                Model %in% "Diagnosis") %>%
+  dplyr::select(ensembl_gene_id, Tissue, t) %>%
+  dplyr::group_by(ensembl_gene_id, Tissue) %>% 
+  dplyr::filter(row_number() == 1) %>% 
+  tidyr::pivot_wider(names_from = Tissue,
+                     values_from = t) %>%
+  dplyr::rename(`AMP-AD MSSM FP AD t-statistic` = FP,
+                `AMP-AD MSSM IGF AD t-statistic` = IFG,
+                `AMP-AD MSSM PHG AD t-statistic` = PHG,
+                `AMP-AD MSSM STG AD t-statistic` = STG,
+                gene_id = ensembl_gene_id) %>% 
+  dplyr::select(gene_id, `AMP-AD MSSM STG AD t-statistic`) 
+MSSM <- MSSM[!is.na(MSSM$`AMP-AD MSSM STG AD t-statistic`),]
+
+MAYO <- readr::read_tsv(synGet("syn8468023")$path) %>% 
+  dplyr::filter(Comparison %in% "AD-CONTROL",
+                Model %in% "Diagnosis") %>%
+  dplyr::select(ensembl_gene_id, `Tissue.ref`, t) %>%
+  dplyr::group_by(ensembl_gene_id, `Tissue.ref`) %>% 
+  dplyr::filter(row_number() == 1) %>% 
+  tidyr::pivot_wider(names_from = `Tissue.ref`,
+                     values_from = t) %>% 
+  dplyr::rename(`AMP-AD MAYO CBE AD t-statistic` = CBE,
+                `AMP-AD MAYO TCX AD t-statistic` = TCX,
+                gene_id = ensembl_gene_id) %>% 
+  dplyr::select(-`AMP-AD MAYO TCX AD t-statistic`)
+
+ROSMAP <- readr::read_tsv(synGet("syn8456721")$path) %>%
+    dplyr::filter(Comparison %in% "AD-CONTROL",
+                Model %in% "Diagnosis") %>%
+  dplyr::rename(gene_id = ensembl_gene_id, 
+                `AMP-AD ROSMAP DLPFC AD t-statistic` = t) %>% 
+  dplyr::select(gene_id, `AMP-AD ROSMAP DLPFC AD t-statistic`)
+
+out_group <- inner_join(MSSM, MAYO) %>% 
+  inner_join(., ROSMAP) %>% 
+  inner_join(., gandal)
+
 df <- inner_join(mpp_acc, mpp_dlpfc,
@@
   geom_abline() +
+  geom_smooth(method = "lm") +
   theme_bw() +
+  xlim(8,-8) + 
+  ylim(8,-8) +
     labs(
@@
   dplyr::select(-one_of("transcript"))
-df_one <- inner_join(df_one, dplyr::select(gandal, gene_id,  `Gandal, SCZ t-statistic`,`Gandal, BD t-statistic`,`Gandal, ASD t-statistic`), by = c("gene_id")) %>% 
+df_one <- inner_join(df_one, out_group, by = c("gene_id")) %>% 
   group_by(gene_id) %>% 
@@
 
-cor <- corr.test(df_one, use = "pairwise.complete.obs", method = "spearman")
-
+cor <- cor(df_one)
+# clustering based on distance
+hcl <- pvclust(cor)
+# plot clustering
+plot(hcl, hang=-1)
 col2 <- colorRampPalette(rev(c("#67001F", "#B2182B", "#D6604D", "#F4A582",
@@
 df_two <- inner_join(df_two, CMC_HBCC, by = c("gene_id"))
-df_two <- inner_join(df_two, dplyr::select(gandal, gene_id, `Gandal, SCZ t-statistic`, 
-                                           `Gandal, ASD t-statistic`, `Gandal, BD t-statistic`), by = c("gene_id")) %>% 
+df_two <- inner_join(df_two, out_group, by = c("gene_id")) %>% 
   group_by(gene_id) %>% 
@@
 
-cor <- corr.test(df_two, use = "pairwise.complete.obs", method = "spearman")
-
-corrplot(cor$r, col = col2(200))
+cor <- cor(df_two)
+# clustering based on distance
+hcl <- pvclust(cor)
+# plot clustering
+plot(hcl, hang=-1)
+corrplot(reorder, col = col2(200))
 ```
@@
   dplyr::select(-one_of("transcript"))
-df_three <- inner_join(df_three, dplyr::select(gandal, gene_id,  `Gandal, BD t-statistic`,
-                                               `Gandal, ASD t-statistic`, `Gandal, SCZ t-statistic`), by = c("gene_id")) %>% 
+df_three <- inner_join(df_three, out_group, by = c("gene_id")) %>% 
   group_by(gene_id) %>% 
@@
 
-cor <- corr.test(df_three, use = "pairwise.complete.obs", method = "spearman")
-
-corrplot(cor$r, col = col2(200))
+cor <- cor(df_three)
+# clustering based on distance
+hcl <- pvclust(cor)
+# plot clustering
+plot(hcl, hang=-1)
+corrplot(reorder, col = col2(200))
 ```

### 2020-06-19 individual-cohorts/graphics_t_statistics.Rmd
add AD outgroup and pvclust

@@
 # clustering based on distance
-hcl <- pvclust(cor)
+hcl = hclust(as.dist(1-cor))
+# reorder based on clustering
+reorder = cor[hcl$order, hcl$order]
+results <- pvclust(cor)
 # plot clustering
-plot(hcl, hang=-1)
+plot(results, hang=-1)
 col2 <- colorRampPalette(rev(c("#67001F", "#B2182B", "#D6604D", "#F4A582",
@@
 
-cor <- cor(df_two)
+cor <- cor(df_one)
 # clustering based on distance
-hcl <- pvclust(cor)
+hcl = hclust(as.dist(1-cor))
+# reorder based on clustering
+reorder = cor[hcl$order, hcl$order]
+results <- pvclust(cor)
 # plot clustering
-plot(hcl, hang=-1)
+plot(results, hang=-1)
 corrplot(reorder, col = col2(200))
@@
 
-cor <- cor(df_three)
+cor <- cor(df_one)
 # clustering based on distance
-hcl <- pvclust(cor)
+hcl = hclust(as.dist(1-cor))
+# reorder based on clustering
+reorder = cor[hcl$order, hcl$order]
+results <- pvclust(cor)
 # plot clustering
-plot(hcl, hang=-1)
+plot(results, hang=-1)
 corrplot(reorder, col = col2(200))

### 2020-06-19 individual-cohorts/graphics_t_statistics.Rmd
add AD outgroup and pvclust

@@
 # Control vs SCZ
-```{r plotb, fig.width = 10, fig.height = 100}
+```{r plotb, fig.width = 10, fig.height = 150}
 df_two <- df[df$Comparison %in% "Control_vs_SCZ",] %>% 
@@
 
-cor <- cor(df_one)
+cor <- cor(df_two)
 # clustering based on distance
@@
 
-cor <- cor(df_one)
+cor <- cor(df_three)
 # clustering based on distance

### 2020-07-16 individual-cohorts/graphics_t_statistics.Rmd
add meta-analysis

@@
 ---
-title: "Graphics"
-author: "KM"
+title: "Analysis of diferential expression results"
+author: "KM and GH"
 date: "`r date()`"
@@
 ```{r knit2synapse, eval=FALSE}
+# rmarkdown::render("graphics_t_statistics.Rmd")
 library(synapser)
@@
 ```{r setup, include=FALSE}
+library(synapser)
+synLogin()
 knitr::opts_chunk$set(echo = TRUE)
@@
 library(gtools)
+library(corrplot)
 library(pvclust)
@@
 
-```{r synapse.parameters, include=FALSE, cache=TRUE}
+```{r synapse.parameters, include=FALSE, cache=TRUE, eval=FALSE}
 thisFileName <- 'graphics_t_statistics.Rmd'
@@
 thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/',thisFileName))
+```
+
+```{r multiplot}
+# Multiple plot function
+#
+# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
+# - cols:   Number of columns in layout
+# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
+#
+# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
+# then plot 1 will go in the upper left, 2 will go in the upper right, and
+# 3 will go all the way across the bottom.
+#
+multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
+  library(grid)
+
+  # Make a list from the ... arguments and plotlist
+  plots <- c(list(...), plotlist)
+
+  numPlots = length(plots)
+
+  # If layout is NULL, then use 'cols' to determine layout
+  if (is.null(layout)) {
+    # Make the panel
+    # ncol: Number of columns of plots
+    # nrow: Number of rows needed, calculated from # of cols
+    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
+                    ncol = cols, nrow = ceiling(numPlots/cols))
+  }
+
+ if (numPlots==1) {
+    print(plots[[1]])
+
+  } else {
+    # Set up the page
+    grid.newpage()
+    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
+
+    # Make each plot, in the correct location
+    for (i in 1:numPlots) {
+      # Get the i,j matrix positions of the regions that contain this subplot
+      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
+
+      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
+                                      layout.pos.col = matchidx$col))
+    }
+  }
+}
 ```
@@
 mpp_acc <- readr::read_tsv(synGet("syn22079117")$path) %>% 
-  dplyr::select(Comparison, gene_id, t) %>% 
+  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% 
   dplyr::rename(`MSSM-Penn-Pitt ACC t-statistic` = t)
 mpp_dlpfc <- readr::read_tsv(synGet("syn22079472")$path) %>% 
-  dplyr::select(Comparison, gene_id, t) %>% 
+  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% 
   dplyr::rename(`MSSM-Penn-Pitt DLPFC t-statistic` = t)
 hbcc_dsACC <- readr::read_tsv(synGet("syn22086829")$path) %>% 
-  dplyr::select(Comparison, gene_id, t) %>% 
+  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% 
   dplyr::rename(`HBCC dsACC t-statistic` = t)
 hbcc_stg <- readr::read_tsv(synGet("syn22087386")$path) %>% 
-  dplyr::select(Comparison, gene_id, t) %>% 
+  dplyr::select(Comparison, gene_id, AveExpr, t) %>% 
   dplyr::rename(`HBCC STG t-statistic` = t)
 hbcc_sgACC <- readr::read_tsv(synGet("syn22085862")$path) %>% 
-  dplyr::select(Comparison, gene_id, t) %>% 
+  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% 
   dplyr::rename(`HBCC sgACC t-statistic` = t)
 hbcc_dlpfc <- readr::read_tsv(synGet("syn22084205")$path) %>% 
-  dplyr::select(Comparison, gene_id, t) %>% 
+  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% 
   dplyr::rename(`HBCC DLPFC t-statistic` = t)
@@
 # wget https://science.sciencemag.org/highwire/filestream/720110/field_highwire_adjunct_files/1/aat8127_Table_S1.xlsx
-gandal <- readxl::read_xlsx('../aat8127_Table_S1.xlsx', sheet = "DGE") %>% 
+gandal <- readxl::read_xlsx( synGet("syn22255217")$path, sheet = "DGE") %>% 
   dplyr::rename(gene_id = ensembl_gene_id,
@@
 df_one <- df_one %>% 
-  separate(gene_id, c("gene_id", "transcript")) %>% 
+  tidyr::separate(gene_id, c("gene_id", "transcript")) %>% 
   dplyr::select(-one_of("transcript"))
@@
 df_two <- df_two %>% 
-  separate(gene_id, c("gene_id", "transcript")) %>% 
+  tidyr::separate(gene_id, c("gene_id", "transcript")) %>% 
   dplyr::select(-one_of("transcript"))
@@
 plot(results, hang=-1)
+plot(hcl, hang=-1)
 corrplot(reorder, col = col2(200))
+```
+
+```{r custom.functions}
+
+library(metafor)
+
+# perform meta-analysis given fits from limma
+run_meta_analysis = function( tabList, coef, method="REML" ){
+
+  if( length(coef) > 1){
+    stop("Only supports one coefficient term")
+  }
+
+  # extract logFC and se for each model fit
+  resList = lapply( tabList, function(tab){
+    tab$se = tab$logFC /  tab$t
+    tab
+  })
+
+  # extract features names
+  featureNames = lapply(resList, function(res){
+    rownames(res)
+    })
+  featureNames = unique(unlist(featureNames))
+
+  # check that each fit has all features
+  # testIdentical = lapply(resList, function(res){
+  #   identical(sort(rownames(res)), sort(featureNames))
+  #   })
+  # if(  any(!unlist(testIdentical)) ){
+  #   stop("each entry in fitList must have all features with the same names")
+  # }
+
+  # apply meta analysis for each feature
+  resRMA = lapply(featureNames, function(feature){
+    # extract summary statistics for this feature from all studies
+    res = lapply(resList, function(res){
+      res[feature,]
+      })
+    df = do.call("rbind", res)
+    suppressWarnings(rma(yi=logFC, sei=se, data = df, method=method))
+    })
+  names(resRMA) = featureNames
+
+  # extract results
+  # Currently exatract simple results
+  # other results may be relevant if meta-analysing more than two studies
+  resTable = lapply(resRMA, function(x){
+    data.frame( logFC     = x$beta,
+                se        = x$se,
+                P.Value   = x$pval,
+                adj.P.Val = p.adjust(x$pval, "fdr"),
+                Q         = x$QE,
+                N         = x$k,
+                I2        = x$I2)
+    })
+  do.call("rbind", resTable)
+}
+
+plotVolcano = function(df, nshow=10){
+
+  library(org.Hs.eg.db)
+  library(ggrepel)
+
+  # indicate significant genes
+  df$isSignif = c("no","yes")[(df$adj.P.Val < 0.05)+1]
+
+  # show gene names
+  df$EnsID = gsub('\\..*$', '', rownames(df))
+
+  df_hugo = select(org.Hs.eg.db, df$EnsID, "SYMBOL", "ENSEMBL")
+  df = merge(df, df_hugo, by.x='EnsID', by.y="ENSEMBL")
+
+  # df = df[!duplicated(df$gene_id),]
+
+  # top significant genes
+  df2 = df[df$adj.P.Val <= sort(df$adj.P.Val )[nshow], ]
+
+  ggplot(df, aes( logFC, -log10(P.Value), color=isSignif)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, legend.position="none", plot.title = element_text(hjust = 0.5)) + xlab(bquote(log[2]~fold~change)) + ylab(bquote(-log[10]~P)) + scale_color_manual(values=c("grey", "red")) + geom_text_repel(data=df2, aes(logFC, -log10(P.Value), label=SYMBOL), segment.size=.5,  segment.color="black", color="black", force=1, nudge_x=.005, nudge_y=.5)
+}
+
+plot_forrest = function( resList, ensID){
+
+  library(patchwork)
+
+  # create a single data.frame with logFC and se
+  df_forrest = lapply(resList, function(x)x[ensID,])
+  df_forrest = do.call(rbind, df_forrest)
+  df_forrest$Dataset = rownames(df_forrest)
+  df_forrest$se = df_forrest$logFC /  df_forrest$t
+
+  # bind only shared columns
+  df = gtools::smartbind(df_forrest,  res_dlpfc[ensID,])
+  df$Dataset = factor(df$Dataset, rev(df$Dataset))
+
+  # set range to include zero
+  rng = range(c(with(df, logFC-se), with(df, logFC+se)))
+  if( rng[1] > 0 ){
+    rng[1] = 0
+  }
+  if( rng[2] < 0 ){
+    rng[2] = 0
+  }
+
+  # forrest plot
+  fig1 = ggplot(df, aes(Dataset, logFC)) + geom_point(color="navy") + geom_errorbar(aes(ymin = logFC - se, ymax=logFC + se), width=0.1) + theme_bw(12) + theme(aspect.ratio= .2, plot.title = element_text(hjust = 0.5)) + coord_flip() + ylim(rng) + xlab(bquote(log[2]~fold~change)) + ggtitle('RCSD1')
+
+  # Adjusted p-value
+  ymax = max(-log10(df$adj.P.Val))*1.05
+  fig2 = ggplot(df, aes(Dataset, -log10(adj.P.Val))) + geom_bar(stat="identity", fill="navy") + theme_bw(12) + theme( aspect.ratio= .5, plot.margin = unit(c(0,0,0,0), "cm"),
+    axis.title.y=element_blank(),
+          axis.text.y=element_blank(),
+          axis.ticks.y=element_blank()) + coord_flip() + ylab(bquote(-log[10]~adjusted~P)) + scale_y_continuous(expand=c(0,0), limits=c(0, ymax))
+
+  # plot_grid(fig1, fig2, align='vh')
+  fig1 + fig2
+}
+```
+
+# Violin plots
+```{r make.volcano}
+
+df_mpp_dlpfc <- readr::read_tsv(synGet("syn22079472")$path) %>% 
+  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% 
+  dplyr::filter(Comparison == 'Control_vs_SCZ') %>% 
+  dplyr::rename(`MSSM-Penn-Pitt DLPFC t-statistic` = t)  %>% 
+  as.data.frame  %>% unique
+rownames(df_mpp_dlpfc) = df_mpp_dlpfc$gene_id
+
+plotVolcano( df_mpp_dlpfc, 15) + ggtitle("MSSM-Penn-Pitt - DLPFC")
+```
+
+```{r meta_analysis.SZ}
+
+dlpfcList = list(MPP  = mpp_dlpfc %>% unique %>%
+                        dplyr::filter( Comparison == 'Control_vs_SCZ') %>%
+                        dplyr::rename(t = `MSSM-Penn-Pitt DLPFC t-statistic`) %>%
+                        as.data.frame,
+                 HBCC = hbcc_dlpfc %>%  unique %>%
+                        dplyr::filter( Comparison == 'Control_vs_SCZ') %>% 
+                        dplyr::rename(t = `HBCC DLPFC t-statistic`) %>%
+                        as.data.frame)
+
+rownames(dlpfcList[['MPP']]) = dlpfcList[['MPP']]$gene_id
+rownames(dlpfcList[['HBCC']]) = dlpfcList[['HBCC']]$gene_id
+
+res_dlpfc = run_meta_analysis( dlpfcList, "FE")
+res_dlpfc$Dataset = "Meta-analysis"
+
+accList = list(MPP  = mpp_acc %>% unique %>%
+                        dplyr::filter( Comparison == 'Control_vs_SCZ') %>%
+                        dplyr::rename(t = `MSSM-Penn-Pitt ACC t-statistic` )%>%
+                        as.data.frame,
+              HBCC  = hbcc_dsACC %>% unique %>%
+                        dplyr::filter( Comparison == 'Control_vs_SCZ') %>% 
+                        dplyr::rename(t = `HBCC dsACC t-statistic`) %>%
+                        as.data.frame)
+
+rownames(accList[['MPP']]) = accList[['MPP']]$gene_id
+rownames(accList[['HBCC']]) = accList[['HBCC']]$gene_id
+
+res_acc = run_meta_analysis( accList, "FE")
+res_acc$Dataset = "Meta-analysis"
+```
+
+```{r plot.meta_analysis}
+
+library(cowplot)
+
+table(res_dlpfc$adj.P.Val < 0.05)
+table(res_acc$adj.P.Val < 0.05)
+
+# forrest plot of meta-analysis
+# This is a bit of a mess
+ensID = 'ENSG00000198771.11'
+res_dlpfc[ensID,]
+
+lapply(dlpfcList, function(x) x[ensID,])
+
+plotVolcano( res_dlpfc, 20) + ggtitle("Meta analysis - DLPFC")
+
+ensID = rownames(res_dlpfc)[which.min(res_dlpfc$adj.P.Val)][1]
+
+plot_forrest(dlpfcList, ensID)
+
+plotVolcano( res_acc, 20) + ggtitle("Meta analysis - ACC")
 ```
@@
 df_three <- df_three %>% 
-  separate(gene_id, c("gene_id", "transcript")) %>% 
+  tidyr::separate(gene_id, c("gene_id", "transcript")) %>% 
   dplyr::select(-one_of("transcript"))
@@
 
-Code: `r thisFile`
+Code: r thisFile
 
-```{r store.files}
+```{r store.files, eval=FALSE}
 CODE <- Folder(name = "t-statistics across cohorts", parentId = "syn22012942")

### 2020-07-28 individual-cohorts/MSSM-Penn-Pitt_ACC_DE.Rmd
test Dx_Sex

@@
 ---
-title: "Differential Expression - MSSM-Penn-Pitt - ACC - Dx"
+title: "Differential Expression - MSSM-Penn-Pitt - ACC - Dx_Sex"
 author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
@@
                                  parentId ="syn22012953",
-                                 folderName = "Differential Expression - Dx")
+                                 folderName = "Differential Expression - Dx - Sex")
 ```
@@
 ## Load required libraries
-library(CovariateAnalysis) # get the package from devtools::install_github('th1vairam/CovariateAnalysis@dev')
 library(data.table)
-library(tidyr)
-library(plyr)
 library(dplyr)
-library(stringr)
 library(R.utils)
-library(ggplot2)
-library(limma)
 library(psych)
 library(lme4)
-library(biomaRt)
 library(variancePartition)
-library(edgeR)
 library(synapser)
 library(knitr)
-library(githubr) # get the package from devtools::install_github('brian-bot/githubr')
 
@@
 activityName = 'Differential Expression';
-activityDescription = 'Differential Expression - Dx';
+activityDescription = 'Differential Expression - Dx - Sex';
 
@@
 })
+# Test Dx-Sex
+COVARIATES <- COVARIATES %>% 
+  tidyr::unite("Dx_Sex",
+               dplyr::all_of(c("Dx", "Reported_Gender")),
+               remove = TRUE) %>% 
+  dplyr::mutate(`Dx_Sex` = as.factor(`Dx_Sex`))
 
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Dx", "ageOfDeath", "RIN2", "RIN", "rRNARate", "IntronicRate", "IntragenicRate", "IntergenicRate")
-random_effect <- c("Institution", "Reported_Gender")
+adjust.covars <- c("Dx_Sex", "ageOfDeath", "RIN2", "RIN", "rRNARate", "IntronicRate", "IntragenicRate", "IntergenicRate")
+random_effect <- c("Institution")
 ```
 
-### Normalisation - Dx
-1. Dx is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis)
+### Normalisation - Dx_Sex
+1. Dx_Sex is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis)
 
@@
 ### Residual calculation
-Calculate weighted residuals and add back "Dx" to the residuals
+Calculate weighted residuals and add back "Dx_Sex" to the residuals
 ```{r varsToAddBack}
 # Add variable of interest back to the residuals
-varsToAddIn = grep("Dx", colnames(ADJUSTED.FIT$design), value = T)
+varsToAddIn = grep("Dx_Sex", colnames(ADJUSTED.FIT$design), value = T)
 RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION +
@@
                                    data = COVARIATES, 
-                                   coefficient = c("DxOther", "DxControl"))
+                                   coefficient = c("Dx_SexControl_Female", "Dx_SexControl_Male"))
 L2 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
@@
                                    data = COVARIATES, 
-                                   coefficient = c("DxSCZ", "DxControl"))
+                                   coefficient = c("Dx_SexSCZ_Female", "Dx_SexSCZ_Male"))
 L3 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
@@
                                    data = COVARIATES, 
-                                   coefficient = c("DxSCZ", "DxOther"))
+                                   coefficient = c("Dx_SexOther_Female", "Dx_SexOther_Male"))
 
-L = cbind(L1, L2, L3)
+L = cbind(L1, L2, L3, L4)
 
@@
   rbindlist(idcol = 'Comparison') %>%
-  dplyr::mutate(Comparison = gsub('Dx','',Comparison),
+  dplyr::mutate(Comparison = gsub('Dx_Sex','',Comparison),
                 Direction = logFC/abs(logFC),
@@
 # Code
-CODE <- Folder(name = "Differential Expression - Dx", parentId = parentId)
+CODE <- Folder(name = "Differential Expression - Dx - Sex", parentId = parentId)
 CODE <- synStore(CODE)
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_ACC_Dx_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
-nEXP_OBJ = File('CMC_ACC_Dx_netResidualExpression.tsv', 
-                name = 'Dx Normalised, covariates removed residual expression (for network analysis)', 
+  write.table(file = 'CMC_ACC_Dx_Sex_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
+nEXP_OBJ = File('CMC_ACC_Dx_Sex_netResidualExpression.tsv', 
+                name = 'Dx_Sex Normalised, covariates removed residual expression (for network analysis)', 
                 parentId = CODE$properties$id)
@@
 # Store differential expression results
-write.table(DE, file = 'CMC_ACC_Dx_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
-DEXP_OBJ = File('CMC_ACC_Dx_DiffExpression.tsv', 
-                name = 'Differential Expression Results (Dx)', 
+write.table(DE, file = 'CMC_ACC_Dx_Sex_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
+DEXP_OBJ = File('CMC_ACC_Dx_Sex_DiffExpression.tsv', 
+                name = 'Differential Expression Results (Dx_Sex)', 
                 parentId = CODE$properties$id)
@@
 # Store adjusted fit linear mixed model object as RDS - Dx
-saveRDS(ADJUSTED.FIT, file = 'CMC_ACC_Dx_dream-output.rds')
-ADJ_OBJ = File('CMC_ACC_Dx_dream-output.rds',
-               name = 'Dream Output object - Dx', 
+saveRDS(ADJUSTED.FIT, file = 'CMC_ACC_Dx_Sex_dream-output.rds')
+ADJ_OBJ = File('CMC_ACC_Dx_Sex_dream-output.rds',
+               name = 'Dream Output object - Dx_Sex', 
                parentId = CODE$properties$id)

### 2020-07-29 individual-cohorts/MSSM-Penn-Pitt_DLPFC_DE.Rmd
manual contrasts for Ct vs. SCZOther

@@
 library(githubr) # get the package from devtools::install_github('brian-bot/githubr')
+library(WGCNA)
 
@@
 RESIDUAL.GENE_EXPRESSION = residuals(ADJUSTED.FIT, CQN.GENE_EXPRESSION)
-
-# Find PC of residual gene expression and significant covariates that are highly correlated with PCs
-residualSigCovars = runPCAandPlotCorrelations(RESIDUAL.GENE_EXPRESSION, COVARIATES,
-                                              'adjusted design(voom-normalized)',
-                                              isKeyPlot=TRUE)
-```
-
-### Sanity check
-```{r residualSigCovars.manual, fig.width=12, fig.height=8}
-residualSigCovars[["PC_res"]][[2]]$plotData
 ```
@@
                                    coefficient = c("DxSCZ", "DxOther"))
-
-L = cbind(L1, L2, L3)
+# Manually set comparison of SCZ-Other to Control
+L = cbind(L1, L2, L3, data.frame(L4 = c(1, -0.5, -0.5, 0, 0, 0, 0)))
 
@@
               glue::glue_collapse(names(L2[L2 != 0]), "-"),
-              glue::glue_collapse(names(L3[L3 != 0]), "-")
+              glue::glue_collapse(names(L3[L3 != 0]), "-"),
+              glue::glue("DxControl-DxOtherDxSCZ")
 )
@@
 # Get differential expression
-DE = lapply(glue::glue("L{1:3}"), function(i, FIT){
+DE = lapply(glue::glue("L{1:4}"), function(i, FIT){
   topTable(FIT, coef=i, number = Inf) %>%

### 2020-07-29 individual-cohorts/MSSM-Penn-Pitt_DLPFC_DE.Rmd
do not cache githubr

@@
 
-```{r synapse.parameters, include=FALSE, cache=TRUE}
+```{r synapse.parameters, include=FALSE, cache=FALSE}
 parentId = 'syn22043922';

### 2020-07-29 individual-cohorts/MSSM-Penn-Pitt_ACC_DE.Rmd
test DE_Sex

@@
 library(knitr)
+library(githubr)
+library(CovariateAnalysis)
+library(edgeR)
+library(WGCNA)
 
@@
 COVARIATES <- COVARIATES %>%
+  tibble::rownames_to_column(var = "samples") %>% 
   tidyr::unite("Dx_Sex",
@@
                remove = TRUE) %>% 
-  dplyr::mutate(`Dx_Sex` = as.factor(`Dx_Sex`))
+  dplyr::mutate(`Dx_Sex` = as.factor(`Dx_Sex`)) %>% 
+  tibble::column_to_rownames(var = "samples")
 
@@
 
-# Residuals after normalisation
-RESIDUAL.GENE_EXPRESSION = residuals(ADJUSTED.FIT, CQN.GENE_EXPRESSION)
-
-# Find PC of residual gene expression and significant covariates that are highly correlated with PCs
-residualSigCovars = runPCAandPlotCorrelations(RESIDUAL.GENE_EXPRESSION, COVARIATES,
-                                              'adjusted design(voom-normalized)',
-                                              isKeyPlot=TRUE)
-```
-
-### Sanity check
-```{r residualSigCovars.manual, fig.width=12, fig.height=8}
-residualSigCovars[["PC_res"]][[2]]$plotData
+RESIDUAL.GENE_EXPRESSION = stats::residuals(ADJUSTED.FIT, CQN.GENE_EXPRESSION)
 ```
@@
 # Eucledian tree based analysis
-COVARIATES.tmp = data.matrix(COVARIATES[,c('Dx','Institution')])
+COVARIATES.tmp = data.matrix(COVARIATES[,c('Dx_Sex','Institution')])
 
@@
                            main = "Sample dendrogram",
-                           groupLabels = colnames(COVARIATES.tmp[,c('Dx','Institution')]))
+                           groupLabels = colnames(COVARIATES.tmp[,c('Dx_Sex','Institution')]))
 ```
@@
 
-L = cbind(L1, L2, L3, L4)
+# Manually set comparison between SCZ+Other vs. Control 
+L = cbind(L1, L2, L3, 
+          data.frame(L4 = c(1,0,-0.5, 0, -0.5, 0, 0, 0, 0, 0, 0, 0)),
+          data.frame(L5 = c(0, 1, 0, -0.5, 0, -0.5, 0, 0, 0, 0, 0, 0)))
 
@@
               glue::glue_collapse(names(L2[L2 != 0]), "-"),
-              glue::glue_collapse(names(L3[L3 != 0]), "-")
+              glue::glue_collapse(names(L3[L3 != 0]), "-"),
+              glue::glue("Dx_SexControl_Female-Dx_SexOther_FemaleDx_SexSCZ_Female"),
+              glue::glue("Dx_SexControl_Male-Dx_SexOther_MaleDx_SexSCZ_Male"),
+
 )
@@
 # Get differential expression
-DE = lapply(glue::glue("L{1:3}"), function(i, FIT){
+DE = lapply(glue::glue("L{1:5}"), function(i, FIT){
   topTable(FIT, coef=i, number = Inf) %>%

### 2020-07-29 individual-cohorts/MSSM-Penn-Pitt_ACC_DE.Rmd
typo

@@
 
-```{r synapse.parameters, include=FALSE, cache=TRUE}
+```{r synapse.parameters, include=FALSE, cache=FALSE}
 parentId = 'syn22012953';
@@
               glue::glue("Dx_SexControl_Female-Dx_SexOther_FemaleDx_SexSCZ_Female"),
-              glue::glue("Dx_SexControl_Male-Dx_SexOther_MaleDx_SexSCZ_Male"),
-
+              glue::glue("Dx_SexControl_Male-Dx_SexOther_MaleDx_SexSCZ_Male")
 )

### 2020-07-29 individual-cohorts/graphics_t_statistics.Rmd
store meta-analysis to Synapse

@@
 ---
-title: "Analysis of diferential expression results"
+title: "Sex-specific Meta-analysis in MSSM-Penn-Pitt and HBCC ACC/DLPFC"
 author: "KM and GH"
@@
 knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/graphics_t_statistics.Rmd",
-                                          parentId ="syn22012942",
-                                          folderName = 't-statistics across cohorts')
+                                          parentId ="syn21435176",
+                                          folderName = 'Sex-specific Meta-analysis')
 ```
@@
 ```{r store.files, eval=FALSE}
-CODE <- Folder(name = "t-statistics across cohorts", parentId = "syn22012942")
+CODE <- Folder(name = "Sex-specific Meta-analysis", parentId = "syn21435176")
 CODE <- synStore(CODE)

### 2020-07-29 individual-cohorts/MSSM-Penn-Pitt_ACC_DE.Rmd
residuals bug fixed
fix plot scale

@@
 library(WGCNA)
+library(biomaRt)
 
@@
 
-RESIDUAL.GENE_EXPRESSION = stats::residuals(ADJUSTED.FIT, CQN.GENE_EXPRESSION)
+RESIDUAL.GENE_EXPRESSION = residuals(ADJUSTED.FIT, CQN.GENE_EXPRESSION)
 ```
@@
 
-p = ggplot(DE, aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1))
+p = ggplot(DE, aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1)) + ylim(c(0,100))
 p = p + scale_color_manual(values = c('green','grey','red')) + theme_bw() %+replace% theme(legend.position = 'top')

### 2020-07-29 individual-cohorts/graphics_t_statistics.Rmd

@@
 library(pvclust)
+library(metafor)
 
@@
 
-```{r synapse.parameters, include=FALSE, cache=TRUE, eval=FALSE}
+```{r synapse.parameters, include=FALSE, cache=FALSE, eval=FALSE}
 thisFileName <- 'graphics_t_statistics.Rmd'

### 2020-07-30 individual-cohorts/MSSM-Penn-Pitt_DLPFC_DE.Rmd
test condition Dx_Sex

--- 2020-07-29 individual-cohorts/MSSM-Penn-Pitt_DLPFC_DE.Rmd
@@
 ---
-title: "Differential Expression - MSSM-Penn-Pitt - DLPFC - Dx"
+title: "Differential Expression - MSSM-Penn-Pitt - DLPFC - Dx_Sex"
 author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
@@
                                  parentId ="syn22043922",
-                                 folderName = "Differential Expression - Dx")
+                                 folderName = "Differential Expression - Dx - Sex")
 ```
@@
 
-```{r synapse.parameters, include=FALSE, cache=TRUE}
+```{r synapse.parameters, include=FALSE, cache=FALSE}
 parentId = 'syn22043922';
 activityName = 'Differential Expression';
-activityDescription = 'Differential Expression - Dx';
+activityDescription = 'Differential Expression - Dx - Sex';
 
@@
 
+# Test Dx-Sex
+COVARIATES <- COVARIATES %>%
+  tibble::rownames_to_column(var = "samples") %>% 
+  tidyr::unite("Dx_Sex",
+               dplyr::all_of(c("Dx", "Reported_Gender")),
+               remove = TRUE) %>% 
+  dplyr::mutate(`Dx_Sex` = as.factor(`Dx_Sex`)) %>% 
+  tibble::column_to_rownames(var = "samples")
+
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Dx", "ageOfDeath", "RIN2", "RIN", "IntronicRate", "IntragenicRate")
+adjust.covars <- c("Dx_Sex", "ageOfDeath", "RIN2", "RIN", "IntronicRate", "IntragenicRate")
 random_effect <- c("Institution", "Reported_Gender")
@@
 
-### Normalisation - Dx
-1. Dx is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis)
+### Normalisation - Dx_Sex
+1. Dx_Sex is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis)
 
@@
                                    data = COVARIATES, 
-                                   coefficient = c("DxSCZ", "DxOther"))
+                                   coefficient = c("Dx_SexOther_Female", "Dx_SexOther_Male"))
 # Manually set comparison of SCZ-Other to Control
-L = cbind(L1, L2, L3, data.frame(L4 = c(1, -0.5, -0.5, 0, 0, 0, 0)))
+L = cbind(L1, L2, L3, 
+          data.frame(L4 = c(1,0,-0.5, 0, -0.5, 0, 0, 0, 0, 0, 0, 0)),
+          data.frame(L5 = c(0, 1, 0, -0.5, 0, -0.5, 0, 0, 0, 0, 0, 0)))
 
@@
               glue::glue_collapse(names(L3[L3 != 0]), "-"),
-              glue::glue("DxControl-DxOtherDxSCZ")
+              glue::glue("Dx_SexControl_Female-Dx_SexOther_FemaleDx_SexSCZ_Female"),
+              glue::glue("Dx_SexControl_Male-Dx_SexOther_MaleDx_SexSCZ_Male")
 )
@@
 # Get differential expression
-DE = lapply(glue::glue("L{1:4}"), function(i, FIT){
+DE = lapply(glue::glue("L{1:5}"), function(i, FIT){
   topTable(FIT, coef=i, number = Inf) %>%
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_DLPFC_Dx_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
-nEXP_OBJ = File('CMC_DLPFC_Dx_netResidualExpression.tsv', 
-                name = 'Dx Normalised, covariates removed residual expression (for network analysis)', 
+  write.table(file = 'CMC_DLPFC_Dx_Sex_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
+nEXP_OBJ = File('CMC_DLPFC_Dx_Sex_netResidualExpression.tsv', 
+                name = 'Dx_Sex Normalised, covariates removed residual expression (for network analysis)', 
                 parentId = CODE$properties$id)
@@
 # Store differential expression results
-write.table(DE, file = 'CMC_DLPFC_Dx_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
-DEXP_OBJ = File('CMC_DLPFC_Dx_DiffExpression.tsv', 
-                name = 'Differential Expression Results (Dx)', 
+write.table(DE, file = 'CMC_DLPFC_Dx_Sex_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
+DEXP_OBJ = File('CMC_DLPFC_Dx_Sex_DiffExpression.tsv', 
+                name = 'Differential Expression Results (Dx_Sex)', 
                 parentId = CODE$properties$id)
@@
 
-# Store adjusted fit linear mixed model object as RDS - Dx
-saveRDS(ADJUSTED.FIT, file = 'CMC_DLPFC_Dx_dream-output.rds')
-ADJ_OBJ = File('CMC_DLPFC_Dx_dream-output.rds',
-               name = 'Dream Output object - Dx', 
+# Store adjusted fit linear mixed model object as RDS - Dx_Sex
+saveRDS(ADJUSTED.FIT, file = 'CMC_DLPFC_Dx_Sex_dream-output.rds')
+ADJ_OBJ = File('CMC_DLPFC_Dx_Sex_dream-output.rds',
+               name = 'Dream Output object - Dx_Sex', 
                parentId = CODE$properties$id)

### 2020-07-30 individual-cohorts/MSSM-Penn-Pitt_DLPFC_DE.Rmd
remove missing variables

@@
 # Set variable class
-FactorCovariates <- c("Individual_ID", "Institution", "Reported_Gender", "LibraryBatch", "Dx", "RibozeroBatch", "FlowcellBatch")
+FactorCovariates <- c("Individual_ID", "Institution", "LibraryBatch", "Dx_Sex", "RibozeroBatch", "FlowcellBatch")
 COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
@@
 adjust.covars <- c("Dx_Sex", "ageOfDeath", "RIN2", "RIN", "IntronicRate", "IntragenicRate")
-random_effect <- c("Institution", "Reported_Gender")
+random_effect <- c("Institution")
 ```

### 2020-07-30 individual-cohorts/MSSM-Penn-Pitt_DLPFC_DE.Rmd
fix contrasts Dx_Sex

--- 2020-07-30 individual-cohorts/MSSM-Penn-Pitt_DLPFC_DE.Rmd
@@
 L = cbind(L1, L2, L3, 
-          data.frame(L4 = c(1,0,-0.5, 0, -0.5, 0, 0, 0, 0, 0, 0, 0)),
-          data.frame(L5 = c(0, 1, 0, -0.5, 0, -0.5, 0, 0, 0, 0, 0, 0)))
+          data.frame(L4 = c(1,0,-0.5, 0, -0.5, 0, 0, 0, 0, 0)),
+          data.frame(L5 = c(0, 1, 0, -0.5, 0, -0.5, 0, 0, 0, 0)))
 

### 2020-07-30 individual-cohorts/MSSM-Penn-Pitt_DLPFC_DE.Rmd
fix undefined columns

@@
 # Set variable class
-FactorCovariates <- c("Individual_ID", "Institution", "LibraryBatch", "Dx_Sex", "RibozeroBatch", "FlowcellBatch")
+FactorCovariates <- c("Individual_ID", "Institution", "LibraryBatch", "Dx", "Reported_Gender", "RibozeroBatch", "FlowcellBatch")
 COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)

### 2020-07-30 individual-cohorts/MSSM-Penn-Pitt_ACC_DE.Rmd
test Sex

@@
 ---
-title: "Differential Expression - MSSM-Penn-Pitt - ACC - Dx_Sex"
+title: "Differential Expression - MSSM-Penn-Pitt - ACC - Sex"
 author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
@@
                                  parentId ="syn22012953",
-                                 folderName = "Differential Expression - Dx - Sex")
+                                 folderName = "Differential Expression - Sex")
 ```
@@
 activityName = 'Differential Expression';
-activityDescription = 'Differential Expression - Dx - Sex';
+activityDescription = 'Differential Expression - Sex';
 
@@
 })
-# Test Dx-Sex
-COVARIATES <- COVARIATES %>%
-  tibble::rownames_to_column(var = "samples") %>% 
-  tidyr::unite("Dx_Sex",
-               dplyr::all_of(c("Dx", "Reported_Gender")),
-               remove = TRUE) %>% 
-  dplyr::mutate(`Dx_Sex` = as.factor(`Dx_Sex`)) %>% 
-  tibble::column_to_rownames(var = "samples")
+
+COVARIATES <- dplyr::rename(COVARIATES, Sex = Reported_Gender)
 
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Dx_Sex", "ageOfDeath", "RIN2", "RIN", "rRNARate", "IntronicRate", "IntragenicRate", "IntergenicRate")
+adjust.covars <- c("Sex", "ageOfDeath", "RIN2", "RIN", "rRNARate", "IntronicRate", "IntragenicRate", "IntergenicRate")
 random_effect <- c("Institution")
@@
 
-### Normalisation - Dx_Sex
-1. Dx_Sex is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis)
+### Normalisation - Sex
+
+1. Sex is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis)
 
@@
 ### Residual calculation
-Calculate weighted residuals and add back "Dx_Sex" to the residuals
+Calculate weighted residuals and add back "Reported_Gender" to the residuals
 ```{r varsToAddBack}
 # Add variable of interest back to the residuals
-varsToAddIn = grep("Dx_Sex", colnames(ADJUSTED.FIT$design), value = T)
+varsToAddIn = grep("Sex", colnames(ADJUSTED.FIT$design), value = T)
 RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION +
@@
 # Eucledian tree based analysis
-COVARIATES.tmp = data.matrix(COVARIATES[,c('Dx_Sex','Institution')])
+COVARIATES.tmp = data.matrix(COVARIATES[,c('Sex','Institution')])
 
@@
                            main = "Sample dendrogram",
-                           groupLabels = colnames(COVARIATES.tmp[,c('Dx_Sex','Institution')]))
+                           groupLabels = colnames(COVARIATES.tmp[,c('Sex','Institution')]))
 ```
@@
 
-### Differential expression analysis (SCZ - Control - Other)
+### Differential expression analysis (Female - Male)
 Genes that are differentially expressed at an FDR <= 0.05 are:
@@
 ```{r diffExp, fig.height=10, fig.width=15}
-L1 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
+L = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                    formula = formula, 
                                    data = COVARIATES, 
-                                   coefficient = c("Dx_SexControl_Female", "Dx_SexControl_Male"))
-L2 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
-                                   formula = formula, 
-                                   data = COVARIATES, 
-                                   coefficient = c("Dx_SexSCZ_Female", "Dx_SexSCZ_Male"))
-L3 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
-                                   formula = formula, 
-                                   data = COVARIATES, 
-                                   coefficient = c("Dx_SexOther_Female", "Dx_SexOther_Male"))
+                                   coefficient = c("SexFemale", "SexMale"))
 
-# Manually set comparison between SCZ+Other vs. Control 
-L = cbind(L1, L2, L3, 
-          data.frame(L4 = c(1,0,-0.5, 0, -0.5, 0, 0, 0, 0, 0, 0, 0)),
-          data.frame(L5 = c(0, 1, 0, -0.5, 0, -0.5, 0, 0, 0, 0, 0, 0)))
-
-contrast <- c(glue::glue_collapse(names(L1[L1 != 0]), "-"), 
-              glue::glue_collapse(names(L2[L2 != 0]), "-"),
-              glue::glue_collapse(names(L3[L3 != 0]), "-"),
-              glue::glue("Dx_SexControl_Female-Dx_SexOther_FemaleDx_SexSCZ_Female"),
-              glue::glue("Dx_SexControl_Male-Dx_SexOther_MaleDx_SexSCZ_Male")
-)
+contrast <- c(glue::glue_collapse(names(L[L != 0]), "-"))
 
@@
 # Get differential expression
-DE = lapply(glue::glue("L{1:5}"), function(i, FIT){
-  topTable(FIT, coef=i, number = Inf) %>%
+DE = topTable(FIT.CONTR, coef= "L1", number = Inf) %>% 
     rownameToFirstColumn('gene_id') %>%
     left_join(backgroundGenes)
-}, FIT.CONTR) 
+
 names(DE) <- contrast
@@
   rbindlist(idcol = 'Comparison') %>%
-  dplyr::mutate(Comparison = gsub('Dx_Sex','',Comparison),
+  dplyr::mutate(Comparison = gsub('Sex','',Comparison),
                 Direction = logFC/abs(logFC),
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_ACC_Dx_Sex_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
-nEXP_OBJ = File('CMC_ACC_Dx_Sex_netResidualExpression.tsv', 
-                name = 'Dx_Sex Normalised, covariates removed residual expression (for network analysis)', 
+  write.table(file = 'CMC_ACC_Sex_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
+nEXP_OBJ = File('CMC_ACC_Sex_netResidualExpression.tsv', 
+                name = 'Sex Normalised, covariates removed residual expression (for network analysis)', 
                 parentId = CODE$properties$id)
@@
 # Store differential expression results
-write.table(DE, file = 'CMC_ACC_Dx_Sex_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
-DEXP_OBJ = File('CMC_ACC_Dx_Sex_DiffExpression.tsv', 
-                name = 'Differential Expression Results (Dx_Sex)', 
+write.table(DE, file = 'CMC_ACC_Sex_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
+DEXP_OBJ = File('CMC_ACC_Sex_DiffExpression.tsv', 
+                name = 'Differential Expression Results (Sex)', 
                 parentId = CODE$properties$id)
@@
 # Store adjusted fit linear mixed model object as RDS - Dx
-saveRDS(ADJUSTED.FIT, file = 'CMC_ACC_Dx_Sex_dream-output.rds')
-ADJ_OBJ = File('CMC_ACC_Dx_Sex_dream-output.rds',
-               name = 'Dream Output object - Dx_Sex', 
+saveRDS(ADJUSTED.FIT, file = 'CMC_ACC_Sex_dream-output.rds')
+ADJ_OBJ = File('CMC_ACC_Sex_dream-output.rds',
+               name = 'Dream Output object - Sex', 
                parentId = CODE$properties$id)

### 2020-07-30 individual-cohorts/MSSM-Penn-Pitt_ACC_DE.Rmd
fix DE data.frame

@@
 # Get differential expression
-DE = topTable(FIT.CONTR, coef= "L1", number = Inf) %>% 
+DE = lapply(glue::glue("L1"), function(i, FIT){
+  topTable(FIT, coef=i, number = Inf) %>%
     rownameToFirstColumn('gene_id') %>% 
     left_join(backgroundGenes)
+}, FIT.CONTR)
 

### 2020-07-30 individual-cohorts/MSSM-Penn-Pitt_DLPFC_DE.Rmd
test Sex

@@
 ---
-title: "Differential Expression - MSSM-Penn-Pitt - DLPFC - Dx_Sex"
+title: "Differential Expression - MSSM-Penn-Pitt - DLPFC - Sex"
 author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
@@
                                  parentId ="syn22043922",
-                                 folderName = "Differential Expression - Dx - Sex")
+                                 folderName = "Differential Expression - Sex")
 ```
@@
 
-# Test Dx-Sex
-COVARIATES <- COVARIATES %>%
-  tibble::rownames_to_column(var = "samples") %>% 
-  tidyr::unite("Dx_Sex",
-               dplyr::all_of(c("Dx", "Reported_Gender")),
-               remove = TRUE) %>% 
-  dplyr::mutate(`Dx_Sex` = as.factor(`Dx_Sex`)) %>% 
-  tibble::column_to_rownames(var = "samples")
+COVARIATES <- dplyr::rename(COVARIATES, Sex = Reported_Gender)
 
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Dx_Sex", "ageOfDeath", "RIN2", "RIN", "IntronicRate", "IntragenicRate")
+adjust.covars <- c("Sex", "ageOfDeath", "RIN2", "RIN", "IntronicRate", "IntragenicRate")
 random_effect <- c("Institution")
@@
 
-### Normalisation - Dx_Sex
-1. Dx_Sex is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis)
+### Normalisation - Sex
+1. Sex is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis)
 
@@
 ### Residual calculation
-Calculate weighted residuals and add back "Dx_Sex" to the residuals
+Calculate weighted residuals and add back "Sex" to the residuals
 ```{r varsToAddBack}
 # Add variable of interest back to the residuals
-varsToAddIn = grep("Dx_Sex", colnames(ADJUSTED.FIT$design), value = T)
+varsToAddIn = grep("Sex", colnames(ADJUSTED.FIT$design), value = T)
 RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION +
@@
 ```{r diffExp, fig.height=10, fig.width=15}
-L1 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
+L = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                    formula = formula, 
                                    data = COVARIATES, 
-                                   coefficient = c("Dx_SexControl_Female", "Dx_SexControl_Male"))
-L2 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
-                                   formula = formula, 
-                                   data = COVARIATES, 
-                                   coefficient = c("Dx_SexSCZ_Female", "Dx_SexSCZ_Male"))
-L3 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
-                                   formula = formula, 
-                                   data = COVARIATES, 
-                                   coefficient = c("Dx_SexOther_Female", "Dx_SexOther_Male"))
-# Manually set comparison of SCZ-Other to Control
-L = cbind(L1, L2, L3, 
-          data.frame(L4 = c(1,0,-0.5, 0, -0.5, 0, 0, 0, 0, 0)),
-          data.frame(L5 = c(0, 1, 0, -0.5, 0, -0.5, 0, 0, 0, 0)))
-
-contrast <- c(glue::glue_collapse(names(L1[L1 != 0]), "-"), 
-              glue::glue_collapse(names(L2[L2 != 0]), "-"),
-              glue::glue_collapse(names(L3[L3 != 0]), "-"),
-              glue::glue("Dx_SexControl_Female-Dx_SexOther_FemaleDx_SexSCZ_Female"),
-              glue::glue("Dx_SexControl_Male-Dx_SexOther_MaleDx_SexSCZ_Male")
-)
+                                   coefficient = c("SexFemale", "SexMale"))
+contrast <- c(glue::glue_collapse(names(L[L != 0]), "-"))
 
@@
 # Get differential expression
-DE = lapply(glue::glue("L{1:5}"), function(i, FIT){
+DE = lapply(glue::glue("L1"), function(i, FIT){
   topTable(FIT, coef=i, number = Inf) %>%
@@
 # Code
-CODE <- Folder(name = "Differential Expression - Dx - Sex", parentId = parentId)
+CODE <- Folder(name = "Differential Expression - Sex", parentId = parentId)
 CODE <- synStore(CODE)
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_DLPFC_Dx_Sex_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
-nEXP_OBJ = File('CMC_DLPFC_Dx_Sex_netResidualExpression.tsv', 
-                name = 'Dx_Sex Normalised, covariates removed residual expression (for network analysis)', 
+  write.table(file = 'CMC_DLPFC_Sex_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
+nEXP_OBJ = File('CMC_DLPFC_Sex_netResidualExpression.tsv', 
+                name = 'Sex Normalised, covariates removed residual expression (for network analysis)', 
                 parentId = CODE$properties$id)
@@
 # Store differential expression results
-write.table(DE, file = 'CMC_DLPFC_Dx_Sex_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
-DEXP_OBJ = File('CMC_DLPFC_Dx_Sex_DiffExpression.tsv', 
-                name = 'Differential Expression Results (Dx_Sex)', 
+write.table(DE, file = 'CMC_DLPFC_Sex_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
+DEXP_OBJ = File('CMC_DLPFC_Sex_DiffExpression.tsv', 
+                name = 'Differential Expression Results (Sex)', 
                 parentId = CODE$properties$id)
@@
 
-# Store adjusted fit linear mixed model object as RDS - Dx_Sex
-saveRDS(ADJUSTED.FIT, file = 'CMC_DLPFC_Dx_Sex_dream-output.rds')
-ADJ_OBJ = File('CMC_DLPFC_Dx_Sex_dream-output.rds',
-               name = 'Dream Output object - Dx_Sex', 
+# Store adjusted fit linear mixed model object as RDS - Sex
+saveRDS(ADJUSTED.FIT, file = 'CMC_DLPFC_Sex_dream-output.rds')
+ADJ_OBJ = File('CMC_DLPFC_Sex_dream-output.rds',
+               name = 'Dream Output object - Sex', 
                parentId = CODE$properties$id)

### 2020-07-30 individual-cohorts/MSSM-Penn-Pitt_ACC_DE.Rmd
test Dx with 4th contrast

--- 2020-07-30 individual-cohorts/MSSM-Penn-Pitt_ACC_DE.Rmd
@@
 ---
-title: "Differential Expression - MSSM-Penn-Pitt - ACC - Sex"
+title: "Differential Expression - MSSM-Penn-Pitt - ACC - Dx"
 author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
@@
                                  parentId ="syn22012953",
-                                 folderName = "Differential Expression - Sex")
+                                 folderName = "Differential Expression - Dx")
 ```
@@
 ## Load required libraries
+library(CovariateAnalysis) # get the package from devtools::install_github('th1vairam/CovariateAnalysis@dev')
 library(data.table)
+library(tidyr)
+library(plyr)
 library(dplyr)
+library(stringr)
 library(R.utils)
+library(ggplot2)
+library(limma)
 library(psych)
 library(lme4)
+library(biomaRt)
 library(variancePartition)
+library(edgeR)
 library(synapser)
 library(knitr)
-library(githubr)
-library(CovariateAnalysis)
-library(edgeR)
-library(WGCNA)
-library(biomaRt)
+library(githubr) # get the package from devtools::install_github('brian-bot/githubr')
 
@@
 
-```{r synapse.parameters, include=FALSE, cache=FALSE}
+```{r synapse.parameters, include=FALSE, cache=TRUE}
 parentId = 'syn22012953';
 activityName = 'Differential Expression';
-activityDescription = 'Differential Expression - Sex';
+activityDescription = 'Differential Expression - Dx';
 
@@
 
-COVARIATES <- dplyr::rename(COVARIATES, Sex = Reported_Gender)
-
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Sex", "ageOfDeath", "RIN2", "RIN", "rRNARate", "IntronicRate", "IntragenicRate", "IntergenicRate")
-random_effect <- c("Institution")
+adjust.covars <- c("Dx", "ageOfDeath", "RIN2", "RIN", "rRNARate", "IntronicRate", "IntragenicRate", "IntergenicRate")
+random_effect <- c("Institution", "Reported_Gender")
 ```
 
-### Normalisation - Sex
-
-1. Sex is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis)
+### Normalisation - Dx
+1. Dx is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis)
 
@@
 
+# Residuals after normalisation
 RESIDUAL.GENE_EXPRESSION = residuals(ADJUSTED.FIT, CQN.GENE_EXPRESSION)
+
+# Find PC of residual gene expression and significant covariates that are highly correlated with PCs
+residualSigCovars = runPCAandPlotCorrelations(RESIDUAL.GENE_EXPRESSION, COVARIATES,
+                                              'adjusted design(voom-normalized)',
+                                              isKeyPlot=TRUE)
+```
+
+### Sanity check
+```{r residualSigCovars.manual, fig.width=12, fig.height=8}
+residualSigCovars[["PC_res"]][[2]]$plotData
 ```
@@
 ### Residual calculation
-Calculate weighted residuals and add back "Reported_Gender" to the residuals
+Calculate weighted residuals and add back "Dx" to the residuals
 ```{r varsToAddBack}
 # Add variable of interest back to the residuals
-varsToAddIn = grep("Sex", colnames(ADJUSTED.FIT$design), value = T)
+varsToAddIn = grep("Dx", colnames(ADJUSTED.FIT$design), value = T)
 RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION +
@@
 # Eucledian tree based analysis
-COVARIATES.tmp = data.matrix(COVARIATES[,c('Sex','Institution')])
+COVARIATES.tmp = data.matrix(COVARIATES[,c('Dx','Institution')])
 
@@
                            main = "Sample dendrogram",
-                           groupLabels = colnames(COVARIATES.tmp[,c('Sex','Institution')]))
+                           groupLabels = colnames(COVARIATES.tmp[,c('Dx','Institution')]))
 ```
@@
 
-### Differential expression analysis (Female - Male)
+### Differential expression analysis (SCZ - Control - Other)
 Genes that are differentially expressed at an FDR <= 0.05 are:
 ```{r set-formula}
-adjust.covars <- adjust.covars[!(adjust.covars %in% "RIN2")]
 formula <- glue::glue("~ 0 + ", glue::glue_collapse(adjust.covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))    
 ```
-
+#drop redundancy - RIN2 or RIN
+#use canCorPairs() to test colinearity, don't want var of interest to be correlated
+#leave out random effects
 `r formula`
@@
 ```{r diffExp, fig.height=10, fig.width=15}
-L = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
+L1 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                    formula = formula, 
                                    data = COVARIATES, 
-                                   coefficient = c("SexFemale", "SexMale"))
+                                   coefficient = c("DxOther", "DxControl"))
+L2 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
+                                   formula = formula, 
+                                   data = COVARIATES, 
+                                   coefficient = c("DxSCZ", "DxControl"))
+L3 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
+                                   formula = formula, 
+                                   data = COVARIATES, 
+                                   coefficient = c("DxSCZ", "DxOther"))
 
-contrast <- c(glue::glue_collapse(names(L[L != 0]), "-"))
+# Manually set comparison of SCZ-Other to Control
+L = cbind(L1, L2, L3, data.frame(L4 = c(1, -0.5, -0.5, 0, 0, 0, 0)))
+
+contrast <- c(glue::glue_collapse(names(L1[L1 != 0]), "-"), 
+              glue::glue_collapse(names(L2[L2 != 0]), "-"),
+              glue::glue_collapse(names(L3[L3 != 0]), "-"),
+              glue::glue("DxControl-DxOtherDxSCZ")
+)
 
@@
 # Get differential expression
-DE = topTable(FIT.CONTR, coef= "L1", number = Inf) %>% 
+DE = lapply(glue::glue("L{1:4}"), function(i, FIT){
+  topTable(FIT, coef=i, number = Inf) %>%
     rownameToFirstColumn('gene_id') %>% 
     left_join(backgroundGenes)
-
+}, FIT.CONTR) 
 names(DE) <- contrast
@@
   rbindlist(idcol = 'Comparison') %>%
-  dplyr::mutate(Comparison = gsub('Sex','',Comparison),
+  dplyr::mutate(Comparison = gsub('Dx','',Comparison),
                 Direction = logFC/abs(logFC),
@@
 
-p = ggplot(DE, aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1)) + ylim(c(0,100))
+p = ggplot(DE, aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1))
 p = p + scale_color_manual(values = c('green','grey','red')) + theme_bw() %+replace% theme(legend.position = 'top')
@@
 # Code
-CODE <- Folder(name = "Differential Expression - Dx - Sex", parentId = parentId)
+CODE <- Folder(name = "Differential Expression - Dx", parentId = parentId)
 CODE <- synStore(CODE)
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_ACC_Sex_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
-nEXP_OBJ = File('CMC_ACC_Sex_netResidualExpression.tsv', 
-                name = 'Sex Normalised, covariates removed residual expression (for network analysis)', 
+  write.table(file = 'CMC_ACC_Dx_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
+nEXP_OBJ = File('CMC_ACC_Dx_netResidualExpression.tsv', 
+                name = 'Dx Normalised, covariates removed residual expression (for network analysis)', 
                 parentId = CODE$properties$id)
@@
 # Store differential expression results
-write.table(DE, file = 'CMC_ACC_Sex_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
-DEXP_OBJ = File('CMC_ACC_Sex_DiffExpression.tsv', 
-                name = 'Differential Expression Results (Sex)', 
+write.table(DE, file = 'CMC_ACC_Dx_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
+DEXP_OBJ = File('CMC_ACC_Dx_DiffExpression.tsv', 
+                name = 'Differential Expression Results (Dx)', 
                 parentId = CODE$properties$id)
@@
 # Store adjusted fit linear mixed model object as RDS - Dx
-saveRDS(ADJUSTED.FIT, file = 'CMC_ACC_Sex_dream-output.rds')
-ADJ_OBJ = File('CMC_ACC_Sex_dream-output.rds',
-               name = 'Dream Output object - Sex', 
+saveRDS(ADJUSTED.FIT, file = 'CMC_ACC_Dx_dream-output.rds')
+ADJ_OBJ = File('CMC_ACC_Dx_dream-output.rds',
+               name = 'Dream Output object - Dx', 
                parentId = CODE$properties$id)

### 2020-07-30 individual-cohorts/MSSM-Penn-Pitt_ACC_DE.Rmd
remove errant legacy code

@@
 ```
-#drop redundancy - RIN2 or RIN
-#use canCorPairs() to test colinearity, don't want var of interest to be correlated
-#leave out random effects
 `r formula`

### 2020-07-30 individual-cohorts/MSSM-Penn-Pitt_ACC_DE.Rmd
fix manual contrasts
RIN2 too highly correlated for model fit

@@
 ```{r set-formula}
+adjust.covars <- adjust.covars[!(adjust.covars %in% "RIN2")]
 formula <- glue::glue("~ 0 + ", glue::glue_collapse(adjust.covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))
+
 ```
 `r formula`
+
+Note:removed RIN2 from formula as RIN and RIN2 are highly correlated.
 
@@
 # Manually set comparison of SCZ-Other to Control
-L = cbind(L1, L2, L3, data.frame(L4 = c(1, -0.5, -0.5, 0, 0, 0, 0)))
+L = cbind(L1, L2, L3, data.frame(L4 = c(1, -0.5, -0.5, 0, 0, 0, 0, 0, 0, 0)))
 

### 2020-07-30 individual-cohorts/MSSM-Penn-Pitt_ACC_DE.Rmd
scale covars

@@
 adjust.covars <- adjust.covars[!(adjust.covars %in% "RIN2")]
+adjust.covars <- glue::glue("scale({adjust.covars})")
+"Dx" -> adjust.covars[1]
 formula <- glue::glue("~ 0 + ", glue::glue_collapse(adjust.covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))  
-
 ```
@@
 # Manually set comparison of SCZ-Other to Control
-L = cbind(L1, L2, L3, data.frame(L4 = c(1, -0.5, -0.5, 0, 0, 0, 0, 0, 0, 0)))
+L = cbind(L1, L2, L3, data.frame(L4 = c(1, -0.5, -0.5, 0, 0, 0, 0, 0, 0)))
 

### 2020-07-30 individual-cohorts/HBCC_DLPFC_DE.Rmd
add comparison contrast

@@
 
-# Find PC of residual gene expression and significant covariates that are highly correlated with PCs
-residualSigCovars = runPCAandPlotCorrelations(RESIDUAL.GENE_EXPRESSION, COVARIATES,
-                                              'adjusted design(voom-normalized)',
-                                              isKeyPlot=TRUE)
-```
-
-### Sanity check
-```{r residualSigCovars.manual, fig.width=12, fig.height=8}
-residualSigCovars[["PC_res"]][[2]]$plotData
 ```
@@
 
-L = cbind(L1, L2, L3)
+# Manually set comparison of SCZ-Other to Control
+L = cbind(L1, L2, L3, data.frame(L4 = c(1, -0.5, -0.5, 0, 0, 0, 0, 0, 0)))
 

### 2020-07-31 individual-cohorts/HBCC_DLPFC_DE.Rmd
test Dx_Sex

@@
  ---
-title: "Differential Expression - HBCC - DLPFC - Dx"
+title: "Differential Expression - HBCC - DLPFC - Dx - Sex"
 author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
@@
                                  parentId ="syn22050019",
-                                 folderName = "Differential Expression - Dx")
+                                 folderName = "Differential Expression - Dx - Sex")
 ```
@@
 activityName = 'Differential Expression';
-activityDescription = 'Differential Expression - Dx';
+activityDescription = 'Differential Expression - Dx_Sex';
 
@@
 
+# Test Dx-Sex
+COVARIATES <- COVARIATES %>%
+  tibble::rownames_to_column(var = "samples") %>% 
+  tidyr::unite("Dx_Sex",
+               dplyr::all_of(c("Dx", "Reported_Gender")),
+               remove = TRUE) %>% 
+  dplyr::mutate(`Dx_Sex` = as.factor(`Dx_Sex`)) %>% 
+  tibble::column_to_rownames(var = "samples")
+
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Dx", "ageOfDeath", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate")
-random_effect <- c("Reported_Gender")
+adjust.covars <- c("Dx_Sex", "ageOfDeath", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate")
+random.effect <- c("Reported_Gender")
 ```
 
-### Normalisation - Dx
-1. Dx is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis)
+### Normalisation - Dx_Sex
+1. Dx_Sex is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis)
 
@@
                  glue::glue_collapse(adjust.covars, sep = ", ", last = " and "),
-                 ' as fixed effects and',
-                 glue::glue_collapse(random_effect, sep = ", ", last = " and "),' chosen as random effects'))
+                 ' as fixed effects.'))
 
@@
 RESIDUAL.GENE_EXPRESSION = residuals(ADJUSTED.FIT, CQN.GENE_EXPRESSION)
-
 ```
@@
 # Eucledian tree based analysis
-COVARIATES.tmp = data.matrix(COVARIATES[,c('Dx'), drop = FALSE])
+COVARIATES.tmp = data.matrix(COVARIATES[,c('Dx_Sex'), drop = FALSE])
 
@@
                            main = "Sample dendrogram",
-                           groupLabels = colnames(COVARIATES.tmp[,c('Dx'), drop = FALSE]))
+                           groupLabels = colnames(COVARIATES.tmp[,c('Dx_Sex'), drop = FALSE]))
 ```
@@
                                    data = COVARIATES, 
-                                   coefficient = c("DxSCZ", "DxOther"))
+                                   coefficient = c("Dx_SexOther_Female", "Dx_SexOther_Male"))
 
-# Manually set comparison of SCZ-Other to Control
-L = cbind(L1, L2, L3, data.frame(L4 = c(1, -0.5, -0.5, 0, 0, 0, 0, 0, 0)))
+# Manually set comparison between SCZ+Other vs. Control 
+L = cbind(L1, L2, L3, 
+          data.frame(L4 = c(1,0,-0.5, 0, -0.5, 0, 0, 0, 0, 0, 0)),
+          data.frame(L5 = c(0, 1, 0, -0.5, 0, -0.5, 0, 0, 0, 0, 0)))
 
@@
               glue::glue_collapse(names(L3[L3 != 0]), "-"),
-              glue::glue("DxControl-DxOtherDxSCZ")
+              glue::glue("Dx_SexControl_Female-Dx_SexOther_FemaleDx_SexSCZ_Female"),
+              glue::glue("Dx_SexControl_Male-Dx_SexOther_MaleDx_SexSCZ_Male"),
 )
@@
 ```{r associate.de, fig.height=10, fig.width=15}
+stop("check for errors")
 pl = list()

### 2020-07-31 individual-cohorts/HBCC_DLPFC_DE.Rmd
remove random_effects from model - not relevant

@@
 adjust.covars <- c("Dx_Sex", "ageOfDeath", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate")
-random.effect <- c("Reported_Gender")
 ```
@@
 
-col_type <- dplyr::select(COVARIATES, c(adjust.covars, random_effect)) %>%
+col_type <- dplyr::select(COVARIATES, c(adjust.covars)) %>%
   dplyr::summarise_all(class) %>% 
@@
 # Post adjusted formula
-formula <- glue::glue("~ ", glue::glue_collapse(adjust.covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))
+formula <- glue::glue("~ ", glue::glue_collapse(adjust.covars, sep = " + "))
 
@@
 ```{r set-formula}
-formula <- glue::glue("~ 0 + ", glue::glue_collapse(adjust.covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))    
+formula <- glue::glue("~ 0 + ", glue::glue_collapse(adjust.covars, sep = " + "))    
 ```

### 2020-07-31 individual-cohorts/HBCC_DLPFC_DE.Rmd
eBayes for fit

@@
 
-```{r synapse.parameters, include=FALSE, cache=TRUE}
+```{r synapse.parameters, include=FALSE, cache=FALSE}
 parentId = 'syn22050019';
@@
               glue::glue("Dx_SexControl_Female-Dx_SexOther_FemaleDx_SexSCZ_Female"),
-              glue::glue("Dx_SexControl_Male-Dx_SexOther_MaleDx_SexSCZ_Male"),
+              glue::glue("Dx_SexControl_Male-Dx_SexOther_MaleDx_SexSCZ_Male")
 )
@@
                        mart = mart)
+
+FIT.CONTR <- eBayes(FIT.CONTR)
 
@@
 ```{r associate.de, fig.height=10, fig.width=15}
-stop("check for errors")
 pl = list()

### 2020-07-31 individual-cohorts/HBCC_dsACC_DE.Rmd

@@
 
-L = cbind(L1, L2, L3)
-
+# Manually set comparison of SCZ-Other to Control
+L = cbind(L1, L2, L3, data.frame(L4 = c(1, -0.5, -0.5, 0, 0, 0, 0, 0)))
 contrast <- c(glue::glue_collapse(names(L1[L1 != 0]), "-"), 
               glue::glue_collapse(names(L2[L2 != 0]), "-"),
-              glue::glue_collapse(names(L3[L3 != 0]), "-")
+              glue::glue_collapse(names(L3[L3 != 0]), "-"),
+              glue::glue("DxControl-DxOtherDxSCZ")
 )

### 2020-07-31 individual-cohorts/HBCC_dsACC_DE.Rmd
test Dx_Sex

@@
 ---
-title: "Differential Expression - HBCC - dsACC - Dx"
+title: "Differential Expression - HBCC - dsACC - Dx - Sex"
 author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
@@
                                  parentId ="syn22084347",
-                                 folderName = "Differential Expression - Dx")
+                                 folderName = "Differential Expression - Dx - Sex")
 ```
@@
 
-```{r synapse.parameters, include=FALSE, cache=TRUE}
+```{r synapse.parameters, include=FALSE, cache=FALSE}
 parentId = 'syn22084347';
 activityName = 'Differential Expression';
-activityDescription = 'Differential Expression - Dx';
+activityDescription = 'Differential Expression - Dx_Sex';
 
@@
 
+# Test Dx-Sex
+COVARIATES <- COVARIATES %>%
+  tibble::rownames_to_column(var = "samples") %>% 
+  tidyr::unite("Dx_Sex",
+               dplyr::all_of(c("Dx", "Reported_Gender")),
+               remove = TRUE) %>% 
+  dplyr::mutate(`Dx_Sex` = as.factor(`Dx_Sex`)) %>% 
+  tibble::column_to_rownames(var = "samples")
+
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Dx", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate", "ageOfDeath")
-random_effect <- c("Reported_Gender", "IsolationBatch")
+adjust.covars <- c("Dx_Sex", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate", "ageOfDeath")
+random_effect <- c("IsolationBatch")
 ```
 
-### Normalisation - Dx
-1. Dx is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis)
+### Normalisation - Dx_Sex
+1. Dx_Sex is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis)
 
@@
                                    data = COVARIATES, 
-                                   coefficient = c("DxSCZ", "DxOther"))
+                                   coefficient = c("Dx_SexOther_Female", "Dx_SexOther_Male"))
 
-# Manually set comparison of SCZ-Other to Control
-L = cbind(L1, L2, L3, data.frame(L4 = c(1, -0.5, -0.5, 0, 0, 0, 0, 0)))
+# Manually set comparison between SCZ+Other vs. Control 
+L = cbind(L1, L2, L3, 
+          data.frame(L4 = c(1,0,-0.5, 0, -0.5, 0, 0, 0, 0, 0, 0)),
+          data.frame(L5 = c(0, 1, 0, -0.5, 0, -0.5, 0, 0, 0, 0, 0)))
+
 contrast <- c(glue::glue_collapse(names(L1[L1 != 0]), "-"), 
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_dsACC_Dx_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
-nEXP_OBJ = File('CMC_dsACC_Dx_netResidualExpression.tsv', 
-                name = 'Dx Normalised, covariates removed residual expression (for network analysis)', 
+  write.table(file = 'CMC_dsACC_Dx_Sex_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
+nEXP_OBJ = File('CMC_dsACC_Dx_Sex_netResidualExpression.tsv', 
+                name = 'Dx_Sex Normalised, covariates removed residual expression (for network analysis)', 
                 parentId = CODE$properties$id)
@@
 # Store differential expression results
-write.table(DE, file = 'CMC_dsACC_Dx_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
-DEXP_OBJ = File('CMC_dsACC_Dx_DiffExpression.tsv', 
-                name = 'Differential Expression Results (Dx)', 
+write.table(DE, file = 'CMC_dsACC_Dx_Sex_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
+DEXP_OBJ = File('CMC_dsACC_Dx_Sex_DiffExpression.tsv', 
+                name = 'Differential Expression Results (Dx_Sex)', 
                 parentId = CODE$properties$id)
@@
 
-# Store adjusted fit linear mixed model object as RDS - Dx
-saveRDS(ADJUSTED.FIT, file = 'CMC_dsACC_Dx_dream-output.rds')
-ADJ_OBJ = File('CMC_dsACC_Dx_dream-output.rds',
-               name = 'Dream Output object - Dx', 
+# Store adjusted fit linear mixed model object as RDS - Dx_Sex
+saveRDS(ADJUSTED.FIT, file = 'CMC_dsACC_Dx_Sex_dream-output.rds')
+ADJ_OBJ = File('CMC_dsACC_Dx_Sex_dream-output.rds',
+               name = 'Dream Output object - Dx_Sex', 
                parentId = CODE$properties$id)

### 2020-07-31 individual-cohorts/HBCC_DLPFC_DE.Rmd
test Sex

@@
  ---
-title: "Differential Expression - HBCC - DLPFC - Dx - Sex"
+title: "Differential Expression - HBCC - DLPFC - Sex"
 author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
@@
                                  parentId ="syn22050019",
-                                 folderName = "Differential Expression - Dx - Sex")
+                                 folderName = "Differential Expression - Sex")
 ```
@@
 activityName = 'Differential Expression';
-activityDescription = 'Differential Expression - Dx_Sex';
+activityDescription = 'Differential Expression - Sex';
 
@@
 
-# Test Dx-Sex
-COVARIATES <- COVARIATES %>%
-  tibble::rownames_to_column(var = "samples") %>% 
-  tidyr::unite("Dx_Sex",
-               dplyr::all_of(c("Dx", "Reported_Gender")),
-               remove = TRUE) %>% 
-  dplyr::mutate(`Dx_Sex` = as.factor(`Dx_Sex`)) %>% 
-  tibble::column_to_rownames(var = "samples")
+COVARIATES <- dplyr::rename(COVARIATES, Sex = Reported_Gender)
 
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Dx_Sex", "ageOfDeath", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate")
+adjust.covars <- c("Sex", "ageOfDeath", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate")
 ```
 
-### Normalisation - Dx_Sex
-1. Dx_Sex is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis)
+### Normalisation - Sex
+1. Sex is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis)
 
@@
 # Eucledian tree based analysis
-COVARIATES.tmp = data.matrix(COVARIATES[,c('Dx_Sex'), drop = FALSE])
+COVARIATES.tmp = data.matrix(COVARIATES[,c('Sex'), drop = FALSE])
 
@@
                            main = "Sample dendrogram",
-                           groupLabels = colnames(COVARIATES.tmp[,c('Dx_Sex'), drop = FALSE]))
+                           groupLabels = colnames(COVARIATES.tmp[,c('Sex'), drop = FALSE]))
 ```
@@
 ```{r diffExp, fig.height=10, fig.width=15}
-L1 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
+L = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                    formula = formula, 
                                    data = COVARIATES, 
-                                   coefficient = c("Dx_SexControl_Female", "Dx_SexControl_Male"))
-L2 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
-                                   formula = formula, 
-                                   data = COVARIATES, 
-                                   coefficient = c("Dx_SexSCZ_Female", "Dx_SexSCZ_Male"))
-L3 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
-                                   formula = formula, 
-                                   data = COVARIATES, 
-                                   coefficient = c("Dx_SexOther_Female", "Dx_SexOther_Male"))
+                                   coefficient = c("SexFemale", "SexMale"))
 
-# Manually set comparison between SCZ+Other vs. Control 
-L = cbind(L1, L2, L3, 
-          data.frame(L4 = c(1,0,-0.5, 0, -0.5, 0, 0, 0, 0, 0, 0)),
-          data.frame(L5 = c(0, 1, 0, -0.5, 0, -0.5, 0, 0, 0, 0, 0)))
-
-contrast <- c(glue::glue_collapse(names(L1[L1 != 0]), "-"), 
-              glue::glue_collapse(names(L2[L2 != 0]), "-"),
-              glue::glue_collapse(names(L3[L3 != 0]), "-"),
-              glue::glue("Dx_SexControl_Female-Dx_SexOther_FemaleDx_SexSCZ_Female"),
-              glue::glue("Dx_SexControl_Male-Dx_SexOther_MaleDx_SexSCZ_Male")
-)
+contrast <- c(glue::glue_collapse(names(L[L != 0]), "-"))
 
@@
   rbindlist(idcol = 'Comparison') %>%
-  dplyr::mutate(Comparison = gsub('Dx','',Comparison),
+  dplyr::mutate(Comparison = gsub('Sex','',Comparison),
                 Direction = logFC/abs(logFC),

### 2020-07-31 individual-cohorts/HBCC_dsACC_DE.Rmd
test Sex

--- 2020-07-31 individual-cohorts/HBCC_dsACC_DE.Rmd
@@
 ---
-title: "Differential Expression - HBCC - dsACC - Dx"
+title: "Differential Expression - HBCC - dsACC - Sex"
 author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
@@
                                  parentId ="syn22084347",
-                                 folderName = "Differential Expression - Dx")
+                                 folderName = "Differential Expression - Sex")
 ```
@@
 
-```{r synapse.parameters, include=FALSE, cache=TRUE}
+```{r synapse.parameters, include=FALSE, cache=FALSE}
 parentId = 'syn22084347';
 activityName = 'Differential Expression';
-activityDescription = 'Differential Expression - Dx';
+activityDescription = 'Differential Expression - ';
 
@@
 
+COVARIATES <- dplyr::rename(COVARIATES, Sex = Reported_Gender)
+
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Dx", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate", "ageOfDeath")
-random_effect <- c("Reported_Gender", "IsolationBatch")
+adjust.covars <- c("Sex", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate", "ageOfDeath")
+random_effect <- c("IsolationBatch")
 ```
 
-### Normalisation - Dx
-1. Dx is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis)
+### Normalisation - Sex
+1. Sex is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis)
 
@@
 ### Residual calculation
-Calculate weighted residuals and add back "Dx" to the residuals
+Calculate weighted residuals and add back "Sex" to the residuals
 ```{r varsToAddBack}
 # Add variable of interest back to the residuals
-varsToAddIn = grep("Dx", colnames(ADJUSTED.FIT$design), value = T)
+varsToAddIn = grep("Sex", colnames(ADJUSTED.FIT$design), value = T)
 RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION +
@@
 # Eucledian tree based analysis
-COVARIATES.tmp = data.matrix(COVARIATES[,c('Dx'), drop = FALSE])
+COVARIATES.tmp = data.matrix(COVARIATES[,c('Sex'), drop = FALSE])
 
@@
                            main = "Sample dendrogram",
-                           groupLabels = colnames(COVARIATES.tmp[,c('Dx'), drop = FALSE]))
+                           groupLabels = colnames(COVARIATES.tmp[,c('Sex'), drop = FALSE]))
 ```
@@
 ```{r diffExp, fig.height=10, fig.width=15}
-L1 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
+L = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                    formula = formula, 
                                    data = COVARIATES, 
-                                   coefficient = c("DxOther", "DxControl"))
-L2 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
-                                   formula = formula, 
-                                   data = COVARIATES, 
-                                   coefficient = c("DxSCZ", "DxControl"))
-L3 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
-                                   formula = formula, 
-                                   data = COVARIATES, 
-                                   coefficient = c("DxSCZ", "DxOther"))
+                                   coefficient = c("SexFemale", "SexMale"))
 
-# Manually set comparison of SCZ-Other to Control
-L = cbind(L1, L2, L3, data.frame(L4 = c(1, -0.5, -0.5, 0, 0, 0, 0, 0)))
-contrast <- c(glue::glue_collapse(names(L1[L1 != 0]), "-"), 
-              glue::glue_collapse(names(L2[L2 != 0]), "-"),
-              glue::glue_collapse(names(L3[L3 != 0]), "-"),
-              glue::glue("DxControl-DxOtherDxSCZ")
-)
+contrast <- c(glue::glue_collapse(names(L[L != 0]), "-"))
 
@@
 # Get differential expression
-DE = lapply(glue::glue("L{1:4}"), function(i, FIT){
+DE = lapply(glue::glue("L1"), function(i, FIT){
   topTable(FIT, coef=i, number = Inf) %>%
@@
 # Code
-CODE <- Folder(name = "Differential Expression - Dx", parentId = parentId)
+CODE <- Folder(name = "Differential Expression - Sex", parentId = parentId)
 CODE <- synStore(CODE)
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_dsACC_Dx_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
-nEXP_OBJ = File('CMC_dsACC_Dx_netResidualExpression.tsv', 
-                name = 'Dx Normalised, covariates removed residual expression (for network analysis)', 
+  write.table(file = 'CMC_dsACC_Sex_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
+nEXP_OBJ = File('CMC_dsACC_Sex_netResidualExpression.tsv', 
+                name = 'Sex Normalised, covariates removed residual expression (for network analysis)', 
                 parentId = CODE$properties$id)
@@
 # Store differential expression results
-write.table(DE, file = 'CMC_dsACC_Dx_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
-DEXP_OBJ = File('CMC_dsACC_Dx_DiffExpression.tsv', 
-                name = 'Differential Expression Results (Dx)', 
+write.table(DE, file = 'CMC_dsACC_Sex_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
+DEXP_OBJ = File('CMC_dsACC_Sex_DiffExpression.tsv', 
+                name = 'Differential Expression Results (Sex)', 
                 parentId = CODE$properties$id)
@@
 
-# Store adjusted fit linear mixed model object as RDS - Dx
-saveRDS(ADJUSTED.FIT, file = 'CMC_dsACC_Dx_dream-output.rds')
-ADJ_OBJ = File('CMC_dsACC_Dx_dream-output.rds',
-               name = 'Dream Output object - Dx', 
+# Store adjusted fit linear mixed model object as RDS - Sex
+saveRDS(ADJUSTED.FIT, file = 'CMC_dsACC_Sex_dream-output.rds')
+ADJ_OBJ = File('CMC_dsACC_Sex_dream-output.rds',
+               name = 'Dream Output object - Sex', 
                parentId = CODE$properties$id)

### 2020-07-31 individual-cohorts/meta-analysis.Rmd
metaanalysis by diagnosis

--- 2020-07-16 individual-cohorts/graphics_t_statistics.Rmd
@@
 ---
-title: "Analysis of diferential expression results"
+title: "Meta-analysis on diagnosis in MSSM-Penn-Pitt and HBCC ACC/DLPFC"
 author: "KM and GH"
@@
 ```{r knit2synapse, eval=FALSE}
-# rmarkdown::render("graphics_t_statistics.Rmd")
 library(synapser)
@@
 synLogin()
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/graphics_t_statistics.Rmd",
-                                          parentId ="syn22012942",
-                                          folderName = 't-statistics across cohorts')
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/meta-analysis.Rmd",
+                                          parentId ="syn22276904",
+                                          folderName = 'Meta-analysis on diagnosis')
 ```
@@
 library(synapser)
+
 synLogin()
-knitr::opts_chunk$set(echo = TRUE)
+
 library(ggplot2)
@@
 library(pvclust)
-
-synLogin()
+library(metafor)
+library(cowplot)
+library(org.Hs.eg.db)
+library(ggrepel)
+library(metafor)
+library(patchwork)
 
 knitr::opts_chunk$set(
-  echo=FALSE,
+  echo = FALSE,
   warning=FALSE,
@@
 
-```{r synapse.parameters, include=FALSE, cache=TRUE, eval=FALSE}
-thisFileName <- 'graphics_t_statistics.Rmd'
+```{r synapse.parameters, include=FALSE, cache=FALSE, eval=TRUE}
+thisFileName <- 'meta-analysis.Rmd'
 
@@
 thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/',thisFileName))
+```
+
+```{r input_data}
+mpp_acc <- readr::read_tsv(synGet("syn22079117")$path) %>% 
+  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val)
+mpp_dlpfc <- readr::read_tsv(synGet("syn22079472")$path) %>% 
+  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val)
+hbcc_dsACC <- readr::read_tsv(synGet("syn22086829")$path) %>% 
+  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) 
+hbcc_dlpfc <- readr::read_tsv(synGet("syn22084205")$path) %>% 
+  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val)
 ```
@@
 
-## t-statistics from differential expression results
-
-- separate by comparison
-
-```{r data}
-mpp_acc <- readr::read_tsv(synGet("syn22079117")$path) %>% 
-  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% 
-  dplyr::rename(`MSSM-Penn-Pitt ACC t-statistic` = t)
-mpp_dlpfc <- readr::read_tsv(synGet("syn22079472")$path) %>% 
-  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% 
-  dplyr::rename(`MSSM-Penn-Pitt DLPFC t-statistic` = t)
-hbcc_dsACC <- readr::read_tsv(synGet("syn22086829")$path) %>% 
-  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% 
-  dplyr::rename(`HBCC dsACC t-statistic` = t)
-hbcc_stg <- readr::read_tsv(synGet("syn22087386")$path) %>% 
-  dplyr::select(Comparison, gene_id, AveExpr, t) %>% 
-  dplyr::rename(`HBCC STG t-statistic` = t)
-hbcc_sgACC <- readr::read_tsv(synGet("syn22085862")$path) %>% 
-  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% 
-  dplyr::rename(`HBCC sgACC t-statistic` = t)
-hbcc_dlpfc <- readr::read_tsv(synGet("syn22084205")$path) %>% 
-  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% 
-  dplyr::rename(`HBCC DLPFC t-statistic` = t)
-# CMC Analysis 1
-CMC1 <- readr::read_tsv(synGet("syn5607603")$path) %>% 
-  dplyr::rename(gene_id = genes,
-                `CMC 1 t-statistic` = t) %>% 
-  dplyr::select(gene_id, `CMC 1 t-statistic`)
-# HBCC-DLPFC microarray
-CMC_HBCC <- readr::read_tsv(synGet("syn4941553")$path) %>% 
-  dplyr::rename(gene_id = genes,
-                `HBCC microarray t-statistic` = t) %>% 
-  dplyr::select(gene_id, `HBCC microarray t-statistic`)
-# Gandal, Science 2018, Table S1
-# Use Tab:DGE
-# wget https://science.sciencemag.org/highwire/filestream/720110/field_highwire_adjunct_files/1/aat8127_Table_S1.xlsx
-gandal <- readxl::read_xlsx( synGet("syn22255217")$path, sheet = "DGE") %>% 
-  dplyr::rename(gene_id = ensembl_gene_id,
-                `Gandal, SCZ t-statistic` = `SCZ.t.value`,
-                `Gandal, BD t-statistic` = `BD.t.value`,
-                `Gandal, ASD t-statistic` = `ASD.t.value`) %>%
-  dplyr::select(gene_id, `Gandal, SCZ t-statistic`,
-                `Gandal, BD t-statistic`, `Gandal, ASD t-statistic`)
-
-# AMP-AD RNASeq reprocessing GRCh37 - https://www.synapse.org/#!Synapse:syn9702085
-MSSM <- readr::read_tsv(synGet("syn10526259")$path) %>%
-  dplyr::filter(Comparison %in% "AD-CONTROL",
-                Model %in% "Diagnosis") %>%
-  dplyr::select(ensembl_gene_id, Tissue, t) %>%
-  dplyr::group_by(ensembl_gene_id, Tissue) %>% 
-  dplyr::filter(row_number() == 1) %>% 
-  tidyr::pivot_wider(names_from = Tissue,
-                     values_from = t) %>%
-  dplyr::rename(`AMP-AD MSSM FP AD t-statistic` = FP,
-                `AMP-AD MSSM IGF AD t-statistic` = IFG,
-                `AMP-AD MSSM PHG AD t-statistic` = PHG,
-                `AMP-AD MSSM STG AD t-statistic` = STG,
-                gene_id = ensembl_gene_id) %>% 
-  dplyr::select(gene_id, `AMP-AD MSSM STG AD t-statistic`) 
-MSSM <- MSSM[!is.na(MSSM$`AMP-AD MSSM STG AD t-statistic`),]
-
-MAYO <- readr::read_tsv(synGet("syn8468023")$path) %>% 
-  dplyr::filter(Comparison %in% "AD-CONTROL",
-                Model %in% "Diagnosis") %>%
-  dplyr::select(ensembl_gene_id, `Tissue.ref`, t) %>%
-  dplyr::group_by(ensembl_gene_id, `Tissue.ref`) %>% 
-  dplyr::filter(row_number() == 1) %>% 
-  tidyr::pivot_wider(names_from = `Tissue.ref`,
-                     values_from = t) %>% 
-  dplyr::rename(`AMP-AD MAYO CBE AD t-statistic` = CBE,
-                `AMP-AD MAYO TCX AD t-statistic` = TCX,
-                gene_id = ensembl_gene_id) %>% 
-  dplyr::select(-`AMP-AD MAYO TCX AD t-statistic`)
-
-ROSMAP <- readr::read_tsv(synGet("syn8456721")$path) %>%
-    dplyr::filter(Comparison %in% "AD-CONTROL",
-                Model %in% "Diagnosis") %>%
-  dplyr::rename(gene_id = ensembl_gene_id, 
-                `AMP-AD ROSMAP DLPFC AD t-statistic` = t) %>% 
-  dplyr::select(gene_id, `AMP-AD ROSMAP DLPFC AD t-statistic`)
-
-out_group <- inner_join(MSSM, MAYO) %>% 
-  inner_join(., ROSMAP) %>% 
-  inner_join(., gandal)
-
-df <- inner_join(mpp_acc, mpp_dlpfc,
-                 by = c("Comparison", "gene_id"))
-df <- inner_join(df, hbcc_dsACC, by = c("Comparison", "gene_id"))
-df <- inner_join(df, hbcc_stg, by = c("Comparison", "gene_id"))
-df <- inner_join(df, hbcc_sgACC, by = c("Comparison", "gene_id"))
-df <- inner_join(df, hbcc_dlpfc, by = c("Comparison", "gene_id"))
-```
-
-```{r functions}
-plot_t <- function(df, conditions){
-  p = list()
-  for(i in seq_along(conditions[,1])){
-  positions = conditions[i,]
-  x = df[positions[1]]
-  y = df[positions[2]]
-  computed_cor = cor(x, y, method = "spearman")
-  x_var <- names(x)
-  y_var <- names(y)
-  p[[i]] = ggplot(df, aes(x = .data[[x_var]], y = .data[[y_var]])) +
-  geom_hex() +
-  geom_abline() +
-  geom_smooth(method = "lm") +
-  theme_bw() +
-  xlim(8,-8) + 
-  ylim(8,-8) +
-    labs(
-    caption = paste0("spearman correlation between the statistics is ", round(computed_cor, digits = 3))
-  ) +
-    theme(aspect.ratio=1, 
-          axis.text=element_text(size=12),
-          plot.caption = element_text(face = "bold", size = "12"))
-  }
-multiplot(plotlist = p, cols = 2)
-}
-```
-
-# Control vs Other
-```{r plota, fig.width = 10, fig.height = 100}
-df_one <- df[df$Comparison %in% "Control_vs_Other",] %>% 
-  dplyr::select_if( ~ !all(is.na(.))) %>% 
-  dplyr::select(-Comparison)
-
-#Remove transcript Ids
-df_one <- df_one %>% 
-  tidyr::separate(gene_id, c("gene_id", "transcript")) %>% 
-  dplyr::select(-one_of("transcript"))
-df_one <- inner_join(df_one, out_group, by = c("gene_id")) %>% 
-  group_by(gene_id) %>% 
-  dplyr::filter(row_number()==1) %>% 
-  tibble::column_to_rownames(var = "gene_id")
-
-conditions <- gtools::combinations(n = ncol(df_one), 2, seq_len(ncol(df_one)))
-plot_t(df_one, conditions)
-```
-
-```{r heatmapa, fig.width = 8, fig.height = 8 }
-colnames(df_one) <- gsub(" t-statistic", "", colnames(df_one))
-
-cor <- cor(df_one)
-# clustering based on distance
-hcl = hclust(as.dist(1-cor))
-# reorder based on clustering
-reorder = cor[hcl$order, hcl$order]
-results <- pvclust(cor)
-# plot clustering
-plot(results, hang=-1)
-col2 <- colorRampPalette(rev(c("#67001F", "#B2182B", "#D6604D", "#F4A582",
-                           "#FDDBC7", "#FFFFFF", "#D1E5F0", "#92C5DE",
-                           "#4393C3", "#2166AC", "#053061")))
-corrplot(reorder, col = col2(200))
-```
-
-# Control vs SCZ
-```{r plotb, fig.width = 10, fig.height = 150}
-df_two <- df[df$Comparison %in% "Control_vs_SCZ",] %>% 
-  dplyr::select_if( ~ !all(is.na(.))) %>% 
-  dplyr::select(-Comparison)
-
-#Remove transcript Ids
-df_two <- df_two %>% 
-  tidyr::separate(gene_id, c("gene_id", "transcript")) %>% 
-  dplyr::select(-one_of("transcript"))
-df_two <- inner_join(df_two, CMC1, by = c("gene_id")) 
-df_two <- inner_join(df_two, CMC_HBCC, by = c("gene_id"))
-df_two <- inner_join(df_two, out_group, by = c("gene_id")) %>% 
-  group_by(gene_id) %>% 
-  dplyr::filter(row_number()==1) %>% 
-  tibble::column_to_rownames(var = "gene_id")
-
-conditions <- gtools::combinations(n = ncol(df_two), 2, seq_len(ncol(df_two)))
-plot_t(df_two, conditions)
-```
-
-```{r heatmapb, fig.width = 8, fig.height = 8 }
-colnames(df_two) <- gsub(" t-statistic", "", colnames(df_two))
-
-cor <- cor(df_two)
-# clustering based on distance
-hcl = hclust(as.dist(1-cor))
-# reorder based on clustering
-reorder = cor[hcl$order, hcl$order]
-results <- pvclust(cor)
-# plot clustering
-plot(results, hang=-1)
-plot(hcl, hang=-1)
-corrplot(reorder, col = col2(200))
-```
-
 ```{r custom.functions}
-
-library(metafor)
-
 # perform meta-analysis given fits from limma
-run_meta_analysis = function( tabList, coef, method="REML" ){
-
-  if( length(coef) > 1){
-    stop("Only supports one coefficient term")
-  }
+run_meta_analysis = function(tabList, method="FE") {
 
@@
 plotVolcano = function(df, nshow=10){
-
-  library(org.Hs.eg.db)
-  library(ggrepel)
-
   # indicate significant genes
@@
 
-plot_forrest = function( resList, ensID){
-
-  library(patchwork)
-
+plot_forrest = function(resList, ensID) { 
   # create a single data.frame with logFC and se
-  df_forrest = lapply(resList, function(x)x[ensID,])
+  df_forrest = lapply(resList, function(x) if(ensID %in% x[ensID,]) {
+    x[ensID,]
+  } 
+  )
+  df_forrest[sapply(df_forrest, is.null)] <- NULL
   df_forrest = do.call(rbind, df_forrest)
@@
 
-# Violin plots
-```{r make.volcano}
+# Control vs SCZ
+```{r meta_analysis.SZ}
+make_list <- function(mpp, hbcc, string) {
+  list <- list(MPP = mpp %>%
+                 unique() %>% 
+                 dplyr::filter(Comparison == string) %>% 
+                 as.data.frame,
+               HBCC = hbcc %>% 
+                 unique() %>% 
+                 dplyr::filter(Comparison == string) %>% 
+                 as.data.frame)
+}
 
-df_mpp_dlpfc <- readr::read_tsv(synGet("syn22079472")$path) %>% 
-  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% 
-  dplyr::filter(Comparison == 'Control_vs_SCZ') %>% 
-  dplyr::rename(`MSSM-Penn-Pitt DLPFC t-statistic` = t)  %>% 
-  as.data.frame  %>% unique
-rownames(df_mpp_dlpfc) = df_mpp_dlpfc$gene_id
-
-plotVolcano( df_mpp_dlpfc, 15) + ggtitle("MSSM-Penn-Pitt - DLPFC")
-```
-
-```{r meta_analysis.SZ}
-
-dlpfcList = list(MPP  = mpp_dlpfc %>% unique %>%
-                        dplyr::filter( Comparison == 'Control_vs_SCZ') %>%
-                        dplyr::rename(t = `MSSM-Penn-Pitt DLPFC t-statistic`) %>%
-                        as.data.frame,
-                 HBCC = hbcc_dlpfc %>%  unique %>%
-                        dplyr::filter( Comparison == 'Control_vs_SCZ') %>% 
-                        dplyr::rename(t = `HBCC DLPFC t-statistic`) %>%
-                        as.data.frame)
+dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Control_vs_SCZ")
+accList <- make_list(mpp_acc, hbcc_dsACC, string = "Control_vs_SCZ")
 
@@
 res_dlpfc$Dataset = "Meta-analysis"
-
-accList = list(MPP  = mpp_acc %>% unique %>%
-                        dplyr::filter( Comparison == 'Control_vs_SCZ') %>%
-                        dplyr::rename(t = `MSSM-Penn-Pitt ACC t-statistic` )%>%
-                        as.data.frame,
-              HBCC  = hbcc_dsACC %>% unique %>%
-                        dplyr::filter( Comparison == 'Control_vs_SCZ') %>% 
-                        dplyr::rename(t = `HBCC dsACC t-statistic`) %>%
-                        as.data.frame)
 
@@
 
-```{r plot.meta_analysis}
-
-library(cowplot)
-
-table(res_dlpfc$adj.P.Val < 0.05)
-table(res_acc$adj.P.Val < 0.05)
-
-# forrest plot of meta-analysis
-# This is a bit of a mess
-ensID = 'ENSG00000198771.11'
-res_dlpfc[ensID,]
-
-lapply(dlpfcList, function(x) x[ensID,])
+```{r plot.meta_analysisSCZ}
+kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
+kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
 
@@
 
-ensID = rownames(res_dlpfc)[which.min(res_dlpfc$adj.P.Val)][1]
+plotVolcano( res_acc, 20) + ggtitle("Meta analysis - ACC")
+```
 
-plot_forrest(dlpfcList, ensID)
+# Control vs Other/SCZ
+```{r meta_analysis.SCZOther}
+make_list <- function(mpp, hbcc, string) {
+  list <- list(MPP = mpp %>%
+                 unique() %>% 
+                 dplyr::filter(Comparison == string) %>% 
+                 as.data.frame,
+               HBCC = hbcc %>% 
+                 unique() %>% 
+                 dplyr::filter(Comparison == string) %>% 
+                 as.data.frame)
+}
+
+dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Control_vs_OtherSCZ")
+accList <- make_list(mpp_acc, hbcc_dsACC, string = "Control_vs_OtherSCZ")
+
+rownames(dlpfcList[['MPP']]) = dlpfcList[['MPP']]$gene_id
+rownames(dlpfcList[['HBCC']]) = dlpfcList[['HBCC']]$gene_id
+
+res_dlpfc = run_meta_analysis( dlpfcList, "FE")
+res_dlpfc$Dataset = "Meta-analysis"
+
+rownames(accList[['MPP']]) = accList[['MPP']]$gene_id
+rownames(accList[['HBCC']]) = accList[['HBCC']]$gene_id
+
+res_acc = run_meta_analysis( accList, "FE")
+res_acc$Dataset = "Meta-analysis"
+```
+
+```{r plot.meta_analysisSCZOther}
+kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
+kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
+
+plotVolcano( res_dlpfc, 20) + ggtitle("Meta analysis - DLPFC")
+
+plotVolcano( res_acc, 20) + ggtitle("Meta analysis - ACC")
+```
+
+# Control vs Other
+```{r meta_analysis.CTOther}
+make_list <- function(mpp, hbcc, string) {
+  list <- list(MPP = mpp %>%
+                 unique() %>% 
+                 dplyr::filter(Comparison == string) %>% 
+                 as.data.frame,
+               HBCC = hbcc %>% 
+                 unique() %>% 
+                 dplyr::filter(Comparison == string) %>% 
+                 as.data.frame)
+}
+
+dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Control_vs_Other")
+accList <- make_list(mpp_acc, hbcc_dsACC, string = "Control_vs_Other")
+
+rownames(dlpfcList[['MPP']]) = dlpfcList[['MPP']]$gene_id
+rownames(dlpfcList[['HBCC']]) = dlpfcList[['HBCC']]$gene_id
+
+res_dlpfc = run_meta_analysis( dlpfcList, "FE")
+res_dlpfc$Dataset = "Meta-analysis"
+
+rownames(accList[['MPP']]) = accList[['MPP']]$gene_id
+rownames(accList[['HBCC']]) = accList[['HBCC']]$gene_id
+
+res_acc = run_meta_analysis( accList, "FE")
+res_acc$Dataset = "Meta-analysis"
+```
+
+```{r plot.meta_analysisCTOther}
+kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
+kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
+
+plotVolcano( res_dlpfc, 20) + ggtitle("Meta analysis - DLPFC")
 
@@
 # Other vs SCZ
-```{r plotc, fig.width = 10, fig.height = 100}
-df_three <- df[df$Comparison %in% "Other_vs_SCZ",] %>% 
-  dplyr::select_if( ~ !all(is.na(.))) %>% 
-  dplyr::select(-Comparison)
+```{r meta_analysis.OtherSCZ}
+make_list <- function(mpp, hbcc, string) {
+  list <- list(MPP = mpp %>%
+                 unique() %>% 
+                 dplyr::filter(Comparison == string) %>% 
+                 as.data.frame,
+               HBCC = hbcc %>% 
+                 unique() %>% 
+                 dplyr::filter(Comparison == string) %>% 
+                 as.data.frame)
+}
 
-#Remove transcript Ids
-df_three <- df_three %>% 
-  tidyr::separate(gene_id, c("gene_id", "transcript")) %>% 
-  dplyr::select(-one_of("transcript"))
-df_three <- inner_join(df_three, out_group, by = c("gene_id")) %>% 
-  group_by(gene_id) %>% 
-  dplyr::filter(row_number()==1) %>% 
-  tibble::column_to_rownames(var = "gene_id")
+dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Other_vs_SCZ")
+accList <- make_list(mpp_acc, hbcc_dsACC, string = "Other_vs_SCZ")
 
-conditions <- gtools::combinations(n = ncol(df_three), 2, seq_len(ncol(df_three)))
-plot_t(df_three, conditions)
+rownames(dlpfcList[['MPP']]) = dlpfcList[['MPP']]$gene_id
+rownames(dlpfcList[['HBCC']]) = dlpfcList[['HBCC']]$gene_id
+
+res_dlpfc = run_meta_analysis( dlpfcList, "FE")
+res_dlpfc$Dataset = "Meta-analysis"
+
+rownames(accList[['MPP']]) = accList[['MPP']]$gene_id
+rownames(accList[['HBCC']]) = accList[['HBCC']]$gene_id
+
+res_acc = run_meta_analysis( accList, "FE")
+res_acc$Dataset = "Meta-analysis"
 ```
 
-```{r heatmapc, fig.width = 8, fig.height = 8 }
-colnames(df_three) <- gsub(" t-statistic", "", colnames(df_three))
+```{r plot.meta_analysisOtherSCZ}
+kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
+kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
 
-cor <- cor(df_three)
-# clustering based on distance
-hcl = hclust(as.dist(1-cor))
-# reorder based on clustering
-reorder = cor[hcl$order, hcl$order]
-results <- pvclust(cor)
-# plot clustering
-plot(results, hang=-1)
-corrplot(reorder, col = col2(200))
+plotVolcano( res_dlpfc, 20) + ggtitle("Meta analysis - DLPFC")
+
+plotVolcano( res_acc, 20) + ggtitle("Meta analysis - ACC")
 ```
 
-Code: r thisFile
+## RCSD1
 
-```{r store.files, eval=FALSE}
-CODE <- Folder(name = "t-statistics across cohorts", parentId = "syn22012942")
-CODE <- synStore(CODE)
+Forest plot of example top gene.
+
+```{r specific-gene}
+# forrest plot of meta-analysis
+ensID = 'ENSG00000198771.11'
+kable(res_dlpfc[ensID,])
+lapply(dlpfcList, function(x) x[ensID,])
+ensID = rownames(res_dlpfc)[which.min(res_dlpfc$adj.P.Val)][1]
+plot_forrest(dlpfcList, ensID)
 ```
+
+### Source

### 2020-07-31 individual-cohorts/meta-analysis.Rmd
metaanalysis by diagnosis and sex

@@
 ---
-title: "Meta-analysis on diagnosis in MSSM-Penn-Pitt and HBCC ACC/DLPFC"
+title: "Meta-analysis on diagnosis and sex in MSSM-Penn-Pitt and HBCC ACC/DLPFC"
 author: "KM and GH"
@@
                                           parentId ="syn22276904",
-                                          folderName = 'Meta-analysis on diagnosis')
+                                          folderName = 'Meta-analysis on diagnosis and sex')
 ```
@@
 ```{r input_data}
-mpp_acc <- readr::read_tsv(synGet("syn22079117")$path) %>% 
+mpp_acc <- readr::read_tsv(synGet("syn22276990")$path) %>% 
   dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val)
-mpp_dlpfc <- readr::read_tsv(synGet("syn22079472")$path) %>% 
+mpp_dlpfc <- readr::read_tsv(synGet("syn22277179")$path) %>% 
   dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val)
-hbcc_dsACC <- readr::read_tsv(synGet("syn22086829")$path) %>% 
+hbcc_dsACC <- readr::read_tsv(synGet("syn22280364")$path) %>% 
   dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) 
-hbcc_dlpfc <- readr::read_tsv(synGet("syn22084205")$path) %>% 
+hbcc_dlpfc <- readr::read_tsv(synGet("syn22280045")$path) %>% 
   dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val)
@@
 ### Source
+[markdown](`r thisFile`)

### 2020-07-31 individual-cohorts/meta-analysis.Rmd
fix contrasts for all conditions

@@
 
-# Control vs SCZ
+# Control - Female vs Male
 ```{r meta_analysis.SZ}
@@
 
-dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Control_vs_SCZ")
-accList <- make_list(mpp_acc, hbcc_dsACC, string = "Control_vs_SCZ")
+dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Control_Female_vs_Control_Male")
+accList <- make_list(mpp_acc, hbcc_dsACC, string = "Control_Female_vs_Control_Male")
 
@@
 
-# Control vs Other/SCZ
+# SCZ - Female vs Male
 ```{r meta_analysis.SCZOther}
@@
 
-dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Control_vs_OtherSCZ")
-accList <- make_list(mpp_acc, hbcc_dsACC, string = "Control_vs_OtherSCZ")
+dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "SCZ_Female_vs_SCZ_Male")
+accList <- make_list(mpp_acc, hbcc_dsACC, string = "SCZ_Female_vs_SCZ_Male")
 
@@
 
-# Control vs Other
+#Other - Female vs Male
 ```{r meta_analysis.CTOther}
@@
 
-dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Control_vs_Other")
-accList <- make_list(mpp_acc, hbcc_dsACC, string = "Control_vs_Other")
+dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Other_Female_vs_Other_Male")
+accList <- make_list(mpp_acc, hbcc_dsACC, string = "Other_Female_vs_Other_Male")
 
@@
 
-# Other vs SCZ
+# Female - Control vs Other/SCZ
 ```{r meta_analysis.OtherSCZ}
@@
 
-dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Other_vs_SCZ")
-accList <- make_list(mpp_acc, hbcc_dsACC, string = "Other_vs_SCZ")
+dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Control_Female_vs_Other_FemaleSCZ_Female")
+accList <- make_list(mpp_acc, hbcc_dsACC, string = "Control_Female_vs_Other_FemaleSCZ_Female")
+
+rownames(dlpfcList[['MPP']]) = dlpfcList[['MPP']]$gene_id
+rownames(dlpfcList[['HBCC']]) = dlpfcList[['HBCC']]$gene_id
+
+res_dlpfc = run_meta_analysis( dlpfcList, "FE")
+res_dlpfc$Dataset = "Meta-analysis"
+
+rownames(accList[['MPP']]) = accList[['MPP']]$gene_id
+rownames(accList[['HBCC']]) = accList[['HBCC']]$gene_id
+
+res_acc = run_meta_analysis( accList, "FE")
+res_acc$Dataset = "Meta-analysis"
+```
+
+```{r plot.meta_analysisOtherSCZ}
+kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
+kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
+
+plotVolcano( res_dlpfc, 20) + ggtitle("Meta analysis - DLPFC")
+
+plotVolcano( res_acc, 20) + ggtitle("Meta analysis - ACC")
+```
+
+# Male - Control vs Other/SCZ
+```{r meta_analysis.OtherSCZ}
+make_list <- function(mpp, hbcc, string) {
+  list <- list(MPP = mpp %>%
+                 unique() %>% 
+                 dplyr::filter(Comparison == string) %>% 
+                 as.data.frame,
+               HBCC = hbcc %>% 
+                 unique() %>% 
+                 dplyr::filter(Comparison == string) %>% 
+                 as.data.frame)
+}
+
+dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Control_Male_vs_Other_MaleSCZ_Male")
+accList <- make_list(mpp_acc, hbcc_dsACC, string = "Control_Male_vs_Other_MaleSCZ_Male")
 

### 2020-07-31 individual-cohorts/meta-analysis.Rmd
metaanalysis on sex

--- 2020-07-31 individual-cohorts/meta-analysis.Rmd
@@
 ---
-title: "Meta-analysis on diagnosis and sex in MSSM-Penn-Pitt and HBCC ACC/DLPFC"
+title: "Meta-analysis on sex in MSSM-Penn-Pitt and HBCC ACC/DLPFC"
 author: "KM and GH"
@@
                                           parentId ="syn22276904",
-                                          folderName = 'Meta-analysis on diagnosis and sex')
+                                          folderName = 'Meta-analysis on sex')
 ```
@@
 ```{r input_data}
-mpp_acc <- readr::read_tsv(synGet("syn22276990")$path) %>% 
+mpp_acc <- readr::read_tsv(synGet("syn22278494")$path) %>% 
   dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val)
-mpp_dlpfc <- readr::read_tsv(synGet("syn22277179")$path) %>% 
+mpp_dlpfc <- readr::read_tsv(synGet("syn22278597")$path) %>% 
   dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val)
-hbcc_dsACC <- readr::read_tsv(synGet("syn22280364")$path) %>% 
+hbcc_dsACC <- readr::read_tsv(synGet("syn22280959")$path) %>% 
   dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) 
-hbcc_dlpfc <- readr::read_tsv(synGet("syn22280045")$path) %>% 
+hbcc_dlpfc <- readr::read_tsv(synGet("syn22280588")$path) %>% 
   dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val)
@@
 
-# Control vs SCZ
+# Female vs Male
 ```{r meta_analysis.SZ}
@@
 
-dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Control_vs_SCZ")
-accList <- make_list(mpp_acc, hbcc_dsACC, string = "Control_vs_SCZ")
+dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Female_vs_Male")
+accList <- make_list(mpp_acc, hbcc_dsACC, string = "Female_vs_Male")
 
@@
 
-# Control vs Other/SCZ
-```{r meta_analysis.SCZOther}
-make_list <- function(mpp, hbcc, string) {
-  list <- list(MPP = mpp %>%
-                 unique() %>% 
-                 dplyr::filter(Comparison == string) %>% 
-                 as.data.frame,
-               HBCC = hbcc %>% 
-                 unique() %>% 
-                 dplyr::filter(Comparison == string) %>% 
-                 as.data.frame)
-}
-
-dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Control_vs_OtherSCZ")
-accList <- make_list(mpp_acc, hbcc_dsACC, string = "Control_vs_OtherSCZ")
-
-rownames(dlpfcList[['MPP']]) = dlpfcList[['MPP']]$gene_id
-rownames(dlpfcList[['HBCC']]) = dlpfcList[['HBCC']]$gene_id
-
-res_dlpfc = run_meta_analysis( dlpfcList, "FE")
-res_dlpfc$Dataset = "Meta-analysis"
-
-rownames(accList[['MPP']]) = accList[['MPP']]$gene_id
-rownames(accList[['HBCC']]) = accList[['HBCC']]$gene_id
-
-res_acc = run_meta_analysis( accList, "FE")
-res_acc$Dataset = "Meta-analysis"
-```
-
-```{r plot.meta_analysisSCZOther}
-kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
-kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
-
-plotVolcano( res_dlpfc, 20) + ggtitle("Meta analysis - DLPFC")
-
-plotVolcano( res_acc, 20) + ggtitle("Meta analysis - ACC")
-```
-
-# Control vs Other
-```{r meta_analysis.CTOther}
-make_list <- function(mpp, hbcc, string) {
-  list <- list(MPP = mpp %>%
-                 unique() %>% 
-                 dplyr::filter(Comparison == string) %>% 
-                 as.data.frame,
-               HBCC = hbcc %>% 
-                 unique() %>% 
-                 dplyr::filter(Comparison == string) %>% 
-                 as.data.frame)
-}
-
-dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Control_vs_Other")
-accList <- make_list(mpp_acc, hbcc_dsACC, string = "Control_vs_Other")
-
-rownames(dlpfcList[['MPP']]) = dlpfcList[['MPP']]$gene_id
-rownames(dlpfcList[['HBCC']]) = dlpfcList[['HBCC']]$gene_id
-
-res_dlpfc = run_meta_analysis( dlpfcList, "FE")
-res_dlpfc$Dataset = "Meta-analysis"
-
-rownames(accList[['MPP']]) = accList[['MPP']]$gene_id
-rownames(accList[['HBCC']]) = accList[['HBCC']]$gene_id
-
-res_acc = run_meta_analysis( accList, "FE")
-res_acc$Dataset = "Meta-analysis"
-```
-
-```{r plot.meta_analysisCTOther}
-kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
-kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
-
-plotVolcano( res_dlpfc, 20) + ggtitle("Meta analysis - DLPFC")
-
-plotVolcano( res_acc, 20) + ggtitle("Meta analysis - ACC")
-```
-
-# Other vs SCZ
-```{r meta_analysis.OtherSCZ}
-make_list <- function(mpp, hbcc, string) {
-  list <- list(MPP = mpp %>%
-                 unique() %>% 
-                 dplyr::filter(Comparison == string) %>% 
-                 as.data.frame,
-               HBCC = hbcc %>% 
-                 unique() %>% 
-                 dplyr::filter(Comparison == string) %>% 
-                 as.data.frame)
-}
-
-dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Other_vs_SCZ")
-accList <- make_list(mpp_acc, hbcc_dsACC, string = "Other_vs_SCZ")
-
-rownames(dlpfcList[['MPP']]) = dlpfcList[['MPP']]$gene_id
-rownames(dlpfcList[['HBCC']]) = dlpfcList[['HBCC']]$gene_id
-
-res_dlpfc = run_meta_analysis( dlpfcList, "FE")
-res_dlpfc$Dataset = "Meta-analysis"
-
-rownames(accList[['MPP']]) = accList[['MPP']]$gene_id
-rownames(accList[['HBCC']]) = accList[['HBCC']]$gene_id
-
-res_acc = run_meta_analysis( accList, "FE")
-res_acc$Dataset = "Meta-analysis"
-```
-
-```{r plot.meta_analysisOtherSCZ}
-kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
-kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
-
-plotVolcano( res_dlpfc, 20) + ggtitle("Meta analysis - DLPFC")
-
-plotVolcano( res_acc, 20) + ggtitle("Meta analysis - ACC")
-```
-
 ## RCSD1

### 2020-07-31 individual-cohorts/MSSM-Penn-Pitt_DLPFC_DE.Rmd
Test Sex with Dx added to the model

@@
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Sex", "ageOfDeath", "RIN2", "RIN", "IntronicRate", "IntragenicRate")
+adjust.covars <- c("Sex", "Dx", "ageOfDeath", "RIN2", "RIN", "IntronicRate", "IntragenicRate")
 random_effect <- c("Institution")

### 2020-07-31 individual-cohorts/MSSM-Penn-Pitt_ACC_DE.Rmd
Test Sex with Dx added to the model

--- 2020-07-30 individual-cohorts/MSSM-Penn-Pitt_ACC_DE.Rmd
@@
 ---
-title: "Differential Expression - MSSM-Penn-Pitt - ACC - Dx"
+title: "Differential Expression - MSSM-Penn-Pitt - ACC - Sex"
 author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
@@
                                  parentId ="syn22012953",
-                                 folderName = "Differential Expression - Dx")
+                                 folderName = "Differential Expression - Sex")
 ```
@@
 ## Load required libraries
-library(CovariateAnalysis) # get the package from devtools::install_github('th1vairam/CovariateAnalysis@dev')
 library(data.table)
-library(tidyr)
-library(plyr)
 library(dplyr)
-library(stringr)
 library(R.utils)
-library(ggplot2)
-library(limma)
 library(psych)
 library(lme4)
-library(biomaRt)
 library(variancePartition)
-library(edgeR)
 library(synapser)
 library(knitr)
-library(githubr) # get the package from devtools::install_github('brian-bot/githubr')
+library(githubr)
+library(CovariateAnalysis)
+library(edgeR)
+library(WGCNA)
+library(biomaRt)
 
@@
 activityName = 'Differential Expression';
-activityDescription = 'Differential Expression - Dx';
+activityDescription = 'Differential Expression - Sex';
 
@@
 
+COVARIATES <- dplyr::rename(COVARIATES, Sex = Reported_Gender)
+
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Dx", "ageOfDeath", "RIN2", "RIN", "rRNARate", "IntronicRate", "IntragenicRate", "IntergenicRate")
-random_effect <- c("Institution", "Reported_Gender")
+adjust.covars <- c("Sex", "Dx", "ageOfDeath", "RIN2", "RIN", "rRNARate", "IntronicRate", "IntragenicRate", "IntergenicRate")
+random_effect <- c("Institution")
 ```
 
-### Normalisation - Dx
-1. Dx is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis)
+### Normalisation - Sex
+
+1. Sex is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis)
 
@@
 
-# Residuals after normalisation
 RESIDUAL.GENE_EXPRESSION = residuals(ADJUSTED.FIT, CQN.GENE_EXPRESSION)
@@
 ### Residual calculation
-Calculate weighted residuals and add back "Dx" to the residuals
+Calculate weighted residuals and add back "Reported_Gender" to the residuals
 ```{r varsToAddBack}
 # Add variable of interest back to the residuals
-varsToAddIn = grep("Dx", colnames(ADJUSTED.FIT$design), value = T)
+varsToAddIn = grep("Sex", colnames(ADJUSTED.FIT$design), value = T)
 RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION +
@@
 # Eucledian tree based analysis
-COVARIATES.tmp = data.matrix(COVARIATES[,c('Dx','Institution')])
+COVARIATES.tmp = data.matrix(COVARIATES[,c('Sex','Institution')])
 
@@
                            main = "Sample dendrogram",
-                           groupLabels = colnames(COVARIATES.tmp[,c('Dx','Institution')]))
+                           groupLabels = colnames(COVARIATES.tmp[,c('Sex','Institution')]))
 ```
@@
 
-### Differential expression analysis (SCZ - Control - Other)
+### Differential expression analysis (Female - Male)
 Genes that are differentially expressed at an FDR <= 0.05 are:
 ```{r set-formula}
+adjust.covars <- adjust.covars[!(adjust.covars %in% "RIN2")]
 formula <- glue::glue("~ 0 + ", glue::glue_collapse(adjust.covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))    
 ```
+
 `r formula`
@@
 ```{r diffExp, fig.height=10, fig.width=15}
-L1 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
+L = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                    formula = formula, 
                                    data = COVARIATES, 
-                                   coefficient = c("DxOther", "DxControl"))
-L2 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
-                                   formula = formula, 
-                                   data = COVARIATES, 
-                                   coefficient = c("DxSCZ", "DxControl"))
-L3 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
-                                   formula = formula, 
-                                   data = COVARIATES, 
-                                   coefficient = c("DxSCZ", "DxOther"))
+                                   coefficient = c("SexFemale", "SexMale"))
 
-# Manually set comparison of SCZ-Other to Control
-L = cbind(L1, L2, L3, data.frame(L4 = c(1, -0.5, -0.5, 0, 0, 0, 0)))
-
-contrast <- c(glue::glue_collapse(names(L1[L1 != 0]), "-"), 
-              glue::glue_collapse(names(L2[L2 != 0]), "-"),
-              glue::glue_collapse(names(L3[L3 != 0]), "-"),
-              glue::glue("DxControl-DxOtherDxSCZ")
-)
+contrast <- c(glue::glue_collapse(names(L[L != 0]), "-"))
 
@@
 }, FIT.CONTR)
+
 names(DE) <- contrast
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_ACC_Dx_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
-nEXP_OBJ = File('CMC_ACC_Dx_netResidualExpression.tsv', 
-                name = 'Dx Normalised, covariates removed residual expression (for network analysis)', 
+  write.table(file = 'CMC_ACC_Sex_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
+nEXP_OBJ = File('CMC_ACC_Sex_netResidualExpression.tsv', 
+                name = 'Sex Normalised, covariates removed residual expression (for network analysis)', 
                 parentId = CODE$properties$id)
@@
 # Store differential expression results
-write.table(DE, file = 'CMC_ACC_Dx_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
-DEXP_OBJ = File('CMC_ACC_Dx_DiffExpression.tsv', 
-                name = 'Differential Expression Results (Dx)', 
+write.table(DE, file = 'CMC_ACC_Sex_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
+DEXP_OBJ = File('CMC_ACC_Sex_DiffExpression.tsv', 
+                name = 'Differential Expression Results (Sex)', 
                 parentId = CODE$properties$id)
@@
 # Store adjusted fit linear mixed model object as RDS - Dx
-saveRDS(ADJUSTED.FIT, file = 'CMC_ACC_Dx_dream-output.rds')
-ADJ_OBJ = File('CMC_ACC_Dx_dream-output.rds',
-               name = 'Dream Output object - Dx', 
+saveRDS(ADJUSTED.FIT, file = 'CMC_ACC_Sex_dream-output.rds')
+ADJ_OBJ = File('CMC_ACC_Sex_dream-output.rds',
+               name = 'Dream Output object - Sex', 
                parentId = CODE$properties$id)

### 2020-08-01 individual-cohorts/HBCC_DLPFC_DE.Rmd
Test Sex with Dx added to the model

@@
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Sex", "ageOfDeath", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate")
+adjust.covars <- c("Sex", "Dx", "ageOfDeath", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate")
 ```

### 2020-08-01 individual-cohorts/HBCC_dsACC_DE.Rmd
Test Sex with Dx added to the model

@@
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Sex", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate", "ageOfDeath")
+adjust.covars <- c("Sex", "Dx", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate", "ageOfDeath")
 random_effect <- c("IsolationBatch")

### 2020-08-01 individual-cohorts/meta-analysis.Rmd
save meta-analysis output to Synapse

@@
   dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val)
+Ids <- c("syn22278494", "syn22278597", "syn22280959", "syn22280588")
 ```
@@
 
+```{r store}
+# Code
+CODE <- Folder(name = "Meta-analysis on sex", parentId = parentId)
+CODE <- synStore(CODE)
+
+# Store meta-analysis
+res_dlpfc %>%
+  write.table(file = 'CMC_dlpfc_Sex_meta_analysis.tsv', sep = '\t', row.names=F, quote=F)
+nEXP_OBJ = File('CMC_dlpfc_Sex_meta_analysis.tsv', 
+                name = 'Sex Normalised, meta-analysis (DLPFC)', 
+                parentId = CODE$properties$id)
+nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, activityName = activityName, 
+                    executed = thisFile, activityDescription = activityDescription)
+res_acc %>%
+  write.table(file = 'CMC_acc_Sex_meta_analysis.tsv', sep = '\t', row.names=F, quote=F)
+nEXP_OBJ = File('CMC_acc_Sex_meta_analysis.tsv', 
+                name = 'Sex Normalised, meta-analysis (ACC)', 
+                parentId = CODE$properties$id)
+nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, activityName = activityName, 
+                    executed = thisFile, activityDescription = activityDescription)
+```
+
 ### Source

### 2020-08-01 individual-cohorts/meta-analysis.Rmd
preserve rownames

@@
 # Code
-CODE <- Folder(name = "Meta-analysis on sex", parentId = parentId)
+CODE <- Folder(name = "Meta-analysis on sex", parentId = "syn22276904") # Make sure these match the folder name and parent Id in the first chunk.
 CODE <- synStore(CODE)
@@
 res_dlpfc %>%
+  tibble::rownames_to_column(var = "gene_id") %>% 
   write.table(file = 'CMC_dlpfc_Sex_meta_analysis.tsv', sep = '\t', row.names=F, quote=F)
@@
 res_acc %>%
+  tibble::rownames_to_column(var = "gene_id") %>% 
   write.table(file = 'CMC_acc_Sex_meta_analysis.tsv', sep = '\t', row.names=F, quote=F)

### 2020-08-01 individual-cohorts/meta-analysis.Rmd
remove ref to activity

@@
                 parentId = CODE$properties$id)
-nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, activityName = activityName, 
-                    executed = thisFile, activityDescription = activityDescription)
+nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, 
+                    executed = thisFile)
 res_acc %>%
@@
                 parentId = CODE$properties$id)
-nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, activityName = activityName, 
-                    executed = thisFile, activityDescription = activityDescription)
+nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, 
+                    executed = thisFile)
 ```

### 2020-08-12 individual-cohorts/common_functions/common_functions.R
Add common functions

@@ -1,68 +1,4 @@
----
-title: "Meta-analysis on sex in MSSM-Penn-Pitt and HBCC ACC/DLPFC"
-author: "KM and GH"
-date: "`r date()`"
-output: html_document
----
 
-```{r knit2synapse, eval=FALSE}
-library(synapser)
-library(knit2synapse)
-synLogin()
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/meta-analysis.Rmd",
-                                          parentId ="syn22276904",
-                                          folderName = 'Meta-analysis on sex')
-```
-
-```{r setup, include=FALSE}
-library(synapser)
-
-synLogin()
-
-library(ggplot2)
-library(dplyr)
-library(ComplexHeatmap)
-library(readxl)
-library(gtools)
-library(corrplot)
-library(pvclust)
-library(metafor)
-library(cowplot)
-library(org.Hs.eg.db)
-library(ggrepel)
-library(metafor)
-library(patchwork)
-
-knitr::opts_chunk$set(
-  echo = FALSE,
-  warning=FALSE,
-  message=FALSE,
-  error = FALSE,
-  tidy = FALSE,
-  cache = TRUE)
-```
-
-```{r synapse.parameters, include=FALSE, cache=FALSE, eval=TRUE}
-thisFileName <- 'meta-analysis.Rmd'
-
-# Github link
-thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
-thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/',thisFileName))
-```
-
-```{r input_data}
-mpp_acc <- readr::read_tsv(synGet("syn22278494")$path) %>% 
-  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val)
-mpp_dlpfc <- readr::read_tsv(synGet("syn22278597")$path) %>% 
-  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val)
-hbcc_dsACC <- readr::read_tsv(synGet("syn22280959")$path) %>% 
-  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) 
-hbcc_dlpfc <- readr::read_tsv(synGet("syn22280588")$path) %>% 
[…]

### 2020-08-12 individual-cohorts/meta_analysis/meta-analysis_diagnosis.Rmd
Add separate meta analysis folder

---
title: "Meta-analysis on diagnosis in MSSM-Penn-Pitt and HBCC ACC/DLPFC"
author: "KM and GH"
date: "`r date()`"
output: html_document
---

```{r knit2synapse, eval=FALSE}
# rmarkdown::render("meta-analysis_diagnosis.Rmd")

library(synapser)
library(knit2synapse)
synLogin()
knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/meta-analysis.Rmd",
                                          parentId ="syn22276904",
                                          folderName = 'Meta-analysis on diagnosis')
```

```{r setup, include=FALSE}
library(synapser)

synLogin()

library(ggplot2)
library(dplyr)
library(ComplexHeatmap)
library(readxl)
library(gtools)
library(corrplot)
library(pvclust)
library(metafor)
library(cowplot)
library(githubr)
library(org.Hs.eg.db)
library(ggrepel)
library(metafor)
library(patchwork)

source("../../common_functions/common_functions.R")

knitr::opts_chunk$set(
  echo = FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE, cache=FALSE, eval=TRUE}
thisFileName <- 'meta-analysis.Rmd'

# Github link
thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/',thisFileName))
```

```{r input_data}
mpp_acc <- readr::read_tsv(synGet("syn22079117")$path) %>% 
  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val)
mpp_dlpfc <- readr::read_tsv(synGet("syn22079472")$path) %>% 
  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val)
hbcc_dsACC <- readr::read_tsv(synGet("syn22086829")$path) %>% 
  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) 
hbcc_dlpfc <- readr::read_tsv(synGet("syn22084205")$path) %>% 
  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val)
```

# Control vs SCZ
```{r meta_analysis.SZ}
make_list <- function(mpp, hbcc, string) {
  list <- list(MPP = mpp %>%
                 unique() %>% 
                 dplyr::filter(Comparison == string) %>% 
                 as.data.frame,
               HBCC = hbcc %>% 
                 unique() %>% 
                 dplyr::filter(Comparison == string) %>% 
                 as.data.frame)
}

dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Control_vs_SCZ")
accList <- make_list(mpp_acc, hbcc_dsACC, string = "Control_vs_SCZ")

rownames(dlpfcList[['MPP']]) = dlpfcList[['MPP']]$gene_id
rownames(dlpfcList[['HBCC']]) = dlpfcList[['HBCC']]$gene_id

res_dlpfc = run_meta_analysis( dlpfcList, "FE")
res_dlpfc$Dataset = "Meta-analysis"

rownames(accList[['MPP']]) = accList[['MPP']]$gene_id
rownames(accList[['HBCC']]) = accList[['HBCC']]$gene_id

res_acc = run_meta_analysis( accList, "FE")
res_acc$Dataset = "Meta-analysis"
```

```{r plot.meta_analysisSCZ}
kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))

plotVolcano( res_dlpfc, 20) + ggtitle("Meta analysis - DLPFC")

plotVolcano( res_acc, 20) + ggtitle("Meta analysis - ACC")
```

# Control vs Other/SCZ
```{r meta_analysis.SCZOther}
make_list <- function(mpp, hbcc, string) {
  list <- list(MPP = mpp %>%
                 unique() %>% 
                 dplyr::filter(Comparison == string) %>% 
                 as.data.frame,
               HBCC = hbcc %>% 
                 unique() %>% 
                 dplyr::filter(Comparison == string) %>% 
                 as.data.frame)
}

dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Control_vs_OtherSCZ")
accList <- make_list(mpp_acc, hbcc_dsACC, string = "Control_vs_OtherSCZ")

rownames(dlpfcList[['MPP']]) = dlpfcList[['MPP']]$gene_id
rownames(dlpfcList[['HBCC']]) = dlpfcList[['HBCC']]$gene_id

res_dlpfc = run_meta_analysis( dlpfcList, "FE")
res_dlpfc$Dataset = "Meta-analysis"

rownames(accList[['MPP']]) = accList[['MPP']]$gene_id
rownames(accList[['HBCC']]) = accList[['HBCC']]$gene_id

res_acc = run_meta_analysis( accList, "FE")
res_acc$Dataset = "Meta-analysis"
```

```{r plot.meta_analysisSCZOther}
kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))

plotVolcano( res_dlpfc, 20) + ggtitle("Meta analysis - DLPFC")

plotVolcano( res_acc, 20) + ggtitle("Meta analysis - ACC")
```

# Control vs Other
```{r meta_analysis.CTOther}
make_list <- function(mpp, hbcc, string) {
  list <- list(MPP = mpp %>%
                 unique() %>% 
                 dplyr::filter(Comparison == string) %>% 
                 as.data.frame,
               HBCC = hbcc %>% 
                 unique() %>% 
                 dplyr::filter(Comparison == string) %>% 
                 as.data.frame)
}

dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Control_vs_Other")
accList <- make_list(mpp_acc, hbcc_dsACC, string = "Control_vs_Other")

rownames(dlpfcList[['MPP']]) = dlpfcList[['MPP']]$gene_id
rownames(dlpfcList[['HBCC']]) = dlpfcList[['HBCC']]$gene_id

res_dlpfc = run_meta_analysis( dlpfcList, "FE")
res_dlpfc$Dataset = "Meta-analysis"

rownames(accList[['MPP']]) = accList[['MPP']]$gene_id
rownames(accList[['HBCC']]) = accList[['HBCC']]$gene_id

res_acc = run_meta_analysis( accList, "FE")
res_acc$Dataset = "Meta-analysis"
```

```{r plot.meta_analysisCTOther}
kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))

plotVolcano( res_dlpfc, 20) + ggtitle("Meta analysis - DLPFC")

plotVolcano( res_acc, 20) + ggtitle("Meta analysis - ACC")
```

# Other vs SCZ
```{r meta_analysis.OtherSCZ}
make_list <- function(mpp, hbcc, string) {
  list <- list(MPP = mpp %>%
                 unique() %>% 
                 dplyr::filter(Comparison == string) %>% 
                 as.data.frame,
               HBCC = hbcc %>% 
                 unique() %>% 
                 dplyr::filter(Comparison == string) %>% 
                 as.data.frame)
}

dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Other_vs_SCZ")
accList <- make_list(mpp_acc, hbcc_dsACC, string = "Other_vs_SCZ")

rownames(dlpfcList[['MPP']]) = dlpfcList[['MPP']]$gene_id
rownames(dlpfcList[['HBCC']]) = dlpfcList[['HBCC']]$gene_id

res_dlpfc = run_meta_analysis( dlpfcList, "FE")
res_dlpfc$Dataset = "Meta-analysis"

rownames(accList[['MPP']]) = accList[['MPP']]$gene_id
rownames(accList[['HBCC']]) = accList[['HBCC']]$gene_id

res_acc = run_meta_analysis( accList, "FE")
res_acc$Dataset = "Meta-analysis"
```

```{r plot.meta_analysisOtherSCZ}
kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))

plotVolcano( res_dlpfc, 20) + ggtitle("Meta analysis - DLPFC")

plotVolcano( res_acc, 20) + ggtitle("Meta analysis - ACC")
```

## RCSD1

Forest plot of example top gene.

```{r specific-gene}
# forrest plot of meta-analysis
ensID = 'ENSG00000198771.11'
kable(res_dlpfc[ensID,])
lapply(dlpfcList, function(x) x[ensID,])
ensID = rownames(res_dlpfc)[which.min(res_dlpfc$adj.P.Val)][1]
plot_forrest(dlpfcList, ensID)
```

### Source

### 2020-08-12 individual-cohorts/meta_analysis/meta-analysis_sex.Rmd
Add separate meta analysis folder

--- 2020-08-01 individual-cohorts/meta-analysis.Rmd
@@
 ```{r knit2synapse, eval=FALSE}
+# rmarkdown::render("meta-analysis_sex.Rmd")
+
 library(synapser)
@@
 library(ggplot2)
+library(githubr)
 library(dplyr)
@@
 library(ggrepel)
+library(knitr)
 library(metafor)

### 2020-08-17 common_functions/common_functions.R
add plot_sex_compare

@@ -1,4 +1,113 @@
 
+library(qvalue)
+# library(EnsDb.Hsapiens.v86)
+library(AnnotationHub)
+library(ensembldb)
+library(MASS)
+library(ggplot2)
+library(viridis)
+library(data.table)
+
+# For ENSEMBL id ENSG00000279457.4, return ENSG00000279457
+trim_ensembl_ids = function(x){
+  gsub("(.*)\\.(.*)", "\\1", x) 
+}
+
+# Add column Symbol using column gene_id storing ENSEMBL id
+getGeneSymbol = function( df, column="row.names"){
+
+  # load load since data base can fail with rmarkdown run multiple times
+  # detach("package:EnsDb.Hsapiens.v86", unload=TRUE)
+  # library(EnsDb.Hsapiens.v86)
+
+  # Load ENSEMBL v96 database
+  # see https://www.bioconductor.org/packages/release/bioc/vignettes/ensembldb/inst/doc/ensembldb.html#10_getting_or_building_ensdb_databasespackages
+  ah <- AnnotationHub()
+  ensdb = ah[["AH69187"]] # ENSEMBL v96
+
+  if( column == "row.names"){    
+    geneID = rownames(df)
+  }else{
+    geneID = df[[column]]
+  }
+
+  ensIDs = trim_ensembl_ids( geneID )
+
+  geneSymbol = mapIds(ensdb, keys=ensIDs, keytype="GENEID", column="GENENAME")
+  df$Symbol = geneSymbol
+
+  chroms = ensembldb::select(ensdb, keys=ensIDs, keytype="GENEID", column="SEQNAME")
+
+  df2 = merge( data.frame(GENEID = ensIDs, stringsAsFactors=FALSE), chroms, by='GENEID')
+
+  idx = match(ensIDs, chroms$GENEID)
+  df$Chrom = c()
+  df$Chrom = chroms$SEQNAME[idx]
+
+  df
+}
+
+get_density <- function(x, y, n = 250) {
+  dens <- MASS::kde2d(x = x, y = y, n = n)
+  ix <- findInterval(x, dens$x)
+  iy <- findInterval(y, dens$y)
+  ii <- cbind(ix, iy)
+  return(dens$z[ii])
+}
+
+plot_concordance = function( resList, showGenes = NULL, idx=c(1,2), size=15){
+
+  df = merge( resList[[idx[1]]], resList[[idx[2]]], by="row.names")
+
+  lab = "log2 fold change"
[…]

### 2020-08-17 individual-cohorts/meta_analysis/meta-analysis_sex.Rmd
Sex analysis

--- 2020-08-01 individual-cohorts/meta-analysis.Rmd
@@
 date: "`r date()`"
-output: html_document
+output:
+  html_document:
+    toc: true
+    toc_float: true
 ---
 
-```{r knit2synapse, eval=FALSE}
+<!---
+
+# individual-cohorts/meta_analysis/
+
+# rmarkdown::render("meta-analysis_sex.Rmd")
+--->
+
+```{r knit2synapse, eval=TRUE}
 library(synapser)
@@
 synLogin()
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/meta-analysis.Rmd",
-                                          parentId ="syn22276904",
-                                          folderName = 'Meta-analysis on sex')
+knit2synapse::createAndKnitToFolderEntity(file = "meta-analysis_sex.Rmd",
+                                          parentId ="syn22281021",
+                                          folderName = 'Meta-analysis on sex',
+                                          overwrite=FALSE)
 ```
@@
 library(cowplot)
+library(kableExtra)
+library(githubr)
 library(org.Hs.eg.db)
@@
 library(metafor)
+library(knitr)
 library(patchwork)
+
+source("../../common_functions/common_functions.R")
 
@@
 
-```{r synapse.parameters, include=FALSE, cache=FALSE, eval=TRUE}
-thisFileName <- 'meta-analysis.Rmd'
-
+```{r synapse.parameters, include=FALSE, cache=TRUE, eval=TRUE}
+thisFileName <- 'meta-analysis_sex.Rmd'
+# thisFile = thisFileName
 # Github link
 thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
-thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/',thisFileName))
+thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/meta_analysis/',thisFileName))
 ```
@@
 mpp_acc <- readr::read_tsv(synGet("syn22278494")$path) %>% 
-  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val)
+  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% getGeneSymbol( 'gene_id' )
 mpp_dlpfc <- readr::read_tsv(synGet("syn22278597")$path) %>% 
-  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val)
+  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% getGeneSymbol( 'gene_id' )
 hbcc_dsACC <- readr::read_tsv(synGet("syn22280959")$path) %>% 
-  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) 
+  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% getGeneSymbol( 'gene_id' )
 hbcc_dlpfc <- readr::read_tsv(synGet("syn22280588")$path) %>% 
-  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val)
+  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% getGeneSymbol( 'gene_id' )
 Ids <- c("syn22278494", "syn22278597", "syn22280959", "syn22280588")
@@
 
-```{r multiplot}
-# Multiple plot function
-#
-# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
-# - cols:   Number of columns in layout
-# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
-#
-# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
-# then plot 1 will go in the upper left, 2 will go in the upper right, and
-# 3 will go all the way across the bottom.
-#
-multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
-  library(grid)
-
-  # Make a list from the ... arguments and plotlist
-  plots <- c(list(...), plotlist)
-
-  numPlots = length(plots)
-
-  # If layout is NULL, then use 'cols' to determine layout
-  if (is.null(layout)) {
-    # Make the panel
-    # ncol: Number of columns of plots
-    # nrow: Number of rows needed, calculated from # of cols
-    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
-                    ncol = cols, nrow = ceiling(numPlots/cols))
-  }
-
- if (numPlots==1) {
-    print(plots[[1]])
-
-  } else {
-    # Set up the page
-    grid.newpage()
-    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
-
-    # Make each plot, in the correct location
-    for (i in 1:numPlots) {
-      # Get the i,j matrix positions of the regions that contain this subplot
-      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
-
-      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
-                                      layout.pos.col = matchidx$col))
-    }
-  }
-}
-```
-
-```{r custom.functions}
-# perform meta-analysis given fits from limma
-run_meta_analysis = function(tabList, method="FE") {
-
-  # extract logFC and se for each model fit
-  resList = lapply( tabList, function(tab){
-    tab$se = tab$logFC /  tab$t
-    tab
-  })
-
-  # extract features names
-  featureNames = lapply(resList, function(res){
-    rownames(res)
-    })
-  featureNames = unique(unlist(featureNames))
-
-  # check that each fit has all features
-  # testIdentical = lapply(resList, function(res){
-  #   identical(sort(rownames(res)), sort(featureNames))
-  #   })
-  # if(  any(!unlist(testIdentical)) ){
-  #   stop("each entry in fitList must have all features with the same names")
-  # }
-
-  # apply meta analysis for each feature
-  resRMA = lapply(featureNames, function(feature){
-    # extract summary statistics for this feature from all studies
-    res = lapply(resList, function(res){
-      res[feature,]
-      })
-    df = do.call("rbind", res)
-    suppressWarnings(rma(yi=logFC, sei=se, data = df, method=method))
-    })
-  names(resRMA) = featureNames
-
-  # extract results
-  # Currently exatract simple results
-  # other results may be relevant if meta-analysing more than two studies
-  resTable = lapply(resRMA, function(x){
-    data.frame( logFC     = x$beta,
-                se        = x$se,
-                P.Value   = x$pval,
-                adj.P.Val = p.adjust(x$pval, "fdr"),
-                Q         = x$QE,
-                N         = x$k,
-                I2        = x$I2)
-    })
-  do.call("rbind", resTable)
-}
-
-plotVolcano = function(df, nshow=10){
-  # indicate significant genes
-  df$isSignif = c("no","yes")[(df$adj.P.Val < 0.05)+1]
-
-  # show gene names
-  df$EnsID = gsub('\\..*$', '', rownames(df))
-
-  df_hugo = select(org.Hs.eg.db, df$EnsID, "SYMBOL", "ENSEMBL")
-  df = merge(df, df_hugo, by.x='EnsID', by.y="ENSEMBL")
-
-  # df = df[!duplicated(df$gene_id),]
-
-  # top significant genes
-  df2 = df[df$adj.P.Val <= sort(df$adj.P.Val )[nshow], ]
-
-  ggplot(df, aes( logFC, -log10(P.Value), color=isSignif)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, legend.position="none", plot.title = element_text(hjust = 0.5)) + xlab(bquote(log[2]~fold~change)) + ylab(bquote(-log[10]~P)) + scale_color_manual(values=c("grey", "red")) + geom_text_repel(data=df2, aes(logFC, -log10(P.Value), label=SYMBOL), segment.size=.5,  segment.color="black", color="black", force=1, nudge_x=.005, nudge_y=.5)
-}
-
-plot_forrest = function(resList, ensID) { 
-  # create a single data.frame with logFC and se
-  df_forrest = lapply(resList, function(x) if(ensID %in% x[ensID,]) {
-    x[ensID,]
-  } 
-  )
-  df_forrest[sapply(df_forrest, is.null)] <- NULL
-  df_forrest = do.call(rbind, df_forrest)
-  df_forrest$Dataset = rownames(df_forrest)
-  df_forrest$se = df_forrest$logFC /  df_forrest$t
-
-  # bind only shared columns
-  df = gtools::smartbind(df_forrest,  res_dlpfc[ensID,])
-  df$Dataset = factor(df$Dataset, rev(df$Dataset))
-
-  # set range to include zero
-  rng = range(c(with(df, logFC-se), with(df, logFC+se)))
-  if( rng[1] > 0 ){
-    rng[1] = 0
-  }
-  if( rng[2] < 0 ){
-    rng[2] = 0
-  }
-
-  # forrest plot
-  fig1 = ggplot(df, aes(Dataset, logFC)) + geom_point(color="navy") + geom_errorbar(aes(ymin = logFC - se, ymax=logFC + se), width=0.1) + theme_bw(12) + theme(aspect.ratio= .2, plot.title = element_text(hjust = 0.5)) + coord_flip() + ylim(rng) + xlab(bquote(log[2]~fold~change)) + ggtitle('RCSD1')
-
-  # Adjusted p-value
-  ymax = max(-log10(df$adj.P.Val))*1.05
-  fig2 = ggplot(df, aes(Dataset, -log10(adj.P.Val))) + geom_bar(stat="identity", fill="navy") + theme_bw(12) + theme( aspect.ratio= .5, plot.margin = unit(c(0,0,0,0), "cm"),
-    axis.title.y=element_blank(),
-          axis.text.y=element_blank(),
-          axis.ticks.y=element_blank()) + coord_flip() + ylab(bquote(-log[10]~adjusted~P)) + scale_y_continuous(expand=c(0,0), limits=c(0, ymax))
-
-  # plot_grid(fig1, fig2, align='vh')
-  fig1 + fig2
-}
-```
-
 # Female vs Male
-```{r meta_analysis.SZ}
+```{r meta_analysis.Sex}
 make_list <- function(mpp, hbcc, string) {
@@
 
-res_dlpfc = run_meta_analysis( dlpfcList, "FE")
+res_dlpfc = run_meta_analysis( dlpfcList, "FE") %>% getGeneSymbol
 res_dlpfc$Dataset = "Meta-analysis"
@@
 
-res_acc = run_meta_analysis( accList, "FE")
+res_acc = run_meta_analysis( accList, "FE") %>% getGeneSymbol
 res_acc$Dataset = "Meta-analysis"
@@
 
-```{r plot.meta_analysisSCZ}
-kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
-kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
-
-plotVolcano( res_dlpfc, 20) + ggtitle("Meta analysis - DLPFC")
-
-plotVolcano( res_acc, 20) + ggtitle("Meta analysis - ACC")
+## DLPFC
+```{r plot.meta_analysis.dlpfc.Sex}
+kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05")) %>%
+  kable_styling(full_width = FALSE)
 ```
 
-## RCSD1
+```{r plot.concordance.dlpfc.Sex, fig.height=20}
 
-Forest plot of example top gene.
+# get genes to highlight
+idx = order(res_dlpfc$adj.P.Val)
+showGenes = res_dlpfc$Symbol[idx[1:20]]
+showGenes2 = res_dlpfc$Symbol[!(res_dlpfc$Chrom %in% c("X", "Y"))][idx[1:30]]
 
-```{r specific-gene}
-# forrest plot of meta-analysis
-ensID = 'ENSG00000198771.11'
-kable(res_dlpfc[ensID,])
-lapply(dlpfcList, function(x) x[ensID,])
-ensID = rownames(res_dlpfc)[which.min(res_dlpfc$adj.P.Val)][1]
-plot_forrest(dlpfcList, ensID)
+dlpfcList_autosome = lapply(dlpfcList, function(x){
+  x[!(x$Chrom %in% c("X", "Y")),]
+  })
+
+resConcord_dlpfc = plot_concordance( dlpfcList_autosome, showGenes )
+
+plot_grid(resConcord_dlpfc$fig_logFC, 
+          resConcord_dlpfc$fig_tstat, 
+          plotVolcano( res_dlpfc, showGenes),   
+          plotVolcano( res_dlpfc[!(res_dlpfc$Chrom %in% c("X", "Y")),], showGenes2),
+          ncol=1)
 ```
 
-```{r store}
+```{r plot.concordance.table.dlpfc.Sex, fig.width=6}
+kable( resConcord_dlpfc$tab_pi ) %>%
+  kable_styling(full_width = FALSE)
+
+resConcord_dlpfc$tab_corr$P = format(resConcord_dlpfc$tab_corr$P, scientific=TRUE, digits=3)
+kable( resConcord_dlpfc$tab_corr )  %>%
+  kable_styling(full_width = FALSE)
+```
+
+```{r plot.forrest.dlpfc.Sex, fig.width=5, fig.height=20}  
+figList = lapply(showGenes2[1:10], function(geneSymbol){
+  plot_forrest( c(dlpfcList, list("Meta-analysis" = res_dlpfc)), geneSymbol)
+})
+
+plot_grid(plotlist=figList, ncol=1)
+```
+
+### MPP
+```{r DE.results.dlpfc.mpp, fig.width=10, fig.height=50}
+figList = plot_sex_compare(mpp_dlpfc)
+plot_grid(plotlist=figList, ncol=2) 
+```
+
+### HBCC
+```{r DE.results.dlpfc.hbcc, fig.width=10, fig.height=50}
+figList = plot_sex_compare(hbcc_dlpfc)
+plot_grid(plotlist=figList, ncol=2) 
+```
+
+### Combine cohorts
+```{r DE.results.dlpfc.combine, fig.width=10, fig.height=50}
+res = merge(res_dlpfc, mpp_dlpfc[,c("gene_id", "AveExpr")], by.x="row.names", by.y="gene_id")
+
+figList = plot_sex_compare(res)
+plot_grid(plotlist=figList, ncol=2)
+
+```
+
+## ACC
+```{r plot.meta_analysis.acc.Sex}
+kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05")) %>%
+  kable_styling(full_width = FALSE)
+```
+
+```{r plot.concordance.acc.Sex, fig.height=20}
+
+# get genes to highlight
+idx = order(res_acc$adj.P.Val)
+showGenes = res_acc$Symbol[idx[1:20]]
+showGenes2 = res_acc$Symbol[!(res_acc$Chrom %in% c("X", "Y"))][idx[1:30]]
+
+accList_autosome = lapply(accList, function(x){
+  x[!(x$Chrom %in% c("X", "Y")),]
+  })
+
+resConcord_acc = plot_concordance( accList_autosome, showGenes )
+
+plot_grid(resConcord_acc$fig_logFC, 
+          resConcord_acc$fig_tstat, 
+          plotVolcano( res_acc, showGenes),   
+          plotVolcano( res_acc[!(res_acc$Chrom %in% c("X", "Y")),], showGenes2),
+          ncol=1)
+```
+
+```{r plot.concordance.table.acc.Sex, fig.width=6}
+kable( resConcord_acc$tab_pi ) %>%
+  kable_styling(full_width = FALSE)
+
+resConcord_acc$tab_corr$P = format(resConcord_acc$tab_corr$P, scientific=TRUE, digits=3)
+kable( resConcord_acc$tab_corr )  %>%
+  kable_styling(full_width = FALSE)
+```
+
+```{r plot.forrest.acc.Sex, fig.width=5, fig.height=20}  
+figList = lapply(showGenes2[1:10], function(geneSymbol){
+  plot_forrest( c(accList, list("Meta-analysis" = res_acc)), geneSymbol)
+})
+
+plot_grid(plotlist=figList, ncol=1)
+```
+
+### MPP
+```{r DE.results.acc.mpp, fig.width=10, fig.height=50}
+figList = plot_sex_compare(mpp_acc)
+plot_grid(plotlist=figList, ncol=2) 
+```
+
+### HBCC
+```{r DE.results.acc.hbcc, fig.width=10, fig.height=50}
+figList = plot_sex_compare(hbcc_dsACC)
+plot_grid(plotlist=figList, ncol=2) 
+```
+
+### Combine cohorts
+```{r DE.results.acc.combine, fig.width=10, fig.height=50}
+res = merge(res_acc, mpp_acc[,c("gene_id", "AveExpr")], by.x="row.names", by.y="gene_id")
+
+figList = plot_sex_compare(res)
+plot_grid(plotlist=figList, ncol=2)
+
+```
+
+## Compare across brain regions
+```{r plot.concordance.combine, fig.height=15}
+
+resCombine = merge(res_dlpfc, res_acc, by="row.names")
+
+res_dlpfc$t = with(res_dlpfc, logFC / se)
+res_acc$t = with(res_acc, logFC / se)
+
+resConcord = plot_concordance( list(
+        DLPFC = res_dlpfc[!(res_dlpfc$Chrom %in% c("X", "Y")),], 
+        ACC = res_acc[!(res_acc$Chrom %in% c("X", "Y")),]), showGenes2 )
+
+plot_grid(resConcord$fig_logFC, 
+          resConcord$fig_tstat,     
+          nrow=2) 
+```
+
+```{r plot.concordance.table.combine, fig.width=6}
+kable( resConcord$tab_pi ) %>%
+  kable_styling(full_width = FALSE)
+
+resConcord$tab_corr$P = format(resConcord$tab_corr$P, scientific=TRUE, digits=3)
+kable( resConcord$tab_corr )  %>%
+  kable_styling(full_width = FALSE)
+```
+
+```{r store, eval=TRUE}
 # Code
-CODE <- Folder(name = "Meta-analysis on sex", parentId = "syn22276904") # Make sure these match the folder name and parent Id in the first chunk.
+CODE <- Folder(name = "Meta-analysis on sex", parentId = "syn22281021") # Make sure these match the folder name and parent Id in the first chunk.
 CODE <- synStore(CODE)
@@
 ### Source
-[markdown](`r thisFile`)
+#[markdown](`r thisFile`)

### 2020-08-18 individual-cohorts/HBCC_DLPFC_DE.Rmd
dx*sex interaction, revised model and contrasts, fix conflict

--- 2020-07-31 individual-cohorts/HBCC_DLPFC_DE.Rmd
@@
 
-# Test Dx-Sex
-COVARIATES <- COVARIATES %>%
-  tibble::rownames_to_column(var = "samples") %>% 
-  tidyr::unite("Dx_Sex",
-               dplyr::all_of(c("Dx", "Reported_Gender")),
-               remove = TRUE) %>% 
-  dplyr::mutate(`Dx_Sex` = as.factor(`Dx_Sex`)) %>% 
-  tibble::column_to_rownames(var = "samples")
-
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Dx_Sex", "ageOfDeath", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate")
+interaction <- c("Dx", "Reported_Gender")
+adjust.covars <- c("ageOfDeath", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate")
 ```
 
-### Normalisation - Dx_Sex
-1. Dx_Sex is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis)
+### Normalisation - Dx*Sex
+1. Dx*Sex is chosen as the primary variable of interest
 
@@
 writeLines(glue::glue('Using following covariates in the model:',
+                      glue::glue_collapse(interaction, sep = ", "),
+                      ', ',
                       glue::glue_collapse(adjust.covars, sep = ", ", last = " and "),
@@
 # Post adjusted formula
-formula <- glue::glue("~ ", glue::glue_collapse(adjust.covars, sep = " + "))
+formula <- glue::glue("~ ", glue::glue_collapse(interaction, sep = "*")," + ", glue::glue_collapse(adjust.covars, sep = " + "))
 
@@
 
-formula_for_variance_fractions <- glue::glue("~ ", glue::glue_collapse(formula_for_variance_fractions, sep = " + "))
+# Manually specify interaction
+interaction <- c(interaction, "Dx:Reported_Gender")
+interaction_variance <- glue::glue("(1|{interaction})")
+all <- c(interaction_variance, formula_for_variance_fractions)
+
+formula_for_variance_fractions <- glue::glue("~ ", glue::glue_collapse(all, sep = " + "))
 
@@
 ```
-
+`r formula`
 ### Residual calculation
-Calculate weighted residuals and add back "Dx_Sex" to the residuals
+Calculate weighted residuals and add back "Dx:Sex" to the residuals
 ```{r varsToAddBack}
 # Add variable of interest back to the residuals
-varsToAddIn = grep("Dx_Sex", colnames(ADJUSTED.FIT$design), value = T)
+varsToAddIn = grep("DxOther:Reported_GenderMale", "DxSCZ:Reported_GenderMale", colnames(ADJUSTED.FIT$design), value = T)
 RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION +
@@
 # Eucledian tree based analysis
-COVARIATES.tmp = data.matrix(COVARIATES[,c('Dx_Sex'), drop = FALSE])
+COVARIATES.tmp = data.matrix(COVARIATES[,c('Dx', 'Reported_Gender'), drop = FALSE])
 
@@
                            main = "Sample dendrogram",
-                           groupLabels = colnames(COVARIATES.tmp[,c('Dx_Sex'), drop = FALSE]))
+                           groupLabels = colnames(COVARIATES.tmp[,c('Dx','Reported_Gender'), drop = FALSE]))
 ```
@@
 ```{r set-formula}
-formula <- glue::glue("~ 0 + ", glue::glue_collapse(adjust.covars, sep = " + "))    
+formula <- glue::glue("~ 0 + Dx_Sex + ", glue::glue_collapse(adjust.covars, sep = " + "))
 ```
@@
 
+Dx_Sex is 6 levels, SCZ_Female, SCZ_Male, Other_Female, Other_Male, Control_Female, Control_Male
+
 ```{r diffExp, fig.height=10, fig.width=15}
+# Test Dx-Sex
+COVARIATES <- COVARIATES %>%
+  tibble::rownames_to_column(var = "samples") %>% 
+  tidyr::unite("Dx_Sex",
+               dplyr::all_of(c("Dx", "Reported_Gender")),
+               remove = TRUE) %>% 
+  dplyr::mutate(`Dx_Sex` = as.factor(`Dx_Sex`)) %>% 
+  tibble::column_to_rownames(var = "samples")
+
 L1 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
@@
                                    data = COVARIATES, 
-                                   coefficient = c("Dx_SexControl_Female", "Dx_SexControl_Male"))
-L2 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
-                                   formula = formula, 
-                                   data = COVARIATES, 
-                                   coefficient = c("Dx_SexSCZ_Female", "Dx_SexSCZ_Male"))
-L3 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
-                                   formula = formula, 
-                                   data = COVARIATES, 
-                                   coefficient = c("Dx_SexOther_Female", "Dx_SexOther_Male"))
+                                   coefficient = c("Dx_SexControl_Male", "Dx_SexControl_Female"))
+L2 = c(1, -1, 0, 0, -1, 1, 0, 0, 0, 0, 0)
+L3 = c(1, -1, -1, 1, 0, 0, 0, 0, 0, 0, 0)
 
 # Manually set comparison between SCZ+Other vs. Control 
-L = cbind(L1, L2, L3, 
-          data.frame(L4 = c(1,0,-0.5, 0, -0.5, 0, 0, 0, 0, 0, 0)),
-          data.frame(L5 = c(0, 1, 0, -0.5, 0, -0.5, 0, 0, 0, 0, 0)))
+L = cbind(L1, L2, L3)
 
 contrast <- c(glue::glue_collapse(names(L1[L1 != 0]), "-"), 
-              glue::glue_collapse(names(L2[L2 != 0]), "-"),
-              glue::glue_collapse(names(L3[L3 != 0]), "-"),
-              glue::glue("Dx_SexControl_Female-Dx_SexOther_FemaleDx_SexSCZ_Female"),
-              glue::glue("Dx_SexControl_Male-Dx_SexOther_MaleDx_SexSCZ_Male")
+              glue::glue("Dx_SexSCZ_MaleDx_SexControl_Male-Dx_SexSCZ_FemaleDx_SexControl_Female"),
+              glue::glue("Dx_SexOther_MaleDx_SexControl_Male-Dx_SexOther_FemaleDx_SexControl_Female")
 )
@@
 counts = 'syn21886235'
-ALL_USED_IDs = counts
+ALL_USED_IDs = c(counts, ALL_USED_IDs)
 counts = downloadFile(counts) %>% data.frame()
@@
 # Get differential expression
-DE = lapply(glue::glue("L{1:5}"), function(i, FIT){
+DE = lapply(glue::glue("L{1:3}"), function(i, FIT){
   topTable(FIT, coef=i, number = Inf) %>%
@@
 ```
+
+### Source Code
+`r thisFile`
 

### 2020-08-19 individual-cohorts/HBCC_dsACC_DE.Rmd
dx*sex interaction, revised model and contrasts HBCC dsACC

--- 2020-07-31 individual-cohorts/HBCC_dsACC_DE.Rmd
@@
 
-# Test Dx-Sex
-COVARIATES <- COVARIATES %>%
-  tibble::rownames_to_column(var = "samples") %>% 
-  tidyr::unite("Dx_Sex",
-               dplyr::all_of(c("Dx", "Reported_Gender")),
-               remove = TRUE) %>% 
-  dplyr::mutate(`Dx_Sex` = as.factor(`Dx_Sex`)) %>% 
-  tibble::column_to_rownames(var = "samples")
-
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Dx_Sex", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate", "ageOfDeath")
+interaction <- c("Dx", "Reported_Gender")
+adjust.covars <- c("ageOfDeath", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate")
 random_effect <- c("IsolationBatch")
@@
 # Post adjusted formula
-formula <- glue::glue("~ ", glue::glue_collapse(scaled_adjust_covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))
+formula <- glue::glue("~ ", glue::glue_collapse(interaction, sep = "*")," + ", glue::glue_collapse(scaled_adjust_covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))
 
@@
 
-formula_for_variance_fractions <- glue::glue("~ ", glue::glue_collapse(formula_for_variance_fractions, sep = " + "))
-
+# Manually specify interaction
+interaction <- c(interaction, "Dx:Reported_Gender")
+interaction_variance <- glue::glue("(1|{interaction})")
+all <- c(interaction_variance, formula_for_variance_fractions)
+formula_for_variance_fractions <- glue::glue("~ ", glue::glue_collapse(all, sep = " + "))
 violin_plot <- variancePartition::fitExtractVarPartModel(VOOM.GENE_EXPRESSION$E,
@@
 ```
+`r formula`
 
 ### Residual calculation
-Calculate weighted residuals and add back "Dx_Sex" to the residuals
+Calculate weighted residuals and add back "Dx:Sex" to the residuals
 ```{r varsToAddBack}
 # Add variable of interest back to the residuals
-varsToAddIn = grep("Dx_Sex", colnames(ADJUSTED.FIT$design), value = T)
+varsToAddIn = grep("DxOther:Reported_GenderMale", "DxSCZ:Reported_GenderMale", colnames(ADJUSTED.FIT$design), value = T)
 RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION +
@@
 ```{r set-formula}
-formula <- glue::glue("~ 0 + ", glue::glue_collapse(adjust.covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))    
+formula <- glue::glue("~ 0 + Dx_Sex + ", glue::glue_collapse(adjust.covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))
 ```
@@
 
+Dx_Sex is 6 levels, SCZ_Female, SCZ_Male, Other_Female, Other_Male, Control_Female, Control_Male
+
 ```{r diffExp, fig.height=10, fig.width=15}
+# Test Dx-Sex
+COVARIATES <- COVARIATES %>%
+  tibble::rownames_to_column(var = "samples") %>% 
+  tidyr::unite("Dx_Sex",
+               dplyr::all_of(c("Dx", "Reported_Gender")),
+               remove = TRUE) %>% 
+  dplyr::mutate(`Dx_Sex` = as.factor(`Dx_Sex`)) %>% 
+  tibble::column_to_rownames(var = "samples")
 L1 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
@@
                                    data = COVARIATES, 
-                                   coefficient = c("Dx_SexControl_Female", "Dx_SexControl_Male"))
-L2 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
-                                   formula = formula, 
-                                   data = COVARIATES, 
-                                   coefficient = c("Dx_SexSCZ_Female", "Dx_SexSCZ_Male"))
-L3 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
-                                   formula = formula, 
-                                   data = COVARIATES, 
-                                   coefficient = c("Dx_SexOther_Female", "Dx_SexOther_Male"))
-
+                                   coefficient = c("Dx_SexControl_Male", "Dx_SexControl_Female"))
+L2 = c(1, -1, 0, 0, -1, 1, 0, 0, 0, 0, 0)
+L3 = c(1, -1, -1, 1, 0, 0, 0, 0, 0, 0, 0)
 # Manually set comparison between SCZ+Other vs. Control 
-L = cbind(L1, L2, L3, 
-          data.frame(L4 = c(1,0,-0.5, 0, -0.5, 0, 0, 0, 0, 0, 0)),
-          data.frame(L5 = c(0, 1, 0, -0.5, 0, -0.5, 0, 0, 0, 0, 0)))
+L = cbind(L1, L2, L3)
 
 contrast <- c(glue::glue_collapse(names(L1[L1 != 0]), "-"), 
-              glue::glue_collapse(names(L2[L2 != 0]), "-"),
-              glue::glue_collapse(names(L3[L3 != 0]), "-"),
-              glue::glue("Dx_SexControl_Female-Dx_SexOther_FemaleDx_SexSCZ_Female"),
-              glue::glue("Dx_SexControl_Male-Dx_SexOther_MaleDx_SexSCZ_Male")
+              glue::glue("Dx_SexSCZ_MaleDx_SexControl_Male-Dx_SexSCZ_FemaleDx_SexControl_Female"),
+              glue::glue("Dx_SexOther_MaleDx_SexControl_Male-Dx_SexOther_FemaleDx_SexControl_Female")
 )
-
 # Visualize contrast matrix
@@
 counts = 'syn21867866'
-ALL_USED_IDs = counts
+ALL_USED_IDs = c(counts, ALL_USED_IDs)
 counts = downloadFile(counts) %>% data.frame()
@@
 ```
+### Source Code
+`r thisFile`
 

### 2020-08-19 individual-cohorts/meta_analysis/meta-analysis_diagnosis.Rmd
Write results to Synapse

@@
 date: "`r date()`"
-output: html_document
+output:
+  html_document:
+    toc: true
+    toc_float: true
 ---
 
-```{r knit2synapse, eval=FALSE}
+```{r knit2synapse, eval=FALSE, echo=FALSE}
 # rmarkdown::render("meta-analysis_diagnosis.Rmd")
@@
 synLogin()
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/meta-analysis.Rmd",
-                                          parentId ="syn22276904",
-                                          folderName = 'Meta-analysis on diagnosis')
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/meta_analysis/meta-analysis_diagnosis.Rmd",
+                                          parentId ="syn22315403",
+                                          folderName = 'Meta-analysis on diagnosis',
+                                          overwrite=TRUE)
 ```
@@
 synLogin()
-
+suppressPackageStartupMessages({
 library(ggplot2)
@@
 library(cowplot)
+library(kableExtra)
 library(githubr)
@@
 library(metafor)
+library(knitr)
 library(patchwork)
+})
 
@@
 ```{r synapse.parameters, include=FALSE, cache=FALSE, eval=TRUE}
-thisFileName <- 'meta-analysis.Rmd'
+thisFileName <- 'meta-analysis_diagnosis.Rmd'
 
@@
 thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
-thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/',thisFileName))
+thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/meta_analysis/',thisFileName))
 ```
@@
 mpp_acc <- readr::read_tsv(synGet("syn22079117")$path) %>% 
-  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val)
+  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% getGeneSymbol( 'gene_id' )
+
 mpp_dlpfc <- readr::read_tsv(synGet("syn22079472")$path) %>% 
-  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val)
+  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% getGeneSymbol( 'gene_id' )
+
 hbcc_dsACC <- readr::read_tsv(synGet("syn22086829")$path) %>% 
-  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) 
+  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val)  %>% getGeneSymbol( 'gene_id' )
+
 hbcc_dlpfc <- readr::read_tsv(synGet("syn22084205")$path) %>% 
-  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val)
+  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% getGeneSymbol( 'gene_id' )
+Ids <- c("syn22079117", "syn22079472", "syn22086829", "syn22084205")
 ```
@@
 
+## DLPFC
+
 ```{r plot.meta_analysisSCZ}
-kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
-kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
-
-plotVolcano( res_dlpfc, 20) + ggtitle("Meta analysis - DLPFC")
-
-plotVolcano( res_acc, 20) + ggtitle("Meta analysis - ACC")
+kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05")) %>%
+  kable_styling(full_width = FALSE)
 ```
 
-# Control vs Other/SCZ
-```{r meta_analysis.SCZOther}
-make_list <- function(mpp, hbcc, string) {
-  list <- list(MPP = mpp %>%
-                 unique() %>% 
-                 dplyr::filter(Comparison == string) %>% 
-                 as.data.frame,
-               HBCC = hbcc %>% 
-                 unique() %>% 
-                 dplyr::filter(Comparison == string) %>% 
-                 as.data.frame)
-}
+```{r plot.concordance.dlpfc, fig.height=20}
 
-dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Control_vs_OtherSCZ")
-accList <- make_list(mpp_acc, hbcc_dsACC, string = "Control_vs_OtherSCZ")
+# get genes to highlight
+idx = order(res_dlpfc$adj.P.Val)
+showGenes = res_dlpfc$Symbol[idx[1:20]]
 
-rownames(dlpfcList[['MPP']]) = dlpfcList[['MPP']]$gene_id
-rownames(dlpfcList[['HBCC']]) = dlpfcList[['HBCC']]$gene_id
+resConcord_dlpfc = plot_concordance( dlpfcList, showGenes )
 
-res_dlpfc = run_meta_analysis( dlpfcList, "FE")
-res_dlpfc$Dataset = "Meta-analysis"
-
-rownames(accList[['MPP']]) = accList[['MPP']]$gene_id
-rownames(accList[['HBCC']]) = accList[['HBCC']]$gene_id
-
-res_acc = run_meta_analysis( accList, "FE")
-res_acc$Dataset = "Meta-analysis"
+plot_grid(resConcord_dlpfc$fig_logFC, 
+          resConcord_dlpfc$fig_tstat, 
+          plotVolcano( res_dlpfc, showGenes),   
+          nrow=3)
 ```
 
-```{r plot.meta_analysisSCZOther}
-kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
-kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
+```{r plot.concordance.table.dlpfc, fig.width=6}
+kable( resConcord_dlpfc$tab_pi ) %>%
+  kable_styling(full_width = FALSE)
 
-plotVolcano( res_dlpfc, 20) + ggtitle("Meta analysis - DLPFC")
+resConcord_dlpfc$tab_corr$P = format(resConcord_dlpfc$tab_corr$P, scientific=TRUE, digits=3)
+kable( resConcord_dlpfc$tab_corr )  %>%
+  kable_styling(full_width = FALSE)
+```
 
-plotVolcano( res_acc, 20) + ggtitle("Meta analysis - ACC")
+```{r test, eval=FALSE}
+# # forrest plot of meta-analysis
+# geneSymbol = 'RCSD1'
+
+# idx = with(res_dlpfc, (Symbol == geneSymbol) & !is.na(Symbol) )
+# kable(res_dlpfc[idx,] )
+```
+
+```{r plot.forrest.dlpfc, fig.width=5, fig.height=20}  
+figList = lapply(showGenes[1:10], function(geneSymbol){
+  plot_forrest( c(dlpfcList, list("Meta-analysis" = res_dlpfc)), geneSymbol)
+})
+
+plot_grid(plotlist=figList, ncol=1)
+```
+
+## ACC
+
+```{r plot.meta_analysisSCZ.acc}
+kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))%>%
+  kable_styling(full_width = FALSE)
+```
+```{r plot.concordance.acc, fig.height=20}
+
+# get genes to highlight
+idx = order(res_acc$adj.P.Val)
+showGenes = res_acc$Symbol[idx[1:20]]
+
+resConcord_acc = plot_concordance( accList, showGenes )
+
+plot_grid(resConcord_acc$fig_logFC, 
+          resConcord_acc$fig_tstat, 
+          plotVolcano( res_acc, showGenes),   
+          nrow=3)
+```
+
+```{r plot.concordance.table.acc, fig.width=6}
+kable( resConcord_acc$tab_pi ) %>%
+  kable_styling(full_width = FALSE)
+
+resConcord_acc$tab_corr$P = format(resConcord_acc$tab_corr$P, scientific=TRUE, digits=3)
+kable( resConcord_acc$tab_corr )  %>%
+  kable_styling(full_width = FALSE)
+```
+
+```{r plot.forrest.acc, fig.width=5, fig.height=20}  
+figList = lapply(showGenes[1:10], function(geneSymbol){
+  plot_forrest( c(accList, list("Meta-analysis" = res_acc)), geneSymbol)
+})
+
+plot_grid(plotlist=figList, ncol=1)
+```
+
+## Compare across brain regions
+```{r plot.concordance.combine, fig.height=15}
+
+resCombine = merge(res_dlpfc, res_acc, by="row.names")
+
+res_dlpfc$t = with(res_dlpfc, logFC / se)
+res_acc$t = with(res_acc, logFC / se)
+
+resConcord = plot_concordance( list(DLPFC = res_dlpfc, ACC = res_acc), showGenes )
+
+plot_grid(resConcord$fig_logFC, 
+          resConcord$fig_tstat,     
+          nrow=2) 
+```
+
+```{r plot.concordance.table.combine, fig.width=6}
+kable( resConcord$tab_pi ) %>%
+  kable_styling(full_width = FALSE)
+
+resConcord$tab_corr$P = format(resConcord$tab_corr$P, scientific=TRUE, digits=3)
+kable( resConcord$tab_corr )  %>%
+  kable_styling(full_width = FALSE)
 ```
@@
 # Control vs Other
-```{r meta_analysis.CTOther}
-make_list <- function(mpp, hbcc, string) {
-  list <- list(MPP = mpp %>%
-                 unique() %>% 
-                 dplyr::filter(Comparison == string) %>% 
-                 as.data.frame,
-               HBCC = hbcc %>% 
-                 unique() %>% 
-                 dplyr::filter(Comparison == string) %>% 
-                 as.data.frame)
-}
-
+```{r meta_analysis.Other}
 dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Control_vs_Other")
@@
 
-```{r plot.meta_analysisCTOther}
-kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
-kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
+## DLPFC
+```{r plot.meta_analysisOther}
+kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05")) %>%
+  kable_styling(full_width = FALSE)
+```
 
-plotVolcano( res_dlpfc, 20) + ggtitle("Meta analysis - DLPFC")
+```{r plot.concordance.dlpfc.Other, fig.height=20}
 
-plotVolcano( res_acc, 20) + ggtitle("Meta analysis - ACC")
+# get genes to highlight
+idx = order(res_dlpfc$adj.P.Val)
+showGenes = res_dlpfc$Symbol[idx[1:20]]
+
+resConcord_dlpfc = plot_concordance( dlpfcList, showGenes )
+
+plot_grid(resConcord_dlpfc$fig_logFC, 
+          resConcord_dlpfc$fig_tstat, 
+          plotVolcano( res_dlpfc, showGenes),   
+          nrow=3)
+```
+
+```{r plot.concordance.table.dlpfc.Other, fig.width=6}
+kable( resConcord_dlpfc$tab_pi ) %>%
+  kable_styling(full_width = FALSE)
+
+resConcord_dlpfc$tab_corr$P = format(resConcord_dlpfc$tab_corr$P, scientific=TRUE, digits=3)
+kable( resConcord_dlpfc$tab_corr )  %>%
+  kable_styling(full_width = FALSE)
+```
+
+```{r plot.forrest.dlpfc.Other, fig.width=5, fig.height=20}  
+figList = lapply(showGenes[1:10], function(geneSymbol){
+  plot_forrest( c(dlpfcList, list("Meta-analysis" = res_dlpfc)), geneSymbol)
+})
+
+plot_grid(plotlist=figList, ncol=1)
+```
+
+## ACC
+
+```{r plot.meta_analysisSCZ.acc.Other}
+kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))%>%
+  kable_styling(full_width = FALSE)
+```
+```{r plot.concordance.acc.Other, fig.height=20}
+
+# get genes to highlight
+idx = order(res_acc$adj.P.Val)
+showGenes = res_acc$Symbol[idx[1:20]]
+
+resConcord_acc = plot_concordance( accList, showGenes )
+
+plot_grid(resConcord_acc$fig_logFC, 
+          resConcord_acc$fig_tstat, 
+          plotVolcano( res_acc, showGenes),   
+          nrow=3)
+```
+
+```{r plot.concordance.table.acc.Other, fig.width=6}
+kable( resConcord_acc$tab_pi ) %>%
+  kable_styling(full_width = FALSE)
+
+resConcord_acc$tab_corr$P = format(resConcord_acc$tab_corr$P, scientific=TRUE, digits=3)
+kable( resConcord_acc$tab_corr )  %>%
+  kable_styling(full_width = FALSE)
+```
+
+```{r plot.forrest.acc.Other, fig.width=5, fig.height=20}  
+figList = lapply(showGenes[1:10], function(geneSymbol){
+  plot_forrest( c(accList, list("Meta-analysis" = res_acc)), geneSymbol)
+})
+
+plot_grid(plotlist=figList, ncol=1)
+```
+
+## Compare across brain regions
+```{r plot.concordance.combine.Other, fig.height=15}
+
+resCombine = merge(res_dlpfc, res_acc, by="row.names")
+
+res_dlpfc$t = with(res_dlpfc, logFC / se)
+res_acc$t = with(res_acc, logFC / se)
+
+resConcord = plot_concordance( list(DLPFC = res_dlpfc, ACC = res_acc), showGenes )
+
+plot_grid(resConcord$fig_logFC, 
+          resConcord$fig_tstat,     
+          nrow=2) 
+```
+
+```{r plot.concordance.table.combine.Other, fig.width=6}
+kable( resConcord$tab_pi ) %>%
+  kable_styling(full_width = FALSE)
+
+resConcord$tab_corr$P = format(resConcord$tab_corr$P, scientific=TRUE, digits=3)
+kable( resConcord$tab_corr )  %>%
+  kable_styling(full_width = FALSE)
 ```
@@
 # Other vs SCZ
-```{r meta_analysis.OtherSCZ}
-make_list <- function(mpp, hbcc, string) {
-  list <- list(MPP = mpp %>%
-                 unique() %>% 
-                 dplyr::filter(Comparison == string) %>% 
-                 as.data.frame,
-               HBCC = hbcc %>% 
-                 unique() %>% 
-                 dplyr::filter(Comparison == string) %>% 
-                 as.data.frame)
-}
-
+```{r meta_analysis.OtherSz}
 dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Other_vs_SCZ")
@@
 
-```{r plot.meta_analysisOtherSCZ}
-kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
-kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
-
-plotVolcano( res_dlpfc, 20) + ggtitle("Meta analysis - DLPFC")
-
-plotVolcano( res_acc, 20) + ggtitle("Meta analysis - ACC")
+## DLPFC
+```{r plot.meta_analysis.OtherSz}
+kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05")) %>%
+  kable_styling(full_width = FALSE)
 ```
 
-## RCSD1
+```{r plot.concordance.dlpfc..OtherSz, fig.height=20}
 
-Forest plot of example top gene.
+# get genes to highlight
+idx = order(res_dlpfc$adj.P.Val)
+showGenes = res_dlpfc$Symbol[idx[1:20]]
 
-```{r specific-gene}
-# forrest plot of meta-analysis
-ensID = 'ENSG00000198771.11'
-kable(res_dlpfc[ensID,])
-lapply(dlpfcList, function(x) x[ensID,])
-ensID = rownames(res_dlpfc)[which.min(res_dlpfc$adj.P.Val)][1]
-plot_forrest(dlpfcList, ensID)
+resConcord_dlpfc = plot_concordance( dlpfcList, showGenes )
+
+plot_grid(resConcord_dlpfc$fig_logFC, 
+          resConcord_dlpfc$fig_tstat, 
+          plotVolcano( res_dlpfc, showGenes),   
+          nrow=3)
+```
+
+```{r plot.concordance.table.dlpfc.OtherSz, fig.width=6}
+kable( resConcord_dlpfc$tab_pi ) %>%
+  kable_styling(full_width = FALSE)
+
+resConcord_dlpfc$tab_corr$P = format(resConcord_dlpfc$tab_corr$P, scientific=TRUE, digits=3)
+kable( resConcord_dlpfc$tab_corr )  %>%
+  kable_styling(full_width = FALSE)
+```
+
+```{r plot.forrest.dlpfc..OtherSz, fig.width=5, fig.height=20}  
+figList = lapply(showGenes[1:10], function(geneSymbol){
+  plot_forrest( c(dlpfcList, list("Meta-analysis" = res_dlpfc)), geneSymbol)
+})
+
+plot_grid(plotlist=figList, ncol=1)
+```
+
+## ACC
+
+```{r plot.meta_analysisSCZ.acc.OtherSz}
+kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))%>%
+  kable_styling(full_width = FALSE)
+```
+```{r plot.concordance.acc.OtherSz, fig.height=20}
+
+# get genes to highlight
+idx = order(res_acc$adj.P.Val)
+showGenes = res_acc$Symbol[idx[1:20]]
+
+resConcord_acc = plot_concordance( accList, showGenes )
+
+plot_grid(resConcord_acc$fig_logFC, 
+          resConcord_acc$fig_tstat, 
+          plotVolcano( res_acc, showGenes),   
+          nrow=3)
+```
+
+```{r plot.concordance.table.acc.OtherSz, fig.width=6}
+kable( resConcord_acc$tab_pi ) %>%
+  kable_styling(full_width = FALSE)
+
+resConcord_acc$tab_corr$P = format(resConcord_acc$tab_corr$P, scientific=TRUE, digits=3)
+kable( resConcord_acc$tab_corr )  %>%
+  kable_styling(full_width = FALSE)
+```
+
+```{r plot.forrest.acc.OtherSz, fig.width=5, fig.height=20}  
+figList = lapply(showGenes[1:10], function(geneSymbol){
+  plot_forrest( c(accList, list("Meta-analysis" = res_acc)), geneSymbol)
+})
+
+plot_grid(plotlist=figList, ncol=1)
+```
+
+## Compare across brain regions
+```{r plot.concordance.combine.OtherSz, fig.height=15}
+
+resCombine = merge(res_dlpfc, res_acc, by="row.names")
+
+res_dlpfc$t = with(res_dlpfc, logFC / se)
+res_acc$t = with(res_acc, logFC / se)
+
+resConcord = plot_concordance( list(DLPFC = res_dlpfc, ACC = res_acc), showGenes )
+
+plot_grid(resConcord$fig_logFC, 
+          resConcord$fig_tstat,     
+          nrow=2) 
+```
+
+```{r plot.concordance.table.combine.OtherSz, fig.width=6}
+kable( resConcord$tab_pi ) %>%
+  kable_styling(full_width = FALSE)
+
+resConcord$tab_corr$P = format(resConcord$tab_corr$P, scientific=TRUE, digits=3)
+kable( resConcord$tab_corr )  %>%
+  kable_styling(full_width = FALSE)
+```
+
+```{r store, eval=TRUE}
+# Code
+CODE <- Folder(name = "Meta-analysis on diagnosis", parentId = "syn22315403") # Make sure these match the folder name and parent Id in the first chunk.
+CODE <- synStore(CODE)
+
+# Store meta-analysis
+res_dlpfc %>%
+  tibble::rownames_to_column(var = "gene_id") %>% 
+  write.table(file = 'CMC_dlpfc_diagnosis_meta_analysis.tsv', sep = '\t', row.names=F, quote=F)
+nEXP_OBJ = File('CMC_dlpfc_diagnosis_meta_analysis.tsv', 
+                name = 'diagnosis Normalised, meta-analysis (DLPFC)', 
+                parentId = CODE$properties$id)
+nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, 
+                    executed = thisFile)
+res_acc %>%
+  tibble::rownames_to_column(var = "gene_id") %>% 
+  write.table(file = 'CMC_acc_diagnosis_meta_analysis.tsv', sep = '\t', row.names=F, quote=F)
+nEXP_OBJ = File('CMC_acc_diagnosis_meta_analysis.tsv', 
+                name = 'diagnosis Normalised, meta-analysis (ACC)', 
+                parentId = CODE$properties$id)
+nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, 
+                    executed = thisFile)
 ```

### 2020-08-19 individual-cohorts/meta_analysis/meta-analysis_sex.Rmd
Write results to Synapse

@@
 
-# rmarkdown::render("meta-analysis_sex.Rmd")
+# rmarkdown::render("individual-cohorts/meta_analysis/meta-analysis_sex.Rmd")
+
 --->
 
-```{r knit2synapse, eval=TRUE}
+```{r knit2synapse, eval=FALSE}
 library(synapser)
@@
 synLogin()
-knit2synapse::createAndKnitToFolderEntity(file = "meta-analysis_sex.Rmd",
-                                          parentId ="syn22281021",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/meta_analysis/meta-analysis_sex.Rmd",
+                                          parentId ="syn22315403",
                                           folderName = 'Meta-analysis on sex',
-                                          overwrite=FALSE)
+                                          overwrite=TRUE)
 ```
@@
 
+```{r plot.logFC_by_chr.dlpfc}
+plot_circos_logFC( res_dlpfc ) + ggtitle("DLPFC sex differential expression")
+```
+
 ### MPP
@@
 
+```{r plot.logFC_by_chr.acc}
+plot_circos_logFC( res_acc) + ggtitle("ACC sex differential expression")
+```
+
 ### MPP
@@
 # Code
-CODE <- Folder(name = "Meta-analysis on sex", parentId = "syn22281021") # Make sure these match the folder name and parent Id in the first chunk.
+CODE <- Folder(name = "Meta-analysis on sex", parentId = "syn22315403") # Make sure these match the folder name and parent Id in the first chunk.
 CODE <- synStore(CODE)
@@
 ### Source
-#[markdown](`r thisFile`)
+[markdown](`r thisFile`)

### 2020-08-19 individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_Diagnosis.Rmd
inital commit

--- 2020-05-20 individual-cohorts/HBCC_sgACC_DE.Rmd
@@
  ---
-title: "Differential Expression - HBCC - sgACC - Dx"
+title: "Differential Expression - HBCC - DLPFC - Dx"
 author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/HBCC_sgACC_DE.Rmd",
-                                 parentId ="syn22084216",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_Diagnosis.Rmd",
+                                 parentId ="syn22050019",
                                  folderName = "Differential Expression - Dx")
@@
 ```{r synapse.parameters, include=FALSE, cache=TRUE}
-parentId = 'syn22084216';
+parentId = 'syn22050019';
 activityName = 'Differential Expression';
@@
 
-thisFileName <- 'HBCC_sgACC_DE.Rmd'
+thisFileName <- 'HBCC_DLPFC_DE.Rmd'
 
@@
 # Download CQN counts
-CQN_ID = "syn22084946"
+CQN_ID = "syn22082460"
 ALL_USED_IDs = c(CQN_ID)
@@
 # Download filtered counts
-COUNTS_ID = "syn22084938"
+COUNTS_ID = "syn22082459"
 ALL_USED_IDs = c(ALL_USED_IDs, COUNTS_ID)
@@
 # Get covariates
-COVAR_ID = "syn22084935"
+COVAR_ID = "syn22082458"
 ALL_USED_IDs = c(ALL_USED_IDs, COVAR_ID)
@@
 # Set variable class
-FactorCovariates <- c("Individual_ID", "Reported_Gender", "LibraryBatch", "Dx", "IsolationBatch")
+FactorCovariates <- c("Individual_ID", "Reported_Gender", "LibraryBatch", "Dx", "RibozeroBatch", "FlowcellBatch")
 COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
@@
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Dx", "RIN2", "IntronicRate")
-random_effect <- c("Reported_Gender", "LibraryBatch")
+adjust.covars <- c("Dx", "ageOfDeath", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate")
+random_effect <- c("Reported_Gender")
 ```
@@
                  glue::glue_collapse(adjust.covars, sep = ", ", last = " and "),
-                 ' as fixed effects and ',
+                 ' as fixed effects and',
                  glue::glue_collapse(random_effect, sep = ", ", last = " and "),' chosen as random effects'))
 
-scaled_col_type <- dplyr::select(COVARIATES, c(adjust.covars)) %>%
+col_type <- dplyr::select(COVARIATES, c(adjust.covars, random_effect)) %>%
   dplyr::summarise_all(class) %>% 
@@
 # Scale continuous variables
-scaled_adjust_covars = sapply(1:length(scaled_col_type$class), function(i){
-  switch(scaled_col_type$class[i], 
-    "factor" = paste0(scaled_col_type$variable[i]),
-    "numeric" = paste0('scale(', scaled_col_type$variable[i], ')')
+adjust.covars = sapply(1:length(col_type$class), function(i){
+  switch(col_type$class[i], 
+    "factor" = paste0(col_type$variable[i]),
+    "numeric" = paste0('scale(', col_type$variable[i], ')')
   )
@@
 # Post adjusted formula
-formula <- glue::glue("~ ", glue::glue_collapse(scaled_adjust_covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))
+formula <- glue::glue("~ ", glue::glue_collapse(adjust.covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))
 
@@
 VOOM.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION
-ADJUSTED.FIT = variancePartition::dream(exprObj = VOOM.GENE_EXPRESSION$E,
+ADJUSTED.FIT = variancePartition::dream(exprObj = VOOM.GENE_EXPRESSION,
                                         formula = formula,
@@
 plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>% 
-  dplyr::mutate(Tissue = "sgACC")
+  dplyr::mutate(Tissue = "DLPFC")
 
@@
 # Get gene lengths
-counts = 'syn21867762'
+counts = 'syn21886235'
 ALL_USED_IDs = counts
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_sgACC_Dx_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
-nEXP_OBJ = File('CMC_sgACC_Dx_netResidualExpression.tsv', 
+  write.table(file = 'CMC_DLPFC_Dx_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
+nEXP_OBJ = File('CMC_DLPFC_Dx_netResidualExpression.tsv', 
                 name = 'Dx Normalised, covariates removed residual expression (for network analysis)', 
@@
 # Store differential expression results
-write.table(DE, file = 'CMC_sgACC_Dx_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
-DEXP_OBJ = File('CMC_sgACC_Dx_DiffExpression.tsv', 
+write.table(DE, file = 'CMC_DLPFC_Dx_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
+DEXP_OBJ = File('CMC_DLPFC_Dx_DiffExpression.tsv', 
                 name = 'Differential Expression Results (Dx)', 
@@
 # Store adjusted fit linear mixed model object as RDS - Dx
-saveRDS(ADJUSTED.FIT, file = 'CMC_sgACC_Dx_dream-output.rds')
-ADJ_OBJ = File('CMC_sgACC_Dx_dream-output.rds',
+saveRDS(ADJUSTED.FIT, file = 'CMC_DLPFC_Dx_dream-output.rds')
+ADJ_OBJ = File('CMC_DLPFC_Dx_dream-output.rds',
                name = 'Dream Output object - Dx', 
@@
 synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store variancePartition
+saveRDS(sortCols(violin_plot), file = 'CMC_DLPFC_Dx_variancePartition.rds')
+ADJ_OBJ = File('CMC_DLPFC_Dx_variancePartition.rds',
+               name = 'Variance Partition result', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store voom object
+saveRDS(VOOM.GENE_EXPRESSION, file = 'CMC_DLPFC_Dx_voom.rds')
+ADJ_OBJ = File('CMC_DLPFC_Dx_voom.rds',
+               name = 'voom object', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store contrasts
+saveRDS(L, file = 'CMC_DLPFC_Dx_contrasts.rds')
+ADJ_OBJ = File('CMC_DLPFC_Dx_contrasts.rds',
+               name = 'contrast matrix', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store contrasts
+saveRDS(formula, file = 'CMC_DLPFC_Dx_formula.rds')
+ADJ_OBJ = File('CMC_DLPFC_Dx_formula.rds',
+               name = 'formula', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
 ```

### 2020-08-19 individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_Sex.Rmd
inital commit

--- 2020-07-30 individual-cohorts/MSSM-Penn-Pitt_DLPFC_DE.Rmd
@@
  ---
-title: "Differential Expression - MSSM-Penn-Pitt - DLPFC - Sex"
+title: "Differential Expression - HBCC - DLPFC - Sex"
 author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/MSSM-Penn-Pitt_DLPFC_DE.Rmd",
-                                 parentId ="syn22043922",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/HBCC_DLPFC_DE.Rmd",
+                                 parentId ="syn22050019",
                                  folderName = "Differential Expression - Sex")
@@
 library(githubr) # get the package from devtools::install_github('brian-bot/githubr')
-library(WGCNA)
 
@@
 ```{r synapse.parameters, include=FALSE, cache=FALSE}
-parentId = 'syn22043922';
+parentId = 'syn22050019';
 activityName = 'Differential Expression';
@@
 # Get covariates
-COVAR_ID = "syn22045728"
+COVAR_ID = "syn22082458"
 ALL_USED_IDs = c(ALL_USED_IDs, COVAR_ID)
-COVARIATES = downloadFile(COVAR_ID)
-
-# dropped individual id 
-MD_ID = "syn16816488"
-ALL_USED_IDs = c(ALL_USED_IDs, COUNTS_ID)
-METADATA = downloadFile(MD_ID) %>% 
-  dplyr::select(Individual_ID, Sample_RNA_ID)
-
-COVARIATES = right_join(METADATA, COVARIATES, by = c("Sample_RNA_ID" = "SampleID")) %>% 
-  tibble::column_to_rownames(var = "Sample_RNA_ID")
+COVARIATES = downloadFile(COVAR_ID) %>% 
+  tibble::column_to_rownames(var = "SampleID")
 
 # Set variable class
-FactorCovariates <- c("Individual_ID", "Institution", "LibraryBatch", "Dx", "Reported_Gender", "RibozeroBatch", "FlowcellBatch")
+FactorCovariates <- c("Individual_ID", "Reported_Gender", "LibraryBatch", "Dx", "RibozeroBatch", "FlowcellBatch")
 COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
@@
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Sex", "ageOfDeath", "RIN2", "RIN", "IntronicRate", "IntragenicRate")
-random_effect <- c("Institution")
+adjust.covars <- c("Sex", "Dx", "ageOfDeath", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate")
 ```
@@
                  glue::glue_collapse(adjust.covars, sep = ", ", last = " and "),
-                 ' as fixed effects and',
-                 glue::glue_collapse(random_effect, sep = ", ", last = " and "),' chosen as random effects'))
+                 ' as fixed effects.'))
+
+col_type <- dplyr::select(COVARIATES, c(adjust.covars)) %>%
+  dplyr::summarise_all(class) %>% 
+  tidyr::gather(variable, class)
+
+# Scale continuous variables
+adjust.covars = sapply(1:length(col_type$class), function(i){
+  switch(col_type$class[i], 
+    "factor" = paste0(col_type$variable[i]),
+    "numeric" = paste0('scale(', col_type$variable[i], ')')
+  )
+  })
 
 # Post adjusted formula
-formula <- glue::glue("~ ", glue::glue_collapse(adjust.covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))
+formula <- glue::glue("~ ", glue::glue_collapse(adjust.covars, sep = " + "))
 
@@
 # Eucledian tree based analysis
-COVARIATES.tmp = data.matrix(COVARIATES[,c('Sex','Institution')])
+COVARIATES.tmp = data.matrix(COVARIATES[,c('Sex'), drop = FALSE])
 
@@
                            main = "Sample dendrogram",
-                           groupLabels = colnames(COVARIATES.tmp[,c('Sex','Institution')]))
+                           groupLabels = colnames(COVARIATES.tmp[,c('Sex'), drop = FALSE]))
 ```
@@
 
-### Differential expression analysis (Female - Male)
+### Differential expression analysis (SCZ - Control - Other)
 Genes that are differentially expressed at an FDR <= 0.05 are:
 ```{r set-formula}
-adjust.covars <- adjust.covars[!(adjust.covars %in% "RIN2")]
-formula <- glue::glue("~ 0 + ", glue::glue_collapse(adjust.covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))    
+formula <- glue::glue("~ 0 + ", glue::glue_collapse(adjust.covars, sep = " + "))    
 ```
@@
                                    coefficient = c("SexFemale", "SexMale"))
+
 contrast <- c(glue::glue_collapse(names(L[L != 0]), "-"))
@@
 synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store variancePartition
+saveRDS(sortCols(violin_plot), file = 'CMC_DLPFC_Sex_variancePartition.rds')
+ADJ_OBJ = File('CMC_DLPFC_Sex_variancePartition.rds',
+               name = 'Variance Partition result', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store voom object
+saveRDS(VOOM.GENE_EXPRESSION, file = 'CMC_DLPFC_Sex_voom.rds')
+ADJ_OBJ = File('CMC_DLPFC_Sex_voom.rds',
+               name = 'voom object', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store contrasts
+saveRDS(L, file = 'CMC_DLPFC_Sex_contrasts.rds')
+ADJ_OBJ = File('CMC_DLPFC_Sex_contrasts.rds',
+               name = 'contrast matrix', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store contrasts
+saveRDS(formula, file = 'CMC_DLPFC_Sex_formula.rds')
+ADJ_OBJ = File('CMC_DLPFC_Sex_formula.rds',
+               name = 'formula', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
 ```

### 2020-08-19 individual-cohorts/split_by_analysis/HBCC_dsACC_DE_Sex.Rmd
Add all

--- 2020-05-20 individual-cohorts/HBCC_dsACC_DE.Rmd
@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/HBCC_dsACC_DE.Rmd",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/split_by_analysis/HBCC_dsACC_DE_Sex.Rmd",
                                  parentId ="syn22084347",
-                                 folderName = "Differential Expression - Dx")
+                                 folderName = "Differential Expression - Sex")
 ```
@@
 
+COVARIATES <- dplyr::rename(COVARIATES, Sex = Reported_Gender)
+
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Dx", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate", "ageOfDeath")
-random_effect <- c("Reported_Gender", "IsolationBatch")
+adjust.covars <- c("Sex", "Dx", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate", "ageOfDeath")
+random_effect <- c("IsolationBatch")
 ```
 
-### Normalisation - Dx
-1. Dx is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis)
+### Normalisation - Sex
+1. Sex is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis)
 
@@
 ```{r diffExp, fig.height=10, fig.width=15}
-L1 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
+L = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
                                    formula = formula, 
                                    data = COVARIATES, 
-                                   coefficient = c("DxOther", "DxControl"))
-L2 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
-                                   formula = formula, 
-                                   data = COVARIATES, 
-                                   coefficient = c("DxSCZ", "DxControl"))
-L3 = variancePartition::getContrast(exprObj = VOOM.GENE_EXPRESSION, 
-                                   formula = formula, 
-                                   data = COVARIATES, 
-                                   coefficient = c("DxSCZ", "DxOther"))
+                                   coefficient = c("SexFemale", "SexMale"))
 
-L = cbind(L1, L2, L3)
-
-contrast <- c(glue::glue_collapse(names(L1[L1 != 0]), "-"), 
-              glue::glue_collapse(names(L2[L2 != 0]), "-"),
-              glue::glue_collapse(names(L3[L3 != 0]), "-")
-)
+contrast <- c(glue::glue_collapse(names(L[L != 0]), "-"))
 
@@
 # Get differential expression
-DE = lapply(glue::glue("L{1:3}"), function(i, FIT){
+DE = lapply(glue::glue("L1"), function(i, FIT){
   topTable(FIT, coef=i, number = Inf) %>%
@@
 synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store variancePartition
+saveRDS(sortCols(violin_plot), file = 'CMC_dsACC_Sex_variancePartition.rds')
+ADJ_OBJ = File('CMC_dsACC_Sex_variancePartition.rds',
+               name = 'Variance Partition result', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store voom object
+saveRDS(VOOM.GENE_EXPRESSION, file = 'CMC_dsACC_Sex_voom.rds')
+ADJ_OBJ = File('CMC_dsACC_Sex_voom.rds',
+               name = 'voom object', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store contrasts
+saveRDS(L, file = 'CMC_dsACC_Sex_contrasts.rds')
+ADJ_OBJ = File('CMC_dsACC_Sex_contrasts.rds',
+               name = 'contrast matrix', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store contrasts
+saveRDS(formula, file = 'CMC_dsACC_Sex_formula.rds')
+ADJ_OBJ = File('CMC_dsACC_Sex_formula.rds',
+               name = 'formula', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
 ```

### 2020-08-19 individual-cohorts/split_by_analysis/HBCC_sgACC_DE_Dx.Rmd
Add all

--- 2020-05-20 individual-cohorts/HBCC_sgACC_DE.Rmd
@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/HBCC_sgACC_DE.Rmd",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/split_by_analysis/HBCC_sgACC_DE_Dx.Rmd",
                                  parentId ="syn22084216",
@@
 synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store variancePartition
+saveRDS(sortCols(violin_plot), file = 'CMC_sgACC_Dx_variancePartition.rds')
+ADJ_OBJ = File('CMC_sgACC_Dx_variancePartition.rds',
+               name = 'Variance Partition result', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store voom object
+saveRDS(VOOM.GENE_EXPRESSION, file = 'CMC_sgACC_Dx_voom.rds')
+ADJ_OBJ = File('CMC_sgACC_Dx_voom.rds',
+               name = 'voom object', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store contrasts
+saveRDS(L, file = 'CMC_sgACC_Dx_contrasts.rds')
+ADJ_OBJ = File('CMC_sgACC_Dx_contrasts.rds',
+               name = 'contrast matrix', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store contrasts
+saveRDS(formula, file = 'CMC_sgACC_Dx_formula.rds')
+ADJ_OBJ = File('CMC_sgACC_Dx_formula.rds',
+               name = 'formula', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
 ```
+
+```

### 2020-08-19 individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_ACC_DE_Diagnosis.Rmd
Add all

--- 2020-08-19 individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_Diagnosis.Rmd
@@
 ---
-title: "Differential Expression - HBCC - DLPFC - Dx"
+title: "Differential Expression - MSSM-Penn-Pitt - ACC - Dx"
 author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_Diagnosis.Rmd",
-                                 parentId ="syn22050019",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/MSSM-Penn-Pitt_ACC_DE.Rmd",
+                                 parentId ="syn22012953",
                                  folderName = "Differential Expression - Dx")
@@
 
-```{r synapse.parameters, include=FALSE, cache=TRUE}
-parentId = 'syn22050019';
+```{r synapse.parameters, include=FALSE, cache=FALSE}
+parentId = 'syn22012953';
 activityName = 'Differential Expression';
@@
 
-thisFileName <- 'HBCC_DLPFC_DE.Rmd'
+thisFileName <- 'MSSM-Penn-Pitt_ACC_DE.Rmd'
 
@@
 # Download CQN counts
-CQN_ID = "syn22082460"
+CQN_ID = "syn22013154"
 ALL_USED_IDs = c(CQN_ID)
@@
 # Download filtered counts
-COUNTS_ID = "syn22082459"
+COUNTS_ID = "syn22013153"
 ALL_USED_IDs = c(ALL_USED_IDs, COUNTS_ID)
@@
 # Get covariates
-COVAR_ID = "syn22082458"
+COVAR_ID = "syn22013150"
 ALL_USED_IDs = c(ALL_USED_IDs, COVAR_ID)
-COVARIATES = downloadFile(COVAR_ID) %>% 
-  tibble::column_to_rownames(var = "SampleID")
+COVARIATES = downloadFile(COVAR_ID)
+
+# dropped individual id 
+MD_ID = "syn16816488"
+ALL_USED_IDs = c(ALL_USED_IDs, COUNTS_ID)
+METADATA = downloadFile(MD_ID) %>% 
+  dplyr::select(Individual_ID, Sample_RNA_ID)
+
+COVARIATES = right_join(METADATA, COVARIATES, by = c("Sample_RNA_ID" = "SampleID")) %>% 
+  tibble::column_to_rownames(var = "Sample_RNA_ID")
 
 # Set variable class
-FactorCovariates <- c("Individual_ID", "Reported_Gender", "LibraryBatch", "Dx", "RibozeroBatch", "FlowcellBatch")
+FactorCovariates <- c("Individual_ID", "Institution", "Reported_Gender", "LibraryBatch", "Dx", "RibozeroBatch", "FlowcellBatch")
 COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
@@
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Dx", "ageOfDeath", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate")
-random_effect <- c("Reported_Gender")
+adjust.covars <- c("Dx", "ageOfDeath", "RIN2", "RIN", "rRNARate", "IntronicRate", "IntragenicRate", "IntergenicRate")
+random_effect <- c("Institution", "Reported_Gender")
 ```
@@
                  glue::glue_collapse(random_effect, sep = ", ", last = " and "),' chosen as random effects'))
-
-col_type <- dplyr::select(COVARIATES, c(adjust.covars, random_effect)) %>%
-  dplyr::summarise_all(class) %>% 
-  tidyr::gather(variable, class)
-
-# Scale continuous variables
-adjust.covars = sapply(1:length(col_type$class), function(i){
-  switch(col_type$class[i], 
-    "factor" = paste0(col_type$variable[i]),
-    "numeric" = paste0('scale(', col_type$variable[i], ')')
-  )
-  })
 
@@
 
+col_type <- dplyr::select(COVARIATES, c(adjust.covars, random_effect)) %>%
+  dplyr::summarise_all(class) %>% 
+  tidyr::gather(variable, class)
+
 # Model categorical variables as random to report variance fractions
@@
     "factor" = paste0('(1|', col_type$variable[i], ')'),
-    "numeric" = paste0('scale(', col_type$variable[i], ')')
-  )
+    "numeric" = col_type$variable[i])
   })
@@
 plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>% 
-  dplyr::mutate(Tissue = "DLPFC")
+  dplyr::mutate(Tissue = "ACC")
 
 p <- ggplot(plotdata, aes(x=PC1, y=PC2))
-p <- p + geom_point(aes(shape=Tissue, size=ageOfDeath))
+p <- p + geom_point(aes(color=Institution, shape=Tissue, size=ageOfDeath))
 p <- p + theme_bw() + theme(legend.position="right") + facet_grid(Tissue~., scales = 'free_y')
@@
 # Eucledian tree based analysis
-COVARIATES.tmp = data.matrix(COVARIATES[,c('Dx'), drop = FALSE])
+COVARIATES.tmp = data.matrix(COVARIATES[,c('Dx','Institution')])
 
@@
                            main = "Sample dendrogram",
-                           groupLabels = colnames(COVARIATES.tmp[,c('Dx'), drop = FALSE]))
+                           groupLabels = colnames(COVARIATES.tmp[,c('Dx','Institution')]))
 ```
@@
 ```{r set-formula}
+adjust.covars <- adjust.covars[!(adjust.covars %in% "RIN2")]
+adjust.covars <- glue::glue("scale({adjust.covars})")
+"Dx" -> adjust.covars[1]
 formula <- glue::glue("~ 0 + ", glue::glue_collapse(adjust.covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))  
 ```
+`r formula`
 
-`r formula`
+Note:removed RIN2 from formula as RIN and RIN2 are highly correlated.
 
@@
 # Get gene lengths
-counts = 'syn21886235'
+counts = 'syn21867938'
 ALL_USED_IDs = counts
@@
 
-  tissue	= 'dorsolateral prefrontal cortex', 
+  tissue  = 'anterior cingulate cortex', 
   study = 'CMC',
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_DLPFC_Dx_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
-nEXP_OBJ = File('CMC_DLPFC_Dx_netResidualExpression.tsv', 
+  write.table(file = 'CMC_ACC_Dx_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
+nEXP_OBJ = File('CMC_ACC_Dx_netResidualExpression.tsv', 
                 name = 'Dx Normalised, covariates removed residual expression (for network analysis)', 
@@
 # Store differential expression results
-write.table(DE, file = 'CMC_DLPFC_Dx_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
-DEXP_OBJ = File('CMC_DLPFC_Dx_DiffExpression.tsv', 
+write.table(DE, file = 'CMC_ACC_Dx_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
+DEXP_OBJ = File('CMC_ACC_Dx_DiffExpression.tsv', 
                 name = 'Differential Expression Results (Dx)', 
@@
 # Store adjusted fit linear mixed model object as RDS - Dx
-saveRDS(ADJUSTED.FIT, file = 'CMC_DLPFC_Dx_dream-output.rds')
-ADJ_OBJ = File('CMC_DLPFC_Dx_dream-output.rds',
+saveRDS(ADJUSTED.FIT, file = 'CMC_ACC_Dx_dream-output.rds')
+ADJ_OBJ = File('CMC_ACC_Dx_dream-output.rds',
                name = 'Dream Output object - Dx', 
@@
 # Store variancePartition
-saveRDS(sortCols(violin_plot), file = 'CMC_DLPFC_Dx_variancePartition.rds')
-ADJ_OBJ = File('CMC_DLPFC_Dx_variancePartition.rds',
+saveRDS(sortCols(violin_plot), file = 'CMC_ACC_Dx_variancePartition.rds')
+ADJ_OBJ = File('CMC_ACC_Dx_variancePartition.rds',
                name = 'Variance Partition result', 
@@
 # Store voom object
-saveRDS(VOOM.GENE_EXPRESSION, file = 'CMC_DLPFC_Dx_voom.rds')
-ADJ_OBJ = File('CMC_DLPFC_Dx_voom.rds',
+saveRDS(VOOM.GENE_EXPRESSION, file = 'CMC_ACC_Dx_voom.rds')
+ADJ_OBJ = File('CMC_ACC_Dx_voom.rds',
                name = 'voom object', 
@@
 # Store contrasts
-saveRDS(L, file = 'CMC_DLPFC_Dx_contrasts.rds')
-ADJ_OBJ = File('CMC_DLPFC_Dx_contrasts.rds',
+saveRDS(L, file = 'CMC_ACC_Dx_contrasts.rds')
+ADJ_OBJ = File('CMC_ACC_Dx_contrasts.rds',
                name = 'contrast matrix', 
@@
 # Store contrasts
-saveRDS(formula, file = 'CMC_DLPFC_Dx_formula.rds')
-ADJ_OBJ = File('CMC_DLPFC_Dx_formula.rds',
+saveRDS(formula, file = 'CMC_ACC_Dx_formula.rds')
+ADJ_OBJ = File('CMC_ACC_Dx_formula.rds',
                name = 'formula', 

### 2020-08-19 individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_ACC_DE_Sex.Rmd
Add all

--- 2020-07-31 individual-cohorts/MSSM-Penn-Pitt_ACC_DE.Rmd
@@
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Sex", "Dx", "ageOfDeath", "RIN2", "RIN", "rRNARate", "IntronicRate", "IntragenicRate", "IntergenicRate")
+adjust.covars <- c("Sex", "ageOfDeath", "RIN2", "RIN", "rRNARate", "IntronicRate", "IntragenicRate", "IntergenicRate")
 random_effect <- c("Institution")
@@
 synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store variancePartition
+saveRDS(sortCols(violin_plot), file = 'CMC_ACC_Sex_variancePartition.rds')
+ADJ_OBJ = File('CMC_ACC_Sex_variancePartition.rds',
+               name = 'Variance Partition result', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store voom object
+saveRDS(VOOM.GENE_EXPRESSION, file = 'CMC_ACC_Sex_voom.rds')
+ADJ_OBJ = File('CMC_ACC_Sex_voom.rds',
+               name = 'voom object', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store contrasts
+saveRDS(L, file = 'CMC_ACC_Sex_contrasts.rds')
+ADJ_OBJ = File('CMC_ACC_Sex_contrasts.rds',
+               name = 'contrast matrix', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store contrasts
+saveRDS(formula, file = 'CMC_ACC_Sex_formula.rds')
+ADJ_OBJ = File('CMC_ACC_Sex_formula.rds',
+               name = 'formula', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
 ```

### 2020-08-19 individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_DLPFC_DE_Diagnosis.Rmd
Add all

--- 2020-08-19 individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_Diagnosis.Rmd
@@
 ---
-title: "Differential Expression - HBCC - DLPFC - Dx"
+title: "Differential Expression - MSSM-Penn-Pitt - DLPFC - Dx"
 author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_Diagnosis.Rmd",
-                                 parentId ="syn22050019",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_DLPFC_DE_Diagnosis.Rmd",
+                                 parentId ="syn22043922",
                                  folderName = "Differential Expression - Dx")
@@
 
-```{r synapse.parameters, include=FALSE, cache=TRUE}
-parentId = 'syn22050019';
+```{r synapse.parameters, include=FALSE, cache=FALSE}
+parentId = 'syn22043922';
 activityName = 'Differential Expression';
@@
 
-thisFileName <- 'HBCC_DLPFC_DE.Rmd'
+thisFileName <- 'MSSM-Penn-Pitt_DLPFC_DE.Rmd'
 
@@
 # Download CQN counts
-CQN_ID = "syn22082460"
+CQN_ID = "syn22045741"
 ALL_USED_IDs = c(CQN_ID)
@@
 # Download filtered counts
-COUNTS_ID = "syn22082459"
+COUNTS_ID = "syn22045729"
 ALL_USED_IDs = c(ALL_USED_IDs, COUNTS_ID)
@@
 # Get covariates
-COVAR_ID = "syn22082458"
+COVAR_ID = "syn22045728"
 ALL_USED_IDs = c(ALL_USED_IDs, COVAR_ID)
-COVARIATES = downloadFile(COVAR_ID) %>% 
-  tibble::column_to_rownames(var = "SampleID")
+COVARIATES = downloadFile(COVAR_ID)
+
+# dropped individual id 
+MD_ID = "syn16816488"
+ALL_USED_IDs = c(ALL_USED_IDs, COUNTS_ID)
+METADATA = downloadFile(MD_ID) %>% 
+  dplyr::select(Individual_ID, Sample_RNA_ID)
+
+COVARIATES = right_join(METADATA, COVARIATES, by = c("Sample_RNA_ID" = "SampleID")) %>% 
+  tibble::column_to_rownames(var = "Sample_RNA_ID")
 
 # Set variable class
-FactorCovariates <- c("Individual_ID", "Reported_Gender", "LibraryBatch", "Dx", "RibozeroBatch", "FlowcellBatch")
+FactorCovariates <- c("Individual_ID", "Institution", "Reported_Gender", "LibraryBatch", "Dx", "RibozeroBatch", "FlowcellBatch")
 COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
@@
 # Set covariates to adjust for from model selection
-adjust.covars <- c("Dx", "ageOfDeath", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate")
-random_effect <- c("Reported_Gender")
+adjust.covars <- c("Dx", "ageOfDeath", "RIN2", "RIN", "IntronicRate", "IntragenicRate")
+random_effect <- c("Institution", "Reported_Gender")
 ```
@@
 p <- ggplot(plotdata, aes(x=PC1, y=PC2))
-p <- p + geom_point(aes(shape=Tissue, size=ageOfDeath))
+p <- p + geom_point(aes(color=Institution, shape=Tissue, size=ageOfDeath))
 p <- p + theme_bw() + theme(legend.position="right") + facet_grid(Tissue~., scales = 'free_y')
@@
                                    coefficient = c("DxSCZ", "DxOther"))
-
 # Manually set comparison of SCZ-Other to Control
-L = cbind(L1, L2, L3, data.frame(L4 = c(1, -0.5, -0.5, 0, 0, 0, 0, 0, 0)))
+L = cbind(L1, L2, L3, data.frame(L4 = c(1, -0.5, -0.5, 0, 0, 0, 0)))
 

### 2020-08-19 individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_DLPFC_DE_Sex.Rmd
Add all

--- 2020-07-31 individual-cohorts/MSSM-Penn-Pitt_DLPFC_DE.Rmd
@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/MSSM-Penn-Pitt_DLPFC_DE.Rmd",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_DLPFC_DE_Sex.Rmd",
                                  parentId ="syn22043922",

### 2020-08-19 individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_Diagnosis.Rmd
fix cpm

@@
 # Fit linear model using new weights and new design
-VOOM.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION
+# VOOM.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION
 ADJUSTED.FIT = variancePartition::dream(exprObj = VOOM.GENE_EXPRESSION,

### 2020-08-19 individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_Sex.Rmd
fix cpm

@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/HBCC_DLPFC_DE.Rmd",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_Sex.Rmd",
                                  parentId ="syn22050019",

### 2020-08-19 individual-cohorts/split_by_analysis/HBCC_STG_DE_Diagnosis.Rmd
fix cpm

--- 2020-05-21 individual-cohorts/HBCC_STG_DE.Rmd
@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/HBCC_STG_DE.Rmd",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/split_by_analysis/HBCC_STG_DE_Diagnosis.Rmd",
                                  parentId ="syn22086820",
@@
 # Fit linear model using new weights and new design
-VOOM.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION
+# VOOM.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION
 ADJUSTED.FIT = variancePartition::dream(exprObj = VOOM.GENE_EXPRESSION$E,
@@
 synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store variancePartition
+saveRDS(sortCols(violin_plot), file = 'CMC_STG_Dx_variancePartition.rds')
+ADJ_OBJ = File('CMC_STG_Dx_variancePartition.rds',
+               name = 'Variance Partition result', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store voom object
+saveRDS(VOOM.GENE_EXPRESSION, file = 'CMC_STG_Dx_voom.rds')
+ADJ_OBJ = File('CMC_STG_Dx_voom.rds',
+               name = 'voom object', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store contrasts
+saveRDS(L, file = 'CMC_STG_Dx_contrasts.rds')
+ADJ_OBJ = File('CMC_STG_Dx_contrasts.rds',
+               name = 'contrast matrix', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store contrasts
+saveRDS(formula, file = 'CMC_STG_Dx_formula.rds')
+ADJ_OBJ = File('CMC_STG_Dx_contrasts.rds',
+               name = 'formula', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
 ```

### 2020-08-19 individual-cohorts/split_by_analysis/HBCC_dsACC_DE_Diagnosis.Rmd
fix cpm

--- 2020-05-20 individual-cohorts/HBCC_dsACC_DE.Rmd
@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/HBCC_dsACC_DE.Rmd",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/split_by_analysis/HBCC_dsACC_DE_Diagnosis.Rmd",
                                  parentId ="syn22084347",
@@
 synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store variancePartition
+saveRDS(sortCols(violin_plot), file = 'CMC_dsACC_Dx_variancePartition.rds')
+ADJ_OBJ = File('CMC_dsACC_Dx_variancePartition.rds',
+               name = 'Variance Partition result', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store voom object
+saveRDS(VOOM.GENE_EXPRESSION, file = 'CMC_dsACC_Dx_voom.rds')
+ADJ_OBJ = File('CMC_dsACC_Dx_voom.rds',
+               name = 'voom object', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store contrasts
+saveRDS(L, file = 'CMC_dsACC_Dx_contrasts.rds')
+ADJ_OBJ = File('CMC_dsACC_Dx_contrasts.rds',
+               name = 'contrast matrix', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store contrasts
+saveRDS(formula, file = 'CMC_dsACC_Dx_formula.rds')
+ADJ_OBJ = File('CMC_dsACC_Dx_formula.rds',
+               name = 'formula', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
 ```

### 2020-08-19 individual-cohorts/split_by_analysis/HBCC_sgACC_DE_Diagnosis.Rmd
rename
fix cpm

--- 2020-08-19 individual-cohorts/split_by_analysis/HBCC_sgACC_DE_Dx.Rmd
@@
 ```
-
-```

### 2020-08-19 individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_Diagnosis.Rmd
revert

--- 2020-08-19 individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_Diagnosis.Rmd
@@
 ```{r knit2synapse, eval=FALSE}
+render::rmarkdown("individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_Diagnosis.Rmd")
+
 library(synapser)
@@
                                  parentId ="syn22050019",
-                                 folderName = "Differential Expression - Dx")
+                                 folderName = "Differential Expression - Dx",
+                                 knitmd=FALSE)
 ```

### 2020-08-19 individual-cohorts/split_by_analysis/HBCC_STG_DE_Diagnosis.Rmd
revert

@@
 # Fit linear model using new weights and new design
-# VOOM.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION
+VOOM.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION
 ADJUSTED.FIT = variancePartition::dream(exprObj = VOOM.GENE_EXPRESSION$E,

### 2020-08-19 individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_Diagnosis.Rmd
revert

@@
 ```{r knit2synapse, eval=FALSE}
-render::rmarkdown("individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_Diagnosis.Rmd")
+ml git pandoc
+rmarkdown::render("individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_Diagnosis.Rmd")
 
@@
 
-thisFileName <- 'HBCC_DLPFC_DE.Rmd'
+thisFileName <- 'HBCC_DLPFC_DE_Diagnosis.Rmd'
 
@@
 thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
-thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/',thisFileName))
+thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/split_by_analysis/',thisFileName))
 ```

### 2020-08-19 individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_Sex.Rmd
revert

@@
 ```{r knit2synapse, eval=FALSE}
+ml git pandoc
+rmarkdown::render("individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_Sex.Rmd")
+
 library(synapser)
@@
 
-thisFileName <- 'HBCC_DLPFC_DE.Rmd'
+thisFileName <- 'HBCC_DLPFC_DE_Sex.Rmd'
 

### 2020-08-20 testing/test_CQN_HBCC.Rmd
initial commit

---
title: "Differential Expression - testing CQN"
author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
date: "`r date()`"
output: html_document
---

```{r knit2synapse, eval=FALSE}
ml git pandoc
rmarkdown::render("test_CQN_HBCC.Rmd")

```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## Load required libraries
suppressPackageStartupMessages({
library(CovariateAnalysis) # get the package from devtools::install_github('th1vairam/CovariateAnalysis@dev')
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)
library(R.utils)
library(ggplot2)
library(limma)
library(psych)
library(lme4)
library(biomaRt)
library(variancePartition)
library(edgeR)
library(cowplot)
library(synapser)
library(knitr)
library(githubr) # get the package from devtools::install_github('brian-bot/githubr')
})
options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = FALSE)
```

```{r coef}
# Specify which variable to test
coef = "DxSCZ"
# coef = "Reported_GenderMale"

# from HBCC_DLPFC_FEATURECOUNTS_GRCh38_CQN.Rmd
 # save(COVARIATES, NEW.COUNTS, CQN.GENE_EXPRESSION, GENE.GC.CONT, gene_lengths, file="save.RDATA")
load("save.RDATA")
```

# Standard normalization - TMM
```{r test.main, fig.height=3, fig.width=9}
# Standard normalization
dge = DGEList( counts = NEW.COUNTS )
dge = calcNormFactors(dge)
geneExpr = cpm(dge, log=TRUE)

# standard design
form = ~ Dx + scale(RIN) + ageOfDeath + Reported_Gender + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate)
design = model.matrix(form, COVARIATES)

# standard voom and lmFit
vobj = voom(dge, design)
fit = lmFit(vobj, design)
fit = eBayes(fit)

tab = topTable(fit, coef=coef, number=Inf)

# plot exression vs logFC (threshold at 1 for sex)
fig1 = ggplot(tab, aes(AveExpr, pmin(1, pmax(-1,logFC)))) + geom_point() + theme_bw() + geom_smooth(method="loess", se=FALSE) + ggtitle("NEW.COUNTS") + ylab("logFC") + theme(aspect.ratio=1)

# plot vs GC content and gene length
df = merge(tab, GENE.GC.CONT, by="row.names")
df = merge(df, gene_lengths, by.x = "Row.names", by.y="row.names")

fig1a = ggplot(df, aes(percentage_gene_gc_content, pmin(1, pmax(-1,logFC)))) + geom_point() + theme_bw() + geom_smooth(method="loess", se=FALSE) + ggtitle("NEW.COUNTS") + ylab("logFC")+ theme(aspect.ratio=1)

fig1b = ggplot(df, aes(log10(Length), pmin(1, pmax(-1,logFC)))) + geom_point() + theme_bw() + geom_smooth(method="loess", se=FALSE) + ggtitle("NEW.COUNTS") + ylab("logFC") + theme(aspect.ratio=1)

plot_grid(fig1, fig1a, fig1b, ncol=3)
```

# Current CQN (problematic)
```{r current.CQN, fig.height=3, fig.width=9}
geneExpr = DGEList(NEW.COUNTS)
geneExpr = edgeR::calcNormFactors(geneExpr)

form = ~ Dx + scale(RIN) + ageOfDeath + Reported_Gender + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate)
design = model.matrix(form, COVARIATES)

vobj_cqn = voom(geneExpr, design)

vobj_cqn$E = CQN.GENE_EXPRESSION$E

fit_cqn = lmFit(vobj_cqn, design)
fit_cqn = eBayes(fit_cqn)

tab_cqn = topTable(fit_cqn, coef=coef, number=Inf)

fig2 = ggplot(tab_cqn, aes(AveExpr, pmin(1, pmax(-1,logFC)))) + geom_point() + theme_bw() + geom_smooth(method="loess", se=FALSE) + ggtitle("CQN two step") + ylab("logFC") + theme(aspect.ratio=1)

df = merge(tab_cqn, GENE.GC.CONT, by="row.names")
df = merge(df, gene_lengths, by.x = "Row.names", by.y="row.names")

fig2a = ggplot(df, aes(percentage_gene_gc_content, pmin(1, pmax(-1,logFC)))) + geom_point() + theme_bw() + geom_smooth(method="loess", se=FALSE)  + ggtitle("CQN two step") + ylab("logFC") + theme(aspect.ratio=1)

fig2b = ggplot(df, aes(log10(Length), pmin(1, pmax(-1,logFC)))) + geom_point() + theme_bw() + geom_smooth(method="loess", se=FALSE) + ggtitle("CQN two step") + ylab("logFC") + theme(aspect.ratio=1)

plot_grid(fig2, fig2a, fig2, ncol=3)
```

# Correct offset CQN
CQN computes an offset that can be used the get the corrected log2 CPM corrected for bias caused by gene length and GC content.  While adding the offset to the counts is fine for computing the corrected magnitude, this affects the mean-variance trend used by voom and cant be used directly by voom.  edgeR can take this offset into account when estimating the mean variance trend, but standard voom cannot.   Ryan Thompson (at Sinai) wrote an extension of voom that correctly handles this offset: rctutils::voomWithOffset installable from https://github.com/DarwinAwardWinner/rctutils/. The lines A,B,C implement this offset using voom.

```{r cqn.offset, echo=TRUE}

dge_cqn = DGEList( counts = NEW.COUNTS )

# Extract default offset
offset_default = edgeR::expandAsMatrix(edgeR::getOffset(dge_cqn), dim(dge_cqn))

# Build custom offset from default offset and the CQN offset
dge_cqn$offset = offset_default - CQN.GENE_EXPRESSION$offset

# Run rctutils::voomWithOffset
vobj_cqn_offset = rctutils::voomWithOffset(dge_cqn, design)
```

```{r continue, , fig.height=3, fig.width=9}

form = ~ Dx + scale(RIN) + ageOfDeath + Reported_Gender + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate)
design = model.matrix(form, COVARIATES)

vobj_cqn_offset = voom(dge_cqn, design)
fit_cqn_offset = lmFit(vobj_cqn_offset, design)
fit_cqn_offset = eBayes(fit_cqn_offset)

tab_cqn_offset = topTable(fit_cqn_offset, coef=coef, number=Inf)

fig3 = ggplot(tab_cqn_offset, aes(AveExpr, pmin(1, pmax(-1,logFC)))) + geom_point() + theme_bw() + geom_smooth(method="loess", se=FALSE) + ggtitle("NEW.COUNTS + CQN (offset)") + ylab("logFC") + theme(aspect.ratio=1)

df = merge(tab_cqn_offset, GENE.GC.CONT, by="row.names")
df = merge(df, gene_lengths, by.x = "Row.names", by.y="row.names")

fig2a = ggplot(df, aes(percentage_gene_gc_content, pmin(1, pmax(-1,logFC)))) + geom_point() + theme_bw() + geom_smooth(method="loess", se=FALSE) + ggtitle("NEW.COUNTS + CQN (offset)") + ylab("logFC")  + theme(aspect.ratio=1)

fig2b = ggplot(df, aes(log10(Length), pmin(1, pmax(-1,logFC)))) + geom_point() + theme_bw() + geom_smooth(method="loess", se=FALSE) + ggtitle("NEW.COUNTS + CQN (offset)") + ylab("logFC") + theme(aspect.ratio=1)

plot_grid(fig2, fig2a, fig2b, ncol=3)
```

# Compare methods
```{r compare.methods}

df = merge(tab, tab_cqn, by="row.names")

ggplot(df, aes(t.x, t.y)) + geom_point() + theme_bw() + geom_smooth(method="loess", se=FALSE) + ggtitle("Compare t-statistics") + xlab("NEW.COUNTS") + ylab("CQN (problematic)") + theme(aspect.ratio=1)
with(df, cor.test(t.x, t.y))

df = merge(tab, tab_cqn_offset, by="row.names")

ggplot(df, aes(t.x, t.y)) + geom_point() + theme_bw() + geom_smooth(method="loess", se=FALSE) + xlab("NEW.COUNTS") + ylab("CQN  offset") + theme(aspect.ratio=1)
with(df, cor.test(t.x, t.y))

```

### 2020-08-21 deconvolution/fans4/molec_profiling.Rmd
Initial commit

---
title: "RNA-seq from 4 cell populations"
subtitle: ''
author: "Developed by [Gabriel Hoffman](http://gabrielhoffman.github.io/)"
date: "Run on `r Sys.time()`"
documentclass: article
output: 
  html_document:
  toc: true
  smart: false
---

<!---

cd /hpc/users/hoffmg01/work/cmc_analysis/src
# rm -rf molec_profiling_cache/
ml git pandoc
git pull
R

system("git pull")
rmarkdown::render("molec_profiling.Rmd", output_dir='./', intermediates_dir='./')

# http
# https://hoffmg01.u.hpc.mssm.edu/molec_profiling.html

--->

--->

```{r knit2synapse, eval=FALSE}
library(synapser)
library(knit2synapse)
synLogin()
knit2synapse::createAndKnitToFolderEntity(file = "deconvolution/fans4/molec_profiling.Rmd",
                                          parentId ="syn22231629",
                                          folderName = 'FANS4 processing and testing',
                                          overwrite=TRUE, 
                                          knitmd=TRUE)
```

```{r load.packages, echo=FALSE, message=FALSE, results='hide', cache=FALSE}
suppressPackageStartupMessages({
library(variancePartition)
library(limma)
library(edgeR)
library(mvIC)
library(reshape2)
library(qvalue)
library(data.table)
library(BiocParallel)
library(knitr)
library(org.Hs.eg.db)
library(kableExtra)
library(ggplot2)
library(cowplot)
library(scran)
library(SummarizedExperiment)
library(gridExtra)
library(openxlsx)
library(synapser)
synLogin()
library(EnsDb.Hsapiens.v86)
detach("package:EnsDb.Hsapiens.v86", unload=TRUE)
library(EnsDb.Hsapiens.v86)
ensdb = EnsDb.Hsapiens.v86
})

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE,
  cache.lazy = FALSE,  dev = c("png", "pdf"), 
  fig.width=7, fig.height=7)

options(markdown.HTML.stylesheet = 'css/custom.css')
```

```{r synapse.parameters, include=FALSE, cache=TRUE, eval=TRUE}
thisFileName <- 'molec_profiling.Rmd'
# thisFile = thisFileName
# Github link
thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('deconvolution/fans4/',thisFileName))
```

### Reading RNA-seq data
```{r read.data, message=TRUE}
Ids = c('syn22321012', 'syn22321013', 'syn22321014', 'syn22321016')
# file = "/sc/arion/projects/CommonMind/roussp01a/MOLECULAR_PROFILING/rnaseq/step1-merged/featureCount/featureCount.gene.tsv.gz"
file = synGet('syn22321012')$path
df_featureCounts_gene = fread( file )

geneInfo = df_featureCounts_gene[,1:6]
geneInfo$EnsId = gsub("^(.*)(\\.\\S+)$", "\\1", geneInfo$Geneid)

geneCounts = df_featureCounts_gene[,7:ncol(df_featureCounts_gene)]
geneCounts = as.data.frame(geneCounts)
rownames(geneCounts) = geneInfo$Geneid

# identify samples with improper names
idx = grep("Undetermined", colnames(geneCounts))
message("Dropping samples with improper names:\n",  paste(colnames(geneCounts)[idx], collapse="\n"))
geneCounts = geneCounts[,-idx]

idx = which(colSums(geneCounts) < 1e5)
message("Dropping samples with zero read counts:\n",  paste(colnames(geneCounts)[idx], collapse="\n"))
geneCounts = geneCounts[,-idx]

# remove genes with duplicated EnsId
exclude = duplicated(geneInfo$EnsId)
geneInfo = geneInfo[!exclude,]
geneCounts = geneCounts[!exclude,]
rownames(geneCounts) = gsub("^(.*)(\\.\\S+)$", "\\1", rownames(geneCounts))

# convert to HUGO names
df_hugo = select(org.Hs.eg.db, geneInfo$EnsId, "SYMBOL", "ENSEMBL")
geneInfo = merge(geneInfo, df_hugo, by.x='EnsId', by.y="ENSEMBL")

```

```{r metadata}
# file = "/sc/arion/projects/CommonMind/roussp01a/MOLECULAR_PROFILING/meta-files/MolecularProfiling_MSSM_120samples_05.24.17-4.xlsx"
file = synGet('syn22321013')$path
# focused on this experiment
metadata1 = read.xlsx( file )

# from Synapse
# file = "/sc/arion/projects/CommonMind/roussp01a/MOLECULAR_PROFILING/meta-files/CMC_METADATA.csv"
file = synGet('syn22321014')$path
metadataCMC = fread( file )
```

### RNA-seq QC metrics
```{r QC, fig.width=18, fig.height=74}
# file = '/sc/arion/projects/CommonMind/roussp01a/MOLECULAR_PROFILING/rnaseq/step1-merged/QCmetrics/qc_metrics.RDS'
file = synGet('syn22321016')$path
qc_metrics.in = t(readRDS( file ))
colnames(qc_metrics.in) = qc_metrics.in[1,]
qc_metrics.in = as.data.frame(qc_metrics.in[-c(1:2),], stringsAsFactors=FALSE)
qc_metrics = apply(qc_metrics.in, 2, as.numeric)
rownames(qc_metrics) = gsub('-', '.', rownames(qc_metrics.in))

# extract information from sample names
info = data.frame( Name = colnames(geneCounts), stringsAsFactors=FALSE)
rownames(info) = info$Name
info$DissectionID = sapply(strsplit( info$Name, '\\.'), function(x) x[1])
info$CellType = sapply(strsplit( info$Name, '\\.'), function(x) x[2])
info$Assay = sapply(strsplit( info$Name, '\\.'), function(x) x[3])

# set color codes for each cell type
colorCodes = c(
  GABA = '#66A61E',
  GLU = '#E6AB02',
  Olig = '#E7298A',
  MgAs = '#7570B3')

# set factor levels to be the same as the order of the colors
# this makes plotting easier
info$CellType = factor(info$CellType,names(colorCodes))

# add information about each sample
info = merge(info, metadata1, by.x="DissectionID", by.y='Dissection.ID')
info = merge(info, qc_metrics, by.x="Name", by.y='row.names')
rownames(info) = info$Name

# info = merge(info, metadataCMC, by.x='Individual.ID', by.y="Individual ID")
#  info[1:4, 1:20]

# set order to be the same as geneCounts
info = info[match(colnames(geneCounts), info$Name),]

# exclude columns with no variance
index = apply(qc_metrics, 2, function(x) var(x, na.rm=TRUE)) > 0
include = which(index==TRUE)
include = include[! names(include) %in% "PF_BASES"]

df = reshape2::melt(info[,c('DissectionID', 'CellType', colnames(qc_metrics)[include])], id.vars=c('DissectionID', 'CellType'))

ggplot(df, aes( x=CellType, y=value, fill=CellType)) + geom_boxplot() + facet_wrap(~variable, scales="free_y", ncol=5) + theme_bw(16) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + scale_fill_manual("Cell Type", values=colorCodes) 
```

### Study Summary
```{r}
kable(table(info$Dx, info$CellType)) %>% kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

```{r}
kable(table(info$Individual.ID)) %>% kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

```{r preprocess}
isexpr = rowSums(cpm(geneCounts)>1) >= 0.1*ncol(geneCounts)
table(isexpr)

dge = DGEList( geneCounts[isexpr,] )
dge = calcNormFactors(dge)
```

```{r TMM}
df = merge(dge$samples, info, by="row.names")

ggplot(df, aes(CellType, norm.factors, fill=CellType)) + geom_boxplot() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=colorCodes) + ylab("TMM factor")
```

```{r expression.density}

geneExpr = cpm(dge, log=TRUE)

df = reshape2::melt(geneExpr)
df = merge(df, info[,'CellType',drop=FALSE], by.x = "Var2", by.y="row.names")

ggplot(df, aes(value, color=CellType)) + geom_density() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=colorCodes) + xlab(bquote(log[2]~CPM))

ggplot(df, aes(value, color=Var2)) + geom_density() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + facet_wrap(~CellType, ncol=4) + xlab(bquote(log[2]~CPM))
```

### PCA
```{r PCA}

geneExpr = cpm(dge, log=TRUE)

dcmp = prcomp( t(geneExpr), scale=TRUE)

df_pca = data.frame(dcmp$x)

df_pca = merge(df_pca, info, by.x="row.names", by.y="Name")

varFrac = dcmp$sdev^2 / sum(dcmp$sdev^2)

xlab = paste0('PC1 (', format(varFrac[1]*100, digits=3), '%)')
ylab = paste0('PC2 (', format(varFrac[2]*100, digits=3), '%)')
ggplot(df_pca, aes(PC1, PC2, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=colorCodes) + xlab(xlab) + ylab(ylab)

xlab = paste0('PC2 (', format(varFrac[2]*100, digits=3), '%)')
ylab = paste0('PC3 (', format(varFrac[3]*100, digits=3), '%)')
ggplot(df_pca, aes(PC2, PC3, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=colorCodes) + xlab(xlab) + ylab(ylab)
```

### Sex check
```{r sex.check}

XIST = 'ENSG00000229807'
UTY = 'ENSG00000183878'

df = data.frame(  Name  = colnames(geneExpr), 
                  XIST  = geneExpr[XIST,],
                  UTY   = geneExpr[UTY,])
df = merge( df, info[,c('Name', 'Gender', 'Individual.ID')], by="Name" )
df$Gender = factor(df$Gender, c("Male", 'Female'))

ggplot(df, aes(XIST, UTY, color=Gender)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=c("blue", "red")) + ggtitle("Sex check")
```

Samples with wrong sex
```{r sex.off}
kable(df[df$Gender == 'Female' & df$UTY > 5,], digits=2, row.names=FALSE) %>% kable_styling(bootstrap_options = "striped", full_width = FALSE)  
```

### Esimate precision weights
```{r voom}
form = ~ (1|CellType) + (1|Individual.ID) + (1|Gender)
vobj = voomWithDreamWeights(dge, form, info, plot=TRUE, save.plot=TRUE, BPPARAM=SnowParam(6, progressbar=TRUE))
```

### mvIC
Need to get proper covariates here

```{r mvIC}
colnames(info) = gsub(' ', '_', colnames(info))
colnames(info) = gsub(':', '', colnames(info))
colnames(info) = gsub('\\(', '', colnames(info))
colnames(info) = gsub(')', '', colnames(info))
colnames(info) = gsub('/', '', colnames(info))
colnames(info) = gsub('-', '_', colnames(info))
colnames(info) = gsub(',', '', colnames(info))

include = apply(info, 2, function(x) all(!is.na(x)) && var(x, na.rm=TRUE) >0 )

variables = colnames(info)[colnames(info) %in% names(include[include])]

mvSearch = mvForwardStepwise( vobj, ~ CellType, info, variables)
```

### variancePartition
```{r vp}
form = ~ (1|CellType) + (1|Dx) + (1|CellType:Dx) + (1|Individual.ID) + (1|Gender)
vp = fitExtractVarPartModel( geneExpr, form, info, BPPARAM=SnowParam(6, progressbar=TRUE))

plotVarPart( sortCols(vp) )
```

## Differential expression analysis
### Cell type
```{r DE_process}
form = ~ 0 + CellType + (1|Individual.ID) + (1|Gender)

# Prepare all pairwise contrasts
index = combn(nlevels(info$CellType),2)

L = lapply( 1:ncol(index), function(j){

  CT1 = levels(info$CellType)[index[1,j]]
  CT2 = levels(info$CellType)[index[2,j]]

  L = getContrast( vobj, form, info, paste0('CellType', c(CT1, CT2)))

  L = as.matrix(L)
  colnames(L) = paste0(CT1, '.', CT2)

  L
})
L = do.call(cbind, L)

L = cbind(L, data.frame(GABAGLU.Other = c(1, 1, -1, -1),
                    MgAs.Other = c(-1/3, -1/3, -1/3, 1),
                    Olig.Other = c(-1/3, -1/3, 1, -1/3) ))

fitBasic = dream( vobj, form, info, L=L, BPPARAM=SnowParam(6, progressbar=TRUE))
```

```{r plotContrasts}
plotContrasts(round(L,2)) + theme(axis.text.x=element_text(angle=45, hjust=1))   
```

```{r DE.analysis}

index = combn(nlevels(info$CellType),2)

df_deCount = lapply( 1:ncol(index), function(j){

    CT1 = levels(info$CellType)[index[1,j]]
    CT2 = levels(info$CellType)[index[2,j]]

    key = paste0(CT1, '.', CT2)

    tabl = topTable(fitBasic, coef=key, number=Inf)

    data.frame(CT1, CT2, key, count = sum(tabl$adj.P.Val < 0.05), pi1 = 1 - qvalue(tabl$P.Value)$pi0)
})
df_deCount = do.call(rbind, df_deCount)

ggplot(df_deCount, aes(CT1, CT2, fill=count)) + geom_tile() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + scale_fill_gradient(low="white", high="red", limits=c(0, 30000)) + geom_text(aes(label=format(count, big.mark=','))) + xlab('') + ylab('') + ggtitle("Count differentiall expression genes")

ggplot(df_deCount, aes(CT1, CT2, fill=pi1)) + geom_tile() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + scale_fill_gradient(name = quote(pi[1]), low="white", high="red", limits=c(0, 1)) + geom_text(aes(label=format(pi1, digits=3))) + xlab('') + ylab('') + ggtitle(bquote(pi[1]))
```

```{r DE2}

index = combn(nlevels(info$CellType),2)

df = lapply( 1:ncol(index), function(j){

    CT1 = levels(info$CellType)[index[1,j]]
    CT2 = levels(info$CellType)[index[2,j]]

    key = paste0(CT1, '.', CT2)

    tabl = topTable(fitBasic, coef=key, number=Inf)

    data.frame(CT1, CT2, key, tabl)
})
df = do.call(rbind, df)

ggplot(df, aes(logFC, ..count..)) + geom_density(alpha = .2) + facet_wrap(~key) + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) 
```

## Plots of top differentially expressed genes
```{r DE.marker_genes, fig.width=15, fig.height=30}

index = combn(nlevels(info$CellType),2)

ensIDs = sapply( 1:ncol(index), function(j){

    CT1 = levels(info$CellType)[index[1,j]]
    CT2 = levels(info$CellType)[index[2,j]]

    key = paste0(CT1, '.', CT2)

    tabl = topTable(fitBasic, coef=key, number=Inf)

    tabl = tabl[order(tabl$t),]

    c(head(rownames(tabl), 10), tail(rownames(tabl), 10))
})

ensIDs = unique(c(ensIDs, 'ENSG00000110693', 'ENSG00000167281'))

df = reshape2::melt(geneExpr[ensIDs,])
df = merge(df, info[,1:3], by.x="Var2", by.y="row.names")

df_hugo = select(org.Hs.eg.db, unique(as.character(df$Var1)), "SYMBOL", "ENSEMBL")
df = merge(df, df_hugo, by.x="Var1", by.y="ENSEMBL")
df = df[!is.na(df$SYMBOL),]

ggplot(df, aes(CellType, value, fill=CellType)) + geom_boxplot() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position = "none") + scale_fill_manual("Cell Type", values=colorCodes) + facet_wrap(~SYMBOL, ncol=5) + ylab(bquote(log[2]~CPM))
```

## Analysis using just pairs of samples
```{r pairwise.analysis}

index = combn(nlevels(info$CellType),2)

# for all pairs of cell types
tabList = lapply( 1:ncol(index), function(j){

  CT1 = levels(info$CellType)[index[1,j]]
  CT2 = levels(info$CellType)[index[2,j]]
  key = paste0(CT1, '.', CT2)

  idx = which(info$CellType %in% c(CT1, CT2))

  isexpr = rowSums(cpm(geneCounts[,idx])>1) >= 0.3*ncol(geneCounts[,idx])
  table(isexpr)
  infoSub = droplevels(info[idx,])

  dge = DGEList( geneCounts[isexpr,idx] )
  dge = calcNormFactors(dge)

  form = ~ CellType + Gender
  vobj = voomWithDreamWeights(dge, form, infoSub, plot=FALSE, save.plot=TRUE, BPPARAM=SnowParam(6, progressbar=TRUE))

  form = ~ CellType + Gender
  fit = dream( vobj, form, infoSub, BPPARAM=SnowParam(6, progressbar=TRUE))
  fit = eBayes(fit)

  coef = paste0('CellType',levels(infoSub$CellType)[2])

  topTable(fit, coef=coef, number=Inf)
})

names(tabList) = apply(matrix(levels(info$CellType)[index], nrow=2), 2, function(x) paste0(x, collapse='.'))

df_de_summary = lapply( names(tabList), function(key){
  tab = data.table(tabList[[key]])

  data.frame( Comparison  = key, 
              nGenes      = nrow(tab), 
              nDE         = sum(tab$adj.P.Val < 0.05),
              nDE_gt_1    =  nrow(tab[adj.P.Val < 0.05 & abs(logFC) > 1,]), 
              nDE_gt_2    =  nrow(tab[adj.P.Val < 0.05 & abs(logFC) > 2,]), 
              nDE_gt_3    =  nrow(tab[adj.P.Val < 0.05 & abs(logFC) > 3,]), 
              nDE_gt_4    =  nrow(tab[adj.P.Val < 0.05 & abs(logFC) > 4,]), 
              pi1         = 1-qvalue(tab$P.Value)$pi0)
  })
df_de_summary = do.call(rbind, df_de_summary)

kable(df_de_summary) %>% kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

```{r pairwise.density}

index = combn(nlevels(info$CellType),2)

# for all pairs of cell types
tabList = lapply( 1:ncol(index), function(j){

  CT1 = levels(info$CellType)[index[1,j]]
  CT2 = levels(info$CellType)[index[2,j]]
  key = paste0(CT1, '.', CT2)

  idx = which(info$CellType %in% c(CT1, CT2))

  isexpr = rowSums(cpm(geneCounts[,idx])>1) >= 0.3*ncol(geneCounts[,idx])
  table(isexpr)
  infoSub = droplevels(info[idx,])

  dge = DGEList( geneCounts[isexpr,idx] )
  dge = calcNormFactors(dge)

  form = ~ CellType + Gender
  vobj = voomWithDreamWeights(dge, form, infoSub, plot=FALSE, save.plot=TRUE, BPPARAM=SnowParam(6, progressbar=TRUE))

  form = ~ CellType + Gender
  fit = dream( vobj, form, infoSub, BPPARAM=SnowParam(6, progressbar=TRUE))
  fit = eBayes(fit)

  coef = paste0('CellType',levels(infoSub$CellType)[2])

  data.frame(key, topTable(fit, coef=coef, number=Inf))
})
df = do.call(rbind, tabList)

ggplot(df, aes(logFC)) + geom_density(alpha = .2) + facet_wrap(~key) + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) 
```

```{r compare.voom}

index = combn(nlevels(info$CellType),2)

# for all pairs of cell types
df_voomTrend = lapply( 1:ncol(index), function(j){

  CT1 = levels(info$CellType)[index[1,j]]
  CT2 = levels(info$CellType)[index[2,j]]
  key = paste0(CT1, '.', CT2)

  idx = which(info$CellType %in% c(CT1, CT2))

  isexpr = rowSums(cpm(geneCounts[,idx])>1) >= 0.3*ncol(geneCounts[,idx])
  table(isexpr)
  infoSub = droplevels(info[idx,])

  dge = DGEList( geneCounts[isexpr,idx] )
  dge = calcNormFactors(dge)

  form = ~ CellType + Gender
  vobj = voomWithDreamWeights(dge, form, infoSub, plot=FALSE, save.plot=TRUE, BPPARAM=SnowParam(6, progressbar=TRUE))

  data.frame( key, x = vobj$voom.line[['x']], y = vobj$voom.line[['y']])
})
df_voomTrend = do.call(rbind, df_voomTrend)

ggplot(df_voomTrend, aes(x, y, color=key)) + geom_line() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + ylab("sqrt( standard deviation)") + xlab("log2( count size + 0.5 )") + xlim(0, 22) + ylim(0.4, 1.4)
```

## Identify marker genes
```{r tabList.markerGenes}

tabList = lapply(colnames(L), function(coef){
  tab = topTable(fitBasic, coef=coef, number=Inf)
  # tab[which((tab$adj.P.Val < 0.05) & (abs(tab$logFC) > 5)),]
  tab
})
names(tabList) = colnames(L)

saveRDS(tabList, file='./Roussos_MSSM_MolProf_FANS_4_diffExpression.RDS')

system("ml python; synapse add --parentid syn22243354 Roussos_MSSM_MolProf_FANS_4_diffExpression.RDS")
```

## Plots of top differentially expressed genes
```{r markerGenes, fig.width=15, fig.height=30}

lapply( names(tabList), function(key){tab = tabList[[key]]; nrow(tab[tab$logFC > 6,]) })

boxplot_gene = function(ensIDs, tab, ngenes=5){
  df = reshape2::melt(geneExpr[ensIDs,,drop=FALSE])
  df = merge(df, info[,1:3], by.x="Var2", by.y="row.names")

  df_hugo = select(org.Hs.eg.db, unique(as.character(df$Var1)), "SYMBOL", "ENSEMBL")
  df = merge(df, df_hugo, by.x="Var1", by.y="ENSEMBL")
  df = df[!is.na(df$SYMBOL),]

  df = df[df$SYMBOL %in% df_hugo$SYMBOL[!is.na(df_hugo$SYMBOL)][1:ngenes], ]

  ggplot(df, aes(CellType, value, fill=CellType)) + geom_boxplot() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position = "none") + scale_fill_manual("Cell Type", values=colorCodes) + facet_wrap(~SYMBOL, nrow=1) + ylab(bquote(log[2]~CPM))
}

figList = lapply( names(tabList), function(key){

  tab = tabList[[key]]

  # only select up regulated genes
  tab = tab[tab$logFC > 6,]

  ensIDs = rownames(tab)

  boxplot_gene( ensIDs, tab, 5) + ggtitle(key)
  } )
plot_grid(plotlist = figList, ncol=1)

```

```{r FANS.markers}
ensIDs =  c('ENSG00000110693','ENSG00000167281', 'ENSG00000100146')
boxplot_gene(ensIDs, tab) + ggtitle("FANS markers")
```

# Compare to GTEx
```{r GTEX.voom}

file_SampleAttributesDS = "/sc/arion/projects/CommonMind/data/resources/GTEx/GTEx_Analysis_v8_Annotations_SampleAttributesDS.txt"
file_SubjectPhenotypesDS= "/sc/arion/projects/CommonMind/data/resources/GTEx/GTEx_Analysis_v8_Annotations_SubjectPhenotypesDS.txt"
file_counts = "/sc/arion/projects/CommonMind/data/resources/GTEx/GTEx_Analysis_2017-06-05_v8_RNASeQCv1.1.9_gene_reads.gct.gz"

gtex_sample = read.delim(file_SampleAttributesDS, header = TRUE, stringsAsFactors = FALSE)
# gtex_sample = gtex_sample[gtex_sample$SMTS == 'Brain',]

rnaseq_sample_id = fread(file_counts, header = TRUE, skip = 2, nrows = 1)
gtex_sample = gtex_sample[gtex_sample$SAMPID %in% colnames(rnaseq_sample_id),]

# read count
gtex_count <- fread(file_counts, header = TRUE, skip = 2, select = c('Name', unique(gtex_sample$SAMPID)))
gtex_count = as.data.frame(gtex_count, stringsAsFactors=FALSE)

gtex_count$Name = gsub("^(.*)(\\.\\S+)$", "\\1", gtex_count$Name )
gtex_count = gtex_count[!duplicated(gtex_count$Name), ]
rownames(gtex_count) = gtex_count$Name
gtex_count$Name = NULL
gtex_count = as.matrix(gtex_count)

# isexpr = rowSums(cpm(gtex_count>1)) >= 0.3*ncol(gtex_count)
# dge = DGEList( gtex_count[isexpr,] )
# dge = calcNormFactors(dge)
# gtex_log2cpm = cpm(dge, log=TRUE)

# df_voomTrend_GTEx = lapply( unique(gtex_sample$SMTSD), function(Tissue){

#   idx = which( gtex_sample$SMTSD == Tissue)

#   isexpr = rowSums(cpm(gtex_count[,idx])>1) >= 0.3*ncol(gtex_count[,idx])
#   table(isexpr)

#   dge = DGEList( gtex_count[isexpr,idx] )
#   dge = calcNormFactors(dge)

#   vobj = voomWithDreamWeights(dge, ~1, gtex_sample[idx,], plot=FALSE, save.plot=TRUE, BPPARAM=SnowParam(6, progressbar=TRUE))

#   data.frame( key = Tissue, x = vobj$voom.line[['x']], y = vobj$voom.line[['y']])
# })
# df_voomTrend_GTEx = do.call(rbind, df_voomTrend_GTEx)
```

```{r GTEX.voom.plot}
# ggplot(df_voomTrend_GTEx, aes(x, y, color=key)) + geom_line() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="bottom") + ylab("sqrt( standard deviation)") + xlab("log2( count size + 0.5 )") + ggtitle("GTEx brain regions") + xlim(2, 22) + ylim(0.45, 1.36)
```

```{r gtex.colors}
# https://genome.ucsc.edu/cgi-bin/hgTables?db=hg19&hgta_track=hgFixed&hgta_group=allTables&hgta_table=hgFixed.gtexTissue

gtexColors = read.csv('/sc/arion/projects/CommonMind/data/resources/GTEx/GTEx_Tissue_colors.csv', header=TRUE)
```

```{r CV.all.in.one, eval=TRUE}
library(scran)
# FANS4
isexpr = rowSums(edgeR::cpm(geneCounts)>1) >= 0.3*ncol(geneCounts)

se <- SummarizedExperiment( assays=list(counts = as.matrix(geneCounts[isexpr,])), colData = info)
res = modelGeneCV2( se, block=info$CellType )
fit = metadata(res)

df_fans4 = lapply( names(res$per.block), function(key){
  with( res$per.block[[key]], data.frame(key, mean, total, trend, ratio, FDR))
  })
df_fans4 = data.table(do.call(rbind, df_fans4))

df_fans4 = df_fans4[order(df_fans4$mean),]
ggplot(df_fans4, aes(log2(mean), log2(trend), color=key)) + geom_line() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + xlab(bquote(log[2]~Mean~expression)) + ylab(bquote(log[2]~CV~trend)) + scale_color_manual(values=colorCodes)+ xlim(0, 20) + ylim(-5, 5)

# GTEX
idx = which(gtex_sample$SMTSD %in% names(which(table(gtex_sample$SMTSD) > 30)))

isexpr = rowSums(edgeR::cpm(gtex_count[,idx])>1) >= 0.3*ncol(gtex_count[,idx])

se <- SummarizedExperiment( assays=list(counts = gtex_count[isexpr,idx]), colData = gtex_sample[idx,])
res = modelGeneCV2( se, block=gtex_sample$SMTSD[idx] )

df_gtex = lapply( names(res$per.block), function(key){
  with( res$per.block[[key]], data.frame(key, mean, total, trend, ratio, FDR))
  })
df_gtex = data.table(do.call(rbind, df_gtex))
df_gtex$key = factor(df_gtex$key)

i = match(levels(df_gtex$key), gtexColors$description)
colsGTEx = as.hexmode(gtexColors$color)[i[!is.na(i)]]

df_gtex = df_gtex[order(df_gtex$mean),]
df_gtex = df_gtex[df_gtex$mean > 0,]

ggplot(df_gtex, aes(log2(mean), log2(trend), color=key)) + geom_line() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + xlab(bquote(log[2]~Mean~expression)) + ylab(bquote(log[2]~CV~trend))+ scale_color_manual(values=colsGTEx) + xlim(0, 20) + ylim(-5, 5)

df_gtex$dataset = 'GTEx'
df_fans4$dataset = "FANS4"
df_combined = rbind(df_gtex, df_fans4)

ggplot(df_combined, aes(log2(mean), log2(trend), color=key, size=dataset)) + geom_line() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + xlab(bquote(log[2]~Mean~expression)) + ylab(bquote(log[2]~CV~trend)) + scale_size_manual(values=c(2, .1)) + scale_color_manual(values=c(rep("black", length(colsGTEx)), colorCodes)) + xlim(-8, 20) + ylim(-5, 3) 
```

```{r coef.variation, eval=TRUE}

CTarray = as.list(levels(info$CellType))
# CTarray[[5]] = c("GABA", "GLU")
# CTarray[[6]] = levels(info$CellType)

# different because only genes expressed in given cell type are used
res_cv_FANS = lapply(CTarray, function(CT){

  id = info$Name[which(info$CellType %in% CT)]

  isexpr = rowSums(edgeR::cpm(geneCounts[,id])>1) >= 0.1*ncol(geneCounts[,id])

  se <- SummarizedExperiment( assays=list(counts = as.matrix(geneCounts[isexpr,id])), colData = info[id,])
  res = modelGeneCV2( se ) # returns CV^2
  fit = metadata(res)

  geneTab = select(ensdb, keys=names(isexpr[isexpr]), columns=c("GENEID", "GENEBIOTYPE"), keytype="GENEID")

  list(df = data.frame( key = paste0(CT, collapse="."), res ),
       f = fit$trend,
       nExpressed = sum(isexpr),
       nExpressed.CDS = sum(geneTab$GENEBIOTYPE == 'protein_coding'))
})

x = 2^seq(-3, 15, length.out=500)
df_trend_FANS = lapply(res_cv_FANS, function(res){ 
  data.frame( Dataset="FANS4", key = res$df$key[1], mean = x, trend=res$f(x))
  })
df_trend_FANS = do.call(rbind, df_trend_FANS)

df_cv_FANS = lapply(res_cv_FANS, function(res){ 
  res$df
  })
df_cv_FANS = do.call(rbind, df_cv_FANS)

colorCodes['GABA.GLU'] = "orange"
colorCodes['GABA.GLU.Olig.MgAs'] = "dodgerblue"

# Take sqrt(CV^2)
ggplot(df_trend_FANS, aes(log2(mean), log2(sqrt(trend)), color=key, size=Dataset)) + geom_line() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position=c(.7, .9)) + xlab(bquote(log[2]~Mean~expression)) + ylab(bquote(log[2]~CV~trend)) + scale_size_manual(values=c(2, .1)) + scale_color_manual(values=colorCodes) + scale_x_continuous(expand=c(0.01, 0))

ggplot(df_cv_FANS[df_cv_FANS$key %in% names(colorCodes)[1:4],], aes(log2(sqrt(total)), color=key)) + geom_density(size=2) + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="right")  + scale_color_manual(values=colorCodes) + xlab(bquote(log[2]~CV))

ggplot(df_cv_FANS, aes(log2(sqrt(total)), color=key)) + geom_density(size=2) + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="right")  + scale_color_manual(values=colorCodes) + xlab(bquote(log[2]~CV))

#  GTEX
res_cv_GTEx = lapply( names(which(table(gtex_sample$SMTSD) > 30)), function(Tissue){

  id = gtex_sample$SAMPID[which( gtex_sample$SMTSD == Tissue)]

  isexpr = rowSums(edgeR::cpm(gtex_count[,id])>1) >= 0.1*ncol(gtex_count[,id])

  se <- SummarizedExperiment( assays=list(counts = gtex_count[isexpr,id]), colData = gtex_sample[id,])
  res = modelGeneCV2( se )
  fit = metadata(res)

  geneTab = select(ensdb, keys=names(isexpr[isexpr]), columns=c("GENEID", "GENEBIOTYPE"), keytype="GENEID")

  list(df = data.frame( key = Tissue, res ),
       f = fit$trend,
       nExpressed = sum(isexpr),
       nExpressed.CDS = sum(geneTab$GENEBIOTYPE == 'protein_coding'))
})

df_trend_GTEx = lapply(res_cv_GTEx, function(res){ 
  data.frame( Dataset="GTEx", key = res$df$key[1], mean = x, trend=res$f(x))
  })
df_trend_GTEx = do.call(rbind, df_trend_GTEx)
df_trend_GTEx$key = factor(df_trend_GTEx$key)

df_cv_GTEx = lapply(res_cv_GTEx, function(res){ 
  res$df
  })
df_cv_GTEx = do.call(rbind, df_cv_GTEx)
df_cv_GTEx$key = factor(df_cv_GTEx$key)

i = match(levels(df_trend_GTEx$key), gtexColors$description)
colsGTEx = as.hexmode(gtexColors$color)[i[!is.na(i)]]

ggplot(df_trend_GTEx, aes(log2(mean), log2(sqrt(trend)), color=key)) + geom_line() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + xlab(bquote(log[2]~Mean~expression)) + ylab(bquote(log[2]~CV~trend)) + scale_color_manual(values=colsGTEx) + scale_x_continuous(expand=c(.01,0))

ggplot(df_cv_GTEx, aes(log2(sqrt(total)), color=key)) + geom_density(size=2) + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + xlab(bquote(log[2]~CV~trend)) + scale_color_manual(values=colsGTEx)

# combine
df_combined = rbind(df_trend_GTEx, df_trend_FANS)

ggplot(df_combined, aes(log2(mean), log2(trend), color=key, size=Dataset)) + geom_line() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + xlab(bquote(log[2]~Mean~expression)) + ylab(bquote(log[2]~CV~trend)) + scale_size_manual(values=c(2, .2)) + scale_color_manual(values=c(rep("grey60", length(colsGTEx)), colorCodes)) + scale_x_continuous(expand=c(.01,0))

df_cv_GTEx$Dataset = "GTEx"
df_cv_GTEx$Type = "Real"
df_cv_FANS$Dataset = "FANS4"
df_cv_FANS$Type = "Real"
df_cv_FANS$Type[df_cv_FANS$key %in% names(colorCodes)[5:6]] = "Synthetic"

df = data.table(rbind(df_cv_GTEx, df_cv_FANS))

ggplot(df, aes(log2(total), color=key, size=Dataset, linetype=Type)) + geom_density(size=2) + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + xlab(bquote(log[2]~CV~trend)) + scale_color_manual(values=c(rep("grey", length(colsGTEx)), colorCodes)) + scale_size_manual(values=c(2, .2)) + scale_linetype_discrete(c("solid", "dashed"))

# number of expressed genes
 df[,sum(total > 0), by='key']

df2 = lapply(sort(unique(df$key)), function(KEY){
  data.frame(key = KEY, Dataset = df$Dataset[df$key == KEY][1],
    Type = df$Type[df$key == KEY][1],
    CV2=df[key==KEY,][order(mean,decreasing=TRUE)[1:15000],]$total)
  })
df2 = data.table(do.call(rbind, df2))
df2$key = factor(df2$key, (sort(unique(df2$key))))
levels(df2$key)

i = match(levels(df2$key), gtexColors$description)
colsGTEx = paste0("#", as.hexmode(gtexColors$color)[i[!is.na(i)]])
names(colsGTEx) = gtexColors$description[i[!is.na(i)]]

cols = c(colsGTEx, colorCodes)[levels(df2$key)]
cols[names(colsGTEx)] = "grey"

ggplot(df2, aes(log2(sqrt(CV2)), color=key, size=Dataset, linetype=Type)) + geom_density(size=2) + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + xlab(bquote(log[2]~CV~trend)) + scale_color_manual(values=cols) + scale_size_manual(values=c(2, .1)) + scale_linetype_discrete(c("solid", "dashed"))

df3 = df2[,median(sqrt(CV2)), by='key']
df3$key = factor(df3$key, rev(levels(df3$key)))

ggplot( df3[!is.na(df3$V1)], aes(key, V1, fill=key)) + geom_bar(stat='identity') + theme_bw(10) + theme(aspect.ratio=5, plot.title = element_text(hjust = 0.5), legend.position="none")  + scale_fill_manual(values=c(colsGTEx, colorCodes)[levels(df3$key)]) + coord_flip() + ylab(bquote(Median~CV)) + scale_y_continuous(expand=c(0,0)) + xlab('') + ylim(0, max(df3$V1, na.rm=TRUE)*1.05)

# Compare to IPS and LCL

# genes expressed

df_count_fans4 = lapply(res_cv_FANS, function(res){  
  data.frame( Dataset="FANS4", key = res$df$key[1], 
    "Expressed genes"=res$nExpressed, "Expressed coding genes"=res$nExpressed.CDS)
 })
df_count_fans4 = do.call(rbind, df_count_fans4)

df_count_gtex = lapply(res_cv_GTEx, function(res){  
  data.frame( Dataset="GTEx", key = res$df$key[1], 
    "Expressed genes"=res$nExpressed, "Expressed coding genes"=res$nExpressed.CDS)
 })
df_count_gtex = do.call(rbind, df_count_gtex)

df_count = rbind(df_count_gtex, df_count_fans4)
df_count$key = factor(df_count$key, rev(levels(df2$key)))

i = match(levels(df_count$key), gtexColors$description)
colsGTEx = paste0("#", as.hexmode(gtexColors$color)[i[!is.na(i)]])
names(colsGTEx) = gtexColors$description[i[!is.na(i)]]

df_melt = reshape2::melt(df_count, id.vars=c("Dataset", "key"))

ggplot( df_melt, aes(key, value, fill=key)) + geom_bar(stat='identity') + theme_bw(10) + theme(aspect.ratio=5, plot.title = element_text(hjust = 0.5), legend.position="none")  + scale_fill_manual(values=c(colsGTEx, colorCodes)[levels(df3$key)]) + coord_flip() + ylab("# genes expressed") + scale_y_continuous(expand=c(0,0)) + xlab('') + facet_wrap(~variable)

```

```{r plot.coef.variation, fig.width=12, fig.height=12, eval=FALSE}
# use colors here
# gtexColors and colorCodes

df = rbind(df_cv_GTEx, df_cv_FANS)
df$key = factor(df$key, c(rev(sort(unique(df_cv_FANS$key))), rev(sort(unique(df_cv_GTEx$key)))))

i = match(levels(df$key), gtexColors$description)
cols = as.hexmode(gtexColors$color)[i[!is.na(i)]]
cols = paste0('#', cols)
names(cols) = gtexColors$description[i[!is.na(i)]]

cols = c(cols, colorCodes)[levels(df$key)]

cols[!names(cols) %in% unique(df_cv_FANS$key)] = "black"

# ggplot( df, aes(log2(BCV), color=key)) + geom_density() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="bottom") + facet_wrap(~Dataset, ncol=1) + scale_color_manual("Type", values=cols)

ggplot(df[,nrow(.SD), by="key"], aes(key, V1, fill=key)) + geom_bar(stat="identity") + theme_bw(10) + theme(aspect.ratio=5, plot.title = element_text(hjust = 0.5), legend.position="none") + scale_fill_manual("Cell Type", values=cols) + coord_flip() + xlab('') + ylab("# genes exprressed")

ggplot( df[CV.ratio>=1,], aes(CV.ratio, ..density.., color=key)) + geom_density(size=2) + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + scale_color_manual("Cell Type", values=cols) + xscale("log2", .format = TRUE) + facet_wrap(~Dataset, ncol=1)

ggplot( df[CV.ratio>=1,], aes(log2(CV.ratio), color=key)) + stat_ecdf() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + scale_color_manual("Cell Type", values=cols) + facet_wrap(~Dataset, ncol=2) + ylab("eCDF") + scale_x_continuous(expand=c(0,0), limits=c(0, 10))

df2 = df[,data.frame(Dataset=Dataset[1], Count = sum(CV.ratio > 2) / length(CV.ratio)),by="key"]

v = df2[Dataset=="GTEx", median(Count)]

ggplot(df2, aes(key, Count, fill=key))  + geom_bar(stat="identity") + theme_bw(10) + theme(aspect.ratio=5, plot.title = element_text(hjust = 0.5), legend.position="none") + scale_fill_manual("Cell Type", values=cols) + coord_flip() + xlab('') + ylab('# genes passing cutoff') + geom_hline(yintercept=v, linetype="dashed", color="grey70")

df2 = df[,data.frame(Dataset=Dataset[1], Count = sum(FDR < .05)/length(FDR)),by="key"]
v = df2[Dataset=="GTEx", median(Count)]

ggplot(df[,sum(FDR < .05),by='key'], aes(key, V1, fill=key))  + geom_bar(stat="identity") + theme_bw(8) + theme(aspect.ratio=5, plot.title = element_text(hjust = 0.5), legend.position="none") + scale_fill_manual("Cell Type", values=cols) + coord_flip() + xlab('') + ylab('# genes passing cutoff') + geom_hline(yintercept=v, linetype="dashed", color="grey70")

```

```{r cv.old}

# cv = function(x){
#   sd(x) / mean(x)
# }

# #   cv_est = apply(geneExpr[,id], 1, cv)

# #   data.frame( key = CT, Gene = names(cv_est), CV = cv_est, stringsAsFactors=FALSE)
# # })
# # df_cv_FANS = do.call(rbind, df_cv_FANS)

# # CV for GTEX
# df_cv_GTEx = lapply( unique(gtex_sample$SMTSD), function(Tissue){

#   id = gtex_sample$SAMPID[which( gtex_sample$SMTSD == Tissue)]

#   cv_est = apply(gtex_log2cpm[,id], 1, cv)

#   data.frame( key = Tissue, Gene = names(cv_est), CV = cv_est, stringsAsFactors=FALSE)
# })
# df_cv_GTEx = do.call(rbind, df_cv_GTEx)

# df = rbind(df_cv_FANS, df_cv_GTEx)

# ggplot( df_cv_FANS, aes(log10(CV), color=key)) + geom_density() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=colorCodes) + xlim(-2, 5) + ylim(0,1.4)

# ggplot( df_cv_GTEx, aes(log10(CV), color=key)) + geom_density() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + xlim(-2, 5) + ylim(0,1.4)

# ggplot( df, aes(log10(CV), color=key)) + geom_density() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + xlim(-2, 5) + ylim(0,1.4)

```

### PCA including GTEx
Computed on residuals

```{r PCA.gtex}

id = gtex_sample$SAMPID[which( gtex_sample$SMTSD == 'Brain - Frontal Cortex (BA9)')]
countsCombined = merge(geneCounts, gtex_count[,id], by="row.names")
rownames(countsCombined) = countsCombined$Row.names
countsCombined$Row.names = NULL

isexpr = rowSums(cpm(countsCombined)>1) >= 0.3*ncol(countsCombined)

dge = DGEList(countsCombined[isexpr,])
dge = calcNormFactors(dge)

geneExprCombined = cpm(dge, log=TRUE)

infoExpand = rbind(data.frame(info[,c(1,3)], dataset="FANS4", stringsAsFactors=FALSE), 
  data.frame(Name=colnames(gtex_count[,id]), CellType = "GTEx BA9", dataset="GTEx", stringsAsFactors=FALSE))
rownames(infoExpand) = infoExpand$Name

# the first PC separates FANS from bulk, so subtract that and use residuals
fit = dream(geneExprCombined, ~ dataset, infoExpand)
resid = residuals(fit, geneExprCombined)

dcmp = prcomp( t(resid), scale=TRUE)

df_pca = data.frame(dcmp$x)

df_pca = merge(df_pca, infoExpand, by.x="row.names", by.y="Name")

varFrac = dcmp$sdev^2 / sum(dcmp$sdev^2)

xlab = paste0('PC1 (', format(varFrac[1]*100, digits=3), '%)')
ylab = paste0('PC2 (', format(varFrac[2]*100, digits=3), '%)')
ggplot(df_pca, aes(PC1, PC2, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=c(colorCodes, "grey")) + xlab(xlab) + ylab(ylab)

# xlab = paste0('PC2 (', format(varFrac[2]*100, digits=3), '%)')
# ylab = paste0('PC3 (', format(varFrac[3]*100, digits=3), '%)')
# ggplot(df_pca, aes(PC2, PC3, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=c(colorCodes, "grey")) + xlab(xlab) + ylab(ylab)

```

### MDS
computed on log cpm

```{r GTEX.MDS}

d = as.dist(1 - cor(geneExprCombined)^2)
# d = as.dist(1 - cor(apply(geneExprCombined, 2, RNOmni::rankNorm), method="sp")^2)

df_mds = data.frame(cmdscale( d, k=3))

df_mds = merge(df_mds, infoExpand, by.x="row.names", by.y="Name")

ggplot(df_mds, aes(X1, X2, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=c(colorCodes, "grey")) + xlab('Coordinate 1') + ylab('Coordinate 2')

ggplot(df_mds, aes(X2, X3, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=c(colorCodes, "grey")) + xlab('Coordinate 2') + ylab('Coordinate 3')
```

## Differential expression with Increasing sample size
```{r subset.analysis}

n_array =  c(3, 5, 10, 20, 28)
sampleRamp = lapply(n_array, function(N){

  df = sapply(levels(info$CellType), function(CT){
    info$Name[which(info$CellType == CT)[1:N]]
    })
  data.frame(df, stringsAsFactors=FALSE)
  })

index = combn(nlevels(info$CellType),2)

tabRampList = lapply( sampleRamp, function(sampleSet){

  infoSub = droplevels(info[info$Name %in% unlist(sampleSet),])
  geneCountSub = geneCounts[,infoSub$Name]

  # for all pairs of cell types
  tabList = lapply( 1:ncol(index), function(j){

    CT1 = levels(infoSub$CellType)[index[1,j]]
    CT2 = levels(infoSub$CellType)[index[2,j]]
    key = paste0(CT1, '.', CT2)

    idx = which(infoSub$CellType %in% c(CT1, CT2))

    message(key, " ", nrow(sampleSet) )
    infoSub2 = droplevels(infoSub[idx,])

    isexpr = rowSums(cpm(geneCountSub[,idx])>1) >= 0.3*ncol(geneCountSub[,idx])
    table(isexpr)

    dge = DGEList( geneCountSub[isexpr,idx] )
    dge = calcNormFactors(dge)

    form = ~ CellType
    vobj = voomWithDreamWeights(dge, form, infoSub2, plot=FALSE, save.plot=TRUE, BPPARAM=SnowParam(6, progressbar=TRUE), quiet=TRUE)

    form = ~ CellType
    fit = dream( vobj, form, infoSub2, BPPARAM=SnowParam(6, progressbar=TRUE), quiet=TRUE)
    fit = eBayes(fit)

    coef = paste0('CellType',levels(infoSub2$CellType)[2])

    topTable(fit, coef=coef, number=Inf)
  })
  names(tabList) = apply(matrix(levels(infoSub$CellType)[index], nrow=2), 2, function(x) paste0(x, collapse='.'))
  tabList  
})
```

Just counting coding genes, but analysis was for everything

```{r subset.analysis.plots, fig.width=9}

df_de_summary = lapply( 1:length(tabRampList), function(j){

  tabList = tabRampList[[j]]

  df_de_summary = lapply( names(tabList), function(key){
    tab = tabList[[key]]

    geneTab = select(ensdb, keys=rownames(tab), columns=c("GENEID", "GENEBIOTYPE"), keytype="GENEID")

    tab = merge(tab, geneTab, by.x="row.names", by.y="GENEID")

    tab = data.table(tab)
    tab = tab[GENEBIOTYPE=='protein_coding',]

    df_de_summary = data.frame( Comparison  = key, 
                N           = n_array[j],
                nGenes      = nrow(tab), 
                nDE         = sum(tab$adj.P.Val < 0.05),
                nDE_gt_1    =  nrow(tab[adj.P.Val < 0.05 & abs(logFC) > 1,]), 
                nDE_gt_2    =  nrow(tab[adj.P.Val < 0.05 & abs(logFC) > 2,]), 
                nDE_gt_3    =  nrow(tab[adj.P.Val < 0.05 & abs(logFC) > 3,]), 
                nDE_gt_4    =  nrow(tab[adj.P.Val < 0.05 & abs(logFC) > 4,]), 
                pi1         = 1-qvalue(tab$P.Value)$pi0)
    })
   do.call(rbind, df_de_summary)
})
df_de_summary = do.call(rbind, df_de_summary)

ggplot(df_de_summary, aes(N, nDE_gt_1, color=Comparison)) + geom_point() + geom_line() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position = "right") + ylab("# DE and |logFC| > 1") + xlab("Sample size") + ylim(0,NA)
```

# Disease
```{r DE.disease}
# Disease
fitDisease = dream( vobj, ~ 0 + CellType*Dx + (1|Individual.ID) + (1|Gender), info, BPPARAM=SnowParam(6, progressbar=TRUE))
```

```{r Disease.analysis}
df_disease = lapply(levels(info$CellType), function(CT){

  key = paste0('CellType', CT, ':DxSCZ')

  if( CT == levels(info$CellType)[1]){
    key = 'DxSCZ'
  }

  tabl = topTable( fitDisease, coef = key, number=Inf)

  data.frame(CellType = CT, key, count = sum(tabl$adj.P.Val < 0.05), pi1 = 1 - qvalue(tabl$P.Value)$pi0)
  })
df_disease = do.call(rbind, df_disease)

kable(df_disease, digits=2, row.names=FALSE) %>% kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

Volcano plot colored by cell type

DE from null model
Power vs sample size
Paired versus non-paired
Plot examples of marker genes
distribution of effect size
GTEX brain span.
Compare cell types
Homogeneity of variance for each cell type

For next week: deconvolution of GTEx, PEC, CMC, BrainSpan, BrainVar

bMIND can incorporate the prior of cell type variance/covariance. That is, the K x K covariance matrix for K cell types for each gene. I am wondering if Gabriel can also share this info, if sharing the sample-level data is not possible.

### Save mean expression per cell type
https://www.synapse.org/#!Synapse:syn22231629
```{r save.reference.panel}

# compute mean for each cell type, using precision weights
fitCT = dream( vobj, ~ 0 + CellType, info, BPPARAM=SnowParam(6, progressbar=TRUE))
fitCT = eBayes(fitCT)

df_RefPanel = lapply( colnames(fitCT), function(coef){

  # Extract mean for this cell type
  tabl = topTable(fitCT, coef=coef, number=Inf, sort.by='none')
  tabl$se = tabl$logFC / tabl$t # compute standard error given beta and t

  df = data.frame(tabl[,c('logFC', 'se')])
  rownames(df) = rownames(tabl)
  colnames(df) = paste0(gsub('CellType', '', coef), c('', '.se'))
  df

  })
df_RefPanel = do.call(cbind, df_RefPanel)
df_RefPanel = data.frame(EnsID = rownames(df_RefPanel), df_RefPanel)

exclude = grep("\\.se$", colnames(df_RefPanel))
```

```{r store, eval=TRUE}
# Code
CODE <- Folder(name = "FANS4 processing and testing", parentId = "syn22231629") # Make sure these match the folder name and parent Id in the first chunk.
CODE <- synStore(CODE)

# write mean expression
file = 'Roussos_MSSM_MolProf_FANS_4_log2CPM_reference_panel.csv.gz'
gz1 <- gzfile(file, "w")
write.table(df_RefPanel[,-exclude], gz1, quote=FALSE, sep=",", row.names=FALSE)
close(gz1)

nEXP_OBJ = File(file, 
                name = 'FANS4 reference panel: log2CPM', 
                parentId = CODE$properties$id)
nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, 
                    executed = thisFile)

# write standard error
file = 'Roussos_MSSM_MolProf_FANS_4_log2CPM_se_reference_panel.csv.gz'
gz1 <- gzfile(file, "w")
write.table(df_RefPanel[,c(1,exclude)], gz1, quote=FALSE, sep=",", row.names=FALSE)
close(gz1)

nEXP_OBJ = File(file, 
                name = 'FANS4 reference panel: log2CPM standard error', 
                parentId = CODE$properties$id)
nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, 
                    executed = thisFile)

# full counts
file = '/hpc/users/hoffmg01/work/cmc_analysis/src/Roussos_MSSM_MolProf_FANS_4_geneCounts.csv.gz'
gz1 <- gzfile(file, "w")
write.table(data.table(EnsID = rownames(geneCounts), geneCounts), gz1, quote=FALSE, sep=",", row.names=FALSE)
close(gz1)

nEXP_OBJ = File(file, 
                name = 'FANS4 reference panel: full counts', 
                parentId = CODE$properties$id)
nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, 
                    executed = thisFile)

# full counts
file = '/hpc/users/hoffmg01/work/cmc_analysis/src/Roussos_MSSM_MolProf_FANS_4_info.csv.gz'
gz1 <- gzfile(file, "w")
write.table(info, gz1, quote=FALSE, sep=",", row.names=FALSE)
close(gz1)

nEXP_OBJ = File(file, 
                name = 'FANS4 reference panel: metadata', 
                parentId = CODE$properties$id)
nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, 
                    executed = thisFile)

```

### Source
[markdown](`r thisFile`)

### 2020-08-21 deconvolution/fans4/molec_profiling.Rmd
Initial commit

@@
 
+source("../../common_functions/common_functions.R")
+
 knitr::opts_chunk$set(
@@
 geneInfo = df_featureCounts_gene[,1:6]
-geneInfo$EnsId = gsub("^(.*)(\\.\\S+)$", "\\1", geneInfo$Geneid)
+geneInfo$EnsId = trim_ensembl_ids( geneInfo$Geneid )
+geneInfo = geneInfo %>% getGeneSymbol( 'Geneid' )
 
@@
 geneCounts = geneCounts[!exclude,]
-rownames(geneCounts) = gsub("^(.*)(\\.\\S+)$", "\\1", rownames(geneCounts))
+rownames(geneCounts) = trim_ensembl_ids( rownames(geneCounts) )
 
 # convert to HUGO names
-df_hugo = select(org.Hs.eg.db, geneInfo$EnsId, "SYMBOL", "ENSEMBL")
-geneInfo = merge(geneInfo, df_hugo, by.x='EnsId', by.y="ENSEMBL")
+# df_hugo = select(org.Hs.eg.db, geneInfo$EnsId, "SYMBOL", "ENSEMBL")
+# geneInfo = merge(geneInfo, df_hugo, by.x='EnsId', by.y="ENSEMBL")
 

### 2020-08-21 deconvolution/fans4/molec_profiling.Rmd
Initial commit

@@
 geneCounts = df_featureCounts_gene[,7:ncol(df_featureCounts_gene)]
-geneCounts = as.data.frame(geneCounts)
+geneCounts = as.matrix(geneCounts)
 rownames(geneCounts) = geneInfo$Geneid
@@
 ```{r preprocess}
-isexpr = rowSums(cpm(geneCounts)>1) >= 0.1*ncol(geneCounts)
+isexpr = rowSums(edgeR::cpm(geneCounts)>1) >= 0.1*ncol(geneCounts)
 table(isexpr)
@@
 
-geneExpr = cpm(dge, log=TRUE)
+geneExpr = edgeR::cpm(dge, log=TRUE)
 
@@
 
-  isexpr = rowSums(cpm(geneCounts[,idx])>1) >= 0.3*ncol(geneCounts[,idx])
+  isexpr = rowSums(edgeR::cpm(geneCounts[,idx])>1) >= 0.3*ncol(geneCounts[,idx])
   table(isexpr)
@@
 
-system("ml python; synapse add --parentid syn22243354 Roussos_MSSM_MolProf_FANS_4_diffExpression.RDS")
+# system("ml python; synapse add --parentid syn22243354 Roussos_MSSM_MolProf_FANS_4_diffExpression.RDS")
 ```
@@
 
-file_SampleAttributesDS = "/sc/arion/projects/CommonMind/data/resources/GTEx/GTEx_Analysis_v8_Annotations_SampleAttributesDS.txt"
-file_SubjectPhenotypesDS= "/sc/arion/projects/CommonMind/data/resources/GTEx/GTEx_Analysis_v8_Annotations_SubjectPhenotypesDS.txt"
-file_counts = "/sc/arion/projects/CommonMind/data/resources/GTEx/GTEx_Analysis_2017-06-05_v8_RNASeQCv1.1.9_gene_reads.gct.gz"
+# file_SampleAttributesDS = "/sc/arion/projects/CommonMind/data/resources/GTEx/GTEx_Analysis_v8_Annotations_SampleAttributesDS.txt"
+file_SampleAttributesDS = synGet('syn22321103')$path
+
+# file_SubjectPhenotypesDS= "/sc/arion/projects/CommonMind/data/resources/GTEx/GTEx_Analysis_v8_Annotations_SubjectPhenotypesDS.txt"
+file_SubjectPhenotypesDS = synGet('syn22321104')$path
+
+# file_counts = "/sc/arion/projects/CommonMind/data/resources/GTEx/GTEx_Analysis_2017-06-05_v8_RNASeQCv1.1.9_gene_reads.gct.gz"
+file_counts = synGet('syn22321107')$path
 
@@
 
-gtexColors = read.csv('/sc/arion/projects/CommonMind/data/resources/GTEx/GTEx_Tissue_colors.csv', header=TRUE)
+# file= '/sc/arion/projects/CommonMind/data/resources/GTEx/GTEx_Tissue_colors.csv'
+file = synGet('syn22321106')$path
+gtexColors = read.csv( file, header=TRUE)
 ```
@@
 # full counts
-file = '/hpc/users/hoffmg01/work/cmc_analysis/src/Roussos_MSSM_MolProf_FANS_4_geneCounts.csv.gz'
+file = 'Roussos_MSSM_MolProf_FANS_4_geneCounts.csv.gz'
 gz1 <- gzfile(file, "w")
@@
 # full counts
-file = '/hpc/users/hoffmg01/work/cmc_analysis/src/Roussos_MSSM_MolProf_FANS_4_info.csv.gz'
+file = 'Roussos_MSSM_MolProf_FANS_4_info.csv.gz'
 gz1 <- gzfile(file, "w")
@@
 
+# Differential expression
+file = Roussos_MSSM_MolProf_FANS_4_diffExpression.RDS
+
+nEXP_OBJ = File(file, 
+                name = 'FANS4 Differential Expression results', 
+                parentId = CODE$properties$id)
+nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, 
+                    executed = thisFile)
+
 ```

### 2020-08-21 deconvolution/fans4/molec_profiling.Rmd
Initial commit

@@
 
-```{r mvIC}
+```{r mvIC, eval=FALSE}
 colnames(info) = gsub(' ', '_', colnames(info))

### 2020-08-21 deconvolution/fans4/molec_profiling.Rmd
Initial commit

@@
 
-isexpr = rowSums(cpm(countsCombined)>1) >= 0.3*ncol(countsCombined)
+isexpr = rowSums(edgeR::cpm(countsCombined)>1) >= 0.3*ncol(countsCombined)
 
@@
 
-geneExprCombined = cpm(dge, log=TRUE)
+geneExprCombined = edgeR::cpm(dge, log=TRUE)
 
@@
 
-    isexpr = rowSums(cpm(geneCountSub[,idx])>1) >= 0.3*ncol(geneCountSub[,idx])
+    isexpr = rowSums(edgeR::cpm(geneCountSub[,idx])>1) >= 0.3*ncol(geneCountSub[,idx])
     table(isexpr)

### 2020-08-21 deconvolution/fans4/molec_profiling.Rmd
Initial commit

@@
 nEXP_OBJ = File(file, 
-                name = 'FANS4 reference panel: log2CPM', 
+                name = 'FANS4 reference panel - log2CPM', 
                 parentId = CODE$properties$id)
@@
 nEXP_OBJ = File(file, 
-                name = 'FANS4 reference panel: log2CPM standard error', 
+                name = 'FANS4 reference panel - log2CPM standard error', 
                 parentId = CODE$properties$id)
@@
 nEXP_OBJ = File(file, 
-                name = 'FANS4 reference panel: full counts', 
+                name = 'FANS4 reference panel - full counts', 
                 parentId = CODE$properties$id)
@@
 nEXP_OBJ = File(file, 
-                name = 'FANS4 reference panel: metadata', 
+                name = 'FANS4 reference panel - metadata', 
                 parentId = CODE$properties$id)

### 2020-08-24 deconvolution/apply_dtangle_FANS4.Rmd
Initial commit

---
title: "Deconvolution to FANS4"
author: "Gabriel Hoffman"
date: "`r date()`"
output: html_document
---

```{r knit2synapse, eval=FALSE}
ml git pandoc
rmarkdown::render("deconvolution/apply_dtangle_FANS4.Rmd")

library(synapser)
library(knit2synapse)

synLogin()

knit2synapse::createAndKnitToFolderEntity(file = "deconvolution/apply_dtangle_FANS4.Rmd",
                                 parentId ="syn22321035",
                                 folderName = "Deconvolution results",
                                 knitmd=FALSE)
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
library(data.table)
library(dtangle)
library(edgeR)
library(variancePartition)
library(BiocParallel)

library(synapser)
library(future)
library(knitr)
library(githubr) # get the package from devtools::install_github('brian-bot/githubr')

source("../common_functions/common_functions.R")

synLogin()

# See dream documentation to customize parallel processing       http://bioconductor.org/packages/release/bioc/vignettes/variancePartition/inst/doc/dream.html#parallel-processing
param = SnowParam((availableCores()-2)/2, "SOCK", progressbar=TRUE)

register(param)

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
parentId = 'syn22321035';
activityName = 'Deconvolution';
activityDescription = 'Deconvolution';

thisFileName <- 'apply_dtangle_FANS4.Rmd'

# Github link
thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('deconvolution/',thisFileName))
```

```{r read.data}

# Read reference panel
######################
geneExpr_FANS4_lcpm = read.csv(synGet('syn22321183')$path, header=TRUE)
rownames(geneExpr_FANS4_lcpm) = geneExpr_FANS4_lcpm[,1]
geneExpr_FANS4_lcpm = geneExpr_FANS4_lcpm[,-1]
metadata_FANS4 = read.csv(synGet("syn22321189")$path, header=TRUE, row.names=1)

# read Raw counts
#################
file_counts_synid = c('syn21888969', 'syn21862455', 'syn21887889', 'syn21825508')

# read one file at a time
geneCounts_list = lapply(file_counts_synid, function(id){
	read.table(synGet(id)$path, header=TRUE, check.names=FALSE)
	})

# lapply(geneCounts_list, function(x) x[1:3, 1:3])

# merge files based on gene name
geneCounts = Reduce(function(x,y) merge(x = x, y = y, by = "ensembl_gene_id"), geneCounts_list)
rownames(geneCounts) = trim_ensembl_ids(geneCounts[,1])
geneCounts = geneCounts[,-1]

# TMM normalization
geneExpr = DGEList( as.matrix(geneCounts))
geneExpr = calcNormFactors(geneExpr)
geneExpr_logCPM = cpm(geneExpr, log=TRUE)

# only keep genes measured in FANS4
keep = rownames(geneExpr_logCPM) %in% rownames(geneExpr_FANS4_lcpm)
geneExpr_logCPM = geneExpr_logCPM[keep,]

```

```{r process.counts}

idx = match(rownames(geneExpr_logCPM), rownames(geneExpr_FANS4_lcpm))

# a = cbind( rownames(geneExpr_logCPM), rownames(geneExpr_FANS4_lcpm[idx,]) )
# table(a[,1] == a[,2])
# which(a[,1] != a[,2])

cellFracEstimates = dtangle( Y = t(geneExpr_logCPM), reference = t(geneExpr_FANS4_lcpm[idx,]))

cellFracEstimates2 = dtangle2( Y = t(geneExpr_logCPM), reference = t(geneExpr_FANS4_lcpm[idx,]))

cor(cellFracEstimates$estimates, cellFracEstimates2$estimates, method="spearman")
```

### Data preprocessing
```{r preprocess.data}
# Get genotype ids from synapse
GENOTYPE_ID = 'syn16816490'
ALL_USED_IDs = c(ALL_USED_IDs, GENOTYPE_ID)
GENOTYPE = downloadFile(GENOTYPE_ID) %>% 
  dplyr::select(Individual_ID, `SNP_report:Genotyping_Sample_ID`, `SNP_report:Exclude?`) %>% 
  dplyr::inner_join(ANCESTRY) %>% 
  dplyr::filter(is.na(`SNP_report:Exclude?`))

# Merge metadata
METADATA = metadata %>%
  dplyr::left_join(GENOTYPE) %>% 
  dplyr::rename(Region = `rnaSeq_dissection:Brain_Region`,
                PMI = `PMI_(in_hours)`,
                RIN = `rnaSeq_isolation:RIN`, 
                ReportExclude = `rnaSeq_report:Exclude?`,
                GenotypeExclude = `SNP_report:Exclude?`,
                SampleID = Sample_RNA_ID,
                LibraryBatch = `rnaSeq_report:Library_Batch`,
                FlowcellBatch = `rnaSeq_report:Flowcell_Batch`,
                RibozeroBatch = `rnaSeq_report:Ribozero_Batch`,
                MappedReads = `rnaSeq_report:Mapped_Reads`,
                IntragenicRate = `rnaSeq_report:Intragenic_Rate`, IntronicRate = `rnaSeq_report:Intronic_Rate`, 
                IntergenicRate = `rnaSeq_report:Intergenic_Rate`, GenesDetected = `rnaSeq_report:Genes_Detected`,
                ExpProfEfficiency = `rnaSeq_report:Expression_Profiling_Efficiency`, rRNARate = `rnaSeq_report:rRNA_Rate`,
                TotalReads = `rnaSeq_report:Total_Reads`, AlignmentRate = `rnaSeq_report:Percent_Aligned`, TranscriptsDetected = `rnaSeq_report:Transcripts_Detected`) %>%
  dplyr::select(SampleID, Individual_ID, Institution, Cohort, Reported_Gender,Sex, Ethnicity, ageOfDeath, PMI, Dx, 
                RIN, ReportExclude, GenotypeExclude, EV.1, EV.2, EV.3, EV.4, EV.5,
                LibraryBatch, FlowcellBatch, RibozeroBatch, MappedReads, TotalReads, GenesDetected, AlignmentRate,
                IntragenicRate, IntergenicRate, IntronicRate, ExpProfEfficiency, rRNARate, TranscriptsDetected) %>%
  dplyr::filter(Cohort %in% "NIMH-HBCC") %>% 
  dplyr::filter(SampleID %in% colnames(COUNT), !is.na(SampleID)) %>% 
  dplyr::mutate(Dx = forcats::fct_recode(Dx, Other = "BP", Other = "undetermined", Control = "Control", SCZ = "SCZ"))

ind = METADATA$SampleID[which(METADATA$ReportExclude == 1 | METADATA$GenotypeExclude)]
writeLines(paste('Following',length(ind),'samples are marked exclude'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$Ethnicity) | is.na(METADATA$Institution) | is.na(METADATA$Dx)]
writeLines(paste('Following', length(ind), 'counts are missing any metadata'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$PMI)]
writeLines(paste('Following', length(ind), 'counts are missing PMI'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$Reported_Gender)]
writeLines(paste('Following', length(ind), 'counts are missing gender'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$ageOfDeath)]
writeLines(paste('Following', length(ind), 'counts are missing age of death'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$EV.1)]
writeLines(paste('Following', length(ind), 'counts are missing ancestry information'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))
```

```{r preprocess.data1, results='asis'}
# Match covariates to expression data
indToRetain = intersect(METADATA$SampleID, colnames(COUNT))

rownames(COUNT) <- COUNT$Geneid
COUNT = COUNT[,indToRetain]

rownames(METADATA) = METADATA$SampleID
METADATA = METADATA[indToRetain,]

METADATA %>% 
  group_by(Dx) %>% 
  dplyr::summarise(count = n()) %>% 
  kable()

FactorCovariates <- c('Individual_ID', "Reported_Gender", "LibraryBatch", "Dx", "FlowcellBatch", "RibozeroBatch")
ContCovariates <- c("ageOfDeath", "PMI", "RIN", "MappedReads", "IntragenicRate", "IntronicRate", "IntergenicRate",
                    "GenesDetected", "ExpProfEfficiency", "rRNARate", "TotalReads", "AlignmentRate",
                    "EV.1", "EV.2", "EV.3", "EV.4", "EV.5")

# Find inter relation between factor covariates
COVARIATES = METADATA[,c(FactorCovariates,ContCovariates),drop=F]
COVARIATES[,FactorCovariates] <- data.frame(lapply(COVARIATES[,FactorCovariates],function(x){
  x <- sapply(x,function(y){str_replace_all(as.character(y),'[^[:alnum:]]','_')})}))
rownames(COVARIATES) <- METADATA$SampleID

# Convert factor covariates to factors
COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], function(x){
  x = as.numeric(as.character(gsub('[\\,\\%]','',x)))
})

```{r store, eval=FALSE}
# Code
CODE <- Folder(name = "Meta-analysis on sex", parentId = "syn22315403") # Make sure these match the folder name and parent Id in the first chunk.
CODE <- synStore(CODE)

# Store meta-analysis
file = "cellFracEstimates.RDS"
saveRDS(cellFracEstimates, file=file)
nEXP_OBJ = File(file, 
                name = 'Result of dtangle', 
                parentId = CODE$properties$id)
nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, 
                    executed = thisFile)
res_acc %>%
  tibble::rownames_to_column(var = "gene_id") %>% 
  write.table(file = 'CMC_acc_Sex_meta_analysis.tsv', sep = '\t', row.names=F, quote=F)
nEXP_OBJ = File('CMC_acc_Sex_meta_analysis.tsv', 
                name = 'Sex Normalised, meta-analysis (ACC)', 
                parentId = CODE$properties$id)
nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, 
                    executed = thisFile)
```

### Source
[markdown](`r thisFile`)

### 2020-08-24 testing/implementation_prcomp.R
describe implementation details of prcomp

@@ -1,0 +1,22 @@
+# This just an implementation detail: x and rotation are complimentary.  Since we ran prcomp with genes as rows, the rotation matrix stores the relevant PCs:
+Y = matrix(rnorm(10000), ncol=50)
+rownames(Y) = paste0('gene_', 1:nrow(Y))
+colnames(Y) = paste0('ID_', 1:ncol(Y))
+
+dcmp = prcomp( Y )
+
+# Components for each Gene
+dcmp$x[1:4, 1:4]
+
+# Components for each Individual
+dcmp$rotation[1:4, 1:4]
+
+
+# PCA on the transpose
+dcmp2 = prcomp( t(Y) )
+
+# Components for each individual
+dcmp2$x[1:4, 1:4]
+
+# Components for each gene
+dcmp2$rotation[1:4, 1:4]

### 2020-08-24 individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_Diagnosis.Rmd
integrate CQN using voomWithOffset

--- 2020-08-19 individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_Diagnosis.Rmd
@@
 library(githubr) # get the package from devtools::install_github('brian-bot/githubr')
+library(rctutils) # get the packages from BiocManager::install("DarwinAwardWinner/rctutils")
 
@@
 writeLines(glue::glue('Using following covariates in the model:',
-                 glue::glue_collapse(adjust.covars, sep = ", ", last = " and "),
-                 ' as fixed effects and',
-                 glue::glue_collapse(random_effect, sep = ", ", last = " and "),' chosen as random effects'))
+                 glue::glue_collapse(c(adjust.covars, random_effect), sep = ", ", last = " and "),
+                 ' as fixed effects.'))
 
@@
 # Scale continuous variables
-adjust.covars = sapply(1:length(col_type$class), function(i){
+covars = sapply(1:length(col_type$class), function(i){
   switch(col_type$class[i], 
@@
 # Post adjusted formula
-formula <- glue::glue("~ ", glue::glue_collapse(adjust.covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))
+formula <- formula(glue::glue("~", glue::glue_collapse(covars, sep = " + ")))
+design <- model.matrix(formula, COVARIATES)
 
@@
 geneExpr = DGEList(NEW.COUNTS)
-geneExpr = edgeR::calcNormFactors(geneExpr)
+
+# Extract default offset
+offset_default = edgeR::expandAsMatrix(edgeR::getOffset(geneExpr), dim(geneExpr))
+
+# Build custom offset from default offset and the CQN offset
+geneExpr$offset = offset_default - CQN.GENE_EXPRESSION
+
+# Run rctutils::voomWithOffset
+vobj_cqn_offset = rctutils::voomWithOffset(geneExpr, design)
 
@@
 # Fit linear model using new weights and new design
-# VOOM.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION
-ADJUSTED.FIT = variancePartition::dream(exprObj = VOOM.GENE_EXPRESSION,
+ADJUSTED.FIT = variancePartition::dream(exprObj = vobj_cqn_offset,
                                         formula = formula,
@@
 synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
-
 ```

### 2020-08-24 individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_Diagnosis.Rmd
remove samples based on age threshold
update provenance

@@
 # Github link
-thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
+thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='modify-cqn-implementation')
 thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/split_by_analysis/',thisFileName))
@@
   tibble::column_to_rownames(var = "SampleID")
+
+# Remove ages under 17
+COVARIATES <- COVARIATES[COVARIATES$ageOfDeath >= 17,]
+NEW.COUNTS <- NEW.COUNTS[names(NEW.COUNTS) %in% rownames(COVARIATES)]
+CQN.GENE_EXPRESSION <- CQN.GENE_EXPRESSION[names(CQN.GENE_EXPRESSION) %in% rownames(COVARIATES)]
 
@@
 counts = 'syn21886235'
-ALL_USED_IDs = counts
+ALL_USED_IDs = c(ALL_USED_IDs, counts)
 counts = downloadFile(counts) %>% data.frame()
@@
 ```
+### Source Code
+[markdown file](`r thisFile`)
 

### 2020-08-24 individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_Diagnosis.Rmd
fix type error

@@
 CQN.GENE_EXPRESSION = downloadFile(CQN_ID) %>% 
-  tibble::column_to_rownames(var = "ensembl_gene_id") %>% 
-  data.matrix()
+  tibble::column_to_rownames(var = "ensembl_gene_id")
 
@@
 NEW.COUNTS <- NEW.COUNTS[names(NEW.COUNTS) %in% rownames(COVARIATES)]
-CQN.GENE_EXPRESSION <- CQN.GENE_EXPRESSION[names(CQN.GENE_EXPRESSION) %in% rownames(COVARIATES)]
+CQN.GENE_EXPRESSION <- CQN.GENE_EXPRESSION[names(CQN.GENE_EXPRESSION) %in% rownames(COVARIATES)] %>% 
+  data.matrix()
 

### 2020-08-24 individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_Diagnosis.Rmd
fix contrasts
scale vars to fix correlation

@@
 ```{r set-formula}
-formula <- glue::glue("~ 0 + ", glue::glue_collapse(adjust.covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))    
+adjust.covars <- glue::glue('scale({adjust.covars})')[-1]
+formula <- glue::glue("~ 0 + Dx + ", glue::glue_collapse(adjust.covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))    
 ```
@@
 # Manually set comparison of SCZ-Other to Control
-L = cbind(L1, L2, L3, data.frame(L4 = c(1, -0.5, -0.5, 0, 0, 0, 0, 0, 0)))
+L = cbind(L1, L2, L3, data.frame(L4 = c(1, -0.5, -0.5, 0, 0, 0, 0, 0)))
 

### 2020-08-25 individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_DxSex.Rmd
stage interaction analysis

--- 2020-08-18 individual-cohorts/HBCC_DLPFC_DE.Rmd
@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/HBCC_DLPFC_DE.Rmd",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_DxSex.Rmd",
                                  parentId ="syn22050019",
@@
 
-thisFileName <- 'HBCC_DLPFC_DE.Rmd'
+thisFileName <- 'HBCC_DLPFC_DE_DxSex.Rmd'
 
 # Github link
-thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
-thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/',thisFileName))
+thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='modify-cqn-implementation')
+thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/split_by_analysis',thisFileName))
 ```
@@
 # Post adjusted formula
-formula <- glue::glue("~ ", glue::glue_collapse(interaction, sep = "*")," + ", glue::glue_collapse(adjust.covars, sep = " + "))
+formula <- formula(glue::glue("~ ", glue::glue_collapse(interaction, sep = "*")," + ", glue::glue_collapse(adjust.covars, sep = " + ")))
+design <- model.matrix(formula, COVARIATES)
 
@@
 geneExpr = DGEList(NEW.COUNTS)
-geneExpr = edgeR::calcNormFactors(geneExpr)
+
+# Extract default offset
+offset_default = edgeR::expandAsMatrix(edgeR::getOffset(geneExpr), dim(geneExpr))
+# Build custom offset from default offset and the CQN offset
+geneExpr$offset = offset_default - CQN.GENE_EXPRESSION
+# Run rctutils::voomWithOffset
+vobj_cqn_offset = rctutils::voomWithOffset(geneExpr, design)
 
@@
 # Fit linear model using new weights and new design
-VOOM.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION
-ADJUSTED.FIT = variancePartition::dream(exprObj = VOOM.GENE_EXPRESSION,
+ADJUSTED.FIT = variancePartition::dream(exprObj = vobj_cqn_offset,
                                         formula = formula,
@@
 ```
-
 ### Source Code
@@
 synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store variancePartition
+saveRDS(sortCols(violin_plot), file = 'CMC_DLPFC_Dx_Sex_variancePartition.rds')
+ADJ_OBJ = File('CMC_DLPFC_Dx_Sex_variancePartition.rds',
+               name = 'Variance Partition result', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store voom object
+saveRDS(VOOM.GENE_EXPRESSION, file = 'CMC_DLPFC_Dx_Sex_voom.rds')
+ADJ_OBJ = File('CMC_DLPFC_Dx_Sex_voom.rds',
+               name = 'voom object', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store contrasts
+saveRDS(L, file = 'CMC_DLPFC_Dx_Sex_contrasts.rds')
+ADJ_OBJ = File('CMC_DLPFC_Dx_Sex_contrasts.rds',
+               name = 'contrast matrix', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store formula
+saveRDS(formula, file = 'CMC_DLPFC_Dx_Sex_formula.rds')
+ADJ_OBJ = File('CMC_DLPFC_Dx_Sex_formula.rds',
+               name = 'formula', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
 ```

### 2020-08-25 individual-cohorts/split_by_analysis/HBCC_dsACC_DE_DxSex.Rmd
stage interaction analysis

--- 2020-08-19 individual-cohorts/HBCC_dsACC_DE.Rmd
@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/HBCC_dsACC_DE.Rmd",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/split_by_analysis/HBCC_dsACC_DE_DxSex.Rmd",
                                  parentId ="syn22084347",
@@
 
-thisFileName <- 'HBCC_dsACC_DE.Rmd'
+thisFileName <- 'HBCC_dsACC_DE_DxSex.Rmd'
 
 # Github link
-thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
-thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/',thisFileName))
+thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='modify-cqn-implementation')
+thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/split_by_analysis',thisFileName))
 ```
@@
 
-### Normalisation - Dx_Sex
-1. Dx_Sex is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis)
+### Normalisation - Dx*Sex
+1. Dx*Sex is chosen as the primary variable of interest
 
@@
 writeLines(glue::glue('Using following covariates in the model:',
+                      glue::glue_collapse(interaction, sep = ", "),
+                      ', ',
                       glue::glue_collapse(adjust.covars, sep = ", ", last = " and "),
-                 ' as fixed effects and ',
-                 glue::glue_collapse(random_effect, sep = ", ", last = " and "),' chosen as random effects'))
+                      ' as fixed effects.'))
 
-scaled_col_type <- dplyr::select(COVARIATES, c(adjust.covars)) %>%
+scaled_col_type <- dplyr::select(COVARIATES, c(adjust.covars, random_effect)) %>%
   dplyr::summarise_all(class) %>% 
@@
 # Post adjusted formula
-formula <- glue::glue("~ ", glue::glue_collapse(interaction, sep = "*")," + ", glue::glue_collapse(scaled_adjust_covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))
+formula <- formula(glue::glue("~ ", glue::glue_collapse(interaction, sep = "*")," + ", glue::glue_collapse(scaled_adjust_covars, sep = " + ")))   
+design <- model.matrix(formula, COVARIATES)
 
@@
 # Fit linear model using new weights and new design
-VOOM.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION
-ADJUSTED.FIT = variancePartition::dream(exprObj = VOOM.GENE_EXPRESSION$E,
+ADJUSTED.FIT = variancePartition::dream(exprObj = vobj_cqn_offset,
                                         formula = formula,
@@
 `r formula`
-
 ### Residual calculation
@@
 synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store variancePartition
+saveRDS(sortCols(violin_plot), file = 'CMC_dsACC_Dx_Sex_variancePartition.rds')
+ADJ_OBJ = File('CMC_dsACC_Dx_Sex_variancePartition.rds',
+               name = 'Variance Partition result', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store voom object
+saveRDS(VOOM.GENE_EXPRESSION, file = 'CMC_dsACC_Dx_Sex_voom.rds')
+ADJ_OBJ = File('CMC_dsACC_Dx_Sex_voom.rds',
+               name = 'voom object', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store contrasts
+saveRDS(L, file = 'CMC_dsACC_Dx_Sex_contrasts.rds')
+ADJ_OBJ = File('CMC_dsACC_Dx_Sex_contrasts.rds',
+               name = 'contrast matrix', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store formula
+saveRDS(formula, file = 'CMC_dsACC_Dx_Sex_formula.rds')
+ADJ_OBJ = File('CMC_dsACC_Dx_Sex_formula.rds',
+               name = 'formula', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
 ```

### 2020-08-25 individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_ACC_DE_DxSex.Rmd
stage interaction analysis

--- 2020-08-18 individual-cohorts/HBCC_DLPFC_DE.Rmd
@@
 ---
-title: "Differential Expression - HBCC - DLPFC - Dx - Sex"
+title: "Differential Expression - MSSM-Penn-Pitt - ACC - Dx - Sex"
 author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/HBCC_DLPFC_DE.Rmd",
-                                 parentId ="syn22050019",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_ACC_DE_DxSex.Rmd",
+                                 parentId ="syn22012953",
                                  folderName = "Differential Expression - Dx - Sex")
@@
 ## Load required libraries
-library(CovariateAnalysis) # get the package from devtools::install_github('th1vairam/CovariateAnalysis@dev')
 library(data.table)
-library(tidyr)
-library(plyr)
 library(dplyr)
-library(stringr)
 library(R.utils)
-library(ggplot2)
-library(limma)
 library(psych)
 library(lme4)
-library(biomaRt)
 library(variancePartition)
-library(edgeR)
 library(synapser)
 library(knitr)
-library(githubr) # get the package from devtools::install_github('brian-bot/githubr')
+library(githubr)
+library(CovariateAnalysis)
+library(edgeR)
+library(WGCNA)
+library(biomaRt)
+library(rctutils) # get the packages from BiocManager::install("DarwinAwardWinner/rctutils")
 
@@
 ```{r synapse.parameters, include=FALSE, cache=FALSE}
-parentId = 'syn22050019';
+parentId = 'syn22012953';
 activityName = 'Differential Expression';
-activityDescription = 'Differential Expression - Dx_Sex';
+activityDescription = 'Differential Expression - Dx - Sex';
 
-thisFileName <- 'HBCC_DLPFC_DE.Rmd'
+thisFileName <- 'MSSM-Penn-Pitt_ACC_DE_DxSex.Rmd'
 
 # Github link
-thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
-thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/',thisFileName))
+thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='modify-cqn-implementation')
+thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/split_by_analysis',thisFileName))
 ```
@@
 interaction <- c("Dx", "Reported_Gender")
-adjust.covars <- c("ageOfDeath", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate")
+adjust.covars <- c("ageOfDeath", "RIN2", "RIN", "rRNARate", "IntronicRate", "IntragenicRate", "IntergenicRate")
+random_effect <- c("Institution")
 ```
@@
                       glue::glue_collapse(adjust.covars, sep = ", ", last = " and "),
-                      ' as fixed effects.'))
+                      ' as fixed effects and',
+                      glue::glue_collapse(random_effect, sep = ", ", last = " and "),' chosen as random effects'))
 
-col_type <- dplyr::select(COVARIATES, c(adjust.covars)) %>%
+col_type <- dplyr::select(COVARIATES, c(adjust.covars, random_effect)) %>%
   dplyr::summarise_all(class) %>% 
@@
 # Post adjusted formula
-formula <- glue::glue("~ ", glue::glue_collapse(interaction, sep = "*")," + ", glue::glue_collapse(adjust.covars, sep = " + "))
+formula <- formula(glue::glue("~ ", glue::glue_collapse(interaction, sep = "*")," + ", glue::glue_collapse(covars, sep = " + ")))
+design <- model.matrix(formula, COVARIATES)
 
@@
 # Eucledian tree based analysis
-COVARIATES.tmp = data.matrix(COVARIATES[,c('Dx', 'Reported_Gender'), drop = FALSE])
+COVARIATES.tmp = data.matrix(COVARIATES[,c('Dx', 'Reported_Gender', 'Institution'), drop = FALSE])
 
@@
                            main = "Sample dendrogram",
-                           groupLabels = colnames(COVARIATES.tmp[,c('Dx','Reported_Gender'), drop = FALSE]))
+                           groupLabels = colnames(COVARIATES.tmp[,c('Dx','Reported_Gender', 'Institution'), drop = FALSE]))
 ```
@@
 ```{r set-formula}
-formula <- glue::glue("~ 0 + Dx_Sex + ", glue::glue_collapse(adjust.covars, sep = " + "))
+adjust.covars <- adjust.covars[!(adjust.covars %in% "RIN2")]
+formula <- glue::glue("~ 0 + Dx_Sex + ", glue::glue_collapse(adjust.covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))    
 ```
@@
                                    coefficient = c("Dx_SexControl_Male", "Dx_SexControl_Female"))
-L2 = c(1, -1, 0, 0, -1, 1, 0, 0, 0, 0, 0)
-L3 = c(1, -1, -1, 1, 0, 0, 0, 0, 0, 0, 0)
+L2 = c(1, -1, 0, 0, -1, 1, 0, 0, 0, 0, 0, 0)
+L3 = c(1, -1, -1, 1, 0, 0, 0, 0, 0, 0, 0, 0)
 
@@
 # Get gene lengths
-counts = 'syn21886235'
+counts = 'syn21867938'
 ALL_USED_IDs = c(counts, ALL_USED_IDs)
@@
                        mart = mart)
-
-FIT.CONTR <- eBayes(FIT.CONTR)
 
@@
 
-  tissue	= 'dorsolateral prefrontal cortex', 
+  tissue	= 'anterior cingulate cortex', 
   study = 'CMC',
@@
   rownameToFirstColumn('ensembl_gene_id') %>%
-  write.table(file = 'CMC_DLPFC_Dx_Sex_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
-nEXP_OBJ = File('CMC_DLPFC_Dx_Sex_netResidualExpression.tsv', 
+  write.table(file = 'CMC_ACC_Dx_Sex_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
+nEXP_OBJ = File('CMC_ACC_Dx_Sex_netResidualExpression.tsv', 
                 name = 'Dx_Sex Normalised, covariates removed residual expression (for network analysis)', 
@@
 # Store differential expression results
-write.table(DE, file = 'CMC_DLPFC_Dx_Sex_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
-DEXP_OBJ = File('CMC_DLPFC_Dx_Sex_DiffExpression.tsv', 
+write.table(DE, file = 'CMC_ACC_Dx_Sex_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
+DEXP_OBJ = File('CMC_ACC_Dx_Sex_DiffExpression.tsv', 
                 name = 'Differential Expression Results (Dx_Sex)', 
@@
 
-# Store adjusted fit linear mixed model object as RDS - Dx_Sex
-saveRDS(ADJUSTED.FIT, file = 'CMC_DLPFC_Dx_Sex_dream-output.rds')
-ADJ_OBJ = File('CMC_DLPFC_Dx_Sex_dream-output.rds',
+# Store adjusted fit linear mixed model object as RDS - Dx
+saveRDS(ADJUSTED.FIT, file = 'CMC_ACC_Dx_Sex_dream-output.rds')
+ADJ_OBJ = File('CMC_ACC_Dx_Sex_dream-output.rds',
                name = 'Dream Output object - Dx_Sex', 
@@
 synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store variancePartition
+saveRDS(sortCols(violin_plot), file = 'CMC_ACC_Dx_Sex_variancePartition.rds')
+ADJ_OBJ = File('CMC_ACC_Dx_Sex_variancePartition.rds',
+               name = 'Variance Partition result', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store voom object
+saveRDS(VOOM.GENE_EXPRESSION, file = 'CMC_ACC_Dx_Sex_voom.rds')
+ADJ_OBJ = File('CMC_ACC_Dx_Sex_voom.rds',
+               name = 'voom object', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store contrasts
+saveRDS(L, file = 'CMC_ACC_Dx_Sex_contrasts.rds')
+ADJ_OBJ = File('CMC_ACC_Dx_Sex_contrasts.rds',
+               name = 'contrast matrix', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
+
+# Store formula
+saveRDS(formula, file = 'CMC_ACC_Dx_Sex_formula.rds')
+ADJ_OBJ = File('CMC_ACC_Dx_Sex_formula.rds',
+               name = 'formula', 
+               parentId = CODE$properties$id)
+ADJ_OBJ = synStore(ADJ_OBJ, used = ALL_USED_IDs, activityName = activityName,
+                   executed = thisFile, activityDescription = activityDescription)
+synSetAnnotations(ADJ_OBJ, annotations = all.annotations)
 ```

### 2020-08-25 individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_DLPFC_DE_DxSex.Rmd
stage interaction analysis

--- 2020-08-25 individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_DxSex.Rmd
@@
 ---
-title: "Differential Expression - HBCC - DLPFC - Dx - Sex"
+title: "Differential Expression - MSSM-Penn-Pitt - DLPFC - Dx - Sex"
 author: "Kelsey Montgomery, Gabriel Hoffman, Thanneer Perumal"
@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_DxSex.Rmd",
-                                 parentId ="syn22050019",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_DLPFC_DE_DxSex.Rmd",
+                                 parentId ="syn22043922",
                                  folderName = "Differential Expression - Dx - Sex")
@@
 library(githubr) # get the package from devtools::install_github('brian-bot/githubr')
+library(WGCNA)
 library(rctutils) # get the packages from BiocManager::install("DarwinAwardWinner/rctutils")
@@
 ```{r synapse.parameters, include=FALSE, cache=FALSE}
-parentId = 'syn22050019';
+parentId = 'syn22043922';
 activityName = 'Differential Expression';
-activityDescription = 'Differential Expression - Dx_Sex';
+activityDescription = 'Differential Expression - Dx - Sex';
 
-thisFileName <- 'HBCC_DLPFC_DE_DxSex.Rmd'
+thisFileName <- 'MSSM-Penn-Pitt_DLPFC_DE_DxSex.Rmd'
 
@@
 # Get covariates
-COVAR_ID = "syn22082458"
+COVAR_ID = "syn22045728"
 ALL_USED_IDs = c(ALL_USED_IDs, COVAR_ID)
-COVARIATES = downloadFile(COVAR_ID) %>% 
-  tibble::column_to_rownames(var = "SampleID")
+COVARIATES = downloadFile(COVAR_ID)
+
+# dropped individual id 
+MD_ID = "syn16816488"
+ALL_USED_IDs = c(ALL_USED_IDs, COUNTS_ID)
+METADATA = downloadFile(MD_ID) %>% 
+  dplyr::select(Individual_ID, Sample_RNA_ID)
+
+COVARIATES = right_join(METADATA, COVARIATES, by = c("Sample_RNA_ID" = "SampleID")) %>% 
+  tibble::column_to_rownames(var = "Sample_RNA_ID")
 
 # Set variable class
-FactorCovariates <- c("Individual_ID", "Reported_Gender", "LibraryBatch", "Dx", "RibozeroBatch", "FlowcellBatch")
+FactorCovariates <- c("Individual_ID", "Institution", "LibraryBatch", "Dx", "Reported_Gender", "RibozeroBatch", "FlowcellBatch")
 COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
@@
   x = as.numeric(as.character(gsub('[\\,\\%]','',x)))
-})
+})c
 
@@
 interaction <- c("Dx", "Reported_Gender")
-adjust.covars <- c("ageOfDeath", "RIN", "IntronicRate", "IntragenicRate", "IntergenicRate")
+adjust.covars <- c("ageOfDeath", "RIN2", "RIN", "IntronicRate", "IntragenicRate")
+random_effect <- c("Institution")
 ```
@@
 # Post adjusted formula
-formula <- formula(glue::glue("~ ", glue::glue_collapse(interaction, sep = "*")," + ", glue::glue_collapse(adjust.covars, sep = " + ")))
+formula <- formula(glue::glue("~ ", glue::glue_collapse(interaction, sep = "*")," + ", glue::glue_collapse(covars, sep = " + ")))
 design <- model.matrix(formula, COVARIATES)
@@
     "factor" = paste0('(1|', col_type$variable[i], ')'),
-    "numeric" = paste0('scale(', col_type$variable[i], ')')
-  )
+    "numeric" = col_type$variable[i])
   })
 
-# Manually specify interaction
-interaction <- c(interaction, "Dx:Reported_Gender")
-interaction_variance <- glue::glue("(1|{interaction})")
-all <- c(interaction_variance, formula_for_variance_fractions)
-
-formula_for_variance_fractions <- glue::glue("~ ", glue::glue_collapse(all, sep = " + "))
+formula_for_variance_fractions <- glue::glue("~ ", glue::glue_collapse(formula_for_variance_fractions, sep = " + "))
 
@@
                                    coefficient = c("Dx_SexControl_Male", "Dx_SexControl_Female"))
-L2 = c(1, -1, 0, 0, -1, 1, 0, 0, 0, 0, 0)
-L3 = c(1, -1, -1, 1, 0, 0, 0, 0, 0, 0, 0)
+L2 = c(1, -1, 0, 0, -1, 1, 0, 0, 0, 0)
+L3 = c(1, -1, -1, 1, 0, 0, 0, 0, 0, 0)
 

### 2020-08-25 individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_Sex.Rmd
integrate CQN using voomWithOffset - all cohorts

@@
 ```{r knit2synapse, eval=FALSE}
-ml git pandoc
-rmarkdown::render("individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_Sex.Rmd")
-
 library(synapser)
@@
 ### Normalisation - Sex
-1. Sex is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis)
+1. Sex is chosen as the primary variable of interest
 
@@
 # Post adjusted formula
-formula <- glue::glue("~ ", glue::glue_collapse(adjust.covars, sep = " + "))
+formula <- formula(glue::glue("~ ", glue::glue_collapse(adjust.covars, sep = " + ")))    
+design <- model.matrix(formula, COVARIATES)
 

### 2020-08-25 individual-cohorts/split_by_analysis/HBCC_dsACC_DE_Diagnosis.Rmd
integrate CQN using voomWithOffset - all cohorts

@@
                  glue::glue_collapse(adjust.covars, sep = ", ", last = " and "),
-                 ' as fixed effects and ',
-                 glue::glue_collapse(random_effect, sep = ", ", last = " and "),' chosen as random effects'))
+                 ' as fixed effects.'))
 
-scaled_col_type <- dplyr::select(COVARIATES, c(adjust.covars)) %>%
+scaled_col_type <- dplyr::select(COVARIATES, c(adjust.covars, random_effect)) %>%
   dplyr::summarise_all(class) %>% 
@@
 # Post adjusted formula
-formula <- glue::glue("~ ", glue::glue_collapse(scaled_adjust_covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))
+formula <- formula(glue::glue("~ ", glue::glue_collapse(scaled_adjust_covars, sep = " + ")))
+design <- model.matrix(formula, COVARIATES)
 

### 2020-08-25 individual-cohorts/split_by_analysis/HBCC_dsACC_DE_Sex.Rmd
integrate CQN using voomWithOffset - all cohorts

--- 2020-08-19 individual-cohorts/split_by_analysis/HBCC_dsACC_DE_Sex.Rmd
@@
 writeLines(glue::glue('Using following covariates in the model:',
-                 glue::glue_collapse(adjust.covars, sep = ", ", last = " and "),
-                 ' as fixed effects and ',
-                 glue::glue_collapse(random_effect, sep = ", ", last = " and "),' chosen as random effects'))
+                 glue::glue_collapse(c(adjust.covars,random_effect), sep = ", ", last = " and "),
+                 ' as fixed effects.'))
 
-scaled_col_type <- dplyr::select(COVARIATES, c(adjust.covars)) %>%
+col_type <- dplyr::select(COVARIATES, c(adjust.covars, random_effect)) %>%
   dplyr::summarise_all(class) %>% 
@@
 # Scale continuous variables
-scaled_adjust_covars = sapply(1:length(scaled_col_type$class), function(i){
-  switch(scaled_col_type$class[i], 
-    "factor" = paste0(scaled_col_type$variable[i]),
-    "numeric" = paste0('scale(', scaled_col_type$variable[i], ')')
+covars = sapply(1:length(col_type$class), function(i){
+  switch(col_type$class[i], 
+    "factor" = paste0(col_type$variable[i]),
+    "numeric" = paste0('scale(', col_type$variable[i], ')')
   )
@@
 # Post adjusted formula
-formula <- glue::glue("~ ", glue::glue_collapse(scaled_adjust_covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))
+formula <- formula(glue::glue("~", glue::glue_collapse(covars, sep = " + ")))
+design <- model.matrix(formula, COVARIATES)
 

### 2020-08-25 individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_ACC_DE_Diagnosis.Rmd
integrate CQN using voomWithOffset - all cohorts

--- 2020-08-19 individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_ACC_DE_Diagnosis.Rmd
@@
                  glue::glue_collapse(adjust.covars, sep = ", ", last = " and "),
-                 ' as fixed effects and',
-                 glue::glue_collapse(random_effect, sep = ", ", last = " and "),' chosen as random effects'))
+                 ' as fixed effects.'))
+
+col_type <- dplyr::select(COVARIATES, c(adjust.covars, random_effect)) %>%
+  dplyr::summarise_all(class) %>% 
+  tidyr::gather(variable, class)
+
+# Scale continuous variables
+covars = sapply(1:length(col_type$class), function(i){
+  switch(col_type$class[i], 
+    "factor" = paste0(col_type$variable[i]),
+    "numeric" = paste0('scale(', col_type$variable[i], ')')
+  )
+  })
 
 # Post adjusted formula
-formula <- glue::glue("~ ", glue::glue_collapse(adjust.covars, sep = " + "), " + ", glue::glue_collapse(glue::glue("(1|{random_effect})"), sep = " + "))
+formula <- formula(glue::glue("~", glue::glue_collapse(covars, sep = " + ")))
+design <- model.matrix(formula, COVARIATES)
 

### 2020-08-25 individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_ACC_DE_Sex.Rmd
integrate CQN using voomWithOffset - all cohorts

--- 2020-08-19 individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_ACC_DE_Sex.Rmd
@@
 library(biomaRt)
+library(rctutils) # get the packages from BiocManager::install("DarwinAwardWinner/rctutils")
 
@@
 ### Normalisation - Sex
-
-1. Sex is chosen as the primary variable of interest (i.e., covariate adjustments is conditioned on diagnosis)
+1. Sex is chosen as the primary variable of interest
 

### 2020-08-25 individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_DLPFC_DE_Diagnosis.Rmd
integrate CQN using voomWithOffset - all cohorts

--- 2020-08-19 individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_DLPFC_DE_Diagnosis.Rmd
@@
 library(WGCNA)
+library(rctutils) # get the packages from BiocManager::install("DarwinAwardWinner/rctutils")
 

### 2020-08-25 individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_DxSex.Rmd
remove samples based on age threshold - all cohorts

@@
   tibble::column_to_rownames(var = "SampleID")
+
+# Remove ages under 17
+COVARIATES <- COVARIATES[COVARIATES$ageOfDeath >= 17,]
+NEW.COUNTS <- NEW.COUNTS[names(NEW.COUNTS) %in% rownames(COVARIATES)]
+CQN.GENE_EXPRESSION <- CQN.GENE_EXPRESSION[names(CQN.GENE_EXPRESSION) %in% rownames(COVARIATES)] %>% 
+  data.matrix()
 

### 2020-08-25 individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_ACC_DE_Diagnosis.Rmd
remove samples based on age threshold - all cohorts

@@
   tibble::column_to_rownames(var = "Sample_RNA_ID")
+
+# Remove ages under 17
+COVARIATES <- COVARIATES[COVARIATES$ageOfDeath >= 17,]
+NEW.COUNTS <- NEW.COUNTS[names(NEW.COUNTS) %in% rownames(COVARIATES)]
+CQN.GENE_EXPRESSION <- CQN.GENE_EXPRESSION[names(CQN.GENE_EXPRESSION) %in% rownames(COVARIATES)] %>% 
+  data.matrix()
 

### 2020-08-25 individual-cohorts/split_by_analysis/HBCC_dsACC_DE_Diagnosis.Rmd
update provenance - all cohorts

@@
 
-thisFileName <- 'HBCC_dsACC_DE.Rmd'
+thisFileName <- 'HBCC_dsACC_DE_Diagnosis.Rmd'
 
 # Github link
-thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
-thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/',thisFileName))
+thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='modify-cqn-implementation')
+thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/split_by_analysis/',thisFileName))
 ```
@@
 counts = 'syn21867866'
-ALL_USED_IDs = counts
+ALL_USED_IDs = c(ALL_USED_IDs, counts)
 counts = downloadFile(counts) %>% data.frame()

### 2020-08-25 individual-cohorts/split_by_analysis/HBCC_dsACC_DE_Sex.Rmd
update provenance - all cohorts

@@
 activityName = 'Differential Expression';
-activityDescription = 'Differential Expression - ';
+activityDescription = 'Differential Expression - Sex';
 
-thisFileName <- 'HBCC_dsACC_DE.Rmd'
+thisFileName <- 'HBCC_dsACC_DE_Sex.Rmd'
 
 # Github link
-thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
-thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/',thisFileName))
+thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='modify-cqn-implementation')
+thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/split_by_analysis/',thisFileName))
 ```

### 2020-08-25 individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_ACC_DE_Diagnosis.Rmd
update provenance - all cohorts

@@
 
-thisFileName <- 'MSSM-Penn-Pitt_ACC_DE.Rmd'
+thisFileName <- 'MSSM-Penn-Pitt_ACC_DE_Diagnosis.Rmd'
 
 # Github link
-thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
-thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/',thisFileName))
+thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='modify-cqn-implementation')
+thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/split_by_analysis/',thisFileName))
 ```
@@
 counts = 'syn21867938'
-ALL_USED_IDs = counts
+ALL_USED_IDs = c(ALL_USED_IDs, counts)
 counts = downloadFile(counts) %>% data.frame()

### 2020-08-25 individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_ACC_DE_Sex.Rmd
update provenance - all cohorts

@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/MSSM-Penn-Pitt_ACC_DE.Rmd",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_ACC_DE_Sex.Rmd",
                                  parentId ="syn22012953",
@@
 
-thisFileName <- 'MSSM-Penn-Pitt_ACC_DE.Rmd'
+thisFileName <- 'MSSM-Penn-Pitt_ACC_DE_Sex.Rmd'
 
 # Github link
-thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
-thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/',thisFileName))
+thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='modify-cqn-implementation')
+thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/split_by_analysis/',thisFileName))
 ```

### 2020-08-25 individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_DLPFC_DE_Diagnosis.Rmd
update provenance - all cohorts

@@
 
-thisFileName <- 'MSSM-Penn-Pitt_DLPFC_DE.Rmd'
+thisFileName <- 'MSSM-Penn-Pitt_DLPFC_DE_Diagnosis.Rmd'
 
 # Github link
-thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
-thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/',thisFileName))
+thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='modify-cqn-implementation')
+thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/split_by_analysis/',thisFileName))
 ```

### 2020-08-25 individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_DLPFC_DE_DxSex.Rmd
update provenance - all cohorts

@@
 thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='modify-cqn-implementation')
-thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/split_by_analysis',thisFileName))
+thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/split_by_analysis/',thisFileName))
 ```
@@
 ### Source Code
-`r thisFile`
+[markdown file](`r thisFile`)
 

### 2020-08-25 individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_DLPFC_DE_Sex.Rmd
update provenance - all cohorts

@@
 
-thisFileName <- 'MSSM-Penn-Pitt_DLPFC_DE.Rmd'
+thisFileName <- 'MSSM-Penn-Pitt_DLPFC_DE_Sex.Rmd'
 
 # Github link
-thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
-thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/',thisFileName))
+thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='modify-cqn-implementation')
+thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/split_by_analysis/',thisFileName))
 ```

### 2020-08-26 individual-cohorts/split_by_analysis/HBCC_dsACC_DE_Diagnosis.Rmd
fix Provenance

@@
 writeLines(glue::glue('Using following covariates in the model:',
-                 glue::glue_collapse(adjust.covars, sep = ", ", last = " and "),
+                 glue::glue_collapse(c(adjust.covars,random_effect), sep = ", ", last = " and "),
                  ' as fixed effects.'))
@@
 ```
-
+`r glue::glue_collapse(formula)`
 ### Residual calculation

### 2020-08-26 individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_ACC_DE_Diagnosis.Rmd
fix Provenance

@@
 
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/MSSM-Penn-Pitt_ACC_DE.Rmd",
+knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_ACC_DE_Diagnosis.Rmd",
                                  parentId ="syn22012953",

### 2020-08-26 common_functions/common_functions.R
Add functions

@@ -1,13 +1,13 @@
-
+suppressPackageStartupMessages({
 library(qvalue)
-# library(EnsDb.Hsapiens.v86)
 library(AnnotationHub)
 library(ensembldb)
 library(MASS)
 library(ggplot2)
+library(ggbio)
 library(viridis)
 library(data.table)
-
+})
 # For ENSEMBL id ENSG00000279457.4, return ENSG00000279457
 trim_ensembl_ids = function(x){
   gsub("(.*)\\.(.*)", "\\1", x) 
@@ -62,7 +62,7 @@ plot_concordance = function( resList, showGenes = NULL, idx=c(1,2), size=15){
   lab = "log2 fold change"
   lim = with(df, max(abs(c(logFC.x, logFC.y))))
   df$density <- get_density(df$logFC.x, df$logFC.y, n = 100)
-  fig1 = ggplot(df, aes(logFC.x, logFC.y, color=density)) + geom_point(size=.4) + theme_bw(size) + theme(aspect.ratio=1, legend.position="bottom") + geom_abline(color="red") + xlab(paste(names(resList)[1], lab)) + ylab(paste(names(resList)[2], lab)) + geom_vline(xintercept=0, col="grey40", linetype="dashed") + geom_hline(yintercept=0, col="grey40", linetype="dashed") + xlim(-lim, lim) + ylim(-lim, lim) + scale_color_viridis() + geom_smooth(method="rlm", se=FALSE, color="darkorange")
+  fig1 = ggplot(df, aes(logFC.x, logFC.y, color=density)) + geom_point(size=.4) + theme_bw(size) + theme(aspect.ratio=1, legend.position="bottom") + geom_abline(color="red") + xlab(paste(names(resList)[1], lab)) + ylab(paste(names(resList)[2], lab)) + geom_vline(xintercept=0, col="grey40", linetype="dashed") + geom_hline(yintercept=0, col="grey40", linetype="dashed") + xlim(-lim, lim) + ylim(-lim, lim) + scale_color_viridis() + geom_smooth(method="lm", se=FALSE, color="darkorange")
 
   if( !is.null(showGenes) ){
     df2 = df[df$Symbol.x %in% showGenes,]
@@ -309,10 +309,13 @@ convert_gene_list_to_ensembl = function( x, geneInfo){
 }
 
 
[…]

### 2020-08-26 deconvolution/apply_dtangle_FANS4.Rmd
add upload

@@
                                  folderName = "Deconvolution results",
-                                 knitmd=FALSE)
+                                 knitmd=TRUE)
 ```
@@
 ```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
+suppressPackageStartupMessages({
 library(data.table)
@@
 library(BiocParallel)
-
+library(RNOmni)
+library(tidyr)
+library(plyr)
+library(dplyr)
+library(stringr)
 library(synapser)
@@
 library(githubr) # get the package from devtools::install_github('brian-bot/githubr')
-
+})
 source("../common_functions/common_functions.R")
@@
 ```{r read.data}
-
 # Read reference panel
@@
 file_counts_synid = c('syn21888969', 'syn21862455', 'syn21887889', 'syn21825508')
+ALL_USED_IDs = c('syn22321183', file_counts_synid)
 
@@
 
-cellFracEstimates2 = dtangle2( Y = t(geneExpr_logCPM), reference = t(geneExpr_FANS4_lcpm[idx,]))
+# cellFracEstimates2 = dtangle2( Y = t(geneExpr_logCPM), reference = t(geneExpr_FANS4_lcpm[idx,]))
 
-cor(cellFracEstimates$estimates, cellFracEstimates2$estimates, method="spearman")
+# cor(cellFracEstimates$estimates, cellFracEstimates2$estimates, method="spearman")
 ```
@@
 ```{r preprocess.data}
-# Get genotype ids from synapse
-GENOTYPE_ID = 'syn16816490'
-ALL_USED_IDs = c(ALL_USED_IDs, GENOTYPE_ID)
-GENOTYPE = downloadFile(GENOTYPE_ID) %>% 
-  dplyr::select(Individual_ID, `SNP_report:Genotyping_Sample_ID`, `SNP_report:Exclude?`) %>% 
-  dplyr::inner_join(ANCESTRY) %>% 
-  dplyr::filter(is.na(`SNP_report:Exclude?`))
+synId_covariates = c('syn21862454', 'syn21887888', 'syn21825507') # 'syn21888968'
 
-# Merge metadata
-METADATA = metadata %>%
-  dplyr::left_join(GENOTYPE) %>% 
-  dplyr::rename(Region = `rnaSeq_dissection:Brain_Region`,
-                PMI = `PMI_(in_hours)`,
-                RIN = `rnaSeq_isolation:RIN`, 
-                ReportExclude = `rnaSeq_report:Exclude?`,
-                GenotypeExclude = `SNP_report:Exclude?`,
-                SampleID = Sample_RNA_ID,
-                LibraryBatch = `rnaSeq_report:Library_Batch`,
-                FlowcellBatch = `rnaSeq_report:Flowcell_Batch`,
-                RibozeroBatch = `rnaSeq_report:Ribozero_Batch`,
-                MappedReads = `rnaSeq_report:Mapped_Reads`,
-                IntragenicRate = `rnaSeq_report:Intragenic_Rate`, IntronicRate = `rnaSeq_report:Intronic_Rate`, 
-                IntergenicRate = `rnaSeq_report:Intergenic_Rate`, GenesDetected = `rnaSeq_report:Genes_Detected`,
-                ExpProfEfficiency = `rnaSeq_report:Expression_Profiling_Efficiency`, rRNARate = `rnaSeq_report:rRNA_Rate`,
-                TotalReads = `rnaSeq_report:Total_Reads`, AlignmentRate = `rnaSeq_report:Percent_Aligned`, TranscriptsDetected = `rnaSeq_report:Transcripts_Detected`) %>%
-  dplyr::select(SampleID, Individual_ID, Institution, Cohort, Reported_Gender,Sex, Ethnicity, ageOfDeath, PMI, Dx, 
-                RIN, ReportExclude, GenotypeExclude, EV.1, EV.2, EV.3, EV.4, EV.5,
-                LibraryBatch, FlowcellBatch, RibozeroBatch, MappedReads, TotalReads, GenesDetected, AlignmentRate,
-                IntragenicRate, IntergenicRate, IntronicRate, ExpProfEfficiency, rRNARate, TranscriptsDetected) %>%
-  dplyr::filter(Cohort %in% "NIMH-HBCC") %>% 
-  dplyr::filter(SampleID %in% colnames(COUNT), !is.na(SampleID)) %>% 
-  dplyr::mutate(Dx = forcats::fct_recode(Dx, Other = "BP", Other = "undetermined", Control = "Control", SCZ = "SCZ"))
+COV_list = lapply(synId_covariates, function(id){
+	df = read.table( synGet(id)$path, header=TRUE, stringsAsFactors=FALSE)
+	df$SampleID = as.character(df$SampleID)
+	if( any(colnames(df) == "Dx.Tissue") ){
+		df$Dx = gsub("^(\\S+)_(\\S+)$", "\\1", df$Dx.Tissue)
+		df$Tissue = gsub("^(\\S+)_(\\S+)$", "\\2", df$Dx.Tissue)
+	}	
+	if( !any("Institution" == colnames(df)) ){
+		df$Institution = "HBCC"
 
-ind = METADATA$SampleID[which(METADATA$ReportExclude == 1 | METADATA$GenotypeExclude)]
-writeLines(paste('Following',length(ind),'samples are marked exclude'))
-writeLines(paste(ind, collapse = ', '))
-METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))
+	}
+	if( id == 'syn21862454' ){
+		df$Tissue = "DLPFC"
+	}
+	if( id == 'syn21887888' ){
+		df$Tissue = "sgACC"
+	}
+	if( !any('Tissue' == colnames(df))){
+		stop(id)
+	}
+	df
+	})
 
-ind = METADATA$SampleID [is.na(METADATA$Ethnicity) | is.na(METADATA$Institution) | is.na(METADATA$Dx)]
-writeLines(paste('Following', length(ind), 'counts are missing any metadata'))
-writeLines(paste(ind, collapse = ', '))
-METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))
+COVARIATES = gtools::smartbind( list = COV_list)
 
-ind = METADATA$SampleID [is.na(METADATA$PMI)]
-writeLines(paste('Following', length(ind), 'counts are missing PMI'))
-writeLines(paste(ind, collapse = ', '))
-METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))
+df = with(COVARIATES, data.frame(SampleID, Study = paste(Institution, Tissue)))
+df$Study = gsub("Penn|Pitt", "MSSM", df$Study)
 
-ind = METADATA$SampleID [is.na(METADATA$Reported_Gender)]
-writeLines(paste('Following', length(ind), 'counts are missing gender'))
-writeLines(paste(ind, collapse = ', '))
-METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))
+ids = unique(df$Study)
+sampleByStudy = lapply( ids, function(stdy){
+	df[df$Study == stdy, "SampleID"]
+	})
+names(sampleByStudy) = ids
 
-ind = METADATA$SampleID [is.na(METADATA$ageOfDeath)]
-writeLines(paste('Following', length(ind), 'counts are missing age of death'))
-writeLines(paste(ind, collapse = ', '))
-METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))
+# only keep individuals older than 17
+COVARIATES = COVARIATES[COVARIATES$ageOfDeath > 17,]
 
-ind = METADATA$SampleID [is.na(METADATA$EV.1)]
-writeLines(paste('Following', length(ind), 'counts are missing ancestry information'))
-writeLines(paste(ind, collapse = ', '))
-METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))
 ```
 
-```{r preprocess.data1, results='asis'}
-# Match covariates to expression data
-indToRetain = intersect(METADATA$SampleID, colnames(COUNT))
+```{r deconv.compare.covariates}
 
-rownames(COUNT) <- COUNT$Geneid
-COUNT = COUNT[,indToRetain]
+deconvEstimates = data.frame(cellFracEstimates$estimates)
 
-rownames(METADATA) = METADATA$SampleID
-METADATA = METADATA[indToRetain,]
+f = log
+# f = rankNorm
 
-METADATA %>% 
-  group_by(Dx) %>% 
-  dplyr::summarise(count = n()) %>% 
-  kable()
+df_deconv_test = lapply( names(sampleByStudy) , function(stdy){
 
-FactorCovariates <- c('Individual_ID', "Reported_Gender", "LibraryBatch", "Dx", "FlowcellBatch", "RibozeroBatch")
-ContCovariates <- c("ageOfDeath", "PMI", "RIN", "MappedReads", "IntragenicRate", "IntronicRate", "IntergenicRate",
-                    "GenesDetected", "ExpProfEfficiency", "rRNARate", "TotalReads", "AlignmentRate",
-                    "EV.1", "EV.2", "EV.3", "EV.4", "EV.5")
+	smplids = sampleByStudy[[stdy]]
 
-# Find inter relation between factor covariates
-COVARIATES = METADATA[,c(FactorCovariates,ContCovariates),drop=F]
-COVARIATES[,FactorCovariates] <- data.frame(lapply(COVARIATES[,FactorCovariates],function(x){
-  x <- sapply(x,function(y){str_replace_all(as.character(y),'[^[:alnum:]]','_')})}))
-rownames(COVARIATES) <- METADATA$SampleID
+	deconvEst = deconvEstimates[smplids,]
 
-# Convert factor covariates to factors
-COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
-COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], function(x){
-  x = as.numeric(as.character(gsub('[\\,\\%]','',x)))
+	df = merge(deconvEst, COVARIATES[COVARIATES$SampleID %in% smplids,], by.x="row.names", by.y="SampleID")
+
+	form = cbind(GABA 	= f(GABA), 
+				GLU 	= f(GLU), 
+				Olig 	= f(Olig),
+				MgAs 	= f(MgAs)) ~ ageOfDeath + Reported_Gender + PMI + RIN + RIN2 + Dx 
+	fit = lm(form, df)
+
+	res = lapply( summary(fit), function(x) data.frame(p = -log10(coef(x)[-1,4]) * sign(coef(x)[-1,1]) ))
+
+	df_return = do.call(cbind, res)
+	colnames(df_return) = gsub("Response ", "", names(res))
+	data.frame(Study = stdy, coef = rownames(df_return), df_return)
+
 	})
 
-```{r store, eval=FALSE}
+df_deconv_test = do.call(rbind, df_deconv_test)
+
+df_deconv_test_melt = reshape2::melt(df_deconv_test, id.vars=c("Study", "coef"))
+n_tests = nrow(df_deconv_test_melt)
+df_deconv_test_melt$isSignif = with(df_deconv_test_melt, abs(value) > -log10(0.05 / n_tests))
+
+lim = max(abs(df_deconv_test_melt$value))
+
+ratio = with(df_deconv_test_melt, length(unique(coef)) / length(unique(Study)))
+
+ggplot(df_deconv_test_melt, aes(Study, coef, fill=value, label=ifelse(isSignif, '*', ''))) + geom_tile() + theme_bw() + theme(aspect.ratio = ratio, panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.text.x=element_text(angle=45, hjust=1)) + scale_fill_gradient2(high = "red", mid = "white", low="blue", limits=c(-lim, lim), name="-log10 p") + facet_wrap( ~ variable) + geom_text()
+```
+
+```{r combine.samples}
+
+df = merge(deconvEstimates, COVARIATES, by.x="row.names", by.y="SampleID")
+
+form = ~ (1|Individual_ID) + ageOfDeath + (1|Reported_Gender) + PMI + RIN + (1|Dx) + (1|Tissue)
+vp = fitExtractVarPartModel( t(log(df[,colnames(deconvEstimates)])), form, df, BPPARAM=SerialParam())
+
+plotPercentBars( vp )
+```
+
+```{r cca, fig.width=5, fig.height=5}
+form = ~ Individual_ID + ageOfDeath + Reported_Gender + PMI + RIN + Dx + Tissue
+C = canCorPairs(form, df)
+plotCorrMatrix(C)
+```
+
+```{r store, eval=TRUE}
 # Code
-CODE <- Folder(name = "Meta-analysis on sex", parentId = "syn22315403") # Make sure these match the folder name and parent Id in the first chunk.
+CODE <- Folder(name = "Deconvolution results", parentId = "syn22321035") # Make sure these match the folder name and parent Id in the first chunk.
 CODE <- synStore(CODE)
@@
 # Store meta-analysis
-file = "cellFracEstimates.RDS"
-saveRDS(cellFracEstimates, file=file)
+file = "cellFracEstimates_dtangle_FANS4.tsv"
+# saveRDS(data.frame(cellFracEstimates$estimates), file=file)
+write.table(cellFracEstimates$estimates, quote=FALSE, sep='\t', file=file)
+
 nEXP_OBJ = File(file, 
-                name = 'Result of dtangle', 
-                parentId = CODE$properties$id)
-nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, 
-                    executed = thisFile)
-res_acc %>%
-  tibble::rownames_to_column(var = "gene_id") %>% 
-  write.table(file = 'CMC_acc_Sex_meta_analysis.tsv', sep = '\t', row.names=F, quote=F)
-nEXP_OBJ = File('CMC_acc_Sex_meta_analysis.tsv', 
-                name = 'Sex Normalised, meta-analysis (ACC)', 
+                name = 'Estimates from dtangle', 
                 parentId = CODE$properties$id)

### 2020-08-26 deconvolution/apply_dtangle_FANS4.Rmd
add upload

@@
 geneExpr = calcNormFactors(geneExpr)
-geneExpr_logCPM = cpm(geneExpr, log=TRUE)
+geneExpr_logCPM = edgeR::cpm(geneExpr, log=TRUE)
 
@@
 synId_covariates = c('syn21862454', 'syn21887888', 'syn21825507') # 'syn21888968'
+ALL_USED_IDs = c(ALL_USED_IDs, synId_covariates)
 
@@
                 parentId = CODE$properties$id)
-nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, 
+nEXP_OBJ = synStore(nEXP_OBJ, used = ALL_USED_IDs, 
                     executed = thisFile)

### 2020-08-26 testing/test_CQN_HBCC.Rmd
initial commit
fix plot labels

@@
 # Specify which variable to test
-coef = "DxSCZ"
-# coef = "Reported_GenderMale"
+# coef = "DxSCZ"
+coef = "Reported_GenderMale"
 
@@
 
-plot_grid(fig2, fig2a, fig2, ncol=3)
+plot_grid(fig2, fig2a, fig2b, ncol=3)
 ```
@@
 
-fig2a = ggplot(df, aes(percentage_gene_gc_content, pmin(1, pmax(-1,logFC)))) + geom_point() + theme_bw() + geom_smooth(method="loess", se=FALSE) + ggtitle("NEW.COUNTS + CQN (offset)") + ylab("logFC")  + theme(aspect.ratio=1)
+fig3a = ggplot(df, aes(percentage_gene_gc_content, pmin(1, pmax(-1,logFC)))) + geom_point() + theme_bw() + geom_smooth(method="loess", se=FALSE) + ggtitle("NEW.COUNTS + CQN (offset)") + ylab("logFC")  + theme(aspect.ratio=1)
 
-fig2b = ggplot(df, aes(log10(Length), pmin(1, pmax(-1,logFC)))) + geom_point() + theme_bw() + geom_smooth(method="loess", se=FALSE) + ggtitle("NEW.COUNTS + CQN (offset)") + ylab("logFC") + theme(aspect.ratio=1)
+fig3b = ggplot(df, aes(log10(Length), pmin(1, pmax(-1,logFC)))) + geom_point() + theme_bw() + geom_smooth(method="loess", se=FALSE) + ggtitle("NEW.COUNTS + CQN (offset)") + ylab("logFC") + theme(aspect.ratio=1)
 
-plot_grid(fig2, fig2a, fig2b, ncol=3)
+plot_grid(fig3, fig3a, fig3b, ncol=3)
 ```

### 2020-08-28 individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_ACC_DE_Sex.Rmd
update ensembl to functional mirror and fix formula output format

@@
 mart <- biomaRt::useMart(biomart = "ENSEMBL_MART_ENSEMBL",
-                host = "uswest.ensembl.org", # Ensembl Release 99 (January 2020)
+                host = "ensembl.org", # Ensembl Release 99 (January 2020)
                 dataset = "hsapiens_gene_ensembl")

### 2020-08-28 individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_DLPFC_DE_Sex.Rmd
update ensembl to functional mirror and fix formula output format

@@
 ```
+`r glue::glue_collapse(formula)`
 

### 2020-08-31 individual-cohorts/meta_analysis/meta-analysis_sex.Rmd
Freeze analysis

@@
                                           folderName = 'Meta-analysis on sex',
-                                          overwrite=TRUE)
+                                          overwrite=TRUE, 
+                                          knitmd=TRUE)
 ```
@@
 Ids <- c("syn22278494", "syn22278597", "syn22280959", "syn22280588")
+```
+
+```{r plotma}
+
+f = function(x){
+  pmin(1, pmax(-1, x))
+}
+
+ggplot(mpp_acc, aes(AveExpr, f(logFC))) + geom_point() + theme_bw() + geom_smooth(method="loess")  + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + ggtitle("mpp_acc")
+
+ggplot(mpp_dlpfc, aes(AveExpr, f(logFC))) + geom_point() + theme_bw() + geom_smooth(method="loess")  + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + ggtitle("mpp_dlpfc")
+
+ggplot(hbcc_dsACC, aes(AveExpr, f(logFC))) + geom_point() + theme_bw() + geom_smooth(method="loess")  + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + ggtitle("hbcc_dsACC")
+
+ggplot(hbcc_dlpfc, aes(AveExpr, f(logFC))) + geom_point() + theme_bw() + geom_smooth(method="loess")  + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + ggtitle("hbcc_dlpfc")
 ```
@@
 ### Combine cohorts
-```{r DE.results.dlpfc.combine, fig.width=10, fig.height=50}
+```{r DE.results.dlpfc.combine.MA}
 res = merge(res_dlpfc, mpp_dlpfc[,c("gene_id", "AveExpr")], by.x="row.names", by.y="gene_id")
 
+ggplot(res, aes(AveExpr, logFC)) + geom_point() + theme_bw() + geom_smooth(method="loess")  + theme(aspect.ratio=1)
+```
+
+```{r DE.results.dlpfc.hbcc.plot_sex_compare, fig.width=10, fig.height=50}
 figList = plot_sex_compare(res)
 plot_grid(plotlist=figList, ncol=2) 
-
 ```

### 2020-09-01 individual-cohorts/meta_analysis/meta-analysis_diagnosis.Rmd
Freeze analysis

@@
 ```{r knit2synapse, eval=FALSE, echo=FALSE}
-# rmarkdown::render("meta-analysis_diagnosis.Rmd")
+# rmarkdown::render("individual-cohorts/meta_analysis/meta-analysis_diagnosis.Rmd")
 
@@
 synLogin()
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/meta_analysis/meta-analysis_diagnosis.Rmd",
+# knit2synapse::
+
+createAndKnitToFolderEntity <- function(file, parentId, folderName,
+                                        wikiName=NULL, overwrite=FALSE, knitmd=TRUE, ...) {
+
+  entity <- synapser::Folder(parentId=parentId, name=folderName)
+  entity <- synapser::synStore(entity, ...)
+
+  knitfile2synapse(file=file, owner=entity, wikiName=wikiName, 
+                   overwrite=overwrite, knitmd=knitmd, parentWikiId=NULL)
+}
+
+createAndKnitToFolderEntity(file = "individual-cohorts/meta_analysis/meta-analysis_diagnosis.Rmd",
                                           parentId ="syn22315403",
                                           folderName = 'Meta-analysis on diagnosis',
-                                          overwrite=TRUE)
+                                          overwrite=TRUE, 
+                                          knitmd=TRUE)
 ```

### 2020-09-01 individual-cohorts/meta_analysis/meta-analysis_sex.Rmd
Freeze analysis

@@
 
-```{r knit2synapse, eval=FALSE}
+```{r knit2synapse, eval=FALSE, echo=FALSE}
+
+createAndKnitToFolderEntity <- function(file, parentId, folderName,
+                                        wikiName=NULL, overwrite=FALSE, knitmd=TRUE, ...) {
+
+  entity <- synapser::Folder(parentId=parentId, name=folderName)
+  entity <- synapser::synStore(entity, ...)
+
+  knitfile2synapse(file=file, owner=entity, wikiName=wikiName, 
+                   overwrite=overwrite, knitmd=knitmd, parentWikiId=NULL)
+}
+
 library(synapser)
@@
 synLogin()
-knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/meta_analysis/meta-analysis_sex.Rmd",
+# knit2synapse::
+createAndKnitToFolderEntity(file = "individual-cohorts/meta_analysis/meta-analysis_sex.Rmd",
                                           parentId ="syn22315403",

### 2020-09-01 deconvolution/example_linking_FANS4_to_bulk.R
Initial commit

@@ -1,0 +1,71 @@
+# Gabriel Hoffman
+#
+# August 31, 2020
+#
+# Link FANS4 identifier to Individual_ID for genotype and Sample_RNA_ID for bulk RNA-seq
+
+library(data.table)
+library(openxlsx)
+library(dplyr)
+library(synapser)
+
+synLogin()
+
+# Get meta-data for FANS4 data
+###############################
+
+file = synGet('syn22321012')$path
+df_featureCounts_gene = fread( file )
+geneCounts = df_featureCounts_gene[,7:ncol(df_featureCounts_gene)]
+
+# read metadata
+file = synGet('syn22321013')$path
+metadata_FANS4 = read.xlsx( file )
+
+# extract information from sample names
+info = data.frame( Name = colnames(geneCounts), stringsAsFactors=FALSE)
+rownames(info) = info$Name
+info$DissectionID = sapply(strsplit( info$Name, '\\.'), function(x) x[1])
+info$CellType = sapply(strsplit( info$Name, '\\.'), function(x) x[2])
+info$Assay = sapply(strsplit( info$Name, '\\.'), function(x) x[3])
+
+# set color codes for each cell type
+colorCodes = c(
+  GABA = '#66A61E',
+  GLU = '#E6AB02',
+  Olig = '#E7298A',
+  MgAs = '#7570B3')
+
+# set factor levels to be the same as the order of the colors
+# this makes plotting easier
+info$CellType = factor(info$CellType,names(colorCodes))
+
+# add information about each sample
+info = merge(info, metadata_FANS4, by.x="DissectionID", by.y='Dissection.ID')
+rownames(info) = info$Name
+
+# set order to be the same as geneCounts
+info = info[match(colnames(geneCounts), info$Name),]
+
+# Intersect with BULK RNA-seq
+#############################
+
+# Get bulk RNA-seq counts
+file = synGet('syn22045729')$path
+geneCountsBulk = data.frame(fread( file ))
+rownames(geneCountsBulk) = geneCountsBulk[,1]
+geneCountsBulk = data.frame(geneCountsBulk[,-1])
+
+# get link between Individual_ID and Sample_RNA_ID 
+METADATA = fread(synGet('syn16816488')$path , data.table=FALSE) %>% 
+	dplyr::select(Individual_ID, Sample_RNA_ID)
+
+info2 = merge(info, METADATA, by.x="Individual.ID", by.y="Individual_ID")
+
[…]

### 2020-09-02 individual-cohorts/meta_analysis/meta-analysis_interaction.Rmd
Initial commit

--- 2020-09-01 individual-cohorts/meta_analysis/meta-analysis_diagnosis.Rmd
@@
 ---
-title: "Meta-analysis on diagnosis in MSSM-Penn-Pitt and HBCC ACC/DLPFC"
+title: "Meta-analysis on Sex*Dianosis interaction in MSSM-Penn-Pitt and HBCC ACC/DLPFC"
 author: "KM and GH"
@@
 ```{r knit2synapse, eval=FALSE, echo=FALSE}
-# rmarkdown::render("individual-cohorts/meta_analysis/meta-analysis_diagnosis.Rmd")
+# rmarkdown::render("individual-cohorts/meta_analysis/meta-analysis_interaction.Rmd")
 
@@
 
-createAndKnitToFolderEntity(file = "individual-cohorts/meta_analysis/meta-analysis_diagnosis.Rmd",
+createAndKnitToFolderEntity(file = "individual-cohorts/meta_analysis/meta-analysis_interaction.Rmd",
                                           parentId ="syn22315403",
-                                          folderName = 'Meta-analysis on diagnosis',
+                                          folderName = 'Meta-analysis on Sex*Diagnosis interaction',
                                           overwrite=TRUE, 
@@
 ```{r synapse.parameters, include=FALSE, cache=FALSE, eval=TRUE}
-thisFileName <- 'meta-analysis_diagnosis.Rmd'
+thisFileName <- 'meta-analysis_interaction.Rmd'
 
@@
 ```{r input_data}
-mpp_acc <- readr::read_tsv(synGet("syn22079117")$path) %>% 
+mpp_acc <- readr::read_tsv(synGet("syn22276990")$path) %>% 
   dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% getGeneSymbol( 'gene_id' )
 
-mpp_dlpfc <- readr::read_tsv(synGet("syn22079472")$path) %>% 
+mpp_dlpfc <- readr::read_tsv(synGet("syn22277179")$path) %>% 
   dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% getGeneSymbol( 'gene_id' )
@@
 
-hbcc_dlpfc <- readr::read_tsv(synGet("syn22084205")$path) %>% 
+hbcc_dlpfc <- readr::read_tsv(synGet("syn22280364")$path) %>% 
   dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% getGeneSymbol( 'gene_id' )
-Ids <- c("syn22079117", "syn22079472", "syn22086829", "syn22084205")
+
+Ids <- c("syn22276990", "syn22277179", "syn22086829", "syn22280364")
 ```
@@
 
-```{r store, eval=TRUE}
+```{r store, eval=FALSE}
 # Code
-CODE <- Folder(name = "Meta-analysis on diagnosis", parentId = "syn22315403") # Make sure these match the folder name and parent Id in the first chunk.
+CODE <- Folder(name = "Meta-analysis on Sex*Diagnosis interaction", parentId = "syn22315403") # Make sure these match the folder name and parent Id in the first chunk.
 CODE <- synStore(CODE)
@@
   tibble::rownames_to_column(var = "gene_id") %>% 
-  write.table(file = 'CMC_dlpfc_diagnosis_meta_analysis.tsv', sep = '\t', row.names=F, quote=F)
-nEXP_OBJ = File('CMC_dlpfc_diagnosis_meta_analysis.tsv', 
-                name = 'diagnosis Normalised, meta-analysis (DLPFC)', 
+  write.table(file = 'CMC_dlpfc_interaction_meta_analysis.tsv', sep = '\t', row.names=F, quote=F)
+nEXP_OBJ = File('CMC_dlpfc_interaction_meta_analysis.tsv', 
+                name = 'Interaction meta-analysis (DLPFC)', 
                 parentId = CODE$properties$id)
@@
   tibble::rownames_to_column(var = "gene_id") %>% 
-  write.table(file = 'CMC_acc_diagnosis_meta_analysis.tsv', sep = '\t', row.names=F, quote=F)
-nEXP_OBJ = File('CMC_acc_diagnosis_meta_analysis.tsv', 
-                name = 'diagnosis Normalised, meta-analysis (ACC)', 
+  write.table(file = 'CMC_acc_Interaction_meta_analysis.tsv', sep = '\t', row.names=F, quote=F)
+nEXP_OBJ = File('CMC_acc_Interaction_meta_analysis.tsv', 
+                name = 'Interaction meta-analysis (ACC)', 
                 parentId = CODE$properties$id)

### 2020-09-04 combined_analysis/combine_all_data.Rmd
Add joint analysis

---
title: "Joint analysis of read count data"
subtitle: ''
author: "Developed by [Gabriel Hoffman](http://gabrielhoffman.github.io/)"
date: "Run on `r Sys.time()`"
documentclass: article
output: 
  html_document:
  toc: true
  smart: false
---

<!---

cd /hpc/users/hoffmg01/work/cmc_analysis/src
# rm -rf molec_profiling_cache/
ml git pandoc
git pull
R

system("git pull")

rmarkdown::render("combined_analysis/combine_all_data.Rmd")

--->

--->

```{r knit2synapse, eval=FALSE}
library(synapser)
library(knit2synapse)
synLogin()
knit2synapse::createAndKnitToFolderEntity(file = "combined_analysis/combine_all_data.Rmd",
                                          parentId ="syn22231629",
                                          folderName = 'Joint analysis of count data',
                                          overwrite=TRUE, 
                                          knitmd=TRUE)
```

GSEA and dream from raw counts

```{r setup, include=FALSE}
suppressPackageStartupMessages({
library(synapser)

synLogin()

library(ggplot2)
library(dplyr)
library(ComplexHeatmap)
library(readxl)
library(gtools)
library(corrplot)
library(pvclust)
library(metafor)
library(data.table)
library(cowplot)
library(kableExtra)
library(githubr)
library(ggrepel)
library(knitr)
library(patchwork)
library(plyr)
library(BiocParallel)

library(limma)
library(edgeR)
library(variancePartition)

})
source("../common_functions/common_functions.R")

knitr::opts_chunk$set(
  echo = FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r read.counts}
# Read raw counts from MPP and HBCC
counts.MPP = fread(synGet('syn21867938')$path, data.table=FALSE)
rownames(counts.MPP) = counts.MPP$Geneid
counts.MPP = counts.MPP[,-c(1:6)]

# HBCC - DLFPC
counts.HBCC = fread(synGet('syn21886235')$path, data.table=FALSE)
rownames(counts.HBCC) = counts.HBCC$Geneid
counts.HBCC = counts.HBCC[,-c(1:6)]

# HBCC - sgACC
counts.HBCC4 = fread(synGet('syn21867866')$path, data.table=FALSE)
rownames(counts.HBCC4) = counts.HBCC4$Geneid
counts.HBCC4 = counts.HBCC4[,-c(1:6)]

geneCountsMerged = cbind(counts.MPP, counts.HBCC, counts.HBCC4)
```

### Data download
#### Obtain effective count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
downloadFile <- function(id){
  fread(synGet(id)$path, data.table = F)
}
downloadFile_version <- function(id , version){
  fread(synGet(id, version = version)$path, data.table = F)
}

# Get RNASeq QCmetadata
METADATA_QC_DLPFC_ID = 'syn16816488' 
ALL_USED_IDs = METADATA_QC_DLPFC_ID
metadata = downloadFile_version(METADATA_QC_DLPFC_ID, version = 19) %>%
  dplyr::select(`Individual_ID`, Institution, Cohort, `Reported_Gender`, Sex, Ethnicity, ageOfDeath, `PMI_(in_hours)`, Dx, 
                Sample_RNA_ID, one_of('rnaSeq_isolation:RIN',
                                      'rnaSeq_dissection:Brain_Region', 
                                      'rnaSeq_isolation:RNA_Isolation_Batch',
                                      'rnaSeq_report:Ribozero_Batch', 'rnaSeq_report:Library_Batch', 
                                      'rnaSeq_report:Flowcell_Batch', 'rnaSeq_report:Exclude?',
                                      'rnaSeq_report:Mapped_Reads',
                                      'rnaSeq_report:Intragenic_Rate','rnaSeq_report:Intronic_Rate',
                                      'rnaSeq_report:Intergenic_Rate','rnaSeq_report:Genes_Detected',
                                      'rnaSeq_report:Expression_Profiling_Efficiency',
                                      'rnaSeq_report:rRNA_Rate',
                                      'rnaSeq_report:Total_Reads', 'rnaSeq_report:Percent_Aligned','rnaSeq_report:Transcripts_Detected'))

# Get ancestry vector calculated using gemtools
ANCESTRY_ID = 'syn9922992'
ALL_USED_IDs = c(ALL_USED_IDs, ANCESTRY_ID)
ANCESTRY.HBCC = downloadFile(ANCESTRY_ID) %>% 
  plyr::rename(c(ID = 'SNP_report:Genotyping_Sample_ID'))

ANCESTRY_ID = 'syn2511399'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = ANCESTRY_ID
ANCESTRY.MPP = downloadFile(ANCESTRY_ID) %>% 
  plyr::rename(c('DNA_report..Genotyping.Sample_ID' = 'SNP_report:Genotyping_Sample_ID'))

ANCESTRY = rbind(ANCESTRY.HBCC[,colnames(ANCESTRY.MPP)[-2]], ANCESTRY.MPP[,colnames(ANCESTRY.MPP)[-2]])

# Get genotype ids from synapse. 
GENOTYPE_ID = 'syn16816490'
ALL_USED_IDs = c(ALL_USED_IDs, GENOTYPE_ID)
GENOTYPE = downloadFile(GENOTYPE_ID) %>% 
  dplyr::select(Individual_ID, `SNP_report:Genotyping_Sample_ID`, `SNP_report:Exclude?`) %>% 
  dplyr::inner_join(ANCESTRY) %>% 
  dplyr::filter(is.na(`SNP_report:Exclude?`))
```

### Data preprocessing
```{r preprocess.data}
# Merge metadata
METADATA = metadata %>%
  dplyr::left_join(GENOTYPE) %>%      
  dplyr::rename(Region = `rnaSeq_dissection:Brain_Region`,
                PMI = `PMI_(in_hours)`,
                RIN = `rnaSeq_isolation:RIN`, 
                ReportExclude = `rnaSeq_report:Exclude?`,
                GenotypeExclude = `SNP_report:Exclude?`,
                SampleID = Sample_RNA_ID,
                LibraryBatch = `rnaSeq_report:Library_Batch`,
                FlowcellBatch = `rnaSeq_report:Flowcell_Batch`,
                IsolationBatch = `rnaSeq_isolation:RNA_Isolation_Batch`, #isolation batch to replace ribozero batch
                RibozeroBatch = `rnaSeq_report:Ribozero_Batch`,
                MappedReads = `rnaSeq_report:Mapped_Reads`,
                IntragenicRate = `rnaSeq_report:Intragenic_Rate`, IntronicRate = `rnaSeq_report:Intronic_Rate`, 
                IntergenicRate = `rnaSeq_report:Intergenic_Rate`, GenesDetected = `rnaSeq_report:Genes_Detected`,
                ExpProfEfficiency = `rnaSeq_report:Expression_Profiling_Efficiency`, rRNARate = `rnaSeq_report:rRNA_Rate`,
                TotalReads = `rnaSeq_report:Total_Reads`, AlignmentRate = `rnaSeq_report:Percent_Aligned`, TranscriptsDetected = `rnaSeq_report:Transcripts_Detected`) %>%
  dplyr::select(SampleID, Individual_ID, Institution, Cohort, Reported_Gender,Sex, Ethnicity, ageOfDeath, PMI, Dx, Region,
                RIN, ReportExclude, GenotypeExclude, EV.1, EV.2, EV.3, EV.4, EV.5,
                LibraryBatch, FlowcellBatch, IsolationBatch, RibozeroBatch, MappedReads, TotalReads, GenesDetected, AlignmentRate,
                IntragenicRate, IntergenicRate, IntronicRate, ExpProfEfficiency, rRNARate, TranscriptsDetected) %>%
  		 		 dplyr::filter(SampleID %in% colnames(geneCountsMerged), !is.na(SampleID)) %>%
  		 		 dplyr::filter(Region %in% c('DLPFC', 'ACC')) %>%
			 	 dplyr::mutate(Cohort = forcats::fct_recode(Cohort, `MSSM-Penn-Pitt` = 'MSSM-Penn-Pitt', `NIMH-HBCC`='NIMH-HBCC', `NIMH-HBCC`='NIMH-HBCC-4')) %>%
			 	 dplyr::mutate(Dx_crude = forcats::fct_recode(Dx, Other = "AFF", Other = "BP", Other = "undetermined", Control = "Control", SCZ = "SCZ"))

ind = METADATA$SampleID[which(METADATA$ReportExclude == 1 | METADATA$GenotypeExclude)]
writeLines(paste('Following',length(ind),'samples are marked exclude'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$Ethnicity) | is.na(METADATA$Institution) | is.na(METADATA$Dx)]
writeLines(paste('Following', length(ind), 'counts are missing any metadata'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$PMI)]
writeLines(paste('Following', length(ind), 'counts are missing PMI'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$Reported_Gender)]
writeLines(paste('Following', length(ind), 'counts are missing gender'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$ageOfDeath)]
writeLines(paste('Following', length(ind), 'counts are missing age of death'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$EV.1)]
writeLines(paste('Following', length(ind), 'counts are missing ancestry information'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))
```

```{r preprocess.data1, results='asis'}
# Match covariates to expression data
indToRetain = intersect(METADATA$SampleID, colnames(geneCountsMerged))

geneCountsMerged = geneCountsMerged[,indToRetain]

rownames(METADATA) = METADATA$SampleID
METADATA = METADATA[indToRetain,]

METADATA %>% 
  group_by(Dx) %>%  
  dplyr::summarise(count = n()) %>% 
  kable() 
```

```{r process.counts}

# keep genes that are express in at least one combination of cohort and gender
isexpr.lst = dlply(METADATA, .(Cohort, Reported_Gender), .fun = function(mtd){
	rowSums(cpm(geneCountsMerged[,mtd$SampleID])>1) >= 0.5*ncol(geneCountsMerged[,mtd$SampleID])
	})

isexpr = do.call(rbind, isexpr.lst)
isexpr = colSums(isexpr) > 0
table(isexpr)

# normalize counts
dge = DGEList( geneCountsMerged[isexpr,] )
dge = calcNormFactors(dge) # TMM

xtabs(~ Region + Cohort, METADATA)
xtabs(~ Institution + Cohort, METADATA)

# FORMULAs

# HBCC - DLPFC: 
# "Dx + scale(RIN) + ageOfDeath + (1 | Reported_Gender) + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate)"

# HBCC dsACC: 
# Dx + scale(RIN) + (1 | Reported_Gender) + ageOfDeath + (1 | IsolationBatch) + scale(IntragenicRate) + scale(IntronicRate) + scale(IntergenicRate)"

# MPP ACC
 # Dx + scale(RIN) + scale(ageOfDeath) + (1 | Institution) + (1 | Reported_Gender) + scale(RIN2) + scale(IntronicRate) + scale(rRNARate) + scale(IntragenicRate) + scale(IntergenicRate)

# MPP DLPFC
# Dx + scale(RIN2) + (1 | Institution) + scale(ageOfDeath) + scale(RIN) + (1 | Reported_Gender) + scale(IntronicRate) + scale(IntragenicRate)

form = ~ Dx + Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Region:Cohort) + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)

vobj = voomWithDreamWeights(dge, form, METADATA, save.plot=TRUE, BPPARAM = SnowParam(8))

plot_voom(vobj)
```

### 2020-09-04 common_functions/common_functions.R
add plot_voom

@@ -96,14 +96,42 @@ plot_concordance = function( resList, showGenes = NULL, idx=c(1,2), size=15){
       NA
       })
 
+
   tab_pi = data.frame(Discovery = names(resList), pi1 = c(pi_discovery_x, pi_discovery_y))
 
   tab_corr = data.frame(R_spearman = res$estimate, P = res$p.value)
 
+  # pi spectrum
+  df_order_x = df[order(df$P.Value.x),]
+  df_order_y = df[order(df$P.Value.y),]
+
+  res_p1_ramp = lapply(1:nrow(df), function(i){
+
+    pi1.x = tryCatch( 
+      1 - qvalue(df_order_x$P.Value.y[1:i])$pi0,
+      error = function(e){
+        NA
+        })
+
+    pi1.y = tryCatch( 
+      1 - qvalue(df_order_y$P.Value.x[1:i])$pi0,
+      error = function(e){
+        NA
+        })
+    res = data.frame( i     = i, 
+                      p.x   = df_order_x$P.Value.x[i], 
+                      pi1.x = pi1.x,
+                      p.y   = df_order_y$P.Value.y[i], 
+                      pi1.y = pi1.y)
+    res
+  })
+  res_p1_ramp = do.call(rbind, res_p1_ramp)
+
   list(fig_logFC  = fig1,
       fig_tstat   = fig2, 
       tab_pi      = tab_pi,
-      tab_corr    = tab_corr)
+      tab_corr    = tab_corr,
+      res_p1_ramp = res_p1_ramp)
 }
 
 
@@ -439,6 +467,53 @@ downloadFile_version <- function(id , version){
   # fig + circle(gr, geom = "point", aes(y = score2, color=factor(sign(score2), -1:1)),size=1, grid = TRUE, radius = 15, trackWidth=23, grid.background="white", grid.line="grey70", space.skip=.002) + scale_size(range = c(1, 2.5)) + scale_size_continuous("|logFC|", limits=c(0,max(limits))) + ylab(bquote(log[2]~fold~change)) + scale_color_manual(values = c("red", "grey60", "blue")) + theme(legend.position="none")
 
 
+
+
+
+#' Plot mean variance trend
+#' 
+#' Plot mean variance trend
+#' 
+#' @param x object returned by voom(), voomWithQualityWeights(), voomWithDreamWeights() or rationale()
+#' @param y missing
+#' @param ... other arguments
+#' 
+#' @import ggplot2
[…]

### 2020-09-04 combined_analysis/combine_all_data.Rmd
Fix Dx

@@
 			 	 dplyr::mutate(Cohort = forcats::fct_recode(Cohort, `MSSM-Penn-Pitt` = 'MSSM-Penn-Pitt', `NIMH-HBCC`='NIMH-HBCC', `NIMH-HBCC`='NIMH-HBCC-4')) %>%
-			 	 dplyr::mutate(Dx_crude = forcats::fct_recode(Dx, Other = "AFF", Other = "BP", Other = "undetermined", Control = "Control", SCZ = "SCZ"))
+			 	 dplyr::filter(Dx %in% c('Control', 'BP', "SCZ"))
+			 	 #%>%
+			 	 # dplyr::mutate(Dx_crude = forcats::fct_recode(Dx, Other = "AFF", Other = "BP", Other = "undetermined", Control = "Control", SCZ = "SCZ"))
 
@@
 
-form = ~ Dx + Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Region:Cohort) + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)
+form = ~ (1|Dx) + Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Region:Cohort) + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)
 
-vobj = voomWithDreamWeights(dge, form, METADATA, save.plot=TRUE, BPPARAM = SnowParam(8))
+vobj = voomWithDreamWeights(dge, form, METADATA, save.plot=TRUE, BPPARAM = SnowParam(18))
 
@@
 ```
+
+```{r differential.expression}
+
+form = ~ Dx*Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Region:Cohort) + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)
+
+Lbase = getContrast( vobj, form, METADATA, "RIN")
+Lbase[] = 0
+
+L = Lbase
+L['Dx_crudeControl'] = 1 
+L['Dx_crudeSCZ'] = -1
+
+fit = dream(vobj[1:10,], form, METADATA, BPPARAM = SnowParam(18))
+
+```

### 2020-09-04 combined_analysis/combine_all_data.Rmd
Fix Dx

@@
 
-```{r differential.expression}
+```{r differential.expression.simple}
 
-form = ~ Dx*Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Region:Cohort) + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)
+form = ~ 0 + Dx + Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Region:Cohort) + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)
 
+# get contrasts
 Lbase = getContrast( vobj, form, METADATA, "RIN")
@@
 
-L = Lbase
-L['Dx_crudeControl'] = 1 
-L['Dx_crudeSCZ'] = -1
+# SCZ
+L_SCZ = Lbase
+L_SCZ['DxControl'] = -1 
+L_SCZ['DxSCZ'] = 1
 
-fit = dream(vobj[1:10,], form, METADATA, BPPARAM = SnowParam(18))
+# BP
+L_BP = Lbase
+L_BP['DxControl'] = -1 
+L_BP['DxBP'] = 1
 
+# BP
+L_sex = Lbase
+L_sex['Reported_GenderMale'] = 1
+
+L = cbind(	Scz = L_SCZ, 
+			BP 	= L_BP, 
+			Sex = L_sex)
+plotContrasts(L) + theme()
+
+fit_marginal = dream(vobj[1:10,], form, METADATA, L=L,BPPARAM = SnowParam(18))
 ```

### 2020-09-04 combined_analysis/combine_all_data.Rmd
Fix Dx

@@
 			Sex = L_sex)
-plotContrasts(L) + theme()
 
-fit_marginal = dream(vobj[1:10,], form, METADATA, L=L,BPPARAM = SnowParam(18))
+plotContrasts(L) + theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust=1))
+
+fit_marginal = dream(vobj, form, METADATA, L=L,BPPARAM = SnowParam(18))
 ```

### 2020-09-04 combined_analysis/combine_all_data.Rmd
Fix Dx

@@
 ```
+
+```{r differential.expression.interaction.sex}
+
+form = ~ 0 + Dx*Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Region:Cohort) + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)
+
+# get contrasts
+Lbase = getContrast( vobj, form, METADATA, "RIN")
+Lbase[] = 0
+
+# SCZ by sex
+L_SCZ = Lbase
+L_SCZ['DxSCZ:Reported_GenderMale'] = 1
+L_SCZ['DxControl:Reported_GenderMale'] = -1
+L_SCZ['DxSCZ'] = -1
+L_SCZ['DxControl'] = 1 
+# (A - B) - (C - D) = A - B - C + D
+
+L = cbind(	SczSex = L_SCZ)
+
+plotContrasts(L) + theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust=1))
+
+fit_interaction = dream(vobj, form, METADATA, L=L,BPPARAM = SnowParam(18))
+```

### 2020-09-04 combined_analysis/combine_all_data.Rmd
Fix Dx

--- 2020-09-04 combined_analysis/combine_all_data.Rmd
@@
 
-form = ~ (1|Dx) + Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Region:Cohort) + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)
+form = ~ (1|Dx) + Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)
 
@@
 
+### Test marginals
 ```{r differential.expression.simple}
 
-form = ~ 0 + Dx + Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Region:Cohort) + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)
+form = ~ 0 + Dx + Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)
 
@@
 ```
+
+```{r plots.marginal}
+# Disease
+fig1 = plotVolcano( topTable(fit_marginal, "Scz", number=Inf) ) + ggtitle("SCZ")
+fig2 = plotVolcano( topTable(fit_marginal, "BP", number=Inf) ) + ggtitle("BP")
+
+cowplot( fig1, fig2, ncol=2)
+
+# Sex
+tab = topTable(fit_marginal, "Reported_GenderMale", number=Inf) %>% getGeneSymbol
+
+fig3 = plotVolcano( tab ) + ggtitle("Sex")
+fig4 = plotVolcano( tab[tab$Chrom %in% 1:22,] ) + ggtitle("Sex")
+
+cowplot( fig3, fig4, ncol=2)
+```
+
+```{r exit, cache=FALSE}
+knitr::knit_exit()
+```
+
+### Test Diagnosis * Sex
+```{r differential.expression.interaction.DxSex}
+
+form = ~ 0 + Dx*Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)
+
+# get contrasts
+Lbase = getContrast( vobj, form, METADATA, "RIN")
+Lbase[] = 0
+
+# SCZ by sex
+L_SCZ = Lbase
+L_SCZ['DxSCZ:Reported_GenderMale'] = 1
+L_SCZ['DxControl:Reported_GenderMale'] = -1
+L_SCZ['DxSCZ'] = -1
+L_SCZ['DxControl'] = 1 
+# (A - B) - (C - D) = A - B - C + D
+
+L = cbind(	SczSex = L_SCZ)
+
+plotContrasts(L) + theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust=1))
+
+fit_interaction_DxSex = dream(vobj, form, METADATA, L=L,BPPARAM = SnowParam(18))
+```
+
+### Test Region * Sex
+```{r differential.expression.interaction.RegionSex, eval=FALSE}
+
+form = ~ 0 + Dx + Region*Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Institution:Cohort) + (1|Cohort) + (1|Individual_ID)
+
+# get contrasts
+Lbase = getContrast( vobj, form, METADATA, "RIN")
+Lbase[] = 0
+
+# SCZ by sex
+L_SCZ = Lbase
+L_SCZ['DxSCZ:Reported_GenderMale'] = 1
+L_SCZ['DxControl:Reported_GenderMale'] = -1
+L_SCZ['DxSCZ'] = -1
+L_SCZ['DxControl'] = 1 
+# (A - B) - (C - D) = A - B - C + D
+
+L = cbind(	SczSex = L_SCZ)
+
+plotContrasts(L) + theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust=1))
+
+fit_interaction_RegionSex = dream(vobj, form, METADATA, L=L,BPPARAM = SnowParam(18))
+```
+
+### Test Region * Disease
+```{r differential.expression.interaction.RegionDx, eval=FALSE}
+
+form = ~ 0 + Region*Dx + Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Institution:Cohort) + (1|Cohort) + (1|Individual_ID)
+
+# get contrasts
+Lbase = getContrast( vobj, form, METADATA, "RIN")
+Lbase[] = 0
+
+# SCZ by sex
+L_SCZ = Lbase
+L_SCZ['DxSCZ:Reported_GenderMale'] = 1
+L_SCZ['DxControl:Reported_GenderMale'] = -1
+L_SCZ['DxSCZ'] = -1
+L_SCZ['DxControl'] = 1 
+# (A - B) - (C - D) = A - B - C + D
+
+L = cbind(	SczSex = L_SCZ)
+
+plotContrasts(L) + theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust=1))
+
+fit_interaction_RegionSex = dream(vobj, form, METADATA, L=L,BPPARAM = SnowParam(18))
+```

### 2020-09-04 combined_analysis/combine_all_data.Rmd
Fix Dx

@@
 
-cd /hpc/users/hoffmg01/work/cmc_analysis/src
+cd /hpc/users/hoffmg01/work/covarr-de
 # rm -rf molec_profiling_cache/
@@
 # Disease
-fig1 = plotVolcano( topTable(fit_marginal, "Scz", number=Inf) ) + ggtitle("SCZ")
-fig2 = plotVolcano( topTable(fit_marginal, "BP", number=Inf) ) + ggtitle("BP")
+tab = topTable(fit_marginal, "Scz", number=Inf) %>% getGeneSymbol
+fig1 = plotVolcano( tab, rownames(tab)[1:20] ) + ggtitle("SCZ")
 
-cowplot( fig1, fig2, ncol=2)
+tab = topTable(fit_marginal, "BP", number=Inf)
+fig2 = plotVolcano( tab, rownames(tab)[1:20] ) + ggtitle("BP")
+
+plot_grid( fig1, fig2, ncol=2)
 
@@
 
-cowplot( fig3, fig4, ncol=2)
+plot_grid( fig3, fig4, ncol=2)
 ```

### 2020-09-04 combined_analysis/combine_all_data.Rmd
Fix Dx

@@
 tab = topTable(fit_marginal, "Reported_GenderMale", number=Inf) %>% getGeneSymbol
+fig3 = plotVolcano( tab, rownames(tab)[1:20] ) + ggtitle("Sex")
 
-fig3 = plotVolcano( tab ) + ggtitle("Sex")
-fig4 = plotVolcano( tab[tab$Chrom %in% 1:22,] ) + ggtitle("Sex")
+tab2 = tab[tab$Chrom %in% 1:22,]
+fig4 = plotVolcano( tab2, rownames(tab2)[1:20] ) + ggtitle("Sex")
 

### 2020-09-04 combined_analysis/combine_all_data.Rmd
Fix Dx

--- 2020-09-04 combined_analysis/combine_all_data.Rmd
@@
 tab = topTable(fit_marginal, "Scz", number=Inf) %>% getGeneSymbol
-fig1 = plotVolcano( tab, rownames(tab)[1:20] ) + ggtitle("SCZ")
+fig1 = plotVolcano( tab, tab$Symbol[1:20] ) + ggtitle("SCZ")
 
 tab = topTable(fit_marginal, "BP", number=Inf)
-fig2 = plotVolcano( tab, rownames(tab)[1:20] ) + ggtitle("BP")
+fig2 = plotVolcano( tab, tab$Symbol[1:20] ) + ggtitle("BP")
 
@@
 tab = topTable(fit_marginal, "Reported_GenderMale", number=Inf) %>% getGeneSymbol
+fig3 = plotVolcano( tab, tab$Symbol[1:20] ) + ggtitle("Sex")
 
-fig3 = plotVolcano( tab ) + ggtitle("Sex")
-fig4 = plotVolcano( tab[tab$Chrom %in% 1:22,] ) + ggtitle("Sex")
+tab2 = tab[tab$Chrom %in% 1:22,]
+fig4 = plotVolcano( tab2, tab2$Symbol[1:20]) + ggtitle("Sex")
 

### 2020-09-04 combined_analysis/combine_all_data.Rmd
Fix Dx

@@
 
-```{r plots.marginal}
+```{r plots.marginal, fig.width=10, fig.height=15}
 # Disease
@@
 
-tab = topTable(fit_marginal, "BP", number=Inf)
+tab = topTable(fit_marginal, "BP", number=Inf)  %>% getGeneSymbol
 fig2 = plotVolcano( tab, tab$Symbol[1:20] ) + ggtitle("BP")

### 2020-09-04 combined_analysis/combine_all_data.Rmd
Fix Dx

@@
 
-```{r plots.marginal, fig.width=10, fig.height=15}
+```{r plots.marginal, fig.width=10, fig.height=10}
 # Disease

### 2020-09-04 combined_analysis/combine_all_data.Rmd
Fix Dx

@@
 
-```{r plots.marginal, fig.width=10, fig.height=10}
+```{r plots.marginal, fig.width=10, fig.height=6}
 # Disease

### 2020-09-04 combined_analysis/combine_all_data.Rmd
Fix Dx

@@
 
-```{r plots.marginal, fig.width=10, fig.height=6}
+```{r plots.marginal, fig.width=10, fig.height=5}
 # Disease
@@
 tab2 = tab[tab$Chrom %in% 1:22,]
-fig4 = plotVolcano( tab2, tab2$Symbol[1:20]) + ggtitle("Sex")
+fig4 = plotVolcano( tab2, tab2$Symbol[1:20]) + ggtitle("Sex: autosomes")
 

### 2020-09-04 combined_analysis/combine_all_data.Rmd
Fix Dx

@@
 
+METADATA$Dx = factor(METADATA$Dx, c('Control', 'BP', "SCZ"))
+
 ind = METADATA$SampleID[which(METADATA$ReportExclude == 1 | METADATA$GenotypeExclude)]
@@
 
-```{r exit, cache=FALSE}
-knitr::knit_exit()
-```
-
 ### Test Diagnosis * Sex
@@
 
+names(Lbase)
+
 # SCZ by sex
@@
 L_SCZ['DxSCZ:Reported_GenderMale'] = 1
-L_SCZ['DxControl:Reported_GenderMale'] = -1
+L_SCZ['Reported_GenderMale'] = -1
 L_SCZ['DxSCZ'] = -1
@@
 
-L = cbind(	SczSex = L_SCZ)
+# BP by sex
+L_BP = Lbase
+L_SCZ['DxBP:Reported_GenderMale'] = 1
+L_SCZ['Reported_GenderMale'] = -1
+L_SCZ['DxBP'] = -1
+L_SCZ['DxControl'] = 1
+
+L = cbind(	SczSex = L_SCZ,
+			BPSex = L_BP)
 
@@
 fit_interaction_DxSex = dream(vobj, form, METADATA, L=L,BPPARAM = SnowParam(18))
+```
+
+```{r plots.interaction.DxSex, fig.width=10, fig.height=5}
+# Disease
+tab = topTable(fit_marginal, "SczSex", number=Inf) %>% getGeneSymbol
+fig1 = plotVolcano( tab, tab$Symbol[1:20] ) + ggtitle("SCZ")
+
+tab = topTable(fit_marginal, "BPSex", number=Inf)  %>% getGeneSymbol
+fig2 = plotVolcano( tab, tab$Symbol[1:20] ) + ggtitle("BP")
+
+plot_grid( fig1, fig2, ncol=2) 
+```
+
+```{r exit, cache=FALSE}
+knitr::knit_exit()
 ```

### 2020-09-04 combined_analysis/combine_all_data.Rmd
Fix Dx

@@
 L_SCZ['DxSCZ:Reported_GenderMale'] = 1
-L_SCZ['Reported_GenderMale'] = -1
+L_SCZ['DxControl:Reported_GenderMale'] = -1
 L_SCZ['DxSCZ'] = -1
@@
 L_BP = Lbase
-L_SCZ['DxBP:Reported_GenderMale'] = 1
-L_SCZ['Reported_GenderMale'] = -1
+L_SCZ['Reported_GenderMale'] = 1
+L_SCZ['DxControl:Reported_GenderMale'] = -1
 L_SCZ['DxBP'] = -1
@@
 
-fit_interaction_DxSex = dream(vobj, form, METADATA, L=L,BPPARAM = SnowParam(18))
+fit_interaction_DxSex = dream(vobj, form, METADATA, L=L, BPPARAM = SnowParam(18))
 ```

### 2020-09-04 combined_analysis/combine_all_data.Rmd
Fix Dx

@@
 # Disease
-tab = topTable(fit_marginal, "SczSex", number=Inf) %>% getGeneSymbol
+tab = topTable(fit_interaction_DxSex, "SczSex", number=Inf) %>% getGeneSymbol
 fig1 = plotVolcano( tab, tab$Symbol[1:20] ) + ggtitle("SCZ")
 
-tab = topTable(fit_marginal, "BPSex", number=Inf)  %>% getGeneSymbol
+tab = topTable(fit_interaction_DxSex, "BPSex", number=Inf)  %>% getGeneSymbol
 fig2 = plotVolcano( tab, tab$Symbol[1:20] ) + ggtitle("BP")

### 2020-09-04 common_functions/common_functions.R
add entrez

@@ -36,6 +36,9 @@ getGeneSymbol = function( df, column="row.names"){
   geneSymbol = mapIds(ensdb, keys=ensIDs, keytype="GENEID", column="GENENAME")
   df$Symbol = geneSymbol
 
+  entrez = mapIds(ensdb, keys=ensIDs, keytype="GENEID", column="ENTREZID")
+  df$Entrez = entrez
+
   chroms = ensembldb::select(ensdb, keys=ensIDs, keytype="GENEID", column="SEQNAME")
 
   df2 = merge( data.frame(GENEID = ensIDs, stringsAsFactors=FALSE), chroms, by='GENEID')
@@ -102,36 +105,36 @@ plot_concordance = function( resList, showGenes = NULL, idx=c(1,2), size=15){
   tab_corr = data.frame(R_spearman = res$estimate, P = res$p.value)
 
   # pi spectrum
-  df_order_x = df[order(df$P.Value.x),]
-  df_order_y = df[order(df$P.Value.y),]
+  # df_order_x = df[order(df$P.Value.x),]
+  # df_order_y = df[order(df$P.Value.y),]
 
-  res_p1_ramp = lapply(1:nrow(df), function(i){
+  # res_p1_ramp = lapply(1:nrow(df), function(i){
 
-    pi1.x = tryCatch( 
-      1 - qvalue(df_order_x$P.Value.y[1:i])$pi0,
-      error = function(e){
-        NA
-        })
+  #   pi1.x = tryCatch( 
+  #     1 - qvalue(df_order_x$P.Value.y[1:i])$pi0,
+  #     error = function(e){
+  #       NA
+  #       })
 
-    pi1.y = tryCatch( 
-      1 - qvalue(df_order_y$P.Value.x[1:i])$pi0,
-      error = function(e){
-        NA
-        })
-    res = data.frame( i     = i, 
-                      p.x   = df_order_x$P.Value.x[i], 
-                      pi1.x = pi1.x,
-                      p.y   = df_order_y$P.Value.y[i], 
-                      pi1.y = pi1.y)
-    res
-  })
-  res_p1_ramp = do.call(rbind, res_p1_ramp)
+  #   pi1.y = tryCatch( 
+  #     1 - qvalue(df_order_y$P.Value.x[1:i])$pi0,
+  #     error = function(e){
+  #       NA
+  #       })
+  #   res = data.frame( i     = i, 
[…]

### 2020-09-09 combined_analysis/combine_all_data.Rmd
Fix Dx
add zenith

@@
 
-### Test Diagnosis * Sex
+```{r gene_set_test.marginal}
+
+library(zenith)
+
+run_zenith = function(fit, coefs, gs.collection, n_genes_mi=10, n_genes_max=5000){
+
+	# Map from Ensembl genes in geneSets_GO to 
+	# from trimmed Ensembl names from RNA-seq data 
+	geneSets.index = ids2indices( gs.collection, trim_ensembl_ids(rownames(fit)))
+
+	# filter by size of gene set
+	geneSets.index = geneSets.index[sapply(geneSets.index, function(x) (length(x) >= n_genes_min) & (length(x) <= n_genes_max)) ]
+
+	# run zenith for each coefficient
+	res = lapply( coefs, function(coef){
+		zenith(fit, coef, geneSets.index, squaredStats=TRUE)
+		} )
+	names(res) = coefs
+
+	res
+}
+
+# Load Gene Ontology database 
+gs.GO = get_GeneOntology()
+gs.MSigDB = get_MSigDB()
+
+# convert from GeneSetCollection to list used by camera and zenith
+gs.GO.collection = recodeToList( gs.GO )
+gs.MSigDB.collection = recodeToList( gs.MSigDB )
+
+res_zenith_marginal.GO = run_zenith( fit_marginal, colnames(L), gs.GO.collection )
+res_zenith_marginal.MSigDB = run_zenith( fit_marginal, colnames(L), gs.MSigDB.collection )
+```
+
+```{r exit, cache=FALSE}
+knitr::knit_exit()
+```
+
+### Test Diagnosis * Sex interaction
 ```{r differential.expression.interaction.DxSex}
@@
 L_BP = Lbase
-L_SCZ['Reported_GenderMale'] = 1
-L_SCZ['DxControl:Reported_GenderMale'] = -1
-L_SCZ['DxBP'] = -1
-L_SCZ['DxControl'] = 1
+L_BP['DxBP:Reported_GenderMale'] = 1
+L_BP['Reported_GenderMale'] = -1
+L_BP['DxBP'] = -1
+L_BP['DxControl'] = 1
 
@@
 
-```{r exit, cache=FALSE}
-knitr::knit_exit()
-```
-
-### Test Region * Sex
+### Test Region * Sex interaction
 ```{r differential.expression.interaction.RegionSex, eval=FALSE}

### 2020-09-09 combined_analysis/combine_all_data.Rmd
add zenith

@@
 
-run_zenith = function(fit, coefs, gs.collection, n_genes_mi=10, n_genes_max=5000){
+run_zenith = function(fit, coefs, gs.collection, n_genes_min=10, n_genes_max=5000){
 
@@
 res_zenith_marginal.MSigDB = run_zenith( fit_marginal, colnames(L), gs.MSigDB.collection )
+```
+
+```{r show.gene_set_test.marginal}
+res_zenith_marginal.GO[['Scz']][1:20,] %>% kable(caption="Scz") %>% kable_styling(full_width=FALSE)
+res_zenith_marginal.GO[['BP']][1:20,] %>% kable(caption="BP") %>% kable_styling(full_width=FALSE)
+res_zenith_marginal.GO[['Scz']][1:Sex,] %>% kable(caption="Sex") %>% kable_styling(full_width=FALSE)
+
 ```

### 2020-09-09 combined_analysis/combine_all_data.Rmd
add zenith

--- 2020-09-09 combined_analysis/combine_all_data.Rmd
@@
 res_zenith_marginal.MSigDB = run_zenith( fit_marginal, colnames(L), gs.MSigDB.collection )
+```
+
+```{r show.gene_set_test.marginal}
+res_zenith_marginal.GO[['Scz']][1:20,] %>% kable(caption="Scz") %>% kable_styling(full_width=FALSE)
+res_zenith_marginal.GO[['BP']][1:20,] %>% kable(caption="BP") %>% kable_styling(full_width=FALSE)
+res_zenith_marginal.GO[['Sex']][1:20,] %>% kable(caption="Sex") %>% kable_styling(full_width=FALSE)
 ```

### 2020-09-09 combined_analysis/combine_all_data.Rmd
add zenith

--- 2020-09-09 combined_analysis/combine_all_data.Rmd
@@
 res_zenith_marginal.MSigDB = run_zenith( fit_marginal, colnames(L), gs.MSigDB.collection )
+```
+
+```{r show.gene_set_test.marginal}
+res_zenith_marginal.GO[['Scz']][1:20,] %>% kable(caption="Scz") %>% kable_styling(full_width=FALSE)
+res_zenith_marginal.MSigDB[['Scz']][1:20,] %>% kable(caption="Scz") %>% kable_styling(full_width=FALSE)
+res_zenith_marginal.GO[['BP']][1:20,] %>% kable(caption="BP") %>% kable_styling(full_width=FALSE)
+res_zenith_marginal.MSigDB[['BP']][1:20,] %>% kable(caption="BP") %>% kable_styling(full_width=FALSE)
+res_zenith_marginal.GO[['Sex']][1:20,] %>% kable(caption="Sex") %>% kable_styling(full_width=FALSE)
+res_zenith_marginal.MSigDB[['Sex']][1:20,] %>% kable(caption="Sex") %>% kable_styling(full_width=FALSE)
 ```

### 2020-09-09 combined_analysis/combine_all_data.Rmd
add zenith

--- 2020-09-09 combined_analysis/combine_all_data.Rmd
@@
 ```{r show.gene_set_test.marginal}
-res_zenith_marginal.GO[['Scz']][1:20,] %>% kable(caption="Scz") %>% kable_styling(full_width=FALSE)
-res_zenith_marginal.GO[['BP']][1:20,] %>% kable(caption="BP") %>% kable_styling(full_width=FALSE)
-res_zenith_marginal.GO[['Sex']][1:20,] %>% kable(caption="Sex") %>% kable_styling(full_width=FALSE)
+
+show_zenith_results = function( resList, key, FDR = 0.05){
+	df = resList[[key]]
+
+	# filter by FDR
+	df = df[df$FDR <= FDR,]
+
+	df %>% kable(caption=key) %>% kable_styling(full_width=FALSE)
+}
+
+show_zenith_results( res_zenith_marginal.GO, "Scz")
+show_zenith_results( res_zenith_marginal.MSigDB, "Scz")
+
+show_zenith_results( res_zenith_marginal.GO, "BP")
+show_zenith_results( res_zenith_marginal.MSigDB, "BP")
+
+show_zenith_results( res_zenith_marginal.GO, "Sex")
+show_zenith_results( res_zenith_marginal.MSigDB, "Sex")
 ```

### 2020-09-09 combined_analysis/combine_all_data.Rmd
add zenith

@@
 
-```{r exit, cache=FALSE}
-knitr::knit_exit()
-```
-
 ### Test Diagnosis * Sex interaction
@@
 plot_grid( fig1, fig2, ncol=2) 
+```
+
+```{r gene_set_test.marginal}
+res_zenith_DxSex.GO = run_zenith( fit_interaction_DxSex, colnames(L), gs.GO.collection )
+res_zenith_DxSex.MSigDB = run_zenith( fit_interaction_DxSex, colnames(L), gs.MSigDB.collection )
+```
+
+```{r show.gene_set_test.marginal}
+show_zenith_results( res_zenith_DxSex.GO, "Scz")
+show_zenith_results( res_zenith_DxSex.MSigDB, "Scz")
+
+show_zenith_results( res_zenith_DxSex.GO, "BP")
+show_zenith_results( res_zenith_DxSex.MSigDB, "BP")
+
+show_zenith_results( res_zenith_DxSex.GO, "Sex")
+show_zenith_results( res_zenith_DxSex.MSigDB, "Sex")
+```
+
+```{r exit, cache=FALSE}
+knitr::knit_exit()
 ```

### 2020-09-09 combined_analysis/combine_all_data.Rmd
add zenith

--- 2020-09-09 combined_analysis/combine_all_data.Rmd
@@
 plot_grid( fig1, fig2, ncol=2) 
+```
+
+```{r gene_set_test.DxSex}
+res_zenith_DxSex.GO = run_zenith( fit_interaction_DxSex, colnames(L), gs.GO.collection )
+res_zenith_DxSex.MSigDB = run_zenith( fit_interaction_DxSex, colnames(L), gs.MSigDB.collection )
+```
+
+```{r show.gene_set_test.DxSex}
+show_zenith_results( res_zenith_DxSex.GO, "SczSex")
+show_zenith_results( res_zenith_DxSex.MSigDB, "SczSex")
+
+show_zenith_results( res_zenith_DxSex.GO, "BPSex")
+show_zenith_results( res_zenith_DxSex.MSigDB, "BPSex")
+```
+
+```{r exit, cache=FALSE}
+knitr::knit_exit()
 ```

### 2020-09-09 combined_analysis/combine_all_data.Rmd
add zenith

@@
 
+METADATA$Reported_Gender = factor(METADATA$Reported_Gender)
+METADATA$Dx = factor(METADATA$Dx, c("BP", "Control", "SCZ"))
+
 form = ~ 0 + Dx*Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)
@@
 L_SCZ['DxSCZ:Reported_GenderMale'] = 1
-L_SCZ['Reported_GenderMale'] = -1
+L_SCZ['DxControl:Reported_GenderMale"'] = -1
 L_SCZ['DxSCZ'] = -1
@@
 
-# BP by sex
-L_BP = Lbase
-L_BP['DxBP:Reported_GenderMale'] = 1
-L_BP['Reported_GenderMale'] = -1
-L_BP['DxBP'] = -1
-L_BP['DxControl'] = 1
+# # BP by sex
+# L_BP = Lbase
+# L_BP['DxBP:Reported_GenderMale'] = 1
+# L_BP['Reported_GenderMale'] = -1
+# L_BP['DxBP'] = -1
+# L_BP['DxControl'] = 1
 
-L = cbind(	SczSex = L_SCZ,
-			BPSex = L_BP)
+L = cbind(	SczSex = L_SCZ)#,
+			# BPSex = L_BP)
 
@@
 ```{r plots.interaction.DxSex, fig.width=10, fig.height=5}
-# Disease
+# Disease. 
 tab = topTable(fit_interaction_DxSex, "SczSex", number=Inf) %>% getGeneSymbol
@@
 ```{r gene_set_test.DxSex}
-res_zenith_DxSex.GO = run_zenith( fit_interaction_DxSex, colnames(L), gs.GO.collection )
+res_zenith_DxSex.GO = run_zenith( fit_interaction_DxSex, colnames(L), gs.GO.collection ). 
 res_zenith_DxSex.MSigDB = run_zenith( fit_interaction_DxSex, colnames(L), gs.MSigDB.collection )

### 2020-09-09 combined_analysis/combine_all_data.Rmd
add zenith

@@
 
-METADATA$Reported_Gender = factor(METADATA$Reported_Gender)
-METADATA$Dx = factor(METADATA$Dx, c("BP", "Control", "SCZ"))
+# METADATA$Reported_Gender = factor(METADATA$Reported_Gender)
+# METADATA$Dx = factor(METADATA$Dx, c("BP", "Control", "SCZ"))
 
-form = ~ 0 + Dx*Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)
+form = ~ Dx*Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)
 
@@
 # SCZ by sex
-L_SCZ = Lbase
-L_SCZ['DxSCZ:Reported_GenderMale'] = 1
-L_SCZ['DxControl:Reported_GenderMale"'] = -1
-L_SCZ['DxSCZ'] = -1
-L_SCZ['DxControl'] = 1 
+# L_SCZ = Lbase
+# L_SCZ['DxSCZ:Reported_GenderMale'] = 1
+# L_SCZ['DxControl:Reported_GenderMale'] = -1
+# L_SCZ['DxSCZ'] = -1
+# L_SCZ['DxControl'] = 1 
 # (A - B) - (C - D) = A - B - C + D
@@
 
-L = cbind(	SczSex = L_SCZ)#,
-			# BPSex = L_BP)
+L_SCZ = Lbase
+L_SCZ['DxSCZ:Reported_GenderMale'] = 1
+
+L_BP = Lbase
+L_BP['DxBP:Reported_GenderMale'] = 1
+
+L = cbind(	SczSex = L_SCZ,
+			BPSex = L_BP)
 

### 2020-09-09 combined_analysis/combine_all_data.Rmd
add zenith

@@
 
-# SCZ by sex
-# L_SCZ = Lbase
-# L_SCZ['DxSCZ:Reported_GenderMale'] = 1
-# L_SCZ['DxControl:Reported_GenderMale'] = -1
-# L_SCZ['DxSCZ'] = -1
-# L_SCZ['DxControl'] = 1 
-# (A - B) - (C - D) = A - B - C + D
-
-# # BP by sex
-# L_BP = Lbase
-# L_BP['DxBP:Reported_GenderMale'] = 1
-# L_BP['Reported_GenderMale'] = -1
-# L_BP['DxBP'] = -1
-# L_BP['DxControl'] = 1
-
 L_SCZ = Lbase
@@
 plotContrasts(L) + theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust=1))
+
+browser()
 

### 2020-09-09 combined_analysis/combine_all_data.Rmd
add zenith

@@
 
-browser()
-
 fit_interaction_DxSex = dream(vobj, form, METADATA, L=L, BPPARAM = SnowParam(18))
@@
 ```{r gene_set_test.DxSex}
-res_zenith_DxSex.GO = run_zenith( fit_interaction_DxSex, colnames(L), gs.GO.collection ). 
+res_zenith_DxSex.GO = run_zenith( fit_interaction_DxSex, colnames(L), gs.GO.collection )
 res_zenith_DxSex.MSigDB = run_zenith( fit_interaction_DxSex, colnames(L), gs.MSigDB.collection )

### 2020-09-09 combined_analysis/combine_all_data.Rmd
add zenith

@@
 res_zenith_DxSex.GO = run_zenith( fit_interaction_DxSex, colnames(L), gs.GO.collection )
-res_zenith_DxSex.MSigDB = run_zenith( fit_interaction_DxSex, colnames(L), gs.MSigDB.collection )
+res_zenith_DxSex.MSigDB = run_zenith( fit_interaction_DxSex, colnames(L), gs.MSigDB.collection ).   
 ```

### 2020-09-11 combined_analysis/combine_all_DLPFC.Rmd
Inital commit

--- 2020-09-09 combined_analysis/combine_all_data.Rmd
@@
 
-rmarkdown::render("combined_analysis/combine_all_data.Rmd")
+rmarkdown::render("combined_analysis/combine_all_DLPFC.Rmd")
 
@@
 synLogin()
-knit2synapse::createAndKnitToFolderEntity(file = "combined_analysis/combine_all_data.Rmd",
+knit2synapse::createAndKnitToFolderEntity(file = "combined_analysis/combine_all_DLPFC.Rmd",
                                           parentId ="syn22231629",
-                                          folderName = 'Joint analysis of count data',
+                                          folderName = 'Joint analysis of DLPFC count data',
                                           overwrite=TRUE, 
@@
   		 		 dplyr::filter(SampleID %in% colnames(geneCountsMerged), !is.na(SampleID)) %>%
-  		 		 dplyr::filter(Region %in% c('DLPFC', 'ACC')) %>%
+  		 		 dplyr::filter(Region %in% c('DLPFC')) %>%
 			 	 dplyr::mutate(Cohort = forcats::fct_recode(Cohort, `MSSM-Penn-Pitt` = 'MSSM-Penn-Pitt', `NIMH-HBCC`='NIMH-HBCC', `NIMH-HBCC`='NIMH-HBCC-4')) %>%
-			 	 dplyr::filter(Dx %in% c('Control', 'BP', "SCZ"))
-			 	 #%>%
-			 	 # dplyr::mutate(Dx_crude = forcats::fct_recode(Dx, Other = "AFF", Other = "BP", Other = "undetermined", Control = "Control", SCZ = "SCZ"))
+			 	 dplyr::filter(Dx %in% c('Control', "SCZ"))
 
-METADATA$Dx = factor(METADATA$Dx, c('Control', 'BP', "SCZ"))
+# Set Control to baseline
+METADATA$Dx = factor(METADATA$Dx, c('Control', 'SCZ'))
 
@@
   kable() 
+```
+
+```{r exit1, cache=FALSE}
+knitr::knit_exit()
 ```
@@
 ```{r show.gene_set_test.DxSex}
-show_zenith_results( res_zenith_DxSex.GO, "SczSex")
-show_zenith_results( res_zenith_DxSex.MSigDB, "SczSex")
+show_zenith_results( res_zenith_DxSex.GO, "DxSCZ:Reported_GenderMale")
+show_zenith_results( res_zenith_DxSex.MSigDB, "DxSCZ:Reported_GenderMale")
 
-show_zenith_results( res_zenith_DxSex.GO, "BPSex")
-show_zenith_results( res_zenith_DxSex.MSigDB, "BPSex")
+show_zenith_results( res_zenith_DxSex.GO, "DxBP:Reported_GenderMale")
+show_zenith_results( res_zenith_DxSex.MSigDB, "DxBP:Reported_GenderMale")
 ```

### 2020-09-11 combined_analysis/combine_all_DLPFC.Rmd
Inital commit

@@
 
---->
-
 ```{r knit2synapse, eval=FALSE}
@@
 ```
-
-GSEA and dream from raw counts
 
@@
 
-# HBCC - sgACC
-counts.HBCC4 = fread(synGet('syn21867866')$path, data.table=FALSE)
-rownames(counts.HBCC4) = counts.HBCC4$Geneid
-counts.HBCC4 = counts.HBCC4[,-c(1:6)]
+# # HBCC - sgACC
+# counts.HBCC4 = fread(synGet('syn21867866')$path, data.table=FALSE)
+# rownames(counts.HBCC4) = counts.HBCC4$Geneid
+# counts.HBCC4 = counts.HBCC4[,-c(1:6)]
 
@@
 
-METADATA %>% 
-  group_by(Dx) %>%  
-  dplyr::summarise(count = n()) %>% 
-  kable() 
+# METADATA %>% 
+#   group_by(Dx) %>%  
+#   dplyr::summarise(count = n()) %>% 
+#   kable() 
 ```
 
-```{r exit1, cache=FALSE}
-knitr::knit_exit()
+## Summarize samples
+```{r summary}
+METADATA %>% group_by(Dx, Institution) %>% dplyr::summarise(count = n()) %>% 
+  kable(caption="Diagnosis, Institution") %>% kable_styling(full_width=FALSE)
+
+METADATA %>% group_by( Institution, Reported_Gender) %>% dplyr::summarise(count = n()) %>% 
+  kable(caption="Institution, Sex") %>% kable_styling(full_width=FALSE)
+
+METADATA %>% group_by(Dx, Institution, Reported_Gender) %>% dplyr::summarise(count = n()) %>% 
+  kable(caption="Diagnosis, Institution, Sex") %>% kable_styling(full_width=FALSE) 
 ```
@@
 ```{r process.counts}
-
 # keep genes that are express in at least one combination of cohort and gender
@@
 isexpr = do.call(rbind, isexpr.lst)
-isexpr = colSums(isexpr) > 0
+isexpr = colSums(isexpr) >= 3
 table(isexpr)
 
+# Normalize each cohort separately
+dge.lst = lapply( levels(METADATA$Cohort), function(Cohort){
   # normalize counts
-dge = DGEList( geneCountsMerged[isexpr,] )
-dge = calcNormFactors(dge) # TMM
+  dge = DGEList( geneCountsMerged[isexpr,METADATA$Cohort==Cohort] )
+  calcNormFactors(dge) # TMM
+})
+names(dge.lst) = levels(METADATA$Cohort)
 
-xtabs(~ Region + Cohort, METADATA)
-xtabs(~ Institution + Cohort, METADATA)
+# form = ~ (1|Dx) + Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)
+
+form = ~ Dx + Reported_Gender + RIN + ageOfDeath + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate)
+
+vobj.lst = lapply(dge.lst, function(dge){
+
+  i = match(colnames(dge), rownames(METADATA))
+  info = METADATA[i,]
+
+  if( length(unique(info$Institution)) > 1 ){
+    form_use = as.formula(paste('~',as.character(form)[2], "+ Institution", collapse=" "))
+  }else{
+    form_use = form
+  }
+
+  voomWithDreamWeights(dge, form_use, info, save.plot=TRUE, BPPARAM = SnowParam(18))
+  })
+names(vobj.lst) = names(dge.lst)
 
@@
 # Dx + scale(RIN2) + (1 | Institution) + scale(ageOfDeath) + scale(RIN) + (1 | Reported_Gender) + scale(IntronicRate) + scale(IntragenicRate)
+```
 
-form = ~ (1|Dx) + Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)
-
-vobj = voomWithDreamWeights(dge, form, METADATA, save.plot=TRUE, BPPARAM = SnowParam(18))
-
-plot_voom(vobj)
+### Voom trend plots
+```{r plot.voom, fig.height=5, fig.width=10}
+fig1 = plot_voom(vobj.lst[[1]]) + ggtitle(names(vobj.lst)[1])
+fig2 = plot_voom(vobj.lst[[2]]) + ggtitle(names(vobj.lst)[2])
+plot_grid( fig1, fig2, nrow=1) 
 ```
@@
 
-form = ~ 0 + Dx + Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)
+form = ~ Dx + Reported_Gender + RIN + ageOfDeath + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate)
 
-# get contrasts
-Lbase = getContrast( vobj, form, METADATA, "RIN")
-Lbase[] = 0
+fit_marginal.lst = lapply( vobj.lst, function(vobj){
 
-# SCZ
-L_SCZ = Lbase
-L_SCZ['DxControl'] = -1 
-L_SCZ['DxSCZ'] = 1
+  i = match(colnames(vobj), rownames(METADATA))
+  info = METADATA[i,]
 
-# BP
-L_BP = Lbase
-L_BP['DxControl'] = -1 
-L_BP['DxBP'] = 1
+  if( length(unique(info$Institution)) > 1 ){
+    form_use = as.formula(paste('~',as.character(form)[2], "+ Institution", collapse=" "))
+  }else{
+    form_use = form
+  }
 
-# BP
-L_sex = Lbase
-L_sex['Reported_GenderMale'] = 1
-
-L = cbind(	Scz = L_SCZ, 
-			BP 	= L_BP, 
-			Sex = L_sex)
-
-plotContrasts(L) + theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust=1))
-
-fit_marginal = dream(vobj, form, METADATA, L=L,BPPARAM = SnowParam(18))
+  fit = dream(vobj, form, info, BPPARAM = SnowParam(18))  
+  eBayes( fit )
+  })
+names(fit_marginal.lst) = names(vobj.lst)
 ```
 
-```{r plots.marginal, fig.width=10, fig.height=5}
-# Disease
-tab = topTable(fit_marginal, "Scz", number=Inf) %>% getGeneSymbol
-fig1 = plotVolcano( tab, tab$Symbol[1:20] ) + ggtitle("SCZ")
-
-tab = topTable(fit_marginal, "BP", number=Inf)  %>% getGeneSymbol
-fig2 = plotVolcano( tab, tab$Symbol[1:20] ) + ggtitle("BP")
-
-plot_grid( fig1, fig2, ncol=2)
-
-# Sex  
-tab = topTable(fit_marginal, "Reported_GenderMale", number=Inf) %>% getGeneSymbol
-fig3 = plotVolcano( tab, tab$Symbol[1:20] ) + ggtitle("Sex")
-
-tab2 = tab[tab$Chrom %in% 1:22,]
-fig4 = plotVolcano( tab2, tab2$Symbol[1:20]) + ggtitle("Sex: autosomes")
-
-plot_grid( fig3, fig4, ncol=2)
-```
-
-```{r gene_set_test.marginal}
-
+```{r load.genesets}
 library(zenith)
-
-run_zenith = function(fit, coefs, gs.collection, n_genes_min=10, n_genes_max=5000){
-
-	# Map from Ensembl genes in geneSets_GO to 
-	# from trimmed Ensembl names from RNA-seq data 
-	geneSets.index = ids2indices( gs.collection, trim_ensembl_ids(rownames(fit)))
-
-	# filter by size of gene set
-	geneSets.index = geneSets.index[sapply(geneSets.index, function(x) (length(x) >= n_genes_min) & (length(x) <= n_genes_max)) ]
-
-	# run zenith for each coefficient
-	res = lapply( coefs, function(coef){
-		zenith(fit, coef, geneSets.index, squaredStats=TRUE)
-		} )
-	names(res) = coefs
-
-	res
-}
 
@@
 gs.MSigDB.collection = recodeToList( gs.MSigDB )
+```
 
-res_zenith_marginal.GO = run_zenith( fit_marginal, colnames(L), gs.GO.collection )
-res_zenith_marginal.MSigDB = run_zenith( fit_marginal, colnames(L), gs.MSigDB.collection )
+```{r run.zenith}
+
+coefTest = c('DxSCZ', 'Reported_GenderMale')
+res_zenith = lapply( names(fit_marginal.lst), function(key){
+
+  fit = fit_marginal.lst[[key]]
+
+  res = lapply(coefTest, function(coef){
+    res1 = run_zenith( fit, coef, gs.GO.collection )
+    res1$Resource = 'GO'
+
+    res2 = run_zenith( fit, coef, gs.MSigDB.collection )
+    res2$Resource = "MSigDB"
+
+    res = rbind(res1, res2)
+    res$Coef = coef
+    res$Cohort = key
+
+    res
+  })
+  do.call(rbind, res)
+})
+names(res_zenith) = names(fit_marginal.lst)
 ```
+
+```{r get.signatures, fig.width=10, fig.height=5}
+# Get DE signatures
+df_DE_results = lapply(fit_marginal.lst, function(fit){
+  res = lapply(coefTest, function(coef){
+    res = topTable(fit, coef, number=Inf) 
+    res$Coef = coef
+    res
+    })
+  do.call(rbind, res) %>% getGeneSymbol
+  })
+names(df_DE_results) = names(fit_marginal.lst)
+```
+
+```{r test.scz, fig.height=5, fig.width=10}
+# Schizophrenia testing
+# extract from each chort
+coef = 'DxSCZ'
+figList.Scz = lapply( names(df_DE_results), function(key){
+  df = df_DE_results[[key]]
+  df = df[df$Coef == coef,]
+
+  main = paste(key, coef, sep = ': ' )
+  plotVolcano( df, df$Symbol[1:20] ) + ggtitle( main )
+})
+
+# volcano plots
+plot_grid(plotlist=figList.Scz, nrow=1)
+```
+
+```{r meta.analysis.SCZ}
+# Create list preparing for meta-analysis for SCZ
+tab.lst = lapply( df_DE_results, function(df){
+  df[df$Coef == coef,]
+})
+names(tab.lst) = names(df_DE_results)
+
+# Run meta-analysis
+df_meta_SCZ = run_meta_analysis( tab.lst, "FE") %>% getGeneSymbol
+```
+
+```{r concordance.scz, fig.height=6, fig.width=10}
+# Concordance analysis
+resConcord.SCZ = plot_concordance(tab.lst, showGenes = df_meta_SCZ$Symbol[1:20])
+
+main = "Concordance of SCZ"
+plot_grid(  resConcord.SCZ$fig_logFC + ggtitle(main),
+            resConcord.SCZ$fig_tstat + ggtitle(main), nrow=1)
+
+data.frame(pi1 = resConcord.SCZ$pi_single) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)
+resConcord.SCZ$tab_pi %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
+resConcord.SCZ$tab_corr %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
+```
+
+```{r plot.meta.analysis.SCZ, fig.height=5, fig.width=10}
+# volcano plot for meta-analysis
+main = paste('Meta-analysis', coef, sep = ': ' )
+plotVolcano( df_meta_SCZ, df_meta_SCZ$Symbol[1:20] ) + ggtitle( main )
+```
+
+```{r exit2, cache=FALSE}
+knitr::knit_exit()
+```
+
+#   rmarkdown::render("combined_analysis/combine_all_DLPFC.Rmd")
 

### 2020-09-11 combined_analysis/combine_all_DLPFC.Rmd
Inital commit

@@
 # # HBCC - sgACC
-# counts.HBCC4 = fread(synGet('syn21867866')$path, data.table=FALSE)
-# rownames(counts.HBCC4) = counts.HBCC4$Geneid
-# counts.HBCC4 = counts.HBCC4[,-c(1:6)]
+counts.HBCC4 = fread(synGet('syn21867866')$path, data.table=FALSE)
+rownames(counts.HBCC4) = counts.HBCC4$Geneid
+counts.HBCC4 = counts.HBCC4[,-c(1:6)]
 

### 2020-09-11 common_functions/common_functions.R
Add zenith

@@ -8,6 +8,7 @@ library(ggbio)
 library(viridis)
 library(data.table)
 })
+
 # For ENSEMBL id ENSG00000279457.4, return ENSG00000279457
 trim_ensembl_ids = function(x){
   gsub("(.*)\\.(.*)", "\\1", x) 
@@ -65,18 +66,17 @@ plot_concordance = function( resList, showGenes = NULL, idx=c(1,2), size=15){
   lab = "log2 fold change"
   lim = with(df, max(abs(c(logFC.x, logFC.y))))
   df$density <- get_density(df$logFC.x, df$logFC.y, n = 100)
-  fig1 = ggplot(df, aes(logFC.x, logFC.y, color=density)) + geom_point(size=.4) + theme_bw(size) + theme(aspect.ratio=1, legend.position="bottom") + geom_abline(color="red") + xlab(paste(names(resList)[1], lab)) + ylab(paste(names(resList)[2], lab)) + geom_vline(xintercept=0, col="grey40", linetype="dashed") + geom_hline(yintercept=0, col="grey40", linetype="dashed") + xlim(-lim, lim) + ylim(-lim, lim) + scale_color_viridis() + geom_smooth(method="lm", se=FALSE, color="darkorange")
+  fig1 = ggplot(df, aes(logFC.x, logFC.y, color=density)) + geom_point(size=.4) + theme_bw(size) + theme(aspect.ratio=1, legend.position="bottom", plot.title = element_text(hjust = 0.5)) + geom_abline(color="red") + xlab(paste(names(resList)[1], lab)) + ylab(paste(names(resList)[2], lab)) + geom_vline(xintercept=0, col="grey40", linetype="dashed") + geom_hline(yintercept=0, col="grey40", linetype="dashed") + xlim(-lim, lim) + ylim(-lim, lim) + scale_color_viridis() + geom_smooth(method="lm", se=FALSE, color="darkorange")
 
   if( !is.null(showGenes) ){
     df2 = df[df$Symbol.x %in% showGenes,]
     fig1 = fig1 + geom_text_repel(data=df2, aes(logFC.x, logFC.y, label=Symbol.x), segment.size=.5,  segment.color="black", color="black", force=1, nudge_x=.005, nudge_y=.1)
   }
 
-
   lab = "t-statistic"
   lim = with(df, max(abs(c(t.x, t.y))))
   df$density <- get_density(df$t.x, df$t.y, n = 100)
[…]

### 2020-09-11 common_functions/common_functions.R
Add zenith

@@ -511,11 +511,11 @@ run_zenith = function(fit, coefs, gs.collection, n_genes_min=10, n_genes_max=500
 
   # run zenith for each coefficient
   res = lapply( coefs, function(coef){
-    zenith(fit, coef, geneSets.index, squaredStats=TRUE)
-    } )
-  names(res) = coefs
-
-  res
+    df = zenith(fit, coef, geneSets.index, squaredStats=TRUE)
+    df$Coef = coef
+    df
+    })
+  do.call(rbind, res)
 }
  
 

### 2020-09-11 combined_analysis/combine_all_DLPFC.Rmd
Inital commit

@@
 
-  res = lapply(coefTest, function(coef){
-    res1 = run_zenith( fit, coef, gs.GO.collection )
-    res1$Resource = 'GO'
+  res.GO = run_zenith( fit, coefTest, gs.GO.collection )
+  res.GO$Resource = "GO"
+  res.GO$Cohort = key
 
-    res2 = run_zenith( fit, coef, gs.MSigDB.collection )
-    res2$Resource = "MSigDB"
+  res.MSigDB = run_zenith( fit, coefTest, gs.MSigDB.collection )
+  res.MSigDB$Resource = "MSigDB"
+  res.MSigDB$Cohort = key
 
-    res = rbind(res1, res2)
-    res$Coef = coef
-    res$Cohort = key
+  rbind(res.GO, res.MSigDB)
 
-    res
-  })
-  do.call(rbind, res)
 })

### 2020-09-11 combined_analysis/combine_all_DLPFC.Rmd
Inital commit

@@
 
+### Create residuals
+```{r Create residuals}
+
+form = ~ Dx*Reported_Gender + RIN + ageOfDeath + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate)
+
+residuals.lst = lapply( vobj.lst, function(vobj){
+
+  i = match(colnames(vobj), rownames(METADATA))
+  info = METADATA[i,]
+
+  if( length(unique(info$Institution)) > 1 ){
+    form_use = as.formula(paste('~',as.character(form)[2], "+ Institution", collapse=" "))
+  }else{
+    form_use = form
+  }
+
+  # Compute residuals which adding back Diagnosis and Sex effects
+  resid.lst = fitVarPartModel(vobj, form, info, fxn = function(fit){
+    residuals(fit) + get_prediction( fit, ~ Dx*Reported_Gender)
+    }, BPPARAM = SnowParam(18))  
+  do.call(rbind, resid.lst)
+  })
+names(residuals.lst) = names(vobj.lst)
+```
+
 ### Test marginals
@@
 
+# Create residuals using fitVarPart and get_prediction()
+
 ### Test Diagnosis * Sex interaction

### 2020-09-11 combined_analysis/combine_all_DLPFC.Rmd
Inital commit

@@
 
+```{r Create.residuals}
+
 ### Create residuals
-```{r Create residuals}
-
 form = ~ Dx*Reported_Gender + RIN + ageOfDeath + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate)
@@
 
-### Test marginals
+```{r check.correlation}
+ngenes = 2000
+C1 = cor(t(residuals.lst[[1]][1:ngenes,]))
+C2 = cor(t(residuals.lst[[2]][1:ngenes,]))
+
+df = data.table(A   = as.numeric(C1[lower.tri(C1)]),
+                B   = as.numeric(C2[lower.tri(C2)]))
+
+df$density <- get_density(df$A, df$B, n = 100)
+ggplot(df, aes(A, B, color=density)) + geom_point() + geom_point(size=.4) + theme_bw(15) + theme(aspect.ratio=1, legend.position="bottom", plot.title = element_text(hjust = 0.5)) + geom_abline(color="red") + scale_color_viridis() + xlim(-1,1) + ylim(-1,1) + xlab(paste0('Correlation (', names(residuals.lst)[1], ')')) + ylab(paste0('Correlation (', names(residuals.lst)[2], ')'))
+```
+
+### Test marginals for Schizophrneia vs controls
 ```{r differential.expression.simple}
@@
 plotVolcano( df_meta_SCZ, df_meta_SCZ$Symbol[1:20] ) + ggtitle( main )
+
+pi1 = 1 - qvalue(df_meta_SCZ$P)$pi0
+data.frame(pi1) %>% kable(caption="pi1 for SCZ meta-analysis") %>% kable_styling(full_width=FALSE)
+
 ```
+
+# Test RCSD anbd ADAMTS2
 

### 2020-09-16 combined_analysis/combine_all_DLPFC.Rmd
Inital commit

--- 2020-09-11 combined_analysis/combine_all_DLPFC.Rmd
@@
 ---
-title: "Joint analysis of read count data"
-subtitle: ''
-author: "Developed by [Gabriel Hoffman](http://gabrielhoffman.github.io/)"
+title: "CommonMind Consortium: DLPFC"
+subtitle: "Analysis of read count data by cohort"
+author:
+- name: "[Gabriel Hoffman](http://gabrielhoffman.github.io)$^{1}$ and Kelsey Montgomery$^{2}$"
+  affiliation: | 
+    [1] Icahn School of Medicine at Mount Sinai, New York
+    [2] Sage Bionetworks, Seattle, WA
 date: "Run on `r Sys.time()`"
@@
     toc: true
-  smart: false
+    toc_float: true
 ---
@@
 
-```{r knit2synapse, eval=FALSE}
+```{r knit2synapse, eval= FALSE}
 library(synapser)
@@
 ```
+
+TODO
+
+zenith 1 sided
+
+zenith plots
+replication or correlation as a function of expression magnitude
+
+difference networks
+Number of DE genes
+
+<!---
+DONE
+
+Effect of ilr should depend on Instaition
+suffix 1.
+zenith for Sex with autosomes
+
+--->
 
@@
 library(ggrepel)
+library(corrplot)
+library(zenith)
+library(compositions)
 library(knitr)
@@
 			 	 dplyr::mutate(Cohort = forcats::fct_recode(Cohort, `MSSM-Penn-Pitt` = 'MSSM-Penn-Pitt', `NIMH-HBCC`='NIMH-HBCC', `NIMH-HBCC`='NIMH-HBCC-4')) %>%
-			 	 dplyr::filter(Dx %in% c('Control', "SCZ"))
+			 	 dplyr::filter(Dx %in% c('Control', "SCZ"))  %>%
+         dplyr::filter(ageOfDeath > 17)
 
@@
 
+```{r cellFractionEstimates}
+# include estimated cell fractions in METADATA
+df_cellFractions = read.table(synGet('syn22333694')$path, row.names=1)
+
+celFrac_ilr = ilr(df_cellFractions)
+colnames(celFrac_ilr) = paste0("cellFrac_ilr_", 1:3)
+
+METADATA = merge(METADATA, df_cellFractions, by.x = 'SampleID', by.y='row.names')
+METADATA = merge(METADATA, celFrac_ilr, by.x = 'SampleID', by.y='row.names')
+```
+
 ```{r preprocess.data1, results='asis'}
@@
 rownames(METADATA) = METADATA$SampleID
-METADATA = METADATA[indToRetain,]
-
-# METADATA %>% 
-#   group_by(Dx) %>%  
-#   dplyr::summarise(count = n()) %>% 
-#   kable() 
 ```
@@
   kable(caption="Diagnosis, Institution, Sex") %>% kable_styling(full_width=FALSE)
+
+ggplot(METADATA, aes(Reported_Gender, ageOfDeath)) + geom_boxplot() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + facet_wrap(~Institution, nrow=1)
+```
+
+### Sex by cell fraction
+```{r summary.sex, fig.width=8, fig.height=10}
+figList = lapply(colnames(df_cellFractions), function(CT){
+  ggplot(METADATA, aes_string('Reported_Gender', CT)) + geom_boxplot() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + facet_wrap(~Institution, nrow=1)
+})
+plot_grid(plotlist=figList, ncol=1) 
+```
+
+### Age by cell fraction
+```{r summary.Age, fig.width=8, fig.height=10}
+figList = lapply(colnames(df_cellFractions), function(CT){
+  ggplot(METADATA, aes_string('ageOfDeath', CT)) + geom_point() + geom_smooth(method="gam") + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + facet_wrap(~Institution, nrow=1)
+}) 
+plot_grid(plotlist=figList, ncol=1) 
+```
+
+## Age by cell fraction log ratio
+```{r summary.Age.log.ratio, fig.width=8, fig.height=8}
+figList = lapply(colnames(celFrac_ilr), function(CT){
+  ggplot(METADATA, aes_string('ageOfDeath', CT)) + geom_point() + geom_smooth(method="gam") + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + facet_wrap(~Institution, nrow=1)
+})
+plot_grid(plotlist=figList, ncol=1) 
 ```
@@
 # keep genes that are express in at least one combination of cohort and gender
-isexpr.lst = dlply(METADATA, .(Cohort, Reported_Gender), .fun = function(mtd){
-	rowSums(cpm(geneCountsMerged[,mtd$SampleID])>1) >= 0.5*ncol(geneCountsMerged[,mtd$SampleID])
+isexpr.lst = dlply(METADATA, .(Cohort), .fun = function(mtd){
+	rowSums(cpm(geneCountsMerged[,mtd$SampleID])>.5) >= .3*ncol(geneCountsMerged[,mtd$SampleID])
 	})
@@
 isexpr = do.call(rbind, isexpr.lst)
-isexpr = colSums(isexpr) >= 3
+isexpr = colSums(isexpr) == 2 # keep gene expressed in both cohorts 
 table(isexpr)
@@
   dge = DGEList( geneCountsMerged[isexpr,METADATA$Cohort==Cohort] )
-  calcNormFactors(dge) # TMM
+  calcNormFactors(dge, "none") # TMM
 })
@@
 
-form = ~ Dx + Reported_Gender + RIN + ageOfDeath + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate)
+form.lst = list(
+  `MSSM-Penn-Pitt`  = ~ Dx + Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
+  `NIMH-HBCC`       = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)
 
@@
 
-  if( length(unique(info$Institution)) > 1 ){
-    form_use = as.formula(paste('~',as.character(form)[2], "+ Institution", collapse=" "))
-  }else{
-    form_use = form
-  }
+  form_use = form.lst[[info$Cohort[1]]]
 
@@
 names(vobj.lst) = names(dge.lst)
-
-# FORMULAs
-
-# HBCC - DLPFC: 
-# "Dx + scale(RIN) + ageOfDeath + (1 | Reported_Gender) + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate)"
-
-# HBCC dsACC: 
-# Dx + scale(RIN) + (1 | Reported_Gender) + ageOfDeath + (1 | IsolationBatch) + scale(IntragenicRate) + scale(IntronicRate) + scale(IntergenicRate)"
-
-# MPP ACC
- # Dx + scale(RIN) + scale(ageOfDeath) + (1 | Institution) + (1 | Reported_Gender) + scale(RIN2) + scale(IntronicRate) + scale(rRNARate) + scale(IntragenicRate) + scale(IntergenicRate)
-
-# MPP DLPFC
-# Dx + scale(RIN2) + (1 | Institution) + scale(ageOfDeath) + scale(RIN) + (1 | Reported_Gender) + scale(IntronicRate) + scale(IntragenicRate)
 ```
@@
 
-### Test marginals
+<!---
+
+# Examine genes on chrX and chrY
+
+# dge = DGEList(geneCountsMerged)
+# dge = calcNormFactors(dge, "none")
+# geneExpr = cpm(dge, log=TRUE)
+
+# df = merge( data.frame(Expression = geneExpr['ENSG00000229807.11',]), METADATA, by="row.names")
+# ggplot(df, aes(Reported_Gender, Expression)) + geom_violin() + theme_bw() + facet_wrap(~Cohort)
+
+# id = 'ENSG00000229807.11'
+# isexpr[id]
+
+# df = as.data.frame(vobj.lst[[2]]$voom.xy)
+# with(df, plot(x,y, cex=.2, pch=20))
+# with(df[id,], points(x,y, col="red", pch=20, cex=2))
+
+--->
+
+```{r Create.residuals}
+
+### Create residuals
+form.lst = list(
+  `MSSM-Penn-Pitt`  = ~ Dx*Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
+  `NIMH-HBCC`       = ~ Dx*Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)
+
+# Compute residuals, but add back in Dx*Reported_Gender
+f = function(fit){
+  residuals(fit) + get_prediction( fit, ~ Dx*Reported_Gender)
+}
+
+residuals.lst = lapply( vobj.lst, function(vobj){
+
+  i = match(colnames(vobj), rownames(METADATA))
+  info = METADATA[i,]
+
+  form_use = form.lst[[info$Cohort[1]]]
+
+  # Compute residuals which adding back Diagnosis and Sex effects
+  resid.lst = fitVarPartModel(vobj, form_use, info, fxn = f, BPPARAM = SnowParam(18))  
+  do.call(rbind, resid.lst)
+  })
+names(residuals.lst) = names(vobj.lst)
+```
+
+```{r check.correlation}
+ngenes = 300
+C1 = cor(t(residuals.lst[[1]][1:ngenes,]))
+C2 = cor(t(residuals.lst[[2]][1:ngenes,]))
+
+df = data.table(A   = as.numeric(C1[lower.tri(C1)]),
+                B   = as.numeric(C2[lower.tri(C2)]))
+
+df$density <- get_density(df$A, df$B, n = 80)
+ggplot(df, aes(A, B, color=density)) + geom_point() + geom_point(size=.4) + theme_bw(15) + theme(aspect.ratio=1, legend.position="bottom", plot.title = element_text(hjust = 0.5)) + geom_abline(color="red") + scale_color_viridis() + xlim(-1,1) + ylim(-1,1) + xlab(paste0('Correlation (', names(residuals.lst)[1], ')')) + ylab(paste0('Correlation (', names(residuals.lst)[2], ')'))
+```
+
+# Schizophrenia vs controls
+
+### Test marginals for Scz vs controls
 ```{r differential.expression.simple}
-
-form = ~ Dx + Reported_Gender + RIN + ageOfDeath + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate)
+form.lst = list(
+  `MSSM-Penn-Pitt`  = ~ Dx + Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
+  `NIMH-HBCC`       = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)
 
@@
 
-  if( length(unique(info$Institution)) > 1 ){
-    form_use = as.formula(paste('~',as.character(form)[2], "+ Institution", collapse=" "))
-  }else{
-    form_use = form
-  }
+  form_use = form.lst[[info$Cohort[1]]]
 
-  fit = dream(vobj, form, info, BPPARAM = SnowParam(18))  
+  fit = dream(vobj, form_use, info, BPPARAM = SnowParam(18))  
   eBayes( fit )
@@
 names(fit_marginal.lst) = names(vobj.lst)
+
+coefTest = c('DxSCZ', 'Reported_GenderMale', 'ageOfDeath')
 ```
 
-```{r load.genesets}
-library(zenith)
-
+```{r load.genesets, eval=TRUE}
 # Load Gene Ontology database 
@@
 
-```{r run.zenith}
-
-coefTest = c('DxSCZ', 'Reported_GenderMale')
+```{r run.zenith, eval=TRUE}
 res_zenith = lapply( names(fit_marginal.lst), function(key){
@@
 
-  res.GO = run_zenith( fit, coefTest, gs.GO.collection )
+  res.GO = run_zenith(  fit, 
+                        coefTest,
+                        gs.GO.collection)
   res.GO$Resource = "GO"   
@@
 
-  res.MSigDB = run_zenith( fit, coefTest, gs.MSigDB.collection )
+  res.MSigDB = run_zenith(fit, 
+                          coefTest,
+                          gs.MSigDB.collection ) 
   res.MSigDB$Resource = "MSigDB"
@@
   rbind(res.GO, res.MSigDB)
-
 })
 names(res_zenith) = names(fit_marginal.lst)
+
+# only autosomes
+res_zenith_sex_autosomes = lapply( names(fit_marginal.lst), function(key){
+
+  fit = fit_marginal.lst[[key]]
+
+  # only keep autosomes, for Reported_GenderMale
+  tab = coef(fit) %>% getGeneSymbol()
+  geneids = rownames(tab)[tab$Chrom %in% 1:22]
+  fit = fit[geneids,]
+
+  res.GO = run_zenith(  fit, 
+                        'Reported_GenderMale',
+                        gs.GO.collection)
+  res.GO$Resource = "GO"   
+  res.GO$Cohort = key
+
+  res.MSigDB = run_zenith(fit, 
+                          'Reported_GenderMale',
+                          gs.MSigDB.collection ) 
+  res.MSigDB$Resource = "MSigDB"
+  res.MSigDB$Cohort = key
+
+  rbind(res.GO, res.MSigDB)
+})
+names(res_zenith_sex_autosomes) = names(fit_marginal.lst)
+```
+
+## Meta-analysis of zenith results
+```{r zenith.meta, eval=TRUE}
+res_meta_zenith = zenith_meta_analysis( res_zenith )                  
+res_meta_zenith.sex_autosome = zenith_meta_analysis( res_zenith_sex_autosomes )
 ```
@@
     res$Coef = coef
+    res %>% getGeneSymbol()  
+    })
+  names(res) = coefTest
   res
-    })
-  do.call(rbind, res) %>% getGeneSymbol
   })
@@
 # Schizophrenia testing
-# extract from each chort
+# extract from each cohort
 coef = 'DxSCZ'
+# ymin = min(sapply(df_DE_results, function(x) min(x[[coef]]$P.Value)))
+
 figList.Scz = lapply( names(df_DE_results), function(key){
-  df = df_DE_results[[key]]
-  df = df[df$Coef == coef,]
+  df = df_DE_results[[key]][[coef]]
 
   main = paste(key, coef, sep = ': ' )
-  plotVolcano( df, df$Symbol[1:20] ) + ggtitle( main )
+  plotVolcano( df, df$Symbol[1:10] ) + ggtitle( main ) #+ ylim(0, -log10(ymin))
 })
@@
 
-```{r meta.analysis.SCZ}
-# Create list preparing for meta-analysis for SCZ
-tab.lst = lapply( df_DE_results, function(df){
-  df[df$Coef == coef,]
+## Meta-analysis across cohorts
+```{r meta.analysis.DE}
+# Store results for each coef and cohort
+tab.de.lst = lapply(coefTest, function(coef){    
+  tab.lst = lapply( df_DE_results, function(lst){
+    lst[[coef]]
   })
   names(tab.lst) = names(df_DE_results)
+  tab.lst
+  })
+names(tab.de.lst) = coefTest
 
+# Create list preparing for meta-analysis for SCZ
+df_meta = mclapply(coefTest, function(coef){
   # Run meta-analysis
-df_meta_SCZ = run_meta_analysis( tab.lst, "FE") %>% getGeneSymbol
+  run_meta_analysis( tab.de.lst[[coef]], "FE") %>% getGeneSymbol
+}, mc.cores=6)
+names(df_meta) = coefTest
 ```
@@
 # Concordance analysis
-resConcord.SCZ = plot_concordance(tab.lst, showGenes = df_meta_SCZ$Symbol[1:20])
+resConcord.SCZ = plot_concordance(tab.de.lst[[coef]], showGenes = df_meta[[coef]]$Symbol[1:20])
 
@@
 
-data.frame(pi1 = resConcord.SCZ$pi_single) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)
-resConcord.SCZ$tab_pi %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
-resConcord.SCZ$tab_corr %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
+data.frame(Cohort = names(tab.de.lst[[coef]]),
+          "# DE"  = resConcord.SCZ$n.de,
+          pi1     = resConcord.SCZ$pi_single,
+          check.names=FALSE) %>% 
+          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)
+
+resConcord.SCZ$tab_pi %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
+resConcord.SCZ$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
 ```
@@
 main = paste('Meta-analysis', coef, sep = ': ' )
-plotVolcano( df_meta_SCZ, df_meta_SCZ$Symbol[1:20] ) + ggtitle( main )
+plotVolcano( df_meta[[coef]], df_meta[[coef]]$Symbol[1:20] ) + ggtitle( main )
+
+data.frame(pi1 = 1 - qvalue(df_meta[[coef]]$P)$pi0,
+          `# DE` = sum(df_meta[[coef]]$adj.P.Val < 0.05),
+          check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for SCZ meta-analysis") %>% kable_styling(full_width=FALSE)
+
+```
+
+## Plot top examples
+```{r plot.forrest.dlpfc, fig.width=8, fig.height=20}  
+plot_gene = function(residuals.lst, df_meta, geneSymbol, variable, fill=c("#1E90FF", "#FF8D1E"), mode=""){
+
+  ensGene = rownames(df_meta)[which(df_meta$Symbol == geneSymbol)]
+
+  exprResid = lapply( names(residuals.lst), function(key){
+      data.frame(Expression = residuals.lst[[key]][ensGene,])
+    })
+  exprResid = do.call(rbind, exprResid)
+
+  df = merge(exprResid, METADATA, by="row.names")
+
+  if( is.factor(df[[variable]]) || length(unique(df[[variable]])) < 5 ){
+    fig = ggplot(df, aes_string(variable, 'Expression')) + geom_violin(aes_string(fill=variable)) + geom_boxplot(width=0.1, fill="grey80") + scale_fill_manual(values = fill)
+  }else{    
+    fig = ggplot(df, aes_string(variable, 'Expression')) + geom_point() + geom_smooth(method="lm") #+ scale_color_manual(values = fill)
+  }
+
+  fig + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + ylab('Expression residuals') + ggtitle(geneSymbol) + facet_wrap(~Cohort)
+}
+
+coef = 'DxSCZ'
+figList = lapply(df_meta[[coef]]$Symbol[1:10], function(geneSymbol){
+  fig1 = plot_forrest( c(tab.de.lst[[coef]], list("Meta-analysis" = df_meta[[coef]])), geneSymbol)
+
+  fig2 = plot_gene(residuals.lst, df_meta[[coef]], geneSymbol, "Dx") + xlab("Diagnosis")
+
+  plot_grid(fig1, fig2, nrow=1)
+})
+
+plot_grid(plotlist=figList, ncol=1)
+```
+
+### Concordance with other datasets
+```{r CMC.concordance}
+# CMC Analysis 1
+CMC1 <- readr::read_tsv(synGet("syn5607603")$path) %>% 
+  dplyr::rename(gene_id = genes,
+                `CMC 1 t-statistic` = t) %>% 
+  dplyr::select(gene_id, `CMC 1 t-statistic`)
+# HBCC-DLPFC microarray
+CMC_HBCC <- readr::read_tsv(synGet("syn4941553")$path) %>% 
+  dplyr::rename(gene_id = genes,
+                `HBCC microarray t-statistic` = t) %>% 
+  dplyr::select(gene_id, `HBCC microarray t-statistic`)
+# Gandal, Science 2018, Table S1
+# Use Tab:DGE
+# wget https://science.sciencemag.org/highwire/filestream/720110/field_highwire_adjunct_files/1/aat8127_Table_S1.xlsx
+gandal <- readxl::read_xlsx( synGet("syn22255217")$path, sheet = "DGE") %>% 
+  dplyr::rename(gene_id = ensembl_gene_id,
+                `Gandal, SCZ t-statistic` = `SCZ.t.value`,
+                `Gandal, BD t-statistic` = `BD.t.value`,
+                `Gandal, ASD t-statistic` = `ASD.t.value`) %>%
+  dplyr::select(gene_id, `Gandal, SCZ t-statistic`,
+                `Gandal, BD t-statistic`, `Gandal, ASD t-statistic`)
+# AMP-AD RNASeq reprocessing GRCh37 - https://www.synapse.org/#!Synapse:syn9702085
+MSSM <- readr::read_tsv(synGet("syn10526259")$path) %>%
+  dplyr::filter(Comparison %in% "AD-CONTROL",
+                Model %in% "Diagnosis") %>%
+  dplyr::select(ensembl_gene_id, Tissue, t) %>%
+  dplyr::group_by(ensembl_gene_id, Tissue) %>% 
+  dplyr::filter(row_number() == 1) %>% 
+  tidyr::pivot_wider(names_from = Tissue,
+                     values_from = t) %>%
+  dplyr::rename(`AMP-AD MSSM FP AD t-statistic` = FP,
+                `AMP-AD MSSM IGF AD t-statistic` = IFG,
+                `AMP-AD MSSM PHG AD t-statistic` = PHG,
+                `AMP-AD MSSM STG AD t-statistic` = STG,
+                gene_id = ensembl_gene_id) %>% 
+  dplyr::select(gene_id, `AMP-AD MSSM STG AD t-statistic`) 
+MSSM <- MSSM[!is.na(MSSM$`AMP-AD MSSM STG AD t-statistic`),]
+
+MAYO <- readr::read_tsv(synGet("syn8468023")$path) %>% 
+  dplyr::filter(Comparison %in% "AD-CONTROL",
+                Model %in% "Diagnosis") %>%
+  dplyr::select(ensembl_gene_id, `Tissue.ref`, t) %>%
+  dplyr::group_by(ensembl_gene_id, `Tissue.ref`) %>% 
+  dplyr::filter(row_number() == 1) %>% 
+  tidyr::pivot_wider(names_from = `Tissue.ref`,
+                     values_from = t) %>% 
+  dplyr::rename(`AMP-AD MAYO CBE AD t-statistic` = CBE,
+                `AMP-AD MAYO TCX AD t-statistic` = TCX,
+                gene_id = ensembl_gene_id) %>% 
+  dplyr::select(-`AMP-AD MAYO TCX AD t-statistic`)
+
+ROSMAP <- readr::read_tsv(synGet("syn8456721")$path) %>%
+    dplyr::filter(Comparison %in% "AD-CONTROL",
+                Model %in% "Diagnosis") %>%
+  dplyr::rename(gene_id = ensembl_gene_id, 
+                `AMP-AD ROSMAP DLPFC AD t-statistic` = t) %>% 
+  dplyr::select(gene_id, `AMP-AD ROSMAP DLPFC AD t-statistic`)
+
+out_group <- inner_join(MSSM, MAYO) %>% 
+  inner_join(., ROSMAP) %>% 
+  inner_join(., gandal)
+
+# merge with meta-analysis
+
+coef = 'DxSCZ'
+df = df_meta[[coef]]
+df_two = data.frame(gene_id = trim_ensembl_ids(rownames(df)),
+                `CMC - meta-analysis` = with(df, logFC/se), check.names=FALSE)
+
+df = tab.de.lst[[coef]][[1]]
+df.tmp = data.frame(gene_id = trim_ensembl_ids(rownames(df)),
+                `t` = df$t, check.names=FALSE)
+colnames(df.tmp)[2] = names(tab.de.lst[[coef]])[1]
+df_two = df_two %>% inner_join(., df.tmp)
+
+df = tab.de.lst[[coef]][[2]]
+df.tmp = data.frame(gene_id = trim_ensembl_ids(rownames(df)),
+                `t` = df$t, check.names=FALSE)
+colnames(df.tmp)[2] = names(tab.de.lst[[coef]])[2]
+df_two = df_two %>% inner_join(., df.tmp)
+
+# Remove transcript Ids
+df_two <- df_two %>% 
+  tidyr::separate(gene_id, c("gene_id", "transcript")) %>% 
+  dplyr::select(-one_of("transcript"))
+df_two <- inner_join(df_two, CMC1, by = c("gene_id")) 
+df_two <- inner_join(df_two, CMC_HBCC, by = c("gene_id"))
+df_two <- inner_join(df_two, out_group, by = c("gene_id")) %>% 
+  group_by(gene_id) %>% 
+  dplyr::filter(row_number()==1) %>% 
+  tibble::column_to_rownames(var = "gene_id")
+
+colnames(df_two) <- gsub(" t-statistic", "", colnames(df_two))
+
+col2 <- colorRampPalette(rev(c("#67001F", "#B2182B", "#D6604D", "#F4A582",
+                           "#FDDBC7", "#FFFFFF", "#D1E5F0", "#92C5DE",
+                           "#4393C3", "#2166AC", "#053061")))
+
+cor <- cor(df_two)
+# clustering based on distance
+hcl = hclust(as.dist(1-cor), method='ward.D2')
+
+results <- pvclust(cor, method.hclust='ward.D2')
+# reorder based on clustering
+reorder = cor[results$hclust$order, results$hclust$order]
+
+# plot clustering
+plot(results, hang=-1)
+# plot(hcl, hang=-1)
+corrplot(reorder, col = col2(200))
+```
+
+```{r test.multtest, eval=FALSE}
+
+library(irlba)
+library(pinnacle)
+library(poolR)
+library(matrixStats)
+
+i = with(res_meta_zenith, (Resource == 'GO') & (Coef == "DxSCZ"))
+df = res_meta_zenith[i,]
+df$qvalue = qvalue(df$PValue)$qvalue
+# df$p.adjust = p.adjust(df$PValue, method="bon" )
+df$p.adjust = pmin(1, 2805 * df$PValue)
+
+M = recodeToSparseMatrix(gs.GO)
+M = M[rownames(M) %in% trim_ensembl_ids(rownames(df_meta_SCZ)),]
+keep = (colSums(M) >= 10) & (colnames(M) %in% rownames(df) )
+M = M[,keep]
+
+m = colMeans(M)
+s = apply(M, 2, var) * nrow(M) / (nrow(M) + 1)
+
+dcmp = irlba( M, nv = 500, center=m, scale=sqrt(s), right_only=FALSE)
+ev = dcmp$d^2 / nrow(M)
+
+plot(cumsum(ev) / ncol(M))
+
+df_unsort = df[colnames(M),]
+edcmp = list(vectors = dcmp$v, values = dcmp$d^2 / nrow(M))
+
+res = p.corr.adjust( df_unsort$PValue, dcmp = edcmp, method.sim="eigen", B = nrow(df_unsort)*4)
+
+plot( cumsum(dcmp$d^2 / sum(dcmp$d^2)))
+
+C = cor(as.matrix(M[,1:1000]))
+evals = eigen(C)$values
+
+dcmp = svd(scale(M) / sqrt(nrow(M)))
+dcmp$d^2
+
+plot(cumsum(dcmp$d^2 / sum(dcmp$d^2)))
+
+meff(C, method='gao')
+
+dcmp = svd(scale(M[,1:100]))
+
+meff(eigen = dcmp$d^2 , method = "galwey")
+
+```
+
+```{r show.gene_set_test.marginal, eval=TRUE}
+show_zenith_results = function( res, coef, Resource, FDR = 0.05){
+
+  i = which((res$Resource == Resource) & (res$Coef == coef))
+  df = res[i,]
+
+	# filter by FDR
+	df = df[df$FDR <= FDR,]
+  df$Coef = c()
+  df$se = c()
+  df$Resource = c()
+
+	df %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption=paste(Resource, coef), row.names=FALSE) %>% kable_styling(full_width=FALSE)
+}
+
+show_zenith_results( res_meta_zenith, "DxSCZ", "GO", FDR=0.05)
+show_zenith_results( res_meta_zenith, "DxSCZ", "MSigDB", FDR=0.05)
+```
+
+# Sex Analysis
+```{r test.Sex, fig.height=5, fig.width=10}
+# Schizophrenia testing
+# extract from each chort
+coef = 'Reported_GenderMale'
+figList.Sex = lapply( names(df_DE_results), function(key){
+  df = df_DE_results[[key]][[coef]]
+  main = paste(key, coef, sep = ': ' )
+  plotVolcano( df, df$Symbol[1:20] ) + ggtitle( main )
+})
+
+# volcano plots
+plot_grid(plotlist=figList.Sex, nrow=1)
+```
+
+### Autosomes
+```{r test.Sex.autosome, fig.height=5, fig.width=10}
+# Schizophrenia testing
+# extract from each chort
+coef = 'Reported_GenderMale'
+figList.Sex = lapply( names(df_DE_results), function(key){
+  df = df_DE_results[[key]][[coef]]
+  main = paste(key, coef, sep = ': ' )
+  df = df[df$Chrom %in% 1:22,]
+  plotVolcano( df, df$Symbol[1:20] ) + ggtitle( main )
+})
+
+# volcano plots
+plot_grid(plotlist=figList.Sex, nrow=1)
+```
+
+## Meta-analysis across cohorts
+```{r concordance.Sex, fig.height=6, fig.width=10}
+# Concordance analysis
+resConcord.Sex = plot_concordance(tab.de.lst[[coef]], showGenes = df_meta[[coef]]$Symbol[1:20])
+
+main = "Concordance of Sex"
+plot_grid(  resConcord.Sex$fig_logFC + ggtitle(main),
+            resConcord.Sex$fig_tstat + ggtitle(main), nrow=1)
+
+data.frame(Cohort = names(tab.de.lst[[coef]]),
+          "# DE"  = resConcord.Sex$n.de,
+          pi1     = resConcord.Sex$pi_single,
+          check.names=FALSE) %>% 
+          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)
+
+resConcord.Sex$tab_pi  %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
+resConcord.Sex$tab_corr  %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
+```
+
+```{r concordance.Sex.autosome, fig.height=6, fig.width=10}
+
+tab.de.lst.autosome = lapply(tab.de.lst[[coef]], function(x) x[x$Chrom %in% 1:22,])
+
+# Concordance analysis
+resConcord.Sex = plot_concordance(tab.de.lst.autosome, showGenes = tab.de.lst.autosome$Symbol[1:20])
+
+main = "Concordance of Sex - autosomes"
+plot_grid(  resConcord.Sex$fig_logFC + ggtitle(main),
+            resConcord.Sex$fig_tstat + ggtitle(main), nrow=1)
+
+data.frame(Cohort = names(resConcord.Sex$pi_single),
+          "# DE"  = resConcord.Sex$n.de,
+          pi1     = resConcord.Sex$pi_single,
+          check.names=FALSE) %>% 
+          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)
+
+resConcord.Sex$tab_pi %>% mutate_if(is.numeric, format, digits=3)  %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
+resConcord.Sex$tab_corr  %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
+```
+
+```{r plot.meta.analysis.Sex, fig.height=5, fig.width=10}
+# volcano plot for meta-analysis
+main = 'Meta-analysis: Sex'
+fig1 = plotVolcano( df_meta[[coef]], df_meta[[coef]]$Symbol[1:20] ) + ggtitle( main )
+
+df = df_meta[[coef]][df_meta[[coef]]$Chrom %in% 1:22,]
+fig2 = plotVolcano( df, df$Symbol[1:20] ) + ggtitle( paste(main, "autosome") )
+plot_grid(fig1, fig2, ncol=2)
+
+data.frame(pi1 = 1 - qvalue(df_meta[[coef]]$P)$pi0,
+          `# DE` = sum(df_meta[[coef]]$adj.P.Val < 0.05),
+          check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for Sex meta-analysis") %>% kable_styling(full_width=FALSE)
+```
+
+## Sex circos
+```{r sex.circos}
+coef = 'Reported_GenderMale'
+sapply( names(df_DE_results), function(key){
+  plot_circos_logFC( df_DE_results[[key]][[coef]] ) + ggtitle(key)
+  })
+
+plot_circos_logFC( df_meta[[coef]] ) + ggtitle("Meta analysis")
+```
+
+## Plot top examples
+```{r plot.forrest.dlpfc.Sex, fig.width=8, fig.height=20}  
+figList = lapply(df_meta[[coef]]$Symbol[1:10], function(geneSymbol){
+
+  fig1 = plot_forrest( c(tab.de.lst[[coef]], list("Meta-analysis" = df_meta[[coef]])), geneSymbol)
+
+  fig2 = plot_gene( residuals.lst, df_meta[[coef]], geneSymbol, "Reported_Gender", fill=c("red", "blue")) + xlab("Sex")
+
+  plot_grid(fig1, fig2, nrow=1)
+})
+
+plot_grid(plotlist=figList, ncol=1)
+```
+
+## Meta-analysis of zenith results
+```{r show.gene_set_test.marginal.Sex, eval=TRUE}
+show_zenith_results( res_meta_zenith, "Reported_GenderMale", "GO", FDR=0.05)         
+show_zenith_results( res_meta_zenith, "Reported_GenderMale", "MSigDB", FDR=0.05)        
+```
+
+### Only autosomes
+```{r show.gene_set_test.marginal.Sex.autosome, eval=TRUE}
+show_zenith_results( res_meta_zenith.sex_autosome, "Reported_GenderMale", "GO", FDR=0.05) 
+show_zenith_results( res_meta_zenith.sex_autosome, "Reported_GenderMale", "MSigDB", FDR=0.05)   
+```
+
+# Age Analysis
+```{r test.Age, fig.height=5, fig.width=10}
+# Schizophrenia testing
+# extract from each chort
+coef = 'ageOfDeath'
+figList.Age = lapply( names(df_DE_results), function(key){
+  df = df_DE_results[[key]][[coef]]
+  main = paste(key, coef, sep = ': ' )
+  plotVolcano( df, df$Symbol[1:20] ) + ggtitle( main )
+})
+
+# volcano plots
+plot_grid(plotlist=figList.Age, nrow=1)
+```
+
+## Meta-analysis across cohorts
+```{r concordance.Age, fig.height=6, fig.width=10}
+# Concordance analysis
+resConcord.Age = plot_concordance(tab.de.lst[[coef]], showGenes = df_meta[[coef]]$Symbol[1:20])
+
+main = "Concordance of Age"
+plot_grid(  resConcord.Age$fig_logFC + ggtitle(main),
+            resConcord.Age$fig_tstat + ggtitle(main), nrow=1)
+
+data.frame(Cohort = names(tab.de.lst[[coef]]),
+          "# DE"  = resConcord.Age$n.de,
+          pi1     = resConcord.Age$pi_single,
+          check.names=FALSE) %>% 
+          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)
+
+resConcord.Age$tab_pi %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
+resConcord.Age$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
+```
+
+```{r plot.meta.analysis.Age, fig.height=5, fig.width=10}
+# volcano plot for meta-analysis
+main = 'Meta-analysis: Age'
+plotVolcano( df_meta[[coef]], df_meta[[coef]]$Symbol[1:20] ) + ggtitle( main )
+
+data.frame(pi1 = 1 - qvalue(df_meta[[coef]]$P)$pi0,
+          `# DE` = sum(df_meta[[coef]]$adj.P.Val < 0.05),
+          check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for Age meta-analysis") %>% kable_styling(full_width=FALSE)
+```
+
+## Plot top examples
+```{r plot.forrest.dlpfc.Age, fig.width=8, fig.height=20}  
+figList = lapply(df_meta[[coef]]$Symbol[1:10], function(geneSymbol){
+
+  fig1 = plot_forrest( c(tab.de.lst[[coef]], list("Meta-analysis" = df_meta[[coef]])), 
+    geneSymbol)
+
+  expr.list = lapply(vobj.lst, function(x)x$E)
+
+  fig2 = plot_gene( expr.list, df_meta[[coef]], geneSymbol, "ageOfDeath", fill=c("red", "blue")) + xlab("Age")
+
+  plot_grid(fig1, fig2, nrow=1)
+})
+
+plot_grid(plotlist=figList, ncol=1)
+```
+
+## Meta-analysis of zenith results
+```{r show.gene_set_test.marginal.Age, eval=TRUE}
+show_zenith_results( res_meta_zenith, "ageOfDeath", "GO", FDR=0.01)   
+show_zenith_results( res_meta_zenith, "ageOfDeath", "MSigDB", FDR=0.01)    
+```
+
+# Sex by Disease interaction
+```{r dream.sex.disease}
+form.lst = list(
+  `MSSM-Penn-Pitt`  = ~ Dx*Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
+  `NIMH-HBCC`       = ~ Dx*Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)
+
+fit_inter_sex.disease.lst = lapply( vobj.lst, function(vobj){
+
+  i = match(colnames(vobj), rownames(METADATA))
+  info = METADATA[i,]
+
+  form_use = form.lst[[info$Cohort[1]]]
+
+  fit = dream(vobj, form_use, info, BPPARAM = SnowParam(18))  
+  eBayes( fit )
+  })
+names(fit_inter_sex.disease.lst) = names(vobj.lst)
+
+coefTest = c('DxSCZ:Reported_GenderMale')
+```
+
+```{r run.zenith.inter_sex.disease, eval=TRUE}
+res_zenith.inter_sex.disease = lapply( names(fit_inter_sex.disease.lst), function(key){
+
+  fit = fit_inter_sex.disease.lst[[key]]
+
+  res.GO = run_zenith(  fit, 
+                        coefTest,
+                        gs.GO.collection)
+  res.GO$Resource = "GO"   
+  res.GO$Cohort = key
+
+  res.MSigDB = run_zenith(fit, 
+                          coefTest,
+                          gs.MSigDB.collection) 
+  res.MSigDB$Resource = "MSigDB"
+  res.MSigDB$Cohort = key
+
+  rbind(res.GO, res.MSigDB)
+})
+names(res_zenith.inter_sex.disease) = names(fit_inter_sex.disease.lst)
+
+# only autosomes
+res_zenith.inter_sex.disease_autosome = lapply( names(fit_inter_sex.disease.lst), function(key){
+
+  fit = fit_inter_sex.disease.lst[[key]]
+
+  # only keep autosomes, for Reported_GenderMale
+  tab = coef(fit) %>% getGeneSymbol()
+  geneids = rownames(tab)[tab$Chrom %in% 1:22]
+  fit = fit[geneids,]
+
+  res.GO = run_zenith(  fit, 
+                        coefTest,
+                        gs.GO.collection)
+  res.GO$Resource = "GO"   
+  res.GO$Cohort = key
+
+  res.MSigDB = run_zenith(fit, 
+                          coefTest,
+                          gs.MSigDB.collection) 
+  res.MSigDB$Resource = "MSigDB"
+  res.MSigDB$Cohort = key
+
+  rbind(res.GO, res.MSigDB)
+})
+names(res_zenith.inter_sex.disease_autosome) = names(fit_inter_sex.disease.lst)
+```
+
+## Meta-analysis of zenith results
+```{r zenith.meta.inter_sex.disease, eval=TRUE}
+res_meta_zenith.inter_sex.disease = zenith_meta_analysis( res_zenith.inter_sex.disease )  
+res_meta_zenith.inter_sex.disease_autosome = zenith_meta_analysis( res_zenith.inter_sex.disease_autosome )
+```
+
+```{r get.signatures.inter_sex.disease, fig.width=10, fig.height=5}
+# Get DE signatures
+df_DE_results.inter_sex.disease = lapply(fit_inter_sex.disease.lst, function(fit){
+  res = lapply(coefTest, function(coef){
+    res = topTable(fit, coef, number=Inf) 
+    res$Coef = coef
+    res %>% getGeneSymbol()  
+    })
+  names(res) = coefTest
+  res
+  })
+names(df_DE_results.inter_sex.disease) = names(fit_inter_sex.disease.lst)
+```
+
+```{r test.scz.inter_sex.disease, fig.height=5, fig.width=10}
+coef = 'DxSCZ:Reported_GenderMale'
+
+figList.Scz = lapply( names(df_DE_results), function(key){
+  df = df_DE_results.inter_sex.disease[[key]][[coef]]
+
+  main = paste(key, coef, sep = ': ' )
+  plotVolcano( df, df$Symbol[1:10] ) + ggtitle( main )
+})
+
+# volcano plots
+plot_grid(plotlist=figList.Scz, nrow=1)
+```
+
+## Meta-analysis across cohorts
+```{r meta.analysis.DE.inter_sex.disease}
+# Store results for each coef and cohort
+tab.de.inter_sex.disease.lst = lapply(coefTest, function(coef){    
+  tab.de.inter_sex.disease.lst = lapply( df_DE_results.inter_sex.disease, function(lst){
+    lst[[coef]]
+  })
+  names(tab.de.inter_sex.disease.lst) = names(df_DE_results.inter_sex.disease)
+  tab.de.inter_sex.disease.lst
+  })
+names(tab.de.inter_sex.disease.lst) = coefTest
+
+# Create list preparing for meta-analysis for SCZ
+df_meta.inter_sex.disease = mclapply(coefTest, function(coef){
+  # Run meta-analysis
+  run_meta_analysis( tab.de.inter_sex.disease.lst[[coef]], "FE") %>% getGeneSymbol
+}, mc.cores=6) 
+names(df_meta.inter_sex.disease) = coefTest
+```
+
+```{r concordance.scz.inter_sex.disease, fig.height=6, fig.width=10}
+# Concordance analysis.  
+resConcord.inter_sex.disease = plot_concordance(tab.de.inter_sex.disease.lst[[coef]], showGenes = df_meta.inter_sex.disease[[coef]]$Symbol[1:20])
+
+main = "Concordance of DxSCZ:Reported_GenderMale"
+plot_grid(  resConcord.inter_sex.disease$fig_logFC + ggtitle(main),
+            resConcord.inter_sex.disease$fig_tstat + ggtitle(main), nrow=1)
+
+data.frame(Cohort = names(tab.de.inter_sex.disease.lst[[coef]]),
+          "# DE"  = resConcord.inter_sex.disease$n.de,
+          pi1     = resConcord.inter_sex.disease$pi_single,
+          check.names=FALSE) %>% 
+          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)
+
+resConcord.inter_sex.disease$tab_pi %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
+resConcord.inter_sex.disease$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
+```
+
+```{r plot.meta.analysis.SCZ.inter_sex.disease, fig.height=5, fig.width=10}
+# volcano plot for meta-analysis.  
+main = paste('Meta-analysis', coef, sep = ': ' )
+plotVolcano( df_meta.inter_sex.disease[[coef]], df_meta.inter_sex.disease[[coef]]$Symbol[1:20] ) + ggtitle( main )
+
+data.frame(pi1 = 1 - qvalue(df_meta.inter_sex.disease[[coef]]$P)$pi0,
+          `# DE` = sum(df_meta.inter_sex.disease[[coef]]$adj.P.Val < 0.05),
+          check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for SCZ:Sex meta-analysis") %>% kable_styling(full_width=FALSE)
+```
+
+## Plot top examples
+```{r plot.forrest.dlpfc.inter_sex.disease, fig.width=8, fig.height=40}
+
+figList = lapply(df_meta.inter_sex.disease[[coef]]$Symbol[1:20], function(geneSymbol){
+  fig1 = plot_forrest( c(tab.de.inter_sex.disease.lst[[coef]], list("Meta-analysis" = df_meta.inter_sex.disease[[coef]])), geneSymbol, plot2='p')
+
+  fig2 = plot_gene(residuals.lst, df_meta.inter_sex.disease[[coef]], geneSymbol, "Dx") + xlab("Diagnosis") + facet_wrap(~Reported_Gender + Cohort, nrow=1)
+
+  plot_grid(fig1, fig2, nrow=1)
+})
+
+plot_grid(plotlist=figList, ncol=1)
+```
+
+## Meta-analysis of zenith results
+```{r show.gene_set_test.marginal.inter_sex.disease, eval=TRUE}
+coef = 'DxSCZ:Reported_GenderMale'   
+show_zenith_results( res_meta_zenith.inter_sex.disease, coef, "GO", FDR=0.1)  
+show_zenith_results( res_meta_zenith.inter_sex.disease, coef, "MSigDB", FDR=0.1)
+
+show_zenith_results( res_meta_zenith.inter_sex.disease_autosome, coef, "GO", FDR=0.1) 
+show_zenith_results( res_meta_zenith.inter_sex.disease_autosome, coef, "MSigDB", FDR=0.1)    
 ```
@@
 ```{r exit2, cache=FALSE}
-knitr::knit_exit()
-```
+
+# DIFFERENTIAL CORRLEATION NETWORKS
+# CLUSTER on C_Scz - C_Control
+# Like contrastive PCA
 
@@
 
-```{r show.gene_set_test.marginal}
-
-show_zenith_results = function( resList, key, FDR = 0.05){
-	df = resList[[key]]
-
-	# filter by FDR
-	df = df[df$FDR <= FDR,]
-
-	df %>% kable(caption=key) %>% kable_styling(full_width=FALSE)
-}
-
-show_zenith_results( res_zenith_marginal.GO, "Scz")
-show_zenith_results( res_zenith_marginal.MSigDB, "Scz")
-
-show_zenith_results( res_zenith_marginal.GO, "BP")
-show_zenith_results( res_zenith_marginal.MSigDB, "BP")
-
-show_zenith_results( res_zenith_marginal.GO, "Sex")
-show_zenith_results( res_zenith_marginal.MSigDB, "Sex")
-```
-
-### Test Diagnosis * Sex interaction
-```{r differential.expression.interaction.DxSex}
-
-# METADATA$Reported_Gender = factor(METADATA$Reported_Gender)
-# METADATA$Dx = factor(METADATA$Dx, c("BP", "Control", "SCZ"))
-
-form = ~ Dx*Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)
-
-# get contrasts
-Lbase = getContrast( vobj, form, METADATA, "RIN")
-Lbase[] = 0
-
-names(Lbase)
-
-L_SCZ = Lbase
-L_SCZ['DxSCZ:Reported_GenderMale'] = 1
-
-L_BP = Lbase
-L_BP['DxBP:Reported_GenderMale'] = 1
-
-L = cbind(	SczSex = L_SCZ,
-			BPSex = L_BP)
-
-plotContrasts(L) + theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust=1))
-
-fit_interaction_DxSex = dream(vobj, form, METADATA, L=L, BPPARAM = SnowParam(18))
-```
-
-```{r plots.interaction.DxSex, fig.width=10, fig.height=5}
-# Disease. 
-tab = topTable(fit_interaction_DxSex, "SczSex", number=Inf) %>% getGeneSymbol
-fig1 = plotVolcano( tab, tab$Symbol[1:20] ) + ggtitle("SCZ")
-
-tab = topTable(fit_interaction_DxSex, "BPSex", number=Inf)  %>% getGeneSymbol
-fig2 = plotVolcano( tab, tab$Symbol[1:20] ) + ggtitle("BP")
-
-plot_grid( fig1, fig2, ncol=2) 
-```
-
-```{r gene_set_test.DxSex}
-res_zenith_DxSex.GO = run_zenith( fit_interaction_DxSex, colnames(L), gs.GO.collection )
-res_zenith_DxSex.MSigDB = run_zenith( fit_interaction_DxSex, colnames(L), gs.MSigDB.collection )     
-```
-
-```{r show.gene_set_test.DxSex}
-show_zenith_results( res_zenith_DxSex.GO, "DxSCZ:Reported_GenderMale")
-show_zenith_results( res_zenith_DxSex.MSigDB, "DxSCZ:Reported_GenderMale")
-
-show_zenith_results( res_zenith_DxSex.GO, "DxBP:Reported_GenderMale")
-show_zenith_results( res_zenith_DxSex.MSigDB, "DxBP:Reported_GenderMale")
-```
-
-```{r exit, cache=FALSE}
 knitr::knit_exit()
 ```
-
-### Test Region * Sex interaction
-```{r differential.expression.interaction.RegionSex, eval=FALSE}
-
-form = ~ 0 + Dx + Region*Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Institution:Cohort) + (1|Cohort) + (1|Individual_ID)
-
-# get contrasts
-Lbase = getContrast( vobj, form, METADATA, "RIN")
-Lbase[] = 0
-
-# SCZ by sex
-L_SCZ = Lbase
-L_SCZ['DxSCZ:Reported_GenderMale'] = 1
-L_SCZ['DxControl:Reported_GenderMale'] = -1
-L_SCZ['DxSCZ'] = -1
-L_SCZ['DxControl'] = 1 
-# (A - B) - (C - D) = A - B - C + D
-
-L = cbind(	SczSex = L_SCZ)
-
-plotContrasts(L) + theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust=1))
-
-fit_interaction_RegionSex = dream(vobj, form, METADATA, L=L,BPPARAM = SnowParam(18))
-```
-
-### Test Region * Disease
-```{r differential.expression.interaction.RegionDx, eval=FALSE}
-
-form = ~ 0 + Region*Dx + Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Institution:Cohort) + (1|Cohort) + (1|Individual_ID)
-
-# get contrasts
-Lbase = getContrast( vobj, form, METADATA, "RIN")
-Lbase[] = 0
-
-# SCZ by sex
-L_SCZ = Lbase
-L_SCZ['DxSCZ:Reported_GenderMale'] = 1
-L_SCZ['DxControl:Reported_GenderMale'] = -1
-L_SCZ['DxSCZ'] = -1
-L_SCZ['DxControl'] = 1 
-# (A - B) - (C - D) = A - B - C + D
-
-L = cbind(	SczSex = L_SCZ)
-
-plotContrasts(L) + theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust=1))
-
-fit_interaction_RegionSex = dream(vobj, form, METADATA, L=L,BPPARAM = SnowParam(18))
-```

### 2020-09-16 common_functions/common_functions.R
fix bugs

@@ -17,9 +17,9 @@ trim_ensembl_ids = function(x){
 # Add column Symbol using column gene_id storing ENSEMBL id
 getGeneSymbol = function( df, column="row.names"){
 
-  # load load since data base can fail with rmarkdown run multiple times
-  # detach("package:EnsDb.Hsapiens.v86", unload=TRUE)
-  # library(EnsDb.Hsapiens.v86)
+  if( ! is(df, "data.frame") ){
+    df = as.data.frame(df)
+  }
 
   # Load ENSEMBL v96 database
   # see https://www.bioconductor.org/packages/release/bioc/vignettes/ensembldb/inst/doc/ensembldb.html#10_getting_or_building_ensdb_databasespackages
@@ -94,13 +94,16 @@ plot_concordance = function( resList, showGenes = NULL, idx=c(1,2), size=15){
   names(pi_single) = names(resList)
 
   p = with(df, P.Value.y[adj.P.Val.x < 0.05])
+  n.de.x = length(p)
   pi_discovery_x = tryCatch( 
     1 - qvalue(p)$pi0, 
     error = function(e){
-      NA
+      # must have a p-value > 0.95 to work
+      1 - qvalue(c(1,p))$pi0
       })
 
   p = with(df, P.Value.x[adj.P.Val.y < 0.05])
+  n.de.y = length(p)
   pi_discovery_y = tryCatch( 
     1 - qvalue(p)$pi0, 
     error = function(e){
@@ -112,6 +115,7 @@ plot_concordance = function( resList, showGenes = NULL, idx=c(1,2), size=15){
   list(fig_logFC  = fig1,
       fig_tstat   = fig2, 
       pi_single   = pi_single,
+      n.de        = c(n.de.x, n.de.y),
       tab_pi      = tab_pi,
       tab_corr    = tab_corr)
 }
@@ -211,18 +215,21 @@ run_meta_analysis = function(tabList, method="FE", minObs = 2){
   resRMA = resRMA[keep]
 
   # extract results
-  # Currently exatract simple results
+  # Currently extract simple results
   # other results may be relevant if meta-analysing more than two studies
   resTable = lapply(resRMA, function(x){
     data.frame( logFC     = x$beta,
                 se        = x$se,
                 P.Value   = x$pval,
-                adj.P.Val = p.adjust(x$pval, "fdr"),
[…]

### 2020-09-16 combined_analysis/combine_all_DLPFC.Rmd
Inital commit

@@
 cd /hpc/users/hoffmg01/work/covarr-de
-# rm -rf molec_profiling_cache/
+# rm -rf combined_analysis/combine_all_DLPFC_cache
 ml git pandoc
@@
 difference networks
-Number of DE genes
 
@@
 zenith for Sex with autosomes
+check zenith squared test statistics
 
@@
 # Create list preparing for meta-analysis for SCZ
-df_meta = mclapply(coefTest, function(coef){
+df_meta = lapply(coefTest, function(coef){
   # Run meta-analysis
   run_meta_analysis( tab.de.lst[[coef]], "FE") %>% getGeneSymbol
-}, mc.cores=6)
+})#, mc.cores=6)
 names(df_meta) = coefTest
@@
 # Create list preparing for meta-analysis for SCZ
-df_meta.inter_sex.disease = mclapply(coefTest, function(coef){
+df_meta.inter_sex.disease = lapply(coefTest, function(coef){
   # Run meta-analysis
   run_meta_analysis( tab.de.inter_sex.disease.lst[[coef]], "FE") %>% getGeneSymbol
-}, mc.cores=6) 
+}) 
 names(df_meta.inter_sex.disease) = coefTest  

### 2020-09-16 combined_analysis/combine_all_DLPFC.Rmd
bug fixes

@@
 plot_grid( fig1, fig2, nrow=1)
+
+df = lapply( names(vobj.lst), function(key){
+  data.frame(Cohort = key,  vobj.lst[[key]]$voom.line)
+  })
+df = do.call(rbind, df)
+
+ggplot(df, aes(x, y, color=Cohort)) + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + geom_line(size=1.2) + xlab(bquote(log[2](count + 0.5))) + ylab(expression( sqrt("standard deviation"))) + ggtitle("Compare voom trends") + ylim(0, max(df$y)) + scale_color_manual(name="Status", values=c("#3758ba", "#33e65c")) + guides(color = guide_legend(override.aes = list(size = 1)))
 ```
@@
 
-```{r check.correlation}
+```{r check.correlation, eval=FALSE}
 ngenes = 300

### 2020-09-16 common_functions/common_functions.R

@@ -294,6 +294,7 @@ plot_forrest = function(resList, geneSymbol, plot2="p"){
     fig2 = ggplot(df_forrest, aes(Dataset, -log10(adj.P.Val))) + ylab(bquote(-log[10]~adjusted~P))  
   }else{
     ymax = max(-log10(df_forrest$P.Value))*1.05
+    ymax = min(1e-300, ymax)
     fig2 = ggplot(df_forrest, aes(Dataset, -log10(P.Value))) + ylab(bquote(-log[10]~P)) 
   }
 

### 2020-09-16 common_functions/common_functions.R
bug fixes

@@ -293,8 +293,10 @@ plot_forrest = function(resList, geneSymbol, plot2="p"){
     ymax = max(-log10(df_forrest$adj.P.Val))*1.05
     fig2 = ggplot(df_forrest, aes(Dataset, -log10(adj.P.Val))) + ylab(bquote(-log[10]~adjusted~P))  
   }else{
+
+    df_forrest$P.Value = pmax(1e-300, df_forrest$P.Value)
+
     ymax = max(-log10(df_forrest$P.Value))*1.05
-    ymax = min(1e-300, ymax)
     fig2 = ggplot(df_forrest, aes(Dataset, -log10(P.Value))) + ylab(bquote(-log[10]~P)) 
   }
 

### 2020-09-16 combined_analysis/combine_all_DLPFC.Rmd
bug fixes

@@
 ```{r show.gene_set_test.marginal.Sex.autosome, eval=TRUE}
-show_zenith_results( res_meta_zenith.sex_autosome, "Reported_GenderMale", "GO", FDR=0.05) 
-show_zenith_results( res_meta_zenith.sex_autosome, "Reported_GenderMale", "MSigDB", FDR=0.05)   
+show_zenith_results( res_meta_zenith.sex_autosome, "Reported_GenderMale", "GO", FDR=0.1) 
+show_zenith_results( res_meta_zenith.sex_autosome, "Reported_GenderMale", "MSigDB", FDR=0.1)   
 ```

### 2020-09-16 combined_analysis/combine_all_DLPFC.Rmd
bug fixes

@@
           `# DE` = sum(df_meta.inter_sex.disease[[coef]]$adj.P.Val < 0.05),
+          `# DE @ 10%` = sum(df_meta.inter_sex.disease[[coef]]$adj.P.Val < 0.1),
           check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for SCZ:Sex meta-analysis") %>% kable_styling(full_width=FALSE)

### 2020-09-16 combined_analysis/combine_all_DLPFC.Rmd
bug fixes

@@
 
+```{r corr.matrix}
+form = Dx + Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
+
+```
+
 ### Sex by cell fraction
@@
 
+eval_residuals = function( form ){
   # Compute residuals, but add back in Dx*Reported_Gender
+  # for form = ~ Dx*Reported_Gender
   f = function(fit){
-  residuals(fit) + get_prediction( fit, ~ Dx*Reported_Gender)
+    residuals(fit) + get_prediction( fit, form)
   }
@@
   names(residuals.lst) = names(vobj.lst)
+
+  residuals.lst
+}
+
+resid.add = list( Dx_Sex  = eval_residuals( ~ Dx*Reported_Gender ),
+                  Dx      = eval_residuals( ~ Dx ),
+                  Sex     = eval_residuals( ~ Reported_Gender ))
 ```

### 2020-09-16 combined_analysis/combine_all_DLPFC.Rmd
bug fixes

@@
     toc_float: true
+params:
+  upload: TRUE
 ---
@@
 knit2synapse::createAndKnitToFolderEntity(file = "combined_analysis/combine_all_DLPFC.Rmd",
-                                          parentId ="syn22231629",
+                                          parentId ="syn22416298",
                                           folderName = 'Joint analysis of DLPFC count data',
@@
 
+```{r synapse.parameters, include=FALSE, cache=TRUE}
+parentId = 'syn22416298';
+activityName = 'Joint analysis of DLPFC count data';
+activityDescription = 'Joint analysis of DLPFC count data';
+
+thisFileName <- 'combine_all_DLPFC.Rmd'
+
+# Github link
+thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
+thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('combined_analysis/',thisFileName))
+```
+
+<!---
+
 TODO
@@
 difference networks
-
-<!---
+-------------------
+
 DONE
@@
 ggplot(METADATA, aes(Reported_Gender, ageOfDeath)) + geom_boxplot() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + facet_wrap(~Institution, nrow=1)
-```
-
-```{r corr.matrix}
-form = Dx + Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
-
 ```
@@
 
-```{r exit2, cache=FALSE}
-
-# DIFFERENTIAL CORRLEATION NETWORKS
-# CLUSTER on C_Scz - C_Control
-# Like contrastive PCA
+```{r store, eval=TRUE}
+if( params$upload ){
+  # Code
+  # Make sure these match the folder name and parent Id in the first chunk.
+  CODE <- Folder(name = "Joint analysis of DLPFC count data", parentId = "syn22416298")
+  CODE <- synStore(CODE)
+
+  # Store data 
+  storeData = function( name, description){
+    file = paste0(name, ".RDS")
+    saveRDS( eval(parse(text=name)), file=file)
+
+    nEXP_OBJ = File(file, 
+                    name = description, 
+                    parentId = CODE$properties$id)
+    synStore(nEXP_OBJ, used = ALL_USED_IDs, executed = thisFile)
+  }
+
+  # Standard analysis
+  storeData( "geneCountsMerged", "Merged DLPFC counts for both cohorts")
+  storeData( "METADATA", "Merged metadata used here for both cohorts")
+  storeData( "isexpr", "Genes used in this analysis")
+  storeData( "vobj.lst", "Voom results for each cohort")
+  storeData( "resid.add", "Residuals correcting for covariates, add back: Dx_Sex, Dx, Sex")
+  storeData( "fit_marginal.lst", "dream fit for marginal analysis")
+  storeData( "res_zenith", "zenith results from fit_marginal.lst")
+  storeData( "res_meta_zenith", "Meta-analysis of zenith results from fit_marginal.lst")
+  storeData( "res_meta_zenith.sex_autosome", "Meta-analysis of zenith results from fit_marginal.lst: just autosomal genes")
+  storeData( "df_DE_results", "Differential expressed genes within each cohort")
+  storeData( "df_meta", "Meta-analysis of differential expression across cohorts")
+
+  # Sex by Disease interaction
+  storeData( "fit_inter_sex.disease.lst", "dream fit for interaction analysis")
+  storeData( "res_zenith.inter_sex.disease", "zenith results for interaction analysis")
+  storeData( "res_meta_zenith.inter_sex.disease", "Meta-analysis of zenith results for interaction analysis")
+  storeData( "res_meta_zenith.inter_sex.disease_autosome", "Meta-analysis of zenith results for interaction analysis: autosomal genes only")
+  storeData( "df_DE_results.inter_sex.disease", "Differential expression result for each cohort for interaction analysis")
+  storeData( "df_meta.inter_sex.disease", "Meta-analysis for differential expression for interaction analysis")
+}
 
@@
 
-knitr::knit_exit()
-```
+```
+
+### Source
+[markdown](`r thisFile`)

### 2020-09-16 common_functions/common_functions.R

@@ -591,3 +591,10 @@ zenith_meta_analysis = function(res_zenith, coefTest){
 
 
 
+
+
+
+
+
+
+

### 2020-09-16 combined_analysis/combine_all_DLPFC.Rmd
bug fixes

@@
 ```{r synapse.parameters, include=FALSE, cache=TRUE}
+
+library(githubr)
 parentId = 'syn22416298';
@@
   storeData( "vobj.lst", "Voom results for each cohort")
-  storeData( "resid.add", "Residuals correcting for covariates, add back: Dx_Sex, Dx, Sex")
+  storeData( "resid.add", "Residuals correcting for covariates and add back - (Dx_Sex Dx Sex)")
   storeData( "fit_marginal.lst", "dream fit for marginal analysis")
@@
   storeData( "res_meta_zenith", "Meta-analysis of zenith results from fit_marginal.lst")
-  storeData( "res_meta_zenith.sex_autosome", "Meta-analysis of zenith results from fit_marginal.lst: just autosomal genes")
+  storeData( "res_meta_zenith.sex_autosome", "Meta-analysis of zenith results from fit_marginal.lst - just autosomal genes")
   storeData( "df_DE_results", "Differential expressed genes within each cohort")
@@
   storeData( "res_meta_zenith.inter_sex.disease", "Meta-analysis of zenith results for interaction analysis")
-  storeData( "res_meta_zenith.inter_sex.disease_autosome", "Meta-analysis of zenith results for interaction analysis: autosomal genes only")
+  storeData( "res_meta_zenith.inter_sex.disease_autosome", "Meta-analysis of zenith results for interaction analysis -  autosomal genes only")
   storeData( "df_DE_results.inter_sex.disease", "Differential expression result for each cohort for interaction analysis")

### 2020-09-17 combined_analysis/diff.networks.Rmd
Initial commit

---
title: "CommonMind Consortium: DLPFC"
subtitle: "Analysis of read count data by cohort"
author:
- name: "[Gabriel Hoffman](http://gabrielhoffman.github.io)$^{1}$ and Kelsey Montgomery$^{2}$"
  affiliation: | 
    [1] Icahn School of Medicine at Mount Sinai, New York
    [2] Sage Bionetworks, Seattle, WA
date: "Run on `r Sys.time()`"
documentclass: article
output:
  html_document:
    toc: true
    toc_float: true
---

<!---

cd /hpc/users/hoffmg01/work/covarr-de
# rm -rf combined_analysis/diff.networks_cache
ml git pandoc  udunits proj gdal geos
git pull
R

system("git pull")

rmarkdown::render("combined_analysis/diff.networks.Rmd")

--->

Since we are generating networks across cohorts and sexes, identifiying conserved clusters using standard methods becomes non-trivial.  Instead of defining clusters for each subset and evaluating the difference, we first compute the differences in the correlation networks, then identify a single set a clusters based on this meta-network.  Then use sLED test on difference network, or use boxM_permute on original residuals, but with consensus clusters.

But if concensus models are constructed to be differential by definition, then these clusters could show differential correlation even under the null.  Therefore, use set A as discovery for M/F differential network, and replicated in set B.

```{r setup, include=FALSE}
suppressPackageStartupMessages({
library(synapser)

synLogin()

library(WGCNA)
library(variancePartition)
library(decorate)
library(ggrepel)
library(cowplot) 
library(plyr) 
library(qvalue)
}) 
source("../common_functions/common_functions.R")   
# source("common_functions/common_functions.R")

knitr::opts_chunk$set(
  echo = FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r load.data}
resid.add = readRDS(synGet('syn22416489')$path)
METADATA = readRDS(synGet('syn22416411')$path)
METADATA$Reported_Gender = factor(METADATA$Reported_Gender, c("Male", "Female"))
```

## Difference networks
```{r diff.networks}

# Compute difference networks
##############################

resVersion = "Sex"
variable = 'Reported_Gender'

diff.net = get_difference_network(resid.add[[resVersion]], METADATA, variable)

# Discover clusters in first cohort
#----------------------------------
C.diff.discovery = diff.net[[1]]

hcl = hclust( as.dist(C.diff.discovery) )

dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = 20)

dynamicColors = labels2colors(dynamicMods, colorSeq = colors())
tab = table(dynamicColors)
length(tab)

# Hypothesis testing
#-------------------

df_test = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA)
```

```{r plot.discovery}
# plot discovery and testing
ggplot(df_test, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
```

```{r modules.in.discovery, fig.width=8, fig.height=20}

figList = lapply(df_test$Module[1:5], function(mod){

	plot_module( mod, df_test, METADATA, dynamicColors, C.diff.discovery, resid.add[[resVersion]], variable, "MSSM-Penn-Pitt")  
	})

plot_grid( plotlist = figList, ncol=1)
```

```{r second.cohort}
df_cohort2 = df_test[df_test$Cohort == "NIMH-HBCC",]
df_cohort2$Module = factor(df_cohort2$Module, unique(df_cohort2$Module))
table(df_cohort2$FDR < 0.05)

df = df_cohort2[df_cohort2$P.Value < 0.1,]

ggplot(df, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
```

```{r modules.in.second.cohort, fig.width=8, fig.height=40}

i = which(df_cohort2$FDR < 0.05)[1:5]

figList = lapply(df_cohort2$Module[i], function(mod){

	plot_module( mod, df_test, METADATA, dynamicColors, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC")  
	})

plot_grid( plotlist = figList, ncol=1)
```

<!---

# need to generate correct residuals

GWAS hits, TWAS, SCHEMA
ChrX and ChrY genes with autism and Scz

DE within each sex, on chrX and chrY.
	drop chrY genes for computing FDR?

library(mvtnorm)
library(Rfast)

X = matrnorm(300, 1000)
Y = matrnorm(300, 1000)

C_x = cor(X)
C_y = cor(Y)

df_cor = data.frame( 	a = C_x[lower.tri(C_x)],
						b = C_y[lower.tri(C_y)])

with(df_cor, plot(a,b))

a = calc_sde(id=1,points=df_cor);

Sigma = cov(df_cor)

--->

### 2020-09-17 common_functions/common_functions.R
Update diff networks

@@ -588,9 +588,155 @@ zenith_meta_analysis = function(res_zenith, coefTest){
 
 
 
+get_difference_network = function(resid.lst, METADATA, variable){
+
+  # For each cohort
+  diff.net = lapply( resid.lst, function(resid){
+
+    i = match(colnames(resid), rownames(METADATA))
+    info = METADATA[i,]
+
+    # get correlation matrix for each category
+    C.lst = lapply( levels(info[[variable]]), function(key){
+      j = (info[[variable]] == key)
+      cor( t(resid[,j]) )  
+    })
+    names(C.lst) = levels(info[[variable]])
+
+    # Compute difference in correlation matrices
+    C.lst[[1]] - C.lst[[2]]
+    })
+  names(diff.net) = names(resid.lst)
+
+  diff.net
+}
 
 
 
+
+test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicColors, METADATA){
+
+  df_test = lapply( unique(dynamicColors), function(col){
+
+    # get genes in this cluster
+    geneid = rownames(C.diff.discovery)[which(dynamicColors==col)]
+
+    res = lapply( resid.lst, function(resid){
+
+      # extact expression residuals for genes in this cluster
+      Y = t(resid[geneid,])
+
+      # extract metadata
+      i = match(rownames(Y), rownames(METADATA))
+      info = METADATA[i,]
+
+      # eval statistical hypothesis
+      res = boxM_permute( Y, info[[variable]])
+
+      # C1 = cor(Y[info$Reported_Gender =='Male',])
+      # C2 = cor(Y[info$Reported_Gender =='Female',])
+      # sLED(C1, C2)
+
+      data.frame(Module = col, P.Value = res$p.value, n.genes = length(geneid))
+    })
+    res = do.call(rbind, res)
+    res$Cohort = names(resid.lst)
+    rownames(res) = c()
+    res
+  })
+
+  # Format results data.frame
+  df_test = do.call(rbind, df_test)
+  df_test = df_test[!is.na(df_test$P.Value),]
+  df_test = df_test[order(df_test$P.Value),]
+  df_test$Module = factor(df_test$Module, unique(df_test$Module))
+
+  browser()
+
+  # Compute FDR separately for each cohort
[…]

### 2020-09-17 combined_analysis/combine_all_DLPFC.Rmd
bug fixes

@@
                   Sex     = eval_residuals( ~ Reported_Gender ))
+
+residuals.lst = resid.add[['Dx_Sex']]
 ```

### 2020-09-17 common_functions/common_functions.R
Update diff networks

@@ -651,8 +651,6 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
   df_test = df_test[order(df_test$P.Value),]
   df_test$Module = factor(df_test$Module, unique(df_test$Module))
 
-  browser()
-
   # Compute FDR separately for each cohort
   df_test$FDR = rep(NA, nrow(df_test))
   i = which(df_test$Cohort == 'MSSM-Penn-Pitt')

### 2020-09-17 common_functions/common_functions.R
Update diff networks

@@ -651,6 +651,8 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
   df_test = df_test[order(df_test$P.Value),]
   df_test$Module = factor(df_test$Module, unique(df_test$Module))
 
+  browser()
+  
   # Compute FDR separately for each cohort
   df_test$FDR = rep(NA, nrow(df_test))
   i = which(df_test$Cohort == 'MSSM-Penn-Pitt')

### 2020-09-17 combined_analysis/diff.networks.Rmd
Update diff networks

@@
   tidy = FALSE,
-  cache = TRUE)
+  cache = TRUE,
+  cache.lazy = FALSE)
 ```
@@
 length(tab)
+```
+
+```{r testing}
 
@@
 
+Can we predict sex given autosomal genes?
+
+PMID: 26892004
+https://science.sciencemag.org/content/369/6509/eaba3066
+https://www.nature.com/articles/s41576-018-0083-1
+
+test: how much sex differences are given by age, cell type and institution.
+
 library(mvtnorm)

### 2020-09-17 common_functions/common_functions.R

@@ -599,7 +599,7 @@ get_difference_network = function(resid.lst, METADATA, variable){
     # get correlation matrix for each category
     C.lst = lapply( levels(info[[variable]]), function(key){
       j = (info[[variable]] == key)
-      cor( t(resid[,j]) )  
+      cor( t(resid[1:1000,j]) )  
     })
     names(C.lst) = levels(info[[variable]])
 
@@ -652,7 +652,7 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
   df_test$Module = factor(df_test$Module, unique(df_test$Module))
 
   browser()
-  
+
   # Compute FDR separately for each cohort
   df_test$FDR = rep(NA, nrow(df_test))
   i = which(df_test$Cohort == 'MSSM-Penn-Pitt')

### 2020-09-17 common_functions/common_functions.R
Update diff networks

@@ -599,7 +599,7 @@ get_difference_network = function(resid.lst, METADATA, variable){
     # get correlation matrix for each category
     C.lst = lapply( levels(info[[variable]]), function(key){
       j = (info[[variable]] == key)
-      cor( t(resid[1:1000,j]) )  
+      cor( t(resid[1:10000,j]) )  
     })
     names(C.lst) = levels(info[[variable]])
 

### 2020-09-17 common_functions/common_functions.R
Update diff networks

@@ -656,7 +656,7 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
   # Compute FDR separately for each cohort
   df_test$FDR = rep(NA, nrow(df_test))
   i = which(df_test$Cohort == 'MSSM-Penn-Pitt')
-  df_test$FDR[i] = qvalue(df_test$P.Value[i], lambda = seq(0,0.80,0.05))$qvalues
+  df_test$FDR[i] = qvalue(c(df_test$P.Value[i],1))$qvalues
 
   i = which(df_test$Cohort == 'NIMH-HBCC')
   df_test$FDR[i] = qvalue(df_test$P.Value[i])$qvalues

### 2020-09-17 common_functions/common_functions.R
Update diff networks

@@ -656,7 +656,7 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
   # Compute FDR separately for each cohort
   df_test$FDR = rep(NA, nrow(df_test))
   i = which(df_test$Cohort == 'MSSM-Penn-Pitt')
-  df_test$FDR[i] = qvalue(c(df_test$P.Value[i],1))$qvalues
+  df_test$FDR[i] = NA#qvalue(c(df_test$P.Value[i],1))$qvalues
 
   i = which(df_test$Cohort == 'NIMH-HBCC')
   df_test$FDR[i] = qvalue(df_test$P.Value[i])$qvalues

### 2020-09-17 common_functions/common_functions.R
Update diff networks

@@ -599,7 +599,7 @@ get_difference_network = function(resid.lst, METADATA, variable){
     # get correlation matrix for each category
     C.lst = lapply( levels(info[[variable]]), function(key){
       j = (info[[variable]] == key)
-      cor( t(resid[1:10000,j]) )  
+      cor( t(resid[,j]) )  
     })
     names(C.lst) = levels(info[[variable]])
 
@@ -656,7 +656,7 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
   # Compute FDR separately for each cohort
   df_test$FDR = rep(NA, nrow(df_test))
   i = which(df_test$Cohort == 'MSSM-Penn-Pitt')
-  df_test$FDR[i] = NA#qvalue(c(df_test$P.Value[i],1))$qvalues
+  df_test$FDR[i] = NA #qvalue(c(df_test$P.Value[i],1))$qvalues
 
   i = which(df_test$Cohort == 'NIMH-HBCC')
   df_test$FDR[i] = qvalue(df_test$P.Value[i])$qvalues

### 2020-09-17 common_functions/common_functions.R
Update diff networks

@@ -5,6 +5,7 @@ library(ensembldb)
 library(MASS)
 library(ggplot2)
 library(ggbio)
+library(Rfast)
 library(viridis)
 library(data.table)
 })
@@ -599,7 +600,7 @@ get_difference_network = function(resid.lst, METADATA, variable){
     # get correlation matrix for each category
     C.lst = lapply( levels(info[[variable]]), function(key){
       j = (info[[variable]] == key)
-      cor( t(resid[,j]) )  
+      cora( t(resid[,j]) )  
     })
     names(C.lst) = levels(info[[variable]])
 

### 2020-09-17 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 
+rm(diff.net) # saving this to disk is slow, so just delete it
+
 hcl = hclust( as.dist(C.diff.discovery) )
@@
 # plot discovery and testing
-ggplot(df_test, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
+# ggplot(df_test, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
 ```

### 2020-09-17 common_functions/common_functions.R
Update diff networks

@@ -688,7 +688,7 @@ plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.disc
 
   FDR = df_test[with(df_test, Cohort==key & Module == clusterID),'FDR']
 
-  main = paste(key, col, 'FDR =', format(FDR, digits=3))
+  main = paste(key, clusterID, 'FDR =', format(FDR, digits=3))
 
   # Evaluate correlation matrix for each category
   lvl = levels(info[[variable]])

### 2020-09-17 common_functions/common_functions.R
Update diff networks

@@ -668,6 +668,8 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
 # Plots for each module
 plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.discovery, resid.lst, variable, key){
 
+  clusterID = as.character(clusterID)
+
   # get genes in this clusters
   geneid = rownames(C.diff.discovery)[which(dynamicColors==clusterID)]
 

### 2020-09-17 combined_analysis/diff.networks.Rmd
Update diff networks

--- 2020-09-17 combined_analysis/diff.networks.Rmd
@@
 
-```{r modules.in.discovery, fig.width=8, fig.height=20}
+```{r modules.in.discovery, fig.width=10, fig.height=20}
 
@@
 
-```{r second.cohort}
+```{r second.cohort, fig.width=6, fig.height=12}
 df_cohort2 = df_test[df_test$Cohort == "NIMH-HBCC",]
@@
 
-i = which(df_cohort2$FDR < 0.05)[1:5]
+i = which(df_cohort2$FDR < 0.05)
 

### 2020-09-17 common_functions/common_functions.R

@@ -655,9 +655,10 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
   # Compute FDR separately for each cohort
   df_test$FDR = rep(NA, nrow(df_test))
   i = which(df_test$Cohort == 'MSSM-Penn-Pitt')
-  df_test$FDR[i] = NA #qvalue(c(df_test$P.Value[i],1))$qvalues
+  df_test$FDR[i] = p.adjust(df_test$P.Value[i], "fdr")
 
   i = which(df_test$Cohort == 'NIMH-HBCC')
+  # df_test$FDR[i] = p.adjust(df_test$P.Value[i], "fdr")
   df_test$FDR[i] = qvalue(df_test$P.Value[i])$qvalues
 
   df_test

### 2020-09-17 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 
+# Sex network
 ## Difference networks
@@
 
-```{r modules.in.second.cohort, fig.width=8, fig.height=40}
+```{r modules.in.second.cohort, fig.width=8, fig.height=100}
+
+i = which(df_cohort2$FDR < 0.05)
+
+figList = lapply(df_cohort2$Module[i], function(mod){
+
+	plot_module( mod, df_test, METADATA, dynamicColors, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC")  
+	})
+
+plot_grid( plotlist = figList, ncol=1)
+```
+
+# Disease network
+## Difference networks
+```{r diff.networks.Dx}
+
+# Compute difference networks
+##############################
+
+resVersion = "Dx"
+variable = 'Dx'
+
+diff.net = get_difference_network(resid.add[[resVersion]], METADATA, variable)
+
+# Discover clusters in first cohort
+#----------------------------------
+C.diff.discovery = diff.net[[1]]
+
+rm(diff.net) # saving this to disk is slow, so just delete it
+
+hcl = hclust( as.dist(C.diff.discovery) )
+
+dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = 20)
+
+dynamicColors = labels2colors(dynamicMods, colorSeq = colors())
+tab = table(dynamicColors)
+length(tab)
+```
+
+```{r testing.Dx}
+
+# Hypothesis testing
+#-------------------
+
+df_test = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA)  
+```
+
+```{r plot.discovery.Dx}
+# plot discovery and testing
+ggplot(df_test, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
+```
+
+```{r modules.in.discovery.Dx, fig.width=10, fig.height=20}
+
+figList = lapply(df_test$Module[1:5], function(mod){
+
+	plot_module( mod, df_test, METADATA, dynamicColors, C.diff.discovery, resid.add[[resVersion]], variable, "MSSM-Penn-Pitt")  
+	})
+
+plot_grid( plotlist = figList, ncol=1)
+```
+
+```{r second.cohort.Dx, fig.width=6, fig.height=12}
+df_cohort2 = df_test[df_test$Cohort == "NIMH-HBCC",]
+df_cohort2$Module = factor(df_cohort2$Module, unique(df_cohort2$Module))
+table(df_cohort2$FDR < 0.05)
+
+df = df_cohort2[df_cohort2$P.Value < 0.1,]
+
+ggplot(df, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
+```
+
+```{r modules.in.second.cohort.Dx, fig.width=8, fig.height=40}
 

### 2020-09-17 combined_analysis/diff.networks.Rmd
Update diff networks

--- 2020-09-17 combined_analysis/diff.networks.Rmd
@@
 
-```{r modules.in.second.cohort, fig.width=8, fig.height=40}
+```{r modules.in.second.cohort, fig.width=8, fig.height=100}
 
 i = which(df_cohort2$FDR < 0.05)
+
+figList = lapply(df_cohort2$Module[i], function(mod){
+
+	plot_module( mod, df_test, METADATA, dynamicColors, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC")  
+	})
+
+plot_grid( plotlist = figList, ncol=1)
+```
+
+# Disease network
+## Difference networks
+```{r diff.networks.Dx}
+
+# Compute difference networks
+##############################
+
+resVersion = "Dx"
+variable = 'Dx'
+
+diff.net = get_difference_network(resid.add[[resVersion]], METADATA, variable)
+
+# Discover clusters in first cohort
+#----------------------------------
+C.diff.discovery = diff.net[[1]]
+
+rm(diff.net) # saving this to disk is slow, so just delete it
+
+hcl = hclust( as.dist(C.diff.discovery) )
+
+dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = 20)
+
+dynamicColors = labels2colors(dynamicMods, colorSeq = colors())
+tab = table(dynamicColors)
+length(tab)
+```
+
+```{r testing.Dx}
+
+# Hypothesis testing
+#-------------------
+
+df_test = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA)  
+```
+
+```{r plot.discovery.Dx}
+# plot discovery and testing.  
+ggplot(df_test, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
+```
+
+```{r modules.in.discovery.Dx, fig.width=10, fig.height=20}
+
+figList = lapply(df_test$Module[1:5], function(mod){
+
+	plot_module( mod, df_test, METADATA, dynamicColors, C.diff.discovery, resid.add[[resVersion]], variable, "MSSM-Penn-Pitt")  
+	})
+
+plot_grid( plotlist = figList, ncol=1)
+```
+
+```{r second.cohort.Dx, fig.width=6, fig.height=12}
+df_cohort2 = df_test[df_test$Cohort == "NIMH-HBCC",]
+df_cohort2$Module = factor(df_cohort2$Module, unique(df_cohort2$Module))
+table(df_cohort2$FDR < 0.05).
+
+df = df_cohort2[df_cohort2$P.Value < 0.1,]
+
+ggplot(df, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
+```
+
+```{r modules.in.second.cohort.Dx, fig.width=8, fig.height=40}
+
+i = which(df_cohort2$FDR < 0.05)[1:20]
 

### 2020-09-17 common_functions/common_functions.R

@@ -658,8 +658,8 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
   df_test$FDR[i] = p.adjust(df_test$P.Value[i], "fdr")
 
   i = which(df_test$Cohort == 'NIMH-HBCC')
-  # df_test$FDR[i] = p.adjust(df_test$P.Value[i], "fdr")
-  df_test$FDR[i] = qvalue(df_test$P.Value[i])$qvalues
+  df_test$FDR[i] = p.adjust(df_test$P.Value[i], "fdr")
+  # df_test$FDR[i] = qvalue(df_test$P.Value[i])$qvalues
 
   df_test
 }

### 2020-09-17 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 df_cohort2$Module = factor(df_cohort2$Module, unique(df_cohort2$Module))
-table(df_cohort2$FDR < 0.05).
+table(df_cohort2$FDR < 0.05)
 
@@
 
+Apparent differences in cell type composition can be explained by age.
+	After accounting for age, can cell type predict sex?
+
 library(mvtnorm)

### 2020-09-17 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 tab = table(dynamicColors)
-length(tab)
+hist(tab)
 ```
@@
 
-```{r modules.in.second.cohort.Dx, fig.width=8, fig.height=40}
+```{r modules.in.second.cohort.Dx, fig.width=8, fig.height=100}
 

### 2020-09-17 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 
-```{r modules.in.second.cohort.Dx, fig.width=8, fig.height=100}
+```{r modules.in.second.cohort.Dx, fig.width=15, fig.height=100}
 
@@
 
+X = matrnorm(300, 1000)
+
+C_x = cor(X)
+
+C = C_x[1:5, 1:5]
+colnames(C) = paste0('gene_', 1:5)
+rownames(C) = colnames(C)
+
+library(ggraph)
+library(ggplot2)
+library(tidygraph)
+library(Rfast)
+
+load("~/Downloads/C.test.RDATA")
+
+C = C1 - C2
+
+plot_corr_network( C1-C2 )
+
+rstat_nodes <- data.frame(name = c("H
+
+	adley", "David", "Romain", "Julia"))
+rstat_edges <- data.frame(from = c(1, 1, 1, 2, 3, 3, 4, 4, 4),
+                         to = c(2, 3, 4, 1, 1, 2, 1, 2, 3))
+rstat_edges$weight = 1:9
+
+rstat_nodes <- data.frame(name = c("Hadley", "David", "Romain", "Julia"))
+rstat_edges <- data.frame(from = c(1, 1, 1, 2, 3, 3, 4, 4, 4),
+                         to = c(2, 3, 4, 1, 1, 2, 1, 2, 3))
+rstat_edges$weight = 1:9
+
+net = tbl_graph(nodes = rstat_nodes, edges = rstat_edges, directed=FALSE)
+
+ggraph(net, layout = "auto") + 
+  geom_node_point(size=10, color="blue") +
+  geom_edge_link(aes(width = weight), alpha = 0.8) + 
+  scale_edge_width(range = c(0.2, 2)) +
+  geom_node_text(aes(label = name), repel = TRUE) +
+  labs(edge_width = "Correlation") +
+  theme_graph()
+
+net = ggnetwork(C)
+
+set.edge.attribute
+set.edge.value
+
+ggplot(net, aes(x = x, y = y, xend = xend, yend = yend)) + theme_blank() + geom_nodes(aes(x,y)) + geom_edges()
+
+ggplot(
+ggnetwork(emon[[1]], layout = "kamadakawai", arrow.gap = 0.025),
+aes(x, y, xend = xend, yend = yend) )+
+geom_edges(aes(color = Frequency, size=Frequency^2),
+arrow = arrow(length = unit(10, "pt"), type = "closed")
+)+
+geom_nodes(aes(size = Formalization)) + scale_color_gradient(low = "grey50", high = "tomato") + scale_size_area(breaks = 1:3) +
+theme_blank()
+
+net = network(C)
+
+get.edges(net, "gene_1")
+
+rstat_nodes <- data.frame(name = c("Hadley", "David", "Romain", "Julia"))
+rstat_edges <- data.frame(from = c(1, 1, 1, 2, 3, 3, 4, 4, 4),
+                         to = c(2, 3, 4, 1, 1, 2, 1, 2, 3))
+rstat_edges$weight = 1:9
+
+net = tbl_graph(nodes = rstat_nodes, edges = rstat_edges, directed=FALSE)
+
+ggraph(net, layout = "auto") + 
+  geom_node_point(size=10, color="blue") +
+  geom_edge_link(aes(width = weight), alpha = 0.8) + 
+  scale_edge_width(range = c(0.2, 2)) +
+  geom_node_text(aes(label = name), repel = TRUE) +
+  labs(edge_width = "Correlation") +
+  theme_graph()
+
 --->

### 2020-09-17 common_functions/common_functions.R

@@ -8,6 +8,10 @@ library(ggbio)
 library(Rfast)
 library(viridis)
 library(data.table)
+library(ggraph)
+library(ggplot2)
+library(tidygraph)
+library(Rfast)
 })
 
 # For ENSEMBL id ENSG00000279457.4, return ENSG00000279457
@@ -737,12 +741,50 @@ plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.disc
         theme(aspect.ratio = 1, plot.title = element_text(hjust = 0.5), 
             legend.position = "bottom", panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.text.x=element_text(angle=45, hjust=1)) + ggtitle(main) + xlab(lvl[2]) + ylab(lvl[1])
 
-    plot_grid( fig1, fig2 )    
+  # plot network   
+  plot_corr_network( C1 - C2) + ggtitle(main)
+
+  plot_grid( fig1, fig2, fig2, nrow=1 )    
 }
 
 
 
 
+
+plot_corr_network = function( C, zcutoff = 1.3, seed=1){
+
+  df_net = data.frame(t(combn(colnames(C), 2)))
+  colnames(df_net) = c('name.from','name.to')
+
+  df_net$weight = apply(df_net, 1, function(x){
+    C[x['name.from'], x['name.to']]
+  })
+  df_net$z = scale(df_net$weight)
+  df_net = df_net[abs(df_net$z) >= zcutoff,]
+
+  node_names = unique(c(df_net$name.from, df_net$name.to))
+
+  C_sub = C[node_names,node_names]
+
+  df_net$from = match(df_net$name.from, node_names)
+  df_net$to = match(df_net$name.to, node_names)
+
+  net = tbl_graph(nodes = data.frame(name=node_names), edges = df_net[,c('from','to','weight')], directed=FALSE)
+       
+  # igraph_layouts <- c('star', 'circle', 'gem', 'dh', 'graphopt', 'grid', 'mds', 
+  #                     'randomly', 'fr', 'kk', 'drl', 'lgl')
+
+  set.seed(seed)
+  ggraph(net, layout = "igraph", algorithm='graphopt') + 
+    geom_edge_link(aes(width = abs(weight), color=weight), alpha = 1) + 
+    scale_edge_width(name='|Corr Diff|',range = c(0.2, 2)) +
+    geom_node_point(size=10, color="grey80") +
[…]

### 2020-09-17 common_functions/common_functions.R
Update diff networks

@@ -744,7 +744,7 @@ plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.disc
   # plot network   
   plot_corr_network( C1 - C2) + ggtitle(main)
 
-  plot_grid( fig1, fig2, fig2, nrow=1 )    
+  plot_grid( fig1, fig2, fig3, nrow=1 )    
 }
 
 
@@ -766,6 +766,8 @@ plot_corr_network = function( C, zcutoff = 1.3, seed=1){
 
   C_sub = C[node_names,node_names]
 
+  message('Nodes: ', length(node_names))
+
   df_net$from = match(df_net$name.from, node_names)
   df_net$to = match(df_net$name.to, node_names)
 
@@ -785,6 +787,26 @@ plot_corr_network = function( C, zcutoff = 1.3, seed=1){
     theme_graph() + 
     theme(aspect.ratio=1,plot.title = element_text(hjust = 0.5))
 }
+
+# df_net = data.table(df_net)
+
+# df_a = df_net[,data.frame(minAbsZ=max(abs(z))), by="name.from"]
+# df_b = df_net[,data.frame(minAbsZ=max(abs(z))), by="name.to"]
+
+# df = merge(df_a, df_b, by.x="name.from", by.y="name.to")
+
+# df2 = df[,data.frame(z=max(c(minAbsZ.x, minAbsZ.y))),by="name.from"]
+
+# min(df2$z)
+
+# plot_corr_network( C1-C2, zcutoff=1.568965)
+
+
+
+
+
+
+
 
 
 

### 2020-09-17 common_functions/common_functions.R
Update diff networks

@@ -733,7 +733,7 @@ plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.disc
     colsFun = colorRampPalette( c("blue", "white","red"))
     color = colsFun(1000)
 
-  fig2 = ggplot(df, aes(Var1, Var2)) + geom_tile(aes(color = value, 
+  fig3 = ggplot(df, aes(Var1, Var2)) + geom_tile(aes(color = value, 
         fill = value)) + scale_color_gradientn(name = "Correlation", 
         colours = color, limits = c(-1, 1), na.value = "grey") + 
         scale_fill_gradientn(name = "Correlation", colours = color, 

### 2020-09-17 common_functions/common_functions.R
Update diff networks

@@ -730,10 +730,10 @@ plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.disc
   diag(C) = NA
   df = reshape2::melt(C)
 
-    colsFun = colorRampPalette( c("blue", "white","red"))
-    color = colsFun(1000)
+  colsFun = colorRampPalette( c("blue", "white","red"))
+  color = colsFun(1000)
 
-  fig3 = ggplot(df, aes(Var1, Var2)) + geom_tile(aes(color = value, 
+  fig2 = ggplot(df, aes(Var1, Var2)) + geom_tile(aes(color = value, 
         fill = value)) + scale_color_gradientn(name = "Correlation", 
         colours = color, limits = c(-1, 1), na.value = "grey") + 
         scale_fill_gradientn(name = "Correlation", colours = color, 
@@ -742,7 +742,7 @@ plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.disc
             legend.position = "bottom", panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.text.x=element_text(angle=45, hjust=1)) + ggtitle(main) + xlab(lvl[2]) + ylab(lvl[1])
 
   # plot network   
-  plot_corr_network( C1 - C2) + ggtitle(main)
+  fig3 = plot_corr_network( C1 - C2) + ggtitle(main)
 
   plot_grid( fig1, fig2, fig3, nrow=1 )    
 }
@@ -766,8 +766,6 @@ plot_corr_network = function( C, zcutoff = 1.3, seed=1){
 
   C_sub = C[node_names,node_names]
 
-  message('Nodes: ', length(node_names))
-
   df_net$from = match(df_net$name.from, node_names)
   df_net$to = match(df_net$name.to, node_names)
 
@@ -776,7 +774,10 @@ plot_corr_network = function( C, zcutoff = 1.3, seed=1){
   # igraph_layouts <- c('star', 'circle', 'gem', 'dh', 'graphopt', 'grid', 'mds', 
   #                     'randomly', 'fr', 'kk', 'drl', 'lgl')
 
+  # set seed for reproducability
   set.seed(seed)
+
+  # plot
   ggraph(net, layout = "igraph", algorithm='graphopt') + 
     geom_edge_link(aes(width = abs(weight), color=weight), alpha = 1) + 
[…]

### 2020-09-17 common_functions/common_functions.R
Update diff networks

@@ -671,7 +671,7 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
 
 
 # Plots for each module
-plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.discovery, resid.lst, variable, key){
+plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.discovery, resid.lst, variable, key, base_size=11){
 
   clusterID = as.character(clusterID)
 
@@ -721,12 +721,11 @@ plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.disc
 
   lim = range(c(df2$cor1, df2$cor2))
 
-  fig1 = ggplot(df2, aes(cor1, cor2, label=pair)) + geom_point(aes(color=ifelse(pair=='', "grey", "red"))) + theme_bw() + theme(aspect.ratio = 1, legend.position='none', plot.title = element_text(hjust = 0.5))  + geom_abline(color="grey", linetype="dashed") + geom_text_repel(direction='x', nudge_x=1,    hjust=1, segment.size = 0.2, box.padding=.2) + scale_color_manual(values=c("grey50", "red")) + xlim(lim) + ylim(lim) + ggtitle(main) + xlab(paste('Correlation in', lvl[1])) + ylab(paste('Correlation in', lvl[2]))
+  fig1 = ggplot(df2, aes(cor1, cor2, label=pair)) + geom_point(aes(color=ifelse(pair=='', "grey", "red"))) + theme_bw(base_size) + theme(aspect.ratio = 1, legend.position='none', plot.title = element_text(hjust = 0.5))  + geom_abline(color="grey", linetype="dashed") + geom_text_repel(direction='x', nudge_x=1, hjust=1, segment.size = 0.2, box.padding=.2, size=base_size/2) + scale_color_manual(values=c("grey50", "red")) + xlim(lim) + ylim(lim) + ggtitle(main) + xlab(paste('Correlation in', lvl[1])) + ylab(paste('Correlation in', lvl[2]))
 
-
-  # upper is C1 and lower is C2
-  C = C1
-  C[lower.tri(C)] = C2[lower.tri(C2)]
+  # upper is C2 and lower is C1, since C1 is baseline, is lvl[1]
+  C = C2
+  C[lower.tri(C)] = C1[lower.tri(C1)]
   diag(C) = NA
   df = reshape2::melt(C)
 
[…]

### 2020-09-17 common_functions/common_functions.R
Update diff networks

@@ -748,7 +748,7 @@ plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.disc
   plot_grid( fig_merge, fig3, nrow=2, rel_heights=c(1,1.5) )    
 }
 
-plot_module( mod, df_test, METADATA, dynamicColors, C.diff.discovery, resid.add[[resVersion]], variable, "MSSM-Penn-Pitt", 5) 
+# plot_module( mod, df_test, METADATA, dynamicColors, C.diff.discovery, resid.add[[resVersion]], variable, "MSSM-Penn-Pitt", 5) 
 
 
 

### 2020-09-17 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 
-```{r modules.in.discovery, fig.width=10, fig.height=20}
+```{r modules.in.discovery, fig.width=10, fig.height=50}
 
@@
 
-	plot_module( mod, df_test, METADATA, dynamicColors, C.diff.discovery, resid.add[[resVersion]], variable, "MSSM-Penn-Pitt")  
+	plot_module( mod, df_test, METADATA, dynamicColors, C.diff.discovery, resid.add[[resVersion]], variable, "MSSM-Penn-Pitt", 5)   
 	})
@@
 
-```{r modules.in.second.cohort, fig.width=8, fig.height=100}
+```{r modules.in.second.cohort, fig.width=8, fig.height=25}
 
@@
 
-```{r modules.in.discovery.Dx, fig.width=10, fig.height=20}
+```{r modules.in.discovery.Dx, fig.width=10, fig.height=50}
 
@@
 
-X = matrnorm(300, 1000)
+X = matrnorm(1000, 30)
 
@@
 C = C_x[1:5, 1:5]
-colnames(C) = paste0('gene_', 1:5)
+colnames(C) = paste0('gene_', 1:nrow(C))
 rownames(C) = colnames(C)

### 2020-09-17 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 
+## Discovery cohort
 ```{r plot.discovery}
@@
 
+## Replication cohort
 ```{r second.cohort, fig.width=6, fig.height=12}
@@
 
+## Discovery cohort
 ```{r plot.discovery.Dx}
@@
 
+## Replication cohort
 ```{r second.cohort.Dx, fig.width=6, fig.height=12}
@@
 ```
-
 <!---

### 2020-09-17 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 
-```{r modules.in.discovery, fig.width=10, fig.height=50}
+```{r modules.in.discovery, fig.width=8, fig.height=50}
 
@@
 ## Replication cohort
-```{r second.cohort, fig.width=6, fig.height=12}
+```{r second.cohort, fig.width=8, fig.height=12}
 df_cohort2 = df_test[df_test$Cohort == "NIMH-HBCC",]
@@
 
-```{r modules.in.second.cohort.Dx, fig.width=15, fig.height=100}
+```{r modules.in.second.cohort.Dx, fig.width=8, fig.height=100}
 
@@
 ```
+
 <!---

### 2020-09-17 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 tab = table(dynamicColors)
-hist(tab)
+hist(tab, xlab="Cluster size")
 ```
@@
 
-```{r modules.in.discovery, fig.width=8, fig.height=50}
+```{r modules.in.discovery, fig.width=8, fig.height=25}
 

### 2020-09-17 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 
-```{r modules.in.discovery, fig.width=8, fig.height=25}
+```{r modules.in.discovery, fig.width=8, fig.height=40}
 
@@
 
-```{r modules.in.second.cohort.Dx, fig.width=8, fig.height=100}
+```{r modules.in.second.cohort.Dx, fig.width=8, fig.height=500}
 
@@
 
-X = matrnorm(300, 1000)
-Y = matrnorm(300, 1000)
+X = matrnorm(300, 20)
+Y = matrnorm(300, 20)
+
+library(BiocParallel)
+res = decorate:::.sLED(X=X, Y=Y, npermute=100000, verbose=TRUE, BPPARAM = SerialParam())
+
+res$pVal
 

### 2020-09-17 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 library(qvalue)  
+library(BiocParallel)  
 }) 
@@
 
-df_test = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA)  
+df_test = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, BPPARAM=SerialParam())  
 ```
@@
 
-```{r modules.in.second.cohort.Dx, fig.width=8, fig.height=500}
+```{r modules.in.second.cohort.Dx, fig.width=8, fig.height=100}
 
-i = which(df_cohort2$FDR < 0.05)[1:20]
+i = which(df_cohort2$FDR < 0.05)[1:10]
 

### 2020-09-17 common_functions/common_functions.R

@@ -618,8 +618,31 @@ get_difference_network = function(resid.lst, METADATA, variable){
 
 
 
+sLED_adapt = function(Y1, Y2, npermute=c(1000,1e6), BPPARAM=SerialParam()){
 
-test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicColors, METADATA){
+  # perform permutations until p-values is precise enough
+  # if not precise enough
+  a = log10(npermute[1])
+  b = log10(npermute[2])
+  permArray = 10^seq(a,b, length.out=b-a +1)
+
+  for( nperm in round(permArray) ){
+      # compare correlation structure with sLED
+      res = decorate:::.sLED(X=Y1, Y=Y2, npermute=nperm, BPPARAM=ifelse(nperm >= 10000,BPPARAM, SerialParam()))
+
+      if( res$pVal * nperm > 10){
+        break
+      }
+  }
+
+  # stats is always positive, so multiply by -1 if negative
+  res$stats = (-1)^(res$sign=='neg') * res$stats
+
+  res
+}
+
+
+test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicColors, METADATA, BPPARAM=SerialParam()){
 
   df_test = lapply( unique(dynamicColors), function(col){
 
@@ -638,11 +661,17 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
       # eval statistical hypothesis
       res = boxM_permute( Y, info[[variable]])
 
-      # C1 = cor(Y[info$Reported_Gender =='Male',])
-      # C2 = cor(Y[info$Reported_Gender =='Female',])
-      # sLED(C1, C2)
+      # sLED
+      lvl = levels(info[[variable]])
+      Y1 = Y[info[[variable]] == lvl[1],]
+      Y2 = Y[info[[variable]] == lvl[2],]
 
-      data.frame(Module = col, P.Value = res$p.value, n.genes = length(geneid))
+      res_sLED = sLED_adapt(Y1, Y2, npermute=c(1000,1e6), BPPARAM=BPPARAM)
+
+      data.frame( Module        = col, 
+                  P.Value       = res$p.value, 
+                  P.Value.sLED  = res_sLED$pVal,
+                  n.genes       = length(geneid))
     })
     res = do.call(rbind, res)
     res$Cohort = names(resid.lst)

### 2020-09-18 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 ```{r testing}
-
 # Hypothesis testing  
-#-------------------
-
-df_test = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, BPPARAM=SerialParam())  
+df_test = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, BPPARAM=SnowParam(18)) 
 ```
@@
 ```{r testing.Dx}
-
 # Hypothesis testing     
-#-------------------
-
 df_test = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA)  

### 2020-09-18 common_functions/common_functions.R

@@ -627,8 +627,13 @@ sLED_adapt = function(Y1, Y2, npermute=c(1000,1e6), BPPARAM=SerialParam()){
   permArray = 10^seq(a,b, length.out=b-a +1)
 
   for( nperm in round(permArray) ){
+
+      bp = BPPARAM
+
+      if(nperm <= 10000) bp = SerialParam()
+
       # compare correlation structure with sLED
-      res = decorate:::.sLED(X=Y1, Y=Y2, npermute=nperm, BPPARAM=ifelse(nperm >= 10000,BPPARAM, SerialParam()))
+      res = decorate:::.sLED(X=Y1, Y=Y2, npermute=nperm, BPPARAM=bp)
 
       if( res$pVal * nperm > 10){
         break
@@ -649,10 +654,12 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
     # get genes in this cluster
     geneid = rownames(C.diff.discovery)[which(dynamicColors==col)]
 
-    res = lapply( resid.lst, function(resid){
+    res = lapply( names(resid.lst), function(key){
+
+      message(key, ' ', col)
 
       # extact expression residuals for genes in this cluster
-      Y = t(resid[geneid,])
+      Y = t(resid.lst[[key]][geneid,])
 
       # extract metadata
       i = match(rownames(Y), rownames(METADATA))
@@ -661,13 +668,16 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
       # eval statistical hypothesis
       res = boxM_permute( Y, info[[variable]])
 
-      # sLED
-      lvl = levels(info[[variable]])
-      Y1 = Y[info[[variable]] == lvl[1],]
-      Y2 = Y[info[[variable]] == lvl[2],]
+      if( key == names(resid.lst)[2] ){
+        # sLED
+        lvl = levels(info[[variable]])
+        Y1 = Y[info[[variable]] == lvl[1],]
+        Y2 = Y[info[[variable]] == lvl[2],]
 
-      res_sLED = sLED_adapt(Y1, Y2, npermute=c(1000,1e6), BPPARAM=BPPARAM)
-
+        res_sLED = sLED_adapt(Y1, Y2, npermute=c(1000,1e5), BPPARAM=BPPARAM)
+      }else{
+        res_sLED = list(pVal=NA)
+      }
       data.frame( Module        = col, 
                   P.Value       = res$p.value, 
                   P.Value.sLED  = res_sLED$pVal,

### 2020-09-18 common_functions/common_functions.R
Update diff networks

@@ -649,14 +649,16 @@ sLED_adapt = function(Y1, Y2, npermute=c(1000,1e6), BPPARAM=SerialParam()){
 
 test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicColors, METADATA, BPPARAM=SerialParam()){
 
-  df_test = lapply( unique(dynamicColors), function(col){
+  col.array = unique(dynamicColors)
+
+  df_test = mclapply( col.array, function(col){
 
     # get genes in this cluster
     geneid = rownames(C.diff.discovery)[which(dynamicColors==col)]
 
     res = lapply( names(resid.lst), function(key){
 
-      message(key, ' ', col)
+      message(key, ' ', col, ' ', match(col, col.array), '/', length(col.array))
 
       # extact expression residuals for genes in this cluster
       Y = t(resid.lst[[key]][geneid,])
@@ -687,7 +689,7 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
     res$Cohort = names(resid.lst)
     rownames(res) = c()
     res
-  })
+  }, mc.cores=4)
 
   # Format results data.frame
   df_test = do.call(rbind, df_test)

### 2020-09-18 common_functions/common_functions.R
Update diff networks

@@ -633,7 +633,7 @@ sLED_adapt = function(Y1, Y2, npermute=c(1000,1e6), BPPARAM=SerialParam()){
       if(nperm <= 10000) bp = SerialParam()
 
       # compare correlation structure with sLED
-      res = decorate:::.sLED(X=Y1, Y=Y2, npermute=nperm, BPPARAM=bp)
+      res = decorate:::.sLED(X=Y1, Y=Y2, npermute=nperm, BPPARAM=bp, rho=1, sumabs.seq=.2)
 
       if( res$pVal * nperm > 10){
         break

### 2020-09-18 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 # Hypothesis testing      
-df_test = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, BPPARAM=SnowParam(18)) 
+df_test = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA)#, BPPARAM=SnowParam(18)) 
 ```

### 2020-09-18 common_functions/common_functions.R

@@ -676,7 +676,7 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
         Y1 = Y[info[[variable]] == lvl[1],]
         Y2 = Y[info[[variable]] == lvl[2],]
 
-        res_sLED = sLED_adapt(Y1, Y2, npermute=c(1000,1e5), BPPARAM=BPPARAM)
+        res_sLED = sLED_adapt(Y1, Y2, npermute=c(500,1e6), BPPARAM=BPPARAM)
       }else{
         res_sLED = list(pVal=NA)
       }

### 2020-09-18 common_functions/common_functions.R
Update diff networks

@@ -676,7 +676,7 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
         Y1 = Y[info[[variable]] == lvl[1],]
         Y2 = Y[info[[variable]] == lvl[2],]
 
-        res_sLED = sLED_adapt(Y1, Y2, npermute=c(500,1e6), BPPARAM=BPPARAM)
+        res_sLED = sLED_adapt(Y1, Y2, npermute=c(100,100), BPPARAM=BPPARAM)
       }else{
         res_sLED = list(pVal=NA)
       }
@@ -689,7 +689,7 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
     res$Cohort = names(resid.lst)
     rownames(res) = c()
     res
-  }, mc.cores=4)
+  }, mc.cores=12)
 
   # Format results data.frame
   df_test = do.call(rbind, df_test)

### 2020-09-18 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 ```{r plot.discovery}
-# plot discovery and testing   
+# plot discovery and testing.      
 ggplot(df_test, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")

### 2020-09-18 common_functions/common_functions.R

@@ -676,7 +676,7 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
         Y1 = Y[info[[variable]] == lvl[1],]
         Y2 = Y[info[[variable]] == lvl[2],]
 
-        res_sLED = sLED_adapt(Y1, Y2, npermute=c(100,100), BPPARAM=BPPARAM)
+        res_sLED = sLED_adapt(Y1, Y2, npermute=c(500,5e6), BPPARAM=BPPARAM)
       }else{
         res_sLED = list(pVal=NA)
       }

### 2020-09-18 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 METADATA$Reported_Gender = factor(METADATA$Reported_Gender, c("Male", "Female"))
+```
+
+# NULL network
+
+# Sex network
+## Difference networks
+```{r diff.networks.null}
+
+# Compute difference networks
+##############################
+
+resVersion = "Sex"
+
+METADATA$rnd = sample(c('A','B'), nrow(METADATA), replace=TRUE)
+METADATA$rnd = factor(METADATA$rnd, c('A','B'))
+
+variable = 'rnd'
+
+diff.net = get_difference_network(resid.add[[resVersion]], METADATA, variable)
+
+# Discover clusters in first cohort
+#----------------------------------
+C.diff.discovery = diff.net[[1]]
+
+rm(diff.net) # saving this to disk is slow, so just delete it
+
+hcl = hclust( as.dist(C.diff.discovery) )
+
+dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = 20)
+
+dynamicColors = labels2colors(dynamicMods, colorSeq = colors())
+tab = table(dynamicColors)
+hist(tab, xlab="Cluster size")
+```
+
+```{r testing.null}
+# Hypothesis testing         
+df_test = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, useSLED=FALSE)   #, BPPARAM=SnowParam(18)) 
+```
+
+```{r null.results}
+p = df_test$P.Value[df_test$Cohort=="MSSM-Penn-Pitt"]
+print(1 - qvalue(p)$pi0)
+hist(p)
+
+p = df_test$P.Value[df_test$Cohort== "NIMH-HBCC"]
+1 - qvalue(p)$pi0
+hist(p)
 ```

### 2020-09-18 common_functions/common_functions.R

@@ -647,7 +647,7 @@ sLED_adapt = function(Y1, Y2, npermute=c(1000,1e6), BPPARAM=SerialParam()){
 }
 
 
-test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicColors, METADATA, BPPARAM=SerialParam()){
+test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicColors, METADATA, useSLED=TRUE, BPPARAM=SerialParam()){
 
   col.array = unique(dynamicColors)
 
@@ -670,7 +670,7 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
       # eval statistical hypothesis
       res = boxM_permute( Y, info[[variable]])
 
-      if( key == names(resid.lst)[2] ){
+      if( useSLED && (key == names(resid.lst)[2]) ){
         # sLED
         lvl = levels(info[[variable]])
         Y1 = Y[info[[variable]] == lvl[1],]

### 2020-09-18 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 library(cowplot) 
+library(knitr) 
 library(plyr) 
@@
 p = df_test$P.Value[df_test$Cohort=="MSSM-Penn-Pitt"]
-print(1 - qvalue(p)$pi0)
+# print(1 - qvalue(p)$pi0)
 hist(p)
@@
 p = df_test$P.Value[df_test$Cohort== "NIMH-HBCC"]
-1 - qvalue(p)$pi0
+data.frame(pi1 = 1 - qvalue(p)$pi0, "# DE" = sum(p.adjust(p, 'fdr') < 0.05)) %>% kable
 hist(p)

### 2020-09-18 common_functions/common_functions.R
Update diff networks

@@ -676,7 +676,7 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
         Y1 = Y[info[[variable]] == lvl[1],]
         Y2 = Y[info[[variable]] == lvl[2],]
 
-        res_sLED = sLED_adapt(Y1, Y2, npermute=c(500,5e6), BPPARAM=BPPARAM)
+        res_sLED = sLED_adapt( scale(Y1), scale(Y2), npermute=c(500,5000), BPPARAM=BPPARAM)
       }else{
         res_sLED = list(pVal=NA)
       }

### 2020-09-18 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 
-# NULL network
+# Null network
 
-# Sex network
-## Difference networks
 ```{r diff.networks.null}
@@
 p = df_test$P.Value[df_test$Cohort== "NIMH-HBCC"]
-data.frame(pi1 = 1 - qvalue(p)$pi0, "# DE" = sum(p.adjust(p, 'fdr') < 0.05)) %>% kable
+data.frame("# DE" = sum(p.adjust(p, 'fdr') < 0.05), check.names=FALSE) %>% kable
 hist(p)
@@
 plot_grid( plotlist = figList, ncol=1)
+```
+
+```{r exit, cache=FALSE}
+knitr::knit_exit()
 ```

### 2020-09-18 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 ```{r load.data}
+# Load data
 resid.add = readRDS(synGet('syn22416489')$path)
+vobj.lst = readRDS(synGet('syn22416446')$path)
 METADATA = readRDS(synGet('syn22416411')$path)
 METADATA$Reported_Gender = factor(METADATA$Reported_Gender, c("Male", "Female"))
+
+# identify genes to keep: remove chrY genes
+geneEnsID = lapply(resid.add, function(x){
+	lapply(x, rownames)
+	})
+geneEnsID = unique(unlist(geneEnsID))
+
+df_genes = data.frame(ENSEMBL = geneEnsID, stringsAsFactors=FALSE) %>% getGeneSymbol('ENSEMBL')
+table(df_genes$Chrom)
+
+keep.ensID = df_genes$ENSEMBL[df_genes$Chrom %in% c(1:22, 'X', 'MT')]
 ```
@@
 # Null network
-
 ```{r diff.networks.null}
@@
 #----------------------------------
-C.diff.discovery = diff.net[[1]]
+C.diff.discovery = diff.net[[1]][keep.ensID,keep.ensID]
 

### 2020-09-18 common_functions/common_functions.R

@@ -662,6 +662,7 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
 
       # extact expression residuals for genes in this cluster
       Y = t(resid.lst[[key]][geneid,])
+      # W = t(vobj.lst[[key]][geneid,]$weights)
 
       # extract metadata
       i = match(rownames(Y), rownames(METADATA))

### 2020-09-18 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 keep.ensID = df_genes$ENSEMBL[df_genes$Chrom %in% c(1:22, 'X', 'MT')]
+```
+
+```{r sex.confounding}
+form = ~ Reported_Gender + GABA + GLU + Olig + MgAs + ageOfDeath 
+par(mfrow=c(2,2))
+lapply( unique(METADATA$Institution), function(INST){
+	i = METADATA$Institution == INST
+	C = canCorPairs(form, METADATA[i,]) 
+	plotCorrMatrix(C, main=INST)
+})
 ```
@@
 
-```{r exit, cache=FALSE}
-knitr::knit_exit()
-```
-
 # Disease network
@@
 plot_grid( plotlist = figList, ncol=1)
+```
+
+# Interaction analysis
+
+## Difference networks
+```{r diff.networks.Dx_Sex}
+
+# Compute difference networks
+##############################
+
+METADATA$Dx_Sex = with(METADATA, paste(Reported_Gender, Dx, sep='_'))
+METADATA$Dx_Sex = factor(METADATA$Dx_Sex, c('Female_Control', 'Female_SCZ', 'Male_Control', 'Male_SCZ'))
+
+resVersion = "Dx_Sex"
+variable = 'Dx_Sex'
+
+net.Dx_Sex = get_networks(resid.add[[resVersion]], METADATA, variable)
+
+lvls = levels(METADATA$Dx_Sex)
+
+print(paste0('(',lvls[4], ' - ',lvls[3], ') - (',lvls[2], ' - ',lvls[1], ')'))
+
+# Discover clusters in first cohort
+#----------------------------------
+C = net.Dx_Sex[[1]]
+C.diff.discovery = (C[[lvls[4]]] - C[[lvls[3]]]) - (C[[lvls[2]]] - C[[lvls[1]]])
+
+C.diff.discovery = C.diff.discovery[keep.ensID,keep.ensID]
+
+rm(diff.net) # saving this to disk is slow, so just delete it
+
+hcl = hclust( as.dist(C.diff.discovery) )
+
+dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = 20)
+
+dynamicColors.Dx_Sex = labels2colors(dynamicMods, colorSeq = colors())
+tab = table(dynamicColors.Dx_Sex)
+hist(tab, xlab="Cluster size")
+```
+
+```{r testing.Dx_Sex}
+# Hypothesis testing     
+df_test.Dx_Sex = test_differential_correlation_interaction(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Dx_Sex, METADATA, nperm=100)  
+```
+
+## Discovery cohort
+```{r plot.discovery.Dx_Sex}
+# plot discovery and testing.  
+ggplot(df_test.Dx_Sex, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
+```
+
+```{r modules.in.discovery.Dx_Sex, fig.width=10, fig.height=50}
+
+figList = lapply(df_test$Module[1:5], function(mod){
+
+	plot_module( mod, df_test.Dx_Sex, METADATA, dynamicColors.Dx_Sex, C.diff.discovery, resid.add[[resVersion]], variable, "MSSM-Penn-Pitt")   
+	})
+
+plot_grid( plotlist = figList, ncol=1)
+```
+
+## Replication cohort
+```{r second.cohort.Dx_Sex, fig.width=6, fig.height=12}
+df_cohort2 = df_test.Dx_Sex[df_test.Dx_Sex$Cohort == "NIMH-HBCC",]
+df_cohort2$Module = factor(df_cohort2$Module, unique(df_cohort2$Module))
+table(df_cohort2$FDR < 0.05)
+
+df = df_cohort2[df_cohort2$P.Value < 0.1,]
+
+ggplot(df, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
+```
+
+```{r modules.in.second.cohort..Dx_Sex, fig.width=8, fig.height=100}
+
+i = which(df_cohort2$FDR < 0.05)[1:10]
+
+figList = lapply(df_cohort2$Module[i], function(mod){
+
+	plot_module( mod, df_test.Dx_Sex, METADATA, dynamicColors.Dx_Sex, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC")   
+	})
+
+plot_grid( plotlist = figList, ncol=1)
+```
+
+```{r exit, cache=FALSE}
+knitr::knit_exit()
 ```

### 2020-09-18 common_functions/common_functions.R

@@ -11,6 +11,8 @@ library(data.table)
 library(ggraph)
 library(ggplot2)
 library(tidygraph)
+library(irlba)
+library(EnvStats)
 library(Rfast)
 })
 
@@ -616,6 +618,29 @@ get_difference_network = function(resid.lst, METADATA, variable){
   diff.net
 }
 
+# like get_difference_network, but don't compute difference
+get_networks = function(resid.lst, METADATA, variable){
+
+  # For each cohort
+  nets = lapply( resid.lst, function(resid){
+
+    i = match(colnames(resid), rownames(METADATA))
+    info = METADATA[i,]
+
+    # get correlation matrix for each category
+    C.lst = lapply( levels(info[[variable]]), function(key){
+      j = (info[[variable]] == key)
+      cora( t(resid[,j]) )  
+    })
+    names(C.lst) = levels(info[[variable]])
+
+    C.lst
+    })
+  names(nets) = names(resid.lst)
+
+  nets
+}
+
 
 
 sLED_adapt = function(Y1, Y2, npermute=c(1000,1e6), BPPARAM=SerialParam()){
@@ -712,8 +737,93 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
 
 
 
+test_differential_correlation_interaction = function(resid.lst, C.diff.discovery, dynamicColors, METADATA, nperm=1000){
+
+  col.array = unique(dynamicColors)
+
+  df_test = lapply( col.array, function(col){
+
+    # get genes in this cluster
+    geneid = rownames(C.diff.discovery)[which(dynamicColors==col)]
+
+    res = lapply( names(resid.lst), function(key){
+
+      message(key, ' ', col, ' ', match(col, col.array), '/', length(col.array))
+
+      # extact expression residuals for genes in this cluster
+      Y = t(resid.lst[[key]][geneid,])
+      # W = t(vobj.lst[[key]][geneid,]$weights)
+
+      # extract metadata
+      i = match(rownames(Y), rownames(METADATA))
+      info = METADATA[i,]
+
+      eval_stat = function(idx){
+        lvls = levels(info[[variable]])
+
+        C.lst = lapply( levels(info[[variable]]), function(lvl){
[…]

### 2020-09-18 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 
-diff.net = get_difference_network(resid.add[[resVersion]], METADATA, variable)
+diff.net.Sex = get_difference_network(resid.add[[resVersion]], METADATA, variable)
 
@@
 #----------------------------------
-C.diff.discovery = diff.net[[1]][keep.ensID,keep.ensID]
+C.diff.discovery = diff.net.Sex[[1]][keep.ensID,keep.ensID]
 
@@
 
-dynamicColors = labels2colors(dynamicMods, colorSeq = colors())
-tab = table(dynamicColors)
+dynamicColors.Sex = labels2colors(dynamicMods, colorSeq = colors())
+tab = table(dynamicColors.Sex)
 hist(tab, xlab="Cluster size")
@@
 # Hypothesis testing         
-df_test = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA)#, BPPARAM=SnowParam(18)) 
+df_test.Sex = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Sex, METADATA)#, BPPARAM=SnowParam(18)) 
 ```
@@
 # plot discovery and testing.      
-ggplot(df_test, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
+ggplot(df_test.Sex, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
 ```
@@
 
-figList = lapply(df_test$Module[1:5], function(mod){
+figList = lapply(df_test.Sex$Module[1:5], function(mod){
 
-	plot_module( mod, df_test, METADATA, dynamicColors, C.diff.discovery, resid.add[[resVersion]], variable, "MSSM-Penn-Pitt", 5)   
+	plot_module( mod, df_test.Sex, METADATA, dynamicColors.Sex, C.diff.discovery, resid.add[[resVersion]], variable, "MSSM-Penn-Pitt", 5)   
 	})
@@
 ```{r second.cohort, fig.width=8, fig.height=12}
-df_cohort2 = df_test[df_test$Cohort == "NIMH-HBCC",]
+df_cohort2 = df_test.Sex[df_test.Sex$Cohort == "NIMH-HBCC",]
 df_cohort2$Module = factor(df_cohort2$Module, unique(df_cohort2$Module))
@@
 
-	plot_module( mod, df_test, METADATA, dynamicColors, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC")   
+	plot_module( mod, df_test.Sex, METADATA, dynamicColors.Sex, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC")   
 	})
@@
 
-diff.net = get_difference_network(resid.add[[resVersion]], METADATA, variable)
+diff.net.Dx = get_difference_network(resid.add[[resVersion]], METADATA, variable)
 
@@
 #----------------------------------
-C.diff.discovery = diff.net[[1]][keep.ensID,keep.ensID]
+C.diff.discovery = diff.net.Dx[[1]][keep.ensID,keep.ensID]
 
@@
 
-dynamicColors = labels2colors(dynamicMods, colorSeq = colors())
+dynamicColors.Dx = labels2colors(dynamicMods, colorSeq = colors())
 tab = table(dynamicColors)
@@
 # Hypothesis testing     
-df_test = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA)  
+df_test.Dx = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Dx, METADATA)  
 ```
@@
 # plot discovery and testing.  
-ggplot(df_test, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
+ggplot(df_test.Dx, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
 ```
@@
 
-	plot_module( mod, df_test, METADATA, dynamicColors, C.diff.discovery, resid.add[[resVersion]], variable, "MSSM-Penn-Pitt")   
+	plot_module( mod, df_test.Dx, METADATA, dynamicColors, C.diff.discovery, resid.add[[resVersion]], variable, "MSSM-Penn-Pitt")   
 	})
@@
 ```{r second.cohort.Dx, fig.width=6, fig.height=12}
-df_cohort2 = df_test[df_test$Cohort == "NIMH-HBCC",]
+df_cohort2 = df_test.Dx[df_test.Dx$Cohort == "NIMH-HBCC",]
 df_cohort2$Module = factor(df_cohort2$Module, unique(df_cohort2$Module))
@@
 
-	plot_module( mod, df_test, METADATA, dynamicColors, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC")   
+	plot_module( mod, df_test.Dx, METADATA, dynamicColors.Dx, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC")   
 	})
@@
 
-	plot_module( mod, df_test.Dx_Sex, METADATA, dynamicColors.Dx_Sex, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC")   
+	plot_module( mod, df_test.Dx_Sex, METADATA, dynamicColors.Dx_Sex, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC", lvlidx=1:2)
+
+	plot_module( mod, df_test.Dx_Sex, METADATA, dynamicColors.Dx_Sex, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC", lvlidx=3:4)
+
+	plot_module( mod, df_test.Dx_Sex, METADATA, dynamicColors.Dx_Sex, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC", lvlidx=1:4) 
 	})

### 2020-09-18 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 # plot discovery and testing.      
-ggplot(df_test.Sex, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
+ggplot(df_test.Sex, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test.Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
 ```
@@
 # plot discovery and testing.  
-ggplot(df_test.Dx, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
+ggplot(df_test.Dx, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test.Dx$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
 ```
@@
 # plot discovery and testing.  
-ggplot(df_test.Dx_Sex, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
+ggplot(df_test.Dx_Sex, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test.Dx_Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
 ```

### 2020-09-18 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 
-rm(diff.net) # saving this to disk is slow, so just delete it
+rm(diff.net.Sex) # saving this to disk is slow, so just delete it
 
@@
 
-rm(diff.net) # saving this to disk is slow, so just delete it
+rm(diff.net.Dx) # saving this to disk is slow, so just delete it
 
@@
 
-```{r modules.in.second.cohort..Dx_Sex, fig.width=8, fig.height=100}
+```{r modules.in.second.cohort.Dx_Sex, fig.width=8, fig.height=100}
 
-i = which(df_cohort2$FDR < 0.05)[1:10]
+i = which(df_cohort2$FDR < 0.05)
 
@@
 
-	plot_module( mod, df_test.Dx_Sex, METADATA, dynamicColors.Dx_Sex, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC", lvlidx=1:2)
+	fig1 = plot_module( mod, df_test.Dx_Sex, METADATA, dynamicColors.Dx_Sex, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC", lvlidx=1:2)
 
-	plot_module( mod, df_test.Dx_Sex, METADATA, dynamicColors.Dx_Sex, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC", lvlidx=3:4)
+	fig2 = plot_module( mod, df_test.Dx_Sex, METADATA, dynamicColors.Dx_Sex, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC", lvlidx=3:4)
 
-	plot_module( mod, df_test.Dx_Sex, METADATA, dynamicColors.Dx_Sex, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC", lvlidx=1:4) 
+	fig3 = plot_module( mod, df_test.Dx_Sex, METADATA, dynamicColors.Dx_Sex, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC", lvlidx=1:4)
+
+	plot_grid(fig1, fig2, fig2, ncol=1)
 	})

### 2020-09-18 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 Since we are generating networks across cohorts and sexes, identifiying conserved clusters using standard methods becomes non-trivial.
-Instead of defining clusters for each subset and evaluating the difference, we first compute the differences in the correlation networks, then identify a single set a clusters based on this meta-network.
-Then use sLED test on difference network, or use boxM_permute on original residuals, but with consensus clusters.
+Instead of defining clusters for each subset and evaluating the difference, we first compute the differences in the correlation networks, then identify a single set of clusters based on this difference-network.
+Then use hypothesis test of differential correlation in the second cohort to get p-value.
 
@@
 library(cowplot) 
+library(zenith) 
 library(knitr) 
@@
 df_genes = data.frame(ENSEMBL = geneEnsID, stringsAsFactors=FALSE) %>% getGeneSymbol('ENSEMBL')
-table(df_genes$Chrom)
+# table(df_genes$Chrom)
 
@@
 
-```{r sex.confounding}
+```{r load.go}
+# Load Gene Ontology database 
+go.gs = get_GeneOntology()
+```
+
+```{r sex.confounding, fig.width=5, fig.height=20}
 form = ~ Reported_Gender + GABA + GLU + Olig + MgAs + ageOfDeath 
 par(mfrow=c(2,2))
-lapply( unique(METADATA$Institution), function(INST){
+res = lapply( unique(METADATA$Institution), function(INST){
 	i = METADATA$Institution == INST
@@
 
+## Module enrichment
+```{r top.enrich.Dx}
+
+enrich.test = function(df_module, gs){
+
+	res = lapply( unique(df_module$Module), function(mod){
+
+			genes = df_module[df$]
+
+		})
+
+}
+
+```
+
 # Disease network
@@
 
-rm(diff.net) # saving this to disk is slow, so just delete it
+rm(net.Dx_Sex) # saving this to disk is slow, so just delete it
 
@@
 # Hypothesis testing     
-df_test.Dx_Sex = test_differential_correlation_interaction(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Dx_Sex, METADATA, nperm=100)  
+df_test.Dx_Sex = test_differential_correlation_interaction(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Dx_Sex, METADATA, nperm=1000)    
 ```
@@
 # plot discovery and testing.  
-ggplot(df_test.Dx_Sex, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test.Dx_Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
+ggplot(df_test.Dx_Sex, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test.Dx_Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed").
 ```
@@
 
-figList = lapply(df_test$Module[1:5], function(mod){
+figList = lapply(df_test$Module[1:5], function(mod){.
 
@@
 
-```{r modules.in.second.cohort.Dx_Sex, fig.width=8, fig.height=100}
+```{r modules.in.second.cohort.Dx_Sex, fig.width=8, fig.height=80}
 
@@
 
-	plot_grid(fig1, fig2, fig2, ncol=1)
+	plot_grid(fig1, fig2, fig3, ncol=1)
 	})

### 2020-09-18 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 
+```{r exit, cache=FALSE, eval=FALSE}
+knitr::knit_exit()
+```
+
 ## Module enrichment
-```{r top.enrich.Dx}
+```{r top.enrich.Sex}
+df_module = data.frame(	ENSEMBL = rownames(C.diff.discovery) %>% trim_ensembl_ids, 
+						Module = dynamicColors.Sex)
 
-enrich.test = function(df_module, gs){
+df_module_enrich.Sex = enrich.test( df_module, go.gs, df_cohort2$Module[i])
+```
 
-	res = lapply( unique(df_module$Module), function(mod){
-
-			genes = df_module[df$]
-
-		})
-
-}
-
-```
+plot_enrich(df_module_enrich.Sex[1:10,], module) + ggtitle()
 
@@
 plot_grid( plotlist = figList, ncol=1)
+```
+
+## Module enrichment
+```{r top.enrich.Sex}
+df_module = data.frame(	ENSEMBL = rownames(C.diff.discovery) %>% trim_ensembl_ids, 
+						Module = dynamicColors.Dx)
+
+df_module_enrich.Dx = enrich.test( df_module, go.gs, df_cohort2$Module[i])
 ```
@@
 plot_grid( plotlist = figList, ncol=1)
+```
+
+## Module enrichment
+```{r top.enrich.Sex}
+df_module = data.frame(	ENSEMBL = rownames(C.diff.discovery) %>% trim_ensembl_ids, 
+						Module = dynamicColors.Dx_Sex)
+
+df_module_enrich.Dx_Sex = enrich.test( df_module, go.gs, df_cohort2$Module[i])
 ```

### 2020-09-18 common_functions/common_functions.R

@@ -14,6 +14,7 @@ library(tidygraph)
 library(irlba)
 library(EnvStats)
 library(Rfast)
+library(GSEABase)
 })
 
 # For ENSEMBL id ENSG00000279457.4, return ENSG00000279457
@@ -980,6 +981,72 @@ plot_corr_network = function( C, zcutoff = 1.3, seed=1, base_size=11){
 # min(df2$z)
 
 # plot_corr_network( C1-C2, zcutoff=1.568965)
+
+
+
+
+enrich.test = function(df_module, gs_set, testModules = unique(df_module$Module) ){
+
+  # Map from Ensembl genes in geneSets_GO to 
+  # from trimmed Ensembl names from RNA-seq data 
+  gs.list = limma::ids2indices( recodeToList(gs_set), df_module$ENSEMBL)
+     
+  # filter by size of gene set
+  n_genes_in = 10
+  gs.list = gs.list[sapply(gs.list, length) >= n_genes_in]
+
+  # for each module
+  res = lapply( testModules, function(mod){
+
+    # Indicator for Module
+    isInModule = rep(0, nrow(df_module))
+    isInModule[df_module$Module == mod] = 1 
+
+    # for each gene set
+    df = lapply( gs.list, function(gsidx){
+
+      # Indicator for Gene
+      isInGeneset = rep(0, nrow(df_module))
+      isInGeneset[gsidx] = 1
+
+      tab = table(isInGeneset, isInModule)
+        
+      fit = fisher.test(tab)
+
+      data.frame( Module = mod,
+            n.genes = sum(tab[,2]),
+            # Geneset = gsid,
+            n.overlap = tab[2,2],
+            OR = fit$estimate,
+            p.value = fit$p.value)
+    })
+    df = do.call(rbind, df)
+    df$Geneset = names(gs.list) 
+    df = df[,c("Module", "n.genes", "Geneset", "n.overlap", "OR", "p.value")]
+    df = df[!is.na(df$p.value) & (df$p.value < 0.99999) & (df$OR !=0),]
+    df$FDR = p.adjust(df$p.value, 'fdr')
+    df = df[order(df$p.value),]
+
+    df
+  })
+  do.call(rbind, res)
+}
+
+
+
+plot_enrich = function( df, module ){
+
+  df2 = df[df$Module == module,]
+  df2$Module = factor(df2$Module, df2$Module)
+
[…]

### 2020-09-18 common_functions/common_functions.R
Update diff networks

@@ -1037,7 +1037,7 @@ enrich.test = function(df_module, gs_set, testModules = unique(df_module$Module)
 plot_enrich = function( df, module ){
 
   df2 = df[df$Module == module,]
-  df2$Module = factor(df2$Module, df2$Module)
+  df2$Geneset = factor(df2$Geneset, df2$Geneset)
 
   ylim = max(-log10(df$FDR))
   ggplot(df2, aes(Geneset, -log10(FDR))) + geom_bar(stat='identity', fill=module) + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + coord_flip() + ylab(bquote(-log[10]~FDR)) + xlab('') + scale_y_continuous(expand=c(0,0), lim=c(0, ylim*1.05)) 

### 2020-09-18 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 ## Module enrichment
-```{r top.enrich.Sex}
+```{r top.enrich.Dx}
 df_module = data.frame(	ENSEMBL = rownames(C.diff.discovery) %>% trim_ensembl_ids, 
@@
 ## Module enrichment
-```{r top.enrich.Sex}
+```{r top.enrich.Dx_sex}
 df_module = data.frame(	ENSEMBL = rownames(C.diff.discovery) %>% trim_ensembl_ids, 

### 2020-09-18 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 
-```{r exit, cache=FALSE, eval=FALSE}
-knitr::knit_exit()
-```
-
 ## Module enrichment
@@
 ## Module enrichment
-```{r top.enrich.Dx_sex}
+```{r top.enrich.Dx_Sex}
 df_module = data.frame(	ENSEMBL = rownames(C.diff.discovery) %>% trim_ensembl_ids, 

### 2020-09-18 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 # plot discovery and testing.  
-ggplot(df_test.Dx_Sex, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test.Dx_Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed").
+ggplot(df_test.Dx_Sex, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test.Dx_Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
 ```

### 2020-09-18 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 
-```{r sex.confounding, fig.width=5, fig.height=20}
+```{r sex.confounding, fig.width=5, fig.height=8}
 form = ~ Reported_Gender + GABA + GLU + Olig + MgAs + ageOfDeath 
@@
 
-plot_enrich(df_module_enrich.Sex[1:10,], module) + ggtitle()
+```{r plot.enrich.Sex}
+figList = lapply( df_cohort2$Module[i], function(module){
+	plot_enrich(df_module_enrich.Sex, module) + ggtitle(module)
+	})
+plot_grid(plotlist = figList, ncol=2)
+```
 
@@
 
+```{r plot.enrich.Dx}
+figList = lapply( df_cohort2$Module[i], function(module){
+	plot_enrich(df_module_enrich.Dx, module) + ggtitle(module)
+	})
+plot_grid(plotlist = figList, ncol=2)
+```
+
 # Interaction analysis
@@
 
-i = which(df_cohort2$FDR < 0.05)
+i = which(df_cohort2$FDR < 0.1)
 
@@
 df_module_enrich.Dx_Sex = enrich.test( df_module, go.gs, df_cohort2$Module[i])
+```
+
+```{r plot.enrich.Dx_Sex}
+figList = lapply( df_cohort2$Module[i], function(module){
+	plot_enrich(df_module_enrich.Dx_Sex, module) + ggtitle(module)
+	})
+plot_grid(plotlist = figList, ncol=2)
 ```

### 2020-09-18 common_functions/common_functions.R

@@ -893,7 +893,7 @@ plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.disc
 
   lim = range(c(df2$cor1, df2$cor2))
 
-  fig1 = ggplot(df2, aes(cor1, cor2, label=pair)) + geom_point(aes(color=ifelse(pair=='', "grey", "red"))) + theme_bw(base_size) + theme(aspect.ratio = 1, legend.position='none', plot.title = element_text(hjust = 0.5))  + geom_abline(color="grey", linetype="dashed") + geom_text_repel(direction='x', nudge_x=1, hjust=1, segment.size = 0.2, box.padding=.2, size=base_size/2) + scale_color_manual(values=c("grey50", "red")) + xlim(lim) + ylim(lim) + ggtitle(main) + xlab(paste('Correlation in', lvl[lvlidx[1]])) + ylab(paste('Correlation in', lvl[lvlidx[2]]))
+  fig1 = ggplot(df2, aes(cor1, cor2, label=pair)) + geom_abline(color="grey", linetype="dashed") + geom_point(aes(color=ifelse(pair=='', "grey", "red"))) + theme_bw(base_size) + theme(aspect.ratio = 1, legend.position='none', plot.title = element_text(hjust = 0.5))  + geom_text_repel(direction='x', nudge_x=1, hjust=1, segment.size = 0.2, box.padding=.2, size=base_size/2) + scale_color_manual(values=c("grey50", "red")) + xlim(lim) + ylim(lim) + ggtitle(main) + xlab(paste('Correlation in', lvl[lvlidx[1]])) + ylab(paste('Correlation in', lvl[lvlidx[2]])) 
 
   # upper is C2 and lower is C1, since C1 is baseline, is lvl[1]
   C = C2
@@ -933,6 +933,9 @@ plot_corr_network = function( C, zcutoff = 1.3, seed=1, base_size=11){
   df_net$weight = apply(df_net, 1, function(x){
     C[x['name.from'], x['name.to']]
   })
+  # threshold weight at 1
+  df_net$weight = pmax(-1.2, pmin(1.2, df_net$weight))
+
   df_net$z = scale(df_net$weight)
   df_net = df_net[abs(df_net$z) >= zcutoff,]
 
@@ -957,7 +960,7 @@ plot_corr_network = function( C, zcutoff = 1.3, seed=1, base_size=11){
[…]

### 2020-09-18 combined_analysis/diff.networks.Rmd
Update diff networks

--- 2020-09-18 combined_analysis/diff.networks.Rmd
@@
 
-```{r sex.confounding, fig.width=5, fig.height=20}
+```{r sex.confounding, fig.width=5, fig.height=4}
 form = ~ Reported_Gender + GABA + GLU + Olig + MgAs + ageOfDeath 
@@
 
-plot_enrich(df_module_enrich.Sex[1:10,], module) + ggtitle()
+```{r plot.enrich.Sex, fig.width=9}
+figList = lapply( df_cohort2$Module[i], function(module){
+	plot_enrich(df_module_enrich.Sex, module, 8) + ggtitle(module)
+	})
+plot_grid(plotlist = figList, ncol=1)
+```
 
@@
 
+```{r plot.enrich.Dx, fig.width=9}
+figList = lapply( df_cohort2$Module[i], function(module){
+	plot_enrich(df_module_enrich.Dx, module, 8) + ggtitle(module)
+	})
+plot_grid(plotlist = figList, ncol=1)
+```
+
 # Interaction analysis
@@
 df_module_enrich.Dx_Sex = enrich.test( df_module, go.gs, df_cohort2$Module[i])
+```
+
+```{r plot.enrich.Dx_Sex, fig.width=9}
+figList = lapply( df_cohort2$Module[i], function(module){
+	plot_enrich(df_module_enrich.Dx_Sex, module, 8) + ggtitle(module)
+	})
+plot_grid(plotlist = figList, ncol=1)
 ```

### 2020-09-18 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 
-```{r sex.confounding, fig.width=5, fig.height=4}
+```{r sex.confounding, fig.width=5, fig.height=5}
 form = ~ Reported_Gender + GABA + GLU + Olig + MgAs + ageOfDeath 

### 2020-09-18 common_functions/common_functions.R
Update diff networks

@@ -1037,13 +1037,13 @@ enrich.test = function(df_module, gs_set, testModules = unique(df_module$Module)
 
 
 
-plot_enrich = function( df, module ){
+plot_enrich = function( df, module, base_size=11 ){
 
   df2 = df[df$Module == module,]
   df2$Geneset = factor(df2$Geneset, df2$Geneset)
 
   ylim = max(-log10(df$FDR))
-  ggplot(df2[1:10,], aes(Geneset, -log10(FDR))) + geom_bar(stat='identity', fill=module) + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + coord_flip() + ylab(bquote(-log[10]~FDR)) + xlab('') + scale_y_continuous(expand=c(0,0), lim=c(0, ylim*1.05)) 
+  ggplot(df2[1:10,], aes(Geneset, -log10(FDR))) + geom_bar(stat='identity', fill=module) + theme_bw(base_size) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + coord_flip() + ylab(bquote(-log[10]~FDR)) + xlab('') + scale_y_continuous(expand=c(0,0), lim=c(0, ylim*1.05)) 
 }
 
 

### 2020-09-18 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 	plot_enrich(df_module_enrich.Sex, module, 8) + ggtitle(module)
-	})
+	}). 
 plot_grid(plotlist = figList, ncol=1)

### 2020-09-18 common_functions/common_functions.R

@@ -1043,7 +1043,7 @@ plot_enrich = function( df, module, base_size=11 ){
   df2$Geneset = factor(df2$Geneset, df2$Geneset)
 
   ylim = max(-log10(df$FDR))
-  ggplot(df2[1:10,], aes(Geneset, -log10(FDR))) + geom_bar(stat='identity', fill=module) + theme_bw(base_size) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + coord_flip() + ylab(bquote(-log[10]~FDR)) + xlab('') + scale_y_continuous(expand=c(0,0), lim=c(0, ylim*1.05)) 
+  ggplot(df2[1:10,], aes(Geneset, -log10(FDR))) + geom_bar(stat='identity', fill=module) + theme_bw(base_size) + theme(plot.title = element_text(hjust = 0.5)) + coord_flip() + ylab(bquote(-log[10]~FDR)) + xlab('') + scale_y_continuous(expand=c(0,0), lim=c(0, ylim*1.05)) 
 }
 
 

### 2020-09-19 combined_analysis/diff.networks.Rmd
Update diff networks

--- 2020-09-18 combined_analysis/diff.networks.Rmd
@@
 keep.ensID = df_genes$ENSEMBL[df_genes$Chrom %in% c(1:22, 'X', 'MT')]
+
+minClusterSize = 50
 ```
@@
 })
+```
+
+```{r load.dream}
+fit_marginal.lst = readRDS(synGet('syn22416500')$path)
+fit_inter_sex.disease.lst = readRDS(synGet('syn22416524')$path)
 ```
@@
 
-dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = 20)
+dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize)
 
@@
 
-```{r null.results}
+```{r null.results, fig.width=5}
 p = df_test$P.Value[df_test$Cohort=="MSSM-Penn-Pitt"]
 # print(1 - qvalue(p)$pi0)
-hist(p)
+hist(p, main="P-values under the null in the discovery cohort", xlab="p-value")
 
@@
 data.frame("# DE" = sum(p.adjust(p, 'fdr') < 0.05), check.names=FALSE) %>% kable
-hist(p)
+hist(p, main="P-values under the null in the replication cohort", xlab="p-value")
 ```
@@
 ```{r top.enrich.Sex}
-df_module = data.frame(	ENSEMBL = rownames(C.diff.discovery) %>% trim_ensembl_ids, 
+df_module.Sex = data.frame(	ENSEMBL = rownames(C.diff.discovery) %>% trim_ensembl_ids, 
 						Module = dynamicColors.Sex)
 
-df_module_enrich.Sex = enrich.test( df_module, go.gs, df_cohort2$Module[i])
+df_module_enrich.Sex = enrich.test( df_module.Sex, go.gs, df_cohort2$Module[i])
 ```
@@
 ```{r top.enrich.Dx}
-df_module = data.frame(	ENSEMBL = rownames(C.diff.discovery) %>% trim_ensembl_ids, 
+df_module.Dx = data.frame(	ENSEMBL = rownames(C.diff.discovery) %>% trim_ensembl_ids, 
 						Module = dynamicColors.Dx)
 
-df_module_enrich.Dx = enrich.test( df_module, go.gs, df_cohort2$Module[i])
+df_module_enrich.Dx = enrich.test( df_module.Dx, go.gs, df_cohort2$Module[i])
 ```
@@
 ```{r top.enrich.Dx_Sex}
-df_module = data.frame(	ENSEMBL = rownames(C.diff.discovery) %>% trim_ensembl_ids, 
+df_module.Dx_Sex = data.frame(	ENSEMBL = rownames(C.diff.discovery) %>% trim_ensembl_ids, 
 						Module = dynamicColors.Dx_Sex)
 
-df_module_enrich.Dx_Sex = enrich.test( df_module, go.gs, df_cohort2$Module[i])
+df_module_enrich.Dx_Sex = enrich.test( df_module.Dx_Sex, go.gs, df_cohort2$Module[i])
 ```

### 2020-09-19 common_functions/common_functions.R

@@ -995,7 +995,7 @@ enrich.test = function(df_module, gs_set, testModules = unique(df_module$Module)
   gs.list = limma::ids2indices( recodeToList(gs_set), df_module$ENSEMBL)
      
   # filter by size of gene set
-  n_genes_in = 10
+  n_genes_in = 50
   gs.list = gs.list[sapply(gs.list, length) >= n_genes_in]
 
   # for each module

### 2020-09-19 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 cd /hpc/users/hoffmg01/work/covarr-de
-# rm -rf combined_analysis/diff.networks_cache
+# rm -rf combined_analysis/diff.networks_* combined_analysis/diff.networks.html
 ml git pandoc udunits proj gdal geos
@@
 
-minClusterSize = 50
+minClusterSize = 20
 ```
@@
                          to = c(2, 3, 4, 1, 1, 2, 1, 2, 3))
-rstat_edges$weight = 1:9
+rstat_edges$weight = rnorm(9)
 
@@
 
-ggraph(net, layout = "auto") + 
+ggraph(net, layout = "stress") + 
   geom_node_point(size=10, color="blue") +

### 2020-09-20 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 fit_inter_sex.disease.lst = readRDS(synGet('syn22416524')$path)
+df_meta = readRDS(synGet('syn22416508')$path)
+df_meta.inter_sex.disease = readRDS(synGet('syn22416532')$path)
 ```
@@
 
-df_module_enrich.Sex = enrich.test( df_module.Sex, go.gs, df_cohort2$Module[i])
+df_module_enrich.Sex = enrich.test( df_module.Sex, go.gs, df_cohort2$Module[i], 150).  
 ```
@@
 	plot_enrich(df_module_enrich.Sex, module, 8) + ggtitle(module)
+	}) 
+plot_grid(plotlist = figList, ncol=1)
+```
+
+## DE enrichment
+```{r cameraPR.Sex}
+res_enrich_DE = enrich_module_DE( df_module.Sex )
+
+figList = lapply( df_cohort2$Module[i], function(module){
+
+	res = res_enrich_DE[res_enrich_DE$Geneset == module,]
+
+	ggplot(res, aes(Test, -log10(PValue))) + geom_bar(stat="identity", fill=res$Geneset[1]) + coord_flip() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5))  + scale_y_continuous(expand=c(0,0)) + ggtitle(module)
 }) 
@@
 
-df_module_enrich.Dx = enrich.test( df_module.Dx, go.gs, df_cohort2$Module[i])
+df_module_enrich.Dx = enrich.test( df_module.Dx, go.gs, df_cohort2$Module[i], 150)
 ```
@@
 	plot_enrich(df_module_enrich.Dx, module, 8) + ggtitle(module)
+	})    
+plot_grid(plotlist = figList, ncol=1)
+```
+
+## DE enrichment
+```{r cameraPR.Dx}
+res_enrich_DE = enrich_module_DE( df_module.Dx )
+
+figList = lapply( df_cohort2$Module[i], function(module){
+
+	res = res_enrich_DE[res_enrich_DE$Geneset == module,]
+
+	ggplot(res, aes(Test, -log10(PValue))) + geom_bar(stat="identity", fill=res$Geneset[1]) + coord_flip() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5))  + scale_y_continuous(expand=c(0,0)) + ggtitle(module)
 }) 
@@
 
-df_module_enrich.Dx_Sex = enrich.test( df_module.Dx_Sex, go.gs, df_cohort2$Module[i])
+df_module_enrich.Dx_Sex = enrich.test( df_module.Dx_Sex, go.gs, df_cohort2$Module[i], 150)
 ```
@@
 	plot_enrich(df_module_enrich.Dx_Sex, module, 8) + ggtitle(module) 
+	}) 
+plot_grid(plotlist = figList, ncol=1)
+```
+
+## DE enrichment
+```{r cameraPR.Dx_Sex}
+res_enrich_DE = enrich_module_DE( df_module_enrich.Dx_Sex )
+
+figList = lapply( df_cohort2$Module[i], function(module){
+
+	res = res_enrich_DE[res_enrich_DE$Geneset == module,]
+
+	ggplot(res, aes(Test, -log10(PValue))) + geom_bar(stat="identity", fill=res$Geneset[1]) + coord_flip() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5))  + scale_y_continuous(expand=c(0,0)) + ggtitle(module)
 }) 

### 2020-09-20 common_functions/common_functions.R

@@ -988,14 +988,17 @@ plot_corr_network = function( C, zcutoff = 1.3, seed=1, base_size=11){
 
 
 
-enrich.test = function(df_module, gs_set, testModules = unique(df_module$Module) ){
+enrich.test = function(df_module, gs_set, testModules = unique(df_module$Module), minSize=50 ){
+
+
+  df_module = df_module[df_module$ENSEMBL %in% unique(unlist(geneIds(gs_set))),]
 
   # Map from Ensembl genes in geneSets_GO to 
   # from trimmed Ensembl names from RNA-seq data 
   gs.list = limma::ids2indices( recodeToList(gs_set), df_module$ENSEMBL)
      
   # filter by size of gene set
-  n_genes_in = 50
+  n_genes_in = minSize
   gs.list = gs.list[sapply(gs.list, length) >= n_genes_in]
 
   # for each module

### 2020-09-20 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 ```
+
+```{r load.WES}
+df_schema = fread(synGet('syn22343823')$path)
+
+df_schema2 = df_schema[,data.frame(ENSEMBL = Gene,
+									OR1 = `OR (Class I)`,
+									OR2 = `OR (Class II)`,
+									P.value = `P meta`)]
 
@@
 
-df_module_enrich.Sex = enrich.test( df_module.Sex, go.gs, df_cohort2$Module[i], 150).  
+df_module_enrich.Sex = enrich.test( df_module.Sex, go.gs, df_cohort2$Module[i], 150) 
 ```
@@
 ```{r cameraPR.Dx_Sex}
-res_enrich_DE = enrich_module_DE( df_module_enrich.Dx_Sex )
+res_enrich_DE = enrich_module_DE( df_module.Dx_Sex )
 

### 2020-09-20 common_functions/common_functions.R
Update diff networks

@@ -990,7 +990,7 @@ plot_corr_network = function( C, zcutoff = 1.3, seed=1, base_size=11){
 
 enrich.test = function(df_module, gs_set, testModules = unique(df_module$Module), minSize=50 ){
 
-
+  # only keep genes that are in at least one gene set
   df_module = df_module[df_module$ENSEMBL %in% unique(unlist(geneIds(gs_set))),]
 
   # Map from Ensembl genes in geneSets_GO to 
@@ -1053,9 +1053,35 @@ plot_enrich = function( df, module, base_size=11 ){
 
 
 
+enrich_module_DE = function(df_module){ 
+  # create gene set lists from modules
+  mods = unique(df_module$Module)
+  modules.gs = lapply( mods, function(mod){
+    df_module$ENSEMBL[df_module$Module == mod]
+    })
+  names(modules.gs) = unique(df_module$Module)
 
+  keys = c(names(df_meta), names(df_meta.inter_sex.disease))
 
+  res_enrich_DE = lapply( keys, function(key){
 
+    df = df_meta[[key]]
+
+    if( is.null(df) ) df = df_meta.inter_sex.disease[[key]]
+    
+    idx = match( trim_ensembl_ids(rownames(df)), df_module$ENSEMBL)
+    modules.list = limma::ids2indices( modules.gs, df_module$ENSEMBL[idx])
+         
+    res = cameraPR( with(df, logFC/se), modules.list )
+    res$Geneset = rownames(res)
+    rownames(res) = c()
+    res$Test = key
+    res
+  })
+  res_enrich_DE = do.call(rbind, res_enrich_DE)
+
+  res_enrich_DE
+}
 
 
 

### 2020-09-20 common_functions/common_functions.R
Update diff networks

@@ -94,7 +94,6 @@ plot_concordance = function( resList, showGenes = NULL, idx=c(1,2), size=15){
   res = with(df, cor.test(logFC.x, logFC.y, method="spearman"))
   tab_corr = data.frame(R_spearman = res$estimate, P = res$p.value)
   
-
   pi1_x = 1 - qvalue(df$P.Value.x)$pi0
   pi1_y = 1 - qvalue(df$P.Value.y)$pi0
 
@@ -103,20 +102,28 @@ plot_concordance = function( resList, showGenes = NULL, idx=c(1,2), size=15){
 
   p = with(df, P.Value.y[adj.P.Val.x < 0.05])
   n.de.x = length(p)
-  pi_discovery_x = tryCatch( 
-    1 - qvalue(p)$pi0, 
-    error = function(e){
-      # must have a p-value > 0.95 to work
-      1 - qvalue(c(1,p))$pi0
-      })
+  if( n.de.x > 0){
+    pi_discovery_x = tryCatch( 
+      1 - qvalue(p)$pi0, 
+      error = function(e){
+        # must have a p-value > 0.95 to work
+        1 - qvalue(c(1,p))$pi0
+        })
+  }else{
+      pi_discovery_x = 0
+  }
 
   p = with(df, P.Value.x[adj.P.Val.y < 0.05])
   n.de.y = length(p)
-  pi_discovery_y = tryCatch( 
-    1 - qvalue(p)$pi0, 
-    error = function(e){
-      NA
-      })
+  if( n.de.y > 0){
+    pi_discovery_y = tryCatch( 
+      1 - qvalue(p)$pi0, 
+      error = function(e){
+        NA
+        })
+  }else{
+    pi_discovery_y = 0
+  }
 
   tab_pi = data.frame(Discovery = names(resList), pi1 = c(pi_discovery_x, pi_discovery_y))
 

### 2020-09-20 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 
-	ggplot(res, aes(Test, -log10(PValue))) + geom_bar(stat="identity", fill=res$Geneset[1]) + coord_flip() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5))  + scale_y_continuous(expand=c(0,0)) + ggtitle(module)
+	lim = max(-log10(res$PValue))
+
+	ggplot(res, aes(Test, -log10(PValue))) + geom_bar(stat="identity", fill=res$Geneset[1]) + coord_flip() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5))  + scale_y_continuous(expand=c(0,0), lim=c(0, lim*1.05)) + ggtitle(module) + geom_hline(yintercept=-log10(0.05), color="red", linetype="dashed")
 }) 
@@
 ## DE enrichment
-```{r cameraPR.Dx}
+```{r cameraPR.Dx, fig.width=7, fig.height=20}
 res_enrich_DE = enrich_module_DE( df_module.Dx )

### 2020-09-20 common_functions/common_functions.R

@@ -1094,5 +1094,3 @@ enrich_module_DE = function(df_module){
 
 
 
-
-

### 2020-09-20 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 									P.value = `P meta`)]
-
-```{r sex.confounding, fig.width=5, fig.height=5}
-form = ~ Reported_Gender + GABA + GLU + Olig + MgAs + ageOfDeath 
-par(mfrow=c(2,2))
-res = lapply( unique(METADATA$Institution), function(INST){
-	i = METADATA$Institution == INST
-	C = canCorPairs(form, METADATA[i,]) 
-	plotCorrMatrix(C, main=INST)
-})
 ```

### 2020-09-20 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 
-# Compute difference networks
+# Compute difference networks for multiple random permutations
 ##############################
@@
 
+df_test = lapply( 1:10, function(k){
 	METADATA$rnd = sample(c('A','B'), nrow(METADATA), replace=TRUE)
@@
 	tab = table(dynamicColors)
-hist(tab, xlab="Cluster size")
-```
 
-```{r testing.null}
+	# hist(tab, xlab="Cluster size")
+
 	# Hypothesis testing         
-df_test = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, useSLED=FALSE)   #, BPPARAM=SnowParam(18)) 
+	df_test = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, useSLED=FALSE)
+
+	df_test
+})
+df_test = do.call(rbind, df_test)
 ```
@@
 
+fit_perm_null = beta.mle(p)
+
+calc_p = function(p, fit){
+	pbeta(p, 	shape1=fit_perm_null$param['alpha'], 
+				shape2=fit_perm_null$param['beta'], 
+				lower.tail=TRUE)
+}
+
 p = df_test$P.Value[df_test$Cohort== "NIMH-HBCC"]
-data.frame("# DE" = sum(p.adjust(p, 'fdr') < 0.05), check.names=FALSE) %>% kable
+data.frame("# DE" = sum(p.adjust(p, 'fdr') < 0.05) / length(p), check.names=FALSE) %>% kable
 hist(p, main="P-values under the null in the replication cohort", xlab="p-value")

### 2020-09-20 common_functions/common_functions.R
Update diff networks

@@ -723,7 +723,7 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
     res$Cohort = names(resid.lst)
     rownames(res) = c()
     res
-  }, mc.cores=12)
+  }, mc.cores=4)
 
   # Format results data.frame
   df_test = do.call(rbind, df_test)

### 2020-09-20 common_functions/common_functions.R
Update diff networks

@@ -684,7 +684,7 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
 
   col.array = unique(dynamicColors)
 
-  df_test = mclapply( col.array, function(col){
+  df_test = lapply( col.array, function(col){
 
     # get genes in this cluster
     geneid = rownames(C.diff.discovery)[which(dynamicColors==col)]
@@ -723,7 +723,7 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
     res$Cohort = names(resid.lst)
     rownames(res) = c()
     res
-  }, mc.cores=4)
+  })#, mc.cores=4)
 
   # Format results data.frame
   df_test = do.call(rbind, df_test)

### 2020-09-20 common_functions/common_functions.R
Update diff networks

@@ -680,7 +680,7 @@ sLED_adapt = function(Y1, Y2, npermute=c(1000,1e6), BPPARAM=SerialParam()){
 }
 
 
-test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicColors, METADATA, useSLED=TRUE, BPPARAM=SerialParam()){
+test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicColors, METADATA, variable, useSLED=TRUE, BPPARAM=SerialParam()){
 
   col.array = unique(dynamicColors)
 
@@ -723,7 +723,7 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
     res$Cohort = names(resid.lst)
     rownames(res) = c()
     res
-  })#, mc.cores=4)
+  }, mc.cores=4)
 
   # Format results data.frame
   df_test = do.call(rbind, df_test)

### 2020-09-20 common_functions/common_functions.R
Update diff networks

@@ -684,7 +684,7 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
 
   col.array = unique(dynamicColors)
 
-  df_test = lapply( col.array, function(col){
+  df_test = mclapply( col.array, function(col){
 
     # get genes in this cluster
     geneid = rownames(C.diff.discovery)[which(dynamicColors==col)]

### 2020-09-20 combined_analysis/combine_all_DLPFC.Rmd
filer

@@
 library(cowplot)
+library(ggraph)
+library(tidygraph)
 library(kableExtra)
@@
 METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))
+
+# drop individuals who where sequenced twice here
+tab = table(METADATA$Individual_ID)
+
+dropSamples = sapply( names(tab[tab > 1]), function(id){
+
+  idx = which(METADATA$Individual_ID == id)
+  df = METADATA[idx,]
+
+  # remove individuals that have less then max reads
+  i = which(df$MappedReads < max(df$MappedReads))
+  rownames(METADATA)[idx[i]]
+  } )
+METADATA = METADATA[!(rownames(METADATA) %in% dropSamples),]
 ```
@@
 kable(caption="Diagnosis, Institution, Sex") %>% kable_styling(full_width=FALSE) 
-
-ggplot(METADATA, aes(Reported_Gender, ageOfDeath)) + geom_boxplot() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + facet_wrap(~Institution, nrow=1)
+```
+
+```{r joint.cca, fig.width=7, fig.height=7}
+variables = c('Reported_Gender', 'CellType', 'ageOfDeath', 'Institution', "Dx")
+
+df_cca = lapply(variables, function(v1){
+
+  if( v1 == "CellType"){
+    form1 = ~ GABA + GLU + Olig + MgAs
+  }else{
+    form1 = as.formula(paste0('~', v1))
+  }
+  M1 = model.matrix(form1, METADATA)
+
+  res = lapply(variables, function(v2){
+
+    if( v2 == "CellType"){
+      form2 = ~ GABA + GLU + Olig + MgAs
+    }else{
+      form2 = as.formula(paste0('~', v2))
+    }
+    M2 = model.matrix(form2, METADATA)
+
+    fit = cancor(M1, M2)
+    V_sqrt = sqrt(mean(fit$cor^2)) # Cramer's sqrt V
+
+    data.frame(v1, v2, V_sqrt)
+  })
+  do.call(rbind, res)
+})
+df_cca = do.call(rbind, df_cca)
+
+C = matrix(df_cca$V_sqrt, ncol=length(variables))
+rownames(C) = variables
+colnames(C) = variables
+
+plotCorrMatrix(C, main='CCA')
+```
+
+```{r explain.CellType}
+CTs = c('GABA', 'GLU', 'Olig', 'MgAs')
+
+form = ~ ageOfDeath:Institution + Reported_Gender + Dx + Institution + Dx
+
+Y = clr(METADATA[,CTs])
+
+vp = fitExtractVarPartModel( t(Y), form, METADATA, BPPARAM=SerialParam())
+plotPercentBars(vp) + ggtitle("Explain cell type composition")
+
+ymax = max(rowSums(vp[,-ncol(vp)]))*110
+plotPercentBars(vp) + scale_y_continuous(expand=c(0,0), limits=c(0, ymax))
+```
+
+```{r sex.graph, fig.width=7, fig.height=7}
+library(ggplot2)
+library(ggraph)
+library(tidygraph)
+
+df_cca2 = df_cca[with(df_cca, v1!=v2),]
+
+rstat_nodes <- data.frame(name = unique(df_cca2$v1))
+
+rstat_edges <- data.frame(from = match(df_cca2$v1, rstat_nodes$name),
+                         to = match(df_cca2$v2, rstat_nodes$name),
+                         Rsq = df_cca2$V_sqrt^2 )
+
+net = tbl_graph(nodes = rstat_nodes, edges = rstat_edges, directed=FALSE)
+
+ggraph(net,  layout = 'igraph', algorithm = 'star') + 
+  geom_edge_link(aes(width = Rsq, color=Rsq), alpha = 1) +
+  geom_node_point(size=20, color="grey70") +
+  geom_node_text(aes(label = name), repel = FALSE) +
+  theme_graph() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) +
+  scale_edge_width(name=bquote(R^2)) +
+  scale_edge_color_gradient(name=bquote(R^2), low="white", high="red", lim=c(0, max(df_cca2$V_sqrt^2))) +
+  ggtitle("Pairwise squared correlation")
+```
+
+```{r sex.confounding, fig.width=5, fig.height=5}
+form = ~ Reported_Gender + GABA + GLU + Olig + MgAs + ageOfDeath 
+par(mfrow=c(2,2))
+res = lapply( unique(METADATA$Institution), function(INST){
+  i = METADATA$Institution == INST
+  C = canCorPairs(form, METADATA[i,]) 
+  plotCorrMatrix(C, main=INST)
+})
+```
+
+```{r sex.by.age}
+ggplot(METADATA, aes(Reported_Gender, ageOfDeath, fill=Reported_Gender)) + geom_boxplot() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + facet_wrap(~Institution, nrow=1) + scale_fill_manual(values = c("red", "blue"))
 ```
@@
 figList = lapply(colnames(df_cellFractions), function(CT){
-  ggplot(METADATA, aes_string('Reported_Gender', CT)) + geom_boxplot() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + facet_wrap(~Institution, nrow=1)
+  ggplot(METADATA, aes_string('Reported_Gender', CT, fill='Reported_Gender')) + geom_boxplot() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + facet_wrap(~Institution, nrow=1) + scale_fill_manual(values = c("red", "blue"))
 })
@@
 
+# drop chrY genes
+df_genes  = data.frame(ENSEMBL = rownames(vobj.lst[[1]])) %>% getGeneSymbol('ENSEMBL')
+keep.ensID = df_genes$ENSEMBL[df_genes$Chrom %in% c(1:22, 'X', 'MT')]
+
 fit_inter_sex.disease.lst = lapply( vobj.lst, function(vobj){
@@
 
-  fit = dream(vobj, form_use, info, BPPARAM = SnowParam(18))  
+  fit = dream(vobj[keep.ensID,], form_use, info, BPPARAM = SnowParam(18))  
   eBayes( fit )
@@
 ```
+
+# Dx test with in Males and then Females
 

### 2020-09-20 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 library(cowplot) 
+library(metap) 
 library(zenith) 
@@
 
-# Null network
+# Null permuations
 ```{r diff.networks.null}
@@
 	# Hypothesis testing         
-	df_test = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, useSLED=FALSE)
+	df_test = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, variable, useSLED=TRUE)
 
@@
 
-fit_perm_null = beta.mle(p)
+beta_perm = list()
+beta_perm[['MSSM-Penn-Pitt']] = beta.mle(p)
 
 calc_p = function(p, fit){
-	pbeta(p, 	shape1=fit_perm_null$param['alpha'], 
-				shape2=fit_perm_null$param['beta'], 
+	pbeta(p, 	shape1=fit$param['alpha'], 
+				shape2=fit$param['beta'], 
 				lower.tail=TRUE)
@@
 hist(p, main="P-values under the null in the replication cohort", xlab="p-value")
+
+beta_perm[['NIMH-HBCC']] = beta.mle(p)
+```
+
+# Null permuations
+```{r diff.networks.null}
+
+# Compute difference networks for multiple random permutations
+##############################
+
+resVersion = "Sex"
+
+METADATA$rnd = sample(c('A','B'), nrow(METADATA), replace=TRUE)
+METADATA$rnd = factor(METADATA$rnd, c('A','B'))
+
+variable = 'rnd'
+
+diff.net = get_difference_network(resid.add[[resVersion]], METADATA, variable)
+
+# Discover clusters in first cohort
+#----------------------------------
+C.diff.discovery = diff.net[[1]][keep.ensID,keep.ensID]
+
+rm(diff.net) # saving this to disk is slow, so just delete it
+
+hcl = hclust( as.dist(C.diff.discovery) )
+
+dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize)
+
+dynamicColors = labels2colors(dynamicMods, colorSeq = colors())
+tab = table(dynamicColors)
+
+# hist(tab, xlab="Cluster size")
+
+# Hypothesis testing         
+df_test_null = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, variable, useSLED=TRUE) 
 ```
@@
 # Hypothesis testing         
-df_test.Sex = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Sex, METADATA)#, BPPARAM=SnowParam(18)) 
+df_test.Sex = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Sex, METADATA, variable, useSLED=TRUE)
+```
+
+```{r test, eval=FALSE}
+df2 = lapply( unique(df_test$Cohort), function(chrt){
+	df = df_test[df_test$Cohort == chrt,]
+	df$P.Value = calc_p(df$P.Value, beta_perm[[chrt]])
+	df$FDR = p.adjust(df$P.Value, "fdr")
+	df
+})
+df2 = do.call(rbind, df2)
+
+df3 = lapply( unique(df2$Module), function(mod){
+
+	message(mod)
+	# combine p-values with Fisher's method
+	p.meta = sumlog(df2$P.Value[df2$Module == mod])$p
+
+	data.frame(Module = mod, P.Value = p.meta)
+} )
+df3 = do.call(rbind, df3)
+df3 = df3[!is.na(df3$P.Value),]
+df3$FDR = qvalue(c(1,df3$P.Value))$qvalues[-1]
+
+df3$FDR = p.adjust(df3$P.Value, "fdr")
 ```
@@
 # Hypothesis testing     
-df_test.Dx = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Dx, METADATA)  
+df_test.Dx = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Dx, METADATA, variable, useSLED=TRUE)  
 ```

### 2020-09-20 common_functions/common_functions.R

@@ -704,13 +704,13 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
       # eval statistical hypothesis
       res = boxM_permute( Y, info[[variable]])
 
-      if( useSLED && (key == names(resid.lst)[2]) ){
+      if( useSLED ){
         # sLED
         lvl = levels(info[[variable]])
         Y1 = Y[info[[variable]] == lvl[1],]
         Y2 = Y[info[[variable]] == lvl[2],]
 
-        res_sLED = sLED_adapt( scale(Y1), scale(Y2), npermute=c(500,5000), BPPARAM=BPPARAM)
+        res_sLED = sLED_adapt( scale(Y1), scale(Y2), npermute=c(500,50000), BPPARAM=BPPARAM)
       }else{
         res_sLED = list(pVal=NA)
       }

### 2020-09-20 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 # Null permuations
-```{r diff.networks.null}
+```{r diff.networks.perm}
 
@@
 
-```{r null.results, fig.width=5}
+```{r perm.results, fig.width=5}
 p = df_test$P.Value[df_test$Cohort=="MSSM-Penn-Pitt"]
@@
 
-# Compute difference networks for multiple random permutations
-##############################
+# Compute difference networks for 1 random permutation
+######################################################
 

### 2020-09-21 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 	# Hypothesis testing         
-	df_test = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, variable, useSLED=TRUE)
+	df_test = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, variable, useSLED=FALSE)
 
@@
 
-# hist(tab, xlab="Cluster size")
+# Hypothesis testing         
+df_test_null = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, variable, useSLED=FALSE) 
+```
 
-# Hypothesis testing         
-df_test_null = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, variable, useSLED=TRUE) 
+```{r recalibrate.null}
+
+df_test_recal = apply_permutation_null( df_test_null )
+
+df_test_meta = module_meta_analsis( df_test_recal )
+
 ```
@@
 # Hypothesis testing         
-df_test.Sex = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Sex, METADATA, variable, useSLED=TRUE)
-```
-
-```{r test, eval=FALSE}
-df2 = lapply( unique(df_test$Cohort), function(chrt){
-	df = df_test[df_test$Cohort == chrt,]
-	df$P.Value = calc_p(df$P.Value, beta_perm[[chrt]])
-	df$FDR = p.adjust(df$P.Value, "fdr")
-	df
-})
-df2 = do.call(rbind, df2)
-
-df3 = lapply( unique(df2$Module), function(mod){
-
-	message(mod)
-	# combine p-values with Fisher's method
-	p.meta = sumlog(df2$P.Value[df2$Module == mod])$p
-
-	data.frame(Module = mod, P.Value = p.meta)
-} )
-df3 = do.call(rbind, df3)
-df3 = df3[!is.na(df3$P.Value),]
-df3$FDR = qvalue(c(1,df3$P.Value))$qvalues[-1]
-
-df3$FDR = p.adjust(df3$P.Value, "fdr")
+df_test.Sex = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Sex, METADATA, variable, useSLED=FALSE)
 ```
@@
 # Hypothesis testing     
-df_test.Dx = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Dx, METADATA, variable, useSLED=TRUE)  
+df_test.Dx = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Dx, METADATA, variable, useSLED=FALSE)  
 ```
@@
 
+X = Rfast::matrnorm(10, 4)
+Y = Rfast::matrnorm(10, 4)
+
+C1 = cor(X)
+C2 = cor(Y)
+D = C1 - C2
+
+pdf("/Users/gabrielhoffman/Dropbox/Presentations/CMC3/CMC_DE_sex/figures/diffCor.pdf")
+variancePartition::plotCorrMatrix(C1, sort=FALSE, dendrogram="none")
+variancePartition::plotCorrMatrix(C2, sort=FALSE, dendrogram="none")
+variancePartition::plotCorrMatrix(D, sort=FALSE, dendrogram="none")
+dev.off()
+
+set.seed(1)
+X1 = Rfast::matrnorm(20, 4)
+X2 = Rfast::matrnorm(20, 4)
+X3 = Rfast::matrnorm(20, 4)
+X4 = Rfast::matrnorm(20, 4)
+
+C1 = cor(X1)
+C2 = cor(X2)
+C3 = cor(X3)
+C4 = cor(X4)
+Da = C1 - C2
+Db = C3 - C4
+
+pdf("/Users/gabrielhoffman/Dropbox/Presentations/CMC3/CMC_DE_sex/figures/diffCor2.pdf")
+variancePartition::plotCorrMatrix(C1, sort=FALSE, dendrogram="none")
+variancePartition::plotCorrMatrix(C2, sort=FALSE, dendrogram="none")
+variancePartition::plotCorrMatrix(C3, sort=FALSE, dendrogram="none")
+variancePartition::plotCorrMatrix(C4, sort=FALSE, dendrogram="none")
+variancePartition::plotCorrMatrix(Da-Db, sort=FALSE, dendrogram="none")
+dev.off()
+
 # need to generate correct residuals

### 2020-09-21 common_functions/common_functions.R

@@ -1092,5 +1092,33 @@ enrich_module_DE = function(df_module){
 
 
 
+apply_permutation_null = function(df_test){
+  df2 = lapply( unique(df_test$Cohort), function(chrt){
+    df = df_test[df_test$Cohort == chrt,]
+    df$P.Value = calc_p(df$P.Value, beta_perm[[chrt]])
+    df$FDR = p.adjust(df$P.Value, "fdr")
+    df
+  })
+  df2 = do.call(rbind, df2)
+  df2
+}
+
+module_meta_analsis = function(df2){
+
+  df3 = lapply( unique(df2$Module), function(mod){
+
+    message(mod)
+    # combine p-values with Fisher's method
+    p.meta = sumlog(df2$P.Value[df2$Module == mod])$p
+
+    data.frame(Module = mod, P.Value = p.meta)
+  } )
+  df3 = do.call(rbind, df3)
+  df3 = df3[!is.na(df3$P.Value),]
+  df3$FDR = qvalue(c(1,df3$P.Value))$qvalues[-1]
+
+  df3$FDR = p.adjust(df3$P.Value, "fdr")
+  df3
+}
 
 

### 2020-09-21 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 
-```{r plot.enrich.Dx, fig.width=9}
+```{r plot.enrich.Dx, fig.width=9, fig.height=25}
 figList = lapply( df_cohort2$Module[i], function(module){
@@
 # plot discovery and testing.  
-ggplot(df_test.Dx_Sex, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test.Dx_Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
+ggplot(df_test.Dx_Sex, aes(Module, -log10(P.Value))) + geom_bar(stat='identity', fill=df_test.Dx_Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~P)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
 ```
@@
 dev.off()
+
+stat = sapply(1:1000, function(i){
+
+	set.seed(i)
+	n = 1000
+	p = 40
+	X = matrix(rnorm(n*p), n,p)
+	Y = matrix(rnorm(n*p), n,p)
+
+	C1 = cor(X)
+	C2 = cor(Y)
+	D = C1 - C2
+
+	stat = sLED:::sLEDTestStat(D, rho=1)$stats
+	max(0, stat)
+})
+
+fit = egamma( stat )
+
+fit$parameters
 

### 2020-09-21 common_functions/common_functions.R

@@ -782,9 +782,11 @@ test_differential_correlation_interaction = function(resid.lst, C.diff.discovery
 
         C.diff = C_alt - C_baseline
         # svd(C.diff, nu=0, nv=0)$d[1]
-        partial_eigen(C.diff, n=1)$values
+        # partial_eigen(C.diff, n=1)$values
+        max(0, sLEDTestStat(C.diff, rho=1, sumabs.seq = .5)$stats)
       }
 
+      # evaluate on real data
       val = eval_stat(1:nrow(info))
 
       # permutation

### 2020-09-21 common_functions/common_functions.R
Update diff networks

@@ -783,7 +783,7 @@ test_differential_correlation_interaction = function(resid.lst, C.diff.discovery
         C.diff = C_alt - C_baseline
         # svd(C.diff, nu=0, nv=0)$d[1]
         # partial_eigen(C.diff, n=1)$values
-        max(0, sLEDTestStat(C.diff, rho=1, sumabs.seq = .5)$stats)
+        max(0, sLED:::sLEDTestStat(C.diff, rho=1, sumabs.seq = .5)$stats)
       }
 
       # evaluate on real data

### 2020-09-21 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 # Hypothesis testing     
-df_test.Dx_Sex = test_differential_correlation_interaction(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Dx_Sex, METADATA, nperm=1000)    
+df_test.Dx_Sex = test_differential_correlation_interaction(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Dx_Sex, METADATA, nperm=20)    
 ```
@@
 
-figList = lapply(df_test$Module[1:5], function(mod){.
+figList = lapply(df_test.Dx_Sex$Module[1:5], function(mod){.
 

### 2020-09-21 common_functions/common_functions.R

@@ -679,12 +679,49 @@ sLED_adapt = function(Y1, Y2, npermute=c(1000,1e6), BPPARAM=SerialParam()){
   res
 }
 
+sLED_perm = function(Y, info, variable, nperm, rho=1, sumabs.seq=.5){
+
+  eval_stat = function(idx){
+    lvls = levels(info[[variable]])
+
+    C.lst = lapply( levels(info[[variable]]), function(lvl){
+      j = (info[[variable]] == lvl)
+      cora( Y[idx,][j,]) 
+    })
+    names(C.lst) = levels(info[[variable]])
+
+    D = C.lst[[lvls[2]]] - C.lst[[lvls[1]]]
+    # svd(C.diff, nu=0, nv=0)$d[1]
+    # partial_eigen(C.diff, n=1)$values
+    max(0, sLED:::sLEDTestStat(D, rho=rho, sumabs.seq = sumabs.seq)$stats)
+  }
+
+  # evaluate on real data
+  val = eval_stat(1:nrow(info))
+
+  # permutation
+  val_perm = sapply( seq_len(nperm), function(k){
+    idx = sample.int(nrow(info), nrow(info))
+    eval_stat( idx )
+  })
+
+   # estimate null distribution as a gamma
+  fit = egamma( val_perm )
+  up = pgamma(val, shape=fit$parameters[1], scale=fit$parameters[2], lower.tail=FALSE)
+  down = pgamma(val, shape=fit$parameters[1], scale=fit$parameters[2])
+  P.Value = 2*min(c(up, down))
+
+  P.Value
+}
+
+
+
 
 test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicColors, METADATA, variable, useSLED=TRUE, BPPARAM=SerialParam()){
 
   col.array = unique(dynamicColors)
 
-  df_test = mclapply( col.array, function(col){
+  df_test = lapply( col.array, function(col){
 
     # get genes in this cluster
     geneid = rownames(C.diff.discovery)[which(dynamicColors==col)]
@@ -702,7 +739,7 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
       info = METADATA[i,]
 
       # eval statistical hypothesis
-      res = boxM_permute( Y, info[[variable]])
+      # res = boxM_permute( Y, info[[variable]])
 
       if( useSLED ){
         # sLED
[…]

### 2020-09-21 common_functions/common_functions.R
Update diff networks

@@ -741,7 +741,7 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
       # eval statistical hypothesis
       # res = boxM_permute( Y, info[[variable]])
 
-      if( useSLED ){
+      # if( useSLED ){
         # sLED
         lvl = levels(info[[variable]])
         Y1 = Y[info[[variable]] == lvl[1],]
@@ -754,9 +754,9 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
 
         p.sled = sLED_perm(Y, info, variable, 20)
 
-      }else{
-        p.sled = NA
-      }
+      # }else{
+      #   p.sled = NA
+      # }
       data.frame( Module  = col, 
                   # P.Value       = res$p.value, 
                   P.Value  = p.sled,

### 2020-09-21 common_functions/common_functions.R
Update diff networks

@@ -826,7 +826,7 @@ test_differential_correlation_interaction = function(resid.lst, C.diff.discovery
         C.diff = C_alt - C_baseline
         # svd(C.diff, nu=0, nv=0)$d[1]
         # partial_eigen(C.diff, n=1)$values
-        max(0, sLED:::sLEDTestStat(C.diff, rho=1, sumabs.seq = .5)$stats)
+        max(0, sLED:::sLEDTestStat(C.diff, rho=1, sumabs.seq = 1)$stats)
       }
 
       # evaluate on real data
@@ -843,7 +843,7 @@ test_differential_correlation_interaction = function(resid.lst, C.diff.discovery
       up = pgamma(val, shape=fit$parameters[1], scale=fit$parameters[2], lower.tail=FALSE)
       down = pgamma(val, shape=fit$parameters[1], scale=fit$parameters[2])
       P.Value = 2*min(c(up, down))
-
+  
       data.frame( Module        = col, 
                   P.Value       = P.Value, 
                   n.genes       = length(geneid))
@@ -1140,7 +1140,9 @@ enrich_module_DE = function(df_module){
 apply_permutation_null = function(df_test){
   df2 = lapply( unique(df_test$Cohort), function(chrt){
     df = df_test[df_test$Cohort == chrt,]
-    df$P.Value = calc_p(df$P.Value, beta_perm[[chrt]])
+    if( chrt != 'NIMH-HBCC' ){
+      df$P.Value = calc_p(df$P.Value, beta_perm[[chrt]])
+    }
     df$FDR = p.adjust(df$P.Value, "fdr")
     df
   })
@@ -1148,11 +1150,9 @@ apply_permutation_null = function(df_test){
   df2
 }
 
-module_meta_analsis = function(df2){
+module_meta_analysis = function(df2){
 
   df3 = lapply( unique(df2$Module), function(mod){
-
-    message(mod)
     # combine p-values with Fisher's method
     p.meta = sumlog(df2$P.Value[df2$Module == mod])$p
 

### 2020-09-21 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 	#----------------------------------
-	C.diff.discovery = diff.net[[1]][keep.ensID,keep.ensID]
+	C = diff.net[[1]]
+	C.diff.discovery = C[[2]] - C[[1]]
+	C.diff.discovery = C.diff.discovery[[1]][keep.ensID,keep.ensID]
 
@@
 
-# Null permuations
+# Null network
 ```{r diff.networks.null}
@@
 #----------------------------------
-C.diff.discovery = diff.net[[1]][keep.ensID,keep.ensID]
+C = diff.net[[1]]
+C.diff.discovery = C[[2]] - C[[1]]
+C.diff.discovery = C.diff.discovery[[1]][keep.ensID,keep.ensID]
 
@@
 ```{r recalibrate.null}
+df_test_recal = apply_permutation_null( df_test_null )
 
-df_test_recal = apply_permutation_null( df_test_null )
+table(df_test_recal$FDR < 0.05)
 
@@
 
+table(df_test_meta$FDR < 0.05)
 ```
@@
 
+res = pickSFT( diff.net.Sex )
+
+ggplot(res, aes(Power, -sign(slope)*SFT.R.sq, color=paste(Cohort, Dataset, sep=' - '))) + geom_point() + theme_bw() + theme(aspect.ratio=1,  plot.title = element_text(hjust = 0.5)) + xlab(="Soft Threshold (power)") + ylab("Scale Free Topology Model Fit,signed R^2") + ggtitle("Scale independence")
+
 # Discover clusters in first cohort
 #----------------------------------
-C.diff.discovery = diff.net.Sex[[1]][keep.ensID,keep.ensID]
+C = diff.net[[1]]
+C.diff.discovery = C[[2]] - C[[1]]
+C.diff.discovery = C.diff.discovery[[1]][keep.ensID,keep.ensID]
 
@@
 df_test.Sex = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Sex, METADATA, variable, useSLED=FALSE)
+```
+
+df_sft = pickSFT( diff.net.Sex )
+
+```{r recalibrate.Sex}
+table(df_test.Sex$FDR < 0.05)
+
+df_test_recal = apply_permutation_null( df_test.Sex )
+
+table(df_test_recal$FDR < 0.05)
+
+df_test_meta = module_meta_analysis( df_test_recal )
+
+table(df_test_meta$FDR < 0.05)
 ```
@@
 
-i = which(df_cohort2$FDR < 0.05)
+# i = which(df_cohort2$FDR < 0.05)
+i = 1:5
 
@@
 df_test.Dx = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Dx, METADATA, variable, useSLED=FALSE)  
+```
+
+```{r recalibrate.Dx}
+table(df_test.Dx$FDR < 0.05)
+
+df_test_recal = apply_permutation_null( df_test.Dx )
+
+table(df_test_recal$FDR < 0.05)
+
+df_test_meta = module_meta_analysis( df_test_recal )
+
+table(df_test_meta$FDR < 0.05)
 ```
@@
 
+```{r recalibrate.Dx_Sex}
+table(df_test.Dx_Sex$FDR < 0.05)
+
+df_test_recal = apply_permutation_null( df_test.Dx_Sex )
+
+table(df_test_recal$FDR < 0.05)
+
+df_test_meta = module_meta_analysis( df_test_recal )
+
+table(df_test_meta$FDR < 0.051)
+```
+
 ## Discovery cohort
@@
 
-i = which(df_cohort2$FDR < 0.1)
+# i = which(df_cohort2$FDR < 0.1) 
+i = 1:5
 

### 2020-09-21 common_functions/common_functions.R

@@ -606,7 +606,7 @@ zenith_meta_analysis = function(res_zenith, coefTest){
 get_difference_network = function(resid.lst, METADATA, variable){
 
   # For each cohort
-  diff.net = lapply( resid.lst, function(resid){
+  C.lst = lapply( resid.lst, function(resid){
 
     i = match(colnames(resid), rownames(METADATA))
     info = METADATA[i,]
@@ -618,12 +618,11 @@ get_difference_network = function(resid.lst, METADATA, variable){
     })
     names(C.lst) = levels(info[[variable]])
 
-    # Compute difference in correlation matrices
-    C.lst[[1]] - C.lst[[2]]
-    })
-  names(diff.net) = names(resid.lst)
+    C.lst
+  })
+  names(C.lst) = names(resid.lst)
 
-  diff.net
+  C.lst
 }
 
 # like get_difference_network, but don't compute difference
@@ -1165,5 +1164,26 @@ module_meta_analysis = function(df2){
   df3$FDR = p.adjust(df3$P.Value, "fdr")
   df3
 }
+
+pickSFT = function( diff.net ){
+
+  df = lapply( names(diff.net), function(key){
+    df = lapply( names(diff.net[[key]]), function(key2){
+      res = pickSoftThreshold( diff.net[[key]][[key2]][1:1000, 1:1000] )
+      df = res$fitIndices
+      df$Cohort = key
+      df$Dataset = key2
+      df
+      })
+    do.call(rbind, df)
+  })
+  do.call(rbind, df)
+}
+
+
+
+
+
+
 
 

### 2020-09-21 common_functions/common_functions.R
Update diff networks

@@ -1169,7 +1169,7 @@ pickSFT = function( diff.net ){
 
   df = lapply( names(diff.net), function(key){
     df = lapply( names(diff.net[[key]]), function(key2){
-      res = pickSoftThreshold( diff.net[[key]][[key2]][1:1000, 1:1000] )
+      res = pickSoftThreshold( diff.net[[key]][[key2]] )
       df = res$fitIndices
       df$Cohort = key
       df$Dataset = key2

### 2020-09-21 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 # Null permuations
-```{r diff.networks.perm}
+```{r diff.networks.perm, eval=FALSE}
 
@@
 
+	pwr = 7
+
 	# Discover clusters in first cohort
@@
 	C = diff.net[[1]]
-	C.diff.discovery = C[[2]] - C[[1]]
-	C.diff.discovery = C.diff.discovery[[1]][keep.ensID,keep.ensID]
+	C.diff.discovery = C[[2]]^pwr - C[[1]]^pwr
+	C.diff.discovery = C.diff.discovery[keep.ensID,keep.ensID]
 
@@
 
-```{r perm.results, fig.width=5}
+```{r perm.results, fig.width=5, eval=FALSE}
 p = df_test$P.Value[df_test$Cohort=="MSSM-Penn-Pitt"]
@@
 # Null network
-```{r diff.networks.null}
+```{r diff.networks.null, eval=FALSE}
 
@@
 C = diff.net[[1]]
-C.diff.discovery = C[[2]] - C[[1]]
-C.diff.discovery = C.diff.discovery[[1]][keep.ensID,keep.ensID]
+C.diff.discovery = C[[2]]^pwr - C[[1]]^pwr
+C.diff.discovery = C.diff.discovery[keep.ensID,keep.ensID]
 
@@
 
-ggplot(res, aes(Power, -sign(slope)*SFT.R.sq, color=paste(Cohort, Dataset, sep=' - '))) + geom_point() + theme_bw() + theme(aspect.ratio=1,  plot.title = element_text(hjust = 0.5)) + xlab(="Soft Threshold (power)") + ylab("Scale Free Topology Model Fit,signed R^2") + ggtitle("Scale independence")
+ggplot(res, aes(Power, -sign(slope)*SFT.R.sq, color=paste(Cohort, Dataset, sep=' - '))) + geom_point() + theme_bw() + theme(aspect.ratio=1,  plot.title = element_text(hjust = 0.5)) + xlab("Soft Threshold (power)") + ylab("Scale Free Topology Model Fit,signed R^2") + ggtitle("Scale independence")
 
@@
 ```
-
-df_sft = pickSFT( diff.net.Sex )
 
@@
 
+res = pickSFT( diff.net.Dx )
+
+ggplot(res, aes(Power, -sign(slope)*SFT.R.sq, color=paste(Cohort, Dataset, sep=' - '))) + geom_point() + theme_bw() + theme(aspect.ratio=1,  plot.title = element_text(hjust = 0.5)) + xlab("Soft Threshold (power)") + ylab("Scale Free Topology Model Fit,signed R^2") + ggtitle("Scale independence")
+
 # Discover clusters in first cohort
 #----------------------------------
-C.diff.discovery = diff.net.Dx[[1]][keep.ensID,keep.ensID]
+
+C = diff.net[[1]]
+C.diff.discovery = C[[2]]^pwr - C[[1]]^pwr
+C.diff.discovery = C.diff.discovery[keep.ensID,keep.ensID]
 
@@
 
+res = pickSFT( net.Dx_Sex )
+
+ggplot(res, aes(Power, -sign(slope)*SFT.R.sq, color=paste(Cohort, Dataset, sep=' - '))) + geom_point() + theme_bw() + theme(aspect.ratio=1,  plot.title = element_text(hjust = 0.5)) + xlab("Soft Threshold (power)") + ylab("Scale Free Topology Model Fit,signed R^2") + ggtitle("Scale independence")
+
+pwr = 7
+
 lvls = levels(METADATA$Dx_Sex)
@@
 C = net.Dx_Sex[[1]]
-C.diff.discovery = (C[[lvls[4]]] - C[[lvls[3]]]) - (C[[lvls[2]]] - C[[lvls[1]]])
+C.diff.discovery = (C[[lvls[4]]]^pwr - C[[lvls[3]]]^pwr) - (C[[lvls[2]]]^pwr - C[[lvls[1]]]^pwr)
 

### 2020-09-21 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 
-```{r recalibrate.null}
+```{r recalibrate.null, eval=FALSE}
 df_test_recal = apply_permutation_null( df_test_null )
@@
 #----------------------------------
-C = diff.net[[1]]
+C = diff.net.Sex[[1]]
 C.diff.discovery = C[[2]]^pwr - C[[1]]^pwr

### 2020-09-21 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 
-minClusterSize = 20
+minClusterSize = 50
+pwr = 7
 ```
@@
 
-	dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize)
+	dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize, deepSplit=2)
 
-	dynamicColors = labels2colors(dynamicMods, colorSeq = colors())
+	dynamicColors = labels2colors(dynamicMods)
 	tab = table(dynamicColors)
@@
 
-dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize)
+dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize, deepSplit=2)
 
-dynamicColors = labels2colors(dynamicMods, colorSeq = colors())
+dynamicColors = labels2colors(dynamicMods)
 tab = table(dynamicColors)
@@
 
-dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize)
+dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize, deepSplit=2)
 
-dynamicColors.Sex = labels2colors(dynamicMods, colorSeq = colors())
+dynamicColors.Sex = labels2colors( dynamicMods )
 tab = table(dynamicColors.Sex)
@@
 # Hypothesis testing         
-df_test.Sex = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Sex, METADATA, variable, useSLED=FALSE)
+df_test.Sex = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Sex, METADATA, variable, .3, 20)
 ```
@@
 
-C = diff.net[[1]]
+C = diff.net.Dx[[1]]
 C.diff.discovery = C[[2]]^pwr - C[[1]]^pwr
@@
 
-dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize)
+dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize, deepSplit=2)
 
-dynamicColors.Dx = labels2colors(dynamicMods, colorSeq = colors())
+dynamicColors.Dx = labels2colors(dynamicMods )
 tab = table(dynamicColors)
@@
 
-dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize)
+dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize, deepSplit=2)
 
-dynamicColors.Dx_Sex = labels2colors(dynamicMods, colorSeq = colors())
+dynamicColors.Dx_Sex = labels2colors(dynamicMods)
 tab = table(dynamicColors.Dx_Sex)

### 2020-09-21 common_functions/common_functions.R

@@ -678,9 +678,9 @@ sLED_adapt = function(Y1, Y2, npermute=c(1000,1e6), BPPARAM=SerialParam()){
   res
 }
 
-sLED_perm = function(Y, info, variable, nperm, rho=1, sumabs.seq=.5){
+sLED_perm = function(Y, info, variable, nperm, rho=1, sumabs.seq=.2){
 
-  eval_stat = function(idx){
+  eval_stat = function(idx, sumabs.seq){
     lvls = levels(info[[variable]])
 
     C.lst = lapply( levels(info[[variable]]), function(lvl){
@@ -692,16 +692,22 @@ sLED_perm = function(Y, info, variable, nperm, rho=1, sumabs.seq=.5){
     D = C.lst[[lvls[2]]] - C.lst[[lvls[1]]]
     # svd(C.diff, nu=0, nv=0)$d[1]
     # partial_eigen(C.diff, n=1)$values
-    max(0, sLED:::sLEDTestStat(D, rho=rho, sumabs.seq = sumabs.seq)$stats)
+
+    res = sLED:::sLEDTestStat(D, rho=rho, sumabs.seq = sumabs.seq)
+
+    data.frame(sumabs.seq, 
+              n.active=rowSums(res$leverage > 0), 
+              stats=pmax(0, res$stats))
   }
 
   # evaluate on real data
-  val = eval_stat(1:nrow(info))
+  res = eval_stat(1:nrow(info), sumabs.seq)#=seq(.2, 1, length.out=5))
+  val = res$stats
 
   # permutation
   val_perm = sapply( seq_len(nperm), function(k){
     idx = sample.int(nrow(info), nrow(info))
-    eval_stat( idx )
+    eval_stat( idx, sumabs.seq )$stats
   })
 
    # estimate null distribution as a gamma
@@ -710,13 +716,14 @@ sLED_perm = function(Y, info, variable, nperm, rho=1, sumabs.seq=.5){
   down = pgamma(val, shape=fit$parameters[1], scale=fit$parameters[2])
   P.Value = 2*min(c(up, down))
 
-  P.Value
+  data.frame(p = P.Value, n.active = res$n.active)
 }
 
 
 
 
-test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicColors, METADATA, variable, useSLED=TRUE, BPPARAM=SerialParam()){
+
+test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicColors, METADATA, variable, sumabs.seq, nperm){
 
[…]

### 2020-09-21 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 # Null permuations
-```{r diff.networks.perm, eval=FALSE}
+```{r diff.networks.perm}
 
@@
 
-	dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize, deepSplit=2)
+	# dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize, deepSplit=2)
+	dynamicMods = cutree(hcl, k=200)
 
@@
 	# Hypothesis testing         
-	df_test = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, variable, useSLED=FALSE)
+	df_test = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, variable, useSLED=TRUE, .5, 20)
 
@@
 
-```{r perm.results, fig.width=5, eval=FALSE}
+```{r perm.results, fig.width=5}
 p = df_test$P.Value[df_test$Cohort=="MSSM-Penn-Pitt"]
@@
 # Null network
-```{r diff.networks.null, eval=FALSE}
+```{r diff.networks.null}
 
@@
 
-dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize, deepSplit=2)
+# dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize, deepSplit=2)
+dynamicMods = cutree(hcl, k=200)
 
-dynamicColors = labels2colors(dynamicMods)
+dynamicColors = labels2colors(dynamicMods, colorSeq=colors())
 tab = table(dynamicColors)
@@
 # Hypothesis testing         
-df_test_null = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, variable, useSLED=FALSE) 
+df_test_null = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, variable, useSLED=TRUE, .5, 20) 
 ```
 
-```{r recalibrate.null, eval=FALSE}
+```{r recalibrate.null}
 df_test_recal = apply_permutation_null( df_test_null )
@@
 
-res = pickSFT( diff.net.Sex )
+# res = pickSFT( diff.net.Sex )
 
-ggplot(res, aes(Power, -sign(slope)*SFT.R.sq, color=paste(Cohort, Dataset, sep=' - '))) + geom_point() + theme_bw() + theme(aspect.ratio=1,  plot.title = element_text(hjust = 0.5)) + xlab("Soft Threshold (power)") + ylab("Scale Free Topology Model Fit,signed R^2") + ggtitle("Scale independence")
+# ggplot(res, aes(Power, -sign(slope)*SFT.R.sq, color=paste(Cohort, Dataset, sep=' - '))) + geom_point() + theme_bw() + theme(aspect.ratio=1,  plot.title = element_text(hjust = 0.5)) + xlab("Soft Threshold (power)") + ylab("Scale Free Topology Model Fit,signed R^2") + ggtitle("Scale independence")
 
@@
 
-dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize, deepSplit=2)
+# dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize, deepSplit=2)
+dynamicMods = cutree(hcl, k=200)
 
-dynamicColors.Sex = labels2colors( dynamicMods )
+# find k so that max size of cluster is
+# k = sapply( seq(100, 1000, by=10), function(k) max(table(cutree(hcl, k=k))))
+
+dynamicMods = cutree(hcl, k=200)
+
+dynamicColors.Sex = labels2colors( dynamicMods, colorSeq=colors() )
 tab = table(dynamicColors.Sex)
@@
 # Hypothesis testing         
-df_test.Sex = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Sex, METADATA, variable, .3, 20)
+df_test.Sex = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Sex, METADATA, variable, useSLED=TRUE, .5, 20)
 ```
@@
 
+<!---
+
 # Disease network
@@
 
-res = pickSFT( diff.net.Dx )
+# res = pickSFT( diff.net.Dx )
 
-ggplot(res, aes(Power, -sign(slope)*SFT.R.sq, color=paste(Cohort, Dataset, sep=' - '))) + geom_point() + theme_bw() + theme(aspect.ratio=1,  plot.title = element_text(hjust = 0.5)) + xlab("Soft Threshold (power)") + ylab("Scale Free Topology Model Fit,signed R^2") + ggtitle("Scale independence")
+# ggplot(res, aes(Power, -sign(slope)*SFT.R.sq, color=paste(Cohort, Dataset, sep=' - '))) + geom_point() + theme_bw() + theme(aspect.ratio=1,  plot.title = element_text(hjust = 0.5)) + xlab("Soft Threshold (power)") + ylab("Scale Free Topology Model Fit,signed R^2") + ggtitle("Scale independence")
 
@@
 
-dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize, deepSplit=2)
+# dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize, deepSplit=2)
+dynamicMods = cutree(hcl, k=200)
 
@@
 
+--->
+
 # Interaction analysis
@@
 
-res = pickSFT( net.Dx_Sex )
+# res = pickSFT( net.Dx_Sex )
 
-ggplot(res, aes(Power, -sign(slope)*SFT.R.sq, color=paste(Cohort, Dataset, sep=' - '))) + geom_point() + theme_bw() + theme(aspect.ratio=1,  plot.title = element_text(hjust = 0.5)) + xlab("Soft Threshold (power)") + ylab("Scale Free Topology Model Fit,signed R^2") + ggtitle("Scale independence")
+# ggplot(res, aes(Power, -sign(slope)*SFT.R.sq, color=paste(Cohort, Dataset, sep=' - '))) + geom_point() + theme_bw() + theme(aspect.ratio=1,  plot.title = element_text(hjust = 0.5)) + xlab("Soft Threshold (power)") + ylab("Scale Free Topology Model Fit,signed R^2") + ggtitle("Scale independence")
 

### 2020-09-21 common_functions/common_functions.R

@@ -723,7 +723,7 @@ sLED_perm = function(Y, info, variable, nperm, rho=1, sumabs.seq=.2){
 
 
 
-test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicColors, METADATA, variable, sumabs.seq, nperm){
+test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicColors, METADATA, variable, useSLED=FALSE, sumabs.seq, nperm){
 
   col.array = unique(dynamicColors)
 
@@ -734,7 +734,7 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
 
     res = lapply( names(resid.lst), function(key){
 
-      message(key, ' ', col, ' ', match(col, col.array), '/', length(col.array))
+      if(runif(1) < .1) message(key, ' ', col, ' ', match(col, col.array), '/', length(col.array))
 
       # extact expression residuals for genes in this cluster
       Y = t(resid.lst[[key]][geneid,])
@@ -743,19 +743,24 @@ test_differential_correlation = function(resid.lst, C.diff.discovery, dynamicCol
       i = match(rownames(Y), rownames(METADATA))
       info = METADATA[i,]
 
-      # eval statistical hypothesis
-      # res = boxM_permute( Y, info[[variable]])
+      if( ! useSLED ){
+        # eval statistical hypothesis
+        res = boxM_permute( Y, info[[variable]])
 
-      lvl = levels(info[[variable]])
-      Y1 = Y[info[[variable]] == lvl[1],]
-      Y2 = Y[info[[variable]] == lvl[2],]
+        res.sled = list(p = res$p.value, n.active=ncol(Y))
 
-      # res_sLED = sLED_adapt( scale(Y1), scale(Y2), npermute=c(500,50000), BPPARAM=BPPARAM)
+      }else{
+        lvl = levels(info[[variable]])
+        Y1 = Y[info[[variable]] == lvl[1],]
+        Y2 = Y[info[[variable]] == lvl[2],]
 
-      # D = cora(Y2) - cora(Y1)
-      # p.sled = max(0, sLED:::sLEDTestStat(D, rho=1, sumabs.seq = .5)$stats)
+        # res_sLED = sLED_adapt( scale(Y1), scale(Y2), npermute=c(500,50000), BPPARAM=BPPARAM)
 
[…]

### 2020-09-21 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 beta_perm = list()
-beta_perm[['MSSM-Penn-Pitt']] = beta.mle(p)
+beta_perm[['MSSM-Penn-Pitt']] = beta.mle(p + 1e-10)
 
@@
 
-p = df_test$P.Value[df_test$Cohort== "NIMH-HBCC"]
+p2 = df_test$P.Value[df_test$Cohort== "NIMH-HBCC"]
 data.frame("# DE" = sum(p.adjust(p, 'fdr') < 0.05) / length(p), check.names=FALSE) %>% kable
-hist(p, main="P-values under the null in the replication cohort", xlab="p-value")
+hist(p2, main="P-values under the null in the replication cohort", xlab="p-value")
 
-beta_perm[['NIMH-HBCC']] = beta.mle(p)
+beta_perm[['NIMH-HBCC']] = beta.mle(p2 + 1e-10)
 ```

### 2020-09-22 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 
-df_test_meta = module_meta_analsis( df_test_recal )
+df_test_meta = module_meta_analysis( df_test_recal )
 
@@
 
-<!---
+# <!---
 
-# Disease network
-## Difference networks
-```{r diff.networks.Dx}
+# # Disease network
+# ## Difference networks
+# ```{r diff.networks.Dx}
 
-# Compute difference networks
-##############################
+# # Compute difference networks
+# ##############################
 
-resVersion = "Dx"
-variable = 'Dx'
+# resVersion = "Dx"
+# variable = 'Dx'
 
-diff.net.Dx = get_difference_network(resid.add[[resVersion]], METADATA, variable)
+# diff.net.Dx = get_difference_network(resid.add[[resVersion]], METADATA, variable)
 
-# res = pickSFT( diff.net.Dx )
+# # res = pickSFT( diff.net.Dx )
 
-# ggplot(res, aes(Power, -sign(slope)*SFT.R.sq, color=paste(Cohort, Dataset, sep=' - '))) + geom_point() + theme_bw() + theme(aspect.ratio=1,  plot.title = element_text(hjust = 0.5)) + xlab("Soft Threshold (power)") + ylab("Scale Free Topology Model Fit,signed R^2") + ggtitle("Scale independence")
+# # ggplot(res, aes(Power, -sign(slope)*SFT.R.sq, color=paste(Cohort, Dataset, sep=' - '))) + geom_point() + theme_bw() + theme(aspect.ratio=1,  plot.title = element_text(hjust = 0.5)) + xlab("Soft Threshold (power)") + ylab("Scale Free Topology Model Fit,signed R^2") + ggtitle("Scale independence")
 
-# Discover clusters in first cohort
-#----------------------------------
+# # Discover clusters in first cohort
+# #----------------------------------
 
-C = diff.net.Dx[[1]]
-C.diff.discovery = C[[2]]^pwr - C[[1]]^pwr
-C.diff.discovery = C.diff.discovery[keep.ensID,keep.ensID]
+# C = diff.net.Dx[[1]]
+# C.diff.discovery = C[[2]]^pwr - C[[1]]^pwr
+# C.diff.discovery = C.diff.discovery[keep.ensID,keep.ensID]
 
-rm(diff.net.Dx) # saving this to disk is slow, so just delete it
+# rm(diff.net.Dx) # saving this to disk is slow, so just delete it
 
-hcl = hclust( as.dist(C.diff.discovery) )
+# hcl = hclust( as.dist(C.diff.discovery) )
 
-# dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize, deepSplit=2)
-dynamicMods = cutree(hcl, k=200)
+# # dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize, deepSplit=2)
+# dynamicMods = cutree(hcl, k=200)
 
-dynamicColors.Dx = labels2colors(dynamicMods )
-tab = table(dynamicColors)
-hist(tab, xlab="Cluster size")
-```
+# dynamicColors.Dx = labels2colors(dynamicMods )
+# tab = table(dynamicColors)
+# hist(tab, xlab="Cluster size")
+# ```
 
-```{r testing.Dx}
-# Hypothesis testing     
-df_test.Dx = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Dx, METADATA, variable, useSLED=FALSE)  
-```
+# ```{r testing.Dx}
+# # Hypothesis testing     
+# df_test.Dx = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Dx, METADATA, variable, useSLED=FALSE)  
+# ```
 
-```{r recalibrate.Dx}
-table(df_test.Dx$FDR < 0.05)
+# ```{r recalibrate.Dx}
+# table(df_test.Dx$FDR < 0.05)
 
-df_test_recal = apply_permutation_null( df_test.Dx )
+# df_test_recal = apply_permutation_null( df_test.Dx )
 
-table(df_test_recal$FDR < 0.05)
+# table(df_test_recal$FDR < 0.05)
 
-df_test_meta = module_meta_analysis( df_test_recal )
+# df_test_meta = module_meta_analysis( df_test_recal )
 
-table(df_test_meta$FDR < 0.05)
-```
+# table(df_test_meta$FDR < 0.05)
+# ```
 
-## Discovery cohort
-```{r plot.discovery.Dx}
-# plot discovery and testing.  
-ggplot(df_test.Dx, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test.Dx$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
-```
+# ## Discovery cohort
+# ```{r plot.discovery.Dx}
+# # plot discovery and testing.  
+# ggplot(df_test.Dx, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test.Dx$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
+# ```
 
-```{r modules.in.discovery.Dx, fig.width=10, fig.height=50}
+# ```{r modules.in.discovery.Dx, fig.width=10, fig.height=50}
 
-figList = lapply(df_test$Module[1:5], function(mod){
+# figList = lapply(df_test$Module[1:5], function(mod){
 
-	plot_module( mod, df_test.Dx, METADATA, dynamicColors, C.diff.discovery, resid.add[[resVersion]], variable, "MSSM-Penn-Pitt")   
-	})
+# 	plot_module( mod, df_test.Dx, METADATA, dynamicColors, C.diff.discovery, resid.add[[resVersion]], variable, "MSSM-Penn-Pitt")   
+# 	})
 
-plot_grid( plotlist = figList, ncol=1)
-```
+# plot_grid( plotlist = figList, ncol=1)
+# ```
 
-## Replication cohort
-```{r second.cohort.Dx, fig.width=6, fig.height=12}
-df_cohort2 = df_test.Dx[df_test.Dx$Cohort == "NIMH-HBCC",]
-df_cohort2$Module = factor(df_cohort2$Module, unique(df_cohort2$Module))
-table(df_cohort2$FDR < 0.05)
+# ## Replication cohort
+# ```{r second.cohort.Dx, fig.width=6, fig.height=12}
+# df_cohort2 = df_test.Dx[df_test.Dx$Cohort == "NIMH-HBCC",]
+# df_cohort2$Module = factor(df_cohort2$Module, unique(df_cohort2$Module))
+# table(df_cohort2$FDR < 0.05)
 
-df = df_cohort2[df_cohort2$P.Value < 0.1,]
+# df = df_cohort2[df_cohort2$P.Value < 0.1,]
 
-ggplot(df, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
-```
+# ggplot(df, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
+# ```
 
-```{r modules.in.second.cohort.Dx, fig.width=8, fig.height=100}
+# ```{r modules.in.second.cohort.Dx, fig.width=8, fig.height=100}
 
-i = which(df_cohort2$FDR < 0.05)[1:10]
+# i = which(df_cohort2$FDR < 0.05)[1:10]
 
-figList = lapply(df_cohort2$Module[i], function(mod){
+# figList = lapply(df_cohort2$Module[i], function(mod){
 
-	plot_module( mod, df_test.Dx, METADATA, dynamicColors.Dx, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC")   
-	})
+# 	plot_module( mod, df_test.Dx, METADATA, dynamicColors.Dx, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC")   
+# 	})
 
-plot_grid( plotlist = figList, ncol=1)
-```
+# plot_grid( plotlist = figList, ncol=1)
+# ```
 
-## Module enrichment
-```{r top.enrich.Dx}
-df_module.Dx = data.frame(	ENSEMBL = rownames(C.diff.discovery) %>% trim_ensembl_ids, 
-						Module = dynamicColors.Dx)
+# ## Module enrichment
+# ```{r top.enrich.Dx}
+# df_module.Dx = data.frame(	ENSEMBL = rownames(C.diff.discovery) %>% trim_ensembl_ids, 
+# 						Module = dynamicColors.Dx)
 
-df_module_enrich.Dx = enrich.test( df_module.Dx, go.gs, df_cohort2$Module[i], 150)
-```
+# df_module_enrich.Dx = enrich.test( df_module.Dx, go.gs, df_cohort2$Module[i], 150)
+# ```
 
-```{r plot.enrich.Dx, fig.width=9, fig.height=25}
-figList = lapply( df_cohort2$Module[i], function(module){
-	plot_enrich(df_module_enrich.Dx, module, 8) + ggtitle(module)
-	})    
-plot_grid(plotlist = figList, ncol=1)
-```
+# ```{r plot.enrich.Dx, fig.width=9, fig.height=25}
+# figList = lapply( df_cohort2$Module[i], function(module){
+# 	plot_enrich(df_module_enrich.Dx, module, 8) + ggtitle(module)
+# 	})    
+# plot_grid(plotlist = figList, ncol=1)
+# ```
 
-## DE enrichment
-```{r cameraPR.Dx, fig.width=7, fig.height=20}
-res_enrich_DE = enrich_module_DE( df_module.Dx )
+# ## DE enrichment
+# ```{r cameraPR.Dx, fig.width=7, fig.height=20}
+# res_enrich_DE = enrich_module_DE( df_module.Dx )
 
-figList = lapply( df_cohort2$Module[i], function(module){
+# figList = lapply( df_cohort2$Module[i], function(module){
 
-	res = res_enrich_DE[res_enrich_DE$Geneset == module,]
+# 	res = res_enrich_DE[res_enrich_DE$Geneset == module,]
 
-	lim = max(-log10(res$PValue))
+# 	lim = max(-log10(res$PValue))
 
-	ggplot(res, aes(Test, -log10(PValue))) + geom_bar(stat="identity", fill=res$Geneset[1]) + coord_flip() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5))  + scale_y_continuous(expand=c(0,0), lim=c(0, lim*1.05)) + ggtitle(module) + geom_hline(yintercept=-log10(0.05), color="red", linetype="dashed")
-}) 
-plot_grid(plotlist = figList, ncol=1)
-```
+# 	ggplot(res, aes(Test, -log10(PValue))) + geom_bar(stat="identity", fill=res$Geneset[1]) + coord_flip() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5))  + scale_y_continuous(expand=c(0,0), lim=c(0, lim*1.05)) + ggtitle(module) + geom_hline(yintercept=-log10(0.05), color="red", linetype="dashed")
+# }) 
+# plot_grid(plotlist = figList, ncol=1)
+# ```
 
---->
+# --->
 

### 2020-09-22 common_functions/common_functions.R

@@ -1013,7 +1013,7 @@ plot_corr_network = function( C, zcutoff = 1.3, seed=1, base_size=11){
   set.seed(seed)
 
   # plot
-  ggraph(net, layout = "igraph", algorithm='graphopt') + 
+  ggraph(net, layout = "stress") + #igraph", algorithm='graphopt') + 
     geom_edge_link(aes(width = abs(weight), color=weight), alpha = 1) + 
     scale_edge_width(name='|Corr Diff|',range = c(0.2, 2)) +
     geom_node_point(size=base_size, color="grey80") +

### 2020-09-22 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 
-```{r recalibrate.Sex}
-table(df_test.Sex$FDR < 0.05)
-
-df_test_recal = apply_permutation_null( df_test.Sex )
-
-table(df_test_recal$FDR < 0.05)
-
-df_test_meta = module_meta_analysis( df_test_recal )
-
-table(df_test_meta$FDR < 0.05)
-```
-
 ## Discovery cohort
@@
 ggplot(df_test.Sex, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test.Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
+```
+
+```{r recalibrate.Sex}
+table(df_test.Sex$FDR < 0.05)
+
+df_test_recal.Sex = apply_permutation_null( df_test.Sex )
+
+table(df_test_recal.Sex$FDR < 0.05)
+
+df_test_meta.Sex = module_meta_analysis( df_test_recal.Sex )
+
+table(df_test_meta.Sex$FDR < 0.05)
+
+# plot discovery and testing.      
+ggplot(df_test_recal.Sex, aes(Module, -log10(P.Value))) + geom_bar(stat='identity', fill=df_test.Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~P)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
+```
+
+```{r plot.meta.Sex}
+ggplot(df_test_meta.Sex, aes(Module, -log10(P.Value))) + geom_bar(stat='identity', fill=df_test.Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~P)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
+
+qqPlot(df_test_meta$P.Value)
 ```
@@
 ## Module enrichment
-```{r top.enrich.Sex}
+```{r top.enrich.Sex, fig.width=7, fig.height=12}
 df_module.Sex = data.frame(	ENSEMBL = rownames(C.diff.discovery) %>% trim_ensembl_ids, 
@@
 ## DE enrichment
-```{r cameraPR.Sex}
+```{r cameraPR.Sex, fig.width=7, fig.height=12}
 res_enrich_DE = enrich_module_DE( df_module.Sex )
@@
 
-df_test_recal = apply_permutation_null( df_test.Dx_Sex )
+df_test_recal.Dx_Sex = apply_permutation_null( df_test.Dx_Sex )
 
-table(df_test_recal$FDR < 0.05)
+table(df_test_recal.Dx_Sex$FDR < 0.05)
 
-df_test_meta = module_meta_analysis( df_test_recal )
+df_test_meta.Dx_Sex = module_meta_analysis( df_test_recal.Dx_Sex )
 
-table(df_test_meta$FDR < 0.051)
+table(df_test_meta.Dx_Sex$FDR < 0.051)
 ```
@@
 ggplot(df, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
+```
+
+```{r plot.meta.Dx_Sex}
+ggplot(df_test_meta.Dx_Sex, aes(Module, -log10(P.Value))) + geom_bar(stat='identity', fill=df_test.Dx_Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~P)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
+
+qqPlot(df_test_meta.Dx_Sex$P.Value)
 ```
@@
 ## DE enrichment
-```{r cameraPR.Dx_Sex}
+```{r cameraPR.Dx_Sex, fig.width=7, fig.height=12}
 res_enrich_DE = enrich_module_DE( df_module.Dx_Sex )

### 2020-09-22 common_functions/common_functions.R

@@ -1104,7 +1104,7 @@ plot_enrich = function( df, module, base_size=11 ){
   df2$Geneset = factor(df2$Geneset, df2$Geneset)
 
   ylim = max(-log10(df$FDR))
-  ggplot(df2[1:10,], aes(Geneset, -log10(FDR))) + geom_bar(stat='identity', fill=module) + theme_bw(base_size) + theme(plot.title = element_text(hjust = 0.5)) + coord_flip() + ylab(bquote(-log[10]~FDR)) + xlab('') + scale_y_continuous(expand=c(0,0), lim=c(0, ylim*1.05)) 
+  ggplot(df2[1:10,], aes(Geneset, -log10(FDR))) + geom_bar(stat='identity', fill=module) + theme_bw(base_size) + theme(aspect.ratio=1,plot.title = element_text(hjust = 0.5)) + coord_flip() + ylab(bquote(-log[10]~FDR)) + xlab('') + scale_y_continuous(expand=c(0,0), lim=c(0, ylim*1.05)) 
 }
 
 

### 2020-09-22 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 ```{r plot.meta.Sex}
-ggplot(df_test_meta.Sex, aes(Module, -log10(P.Value))) + geom_bar(stat='identity', fill=df_test.Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~P)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
+ggplot(df_test_meta.Sex, aes(Module, -log10(P.Value))) + geom_bar(stat='identity', fill=df_test.Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + ylab(bquote(-log[10]~P)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
 
-qqPlot(df_test_meta$P.Value)
+GWASTools::qqPlot(df_test_meta.Sex$P.Value)
 ```
@@
 
-qqPlot(df_test_meta.Dx_Sex$P.Value)
+GWASTools::qqPlot(df_test_meta.Dx_Sex$P.Value)
 ```

### 2020-09-22 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 ```{r plot.meta.Sex}
-ggplot(df_test_meta.Sex, aes(Module, -log10(P.Value))) + geom_bar(stat='identity', fill=df_test.Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + ylab(bquote(-log[10]~P)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
+ggplot(df_test_meta.Sex, aes(Module, -log10(P.Value))) + geom_bar(stat='identity', fill=df_test_meta.Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + ylab(bquote(-log[10]~P)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
 

### 2020-09-22 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 ```{r plot.meta.Dx_Sex}
-ggplot(df_test_meta.Dx_Sex, aes(Module, -log10(P.Value))) + geom_bar(stat='identity', fill=df_test.Dx_Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~P)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
+ggplot(df_test_meta.Dx_Sex, aes(Module, -log10(P.Value))) + geom_bar(stat='identity', fill=df_test.Dx_Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + ylab(bquote(-log[10]~P)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
 

### 2020-09-22 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 ```{r plot.meta.Dx_Sex}
-ggplot(df_test_meta.Dx_Sex, aes(Module, -log10(P.Value))) + geom_bar(stat='identity', fill=df_test.Dx_Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + ylab(bquote(-log[10]~P)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
+ggplot(df_test_meta.Dx_Sex, aes(Module, -log10(P.Value))) + geom_bar(stat='identity', fill=df_test_meta.Dx_Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + ylab(bquote(-log[10]~P)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
 

### 2020-09-22 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 									P.value = `P meta`)]
+
+df_schema2$chisq = qchisq(df_schema2$P.value, 1, lower.tail=FALSE)
+df_schema2 = df_schema2[!is.na(df_schema2$P.value),]
+
+rownames(df_schema2) = df_schema2$ENSEMBL
 ```
@@
 
-```{r plot.enrich.Dx_Sex, fig.width=9}
+```{r plot.enrich.Dx_Sex, fig.width=7, fig.height=12}
 figList = lapply( df_cohort2$Module[i], function(module){

### 2020-09-22 common_functions/common_functions.R

@@ -1103,8 +1103,8 @@ plot_enrich = function( df, module, base_size=11 ){
   df2 = df[df$Module == module,]
   df2$Geneset = factor(df2$Geneset, df2$Geneset)
 
-  ylim = max(-log10(df$FDR))
-  ggplot(df2[1:10,], aes(Geneset, -log10(FDR))) + geom_bar(stat='identity', fill=module) + theme_bw(base_size) + theme(aspect.ratio=1,plot.title = element_text(hjust = 0.5)) + coord_flip() + ylab(bquote(-log[10]~FDR)) + xlab('') + scale_y_continuous(expand=c(0,0), lim=c(0, ylim*1.05)) 
+  ylim = max(-log10(df$p.value))
+  ggplot(df2[1:10,], aes(Geneset, -log10(p.value))) + geom_bar(stat='identity', fill=module) + theme_bw(base_size) + theme(aspect.ratio=1,plot.title = element_text(hjust = 0.5)) + coord_flip() + ylab(bquote(-log[10]~P)) + xlab('') + scale_y_continuous(expand=c(0,0), lim=c(0, ylim*1.05)) 
 }
 
 
@@ -1119,18 +1119,25 @@ enrich_module_DE = function(df_module){
     })
   names(modules.gs) = unique(df_module$Module)
 
-  keys = c(names(df_meta), names(df_meta.inter_sex.disease))
+  keys = c(names(df_meta), names(df_meta.inter_sex.disease), "SCHEMA")
 
   res_enrich_DE = lapply( keys, function(key){
 
-    df = df_meta[[key]]
-
-    if( is.null(df) ) df = df_meta.inter_sex.disease[[key]]
+    if( key == "SCHEMA"){
+      df = df_schema2
+      df$stat = df$chisq
+    }else{
+      df = df_meta[[key]]
+      if( is.null(df) ) df = df_meta.inter_sex.disease[[key]]
+      df$stat = with(df, (logFC/se)^2)
+    }   
     
     idx = match( trim_ensembl_ids(rownames(df)), df_module$ENSEMBL)
     modules.list = limma::ids2indices( modules.gs, df_module$ENSEMBL[idx])
          
-    res = cameraPR( with(df, logFC/se), modules.list )
+    res = cameraPR( df$stat, modules.list )
+    res$PValue = res$PValue / 2
+    res$FDR = c()
     res$Geneset = rownames(res)
     rownames(res) = c()
     res$Test = key

### 2020-09-22 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 rownames(df_schema2) = df_schema2$ENSEMBL
+
+# ASD
+df_asd = fread(synGet('syn22343811')$path)
+colnames(df_asd)[colnames(df_asd)=="Q‑Val" ] = 'Q'
+df_asd2 = data.frame(ENSEMBL = df_asd$Gene, Q = df_asd$Q)
+df_asd2$chisq = qchisq(df_asd2$Q + 1e-15, 1, lower.tail=FALSE)
+rownames(df_asd2) = df_asd2$ENSEMBL
+
 ```

### 2020-09-22 common_functions/common_functions.R

@@ -1119,12 +1119,15 @@ enrich_module_DE = function(df_module){
     })
   names(modules.gs) = unique(df_module$Module)
 
-  keys = c(names(df_meta), names(df_meta.inter_sex.disease), "SCHEMA")
+  keys = c(names(df_meta), names(df_meta.inter_sex.disease), "SCHEMA", "ASD")
 
   res_enrich_DE = lapply( keys, function(key){
 
     if( key == "SCHEMA"){
       df = df_schema2
+      df$stat = df$chisq
+    }else if( key == "ASD"){
+      df = df_asd2
       df$stat = df$chisq
     }else{
       df = df_meta[[key]]

### 2020-09-22 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 
-```{r modules.in.second.cohort, fig.width=8, fig.height=25}
+```{r modules.in.second.cohort, fig.width=8, fig.height=50}
 
@@
 
-	ggplot(res, aes(Test, -log10(PValue))) + geom_bar(stat="identity", fill=res$Geneset[1]) + coord_flip() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5))  + scale_y_continuous(expand=c(0,0), lim=c(0, lim*1.05)) + ggtitle(module) + geom_hline(yintercept=-log10(0.05), color="red", linetype="dashed")
+	ggplot(res, aes(Test, -log10(PValue))) + geom_bar(stat="identity", fill=res$Geneset[1]) + coord_flip() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5))  + scale_y_continuous(expand=c(0,0), lim=c(0, lim*1.05)) + ggtitle(module) + geom_hline(yintercept=-log10(0.05), color="red", linetype="dashed") + ylab(bquote(-log[10]~P))
 }) 
@@
 
-```{r modules.in.discovery.Dx_Sex, fig.width=10, fig.height=50}
+```{r modules.in.discovery.Dx_Sex, fig.width=8, fig.height=50}
 
@@
 # i = which(df_cohort2$FDR < 0.1) 
-i = 1:5
+i = 1:2
 

### 2020-09-22 common_functions/common_functions.R
Update diff networks

@@ -882,7 +882,7 @@ test_differential_correlation_interaction = function(resid.lst, C.diff.discovery
 
 
 # Plots for each module
-plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.discovery, resid.lst, variable, key, base_size=11, lvlidx = 1:2){
+plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.discovery, resid.lst, variable, key, base_size=11, lvlidx = 1:2, zcutoff=1.3){
 
   clusterID = as.character(clusterID)
 
@@ -972,7 +972,7 @@ plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.disc
 
   # plot network   
   main = paste(key, clusterID, paste0('(', lvl[lvlidx[2]], ' - ', lvl[lvlidx[1]], ')'), 'FDR =', format(FDR, digits=3))
-  fig3 = plot_corr_network( C2 - C1 , base_size=base_size) + ggtitle(main)
+  fig3 = plot_corr_network( C2 - C1 , base_size=base_size, zcutoff=zcutoff) + ggtitle(main)
 
   fig_merge = plot_grid( fig1, fig2, nrow=1)
   plot_grid( fig_merge, fig3, nrow=2, rel_heights=c(1,1.5) )    

### 2020-09-22 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 	# Hypothesis testing         
-	df_test = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, variable, useSLED=TRUE, .5, 20)
+	df_test = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, variable, useSLED=TRUE, .5, 1000)
 
@@
 # Hypothesis testing         
-df_test_null = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, variable, useSLED=TRUE, .5, 20) 
+df_test_null = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, variable, useSLED=TRUE, .5, 1000) 
 ```
@@
 # Hypothesis testing         
-df_test.Sex = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Sex, METADATA, variable, useSLED=TRUE, .5, 20)
+df_test.Sex = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Sex, METADATA, variable, useSLED=TRUE, .5, 1000)
 ```
@@
 # Hypothesis testing     
-df_test.Dx_Sex = test_differential_correlation_interaction(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Dx_Sex, METADATA, nperm=20)    
+df_test.Dx_Sex = test_differential_correlation_interaction(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Dx_Sex, METADATA, nperm=1000)    
 ```

### 2020-09-22 combined_analysis/diff.networks.Rmd
Update diff networks

--- 2020-09-22 combined_analysis/diff.networks.Rmd
@@
 	# Hypothesis testing         
-	df_test = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, variable, useSLED=TRUE, .5, 20)
+	df_test = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, variable, useSLED=TRUE, .5, 100)
 
@@
 # Hypothesis testing         
-df_test_null = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, variable, useSLED=TRUE, .5, 20) 
+df_test_null = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, variable, useSLED=TRUE, .5, 100) 
 ```
@@
 # Hypothesis testing         
-df_test.Sex = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Sex, METADATA, variable, useSLED=TRUE, .5, 20)
+df_test.Sex = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Sex, METADATA, variable, useSLED=TRUE, .5, 100)
 ```
@@
 # Hypothesis testing     
-df_test.Dx_Sex = test_differential_correlation_interaction(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Dx_Sex, METADATA, nperm=20)    
+df_test.Dx_Sex = test_differential_correlation_interaction(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Dx_Sex, METADATA, nperm=100)    
 ```

### 2020-09-22 combined_analysis/combine_all_DLPFC.Rmd
freeze

@@
 params:
-  upload: TRUE
+  upload: FALSE
 ---
@@
   tidy = FALSE,
+  dev = c('png', 'pdf'),
   cache = TRUE)
+
+colorArray = colors()[grep("grey|gray", colors(), invert=TRUE)]
+
 ```
@@
 
-form = ~ ageOfDeath:Institution + Reported_Gender + Dx + Institution + Dx
+form = ~ ageOfDeath:Institution + Reported_Gender + Dx + Institution
 
@@
 
-vp = fitExtractVarPartModel( t(Y), form, METADATA, BPPARAM=SerialParam())
-plotPercentBars(vp) + ggtitle("Explain cell type composition")
-
-ymax = max(rowSums(vp[,-ncol(vp)]))*110
-plotPercentBars(vp) + scale_y_continuous(expand=c(0,0), limits=c(0, ymax))
-```
-
-```{r sex.graph, fig.width=7, fig.height=7}
-library(ggplot2)
-library(ggraph)
-library(tidygraph)
-
+df = lapply( colnames(Y), function(id){
+
+  fit = lm(Y[,id] ~ ageOfDeath:Institution + Reported_Gender + Dx + Institution, METADATA)
+
+  data.frame( Test     = rownames(anova(fit))[1:4],
+              CellType = id, 
+              p        = anova(fit)[,'Pr(>F)'][1:4], 
+              frac     =  calcVarPart(fit)[1:4])
+})
+df = data.table(do.call(rbind, df))
+
+df$Test[df$Test=='Reported_Gender'] = "Sex"
+df$Test[df$Test=='ageOfDeath:Institution'] = "Age of death"
+df$p.adj = p.adjust(df$p, "bonferroni")
+
+ymax = max(df[,sum(frac),by="CellType"]$V1)*105
+ggplot(df, aes(CellType, frac*100, fill=Test)) + geom_bar(stat="identity") + coord_flip() + theme_bw() + theme(aspect.ratio=1, panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank()) +  ylab("Variance explained (%)") + theme(axis.line = element_line(colour = "transparent"), 
+        axis.line.x = element_line(colour = "black"), panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(), panel.border = element_blank(), 
+        panel.background = element_blank(), axis.ticks.y = element_blank(), 
+        legend.key = element_blank(), plot.margin = unit(c(0, 
+            0.3, 0, 0.8), "cm")) + scale_y_continuous(expand=c(0,0), limits=c(0, ymax)) + xlab('')
+
+ggplot(df, aes(CellType, Test, fill=-log10(pmax(1e-16,p)), label=ifelse(p.adj<0.05, '*', ''))) + geom_tile() + theme_bw() + scale_fill_gradient(name = bquote(-log[10]~P), low="white", high="red", limit=c(0, NA)) + theme(aspect.ratio=1, panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank()) + geom_text() + coord_flip()
+```
+
+```{r sex.graph, fig.width=7, fig.height=7, dev="png"}
 df_cca2 = df_cca[with(df_cca, v1!=v2),]
@@
 
+```{r plot.gene_set_test.marginal, eval=TRUE}
+ plot_zenith_results = function( res, coef, Resource, FDR = 0.05){
+
+  i = which((res$Resource == Resource) & (res$Coef == coef))
+  df = res[i,]
+
+  # filter by FDR
+  df = df[df$FDR <= FDR,]
+  df$Coef = c()
+  df$se = c()
+  df$Resource = c()
+
+  df = df[order(df$PValue, decreasing=TRUE),]
+  df$Set = factor(df$Set, df$Set)
+  ymax = max(-log10(df$PValue)) * 1.05
+  ggplot(df, aes(Set, -log10(PValue))) + geom_bar(stat="identity", fill="navy") + theme_bw() + theme(aspect.ratio=1,plot.title = element_text(hjust = 0.5)) + coord_flip() + ylab(bquote(-log[10]~P)) + xlab('') + ggtitle(coef) + scale_y_continuous(expand=c(0,0), limits=c(0, ymax))
+}
+
+ plot_zenith_results( res_meta_zenith, "DxSCZ", "GO", FDR=0.05)
+```
+
+```{r, fig.width=9}
+ plot_zenith_results( res_meta_zenith, "DxSCZ", "MSigDB", FDR=0.05)
+```
+
 # Sex Analysis
@@
 
-## Plot top examples
+## Plot top examples - not chrY
+```{r plot.forrest.dlpfc.Sex.autosomes, fig.width=8, fig.height=20}  
+Symbols = df_meta[[coef]][df_meta[[coef]]$Chrom !='Y',]$Symbol[1:10]
+figList = lapply(Symbols, function(geneSymbol){
+
+  fig1 = plot_forrest( c(tab.de.lst[[coef]], list("Meta-analysis" = df_meta[[coef]])), geneSymbol)
+
+  fig2 = plot_gene( residuals.lst, df_meta[[coef]], geneSymbol, "Reported_Gender", fill=c("red", "blue")) + xlab("Sex")
+
+  plot_grid(fig1, fig2, nrow=1)
+})
+
+plot_grid(plotlist=figList, ncol=1)
+```
+
+## Plot top examples - autosomes
 ```{r plot.forrest.dlpfc.Sex, fig.width=8, fig.height=20}  
-figList = lapply(df_meta[[coef]]$Symbol[1:10], function(geneSymbol){
+Symbols = df_meta[[coef]][df_meta[[coef]]$Chrom %in% 1:22,]$Symbol[1:10]
+
+figList = lapply(Symbols, function(geneSymbol){
 
@@
 
+```{r plot.gene_set_test.marginal.Sex, eval=TRUE}
+ plot_zenith_results( res_meta_zenith, "Reported_GenderMale", "GO", FDR=0.05)         
+ plot_zenith_results( res_meta_zenith, "Reported_GenderMale", "MSigDB", FDR=0.05)        
+```
+
 ### Only autosomes
@@
 show_zenith_results( res_meta_zenith.sex_autosome, "Reported_GenderMale", "MSigDB", FDR=0.1)   
+```
+
+```{r plot.gene_set_test.marginal.Sex.autosome, eval=TRUE}
+ plot_zenith_results( res_meta_zenith.sex_autosome, "Reported_GenderMale", "GO", FDR=0.1) 
+ plot_zenith_results( res_meta_zenith.sex_autosome, "Reported_GenderMale", "MSigDB", FDR=0.1)   
 ```
@@
 
+```{r plot.gene_set_test.marginal.Age, eval=TRUE}
+ plot_zenith_results( res_meta_zenith, "ageOfDeath", "GO", FDR=0.01)   
+ plot_zenith_results( res_meta_zenith, "ageOfDeath", "MSigDB", FDR=0.01)    
+```
+
 # Sex by Disease interaction
@@
 
-# Dx test with in Males and then Females
+```{r plot.gene_set_test.marginal.inter_sex.disease, eval=TRUE}
+coef = 'DxSCZ:Reported_GenderMale'   
+ plot_zenith_results( res_meta_zenith.inter_sex.disease, coef, "GO", FDR=0.1)  
+ plot_zenith_results( res_meta_zenith.inter_sex.disease, coef, "MSigDB", FDR=0.1)
+
+ plot_zenith_results( res_meta_zenith.inter_sex.disease_autosome, coef, "GO", FDR=0.1) 
+ plot_zenith_results( res_meta_zenith.inter_sex.disease_autosome, coef, "MSigDB", FDR=0.1)    
+```
+
+# Dx test separately in Males and then Females
+```{r dream.sex.disease.separately}
+form.lst = list(
+  `MSSM-Penn-Pitt`  = ~ Dx + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
+  `NIMH-HBCC`       = ~ Dx + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)
+
+# DE for males and females separately
+fit.separ = lapply( vobj.lst, function(vobj){
+
+  fit.lst = lapply(unique(METADATA$Reported_Gender), function(Sex){
+
+    i = match(colnames(vobj), rownames(METADATA))
+    info = METADATA[i,]
+    idx = which(info$Reported_Gender == Sex)
+
+    form_use = form.lst[[info$Cohort[1]]]
+
+    fit = dream(vobj[,idx], form_use, info[idx,], BPPARAM = SnowParam(18))  
+    eBayes( fit )
+  })
+  names(fit.lst) = unique(METADATA$Reported_Gender)
+  fit.lst
+  })
+names(fit.separ) = names(vobj.lst)
+
+# meta-analysis
+df_meta.separ = lapply(unique(METADATA$Reported_Gender), function(Sex){
+  tab.lst = lapply( names(vobj.lst), function(chrt){
+    topTable(fit.separ[[chrt]][[Sex]], coef="DxSCZ", sort.by='none', number=Inf)
+  })
+
+  res = run_meta_analysis( tab.lst, "FE") %>% getGeneSymbol
+  res$t = with(res, logFC/se)
+  res
+})
+names(df_meta.separ) = unique(METADATA$Reported_Gender)
+```
+
+```{r separ.concordance.scz, fig.height=6, fig.width=10}
+# Concordance analysis
+resConcord.separ = plot_concordance( df_meta.separ, showGenes = df_meta[['DxSCZ']]$Symbol[1:20])
+
+main = "Concordance of SCZ across sexes"
+plot_grid(  resConcord.separ$fig_logFC + ggtitle(main),
+            resConcord.separ$fig_tstat + ggtitle(main), nrow=1)
+
+data.frame(Cohort = names(df_meta.separ),
+          "# DE"  = resConcord.separ$n.de,
+          pi1     = resConcord.separ$pi_single,
+          check.names=FALSE) %>% 
+          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)
+
+resConcord.separ$tab_pi %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
+resConcord.separ$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
+```
+
+```{r compare.separate, fig.width=5, fig.height=5}
+df = merge( df_meta.separ[[1]], df_meta.separ[[2]], by="row.names")
+
+fit = with(df, lm(logFC.y ~ logFC.x))
+
+coef(summary(fit) ) %>%  kable(caption="logFC is larger in males") %>% kable_styling(full_width=FALSE)
+
+ # compare logFC
+tstat = with(df, (logFC.y - logFC.x) / (sqrt(se.y^2 + se.x^2)))
+
+qqt(tstat)
+abline(0, 1, col="red")
+p = 2*pt(abs(tstat), df=778, lower.tail=FALSE)
+table(p.adjust(p) < 0.05)
+```
+
+## Plot top examples
+```{r plot.forrest.dlpfc.separ, fig.width=8, fig.height=40}
+
+figList = lapply(df_meta.separ[[1]]$Symbol[1:20], function(geneSymbol){
+  fig1 = plot_forrest( df_meta.separ, geneSymbol, plot2='p')
+
+  plot_grid(fig1, nrow=1)
+})
+
+plot_grid(plotlist=figList, ncol=1)
+```
 

### 2020-09-22 common_functions/common_functions.R
Update diff networks

@@ -929,6 +929,10 @@ plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.disc
              paste0('(',lvl[2], ' - ',lvl[1], ')'))
     lvlidx = 1:2
 
+    res = sLED:::sLEDTestStat(C2-C1, rho=1, sumabs.seq = .5)
+
+    df_leverage = data.frame(Gene = colnames(C2), leverage = t(res$leverage))
+
   }else{
     stop(length(lvlidx))
   }
@@ -936,6 +940,12 @@ plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.disc
   hcl = hclust(as.dist(1 - C1), method="ward.D2")
   C1 = C1[hcl$order,hcl$order]
   C2 = C2[hcl$order,hcl$order]
+
+  # reorder based on clustering
+  df_leverage$Gene = factor(df_leverage$Gene, rownames(C2))
+
+  ylim = max(df_leverage$leverage)*1.05
+  fig.leverage = ggplot(df_leverage, aes(Gene, leverage)) + geom_bar(stat="identity", fill="navy") + theme_bw() + theme(aspect.ratio=5) + coord_flip() + theme(panel.grid.major.y = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + xlab("") + ylab("Leverage") + scale_y_continuous(limits=c(0, ylim), expand=c(0,0))
 
   # create data.frame to compare correlation values
   ind <- which( upper.tri(C1,diag=FALSE) , arr.ind = TRUE )
@@ -969,6 +979,8 @@ plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.disc
             limits = c(-1, 1), na.value = "grey") + theme_bw(base_size) + 
         theme(aspect.ratio = 1, plot.title = element_text(hjust = 0.5), 
             legend.position = "bottom", panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.text.x=element_text(angle=45, hjust=1)) + ggtitle(main) + ylab(lvl[lvlidx[2]]) + xlab(lvl[lvlidx[1]])
+
+   fig2 = plot_grid(fig2, fig.leverage, ncol=2)     
 
   # plot network   
   main = paste(key, clusterID, paste0('(', lvl[lvlidx[2]], ' - ', lvl[lvlidx[1]], ')'), 'FDR =', format(FDR, digits=3))

### 2020-09-22 common_functions/common_functions.R
Update diff networks

@@ -929,13 +929,13 @@ plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.disc
              paste0('(',lvl[2], ' - ',lvl[1], ')'))
     lvlidx = 1:2
 
-    res = sLED:::sLEDTestStat(C2-C1, rho=1, sumabs.seq = .5)
-
-    df_leverage = data.frame(Gene = colnames(C2), leverage = t(res$leverage))
-
   }else{
     stop(length(lvlidx))
   }
+
+  res = sLED:::sLEDTestStat(C2-C1, rho=1, sumabs.seq = .5)
+  df_leverage = data.frame(Gene = colnames(C2), leverage = t(res$leverage))
+
   # reorder for clustering
   hcl = hclust(as.dist(1 - C1), method="ward.D2")
   C1 = C1[hcl$order,hcl$order]

### 2020-09-22 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 rownames(df_asd2) = df_asd2$ENSEMBL
-
 ```
@@
 
-	plot_module( mod, df_test.Dx_Sex, METADATA, dynamicColors.Dx_Sex, C.diff.discovery, resid.add[[resVersion]], variable, "MSSM-Penn-Pitt")   
+	plot_module( mod, /, METADATA, dynamicColors.Dx_Sex, C.diff.discovery, resid.add[[resVersion]], variable, "MSSM-Penn-Pitt")   
 	})
@@
 
-```{r modules.in.second.cohort.Dx_Sex, fig.width=8, fig.height=80}
+```{r modules.in.second.cohort.Dx_Sex, fig.width=12, fig.height=40}
 
 # i = which(df_cohort2$FDR < 0.1) 
-i = 1:2
+i = 1
 

### 2020-09-22 common_functions/common_functions.R

@@ -934,7 +934,8 @@ plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.disc
   }
 
   res = sLED:::sLEDTestStat(C2-C1, rho=1, sumabs.seq = .5)
-  df_leverage = data.frame(Gene = colnames(C2), leverage = t(res$leverage))
+  df_leverage = data.frame(Gene = colnames(C2), leverage = t(res$leverage), stringsAsFactors=FALSE)
+  genes.has.leverage = df_leverage$Gene[df_leverage$leverage > 0]
 
   # reorder for clustering
   hcl = hclust(as.dist(1 - C1), method="ward.D2")
@@ -980,11 +981,14 @@ plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.disc
         theme(aspect.ratio = 1, plot.title = element_text(hjust = 0.5), 
             legend.position = "bottom", panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.text.x=element_text(angle=45, hjust=1)) + ggtitle(main) + ylab(lvl[lvlidx[2]]) + xlab(lvl[lvlidx[1]])
 
-   fig2 = plot_grid(fig2, fig.leverage, ncol=2)     
+  fig2 = plot_grid(fig2 + theme(legend.position="left", aspect.ratio=1), fig.leverage, ncol=2, align='v', rel_widths=c(4, 1), axis='t')      
 
   # plot network   
+  D = C2 - C1
   main = paste(key, clusterID, paste0('(', lvl[lvlidx[2]], ' - ', lvl[lvlidx[1]], ')'), 'FDR =', format(FDR, digits=3))
-  fig3 = plot_corr_network( C2 - C1 , base_size=base_size, zcutoff=zcutoff) + ggtitle(main)
+
+  D = D[genes.has.leverage,genes.has.leverage]
+  fig3 = plot_corr_network( D , base_size=base_size, zcutoff=zcutoff) + ggtitle(main)
 
   fig_merge = plot_grid( fig1, fig2, nrow=1)
   plot_grid( fig_merge, fig3, nrow=2, rel_heights=c(1,1.5) )    

### 2020-09-22 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 
-```{r plot.meta.Dx_Sex}
+```{r plot.meta.Dx_Sex, fig.width=5, fig.height=9}
 ggplot(df_test_meta.Dx_Sex, aes(Module, -log10(P.Value))) + geom_bar(stat='identity', fill=df_test_meta.Dx_Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + ylab(bquote(-log[10]~P)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")

### 2020-09-22 common_functions/common_functions.R

@@ -981,7 +981,7 @@ plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.disc
         theme(aspect.ratio = 1, plot.title = element_text(hjust = 0.5), 
             legend.position = "bottom", panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.text.x=element_text(angle=45, hjust=1)) + ggtitle(main) + ylab(lvl[lvlidx[2]]) + xlab(lvl[lvlidx[1]])
 
-  fig2 = plot_grid(fig2 + theme(legend.position="left", aspect.ratio=1), fig.leverage, ncol=2, align='v', rel_widths=c(4, 1), axis='t')      
+  # fig2 = plot_grid(fig2 + theme(legend.position="left", aspect.ratio=1), fig.leverage, ncol=2, align='v', rel_widths=c(4, 1), axis='t')      
 
   # plot network   
   D = C2 - C1
@@ -990,8 +990,9 @@ plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.disc
   D = D[genes.has.leverage,genes.has.leverage]
   fig3 = plot_corr_network( D , base_size=base_size, zcutoff=zcutoff) + ggtitle(main)
 
-  fig_merge = plot_grid( fig1, fig2, nrow=1)
-  plot_grid( fig_merge, fig3, nrow=2, rel_heights=c(1,1.5) )    
+  # fig_merge = plot_grid( fig1, fig2, nrow=1)
+  # plot_grid( fig_merge, fig3, nrow=2, rel_heights=c(1,1.5) )  
+  list(fig1, fig2, fig3, fig.leverage)  
 }
 
 # plot_module( mod, df_test, METADATA, dynamicColors, C.diff.discovery, resid.add[[resVersion]], variable, "MSSM-Penn-Pitt", 5) 

### 2020-09-22 common_functions/common_functions.R
Update diff networks

@@ -902,7 +902,7 @@ plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.disc
   info = METADATA[i,]
 
   # perform hypothesis test
-  res = boxM_permute( Y, info[[variable]])
+  # res = boxM_permute( Y, info[[variable]])
 
   FDR = df_test[with(df_test, Cohort==key & Module == clusterID),'FDR']
 

### 2020-09-22 common_functions/common_functions.R
Update diff networks

@@ -992,7 +992,7 @@ plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.disc
 
   # fig_merge = plot_grid( fig1, fig2, nrow=1)
   # plot_grid( fig_merge, fig3, nrow=2, rel_heights=c(1,1.5) )  
-  list(fig1, fig2, fig3, fig.leverage)  
+  list(fig_cor = fig1, fig_heatmap = fig2, fig_network = fig3, fig_leverage = fig.leverage)  
 }
 
 # plot_module( mod, df_test, METADATA, dynamicColors, C.diff.discovery, resid.add[[resVersion]], variable, "MSSM-Penn-Pitt", 5) 

### 2020-09-23 combined_analysis/diff.networks.Rmd
Update diff networks

--- 2020-09-22 combined_analysis/diff.networks.Rmd
@@
     toc_float: true
+upload: FALSE
 ---
@@
 library(plyr) 
+library(githubr)
 library(qvalue)  
@@
 
+```{r synapse.parameters, include=FALSE, cache=TRUE}
+
+parentId = 'syn22416298';
+activityName = 'Differential network analysis - DLPFC';
+activityDescription = 'Differential network analysis - DLPFC';
+
+thisFileName <- 'diff.networks.Rmd'
+
+# Github link
+thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
+thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('combined_analysis/',thisFileName))
+```
+
 ```{r load.data}
@@
 METADATA = readRDS(synGet('syn22416411')$path)
+ALL_USED_IDs = c('syn22416489', 'syn22416446', 'syn22416411')
+
 METADATA$Reported_Gender = factor(METADATA$Reported_Gender, c("Male", "Female"))
@@
 df_schema = fread(synGet('syn22343823')$path)
+ALL_USED_IDs = c(ALL_USED_IDs, 'syn22343823')
 
@@
 df_asd = fread(synGet('syn22343811')$path)
+ALL_USED_IDs = c(ALL_USED_IDs, 'syn22343811')
 colnames(df_asd)[colnames(df_asd)=="Q‑Val" ] = 'Q'
@@
 df_meta.inter_sex.disease = readRDS(synGet('syn22416532')$path)
+
+ALL_USED_IDs = c(ALL_USED_IDs, c('syn22416500', 'syn22416524', 'syn22416508', 'syn22416532'))
 ```
@@
 
-rm(diff.net) # saving this to disk is slow, so just delete it
+# rm(diff.net) # saving this to disk is slow, so just delete it
 
@@
 table(df_test_meta$FDR < 0.05)
+```
+
+```{r save.network.null}
+if( params$upload ){    
+  # Code
+  # Make sure these match the folder name and parent Id in the first chunk.
+  CODE <- Folder(name = "Joint analysis of DLPFC count data", parentId = "syn22416298")
+  CODE <- synStore(CODE)
+
+  # Store data 
+  save_network_results( diff.net, hcl, dynamicColors, df_test_null, "nullTest")
+}
 ```
@@
 
-rm(diff.net.Sex) # saving this to disk is slow, so just delete it
+# rm(diff.net.Sex) # saving this to disk is slow, so just delete it
 
@@
 plot_grid(plotlist = figList, ncol=1)
+```
+
+```{r save.network.Sex}
+if( params$upload ){    
+  # Code
+  # Make sure these match the folder name and parent Id in the first chunk.
+  CODE <- Folder(name = "Joint analysis of DLPFC count data", parentId = "syn22416298")
+  CODE <- synStore(CODE)
+
+  # Store data 
+  save_network_results( diff.net.Sex, hcl, dynamicColors.Sex, df_test.Sex, "SexTest")
+}
 ```
@@
 
-rm(net.Dx_Sex) # saving this to disk is slow, so just delete it
+# rm(net.Dx_Sex) # saving this to disk is slow, so just delete it
 
@@
 
-```{r plot.meta.Dx_Sex}
+```{r plot.meta.Dx_Sex, fig.width=5, fig.height}
 ggplot(df_test_meta.Dx_Sex, aes(Module, -log10(P.Value))) + geom_bar(stat='identity', fill=df_test_meta.Dx_Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + ylab(bquote(-log[10]~P)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
@@
 
-	plot_grid(fig1, fig2, fig3, ncol=1)
+	plot_grid(fig1$fig_cor, fig2$fig_cor, fig3$fig_cor, nrow=1)
+	plot_grid(fig1$fig_heatmap, fig2$fig_heatmap, fig3$fig_heatmap, nrow=1)
+	plot_grid(fig1$fig_network, fig2$fig_network, fig3$fig_network, nrow=1)
+	plot_grid(fig1$fig_cor, fig2$fig_cor, fig3$fig_cor, nrow=1)
 	})
@@
 plot_grid(plotlist = figList, ncol=1)
+```
+
+```{r save.network.Dx_Sex}
+if( params$upload ){    
+  # Code
+  # Make sure these match the folder name and parent Id in the first chunk.
+  CODE <- Folder(name = "Joint analysis of DLPFC count data", parentId = "syn22416298")
+  CODE <- synStore(CODE)
+
+  # Store data 
+  save_network_results( net.Dx_Sex, hcl, dynamicColors.Dx_Sex, df_test.Dx_Sex, "Dx_SexTest")
+}
 ```

### 2020-09-23 common_functions/common_functions.R

@@ -881,8 +881,7 @@ test_differential_correlation_interaction = function(resid.lst, C.diff.discovery
 
 
 
-# Plots for each module
-plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.discovery, resid.lst, variable, key, base_size=11, lvlidx = 1:2, zcutoff=1.3){
+get_module_results = function( clusterID, df_test, METADATA, dynamicColors, C.diff.discovery, resid.lst, variable, key, base_size=11, lvlidx = 1:2, zcutoff=1.3){
 
   clusterID = as.character(clusterID)
 
@@ -900,13 +899,6 @@ plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.disc
   # extract metadata
   i = match(rownames(Y), rownames(METADATA))
   info = METADATA[i,]
-
-  # perform hypothesis test
-  # res = boxM_permute( Y, info[[variable]])
-
-  FDR = df_test[with(df_test, Cohort==key & Module == clusterID),'FDR']
-
-  main = paste(key, clusterID, 'FDR =', format(FDR, digits=3))
 
   # Evaluate correlation matrix for each category
   lvl = levels(info[[variable]])
@@ -945,10 +937,14 @@ plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.disc
   # reorder based on clustering
   df_leverage$Gene = factor(df_leverage$Gene, rownames(C2))
 
-  ylim = max(df_leverage$leverage)*1.05
-  fig.leverage = ggplot(df_leverage, aes(Gene, leverage)) + geom_bar(stat="identity", fill="navy") + theme_bw() + theme(aspect.ratio=5) + coord_flip() + theme(panel.grid.major.y = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + xlab("") + ylab("Leverage") + scale_y_continuous(limits=c(0, ylim), expand=c(0,0))
+  list( C1          = C1,
+        C2          = C2,
+        D           = C2 - C1,
+        df_leverage = df_leverage)
+}
 
-  # create data.frame to compare correlation values
+cor_pair_to_data.frame = function(C1, C2){
+   # create data.frame to compare correlation values
[…]

### 2020-09-23 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 
-```{r plot.meta.Dx_Sex, fig.width=5, fig.height}
+```{r plot.meta.Dx_Sex, fig.width=5, fig.height=8}
 ggplot(df_test_meta.Dx_Sex, aes(Module, -log10(P.Value))) + geom_bar(stat='identity', fill=df_test_meta.Dx_Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + ylab(bquote(-log[10]~P)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")

### 2020-09-23 combined_analysis/diff.networks.Rmd
Update diff networks

@@
     toc_float: true
-upload: FALSE
+upload: TRUE
 ---

### 2020-09-23 combined_analysis/plot_diff.networks.Rmd
Initial commit

--- 2020-09-23 combined_analysis/diff.networks.Rmd
@@
 title: "CommonMind Consortium: DLPFC"
-subtitle: "Analysis of read count data by cohort"
+subtitle: "Plots for Differential network analysis - DLPFC"
 author:
@@
 cd /hpc/users/hoffmg01/work/covarr-de
-# rm -rf combined_analysis/diff.networks_* combined_analysis/diff.networks.html
 ml git pandoc udunits proj gdal geos
@@
 
-rmarkdown::render("combined_analysis/diff.networks.Rmd")
+rmarkdown::render("combined_analysis/plot_diff.networks.Rmd")
 
 --->
-
-Since we are generating networks across cohorts and sexes, identifiying conserved clusters using standard methods becomes non-trivial.
-Instead of defining clusters for each subset and evaluating the difference, we first compute the differences in the correlation networks, then identify a single set of clusters based on this difference-network.
-Then use hypothesis test of differential correlation in the second cohort to get p-value.
-
-But if concensus models are constructed to be differential by definition, then these clusters could show differential correlation even under the null.
-Therefore, use set A as discovery for M/F differential network, and replicated in set B.
 
@@
 
-```{r synapse.parameters, include=FALSE, cache=TRUE}
+```{r synapse.parameters, include=FALSE, cache=TRUE, eval=FALSE}
 
 parentId = 'syn22416298';
-activityName = 'Differential network analysis - DLPFC';
-activityDescription = 'Differential network analysis - DLPFC';
+activityName = 'Plots for Differential network analysis - DLPFC';
+activityDescription = 'Plots for Differential network analysis - DLPFC';
 
@@
 METADATA$Reported_Gender = factor(METADATA$Reported_Gender, c("Male", "Female"))
-
-# identify genes to keep: remove chrY genes
-geneEnsID = lapply(resid.add, function(x){
-	lapply(x, rownames)
-	})
-geneEnsID = unique(unlist(geneEnsID))
-
-df_genes = data.frame(ENSEMBL = geneEnsID, stringsAsFactors=FALSE) %>% getGeneSymbol('ENSEMBL')
-# table(df_genes$Chrom)
-
-keep.ensID = df_genes$ENSEMBL[df_genes$Chrom %in% c(1:22, 'X', 'MT')]
-
-minClusterSize = 50
-pwr = 7
-```
-
-```{r load.go}
-# Load Gene Ontology database 
-go.gs = get_GeneOntology()
-```
-
-```{r load.WES}
-df_schema = fread(synGet('syn22343823')$path)
-ALL_USED_IDs = c(ALL_USED_IDs, 'syn22343823')
-
-df_schema2 = df_schema[,data.frame(ENSEMBL = Gene,
-									OR1 = `OR (Class I)`,
-									OR2 = `OR (Class II)`,
-									P.value = `P meta`)]
-
-df_schema2$chisq = qchisq(df_schema2$P.value, 1, lower.tail=FALSE)
-df_schema2 = df_schema2[!is.na(df_schema2$P.value),]
-
-rownames(df_schema2) = df_schema2$ENSEMBL
-
-# ASD
-df_asd = fread(synGet('syn22343811')$path)
-ALL_USED_IDs = c(ALL_USED_IDs, 'syn22343811')
-colnames(df_asd)[colnames(df_asd)=="Q‑Val" ] = 'Q'
-df_asd2 = data.frame(ENSEMBL = df_asd$Gene, Q = df_asd$Q)
-df_asd2$chisq = qchisq(df_asd2$Q + 1e-15, 1, lower.tail=FALSE)
-rownames(df_asd2) = df_asd2$ENSEMBL
 ```
@@
 ```
-
-# Null permuations
-```{r diff.networks.perm}
-
-# Compute difference networks for multiple random permutations
-##############################
-
-resVersion = "Sex"
-
-df_test = lapply( 1:10, function(k){
-	METADATA$rnd = sample(c('A','B'), nrow(METADATA), replace=TRUE)
-	METADATA$rnd = factor(METADATA$rnd, c('A','B'))
-
-	variable = 'rnd'
-
-	diff.net = get_difference_network(resid.add[[resVersion]], METADATA, variable)
-
-	pwr = 7
-
-	# Discover clusters in first cohort
-	#----------------------------------
-	C = diff.net[[1]]
-	C.diff.discovery = C[[2]]^pwr - C[[1]]^pwr
-	C.diff.discovery = C.diff.discovery[keep.ensID,keep.ensID]
-
-	rm(diff.net) # saving this to disk is slow, so just delete it
-
-	hcl = hclust( as.dist(C.diff.discovery) )
-
-	# dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize, deepSplit=2)
-	dynamicMods = cutree(hcl, k=200)
-
-	dynamicColors = labels2colors(dynamicMods)
-	tab = table(dynamicColors)
-
-	# hist(tab, xlab="Cluster size")
-
-	# Hypothesis testing         
-	df_test = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, variable, useSLED=TRUE, .5, 100)
-
-	df_test
-})
-df_test = do.call(rbind, df_test)
-```
-
-```{r perm.results, fig.width=5}
-p = df_test$P.Value[df_test$Cohort=="MSSM-Penn-Pitt"]
-# print(1 - qvalue(p)$pi0)
-hist(p, main="P-values under the null in the discovery cohort", xlab="p-value")
-
-beta_perm = list()
-beta_perm[['MSSM-Penn-Pitt']] = beta.mle(p + 1e-10)
-
-calc_p = function(p, fit){
-	pbeta(p, 	shape1=fit$param['alpha'], 
-				shape2=fit$param['beta'], 
-				lower.tail=TRUE)
-}
-
-p2 = df_test$P.Value[df_test$Cohort== "NIMH-HBCC"]
-data.frame("# DE" = sum(p.adjust(p, 'fdr') < 0.05) / length(p), check.names=FALSE) %>% kable
-hist(p2, main="P-values under the null in the replication cohort", xlab="p-value")
-
-beta_perm[['NIMH-HBCC']] = beta.mle(p2 + 1e-10)
-```
-
-# Null network
-```{r diff.networks.null}
-
-# Compute difference networks for 1 random permutation
-######################################################
-
-resVersion = "Sex"
-
-METADATA$rnd = sample(c('A','B'), nrow(METADATA), replace=TRUE)
-METADATA$rnd = factor(METADATA$rnd, c('A','B'))
-
-variable = 'rnd'
-
-diff.net = get_difference_network(resid.add[[resVersion]], METADATA, variable)
-
-# Discover clusters in first cohort
-#----------------------------------
-C = diff.net[[1]]
-C.diff.discovery = C[[2]]^pwr - C[[1]]^pwr
-C.diff.discovery = C.diff.discovery[keep.ensID,keep.ensID]
-
-# rm(diff.net) # saving this to disk is slow, so just delete it
-
-hcl = hclust( as.dist(C.diff.discovery) )
-
-# dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize, deepSplit=2)
-dynamicMods = cutree(hcl, k=200)
-
-dynamicColors = labels2colors(dynamicMods, colorSeq=colors())
-tab = table(dynamicColors)
-
-# Hypothesis testing         
-df_test_null = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, variable, useSLED=TRUE, .5, 100) 
-```
-
-```{r recalibrate.null}
-df_test_recal = apply_permutation_null( df_test_null )
-
-table(df_test_recal$FDR < 0.05)
-
-df_test_meta = module_meta_analysis( df_test_recal )
-
-table(df_test_meta$FDR < 0.05)
-```
-
-```{r save.network.null}
-if( params$upload ){    
-  # Code
-  # Make sure these match the folder name and parent Id in the first chunk.
-  CODE <- Folder(name = "Joint analysis of DLPFC count data", parentId = "syn22416298")
-  CODE <- synStore(CODE)
-
-  # Store data 
-  save_network_results( diff.net, hcl, dynamicColors, df_test_null, "nullTest")
-}
-```
-
-# Sex network
-## Difference networks
-```{r diff.networks}
-
-# Compute difference networks
-##############################
-
-resVersion = "Sex"
-variable = 'Reported_Gender'
-
-diff.net.Sex = get_difference_network(resid.add[[resVersion]], METADATA, variable)
-
-# res = pickSFT( diff.net.Sex )
-
-# ggplot(res, aes(Power, -sign(slope)*SFT.R.sq, color=paste(Cohort, Dataset, sep=' - '))) + geom_point() + theme_bw() + theme(aspect.ratio=1,  plot.title = element_text(hjust = 0.5)) + xlab("Soft Threshold (power)") + ylab("Scale Free Topology Model Fit,signed R^2") + ggtitle("Scale independence")
-
-# Discover clusters in first cohort
-#----------------------------------
-C = diff.net.Sex[[1]]
-C.diff.discovery = C[[2]]^pwr - C[[1]]^pwr
-C.diff.discovery = C.diff.discovery[keep.ensID,keep.ensID]
-
-# rm(diff.net.Sex) # saving this to disk is slow, so just delete it
-
-hcl = hclust( as.dist(C.diff.discovery) )
-
-# dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize, deepSplit=2)
-dynamicMods = cutree(hcl, k=200)
-
-# find k so that max size of cluster is
-# k = sapply( seq(100, 1000, by=10), function(k) max(table(cutree(hcl, k=k))))
-
-dynamicMods = cutree(hcl, k=200)
-
-dynamicColors.Sex = labels2colors( dynamicMods, colorSeq=colors() )
-tab = table(dynamicColors.Sex)
-hist(tab, xlab="Cluster size")
-```
-
-```{r testing}
-# Hypothesis testing         
-df_test.Sex = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Sex, METADATA, variable, useSLED=TRUE, .5, 100)
-```
-
-## Discovery cohort
-```{r plot.discovery}
-# plot discovery and testing.      
-ggplot(df_test.Sex, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test.Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
-```
-
-```{r recalibrate.Sex}
-table(df_test.Sex$FDR < 0.05)
-
-df_test_recal.Sex = apply_permutation_null( df_test.Sex )
-
-table(df_test_recal.Sex$FDR < 0.05)
-
-df_test_meta.Sex = module_meta_analysis( df_test_recal.Sex )
-
-table(df_test_meta.Sex$FDR < 0.05)
-
-# plot discovery and testing.      
-ggplot(df_test_recal.Sex, aes(Module, -log10(P.Value))) + geom_bar(stat='identity', fill=df_test.Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~P)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
-```
-
-```{r plot.meta.Sex}
-ggplot(df_test_meta.Sex, aes(Module, -log10(P.Value))) + geom_bar(stat='identity', fill=df_test_meta.Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + ylab(bquote(-log[10]~P)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
-
-GWASTools::qqPlot(df_test_meta.Sex$P.Value)
-```
-
-```{r modules.in.discovery, fig.width=8, fig.height=40}
-
-figList = lapply(df_test.Sex$Module[1:5], function(mod){
-
-	plot_module( mod, df_test.Sex, METADATA, dynamicColors.Sex, C.diff.discovery, resid.add[[resVersion]], variable, "MSSM-Penn-Pitt", 5)   
-	})
-
-plot_grid( plotlist = figList, ncol=1)
-```
-
-## Replication cohort
-```{r second.cohort, fig.width=8, fig.height=12}
-df_cohort2 = df_test.Sex[df_test.Sex$Cohort == "NIMH-HBCC",]
-df_cohort2$Module = factor(df_cohort2$Module, unique(df_cohort2$Module))
-table(df_cohort2$FDR < 0.05)
-
-df = df_cohort2[df_cohort2$P.Value < 0.1,]
-
-ggplot(df, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
-```
-
-```{r modules.in.second.cohort, fig.width=8, fig.height=50}
-
-# i = which(df_cohort2$FDR < 0.05)
-i = 1:5
-
-figList = lapply(df_cohort2$Module[i], function(mod){
-
-	plot_module( mod, df_test.Sex, METADATA, dynamicColors.Sex, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC")   
-	})
-
-plot_grid( plotlist = figList, ncol=1)
-```
-
-## Module enrichment
-```{r top.enrich.Sex, fig.width=7, fig.height=12}
-df_module.Sex = data.frame(	ENSEMBL = rownames(C.diff.discovery) %>% trim_ensembl_ids, 
-						Module = dynamicColors.Sex)
-
-df_module_enrich.Sex = enrich.test( df_module.Sex, go.gs, df_cohort2$Module[i], 150) 
-```
-
-```{r plot.enrich.Sex, fig.width=9}
-figList = lapply( df_cohort2$Module[i], function(module){
-	plot_enrich(df_module_enrich.Sex, module, 8) + ggtitle(module)
-	}) 
-plot_grid(plotlist = figList, ncol=1)
-```
-
-## DE enrichment
-```{r cameraPR.Sex, fig.width=7, fig.height=12}
-res_enrich_DE = enrich_module_DE( df_module.Sex )
-
-figList = lapply( df_cohort2$Module[i], function(module){
-
-	res = res_enrich_DE[res_enrich_DE$Geneset == module,]
-
-	lim = max(-log10(res$PValue))
-
-	ggplot(res, aes(Test, -log10(PValue))) + geom_bar(stat="identity", fill=res$Geneset[1]) + coord_flip() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5))  + scale_y_continuous(expand=c(0,0), lim=c(0, lim*1.05)) + ggtitle(module) + geom_hline(yintercept=-log10(0.05), color="red", linetype="dashed") + ylab(bquote(-log[10]~P))
-}) 
-plot_grid(plotlist = figList, ncol=1)
-```
-
-```{r save.network.Sex}
-if( params$upload ){    
-  # Code
-  # Make sure these match the folder name and parent Id in the first chunk.
-  CODE <- Folder(name = "Joint analysis of DLPFC count data", parentId = "syn22416298")
-  CODE <- synStore(CODE)
-
-  # Store data 
-  save_network_results( diff.net.Sex, hcl, dynamicColors.Sex, df_test.Sex, "SexTest")
-}
-```
-
-# <!---
-
-# # Disease network
-# ## Difference networks
-# ```{r diff.networks.Dx}
-
-# # Compute difference networks
-# ##############################
-
-# resVersion = "Dx"
-# variable = 'Dx'
-
-# diff.net.Dx = get_difference_network(resid.add[[resVersion]], METADATA, variable)
-
-# # res = pickSFT( diff.net.Dx )
-
-# # ggplot(res, aes(Power, -sign(slope)*SFT.R.sq, color=paste(Cohort, Dataset, sep=' - '))) + geom_point() + theme_bw() + theme(aspect.ratio=1,  plot.title = element_text(hjust = 0.5)) + xlab("Soft Threshold (power)") + ylab("Scale Free Topology Model Fit,signed R^2") + ggtitle("Scale independence")
-
-# # Discover clusters in first cohort
-# #----------------------------------
-
-# C = diff.net.Dx[[1]]
-# C.diff.discovery = C[[2]]^pwr - C[[1]]^pwr
-# C.diff.discovery = C.diff.discovery[keep.ensID,keep.ensID]
-
-# rm(diff.net.Dx) # saving this to disk is slow, so just delete it
-
-# hcl = hclust( as.dist(C.diff.discovery) )
-
-# # dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize, deepSplit=2)
-# dynamicMods = cutree(hcl, k=200)
-
-# dynamicColors.Dx = labels2colors(dynamicMods )
-# tab = table(dynamicColors)
-# hist(tab, xlab="Cluster size")
-# ```
-
-# ```{r testing.Dx}
-# # Hypothesis testing     
-# df_test.Dx = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Dx, METADATA, variable, useSLED=FALSE)  
-# ```
-
-# ```{r recalibrate.Dx}
-# table(df_test.Dx$FDR < 0.05)
-
-# df_test_recal = apply_permutation_null( df_test.Dx )
-
-# table(df_test_recal$FDR < 0.05)
-
-# df_test_meta = module_meta_analysis( df_test_recal )
-
-# table(df_test_meta$FDR < 0.05)
-# ```
-
-# ## Discovery cohort
-# ```{r plot.discovery.Dx}
-# # plot discovery and testing.  
-# ggplot(df_test.Dx, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test.Dx$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
-# ```
-
-# ```{r modules.in.discovery.Dx, fig.width=10, fig.height=50}
-
-# figList = lapply(df_test$Module[1:5], function(mod){
-
-# 	plot_module( mod, df_test.Dx, METADATA, dynamicColors, C.diff.discovery, resid.add[[resVersion]], variable, "MSSM-Penn-Pitt")   
-# 	})
-
-# plot_grid( plotlist = figList, ncol=1)
-# ```
-
-# ## Replication cohort
-# ```{r second.cohort.Dx, fig.width=6, fig.height=12}
-# df_cohort2 = df_test.Dx[df_test.Dx$Cohort == "NIMH-HBCC",]
-# df_cohort2$Module = factor(df_cohort2$Module, unique(df_cohort2$Module))
-# table(df_cohort2$FDR < 0.05)
-
-# df = df_cohort2[df_cohort2$P.Value < 0.1,]
-
-# ggplot(df, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
-# ```
-
-# ```{r modules.in.second.cohort.Dx, fig.width=8, fig.height=100}
-
-# i = which(df_cohort2$FDR < 0.05)[1:10]
-
-# figList = lapply(df_cohort2$Module[i], function(mod){
-
-# 	plot_module( mod, df_test.Dx, METADATA, dynamicColors.Dx, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC")   
-# 	})
-
-# plot_grid( plotlist = figList, ncol=1)
-# ```
-
-# ## Module enrichment
-# ```{r top.enrich.Dx}
-# df_module.Dx = data.frame(	ENSEMBL = rownames(C.diff.discovery) %>% trim_ensembl_ids, 
-# 						Module = dynamicColors.Dx)
-
-# df_module_enrich.Dx = enrich.test( df_module.Dx, go.gs, df_cohort2$Module[i], 150)
-# ```
-
-# ```{r plot.enrich.Dx, fig.width=9, fig.height=25}
-# figList = lapply( df_cohort2$Module[i], function(module){
-# 	plot_enrich(df_module_enrich.Dx, module, 8) + ggtitle(module)
-# 	})    
-# plot_grid(plotlist = figList, ncol=1)
-# ```
-
-# ## DE enrichment
-# ```{r cameraPR.Dx, fig.width=7, fig.height=20}
-# res_enrich_DE = enrich_module_DE( df_module.Dx )
-
-# figList = lapply( df_cohort2$Module[i], function(module){
-
-# 	res = res_enrich_DE[res_enrich_DE$Geneset == module,]
-
-# 	lim = max(-log10(res$PValue))
-
-# 	ggplot(res, aes(Test, -log10(PValue))) + geom_bar(stat="identity", fill=res$Geneset[1]) + coord_flip() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5))  + scale_y_continuous(expand=c(0,0), lim=c(0, lim*1.05)) + ggtitle(module) + geom_hline(yintercept=-log10(0.05), color="red", linetype="dashed")
-# }) 
-# plot_grid(plotlist = figList, ncol=1)
-# ```
-
-# --->
-
-# Interaction analysis
-
-## Difference networks
-```{r diff.networks.Dx_Sex}
-
-# Compute difference networks
-##############################
-
-METADATA$Dx_Sex = with(METADATA, paste(Reported_Gender, Dx, sep='_'))
-METADATA$Dx_Sex = factor(METADATA$Dx_Sex, c('Female_Control', 'Female_SCZ', 'Male_Control', 'Male_SCZ'))
-
-resVersion = "Dx_Sex"
-variable = 'Dx_Sex'
-
-net.Dx_Sex = get_networks(resid.add[[resVersion]], METADATA, variable)
-
-# res = pickSFT( net.Dx_Sex )
-
-# ggplot(res, aes(Power, -sign(slope)*SFT.R.sq, color=paste(Cohort, Dataset, sep=' - '))) + geom_point() + theme_bw() + theme(aspect.ratio=1,  plot.title = element_text(hjust = 0.5)) + xlab("Soft Threshold (power)") + ylab("Scale Free Topology Model Fit,signed R^2") + ggtitle("Scale independence")
-
-pwr = 7
-
-lvls = levels(METADATA$Dx_Sex)
-
-print(paste0('(',lvls[4], ' - ',lvls[3], ') - (',lvls[2], ' - ',lvls[1], ')'))
-
-# Discover clusters in first cohort
-#----------------------------------
-C = net.Dx_Sex[[1]]
-C.diff.discovery = (C[[lvls[4]]]^pwr - C[[lvls[3]]]^pwr) - (C[[lvls[2]]]^pwr - C[[lvls[1]]]^pwr)
-
-C.diff.discovery = C.diff.discovery[keep.ensID,keep.ensID]
-
-# rm(net.Dx_Sex) # saving this to disk is slow, so just delete it
-
-hcl = hclust( as.dist(C.diff.discovery) )
-
-# dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize, deepSplit=2)
-dynamicMods = cutree(hcl, k=200)
-
-dynamicColors.Dx_Sex = labels2colors(dynamicMods)
-tab = table(dynamicColors.Dx_Sex)
-hist(tab, xlab="Cluster size")
-```
-
-```{r testing.Dx_Sex}
-# Hypothesis testing     
-df_test.Dx_Sex = test_differential_correlation_interaction(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Dx_Sex, METADATA, nperm=100)    
-```
-
-```{r recalibrate.Dx_Sex}
-table(df_test.Dx_Sex$FDR < 0.05)
-
-df_test_recal.Dx_Sex = apply_permutation_null( df_test.Dx_Sex )
-
-table(df_test_recal.Dx_Sex$FDR < 0.05)
-
-df_test_meta.Dx_Sex = module_meta_analysis( df_test_recal.Dx_Sex )
-
-table(df_test_meta.Dx_Sex$FDR < 0.051)
-```
-
-## Discovery cohort
-```{r plot.discovery.Dx_Sex}
-# plot discovery and testing.  
-ggplot(df_test.Dx_Sex, aes(Module, -log10(P.Value))) + geom_bar(stat='identity', fill=df_test.Dx_Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~P)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
-```
-
-```{r modules.in.discovery.Dx_Sex, fig.width=8, fig.height=50}
-
-figList = lapply(df_test.Dx_Sex$Module[1:5], function(mod){.
-
-	plot_module( mod, /, METADATA, dynamicColors.Dx_Sex, C.diff.discovery, resid.add[[resVersion]], variable, "MSSM-Penn-Pitt")   
-	})
-
-plot_grid( plotlist = figList, ncol=1)
-```
-
-## Replication cohort
-```{r second.cohort.Dx_Sex, fig.width=6, fig.height=12}
-df_cohort2 = df_test.Dx_Sex[df_test.Dx_Sex$Cohort == "NIMH-HBCC",]
-df_cohort2$Module = factor(df_cohort2$Module, unique(df_cohort2$Module))
-table(df_cohort2$FDR < 0.05)
-
-df = df_cohort2[df_cohort2$P.Value < 0.1,]
-
-ggplot(df, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
-```
-
-```{r plot.meta.Dx_Sex, fig.width=5, fig.height}
-ggplot(df_test_meta.Dx_Sex, aes(Module, -log10(P.Value))) + geom_bar(stat='identity', fill=df_test_meta.Dx_Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + ylab(bquote(-log[10]~P)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
-
-GWASTools::qqPlot(df_test_meta.Dx_Sex$P.Value)
-```
-
-```{r modules.in.second.cohort.Dx_Sex, fig.width=12, fig.height=40}
-
-# i = which(df_cohort2$FDR < 0.1) 
-i = 1
-
-figList = lapply(df_cohort2$Module[i], function(mod){
-
-	fig1 = plot_module( mod, df_test.Dx_Sex, METADATA, dynamicColors.Dx_Sex, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC", lvlidx=1:2)
-
-	fig2 = plot_module( mod, df_test.Dx_Sex, METADATA, dynamicColors.Dx_Sex, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC", lvlidx=3:4)
-
-	fig3 = plot_module( mod, df_test.Dx_Sex, METADATA, dynamicColors.Dx_Sex, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC", lvlidx=1:4)
-
-	plot_grid(fig1$fig_cor, fig2$fig_cor, fig3$fig_cor, nrow=1)
-	plot_grid(fig1$fig_heatmap, fig2$fig_heatmap, fig3$fig_heatmap, nrow=1)
-	plot_grid(fig1$fig_network, fig2$fig_network, fig3$fig_network, nrow=1)
-	plot_grid(fig1$fig_cor, fig2$fig_cor, fig3$fig_cor, nrow=1)
-	})
-
-plot_grid( plotlist = figList, ncol=1)
-```
-
-## Module enrichment
-```{r top.enrich.Dx_Sex}
-df_module.Dx_Sex = data.frame(	ENSEMBL = rownames(C.diff.discovery) %>% trim_ensembl_ids, 
-						Module = dynamicColors.Dx_Sex)
-
-df_module_enrich.Dx_Sex = enrich.test( df_module.Dx_Sex, go.gs, df_cohort2$Module[i], 150)
-```
-
-```{r plot.enrich.Dx_Sex, fig.width=7, fig.height=12}
-figList = lapply( df_cohort2$Module[i], function(module){
-	plot_enrich(df_module_enrich.Dx_Sex, module, 8) + ggtitle(module) 
-	}) 
-plot_grid(plotlist = figList, ncol=1)
-```
-
-## DE enrichment
-```{r cameraPR.Dx_Sex, fig.width=7, fig.height=12}
-res_enrich_DE = enrich_module_DE( df_module.Dx_Sex )
-
-figList = lapply( df_cohort2$Module[i], function(module){
-
-	res = res_enrich_DE[res_enrich_DE$Geneset == module,]
-
-	lim = max(-log10(res$PValue))
-
-	ggplot(res, aes(Test, -log10(PValue))) + geom_bar(stat="identity", fill=res$Geneset[1]) + coord_flip() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5))  + scale_y_continuous(expand=c(0,0), lim=c(0, lim*1.05)) + ggtitle(module) + geom_hline(yintercept=-log10(0.05), color="red", linetype="dashed") + ylab(bquote(-log[10]~P))
-}) 
-plot_grid(plotlist = figList, ncol=1)
-```
-
-```{r save.network.Dx_Sex}
-if( params$upload ){    
-  # Code
-  # Make sure these match the folder name and parent Id in the first chunk.
-  CODE <- Folder(name = "Joint analysis of DLPFC count data", parentId = "syn22416298")
-  CODE <- synStore(CODE)
-
-  # Store data 
-  save_network_results( net.Dx_Sex, hcl, dynamicColors.Dx_Sex, df_test.Dx_Sex, "Dx_SexTest")
-}
-```
-
-```{r exit, cache=FALSE}
-knitr::knit_exit()
-```
-
-<!---
-
-X = Rfast::matrnorm(10, 4)
-Y = Rfast::matrnorm(10, 4)
-
-C1 = cor(X)
-C2 = cor(Y)
-D = C1 - C2
-
-pdf("/Users/gabrielhoffman/Dropbox/Presentations/CMC3/CMC_DE_sex/figures/diffCor.pdf")
-variancePartition::plotCorrMatrix(C1, sort=FALSE, dendrogram="none")
-variancePartition::plotCorrMatrix(C2, sort=FALSE, dendrogram="none")
-variancePartition::plotCorrMatrix(D, sort=FALSE, dendrogram="none")
-dev.off()
-
-set.seed(1)
-X1 = Rfast::matrnorm(20, 4)
-X2 = Rfast::matrnorm(20, 4)
-X3 = Rfast::matrnorm(20, 4)
-X4 = Rfast::matrnorm(20, 4)
-
-C1 = cor(X1)
-C2 = cor(X2)
-C3 = cor(X3)
-C4 = cor(X4)
-Da = C1 - C2
-Db = C3 - C4
-
-pdf("/Users/gabrielhoffman/Dropbox/Presentations/CMC3/CMC_DE_sex/figures/diffCor2.pdf")
-variancePartition::plotCorrMatrix(C1, sort=FALSE, dendrogram="none")
-variancePartition::plotCorrMatrix(C2, sort=FALSE, dendrogram="none")
-variancePartition::plotCorrMatrix(C3, sort=FALSE, dendrogram="none")
-variancePartition::plotCorrMatrix(C4, sort=FALSE, dendrogram="none")
-variancePartition::plotCorrMatrix(Da-Db, sort=FALSE, dendrogram="none")
-dev.off()
-
-stat = sapply(1:1000, function(i){
-
-	set.seed(i)
-	n = 1000
-	p = 40
-	X = matrix(rnorm(n*p), n,p)
-	Y = matrix(rnorm(n*p), n,p)
-
-	C1 = cor(X)
-	C2 = cor(Y)
-	D = C1 - C2
-
-	stat = sLED:::sLEDTestStat(D, rho=1)$stats
-	max(0, stat)
-})
-
-fit = egamma( stat )
-
-fit$parameters
-
-# need to generate correct residuals
-
-GWAS hits, TWAS, SCHEMA
-ChrX and ChrY genes with autism and Scz
-
-DE within each sex, on chrX and chrY.
-	drop chrY genes for computing FDR?
-
-Can we predict sex given autosomal genes?
-
-PMID: 26892004
-https://science.sciencemag.org/content/369/6509/eaba3066
-https://www.nature.com/articles/s41576-018-0083-1
-
-test: how much sex differences are given by age, cell type and institution.
-
-Apparent differences in cell type composition can be explained by age.
-	After accounting for age, can cell type predict sex?
-
-library(mvtnorm)
-library(Rfast)
-
-X = matrnorm(300, 20)
-Y = matrnorm(300, 20)
-
-library(BiocParallel)
-res = decorate:::.sLED(X=X, Y=Y, npermute=100000, verbose=TRUE, BPPARAM = SerialParam())
-
-res$pVal
-
-C_x = cor(X)
-C_y = cor(Y)
-
-df_cor = data.frame( 	a = C_x[lower.tri(C_x)],
-						b = C_y[lower.tri(C_y)])
-
-with(df_cor, plot(a,b))
-
-a = calc_sde(id=1,points=df_cor);
-
-Sigma = cov(df_cor)
-
-X = matrnorm(1000, 30)
-
-C_x = cor(X)
-
-C = C_x[1:5, 1:5]
-colnames(C) = paste0('gene_', 1:nrow(C))
-rownames(C) = colnames(C)
-
-library(ggraph)
-library(ggplot2)
-library(tidygraph)
-library(Rfast)
-
-load("~/Downloads/C.test.RDATA")
-
-C = C1 - C2
-
-plot_corr_network( C1-C2 )
-
-rstat_nodes <- data.frame(name = c("H
-
-	adley", "David", "Romain", "Julia"))
-rstat_edges <- data.frame(from = c(1, 1, 1, 2, 3, 3, 4, 4, 4),
-                         to = c(2, 3, 4, 1, 1, 2, 1, 2, 3))
-rstat_edges$weight = 1:9
-
-rstat_nodes <- data.frame(name = c("Hadley", "David", "Romain", "Julia"))
-rstat_edges <- data.frame(from = c(1, 1, 1, 2, 3, 3, 4, 4, 4),
-                         to = c(2, 3, 4, 1, 1, 2, 1, 2, 3))
-rstat_edges$weight = rnorm(9)
-
-net = tbl_graph(nodes = rstat_nodes, edges = rstat_edges, directed=FALSE)
-
-ggraph(net, layout = "stress") + 
-  geom_node_point(size=10, color="blue") +
-  geom_edge_link(aes(width = weight), alpha = 0.8) + 
-  scale_edge_width(range = c(0.2, 2)) +
-  geom_node_text(aes(label = name), repel = TRUE) +
-  labs(edge_width = "Correlation") +
-  theme_graph()
-
-net = ggnetwork(C)
-
-set.edge.attribute
-set.edge.value
-
-ggplot(net, aes(x = x, y = y, xend = xend, yend = yend)) + theme_blank() + geom_nodes(aes(x,y)) + geom_edges()
-
-ggplot(
-ggnetwork(emon[[1]], layout = "kamadakawai", arrow.gap = 0.025),
-aes(x, y, xend = xend, yend = yend) )+
-geom_edges(aes(color = Frequency, size=Frequency^2),
-arrow = arrow(length = unit(10, "pt"), type = "closed")
-)+
-geom_nodes(aes(size = Formalization)) + scale_color_gradient(low = "grey50", high = "tomato") + scale_size_area(breaks = 1:3) +
-theme_blank()
-
-net = network(C)
-
-get.edges(net, "gene_1")
-
-rstat_nodes <- data.frame(name = c("Hadley", "David", "Romain", "Julia"))
-rstat_edges <- data.frame(from = c(1, 1, 1, 2, 3, 3, 4, 4, 4),
-                         to = c(2, 3, 4, 1, 1, 2, 1, 2, 3))
-rstat_edges$weight = 1:9
-
-net = tbl_graph(nodes = rstat_nodes, edges = rstat_edges, directed=FALSE)
-
-ggraph(net, layout = "auto") + 
-  geom_node_point(size=10, color="blue") +
-  geom_edge_link(aes(width = weight), alpha = 0.8) + 
-  scale_edge_width(range = c(0.2, 2)) +
-  geom_node_text(aes(label = name), repel = TRUE) +
-  labs(edge_width = "Correlation") +
-  theme_graph()
-
---->

### 2020-09-23 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 title: "CommonMind Consortium: DLPFC"
-subtitle: "Analysis of read count data by cohort"
+subtitle: "Differential network analysis - DLPFC"
 author:
@@
     toc_float: true
-upload: TRUE
+params:
+	upload: FALSE
 ---

### 2020-09-23 combined_analysis/diff.networks.Rmd
Update diff networks

--- 2020-09-23 combined_analysis/diff.networks.Rmd
@@
     toc_float: true
-upload: TRUE
 ---

### 2020-09-23 combined_analysis/diff.networks.Rmd
Update diff networks

--- 2020-09-23 combined_analysis/diff.networks.Rmd
@@
 
-rmarkdown::render("combined_analysis/diff.networks.Rmd")
+rmarkdown::render("combined_analysis/diff.networks.Rmd", output_dir="/sc/arion/scratch/hoffmg01/network")
 
@@
 
-df_test = lapply( 1:10, function(k){
+df_test = lapply( 1, function(k){
 	METADATA$rnd = sample(c('A','B'), nrow(METADATA), replace=TRUE)

### 2020-09-23 combined_analysis/diff.networks.Rmd
Update diff networks

@@
 
-cd /hpc/users/hoffmg01/work/covarr-de
-# rm -rf combined_analysis/diff.networks_* combined_analysis/diff.networks.html
+#
+
+cd /hpc/users/hoffmg01/work/covarr-de/
+# rm -rf combined_analysis/diff.networks_cache/html/* combined_analysis/diff.networks.html
 ml git pandoc udunits proj gdal geos
@@
 
-rmarkdown::render("combined_analysis/diff.networks.Rmd", output_dir="/sc/arion/scratch/hoffmg01/network")
+rmarkdown::render("combined_analysis/diff.networks.Rmd")
+
+ll /hpc/users/hoffmg01/work/covarr-de/combined_analysis/diff.networks_cache/html/
+
+cd /hpc/users/hoffmg01/work/covarr-de
+# rm -rf combined_analysis/diff.networks_* combined_analysis/diff.networks.html
+ml git pandoc udunits proj gdal geos
+git pull
+cd /sc/arion/scratch/hoffmg01/network/
+\cp -f /hpc/users/hoffmg01/work/covarr-de/combined_analysis/diff.networks.Rmd /sc/arion/scratch/hoffmg01/network
+R
+
+system("cd /hpc/users/hoffmg01/work/covarr-de; git pull")
+
+rmarkdown::render("diff.networks.Rmd", output_dir="/sc/arion/scratch/hoffmg01/network/", intermediates_dir='/sc/arion/scratch/hoffmg01/network/')
+
+ll /hpc/users/hoffmg01/work/covarr-de/combined_analysis/diff.networks_cache/html/
 

### 2020-09-23 common_functions/common_functions.R

@@ -986,6 +986,7 @@ plot_module = function( clusterID, df_test, METADATA, dynamicColors, C.diff.disc
   C1 = res$C1
   C2 = res$C2
  
+  lvl = levels(info[[variable]])
 
   ylim = max(df_leverage$leverage)*1.05
   fig.leverage = ggplot(df_leverage, aes(Gene, leverage)) + geom_bar(stat="identity", fill="navy") + theme_bw() + theme(aspect.ratio=5) + coord_flip() + theme(panel.grid.major.y = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + xlab("") + ylab("Leverage") + scale_y_continuous(limits=c(0, ylim), expand=c(0,0))

### 2020-10-06 combined_analysis/combine_all_DLPFC.Rmd
Biol Psych freeze

@@
 source("../common_functions/common_functions.R")   
+# source("common_functions/common_functions.R")
 
@@
 df_cca2 = df_cca[with(df_cca, v1!=v2),]
-
-rstat_nodes <- data.frame(name = unique(df_cca2$v1))
+df_cca2$key = apply(df_cca2, 1, function(x){
+  x = sort(c(x['v1'], x['v2']))
+  paste(x[1], x[2])
+  }) 
+df_cca2 = df_cca2[!duplicated(df_cca2$key),]
+
+rstat_nodes <- data.frame(name = unique(c(df_cca2$v1, df_cca2$v2)))
 
@@
   scale_edge_width(name=bquote(R^2)) +
-  scale_edge_color_gradient(name=bquote(R^2), low="white", high="red", lim=c(0, max(df_cca2$V_sqrt^2))) +
-  ggtitle("Pairwise squared correlation")
+  scale_edge_color_gradient(name=bquote(R^2), low="white", high="red", lim=c(0, max(df_cca2$V_sqrt^2))) +  ggtitle("Pairwise squared correlation")
 ```
@@
 
+```{r fraction.female}
+df = xtabs(~Reported_Gender + Institution, METADATA)
+df2 = apply(df, 2, function(x) x / sum(x))
+
+plotPercentBars(data.frame(t(df2))) + scale_fill_manual(values=c("red", "blue")) + ylab("Fraction of samples") + xlab("Institution")
+```
+
 ### Sex by cell fraction
@@
 
-  plot_grid(fig1, fig2, nrow=1)
+  plot_grid(fig1, fig2, nrow=1, rel_widths=c(.8, 1))
 })
@@
 
-```{r test.multtest, eval=FALSE}
-
-library(irlba)
-library(pinnacle)
-library(poolR)
-library(matrixStats)
-
-i = with(res_meta_zenith, (Resource == 'GO') & (Coef == "DxSCZ"))
-df = res_meta_zenith[i,]
-df$qvalue = qvalue(df$PValue)$qvalue
-# df$p.adjust = p.adjust(df$PValue, method="bon" )
-df$p.adjust = pmin(1, 2805 * df$PValue)
-
-M = recodeToSparseMatrix(gs.GO)
-M = M[rownames(M) %in% trim_ensembl_ids(rownames(df_meta_SCZ)),]
-keep = (colSums(M) >= 10) & (colnames(M) %in% rownames(df) )
-M = M[,keep]
-
-m = colMeans(M)
-s = apply(M, 2, var) * nrow(M) / (nrow(M) + 1)
-
-dcmp = irlba( M, nv = 500, center=m, scale=sqrt(s), right_only=FALSE)
-ev = dcmp$d^2 / nrow(M)
-
-plot(cumsum(ev) / ncol(M))
-
-df_unsort = df[colnames(M),]
-edcmp = list(vectors = dcmp$v, values = dcmp$d^2 / nrow(M))
-
-res = p.corr.adjust( df_unsort$PValue, dcmp = edcmp, method.sim="eigen", B = nrow(df_unsort)*4)
-
-plot( cumsum(dcmp$d^2 / sum(dcmp$d^2)))
-
-C = cor(as.matrix(M[,1:1000]))
-evals = eigen(C)$values
-
-dcmp = svd(scale(M) / sqrt(nrow(M)))
-dcmp$d^2
-
-plot(cumsum(dcmp$d^2 / sum(dcmp$d^2)))
-
-meff(C, method='gao')
-
-dcmp = svd(scale(M[,1:100]))
-
-meff(eigen = dcmp$d^2 , method = "galwey")
-
-```
-
 ```{r show.gene_set_test.marginal, eval=TRUE}
@@
   df$Set = factor(df$Set, df$Set)
-  ymax = max(-log10(df$PValue)) * 1.05
-  ggplot(df, aes(Set, -log10(PValue))) + geom_bar(stat="identity", fill="navy") + theme_bw() + theme(aspect.ratio=1,plot.title = element_text(hjust = 0.5)) + coord_flip() + ylab(bquote(-log[10]~P)) + xlab('') + ggtitle(coef) + scale_y_continuous(expand=c(0,0), limits=c(0, ymax))
+  ymax = max(-log10(df$FDR)) * 1.05
+  ggplot(df, aes(Set, -log10(FDR))) + geom_bar(stat="identity", fill="navy") + theme_bw() + theme(aspect.ratio=1,plot.title = element_text(hjust = 0.5)) + coord_flip() + ylab(bquote(-log[10]~FDR)) + xlab('') + ggtitle(coef) + scale_y_continuous(expand=c(0,0), limits=c(0, ymax)) + geom_hline(yintercept=-log10(0.05), color="red", linetype="dashed")
 }
@@
 # Concordance analysis
-resConcord.Sex = plot_concordance(tab.de.lst.autosome, showGenes = tab.de.lst.autosome$Symbol[1:20])
+df = df_meta[[coef]][df_meta[[coef]]$Chrom %in% 1:22,]
+resConcord.Sex = plot_concordance(tab.de.lst.autosome, showGenes = df$Symbol[1:20])
 
@@
           check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for Sex meta-analysis") %>% kable_styling(full_width=FALSE)
+
+data.frame(pi1 = 1 - qvalue(df$P)$pi0,
+          `# DE` = sum(df$adj.P.Val < 0.05),
+          check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for Sex meta-analysis - autosomes") %>% kable_styling(full_width=FALSE)
 ```
@@
 
+# Comparison with GTEx
+```{r gtex.sex, fig.height=6, fig.width=10}
+
+# read beta and se from GTEx
+df_gtex.beta = read.table(synGet("syn22891761")$path, header=TRUE, row.names=1)
+df_gtex.se = read.table(synGet("syn22891767")$path, header=TRUE, row.names=1)
+df_gtex.lfsr = read.table(synGet("syn22891755")$path, header=TRUE, row.names=1)
+df_signif = fread(synGet("syn22891771")$path) 
+df_signif = df_signif[tissue =='Brain_Frontal_Cortex_BA9', ]
+# identical(rownames(df_gtex.beta), rownames(df_gtex.se))
+# identical(rownames(df_gtex.lfsr), rownames(df_gtex.se))
+
+# data.frame of GTEx results
+df_gtex_BA9 = data.frame( ENSEMBL.full = rownames(df_gtex.beta),
+                          ENSEMBL = trim_ensembl_ids( rownames(df_gtex.beta) ),
+                          logFC = -1*df_gtex.beta[,'Brain_Frontal_Cortex_BA9'],
+                          se = df_gtex.se[,'Brain_Frontal_Cortex_BA9'],
+                          adj.P.Val = df_gtex.lfsr[,'Brain_Frontal_Cortex_BA9'])
+
+df_gtex_BA9$isSignif = FALSE
+df_gtex_BA9$isSignif[ df_gtex_BA9$ENSEMBL.full %in% df_signif$gene] = TRUE
+
+df_gtex_BA9 = df_gtex_BA9 %>% getGeneSymbol( "ENSEMBL")
+
+df_gtex_BA9$t = with(df_gtex_BA9, logFC / se)
+df_gtex_BA9$P.Value = with(df_gtex_BA9, 2*pnorm(abs(t), lower.tail=FALSE))
+# df_gtex_BA9$adj.P.Val = p.adjust(df_gtex_BA9$P.Value, "fdr")
+df_gtex_BA9 = df_gtex_BA9[!is.na(df_gtex_BA9$logFC),]
+rownames(df_gtex_BA9) = df_gtex_BA9$ENSEMBL
+
+# data.frame of CMC results from meta-analysis
+df = df_meta[['Reported_GenderMale']]
+df$t = with(df, logFC / se)
+rownames(df) = trim_ensembl_ids( rownames(df) )
+
+# all genes
+resConcord.gtex = plot_concordance( list('CMC v3' = df, 'GTEx BA9' = df_gtex_BA9), showGenes=df$Symbol[1:10] )
+
+main = "Concordance with GTEx (all genes)"
+plot_grid(  resConcord.gtex$fig_logFC + ggtitle(main),
+            resConcord.gtex$fig_tstat + ggtitle(main), nrow=1)
+
+# resConcord.gtex$tab_pi %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
+resConcord.gtex$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
+
+# because GTEx uses MASH to combine 44 tissue, standard plot_concordance() does give 
+# correct replication values
+df.merge = merge(df, df_gtex_BA9, by="row.names")
+
+table(df.merge$isSignif)
+
+df_repl = data.frame(Discovery = c("GTEx", "CMC"), rate = c(NA, NA))
+
+# Discover in CMC
+# with(df.merge, table(isSignif[adj.P.Val.x < 0.05]))
+# with(df, 1 - qvalue(P.Value.y[adj.P.Val.x < 0.05])$pi0)
+lfsr = with(df.merge, adj.P.Val.y[adj.P.Val.x < 0.05]) 
+df_repl$rate[1] = sum(1-lfsr) / length(lfsr)
+
+# Discover in GTEx
+# with(df, 1 - qvalue(P.Value.x[adj.P.Val.y < 0.05])$pi0)
+# df_repl$rate[2] = with(df.merge, 1 - qvalue(P.Value.x[isSignif])$pi0)
+
+res = with(df.merge, ash( logFC.x, se.x))
+df.merge$lsfr = res$result$lfsr
+
+lfsr = with(df.merge, lsfr[adj.P.Val.x < 0.05]) 
+df_repl$rate[2] = sum(1-lfsr) / length(lfsr)
+
+df_repl %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)
+
+# how any DEGs?
+# sum(apply(df_gtex.lfsr, 1, function(x) any(x <= 0.05, na.rm=TRUE)))
+
+# df_combine = apply(df_gtex.lfsr, 2, function(x) sum(x <= 0.05, na.rm=TRUE))
+# df_combine = data.frame(Tissue = names(df_combine), Count = df_combine)
+# df_combine$Tissue = factor(df_combine$Tissue, rev(df_combine$Tissue))
+
+# ggplot(df_combine, aes(Tissue, Count)) + geom_bar(stat="identity") + theme_bw() + coord_flip()
+
+# # All genes: 
+# sum(df_gtex.lfsr[,'Brain_Frontal_Cortex_BA9'] < 0.05, na.rm=TRUE)
+# sum(df_gtex.lfsr[,'Brain_Frontal_Cortex_BA9'] < 0.05, na.rm=TRUE)
+
+# # expressed in CMC
+# sum(df_gtex_BA9$adj.P.Val < 0.05, na.rm=TRUE)
+
+# sum( df$adj.P.Val.y < 0.05)
+
+# # get significant genes directly
+# df_combine = df_signif[,data.frame(Count = length(gene)), by="tissue"]
+# df_combine$tissue = factor(df_combine$tissue, rev(df_combine$tissue))
+
+# ggplot(df_combine, aes(tissue, Count)) + geom_bar(stat="identity") + theme_bw() + coord_flip()
+
+```
+
+```{r gtex.sex.2, fig.height=6, fig.width=10}
+# autosomes
+df = df[df$Chrom %in% 1:22,]
+
+resConcord.gtex = plot_concordance( list('CMC v3' = df, 'GTEx BA9' = df_gtex_BA9), showGenes=df$Symbol[1:10] )
+
+main = "Concordance with GTEx (autosomes)"
+plot_grid(  resConcord.gtex$fig_logFC + ggtitle(main),
+            resConcord.gtex$fig_tstat + ggtitle(main), nrow=1)
+
+# resConcord.gtex$tab_pi %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
+resConcord.gtex$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
+
+df.merge = merge(df, df_gtex_BA9, by="row.names")
+
+table(df.merge$isSignif)
+
+df_repl = data.frame(Discovery = c("GTEx", "CMC"), rate = c(NA, NA))
+
+# Discover in CMC
+# with(df.merge, table(isSignif[adj.P.Val.x < 0.05]))
+# with(df, 1 - qvalue(P.Value.y[adj.P.Val.x < 0.05])$pi0)
+lfsr = with(df.merge, adj.P.Val.y[adj.P.Val.x < 0.05]) 
+df_repl$rate[1] = sum(1-lfsr) / length(lfsr)
+
+# Discover in GTEx
+# with(df, 1 - qvalue(P.Value.x[adj.P.Val.y < 0.05])$pi0)
+# df_repl$rate[2] = with(df.merge, 1 - qvalue(P.Value.x[isSignif])$pi0)
+
+res = with(df.merge, ash( logFC.x, se.x))
+df.merge$lsfr = res$result$lfsr
+
+lfsr = with(df.merge, lsfr[adj.P.Val.x < 0.05]) 
+df_repl$rate[2] = sum(1-lfsr) / length(lfsr)
+
+df_repl %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)
+```
+
 # Sex by Disease interaction
@@
 
+```{r separate.test.sz}
+
+df = merge(df_meta.separ[[1]], df_meta.separ[[2]], by="row.names")
+
+df$tstat = with(df, (logFC.x - logFC.y) / sqrt(se.x^2 + se.y^2))
+
+df$p = 2*pnorm(abs(df$tstat), 0, 1, lower.tail=FALSE)
+df$FDR = p.adjust(df$p, "fdr")
+
+head(df[df$FDR < 0.1,])
+
+```
+
 ```{r separ.concordance.scz, fig.height=6, fig.width=10}
@@
 resConcord.separ$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
-```
-
-```{r compare.separate, fig.width=5, fig.height=5}
-df = merge( df_meta.separ[[1]], df_meta.separ[[2]], by="row.names")
-
-fit = with(df, lm(logFC.y ~ logFC.x))
-
-coef(summary(fit) ) %>%  kable(caption="logFC is larger in males") %>% kable_styling(full_width=FALSE)
-
- # compare logFC
-tstat = with(df, (logFC.y - logFC.x) / (sqrt(se.y^2 + se.x^2)))
-
-qqt(tstat)
-abline(0, 1, col="red")
-p = 2*pt(abs(tstat), df=778, lower.tail=FALSE)
-table(p.adjust(p) < 0.05)
 ```

### 2020-10-06 individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_ACC_DE_DxSex.Rmd
update provenance - all cohorts

@@
 ```
-`r formula`
+`r glue::glue_collapse(formula)`
 ### Residual calculation

### 2020-10-06 individual-cohorts/split_by_analysis/MSSM-Penn-Pitt_DLPFC_DE_DxSex.Rmd

@@
   x = as.numeric(as.character(gsub('[\\,\\%]','',x)))
-})c
+})
 

### 2020-10-12 combined_analysis/combine_all_DLPFC_by_celltype.Rmd
initial commit

--- 2020-10-06 combined_analysis/combine_all_DLPFC.Rmd
@@
 
-rmarkdown::render("combined_analysis/combine_all_DLPFC.Rmd")
-
+rmarkdown::render("combined_analysis/combine_all_DLPFC_by_celltype.Rmd")
+
+# load("geneExpr_logCPM_bMIND.RDATA")
+# saveRDS(geneExpr_logCPM_bMIND, file="geneExpr_logCPM_bMIND.RDS")
+# saveRDS(geneExpr_logCPM_bMIND_se, file="geneExpr_logCPM_bMIND_se.RDS")
+
+# ml python
+# synapse add --parentid syn23234712 geneExpr_logCPM_bMIND.RDS
+# synapse add --parentid syn23234712 geneExpr_logCPM_bMIND_se.RDS
 --->
@@
 <!---
-
-TODO
-
-zenith 1 sided
-
-zenith plots
-replication or correlation as a function of expression magnitude
-
-difference networks
--------------------
-
-DONE
-
-Effect of ilr should depend on Instaition
-suffix 1.
-zenith for Sex with autosomes
-check zenith squared test statistics
-
---->
 
@@
 
-```{r read.counts}
+```{r read.counts, eval=FALSE}
 # Read raw counts from MPP and HBCC
@@
 geneCountsMerged = cbind(counts.MPP, counts.HBCC, counts.HBCC4)
+```
+
+```{r read.cts.expression}
+geneExpr.CTS = readRDS(synGet('syn23234833')$path)
+geneExpr.CTS.se = readRDS(synGet('syn23234858')$path)
+
+cellType = "GABA"
+geneExpr = geneExpr.CTS[,cellType,]
+geneExpr.se = geneExpr.CTS.se[,cellType,]
+
 ```
@@
                 IntragenicRate, IntergenicRate, IntronicRate, ExpProfEfficiency, rRNARate, TranscriptsDetected) %>%
-  		 		 dplyr::filter(SampleID %in% colnames(geneCountsMerged), !is.na(SampleID)) %>%
+  		 		 dplyr::filter(SampleID %in% colnames(geneExpr), !is.na(SampleID)) %>%
   		 		 dplyr::filter(Region %in% c('DLPFC')) %>%

### 2020-10-12 combined_analysis/combine_all_DLPFC_by_celltype.Rmd
initial commit

@@
 title: "CommonMind Consortium: DLPFC"
-subtitle: "Analysis of read count data by cohort"
+subtitle: "Cell type specific analysis of read count data by cohort"
 author:
@@
 synLogin()
-knit2synapse::createAndKnitToFolderEntity(file = "combined_analysis/combine_all_DLPFC.Rmd",
+knit2synapse::createAndKnitToFolderEntity(file = "combined_analysis/combine_all_DLPFC_by_celltype.Rmd",
                                           parentId ="syn22416298",
-                                          folderName = 'Joint analysis of DLPFC count data',
+                                          folderName = 'Cell type specific analysis of read count data by cohort',
                                           overwrite=TRUE, 
@@
 parentId = 'syn22416298';
-activityName = 'Joint analysis of DLPFC count data';
-activityDescription = 'Joint analysis of DLPFC count data';
-
-thisFileName <- 'combine_all_DLPFC.Rmd'
+activityName = 'Cell type specific analysis of read count data by cohort';
+activityDescription = 'Cell type specific analysis of read count data by cohort';
+
+thisFileName <- 'combine_all_DLPFC_by_celltype.Rmd'
 
@@
 # Match covariates to expression data
-indToRetain = intersect(METADATA$SampleID, colnames(geneCountsMerged))
-
-geneCountsMerged = geneCountsMerged[,indToRetain]
+indToRetain = intersect(METADATA$SampleID, colnames(geneExpr))
+
+geneExpr = geneExpr[,indToRetain]
 
@@
 plot_grid(plotlist=figList, ncol=1) 
+```
+
+```{r create.voom}
+
 ```
@@
   # Make sure these match the folder name and parent Id in the first chunk.
-  CODE <- Folder(name = "Joint analysis of DLPFC count data", parentId = "syn22416298")
+  CODE <- Folder(name = "Cell type specific analysis of read count data by cohort", parentId = "syn22416298")
   CODE <- synStore(CODE)

### 2020-10-12 combined_analysis/combine_all_DLPFC_by_celltype.Rmd
initial commit

@@
 
-```
-
-```{r process.counts}
-# keep genes that are express in at least one combination of cohort and gender
-isexpr.lst = dlply(METADATA, .(Cohort), .fun = function(mtd){
-	rowSums(cpm(geneCountsMerged[,mtd$SampleID])>.5) >= .3*ncol(geneCountsMerged[,mtd$SampleID])
-	})
-
-isexpr = do.call(rbind, isexpr.lst)
-isexpr = colSums(isexpr) == 2 # keep gene expressed in both cohorts 
-table(isexpr)
-
-# Normalize each cohort separately
-dge.lst = lapply( levels(METADATA$Cohort), function(Cohort){
-  # normalize counts
-  dge = DGEList( geneCountsMerged[isexpr,METADATA$Cohort==Cohort] )
-  calcNormFactors(dge, "none") # TMM
-})
-names(dge.lst) = levels(METADATA$Cohort)
-
-# form = ~ (1|Dx) + Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)
-
-form.lst = list(
-  `MSSM-Penn-Pitt`  = ~ Dx + Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
-  `NIMH-HBCC`       = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)
-
-vobj.lst = lapply(dge.lst, function(dge){
-
-  i = match(colnames(dge), rownames(METADATA))
-  info = METADATA[i,]
-
-  form_use = form.lst[[info$Cohort[1]]]
-
-  voomWithDreamWeights(dge, form_use, info, save.plot=TRUE, BPPARAM = SnowParam(18))
-  })
-names(vobj.lst) = names(dge.lst)
-```
-
-### Voom trend plots
-```{r plot.voom, fig.height=5, fig.width=10}
-fig1 = plot_voom(vobj.lst[[1]]) + ggtitle(names(vobj.lst)[1])
-fig2 = plot_voom(vobj.lst[[2]]) + ggtitle(names(vobj.lst)[2])
-plot_grid( fig1, fig2, nrow=1)
-
-df = lapply( names(vobj.lst), function(key){
-  data.frame(Cohort = key,  vobj.lst[[key]]$voom.line)
-  })
-df = do.call(rbind, df)
-
-ggplot(df, aes(x, y, color=Cohort)) + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + geom_line(size=1.2) + xlab(bquote(log[2](count + 0.5))) + ylab(expression( sqrt("standard deviation"))) + ggtitle("Compare voom trends") + ylim(0, max(df$y)) + scale_color_manual(name="Status", values=c("#3758ba", "#33e65c")) + guides(color = guide_legend(override.aes = list(size = 1)))
-```
-
-<!---
-
-# Examine genes on chrX and chrY
-
-# dge = DGEList(geneCountsMerged)
-# dge = calcNormFactors(dge, "none")
-# geneExpr = cpm(dge, log=TRUE)
-
-# df = merge( data.frame(Expression = geneExpr['ENSG00000229807.11',]), METADATA, by="row.names")
-# ggplot(df, aes(Reported_Gender, Expression)) + geom_violin() + theme_bw() + facet_wrap(~Cohort)
-
-# id = 'ENSG00000229807.11'
-# isexpr[id]
-
-# df = as.data.frame(vobj.lst[[2]]$voom.xy)
-# with(df, plot(x,y, cex=.2, pch=20))
-# with(df[id,], points(x,y, col="red", pch=20, cex=2))
-
---->
+vobj.lst = lapply( levels(METADATA$Cohort), function(chrt){
+
+  idx = METADATA$Cohort == chrt
+
+  new("EList", list(E = geneExpr[,idx], weights = 1/geneExpr.se[rownames(geneExpr), colnames(geneExpr)][,idx]))
+  })
+names(vobj.lst) = levels(METADATA$Cohort)
+```
 
@@
 form.lst = list(
-  `MSSM-Penn-Pitt`  = ~ Dx*Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
-  `NIMH-HBCC`       = ~ Dx*Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)
+  `MSSM-Penn-Pitt`  = ~ Dx*Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*ageOfDeath,
+  `NIMH-HBCC`       = ~ Dx*Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath)
 
@@
 form.lst = list(
-  `MSSM-Penn-Pitt`  = ~ Dx + Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
-  `NIMH-HBCC`       = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)
+  `MSSM-Penn-Pitt`  = ~ Dx + Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + InstitutionageOfDeath,
+  `NIMH-HBCC`       = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath)
 
@@
   # Standard analysis
-  storeData( "geneCountsMerged", "Merged DLPFC counts for both cohorts")
+  # storeData( "geneCountsMerged", "Merged DLPFC counts for both cohorts")
   storeData( "METADATA", "Merged metadata used here for both cohorts")

### 2020-10-12 combined_analysis/combine_all_DLPFC_by_celltype.Rmd
initial commit

@@
 ```{r create.voom}
-
 vobj.lst = lapply( levels(METADATA$Cohort), function(chrt){
@@
 form.lst = list(
-  `MSSM-Penn-Pitt`  = ~ Dx + Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + InstitutionageOfDeath,
+  `MSSM-Penn-Pitt`  = ~ Dx + Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*ageOfDeath,
   `NIMH-HBCC`       = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath)

### 2020-10-12 combined_analysis/combine_all_DLPFC_by_celltype.Rmd
initial commit

@@
 library(kableExtra)
+library(ashr)
 library(githubr)

### 2020-10-12 combined_analysis/CTS/GABA.Rmd
initial commit

--- 2020-10-12 combined_analysis/combine_all_DLPFC_by_celltype.Rmd
@@
   upload: FALSE
+  cellType: "GABA"
 ---
@@
 
-rmarkdown::render("combined_analysis/combine_all_DLPFC_by_celltype.Rmd")
+rmarkdown::render("combined_analysis/CTS/GABA.Rmd")
 
@@
 knit2synapse::createAndKnitToFolderEntity(file = "combined_analysis/combine_all_DLPFC_by_celltype.Rmd",
-                                          parentId ="syn22416298",
-                                          folderName = 'Cell type specific analysis of read count data by cohort',
+                                          parentId ="syn23241817",
+                                          folderName = 'Analysis of read count data by cohort: GABA',
                                           overwrite=TRUE, 
@@
 library(githubr)
-parentId = 'syn22416298';
-activityName = 'Cell type specific analysis of read count data by cohort';
-activityDescription = 'Cell type specific analysis of read count data by cohort';
-
-thisFileName <- 'combine_all_DLPFC_by_celltype.Rmd'
+parentId = 'syn23241817';
+activityName = 'Analysis of read count data by cohort: GABA';
+activityDescription = 'Analysis of read count data by cohort: GABA';
+
+thisFileName <- 'GABA'
 
@@
 thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
-thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('combined_analysis/',thisFileName))
+thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('combined_analysis/CTS/',thisFileName))
 ```
@@
 
-cellType = "GABA"
-geneExpr = geneExpr.CTS[,cellType,]
-geneExpr.se = geneExpr.CTS.se[,cellType,]
+geneExpr = geneExpr.CTS[,params$cellType,]
+geneExpr.se = geneExpr.CTS.se[,params$cellType,]
 
@@
 rownames(METADATA) = METADATA$SampleID
-```
-
-## Summarize samples
-```{r summary}
-METADATA %>% group_by(Dx, Institution) %>% dplyr::summarise(count = n()) %>% 
-  kable(caption="Diagnosis, Institution") %>% kable_styling(full_width=FALSE)
-
-METADATA %>% group_by( Institution, Reported_Gender) %>% dplyr::summarise(count = n()) %>% 
-  kable(caption="Institution, Sex") %>% kable_styling(full_width=FALSE)
-
-METADATA %>% group_by(Dx, Institution, Reported_Gender) %>% dplyr::summarise(count = n()) %>% 
-kable(caption="Diagnosis, Institution, Sex") %>% kable_styling(full_width=FALSE) 
-```
-
-```{r joint.cca, fig.width=7, fig.height=7}
-variables = c('Reported_Gender', 'CellType', 'ageOfDeath', 'Institution', "Dx")
-
-df_cca = lapply(variables, function(v1){
-
-  if( v1 == "CellType"){
-    form1 = ~ GABA + GLU + Olig + MgAs
-  }else{
-    form1 = as.formula(paste0('~', v1))
-  }
-  M1 = model.matrix(form1, METADATA)
-
-  res = lapply(variables, function(v2){
-
-    if( v2 == "CellType"){
-      form2 = ~ GABA + GLU + Olig + MgAs
-    }else{
-      form2 = as.formula(paste0('~', v2))
-    }
-    M2 = model.matrix(form2, METADATA)
-
-    fit = cancor(M1, M2)
-    V_sqrt = sqrt(mean(fit$cor^2)) # Cramer's sqrt V
-
-    data.frame(v1, v2, V_sqrt)
-  })
-  do.call(rbind, res)
-})
-df_cca = do.call(rbind, df_cca)
-
-C = matrix(df_cca$V_sqrt, ncol=length(variables))
-rownames(C) = variables
-colnames(C) = variables
-
-plotCorrMatrix(C, main='CCA')
-```
-
-```{r explain.CellType}
-CTs = c('GABA', 'GLU', 'Olig', 'MgAs')
-
-form = ~ ageOfDeath:Institution + Reported_Gender + Dx + Institution
-
-Y = clr(METADATA[,CTs])
-
-df = lapply( colnames(Y), function(id){
-
-  fit = lm(Y[,id] ~ ageOfDeath:Institution + Reported_Gender + Dx + Institution, METADATA)
-
-  data.frame( Test     = rownames(anova(fit))[1:4],
-              CellType = id, 
-              p        = anova(fit)[,'Pr(>F)'][1:4], 
-              frac     =  calcVarPart(fit)[1:4])
-})
-df = data.table(do.call(rbind, df))
-
-df$Test[df$Test=='Reported_Gender'] = "Sex"
-df$Test[df$Test=='ageOfDeath:Institution'] = "Age of death"
-df$p.adj = p.adjust(df$p, "bonferroni")
-
-ymax = max(df[,sum(frac),by="CellType"]$V1)*105
-ggplot(df, aes(CellType, frac*100, fill=Test)) + geom_bar(stat="identity") + coord_flip() + theme_bw() + theme(aspect.ratio=1, panel.grid.major = element_blank(), 
-        panel.grid.minor = element_blank()) +  ylab("Variance explained (%)") + theme(axis.line = element_line(colour = "transparent"), 
-        axis.line.x = element_line(colour = "black"), panel.grid.major = element_blank(), 
-        panel.grid.minor = element_blank(), panel.border = element_blank(), 
-        panel.background = element_blank(), axis.ticks.y = element_blank(), 
-        legend.key = element_blank(), plot.margin = unit(c(0, 
-            0.3, 0, 0.8), "cm")) + scale_y_continuous(expand=c(0,0), limits=c(0, ymax)) + xlab('')
-
-ggplot(df, aes(CellType, Test, fill=-log10(pmax(1e-16,p)), label=ifelse(p.adj<0.05, '*', ''))) + geom_tile() + theme_bw() + scale_fill_gradient(name = bquote(-log[10]~P), low="white", high="red", limit=c(0, NA)) + theme(aspect.ratio=1, panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank()) + geom_text() + coord_flip()
-```
-
-```{r sex.graph, fig.width=7, fig.height=7, dev="png"}
-df_cca2 = df_cca[with(df_cca, v1!=v2),]
-df_cca2$key = apply(df_cca2, 1, function(x){
-  x = sort(c(x['v1'], x['v2']))
-  paste(x[1], x[2])
-  }) 
-df_cca2 = df_cca2[!duplicated(df_cca2$key),]
-
-rstat_nodes <- data.frame(name = unique(c(df_cca2$v1, df_cca2$v2)))
-
-rstat_edges <- data.frame(from = match(df_cca2$v1, rstat_nodes$name),
-                         to = match(df_cca2$v2, rstat_nodes$name),
-                         Rsq = df_cca2$V_sqrt^2 )
-
-net = tbl_graph(nodes = rstat_nodes, edges = rstat_edges, directed=FALSE)
-
-ggraph(net,  layout = 'igraph', algorithm = 'star') + 
-  geom_edge_link(aes(width = Rsq, color=Rsq), alpha = 1) +
-  geom_node_point(size=20, color="grey70") +
-  geom_node_text(aes(label = name), repel = FALSE) +
-  theme_graph() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) +
-  scale_edge_width(name=bquote(R^2)) +
-  scale_edge_color_gradient(name=bquote(R^2), low="white", high="red", lim=c(0, max(df_cca2$V_sqrt^2))) +  ggtitle("Pairwise squared correlation")
-```
-
-```{r sex.confounding, fig.width=5, fig.height=5}
-form = ~ Reported_Gender + GABA + GLU + Olig + MgAs + ageOfDeath 
-par(mfrow=c(2,2))
-res = lapply( unique(METADATA$Institution), function(INST){
-  i = METADATA$Institution == INST
-  C = canCorPairs(form, METADATA[i,]) 
-  plotCorrMatrix(C, main=INST)
-})
-```
-
-```{r sex.by.age}
-ggplot(METADATA, aes(Reported_Gender, ageOfDeath, fill=Reported_Gender)) + geom_boxplot() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + facet_wrap(~Institution, nrow=1) + scale_fill_manual(values = c("red", "blue"))
-```
-
-```{r fraction.female}
-df = xtabs(~Reported_Gender + Institution, METADATA)
-df2 = apply(df, 2, function(x) x / sum(x))
-
-plotPercentBars(data.frame(t(df2))) + scale_fill_manual(values=c("red", "blue")) + ylab("Fraction of samples") + xlab("Institution")
-```
-
-### Sex by cell fraction
-```{r summary.sex, fig.width=8, fig.height=10}
-figList = lapply(colnames(df_cellFractions), function(CT){
-  ggplot(METADATA, aes_string('Reported_Gender', CT, fill='Reported_Gender')) + geom_boxplot() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + facet_wrap(~Institution, nrow=1) + scale_fill_manual(values = c("red", "blue"))
-})
-plot_grid(plotlist=figList, ncol=1) 
-```
-
-### Age by cell fraction
-```{r summary.Age, fig.width=8, fig.height=10}
-figList = lapply(colnames(df_cellFractions), function(CT){
-  ggplot(METADATA, aes_string('ageOfDeath', CT)) + geom_point() + geom_smooth(method="gam") + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + facet_wrap(~Institution, nrow=1)
-}) 
-plot_grid(plotlist=figList, ncol=1) 
-```
-
-## Age by cell fraction log ratio
-```{r summary.Age.log.ratio, fig.width=8, fig.height=8}
-figList = lapply(colnames(celFrac_ilr), function(CT){
-  ggplot(METADATA, aes_string('ageOfDeath', CT)) + geom_point() + geom_smooth(method="gam") + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + facet_wrap(~Institution, nrow=1)
-})
-plot_grid(plotlist=figList, ncol=1) 
 ```
@@
   # Make sure these match the folder name and parent Id in the first chunk.
-  CODE <- Folder(name = "Cell type specific analysis of read count data by cohort", parentId = "syn22416298")
+  CODE <- Folder(name = activityName, parentId = parentId)
   CODE <- synStore(CODE)
@@
   storeData = function( name, description){
-    file = paste0(name, ".RDS")
+    file = paste0(name, "_", params$cellType, ".RDS")
     saveRDS( eval(parse(text=name)), file=file)
@@
     nEXP_OBJ = File(file, 
-                    name = description, 
+                    name = paste0(params$cellType, ' - ',description), 
                     parentId = CODE$properties$id)

### 2020-10-12 combined_analysis/CTS/MgAs.Rmd
initial commit

--- 2020-10-12 combined_analysis/combine_all_DLPFC_by_celltype.Rmd
@@
   # Store data 
-  storeData = function( name, description){
-    file = paste0(name, ".RDS")
+  storeData = function( name, cellType, description){
+    file = paste0(name, "_", cellType, ".RDS")
     saveRDS( eval(parse(text=name)), file=file)
@@
     nEXP_OBJ = File(file, 
-                    name = description, 
+                    name = paste0(cellType, ' - ',description), 
                     parentId = CODE$properties$id)

### 2020-10-12 combined_analysis/CTS/Glu.Rmd
consistant names

--- 2020-10-12 combined_analysis/CTS/GABA.Rmd
@@
   upload: FALSE
-  cellType: "GABA"
+  cellType: "Glu"
 ---
@@
 
-rmarkdown::render("combined_analysis/CTS/GABA.Rmd")
+rmarkdown::render("combined_analysis/CTS/Glu.Rmd")
 
@@
                                           parentId ="syn23241817",
-                                          folderName = 'Analysis of read count data by cohort: GABA',
+                                          folderName = 'Analysis of read count data by cohort: Glu',
                                           overwrite=TRUE, 
@@
 parentId = 'syn23241817';
-activityName = 'Analysis of read count data by cohort: GABA';
-activityDescription = 'Analysis of read count data by cohort: GABA';
-
-thisFileName <- 'GABA'
+activityName = 'Analysis of read count data by cohort: Glu';
+activityDescription = 'Analysis of read count data by cohort: Glu';
+
+thisFileName <- 'Glu.Rmd'
 
@@
 ```
-
-<!---
 
@@
 }) 
-source("../common_functions/common_functions.R")   
+source("../../common_functions/common_functions.R")   
 # source("common_functions/common_functions.R")

### 2020-10-12 combined_analysis/CTS/MgAs.Rmd
consistant names

@@
   upload: FALSE
-  cellType: "GABA"
+  cellType: "MgAs"
 ---
@@
 
-rmarkdown::render("combined_analysis/combine_all_DLPFC_by_celltype.Rmd")
+rmarkdown::render("combined_analysis/CTS/MgAs.Rmd")
 
@@
 knit2synapse::createAndKnitToFolderEntity(file = "combined_analysis/combine_all_DLPFC_by_celltype.Rmd",
-                                          parentId ="syn22416298",
-                                          folderName = 'Cell type specific analysis of read count data by cohort',
+                                          parentId ="syn23241817",
+                                          folderName = 'Analysis of read count data by cohort: MgAs',
                                           overwrite=TRUE, 
@@
 library(githubr)
-parentId = 'syn22416298';
-activityName = 'Cell type specific analysis of read count data by cohort';
-activityDescription = 'Cell type specific analysis of read count data by cohort';
-
-thisFileName <- 'combine_all_DLPFC_by_celltype.Rmd'
+parentId = 'syn23241817';
+activityName = 'Analysis of read count data by cohort: MgAs';
+activityDescription = 'Analysis of read count data by cohort: MgAs';
+
+thisFileName <- 'MgAs.Rmd'
 
@@
 thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
-thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('combined_analysis/',thisFileName))
-```
-
-<!---
+thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('combined_analysis/CTS/',thisFileName))
+```
 
@@
   # Store data 
-  storeData = function( name, cellType, description){
-    file = paste0(name, "_", cellType, ".RDS")
+  storeData = function( name, description){
+    file = paste0(name, "_", params$cellType, ".RDS")
     saveRDS( eval(parse(text=name)), file=file)
@@
     nEXP_OBJ = File(file, 
-                    name = paste0(cellType, ' - ',description), 
+                    name = paste0(params$cellType, ' - ',description), 
                     parentId = CODE$properties$id)

### 2020-10-12 combined_analysis/CTS/Olig.Rmd
consistant names

--- 2020-10-12 combined_analysis/CTS/Glu.Rmd
@@
   upload: FALSE
-  cellType: "Glu"
+  cellType: "Olig"
 ---
@@
 
-rmarkdown::render("combined_analysis/CTS/Glu.Rmd")
+rmarkdown::render("combined_analysis/CTS/Olig.Rmd")
 
@@
                                           parentId ="syn23241817",
-                                          folderName = 'Analysis of read count data by cohort: Glu',
+                                          folderName = 'Analysis of read count data by cohort: Olig',
                                           overwrite=TRUE, 
@@
 parentId = 'syn23241817';
-activityName = 'Analysis of read count data by cohort: Glu';
-activityDescription = 'Analysis of read count data by cohort: Glu';
-
-thisFileName <- 'Glu.Rmd'
+activityName = 'Analysis of read count data by cohort: Olig';
+activityDescription = 'Analysis of read count data by cohort: Olig';
+
+thisFileName <- 'Olig.Rmd'
 

### 2020-10-12 combined_analysis/CTS/GABA.Rmd
consistant names
drop genes that don't vary

@@
 params:
-  upload: FALSE
+  upload: TRUE
   cellType: "GABA"
@@
 
-thisFileName <- 'GABA'
+thisFileName <- 'GABA.Rmd'
 
@@
 
-  new("EList", list(E = geneExpr[,idx], weights = 1/geneExpr.se[rownames(geneExpr), colnames(geneExpr)][,idx]))
+  # get variance withing cohort
+  # only keep genes with variance > 0
+  v = apply( geneExpr[,idx], 1, var)
+
+  E = geneExpr[v > 0,idx]
+
+  new("EList", list(E = E, weights = 1/geneExpr.se[rownames(E), colnames(E)]))
   })
@@
 ```{r store, eval=TRUE}
+
 if( params$upload ){    
@@
   # Make sure these match the folder name and parent Id in the first chunk.
-  CODE <- Folder(name = activityName, parentId = parentId)
+  CODE <- Folder(name = gsub(":", " -",activityName), parentId = parentId)
   CODE <- synStore(CODE)
@@
   storeData( "METADATA", "Merged metadata used here for both cohorts")
-  storeData( "isexpr", "Genes used in this analysis")
+  # storeData( "isexpr", "Genes used in this analysis")
   storeData( "vobj.lst", "Voom results for each cohort")

### 2020-10-12 combined_analysis/CTS/Glu.Rmd
drop genes that don't vary

@@
 params:
-  upload: FALSE
-  cellType: "Glu"
+  upload: TRUE
+  cellType: "GLU"
 ---

### 2020-10-12 combined_analysis/CTS/MgAs.Rmd
drop genes that don't vary

@@
 geneExpr.se = geneExpr.CTS.se[,params$cellType,]
+
+v = apply(geneExpr, 1, var)
 
@@
     # Compute residuals which adding back Diagnosis and Sex effects
-    resid.lst = fitVarPartModel(vobj, form_use, info, fxn = f, BPPARAM = SnowParam(18))  
+    resid.lst = fitVarPartModel(vobj, form_use, info, fxn = f, BPPARAM = SnowParam(4))  
     do.call(rbind, resid.lst)

### 2020-10-13 combined_analysis/CTS/GABA.Rmd
drop genes that don't vary

@@
 names(vobj.lst) = levels(METADATA$Cohort)
+
+# only keep genes common to both cohorts
+isexpr = intersect(rownames(vobj.lst[[1]]), rownames(vobj.lst[[2]]))
+
+vobj.lst = lapply(vobj.lst, function(vobj){
+  vobj[isexpr,]
+  })
 ```

### 2020-10-13 combined_analysis/CTS/MgAs.Rmd
drop genes that don't vary

@@
 names(vobj.lst) = levels(METADATA$Cohort)
+
+# only keep genes common to both cohorts
+isexpr = intersect(rownames(vobj.lst[[1]]), rownames(vobj.lst[[2]]))
+
 ```
@@
 
-  fit = dream(vobj[keep.ensID,], form_use, info, BPPARAM = SnowParam(18))  
+  fit = dream(vobj[keep.ensID,], form_use, info, BPPARAM = SnowParam(4))  
   eBayes( fit )

### 2020-10-13 combined_analysis/CTS/Olig.Rmd
drop genes that don't vary

@@
 names(vobj.lst) = levels(METADATA$Cohort)
-```
+
+# only keep genes common to both cohorts
+isexpr = intersect(rownames(vobj.lst[[1]]), rownames(vobj.lst[[2]]))
+
+vobj.lst = lapply(vobj.lst, function(vobj){
+  vobj[isexpr,]
+  })
 

### 2020-10-13 combined_analysis/CTS/GABA.Rmd
drop genes that don't vary

@@
 parentId = 'syn23241817';
-activityName = 'Analysis of read count data by cohort: GABA';
-activityDescription = 'Analysis of read count data by cohort: GABA';
-
-thisFileName <- 'GABA.Rmd'
+activityName = paste('Analysis of read count data by cohort:', params$cellType)
+activityDescription = paste('Analysis of read count data by cohort:', params$cellType)
+
+thisFileName <- paste0(params$cellType, '.Rmd')
 

### 2020-10-13 combined_analysis/CTS/Glu.Rmd
drop genes that don't vary

@@
 parentId = 'syn23241817';
-activityName = 'Analysis of read count data by cohort: Glu';
-activityDescription = 'Analysis of read count data by cohort: Glu';
-
-thisFileName <- 'Glu.Rmd'
+activityName = paste('Analysis of read count data by cohort:', params$cellType)
+activityDescription = paste('Analysis of read count data by cohort:', params$cellType)
+
+thisFileName <- paste0(params$cellType, '.Rmd')
 

### 2020-10-13 combined_analysis/CTS/MgAs.Rmd
drop genes that don't vary

@@
 parentId = 'syn23241817';
-activityName = 'Analysis of read count data by cohort: MgAs';
-activityDescription = 'Analysis of read count data by cohort: MgAs';
-
-thisFileName <- 'MgAs.Rmd'
+activityName = paste('Analysis of read count data by cohort:', params$cellType)
+activityDescription = paste('Analysis of read count data by cohort:', params$cellType)
+
+thisFileName <- paste0(params$cellType, '.Rmd')
 
@@
 
+vobj.lst = lapply(vobj.lst, function(vobj){
+  vobj[isexpr,]
+  })
 ```

### 2020-10-13 combined_analysis/CTS/Olig.Rmd
drop genes that don't vary

@@
 parentId = 'syn23241817';
-activityName = 'Analysis of read count data by cohort: Olig';
-activityDescription = 'Analysis of read count data by cohort: Olig';
-
-thisFileName <- 'Olig.Rmd'
+activityName = paste('Analysis of read count data by cohort:', params$cellType)
+activityDescription = paste('Analysis of read count data by cohort:', params$cellType)
+
+thisFileName <- paste0(params$cellType, '.Rmd')
 
@@
   })
+```
 

### 2020-10-13 combined_analysis/CTS/GABA.Rmd
drop genes that don't vary

@@
 
-  idx = METADATA$Cohort == chrt
-
-  # get variance withing cohort
+  # get variance within cohort and sex
   # only keep genes with variance > 0
-  v = apply( geneExpr[,idx], 1, var)
-
-  E = geneExpr[v > 0,idx]
+  # v = apply( geneExpr[,idx], 1, var)
+  v = lapply( unique(METADATA$Reported_Gender), function(Gnd){
+    idx = with(METADATA, (Cohort == chrt) & (Reported_Gender==Gnd))
+    isVar = apply( geneExpr[,idx], 1, var) > 0
+    names(isVar[isVar])
+    })
+
+  # genes with variance in both sexes
+  isexpr = intersect(v[[1]], v[[2]])
+
+  # get expressed genes form this cohort
+  E = geneExpr[isexpr, METADATA$Cohort == chrt]
 
@@
 form.lst = list(
-  `MSSM-Penn-Pitt`  = ~ Dx*Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
-  `NIMH-HBCC`       = ~ Dx*Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)
+  `MSSM-Penn-Pitt`  = ~ Dx*Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*ageOfDeath ,
+  `NIMH-HBCC`       = ~ Dx*Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath )
 
@@
 form.lst = list(
-  `MSSM-Penn-Pitt`  = ~ Dx + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
-  `NIMH-HBCC`       = ~ Dx + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)
+  `MSSM-Penn-Pitt`  = ~ Dx + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*ageOfDeath ,
+  `NIMH-HBCC`       = ~ Dx + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath)
 

### 2020-10-13 combined_analysis/CTS/Glu.Rmd
test
drop genes that don't vary

@@
                                           parentId ="syn23241817",
-                                          folderName = 'Analysis of read count data by cohort: Glu',
+                                          folderName = 'Analysis of read count data by cohort: GABA',
                                           overwrite=TRUE, 
@@
 
-thisFileName <- paste0(params$cellType, '.Rmd')
+thisFileName <- paste0(ifelse(params$cellType=='GLU', 'Glu', 'GLU'), '.Rmd')
 

### 2020-10-13 combined_analysis/CTS/MgAs.Rmd
drop genes that don't vary

@@
 params:
-  upload: FALSE
+  upload: TRUE
   cellType: "MgAs"
@@
                                           parentId ="syn23241817",
-                                          folderName = 'Analysis of read count data by cohort: MgAs',
+                                          folderName = 'Analysis of read count data by cohort: GABA',
                                           overwrite=TRUE, 
@@
 geneExpr.se = geneExpr.CTS.se[,params$cellType,]
-
-v = apply(geneExpr, 1, var)
 
@@
     # Compute residuals which adding back Diagnosis and Sex effects
-    resid.lst = fitVarPartModel(vobj, form_use, info, fxn = f, BPPARAM = SnowParam(4))  
+    resid.lst = fitVarPartModel(vobj, form_use, info, fxn = f, BPPARAM = SnowParam(18))  
     do.call(rbind, resid.lst)
@@
 
-  fit = dream(vobj[keep.ensID,], form_use, info, BPPARAM = SnowParam(4))  
+  fit = dream(vobj[keep.ensID,], form_use, info, BPPARAM = SnowParam(18))  
   eBayes( fit )

### 2020-10-13 combined_analysis/CTS/Olig.Rmd
drop genes that don't vary

@@
 params:
-  upload: FALSE
+  upload: TRUE
   cellType: "Olig"
@@
                                           parentId ="syn23241817",
-                                          folderName = 'Analysis of read count data by cohort: Olig',
+                                          folderName = 'Analysis of read count data by cohort: GABA',
                                           overwrite=TRUE, 

### 2020-10-14 combined_analysis/CTS/combine.Rmd
initial commit

--- 2020-10-13 combined_analysis/CTS/Olig.Rmd
@@
 ---
-title: "CommonMind Consortium: DLPFC"
-subtitle: "Cell type specific analysis of read count data by cohort"
+title: "Combine cell type specific analysis"
+subtitle: ""
 author:
-- name: "[Gabriel Hoffman](http://gabrielhoffman.github.io)$^{1}$ and Kelsey Montgomery$^{2}$"
+- name: "[Gabriel Hoffman](http://gabrielhoffman.github.io)"
   affiliation: | 
-    [1] Icahn School of Medicine at Mount Sinai, New York
-    [2] Sage Bionetworks, Seattle, WA
+    Icahn School of Medicine at Mount Sinai, New York
 date: "Run on `r Sys.time()`"
@@
     toc_float: true
-params:
-  upload: FALSE
-  cellType: "Olig"
 ---
@@
 
-rmarkdown::render("combined_analysis/CTS/Olig.Rmd")
+rmarkdown::render("combined_analysis/CTS/combine.Rmd")
 
-# load("geneExpr_logCPM_bMIND.RDATA")
-# saveRDS(geneExpr_logCPM_bMIND, file="geneExpr_logCPM_bMIND.RDS")
-# saveRDS(geneExpr_logCPM_bMIND_se, file="geneExpr_logCPM_bMIND_se.RDS")
-
-# ml python
-# synapse add --parentid syn23234712 geneExpr_logCPM_bMIND.RDS
-# synapse add --parentid syn23234712 geneExpr_logCPM_bMIND_se.RDS
 --->
 
-```{r knit2synapse, eval= FALSE}
-library(synapser)
-library(knit2synapse)
-synLogin()
-knit2synapse::createAndKnitToFolderEntity(file = "combined_analysis/combine_all_DLPFC_by_celltype.Rmd",
-                                          parentId ="syn23241817",
-                                          folderName = 'Analysis of read count data by cohort: Olig',
-                                          overwrite=TRUE, 
-                                          knitmd=TRUE)
-```
-
-```{r synapse.parameters, include=FALSE, cache=TRUE}
+```{r synapse.parameters, include=FALSE, cache=TRUE, eval=FALSE}
 
@@
 parentId = 'syn23241817';
-activityName = 'Analysis of read count data by cohort: Olig';
-activityDescription = 'Analysis of read count data by cohort: Olig';
+activityName = "Combine cell type specific analysis"
+activityDescription = "Combine cell type specific analysis"
 
-thisFileName <- 'Olig.Rmd'
+thisFileName <- 'combine.Rmd'
 
@@
   cache = TRUE)
-
-colorArray = colors()[grep("grey|gray", colors(), invert=TRUE)]
-
 ```
-
-```{r read.counts, eval=FALSE}
-# Read raw counts from MPP and HBCC
-counts.MPP = fread(synGet('syn21867938')$path, data.table=FALSE)
-rownames(counts.MPP) = counts.MPP$Geneid
-counts.MPP = counts.MPP[,-c(1:6)]
-
-# HBCC - DLFPC
-counts.HBCC = fread(synGet('syn21886235')$path, data.table=FALSE)
-rownames(counts.HBCC) = counts.HBCC$Geneid
-counts.HBCC = counts.HBCC[,-c(1:6)]
-
-# # HBCC - sgACC
-counts.HBCC4 = fread(synGet('syn21867866')$path, data.table=FALSE)
-rownames(counts.HBCC4) = counts.HBCC4$Geneid
-counts.HBCC4 = counts.HBCC4[,-c(1:6)]
-
-geneCountsMerged = cbind(counts.MPP, counts.HBCC, counts.HBCC4)
-```
-
-```{r read.cts.expression}
-geneExpr.CTS = readRDS(synGet('syn23234833')$path)
-geneExpr.CTS.se = readRDS(synGet('syn23234858')$path)
-
-geneExpr = geneExpr.CTS[,params$cellType,]
-geneExpr.se = geneExpr.CTS.se[,params$cellType,]
-
-```
-
-### Data download
-#### Obtain effective count matrix and metadata from synapse.
-```{r download.data, cache=TRUE}
-downloadFile <- function(id){
-  fread(synGet(id)$path, data.table = F)
-}
-downloadFile_version <- function(id , version){
-  fread(synGet(id, version = version)$path, data.table = F)
-}
-
-# Get RNASeq QCmetadata
-METADATA_QC_DLPFC_ID = 'syn16816488' 
-ALL_USED_IDs = METADATA_QC_DLPFC_ID
-metadata = downloadFile_version(METADATA_QC_DLPFC_ID, version = 19) %>%
-  dplyr::select(`Individual_ID`, Institution, Cohort, `Reported_Gender`, Sex, Ethnicity, ageOfDeath, `PMI_(in_hours)`, Dx, 
-                Sample_RNA_ID, one_of('rnaSeq_isolation:RIN',
-                                      'rnaSeq_dissection:Brain_Region', 
-                                      'rnaSeq_isolation:RNA_Isolation_Batch',
-                                      'rnaSeq_report:Ribozero_Batch', 'rnaSeq_report:Library_Batch', 
-                                      'rnaSeq_report:Flowcell_Batch', 'rnaSeq_report:Exclude?',
-                                      'rnaSeq_report:Mapped_Reads',
-                                      'rnaSeq_report:Intragenic_Rate','rnaSeq_report:Intronic_Rate',
-                                      'rnaSeq_report:Intergenic_Rate','rnaSeq_report:Genes_Detected',
-                                      'rnaSeq_report:Expression_Profiling_Efficiency',
-                                      'rnaSeq_report:rRNA_Rate',
-                                      'rnaSeq_report:Total_Reads', 'rnaSeq_report:Percent_Aligned','rnaSeq_report:Transcripts_Detected'))
-
-# Get ancestry vector calculated using gemtools
-ANCESTRY_ID = 'syn9922992'
-ALL_USED_IDs = c(ALL_USED_IDs, ANCESTRY_ID)
-ANCESTRY.HBCC = downloadFile(ANCESTRY_ID) %>% 
-  plyr::rename(c(ID = 'SNP_report:Genotyping_Sample_ID'))
-
-ANCESTRY_ID = 'syn2511399'
-ALL_USED_IDs[length(ALL_USED_IDs)+1] = ANCESTRY_ID
-ANCESTRY.MPP = downloadFile(ANCESTRY_ID) %>% 
-  plyr::rename(c('DNA_report..Genotyping.Sample_ID' = 'SNP_report:Genotyping_Sample_ID'))
-
-ANCESTRY = rbind(ANCESTRY.HBCC[,colnames(ANCESTRY.MPP)[-2]], ANCESTRY.MPP[,colnames(ANCESTRY.MPP)[-2]])
-
-# Get genotype ids from synapse. 
-GENOTYPE_ID = 'syn16816490'
-ALL_USED_IDs = c(ALL_USED_IDs, GENOTYPE_ID)
-GENOTYPE = downloadFile(GENOTYPE_ID) %>% 
-  dplyr::select(Individual_ID, `SNP_report:Genotyping_Sample_ID`, `SNP_report:Exclude?`) %>% 
-  dplyr::inner_join(ANCESTRY) %>% 
-  dplyr::filter(is.na(`SNP_report:Exclude?`))
-```
-
-### Data preprocessing
-```{r preprocess.data}
-# Merge metadata
-METADATA = metadata %>%
-  dplyr::left_join(GENOTYPE) %>%      
-  dplyr::rename(Region = `rnaSeq_dissection:Brain_Region`,
-                PMI = `PMI_(in_hours)`,
-                RIN = `rnaSeq_isolation:RIN`, 
-                ReportExclude = `rnaSeq_report:Exclude?`,
-                GenotypeExclude = `SNP_report:Exclude?`,
-                SampleID = Sample_RNA_ID,
-                LibraryBatch = `rnaSeq_report:Library_Batch`,
-                FlowcellBatch = `rnaSeq_report:Flowcell_Batch`,
-                IsolationBatch = `rnaSeq_isolation:RNA_Isolation_Batch`, #isolation batch to replace ribozero batch
-                RibozeroBatch = `rnaSeq_report:Ribozero_Batch`,
-                MappedReads = `rnaSeq_report:Mapped_Reads`,
-                IntragenicRate = `rnaSeq_report:Intragenic_Rate`, IntronicRate = `rnaSeq_report:Intronic_Rate`, 
-                IntergenicRate = `rnaSeq_report:Intergenic_Rate`, GenesDetected = `rnaSeq_report:Genes_Detected`,
-                ExpProfEfficiency = `rnaSeq_report:Expression_Profiling_Efficiency`, rRNARate = `rnaSeq_report:rRNA_Rate`,
-                TotalReads = `rnaSeq_report:Total_Reads`, AlignmentRate = `rnaSeq_report:Percent_Aligned`, TranscriptsDetected = `rnaSeq_report:Transcripts_Detected`) %>%
-  dplyr::select(SampleID, Individual_ID, Institution, Cohort, Reported_Gender,Sex, Ethnicity, ageOfDeath, PMI, Dx, Region,
-                RIN, ReportExclude, GenotypeExclude, EV.1, EV.2, EV.3, EV.4, EV.5,
-                LibraryBatch, FlowcellBatch, IsolationBatch, RibozeroBatch, MappedReads, TotalReads, GenesDetected, AlignmentRate,
-                IntragenicRate, IntergenicRate, IntronicRate, ExpProfEfficiency, rRNARate, TranscriptsDetected) %>%
-  		 		 dplyr::filter(SampleID %in% colnames(geneExpr), !is.na(SampleID)) %>%
-  		 		 dplyr::filter(Region %in% c('DLPFC')) %>%
-			 	 dplyr::mutate(Cohort = forcats::fct_recode(Cohort, `MSSM-Penn-Pitt` = 'MSSM-Penn-Pitt', `NIMH-HBCC`='NIMH-HBCC', `NIMH-HBCC`='NIMH-HBCC-4')) %>%
-			 	 dplyr::filter(Dx %in% c('Control', "SCZ"))  %>%
-         dplyr::filter(ageOfDeath > 17)
-
-# Set Control to baseline
-METADATA$Dx = factor(METADATA$Dx, c('Control', 'SCZ'))
-
-ind = METADATA$SampleID[which(METADATA$ReportExclude == 1 | METADATA$GenotypeExclude)]
-writeLines(paste('Following',length(ind),'samples are marked exclude'))
-writeLines(paste(ind, collapse = ', '))
-METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))
-
-ind = METADATA$SampleID [is.na(METADATA$Ethnicity) | is.na(METADATA$Institution) | is.na(METADATA$Dx)]
-writeLines(paste('Following', length(ind), 'counts are missing any metadata'))
-writeLines(paste(ind, collapse = ', '))
-METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))
-
-ind = METADATA$SampleID [is.na(METADATA$PMI)]
-writeLines(paste('Following', length(ind), 'counts are missing PMI'))
-writeLines(paste(ind, collapse = ', '))
-METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))
-
-ind = METADATA$SampleID [is.na(METADATA$Reported_Gender)]
-writeLines(paste('Following', length(ind), 'counts are missing gender'))
-writeLines(paste(ind, collapse = ', '))
-METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))
-
-ind = METADATA$SampleID [is.na(METADATA$ageOfDeath)]
-writeLines(paste('Following', length(ind), 'counts are missing age of death'))
-writeLines(paste(ind, collapse = ', '))
-METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))
-
-ind = METADATA$SampleID [is.na(METADATA$EV.1)]
-writeLines(paste('Following', length(ind), 'counts are missing ancestry information'))
-writeLines(paste(ind, collapse = ', '))
-METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))
-
-# drop individuals who where sequenced twice here
-tab = table(METADATA$Individual_ID)
-
-dropSamples = sapply( names(tab[tab > 1]), function(id){
-
-  idx = which(METADATA$Individual_ID == id)
-  df = METADATA[idx,]
-
-  # remove individuals that have less then max reads
-  i = which(df$MappedReads < max(df$MappedReads))
-  rownames(METADATA)[idx[i]]
-  } )
-METADATA = METADATA[!(rownames(METADATA) %in% dropSamples),]
-```
-
-```{r cellFractionEstimates}
-# include estimated cell fractions in METADATA
-df_cellFractions = read.table(synGet('syn22333694')$path, row.names=1)
-
-celFrac_ilr = ilr(df_cellFractions)
-colnames(celFrac_ilr) = paste0("cellFrac_ilr_", 1:3)
-
-METADATA = merge(METADATA, df_cellFractions, by.x = 'SampleID', by.y='row.names')
-METADATA = merge(METADATA, celFrac_ilr, by.x = 'SampleID', by.y='row.names')
-```
-
-```{r preprocess.data1, results='asis'}
-# Match covariates to expression data
-indToRetain = intersect(METADATA$SampleID, colnames(geneExpr))
-
-geneExpr = geneExpr[,indToRetain]
-
-rownames(METADATA) = METADATA$SampleID
-```
-
-```{r create.voom}
-vobj.lst = lapply( levels(METADATA$Cohort), function(chrt){
-
-  idx = METADATA$Cohort == chrt
-
-  # get variance withing cohort
-  # only keep genes with variance > 0
-  v = apply( geneExpr[,idx], 1, var)
-
-  E = geneExpr[v > 0,idx]
-
-  new("EList", list(E = E, weights = 1/geneExpr.se[rownames(E), colnames(E)]))
-  })
-names(vobj.lst) = levels(METADATA$Cohort)
-
-# only keep genes common to both cohorts
-isexpr = intersect(rownames(vobj.lst[[1]]), rownames(vobj.lst[[2]]))
-
-vobj.lst = lapply(vobj.lst, function(vobj){
-  vobj[isexpr,]
-  })
-
-```{r Create.residuals}
-
-### Create residuals
-form.lst = list(
-  `MSSM-Penn-Pitt`  = ~ Dx*Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*ageOfDeath,
-  `NIMH-HBCC`       = ~ Dx*Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath)
-
-eval_residuals = function( form ){
-  # Compute residuals, but add back in Dx*Reported_Gender
-  # for form = ~ Dx*Reported_Gender
-  f = function(fit){
-    residuals(fit) + get_prediction( fit, form)
-  }
-
-  residuals.lst = lapply( vobj.lst, function(vobj){
-
-    i = match(colnames(vobj), rownames(METADATA))
-    info = METADATA[i,]
-
-    form_use = form.lst[[info$Cohort[1]]]
-
-    # Compute residuals which adding back Diagnosis and Sex effects
-    resid.lst = fitVarPartModel(vobj, form_use, info, fxn = f, BPPARAM = SnowParam(18))  
-    do.call(rbind, resid.lst)
-    })
-  names(residuals.lst) = names(vobj.lst)
-
-  residuals.lst
-}
-
-resid.add = list( Dx_Sex  = eval_residuals( ~ Dx*Reported_Gender ),
-                  Dx      = eval_residuals( ~ Dx ),
-                  Sex     = eval_residuals( ~ Reported_Gender ))
-
-residuals.lst = resid.add[['Dx_Sex']]
-```
-
-```{r check.correlation, eval=FALSE}
-ngenes = 300
-C1 = cor(t(residuals.lst[[1]][1:ngenes,]))
-C2 = cor(t(residuals.lst[[2]][1:ngenes,]))
-
-df = data.table(A   = as.numeric(C1[lower.tri(C1)]),
-                B   = as.numeric(C2[lower.tri(C2)]))
-
-df$density <- get_density(df$A, df$B, n = 80)
-ggplot(df, aes(A, B, color=density)) + geom_point() + geom_point(size=.4) + theme_bw(15) + theme(aspect.ratio=1, legend.position="bottom", plot.title = element_text(hjust = 0.5)) + geom_abline(color="red") + scale_color_viridis() + xlim(-1,1) + ylim(-1,1) + xlab(paste0('Correlation (', names(residuals.lst)[1], ')')) + ylab(paste0('Correlation (', names(residuals.lst)[2], ')'))
-```
-
-# Schizophrenia vs controls
-
-### Test marginals for Scz vs controls
-```{r differential.expression.simple}
-form.lst = list(
-  `MSSM-Penn-Pitt`  = ~ Dx + Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*ageOfDeath,
-  `NIMH-HBCC`       = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath)
-
-fit_marginal.lst = lapply( vobj.lst, function(vobj){
-
-  i = match(colnames(vobj), rownames(METADATA))
-  info = METADATA[i,]
-
-  form_use = form.lst[[info$Cohort[1]]]
-
-  fit = dream(vobj, form_use, info, BPPARAM = SnowParam(18))  
-  eBayes( fit )
-  })
-names(fit_marginal.lst) = names(vobj.lst)
-
-coefTest = c('DxSCZ', 'Reported_GenderMale', 'ageOfDeath')
-```
-
-```{r load.genesets, eval=TRUE}
-# Load Gene Ontology database 
-gs.GO = get_GeneOntology()
-gs.MSigDB = get_MSigDB()
-
-# convert from GeneSetCollection to list used by camera and zenith
-gs.GO.collection = recodeToList( gs.GO )
-gs.MSigDB.collection = recodeToList( gs.MSigDB )
-```
-
-```{r run.zenith, eval=TRUE}
-res_zenith = lapply( names(fit_marginal.lst), function(key){
-
-  fit = fit_marginal.lst[[key]]
-
-  res.GO = run_zenith(  fit, 
-                        coefTest,
-                        gs.GO.collection)
-  res.GO$Resource = "GO"   
-  res.GO$Cohort = key
-
-  res.MSigDB = run_zenith(fit, 
-                          coefTest,
-                          gs.MSigDB.collection ) 
-  res.MSigDB$Resource = "MSigDB"
-  res.MSigDB$Cohort = key
-
-  rbind(res.GO, res.MSigDB)
-})
-names(res_zenith) = names(fit_marginal.lst)
-
-# only autosomes
-res_zenith_sex_autosomes = lapply( names(fit_marginal.lst), function(key){
-
-  fit = fit_marginal.lst[[key]]
-
-  # only keep autosomes, for Reported_GenderMale
-  tab = coef(fit) %>% getGeneSymbol()
-  geneids = rownames(tab)[tab$Chrom %in% 1:22]
-  fit = fit[geneids,]
-
-  res.GO = run_zenith(  fit, 
-                        'Reported_GenderMale',
-                        gs.GO.collection)
-  res.GO$Resource = "GO"   
-  res.GO$Cohort = key
-
-  res.MSigDB = run_zenith(fit, 
-                          'Reported_GenderMale',
-                          gs.MSigDB.collection ) 
-  res.MSigDB$Resource = "MSigDB"
-  res.MSigDB$Cohort = key
-
-  rbind(res.GO, res.MSigDB)
-})
-names(res_zenith_sex_autosomes) = names(fit_marginal.lst)
-```
-
-## Meta-analysis of zenith results
-```{r zenith.meta, eval=TRUE}
-res_meta_zenith = zenith_meta_analysis( res_zenith )                  
-res_meta_zenith.sex_autosome = zenith_meta_analysis( res_zenith_sex_autosomes )
-```
-
-```{r get.signatures, fig.width=10, fig.height=5}
-# Get DE signatures
-df_DE_results = lapply(fit_marginal.lst, function(fit){
-  res = lapply(coefTest, function(coef){
-    res = topTable(fit, coef, number=Inf) 
-    res$Coef = coef
-    res %>% getGeneSymbol()  
-    })
-  names(res) = coefTest
-  res
-  })
-names(df_DE_results) = names(fit_marginal.lst)
-```
-
-```{r test.scz, fig.height=5, fig.width=10}
-# Schizophrenia testing
-# extract from each cohort
-coef = 'DxSCZ'
-# ymin = min(sapply(df_DE_results, function(x) min(x[[coef]]$P.Value)))
-
-figList.Scz = lapply( names(df_DE_results), function(key){
-  df = df_DE_results[[key]][[coef]]
-
-  main = paste(key, coef, sep = ': ' )
-  plotVolcano( df, df$Symbol[1:10] ) + ggtitle( main ) #+ ylim(0, -log10(ymin))
-})
-
-# volcano plots
-plot_grid(plotlist=figList.Scz, nrow=1)
-```
-
-## Meta-analysis across cohorts
-```{r meta.analysis.DE}
-# Store results for each coef and cohort
-tab.de.lst = lapply(coefTest, function(coef){    
-  tab.lst = lapply( df_DE_results, function(lst){
-    lst[[coef]]
-  })
-  names(tab.lst) = names(df_DE_results)
-  tab.lst
-  })
-names(tab.de.lst) = coefTest
-
-# Create list preparing for meta-analysis for SCZ
-df_meta = lapply(coefTest, function(coef){
-  # Run meta-analysis
-  run_meta_analysis( tab.de.lst[[coef]], "FE") %>% getGeneSymbol
-})#, mc.cores=6)
-names(df_meta) = coefTest
-```
-
-```{r concordance.scz, fig.height=6, fig.width=10}
-# Concordance analysis
-resConcord.SCZ = plot_concordance(tab.de.lst[[coef]], showGenes = df_meta[[coef]]$Symbol[1:20])
-
-main = "Concordance of SCZ"
-plot_grid(  resConcord.SCZ$fig_logFC + ggtitle(main),
-            resConcord.SCZ$fig_tstat + ggtitle(main), nrow=1)
-
-data.frame(Cohort = names(tab.de.lst[[coef]]),
-          "# DE"  = resConcord.SCZ$n.de,
-          pi1     = resConcord.SCZ$pi_single,
-          check.names=FALSE) %>% 
-          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)
-
-resConcord.SCZ$tab_pi %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
-resConcord.SCZ$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
-```
-
-```{r plot.meta.analysis.SCZ, fig.height=5, fig.width=10}
-# volcano plot for meta-analysis
-main = paste('Meta-analysis', coef, sep = ': ' )
-plotVolcano( df_meta[[coef]], df_meta[[coef]]$Symbol[1:20] ) + ggtitle( main )
-
-data.frame(pi1 = 1 - qvalue(df_meta[[coef]]$P)$pi0,
-          `# DE` = sum(df_meta[[coef]]$adj.P.Val < 0.05),
-          check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for SCZ meta-analysis") %>% kable_styling(full_width=FALSE)
-
-```
-
-## Plot top examples
-```{r plot.forrest.dlpfc, fig.width=8, fig.height=20}  
-plot_gene = function(residuals.lst, df_meta, geneSymbol, variable, fill=c("#1E90FF", "#FF8D1E"), mode=""){
-
-  ensGene = rownames(df_meta)[which(df_meta$Symbol == geneSymbol)]
-
-  exprResid = lapply( names(residuals.lst), function(key){
-      data.frame(Expression = residuals.lst[[key]][ensGene,])
-    })
-  exprResid = do.call(rbind, exprResid)
-
-  df = merge(exprResid, METADATA, by="row.names")
-
-  if( is.factor(df[[variable]]) || length(unique(df[[variable]])) < 5 ){
-    fig = ggplot(df, aes_string(variable, 'Expression')) + geom_violin(aes_string(fill=variable)) + geom_boxplot(width=0.1, fill="grey80") + scale_fill_manual(values = fill)
-  }else{    
-    fig = ggplot(df, aes_string(variable, 'Expression')) + geom_point() + geom_smooth(method="lm") #+ scale_color_manual(values = fill)
-  }
-
-  fig + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + ylab('Expression residuals') + ggtitle(geneSymbol) + facet_wrap(~Cohort)
-}
-
-coef = 'DxSCZ'
-figList = lapply(df_meta[[coef]]$Symbol[1:10], function(geneSymbol){
-  fig1 = plot_forrest( c(tab.de.lst[[coef]], list("Meta-analysis" = df_meta[[coef]])), geneSymbol)
-
-  fig2 = plot_gene(residuals.lst, df_meta[[coef]], geneSymbol, "Dx") + xlab("Diagnosis")
-
-  plot_grid(fig1, fig2, nrow=1, rel_widths=c(.8, 1))
-})
-
-plot_grid(plotlist=figList, ncol=1)
-```
-
-### Concordance with other datasets
-```{r CMC.concordance}
-# CMC Analysis 1
-CMC1 <- readr::read_tsv(synGet("syn5607603")$path) %>% 
-  dplyr::rename(gene_id = genes,
-                `CMC 1 t-statistic` = t) %>% 
-  dplyr::select(gene_id, `CMC 1 t-statistic`)
-# HBCC-DLPFC microarray
-CMC_HBCC <- readr::read_tsv(synGet("syn4941553")$path) %>% 
-  dplyr::rename(gene_id = genes,
-                `HBCC microarray t-statistic` = t) %>% 
-  dplyr::select(gene_id, `HBCC microarray t-statistic`)
-# Gandal, Science 2018, Table S1
-# Use Tab:DGE
-# wget https://science.sciencemag.org/highwire/filestream/720110/field_highwire_adjunct_files/1/aat8127_Table_S1.xlsx
-gandal <- readxl::read_xlsx( synGet("syn22255217")$path, sheet = "DGE") %>% 
-  dplyr::rename(gene_id = ensembl_gene_id,
-                `Gandal, SCZ t-statistic` = `SCZ.t.value`,
-                `Gandal, BD t-statistic` = `BD.t.value`,
-                `Gandal, ASD t-statistic` = `ASD.t.value`) %>%
-  dplyr::select(gene_id, `Gandal, SCZ t-statistic`,
-                `Gandal, BD t-statistic`, `Gandal, ASD t-statistic`)
-# AMP-AD RNASeq reprocessing GRCh37 - https://www.synapse.org/#!Synapse:syn9702085
-MSSM <- readr::read_tsv(synGet("syn10526259")$path) %>%
-  dplyr::filter(Comparison %in% "AD-CONTROL",
-                Model %in% "Diagnosis") %>%
-  dplyr::select(ensembl_gene_id, Tissue, t) %>%
-  dplyr::group_by(ensembl_gene_id, Tissue) %>% 
-  dplyr::filter(row_number() == 1) %>% 
-  tidyr::pivot_wider(names_from = Tissue,
-                     values_from = t) %>%
-  dplyr::rename(`AMP-AD MSSM FP AD t-statistic` = FP,
-                `AMP-AD MSSM IGF AD t-statistic` = IFG,
-                `AMP-AD MSSM PHG AD t-statistic` = PHG,
-                `AMP-AD MSSM STG AD t-statistic` = STG,
-                gene_id = ensembl_gene_id) %>% 
-  dplyr::select(gene_id, `AMP-AD MSSM STG AD t-statistic`) 
-MSSM <- MSSM[!is.na(MSSM$`AMP-AD MSSM STG AD t-statistic`),]
-
-MAYO <- readr::read_tsv(synGet("syn8468023")$path) %>% 
-  dplyr::filter(Comparison %in% "AD-CONTROL",
-                Model %in% "Diagnosis") %>%
-  dplyr::select(ensembl_gene_id, `Tissue.ref`, t) %>%
-  dplyr::group_by(ensembl_gene_id, `Tissue.ref`) %>% 
-  dplyr::filter(row_number() == 1) %>% 
-  tidyr::pivot_wider(names_from = `Tissue.ref`,
-                     values_from = t) %>% 
-  dplyr::rename(`AMP-AD MAYO CBE AD t-statistic` = CBE,
-                `AMP-AD MAYO TCX AD t-statistic` = TCX,
-                gene_id = ensembl_gene_id) %>% 
-  dplyr::select(-`AMP-AD MAYO TCX AD t-statistic`)
-
-ROSMAP <- readr::read_tsv(synGet("syn8456721")$path) %>%
-    dplyr::filter(Comparison %in% "AD-CONTROL",
-                Model %in% "Diagnosis") %>%
-  dplyr::rename(gene_id = ensembl_gene_id, 
-                `AMP-AD ROSMAP DLPFC AD t-statistic` = t) %>% 
-  dplyr::select(gene_id, `AMP-AD ROSMAP DLPFC AD t-statistic`)
-
-out_group <- inner_join(MSSM, MAYO) %>% 
-  inner_join(., ROSMAP) %>% 
-  inner_join(., gandal)
-
-# merge with meta-analysis
-
-coef = 'DxSCZ'
-df = df_meta[[coef]]
-df_two = data.frame(gene_id = trim_ensembl_ids(rownames(df)),
-                `CMC - meta-analysis` = with(df, logFC/se), check.names=FALSE)
-
-df = tab.de.lst[[coef]][[1]]
-df.tmp = data.frame(gene_id = trim_ensembl_ids(rownames(df)),
-                `t` = df$t, check.names=FALSE)
-colnames(df.tmp)[2] = names(tab.de.lst[[coef]])[1]
-df_two = df_two %>% inner_join(., df.tmp)
-
-df = tab.de.lst[[coef]][[2]]
-df.tmp = data.frame(gene_id = trim_ensembl_ids(rownames(df)),
-                `t` = df$t, check.names=FALSE)
-colnames(df.tmp)[2] = names(tab.de.lst[[coef]])[2]
-df_two = df_two %>% inner_join(., df.tmp)
-
-# Remove transcript Ids
-df_two <- df_two %>% 
-  tidyr::separate(gene_id, c("gene_id", "transcript")) %>% 
-  dplyr::select(-one_of("transcript"))
-df_two <- inner_join(df_two, CMC1, by = c("gene_id")) 
-df_two <- inner_join(df_two, CMC_HBCC, by = c("gene_id"))
-df_two <- inner_join(df_two, out_group, by = c("gene_id")) %>% 
-  group_by(gene_id) %>% 
-  dplyr::filter(row_number()==1) %>% 
-  tibble::column_to_rownames(var = "gene_id")
-
-colnames(df_two) <- gsub(" t-statistic", "", colnames(df_two))
-
-col2 <- colorRampPalette(rev(c("#67001F", "#B2182B", "#D6604D", "#F4A582",
-                           "#FDDBC7", "#FFFFFF", "#D1E5F0", "#92C5DE",
-                           "#4393C3", "#2166AC", "#053061")))
-
-cor <- cor(df_two)
-# clustering based on distance
-hcl = hclust(as.dist(1-cor), method='ward.D2')
-
-results <- pvclust(cor, method.hclust='ward.D2')
-# reorder based on clustering
-reorder = cor[results$hclust$order, results$hclust$order]
-
-# plot clustering
-plot(results, hang=-1)
-# plot(hcl, hang=-1)
-corrplot(reorder, col = col2(200))
-```
-
-```{r show.gene_set_test.marginal, eval=TRUE}
-show_zenith_results = function( res, coef, Resource, FDR = 0.05){
-
-  i = which((res$Resource == Resource) & (res$Coef == coef))
-  df = res[i,]
-
-	# filter by FDR
-	df = df[df$FDR <= FDR,]
-  df$Coef = c()
-  df$se = c()
-  df$Resource = c()
-
-	df %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption=paste(Resource, coef), row.names=FALSE) %>% kable_styling(full_width=FALSE)
-}
-
-show_zenith_results( res_meta_zenith, "DxSCZ", "GO", FDR=0.05)
-show_zenith_results( res_meta_zenith, "DxSCZ", "MSigDB", FDR=0.05)
-```
-
-```{r plot.gene_set_test.marginal, eval=TRUE}
- plot_zenith_results = function( res, coef, Resource, FDR = 0.05){
-
-  i = which((res$Resource == Resource) & (res$Coef == coef))
-  df = res[i,]
-
-  # filter by FDR
-  df = df[df$FDR <= FDR,]
-  df$Coef = c()
-  df$se = c()
-  df$Resource = c()
-
-  df = df[order(df$PValue, decreasing=TRUE),]
-  df$Set = factor(df$Set, df$Set)
-  ymax = max(-log10(df$FDR)) * 1.05
-  ggplot(df, aes(Set, -log10(FDR))) + geom_bar(stat="identity", fill="navy") + theme_bw() + theme(aspect.ratio=1,plot.title = element_text(hjust = 0.5)) + coord_flip() + ylab(bquote(-log[10]~FDR)) + xlab('') + ggtitle(coef) + scale_y_continuous(expand=c(0,0), limits=c(0, ymax)) + geom_hline(yintercept=-log10(0.05), color="red", linetype="dashed")
-}
-
- plot_zenith_results( res_meta_zenith, "DxSCZ", "GO", FDR=0.05)
-```
-
-```{r, fig.width=9}
- plot_zenith_results( res_meta_zenith, "DxSCZ", "MSigDB", FDR=0.05)
-```
-
-# Sex Analysis
-```{r test.Sex, fig.height=5, fig.width=10}
-# Schizophrenia testing
-# extract from each chort
-coef = 'Reported_GenderMale'
-figList.Sex = lapply( names(df_DE_results), function(key){
-  df = df_DE_results[[key]][[coef]]
-  main = paste(key, coef, sep = ': ' )
-  plotVolcano( df, df$Symbol[1:20] ) + ggtitle( main )
-})
-
-# volcano plots
-plot_grid(plotlist=figList.Sex, nrow=1)
-```
-
-### Autosomes
-```{r test.Sex.autosome, fig.height=5, fig.width=10}
-# Schizophrenia testing
-# extract from each chort
-coef = 'Reported_GenderMale'
-figList.Sex = lapply( names(df_DE_results), function(key){
-  df = df_DE_results[[key]][[coef]]
-  main = paste(key, coef, sep = ': ' )
-  df = df[df$Chrom %in% 1:22,]
-  plotVolcano( df, df$Symbol[1:20] ) + ggtitle( main )
-})
-
-# volcano plots
-plot_grid(plotlist=figList.Sex, nrow=1)
-```
-
-## Meta-analysis across cohorts
-```{r concordance.Sex, fig.height=6, fig.width=10}
-# Concordance analysis
-resConcord.Sex = plot_concordance(tab.de.lst[[coef]], showGenes = df_meta[[coef]]$Symbol[1:20])
-
-main = "Concordance of Sex"
-plot_grid(  resConcord.Sex$fig_logFC + ggtitle(main),
-            resConcord.Sex$fig_tstat + ggtitle(main), nrow=1)
-
-data.frame(Cohort = names(tab.de.lst[[coef]]),
-          "# DE"  = resConcord.Sex$n.de,
-          pi1     = resConcord.Sex$pi_single,
-          check.names=FALSE) %>% 
-          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)
-
-resConcord.Sex$tab_pi  %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
-resConcord.Sex$tab_corr  %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
-```
-
-```{r concordance.Sex.autosome, fig.height=6, fig.width=10}
-
-tab.de.lst.autosome = lapply(tab.de.lst[[coef]], function(x) x[x$Chrom %in% 1:22,])
-
-# Concordance analysis
-df = df_meta[[coef]][df_meta[[coef]]$Chrom %in% 1:22,]
-resConcord.Sex = plot_concordance(tab.de.lst.autosome, showGenes = df$Symbol[1:20])
-
-main = "Concordance of Sex - autosomes"
-plot_grid(  resConcord.Sex$fig_logFC + ggtitle(main),
-            resConcord.Sex$fig_tstat + ggtitle(main), nrow=1)
-
-data.frame(Cohort = names(resConcord.Sex$pi_single),
-          "# DE"  = resConcord.Sex$n.de,
-          pi1     = resConcord.Sex$pi_single,
-          check.names=FALSE) %>% 
-          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)
-
-resConcord.Sex$tab_pi %>% mutate_if(is.numeric, format, digits=3)  %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
-resConcord.Sex$tab_corr  %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
-```
-
-```{r plot.meta.analysis.Sex, fig.height=5, fig.width=10}
-# volcano plot for meta-analysis
-main = 'Meta-analysis: Sex'
-fig1 = plotVolcano( df_meta[[coef]], df_meta[[coef]]$Symbol[1:20] ) + ggtitle( main )
-
-df = df_meta[[coef]][df_meta[[coef]]$Chrom %in% 1:22,]
-fig2 = plotVolcano( df, df$Symbol[1:20] ) + ggtitle( paste(main, "autosome") )
-plot_grid(fig1, fig2, ncol=2)
-
-data.frame(pi1 = 1 - qvalue(df_meta[[coef]]$P)$pi0,
-          `# DE` = sum(df_meta[[coef]]$adj.P.Val < 0.05),
-          check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for Sex meta-analysis") %>% kable_styling(full_width=FALSE)
-
-data.frame(pi1 = 1 - qvalue(df$P)$pi0,
-          `# DE` = sum(df$adj.P.Val < 0.05),
-          check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for Sex meta-analysis - autosomes") %>% kable_styling(full_width=FALSE)
-```
-
-## Sex circos
-```{r sex.circos}
-coef = 'Reported_GenderMale'
-sapply( names(df_DE_results), function(key){
-  plot_circos_logFC( df_DE_results[[key]][[coef]] ) + ggtitle(key)
-  })
-
-plot_circos_logFC( df_meta[[coef]] ) + ggtitle("Meta analysis")
-```
-
-## Plot top examples - not chrY
-```{r plot.forrest.dlpfc.Sex.autosomes, fig.width=8, fig.height=20}  
-Symbols = df_meta[[coef]][df_meta[[coef]]$Chrom !='Y',]$Symbol[1:10]
-figList = lapply(Symbols, function(geneSymbol){
-
-  fig1 = plot_forrest( c(tab.de.lst[[coef]], list("Meta-analysis" = df_meta[[coef]])), geneSymbol)
-
-  fig2 = plot_gene( residuals.lst, df_meta[[coef]], geneSymbol, "Reported_Gender", fill=c("red", "blue")) + xlab("Sex")
-
-  plot_grid(fig1, fig2, nrow=1)
-})
-
-plot_grid(plotlist=figList, ncol=1)
-```
-
-## Plot top examples - autosomes
-```{r plot.forrest.dlpfc.Sex, fig.width=8, fig.height=20}  
-Symbols = df_meta[[coef]][df_meta[[coef]]$Chrom %in% 1:22,]$Symbol[1:10]
-
-figList = lapply(Symbols, function(geneSymbol){
-
-  fig1 = plot_forrest( c(tab.de.lst[[coef]], list("Meta-analysis" = df_meta[[coef]])), geneSymbol)
-
-  fig2 = plot_gene( residuals.lst, df_meta[[coef]], geneSymbol, "Reported_Gender", fill=c("red", "blue")) + xlab("Sex")
-
-  plot_grid(fig1, fig2, nrow=1)
-})
-
-plot_grid(plotlist=figList, ncol=1)
-```
-
-## Meta-analysis of zenith results
-```{r show.gene_set_test.marginal.Sex, eval=TRUE}
-show_zenith_results( res_meta_zenith, "Reported_GenderMale", "GO", FDR=0.05)         
-show_zenith_results( res_meta_zenith, "Reported_GenderMale", "MSigDB", FDR=0.05)        
-```
-
-```{r plot.gene_set_test.marginal.Sex, eval=TRUE}
- plot_zenith_results( res_meta_zenith, "Reported_GenderMale", "GO", FDR=0.05)               
- plot_zenith_results( res_meta_zenith, "Reported_GenderMale", "MSigDB", FDR=0.05)        
-```
-
-### Only autosomes
-```{r show.gene_set_test.marginal.Sex.autosome, eval=TRUE}
-show_zenith_results( res_meta_zenith.sex_autosome, "Reported_GenderMale", "GO", FDR=0.1) 
-show_zenith_results( res_meta_zenith.sex_autosome, "Reported_GenderMale", "MSigDB", FDR=0.1)   
-```
-
-```{r plot.gene_set_test.marginal.Sex.autosome, eval=TRUE}
- plot_zenith_results( res_meta_zenith.sex_autosome, "Reported_GenderMale", "GO", FDR=0.1) 
- plot_zenith_results( res_meta_zenith.sex_autosome, "Reported_GenderMale", "MSigDB", FDR=0.1)   
-```
-
-# Age Analysis
-```{r test.Age, fig.height=5, fig.width=10}
-# Schizophrenia testing
-# extract from each chort
-coef = 'ageOfDeath'
-figList.Age = lapply( names(df_DE_results), function(key){
-  df = df_DE_results[[key]][[coef]]
-  main = paste(key, coef, sep = ': ' )
-  plotVolcano( df, df$Symbol[1:20] ) + ggtitle( main )
-})
-
-# volcano plots
-plot_grid(plotlist=figList.Age, nrow=1)
-```
-
-## Meta-analysis across cohorts
-```{r concordance.Age, fig.height=6, fig.width=10}
-# Concordance analysis
-resConcord.Age = plot_concordance(tab.de.lst[[coef]], showGenes = df_meta[[coef]]$Symbol[1:20])
-
-main = "Concordance of Age"
-plot_grid(  resConcord.Age$fig_logFC + ggtitle(main),
-            resConcord.Age$fig_tstat + ggtitle(main), nrow=1)
-
-data.frame(Cohort = names(tab.de.lst[[coef]]),
-          "# DE"  = resConcord.Age$n.de,
-          pi1     = resConcord.Age$pi_single,
-          check.names=FALSE) %>% 
-          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)
-
-resConcord.Age$tab_pi %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
-resConcord.Age$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
-```
-
-```{r plot.meta.analysis.Age, fig.height=5, fig.width=10}
-# volcano plot for meta-analysis
-main = 'Meta-analysis: Age'
-plotVolcano( df_meta[[coef]], df_meta[[coef]]$Symbol[1:20] ) + ggtitle( main )
-
-data.frame(pi1 = 1 - qvalue(df_meta[[coef]]$P)$pi0,
-          `# DE` = sum(df_meta[[coef]]$adj.P.Val < 0.05),
-          check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for Age meta-analysis") %>% kable_styling(full_width=FALSE)
-```
-
-## Plot top examples
-```{r plot.forrest.dlpfc.Age, fig.width=8, fig.height=20}  
-figList = lapply(df_meta[[coef]]$Symbol[1:10], function(geneSymbol){
-
-  fig1 = plot_forrest( c(tab.de.lst[[coef]], list("Meta-analysis" = df_meta[[coef]])), 
-    geneSymbol)
-
-  expr.list = lapply(vobj.lst, function(x)x$E)
-
-  fig2 = plot_gene( expr.list, df_meta[[coef]], geneSymbol, "ageOfDeath", fill=c("red", "blue")) + xlab("Age")
-
-  plot_grid(fig1, fig2, nrow=1)
-})
-
-plot_grid(plotlist=figList, ncol=1)
-```
-
-## Meta-analysis of zenith results
-```{r show.gene_set_test.marginal.Age, eval=TRUE}
-show_zenith_results( res_meta_zenith, "ageOfDeath", "GO", FDR=0.01)   
-show_zenith_results( res_meta_zenith, "ageOfDeath", "MSigDB", FDR=0.01)    
-```
-
-```{r plot.gene_set_test.marginal.Age, eval=TRUE}
- plot_zenith_results( res_meta_zenith, "ageOfDeath", "GO", FDR=0.01)   
- plot_zenith_results( res_meta_zenith, "ageOfDeath", "MSigDB", FDR=0.01)    
-```
-
-# Comparison with GTEx
-```{r gtex.sex, fig.height=6, fig.width=10}
-
-# read beta and se from GTEx
-df_gtex.beta = read.table(synGet("syn22891761")$path, header=TRUE, row.names=1)
-df_gtex.se = read.table(synGet("syn22891767")$path, header=TRUE, row.names=1)
-df_gtex.lfsr = read.table(synGet("syn22891755")$path, header=TRUE, row.names=1)
-df_signif = fread(synGet("syn22891771")$path) 
-df_signif = df_signif[tissue =='Brain_Frontal_Cortex_BA9', ]
-# identical(rownames(df_gtex.beta), rownames(df_gtex.se))
-# identical(rownames(df_gtex.lfsr), rownames(df_gtex.se))
-
-# data.frame of GTEx results
-df_gtex_BA9 = data.frame( ENSEMBL.full = rownames(df_gtex.beta),
-                          ENSEMBL = trim_ensembl_ids( rownames(df_gtex.beta) ),
-                          logFC = -1*df_gtex.beta[,'Brain_Frontal_Cortex_BA9'],
-                          se = df_gtex.se[,'Brain_Frontal_Cortex_BA9'],
-                          adj.P.Val = df_gtex.lfsr[,'Brain_Frontal_Cortex_BA9'])
-
-df_gtex_BA9$isSignif = FALSE
-df_gtex_BA9$isSignif[ df_gtex_BA9$ENSEMBL.full %in% df_signif$gene] = TRUE
-
-df_gtex_BA9 = df_gtex_BA9 %>% getGeneSymbol( "ENSEMBL")
-
-df_gtex_BA9$t = with(df_gtex_BA9, logFC / se)
-df_gtex_BA9$P.Value = with(df_gtex_BA9, 2*pnorm(abs(t), lower.tail=FALSE))
-# df_gtex_BA9$adj.P.Val = p.adjust(df_gtex_BA9$P.Value, "fdr")
-df_gtex_BA9 = df_gtex_BA9[!is.na(df_gtex_BA9$logFC),]
-rownames(df_gtex_BA9) = df_gtex_BA9$ENSEMBL
-
-# data.frame of CMC results from meta-analysis
-df = df_meta[['Reported_GenderMale']]
-df$t = with(df, logFC / se)
-rownames(df) = trim_ensembl_ids( rownames(df) )
-
-# all genes
-resConcord.gtex = plot_concordance( list('CMC v3' = df, 'GTEx BA9' = df_gtex_BA9), showGenes=df$Symbol[1:10] )
-
-main = "Concordance with GTEx (all genes)"
-plot_grid(  resConcord.gtex$fig_logFC + ggtitle(main),
-            resConcord.gtex$fig_tstat + ggtitle(main), nrow=1)
-
-# resConcord.gtex$tab_pi %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
-resConcord.gtex$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
-
-# because GTEx uses MASH to combine 44 tissue, standard plot_concordance() does give 
-# correct replication values
-df.merge = merge(df, df_gtex_BA9, by="row.names")
-
-table(df.merge$isSignif)
-
-df_repl = data.frame(Discovery = c("GTEx", "CMC"), rate = c(NA, NA))
-
-# Discover in CMC
-# with(df.merge, table(isSignif[adj.P.Val.x < 0.05]))
-# with(df, 1 - qvalue(P.Value.y[adj.P.Val.x < 0.05])$pi0)
-lfsr = with(df.merge, adj.P.Val.y[adj.P.Val.x < 0.05]) 
-df_repl$rate[1] = sum(1-lfsr) / length(lfsr)
-
-# Discover in GTEx
-# with(df, 1 - qvalue(P.Value.x[adj.P.Val.y < 0.05])$pi0)
-# df_repl$rate[2] = with(df.merge, 1 - qvalue(P.Value.x[isSignif])$pi0)
-
-res = with(df.merge, ash( logFC.x, se.x))
-df.merge$lsfr = res$result$lfsr
-
-lfsr = with(df.merge, lsfr[adj.P.Val.x < 0.05]) 
-df_repl$rate[2] = sum(1-lfsr) / length(lfsr)
-
-df_repl %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)
-
-# how any DEGs?
-# sum(apply(df_gtex.lfsr, 1, function(x) any(x <= 0.05, na.rm=TRUE)))
-
-# df_combine = apply(df_gtex.lfsr, 2, function(x) sum(x <= 0.05, na.rm=TRUE))
-# df_combine = data.frame(Tissue = names(df_combine), Count = df_combine)
-# df_combine$Tissue = factor(df_combine$Tissue, rev(df_combine$Tissue))
-
-# ggplot(df_combine, aes(Tissue, Count)) + geom_bar(stat="identity") + theme_bw() + coord_flip()
-
-# # All genes: 
-# sum(df_gtex.lfsr[,'Brain_Frontal_Cortex_BA9'] < 0.05, na.rm=TRUE)
-# sum(df_gtex.lfsr[,'Brain_Frontal_Cortex_BA9'] < 0.05, na.rm=TRUE)
-
-# # expressed in CMC
-# sum(df_gtex_BA9$adj.P.Val < 0.05, na.rm=TRUE)
-
-# sum( df$adj.P.Val.y < 0.05)
-
-# # get significant genes directly
-# df_combine = df_signif[,data.frame(Count = length(gene)), by="tissue"]
-# df_combine$tissue = factor(df_combine$tissue, rev(df_combine$tissue))
-
-# ggplot(df_combine, aes(tissue, Count)) + geom_bar(stat="identity") + theme_bw() + coord_flip()
-
-```
-
-```{r gtex.sex.2, fig.height=6, fig.width=10}
-# autosomes
-df = df[df$Chrom %in% 1:22,]
-
-resConcord.gtex = plot_concordance( list('CMC v3' = df, 'GTEx BA9' = df_gtex_BA9), showGenes=df$Symbol[1:10] )
-
-main = "Concordance with GTEx (autosomes)"
-plot_grid(  resConcord.gtex$fig_logFC + ggtitle(main),
-            resConcord.gtex$fig_tstat + ggtitle(main), nrow=1)
-
-# resConcord.gtex$tab_pi %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
-resConcord.gtex$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
-
-df.merge = merge(df, df_gtex_BA9, by="row.names")
-
-table(df.merge$isSignif)
-
-df_repl = data.frame(Discovery = c("GTEx", "CMC"), rate = c(NA, NA))
-
-# Discover in CMC
-# with(df.merge, table(isSignif[adj.P.Val.x < 0.05]))
-# with(df, 1 - qvalue(P.Value.y[adj.P.Val.x < 0.05])$pi0)
-lfsr = with(df.merge, adj.P.Val.y[adj.P.Val.x < 0.05]) 
-df_repl$rate[1] = sum(1-lfsr) / length(lfsr)
-
-# Discover in GTEx
-# with(df, 1 - qvalue(P.Value.x[adj.P.Val.y < 0.05])$pi0)
-# df_repl$rate[2] = with(df.merge, 1 - qvalue(P.Value.x[isSignif])$pi0)
-
-res = with(df.merge, ash( logFC.x, se.x))
-df.merge$lsfr = res$result$lfsr
-
-lfsr = with(df.merge, lsfr[adj.P.Val.x < 0.05]) 
-df_repl$rate[2] = sum(1-lfsr) / length(lfsr)
-
-df_repl %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)
-```
-
-# Sex by Disease interaction
-```{r dream.sex.disease}
-form.lst = list(
-  `MSSM-Penn-Pitt`  = ~ Dx*Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
-  `NIMH-HBCC`       = ~ Dx*Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)
-
-# drop chrY genes
-df_genes  = data.frame(ENSEMBL = rownames(vobj.lst[[1]])) %>% getGeneSymbol('ENSEMBL')
-keep.ensID = df_genes$ENSEMBL[df_genes$Chrom %in% c(1:22, 'X', 'MT')]
-
-fit_inter_sex.disease.lst = lapply( vobj.lst, function(vobj){
-
-  i = match(colnames(vobj), rownames(METADATA))
-  info = METADATA[i,]
-
-  form_use = form.lst[[info$Cohort[1]]]
-
-  fit = dream(vobj[keep.ensID,], form_use, info, BPPARAM = SnowParam(18))  
-  eBayes( fit )
-  })
-names(fit_inter_sex.disease.lst) = names(vobj.lst)
-
-coefTest = c('DxSCZ:Reported_GenderMale')
-```
-
-```{r run.zenith.inter_sex.disease, eval=TRUE}
-res_zenith.inter_sex.disease = lapply( names(fit_inter_sex.disease.lst), function(key){
-
-  fit = fit_inter_sex.disease.lst[[key]]
-
-  res.GO = run_zenith(  fit,   
-                        coefTest,
-                        gs.GO.collection)
-  res.GO$Resource = "GO"   
-  res.GO$Cohort = key
-
-  res.MSigDB = run_zenith(fit, 
-                          coefTest,
-                          gs.MSigDB.collection) 
-  res.MSigDB$Resource = "MSigDB"
-  res.MSigDB$Cohort = key
-
-  rbind(res.GO, res.MSigDB)
-})
-names(res_zenith.inter_sex.disease) = names(fit_inter_sex.disease.lst)
-
-# only autosomes
-res_zenith.inter_sex.disease_autosome = lapply( names(fit_inter_sex.disease.lst), function(key){
-
-  fit = fit_inter_sex.disease.lst[[key]]
-
-  # only keep autosomes, for Reported_GenderMale
-  tab = coef(fit) %>% getGeneSymbol()
-  geneids = rownames(tab)[tab$Chrom %in% 1:22]
-  fit = fit[geneids,]
-
-  res.GO = run_zenith(  fit, 
-                        coefTest,
-                        gs.GO.collection)
-  res.GO$Resource = "GO"   
-  res.GO$Cohort = key
-
-  res.MSigDB = run_zenith(fit, 
-                          coefTest,
-                          gs.MSigDB.collection) 
-  res.MSigDB$Resource = "MSigDB"
-  res.MSigDB$Cohort = key
-
-  rbind(res.GO, res.MSigDB)
-})
-names(res_zenith.inter_sex.disease_autosome) = names(fit_inter_sex.disease.lst)
-```
-
-## Meta-analysis of zenith results
-```{r zenith.meta.inter_sex.disease, eval=TRUE}
-res_meta_zenith.inter_sex.disease = zenith_meta_analysis( res_zenith.inter_sex.disease )  
-res_meta_zenith.inter_sex.disease_autosome = zenith_meta_analysis( res_zenith.inter_sex.disease_autosome )  
-```
-
-```{r get.signatures.inter_sex.disease, fig.width=10, fig.height=5}
-# Get DE signatures
-df_DE_results.inter_sex.disease = lapply(fit_inter_sex.disease.lst, function(fit){
-  res = lapply(coefTest, function(coef){
-    res = topTable(fit, coef, number=Inf) 
-    res$Coef = coef
-    res %>% getGeneSymbol()  
-    })
-  names(res) = coefTest 
-  res
-  })
-names(df_DE_results.inter_sex.disease) = names(fit_inter_sex.disease.lst)
-```
-
-```{r test.scz.inter_sex.disease, fig.height=5, fig.width=10}
-coef = 'DxSCZ:Reported_GenderMale'
-
-figList.Scz = lapply( names(df_DE_results), function(key){
-  df = df_DE_results.inter_sex.disease[[key]][[coef]]
-
-  main = paste(key, coef, sep = ': ' )
-  plotVolcano( df, df$Symbol[1:10] ) + ggtitle( main )
-})
-
-# volcano plots
-plot_grid(plotlist=figList.Scz, nrow=1)
-```
-
-## Meta-analysis across cohorts
-```{r meta.analysis.DE.inter_sex.disease}
-# Store results for each coef and cohort
-tab.de.inter_sex.disease.lst = lapply(coefTest, function(coef){    
-  tab.de.inter_sex.disease.lst = lapply( df_DE_results.inter_sex.disease, function(lst){
-    lst[[coef]]
-  })
-  names(tab.de.inter_sex.disease.lst) = names(df_DE_results.inter_sex.disease)
-  tab.de.inter_sex.disease.lst
-  })
-names(tab.de.inter_sex.disease.lst) = coefTest
-
-# Create list preparing for meta-analysis for SCZ
-df_meta.inter_sex.disease = lapply(coefTest, function(coef){
-  # Run meta-analysis
-  run_meta_analysis( tab.de.inter_sex.disease.lst[[coef]], "FE") %>% getGeneSymbol
-}) 
-names(df_meta.inter_sex.disease) = coefTest  
-```
-
-```{r concordance.scz.inter_sex.disease, fig.height=6, fig.width=10}
-# Concordance analysis.  
-resConcord.inter_sex.disease = plot_concordance(tab.de.inter_sex.disease.lst[[coef]], showGenes = df_meta.inter_sex.disease[[coef]]$Symbol[1:20])
-
-main = "Concordance of DxSCZ:Reported_GenderMale"
-plot_grid(  resConcord.inter_sex.disease$fig_logFC + ggtitle(main),
-            resConcord.inter_sex.disease$fig_tstat + ggtitle(main), nrow=1)
-
-data.frame(Cohort = names(tab.de.inter_sex.disease.lst[[coef]]),
-          "# DE"  = resConcord.inter_sex.disease$n.de,
-          pi1     = resConcord.inter_sex.disease$pi_single,
-          check.names=FALSE) %>% 
-          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)
-
-resConcord.inter_sex.disease$tab_pi %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
-resConcord.inter_sex.disease$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
-```
-
-```{r plot.meta.analysis.SCZ.inter_sex.disease, fig.height=5, fig.width=10}
-# volcano plot for meta-analysis.  
-main = paste('Meta-analysis', coef, sep = ': ' )
-plotVolcano( df_meta.inter_sex.disease[[coef]], df_meta.inter_sex.disease[[coef]]$Symbol[1:20] ) + ggtitle( main )
-
-data.frame(pi1 = 1 - qvalue(df_meta.inter_sex.disease[[coef]]$P)$pi0,
-          `# DE` = sum(df_meta.inter_sex.disease[[coef]]$adj.P.Val < 0.05),
-          `# DE @ 10%` = sum(df_meta.inter_sex.disease[[coef]]$adj.P.Val < 0.1),
-          check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for SCZ:Sex meta-analysis") %>% kable_styling(full_width=FALSE)
-```
-
-## Plot top examples
-```{r plot.forrest.dlpfc.inter_sex.disease, fig.width=8, fig.height=40}
-
-figList = lapply(df_meta.inter_sex.disease[[coef]]$Symbol[1:20], function(geneSymbol){
-  fig1 = plot_forrest( c(tab.de.inter_sex.disease.lst[[coef]], list("Meta-analysis" = df_meta.inter_sex.disease[[coef]])), geneSymbol, plot2='p')
-
-  fig2 = plot_gene(residuals.lst, df_meta.inter_sex.disease[[coef]], geneSymbol, "Dx") + xlab("Diagnosis") + facet_wrap(~Reported_Gender + Cohort, nrow=1)
-
-  plot_grid(fig1, fig2, nrow=1)
-})
-
-plot_grid(plotlist=figList, ncol=1)
-```
-
-## Meta-analysis of zenith results
-```{r show.gene_set_test.marginal.inter_sex.disease, eval=TRUE}
-coef = 'DxSCZ:Reported_GenderMale'   
-show_zenith_results( res_meta_zenith.inter_sex.disease, coef, "GO", FDR=0.1)  
-show_zenith_results( res_meta_zenith.inter_sex.disease, coef, "MSigDB", FDR=0.1)
-
-show_zenith_results( res_meta_zenith.inter_sex.disease_autosome, coef, "GO", FDR=0.1) 
-show_zenith_results( res_meta_zenith.inter_sex.disease_autosome, coef, "MSigDB", FDR=0.1)    
-```
-
-```{r plot.gene_set_test.marginal.inter_sex.disease, eval=TRUE}
-coef = 'DxSCZ:Reported_GenderMale'   
- plot_zenith_results( res_meta_zenith.inter_sex.disease, coef, "GO", FDR=0.1)  
- plot_zenith_results( res_meta_zenith.inter_sex.disease, coef, "MSigDB", FDR=0.1)
-
- plot_zenith_results( res_meta_zenith.inter_sex.disease_autosome, coef, "GO", FDR=0.1) 
- plot_zenith_results( res_meta_zenith.inter_sex.disease_autosome, coef, "MSigDB", FDR=0.1)    
-```
-
-# Dx test separately in Males and then Females
-```{r dream.sex.disease.separately}
-form.lst = list(
-  `MSSM-Penn-Pitt`  = ~ Dx + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
-  `NIMH-HBCC`       = ~ Dx + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)
-
-# DE for males and females separately
-fit.separ = lapply( vobj.lst, function(vobj){
-
-  fit.lst = lapply(unique(METADATA$Reported_Gender), function(Sex){
-
-    i = match(colnames(vobj), rownames(METADATA))
-    info = METADATA[i,]
-    idx = which(info$Reported_Gender == Sex)
-
-    form_use = form.lst[[info$Cohort[1]]]
-
-    fit = dream(vobj[,idx], form_use, info[idx,], BPPARAM = SnowParam(18))  
-    eBayes( fit )
-  })
-  names(fit.lst) = unique(METADATA$Reported_Gender)
-  fit.lst
-  })
-names(fit.separ) = names(vobj.lst)
-
-# meta-analysis
-df_meta.separ = lapply(unique(METADATA$Reported_Gender), function(Sex){
-  tab.lst = lapply( names(vobj.lst), function(chrt){
-    topTable(fit.separ[[chrt]][[Sex]], coef="DxSCZ", sort.by='none', number=Inf)
-  })
-
-  res = run_meta_analysis( tab.lst, "FE") %>% getGeneSymbol
-  res$t = with(res, logFC/se)
-  res
-})
-names(df_meta.separ) = unique(METADATA$Reported_Gender)
-```
-
-```{r separate.test.sz}
-
-df = merge(df_meta.separ[[1]], df_meta.separ[[2]], by="row.names")
-
-df$tstat = with(df, (logFC.x - logFC.y) / sqrt(se.x^2 + se.y^2))
-
-df$p = 2*pnorm(abs(df$tstat), 0, 1, lower.tail=FALSE)
-df$FDR = p.adjust(df$p, "fdr")
-
-head(df[df$FDR < 0.1,])
-
-```
-
-```{r separ.concordance.scz, fig.height=6, fig.width=10}
-# Concordance analysis
-resConcord.separ = plot_concordance( df_meta.separ, showGenes = df_meta[['DxSCZ']]$Symbol[1:20])
-
-main = "Concordance of SCZ across sexes"
-plot_grid(  resConcord.separ$fig_logFC + ggtitle(main),
-            resConcord.separ$fig_tstat + ggtitle(main), nrow=1)
-
-data.frame(Cohort = names(df_meta.separ),
-          "# DE"  = resConcord.separ$n.de,
-          pi1     = resConcord.separ$pi_single,
-          check.names=FALSE) %>% 
-          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)
-
-resConcord.separ$tab_pi %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
-resConcord.separ$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
-```
-
-## Plot top examples
-```{r plot.forrest.dlpfc.separ, fig.width=8, fig.height=40}
-
-figList = lapply(df_meta.separ[[1]]$Symbol[1:20], function(geneSymbol){
-  fig1 = plot_forrest( df_meta.separ, geneSymbol, plot2='p')
-
-  plot_grid(fig1, nrow=1)
-})
-
-plot_grid(plotlist=figList, ncol=1)
-```
-
-```{r store, eval=TRUE}
-if( params$upload ){    
-  # Code
-  # Make sure these match the folder name and parent Id in the first chunk.
-  CODE <- Folder(name = gsub(":", " -",activityName), parentId = parentId)
-  CODE <- synStore(CODE)
-
-  # Store data 
-  storeData = function( name, description){
-    file = paste0(name, "_", params$cellType, ".RDS")
-    saveRDS( eval(parse(text=name)), file=file)
-
-    nEXP_OBJ = File(file, 
-                    name = paste0(params$cellType, ' - ',description), 
-                    parentId = CODE$properties$id)
-    synStore(nEXP_OBJ, used = ALL_USED_IDs, executed = thisFile)
-  }
-
-  # Standard analysis
-  # storeData( "geneCountsMerged", "Merged DLPFC counts for both cohorts")
-  storeData( "METADATA", "Merged metadata used here for both cohorts")
-  # storeData( "isexpr", "Genes used in this analysis")
-  storeData( "vobj.lst", "Voom results for each cohort")
-  storeData( "resid.add", "Residuals correcting for covariates and add back - (Dx_Sex Dx Sex)")
-  storeData( "fit_marginal.lst", "dream fit for marginal analysis")
-  storeData( "res_zenith", "zenith results from fit_marginal.lst")
-  storeData( "res_meta_zenith", "Meta-analysis of zenith results from fit_marginal.lst")
-  storeData( "res_meta_zenith.sex_autosome", "Meta-analysis of zenith results from fit_marginal.lst - just autosomal genes")
-  storeData( "df_DE_results", "Differential expressed genes within each cohort")
-  storeData( "df_meta", "Meta-analysis of differential expression across cohorts")
-
-  # Sex by Disease interaction
-  storeData( "fit_inter_sex.disease.lst", "dream fit for interaction analysis")
-  storeData( "res_zenith.inter_sex.disease", "zenith results for interaction analysis")
-  storeData( "res_meta_zenith.inter_sex.disease", "Meta-analysis of zenith results for interaction analysis")
-  storeData( "res_meta_zenith.inter_sex.disease_autosome", "Meta-analysis of zenith results for interaction analysis -  autosomal genes only")
-  storeData( "df_DE_results.inter_sex.disease", "Differential expression result for each cohort for interaction analysis")
-  storeData( "df_meta.inter_sex.disease", "Meta-analysis for differential expression for interaction analysis")
-}
-
-#   rmarkdown::render("combined_analysis/combine_all_DLPFC.Rmd")
-
-```
-
-### Source
-[markdown](`r thisFile`)

### 2020-11-18 combined_analysis/peer_in_each.Rmd
Initial commit

--- 2020-09-20 combined_analysis/combine_all_DLPFC.Rmd
@@
 ---
-title: "CommonMind Consortium: DLPFC"
-subtitle: "Analysis of read count data by cohort"
-author:
-- name: "[Gabriel Hoffman](http://gabrielhoffman.github.io)$^{1}$ and Kelsey Montgomery$^{2}$"
-  affiliation: | 
-    [1] Icahn School of Medicine at Mount Sinai, New York
-    [2] Sage Bionetworks, Seattle, WA
+title: "Run PEER in each dataset"
+subtitle: ''
+author: "Developed by [Gabriel Hoffman](http://gabrielhoffman.github.io/)"
 date: "Run on `r Sys.time()`"
@@
   toc: true
-    toc_float: true
-params:
-  upload: TRUE
+  smart: false
 ---
@@
 cd /hpc/users/hoffmg01/work/covarr-de
-# rm -rf combined_analysis/combine_all_DLPFC_cache
+# rm -rf molec_profiling_cache/
 ml git pandoc
@@
 
-rmarkdown::render("combined_analysis/combine_all_DLPFC.Rmd")
+rmarkdown::render("combined_analysis/peer_in_each.Rmd")
 
@@
 
-```{r knit2synapse, eval= FALSE}
+--->
+
+```{r knit2synapse, eval=FALSE}
 library(synapser)
@@
 synLogin()
-knit2synapse::createAndKnitToFolderEntity(file = "combined_analysis/combine_all_DLPFC.Rmd",
-                                          parentId ="syn22416298",
-                                          folderName = 'Joint analysis of DLPFC count data',
+knit2synapse::createAndKnitToFolderEntity(file = "combined_analysis/peer_in_each.Rmd",
+                                          parentId ="syn22231629",
+                                          folderName = 'Run PEER in each dataset',
                                           overwrite=TRUE, 
@@
 
-```{r synapse.parameters, include=FALSE, cache=TRUE}
-
-library(githubr)
-parentId = 'syn22416298';
-activityName = 'Joint analysis of DLPFC count data';
-activityDescription = 'Joint analysis of DLPFC count data';
-
-thisFileName <- 'combine_all_DLPFC.Rmd'
-
-# Github link
-thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
-thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('combined_analysis/',thisFileName))
-```
-
-<!---
-
-TODO
-
-zenith 1 sided
-
-zenith plots
-replication or correlation as a function of expression magnitude
-
-difference networks
--------------------
-
-DONE
-
-Effect of ilr should depend on Instaition
-suffix 1.
-zenith for Sex with autosomes
-check zenith squared test statistics
-
---->
+GSEA and dream from raw counts
 
@@
 library(cowplot)
-library(ggraph)
-library(tidygraph)
 library(kableExtra)
@@
 library(ggrepel)
-library(corrplot)
-library(zenith)
-library(compositions)
 library(knitr)
@@
            dplyr::filter(SampleID %in% colnames(geneCountsMerged), !is.na(SampleID)) %>%
-  		 		 dplyr::filter(Region %in% c('DLPFC')) %>%
+           # dplyr::filter(Region %in% c('DLPFC')) %>%
          dplyr::mutate(Cohort = forcats::fct_recode(Cohort, `MSSM-Penn-Pitt` = 'MSSM-Penn-Pitt', `NIMH-HBCC`='NIMH-HBCC', `NIMH-HBCC`='NIMH-HBCC-4')) %>%
-			 	 dplyr::filter(Dx %in% c('Control', "SCZ"))  %>%
+         # dplyr::filter(Dx %in% c('Control', "SCZ"))  %>%
          dplyr::filter(ageOfDeath > 17)
@@
 ```
-
-```{r joint.cca, fig.width=7, fig.height=7}
-variables = c('Reported_Gender', 'CellType', 'ageOfDeath', 'Institution', "Dx")
-
-df_cca = lapply(variables, function(v1){
-
-  if( v1 == "CellType"){
-    form1 = ~ GABA + GLU + Olig + MgAs
-  }else{
-    form1 = as.formula(paste0('~', v1))
-  }
-  M1 = model.matrix(form1, METADATA)
-
-  res = lapply(variables, function(v2){
-
-    if( v2 == "CellType"){
-      form2 = ~ GABA + GLU + Olig + MgAs
-    }else{
-      form2 = as.formula(paste0('~', v2))
-    }
-    M2 = model.matrix(form2, METADATA)
-
-    fit = cancor(M1, M2)
-    V_sqrt = sqrt(mean(fit$cor^2)) # Cramer's sqrt V
-
-    data.frame(v1, v2, V_sqrt)
-  })
-  do.call(rbind, res)
-})
-df_cca = do.call(rbind, df_cca)
-
-C = matrix(df_cca$V_sqrt, ncol=length(variables))
-rownames(C) = variables
-colnames(C) = variables
-
-plotCorrMatrix(C, main='CCA')
-```
-
-```{r explain.CellType}
-CTs = c('GABA', 'GLU', 'Olig', 'MgAs')
-
-form = ~ ageOfDeath:Institution + Reported_Gender + Dx + Institution + Dx
-
-Y = clr(METADATA[,CTs])
-
-vp = fitExtractVarPartModel( t(Y), form, METADATA, BPPARAM=SerialParam())
-plotPercentBars(vp) + ggtitle("Explain cell type composition")
-
-ymax = max(rowSums(vp[,-ncol(vp)]))*110
-plotPercentBars(vp) + scale_y_continuous(expand=c(0,0), limits=c(0, ymax))
-```
-
-```{r sex.graph, fig.width=7, fig.height=7}
-library(ggplot2)
-library(ggraph)
-library(tidygraph)
-
-df_cca2 = df_cca[with(df_cca, v1!=v2),]
-
-rstat_nodes <- data.frame(name = unique(df_cca2$v1))
-
-rstat_edges <- data.frame(from = match(df_cca2$v1, rstat_nodes$name),
-                         to = match(df_cca2$v2, rstat_nodes$name),
-                         Rsq = df_cca2$V_sqrt^2 )
-
-net = tbl_graph(nodes = rstat_nodes, edges = rstat_edges, directed=FALSE)
-
-ggraph(net,  layout = 'igraph', algorithm = 'star') + 
-  geom_edge_link(aes(width = Rsq, color=Rsq), alpha = 1) +
-  geom_node_point(size=20, color="grey70") +
-  geom_node_text(aes(label = name), repel = FALSE) +
-  theme_graph() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) +
-  scale_edge_width(name=bquote(R^2)) +
-  scale_edge_color_gradient(name=bquote(R^2), low="white", high="red", lim=c(0, max(df_cca2$V_sqrt^2))) +
-  ggtitle("Pairwise squared correlation")
-```
-
-```{r sex.confounding, fig.width=5, fig.height=5}
-form = ~ Reported_Gender + GABA + GLU + Olig + MgAs + ageOfDeath 
-par(mfrow=c(2,2))
-res = lapply( unique(METADATA$Institution), function(INST){
-  i = METADATA$Institution == INST
-  C = canCorPairs(form, METADATA[i,]) 
-  plotCorrMatrix(C, main=INST)
-})
-```
-
-```{r sex.by.age}
-ggplot(METADATA, aes(Reported_Gender, ageOfDeath, fill=Reported_Gender)) + geom_boxplot() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + facet_wrap(~Institution, nrow=1) + scale_fill_manual(values = c("red", "blue"))
-```
-
-### Sex by cell fraction
-```{r summary.sex, fig.width=8, fig.height=10}
-figList = lapply(colnames(df_cellFractions), function(CT){
-  ggplot(METADATA, aes_string('Reported_Gender', CT, fill='Reported_Gender')) + geom_boxplot() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + facet_wrap(~Institution, nrow=1) + scale_fill_manual(values = c("red", "blue"))
-})
-plot_grid(plotlist=figList, ncol=1) 
-```
-
-### Age by cell fraction
-```{r summary.Age, fig.width=8, fig.height=10}
-figList = lapply(colnames(df_cellFractions), function(CT){
-  ggplot(METADATA, aes_string('ageOfDeath', CT)) + geom_point() + geom_smooth(method="gam") + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + facet_wrap(~Institution, nrow=1)
-}) 
-plot_grid(plotlist=figList, ncol=1) 
-```
-
-## Age by cell fraction log ratio
-```{r summary.Age.log.ratio, fig.width=8, fig.height=8}
-figList = lapply(colnames(celFrac_ilr), function(CT){
-  ggplot(METADATA, aes_string('ageOfDeath', CT)) + geom_point() + geom_smooth(method="gam") + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + facet_wrap(~Institution, nrow=1)
-})
-plot_grid(plotlist=figList, ncol=1) 
-```
-
-```{r process.counts}
-# keep genes that are express in at least one combination of cohort and gender
-isexpr.lst = dlply(METADATA, .(Cohort), .fun = function(mtd){
-	rowSums(cpm(geneCountsMerged[,mtd$SampleID])>.5) >= .3*ncol(geneCountsMerged[,mtd$SampleID])
-	})
-
-isexpr = do.call(rbind, isexpr.lst)
-isexpr = colSums(isexpr) == 2 # keep gene expressed in both cohorts 
-table(isexpr)
-
-# Normalize each cohort separately
-dge.lst = lapply( levels(METADATA$Cohort), function(Cohort){
-  # normalize counts
-  dge = DGEList( geneCountsMerged[isexpr,METADATA$Cohort==Cohort] )
-  calcNormFactors(dge, "none") # TMM
-})
-names(dge.lst) = levels(METADATA$Cohort)
-
-# form = ~ (1|Dx) + Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)
-
-form.lst = list(
-  `MSSM-Penn-Pitt`  = ~ Dx + Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
-  `NIMH-HBCC`       = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)
-
-vobj.lst = lapply(dge.lst, function(dge){
-
-  i = match(colnames(dge), rownames(METADATA))
-  info = METADATA[i,]
-
-  form_use = form.lst[[info$Cohort[1]]]
-
-  voomWithDreamWeights(dge, form_use, info, save.plot=TRUE, BPPARAM = SnowParam(18))
-  })
-names(vobj.lst) = names(dge.lst)
-```
-
-### Voom trend plots
-```{r plot.voom, fig.height=5, fig.width=10}
-fig1 = plot_voom(vobj.lst[[1]]) + ggtitle(names(vobj.lst)[1])
-fig2 = plot_voom(vobj.lst[[2]]) + ggtitle(names(vobj.lst)[2])
-plot_grid( fig1, fig2, nrow=1)
-
-df = lapply( names(vobj.lst), function(key){
-  data.frame(Cohort = key,  vobj.lst[[key]]$voom.line)
-  })
-df = do.call(rbind, df)
-
-ggplot(df, aes(x, y, color=Cohort)) + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + geom_line(size=1.2) + xlab(bquote(log[2](count + 0.5))) + ylab(expression( sqrt("standard deviation"))) + ggtitle("Compare voom trends") + ylim(0, max(df$y)) + scale_color_manual(name="Status", values=c("#3758ba", "#33e65c")) + guides(color = guide_legend(override.aes = list(size = 1)))
-```
-
-<!---
-
-# Examine genes on chrX and chrY
-
-# dge = DGEList(geneCountsMerged)
-# dge = calcNormFactors(dge, "none")
-# geneExpr = cpm(dge, log=TRUE)
-
-# df = merge( data.frame(Expression = geneExpr['ENSG00000229807.11',]), METADATA, by="row.names")
-# ggplot(df, aes(Reported_Gender, Expression)) + geom_violin() + theme_bw() + facet_wrap(~Cohort)
-
-# id = 'ENSG00000229807.11'
-# isexpr[id]
-
-# df = as.data.frame(vobj.lst[[2]]$voom.xy)
-# with(df, plot(x,y, cex=.2, pch=20))
-# with(df[id,], points(x,y, col="red", pch=20, cex=2))
-
---->
-
-```{r Create.residuals}
-
-### Create residuals
-form.lst = list(
-  `MSSM-Penn-Pitt`  = ~ Dx*Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
-  `NIMH-HBCC`       = ~ Dx*Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)
-
-eval_residuals = function( form ){
-  # Compute residuals, but add back in Dx*Reported_Gender
-  # for form = ~ Dx*Reported_Gender
-  f = function(fit){
-    residuals(fit) + get_prediction( fit, form)
-  }
-
-  residuals.lst = lapply( vobj.lst, function(vobj){
-
-    i = match(colnames(vobj), rownames(METADATA))
-    info = METADATA[i,]
-
-    form_use = form.lst[[info$Cohort[1]]]
-
-    # Compute residuals which adding back Diagnosis and Sex effects
-    resid.lst = fitVarPartModel(vobj, form_use, info, fxn = f, BPPARAM = SnowParam(18))  
-    do.call(rbind, resid.lst)
-    })
-  names(residuals.lst) = names(vobj.lst)
-
-  residuals.lst
-}
-
-resid.add = list( Dx_Sex  = eval_residuals( ~ Dx*Reported_Gender ),
-                  Dx      = eval_residuals( ~ Dx ),
-                  Sex     = eval_residuals( ~ Reported_Gender ))
-
-residuals.lst = resid.add[['Dx_Sex']]
-```
-
-```{r check.correlation, eval=FALSE}
-ngenes = 300
-C1 = cor(t(residuals.lst[[1]][1:ngenes,]))
-C2 = cor(t(residuals.lst[[2]][1:ngenes,]))
-
-df = data.table(A   = as.numeric(C1[lower.tri(C1)]),
-                B   = as.numeric(C2[lower.tri(C2)]))
-
-df$density <- get_density(df$A, df$B, n = 80)
-ggplot(df, aes(A, B, color=density)) + geom_point() + geom_point(size=.4) + theme_bw(15) + theme(aspect.ratio=1, legend.position="bottom", plot.title = element_text(hjust = 0.5)) + geom_abline(color="red") + scale_color_viridis() + xlim(-1,1) + ylim(-1,1) + xlab(paste0('Correlation (', names(residuals.lst)[1], ')')) + ylab(paste0('Correlation (', names(residuals.lst)[2], ')'))
-```
-
-# Schizophrenia vs controls
-
-### Test marginals for Scz vs controls
-```{r differential.expression.simple}
-form.lst = list(
-  `MSSM-Penn-Pitt`  = ~ Dx + Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
-  `NIMH-HBCC`       = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)
-
-fit_marginal.lst = lapply( vobj.lst, function(vobj){
-
-  i = match(colnames(vobj), rownames(METADATA))
-  info = METADATA[i,]
-
-  form_use = form.lst[[info$Cohort[1]]]
-
-  fit = dream(vobj, form_use, info, BPPARAM = SnowParam(18))  
-  eBayes( fit )
-  })
-names(fit_marginal.lst) = names(vobj.lst)
-
-coefTest = c('DxSCZ', 'Reported_GenderMale', 'ageOfDeath')
-```
-
-```{r load.genesets, eval=TRUE}
-# Load Gene Ontology database 
-gs.GO = get_GeneOntology()
-gs.MSigDB = get_MSigDB()
-
-# convert from GeneSetCollection to list used by camera and zenith
-gs.GO.collection = recodeToList( gs.GO )
-gs.MSigDB.collection = recodeToList( gs.MSigDB )
-```
-
-```{r run.zenith, eval=TRUE}
-res_zenith = lapply( names(fit_marginal.lst), function(key){
-
-  fit = fit_marginal.lst[[key]]
-
-  res.GO = run_zenith(  fit, 
-                        coefTest,
-                        gs.GO.collection)
-  res.GO$Resource = "GO"   
-  res.GO$Cohort = key
-
-  res.MSigDB = run_zenith(fit, 
-                          coefTest,
-                          gs.MSigDB.collection ) 
-  res.MSigDB$Resource = "MSigDB"
-  res.MSigDB$Cohort = key
-
-  rbind(res.GO, res.MSigDB)
-})
-names(res_zenith) = names(fit_marginal.lst)
-
-# only autosomes
-res_zenith_sex_autosomes = lapply( names(fit_marginal.lst), function(key){
-
-  fit = fit_marginal.lst[[key]]
-
-  # only keep autosomes, for Reported_GenderMale
-  tab = coef(fit) %>% getGeneSymbol()
-  geneids = rownames(tab)[tab$Chrom %in% 1:22]
-  fit = fit[geneids,]
-
-  res.GO = run_zenith(  fit, 
-                        'Reported_GenderMale',
-                        gs.GO.collection)
-  res.GO$Resource = "GO"   
-  res.GO$Cohort = key
-
-  res.MSigDB = run_zenith(fit, 
-                          'Reported_GenderMale',
-                          gs.MSigDB.collection ) 
-  res.MSigDB$Resource = "MSigDB"
-  res.MSigDB$Cohort = key
-
-  rbind(res.GO, res.MSigDB)
-})
-names(res_zenith_sex_autosomes) = names(fit_marginal.lst)
-```
-
-## Meta-analysis of zenith results
-```{r zenith.meta, eval=TRUE}
-res_meta_zenith = zenith_meta_analysis( res_zenith )                  
-res_meta_zenith.sex_autosome = zenith_meta_analysis( res_zenith_sex_autosomes )
-```
-
-```{r get.signatures, fig.width=10, fig.height=5}
-# Get DE signatures
-df_DE_results = lapply(fit_marginal.lst, function(fit){
-  res = lapply(coefTest, function(coef){
-    res = topTable(fit, coef, number=Inf) 
-    res$Coef = coef
-    res %>% getGeneSymbol()  
-    })
-  names(res) = coefTest
-  res
-  })
-names(df_DE_results) = names(fit_marginal.lst)
-```
-
-```{r test.scz, fig.height=5, fig.width=10}
-# Schizophrenia testing
-# extract from each cohort
-coef = 'DxSCZ'
-# ymin = min(sapply(df_DE_results, function(x) min(x[[coef]]$P.Value)))
-
-figList.Scz = lapply( names(df_DE_results), function(key){
-  df = df_DE_results[[key]][[coef]]
-
-  main = paste(key, coef, sep = ': ' )
-  plotVolcano( df, df$Symbol[1:10] ) + ggtitle( main ) #+ ylim(0, -log10(ymin))
-})
-
-# volcano plots
-plot_grid(plotlist=figList.Scz, nrow=1)
-```
-
-## Meta-analysis across cohorts
-```{r meta.analysis.DE}
-# Store results for each coef and cohort
-tab.de.lst = lapply(coefTest, function(coef){    
-  tab.lst = lapply( df_DE_results, function(lst){
-    lst[[coef]]
-  })
-  names(tab.lst) = names(df_DE_results)
-  tab.lst
-  })
-names(tab.de.lst) = coefTest
-
-# Create list preparing for meta-analysis for SCZ
-df_meta = lapply(coefTest, function(coef){
-  # Run meta-analysis
-  run_meta_analysis( tab.de.lst[[coef]], "FE") %>% getGeneSymbol
-})#, mc.cores=6)
-names(df_meta) = coefTest
-```
-
-```{r concordance.scz, fig.height=6, fig.width=10}
-# Concordance analysis
-resConcord.SCZ = plot_concordance(tab.de.lst[[coef]], showGenes = df_meta[[coef]]$Symbol[1:20])
-
-main = "Concordance of SCZ"
-plot_grid(  resConcord.SCZ$fig_logFC + ggtitle(main),
-            resConcord.SCZ$fig_tstat + ggtitle(main), nrow=1)
-
-data.frame(Cohort = names(tab.de.lst[[coef]]),
-          "# DE"  = resConcord.SCZ$n.de,
-          pi1     = resConcord.SCZ$pi_single,
-          check.names=FALSE) %>% 
-          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)
-
-resConcord.SCZ$tab_pi %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
-resConcord.SCZ$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
-```
-
-```{r plot.meta.analysis.SCZ, fig.height=5, fig.width=10}
-# volcano plot for meta-analysis
-main = paste('Meta-analysis', coef, sep = ': ' )
-plotVolcano( df_meta[[coef]], df_meta[[coef]]$Symbol[1:20] ) + ggtitle( main )
-
-data.frame(pi1 = 1 - qvalue(df_meta[[coef]]$P)$pi0,
-          `# DE` = sum(df_meta[[coef]]$adj.P.Val < 0.05),
-          check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for SCZ meta-analysis") %>% kable_styling(full_width=FALSE)
-
-```
-
-## Plot top examples
-```{r plot.forrest.dlpfc, fig.width=8, fig.height=20}  
-plot_gene = function(residuals.lst, df_meta, geneSymbol, variable, fill=c("#1E90FF", "#FF8D1E"), mode=""){
-
-  ensGene = rownames(df_meta)[which(df_meta$Symbol == geneSymbol)]
-
-  exprResid = lapply( names(residuals.lst), function(key){
-      data.frame(Expression = residuals.lst[[key]][ensGene,])
-    })
-  exprResid = do.call(rbind, exprResid)
-
-  df = merge(exprResid, METADATA, by="row.names")
-
-  if( is.factor(df[[variable]]) || length(unique(df[[variable]])) < 5 ){
-    fig = ggplot(df, aes_string(variable, 'Expression')) + geom_violin(aes_string(fill=variable)) + geom_boxplot(width=0.1, fill="grey80") + scale_fill_manual(values = fill)
-  }else{    
-    fig = ggplot(df, aes_string(variable, 'Expression')) + geom_point() + geom_smooth(method="lm") #+ scale_color_manual(values = fill)
-  }
-
-  fig + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + ylab('Expression residuals') + ggtitle(geneSymbol) + facet_wrap(~Cohort)
-}
-
-coef = 'DxSCZ'
-figList = lapply(df_meta[[coef]]$Symbol[1:10], function(geneSymbol){
-  fig1 = plot_forrest( c(tab.de.lst[[coef]], list("Meta-analysis" = df_meta[[coef]])), geneSymbol)
-
-  fig2 = plot_gene(residuals.lst, df_meta[[coef]], geneSymbol, "Dx") + xlab("Diagnosis")
-
-  plot_grid(fig1, fig2, nrow=1)
-})
-
-plot_grid(plotlist=figList, ncol=1)
-```
-
-### Concordance with other datasets
-```{r CMC.concordance}
-# CMC Analysis 1
-CMC1 <- readr::read_tsv(synGet("syn5607603")$path) %>% 
-  dplyr::rename(gene_id = genes,
-                `CMC 1 t-statistic` = t) %>% 
-  dplyr::select(gene_id, `CMC 1 t-statistic`)
-# HBCC-DLPFC microarray
-CMC_HBCC <- readr::read_tsv(synGet("syn4941553")$path) %>% 
-  dplyr::rename(gene_id = genes,
-                `HBCC microarray t-statistic` = t) %>% 
-  dplyr::select(gene_id, `HBCC microarray t-statistic`)
-# Gandal, Science 2018, Table S1
-# Use Tab:DGE
-# wget https://science.sciencemag.org/highwire/filestream/720110/field_highwire_adjunct_files/1/aat8127_Table_S1.xlsx
-gandal <- readxl::read_xlsx( synGet("syn22255217")$path, sheet = "DGE") %>% 
-  dplyr::rename(gene_id = ensembl_gene_id,
-                `Gandal, SCZ t-statistic` = `SCZ.t.value`,
-                `Gandal, BD t-statistic` = `BD.t.value`,
-                `Gandal, ASD t-statistic` = `ASD.t.value`) %>%
-  dplyr::select(gene_id, `Gandal, SCZ t-statistic`,
-                `Gandal, BD t-statistic`, `Gandal, ASD t-statistic`)
-# AMP-AD RNASeq reprocessing GRCh37 - https://www.synapse.org/#!Synapse:syn9702085
-MSSM <- readr::read_tsv(synGet("syn10526259")$path) %>%
-  dplyr::filter(Comparison %in% "AD-CONTROL",
-                Model %in% "Diagnosis") %>%
-  dplyr::select(ensembl_gene_id, Tissue, t) %>%
-  dplyr::group_by(ensembl_gene_id, Tissue) %>% 
-  dplyr::filter(row_number() == 1) %>% 
-  tidyr::pivot_wider(names_from = Tissue,
-                     values_from = t) %>%
-  dplyr::rename(`AMP-AD MSSM FP AD t-statistic` = FP,
-                `AMP-AD MSSM IGF AD t-statistic` = IFG,
-                `AMP-AD MSSM PHG AD t-statistic` = PHG,
-                `AMP-AD MSSM STG AD t-statistic` = STG,
-                gene_id = ensembl_gene_id) %>% 
-  dplyr::select(gene_id, `AMP-AD MSSM STG AD t-statistic`) 
-MSSM <- MSSM[!is.na(MSSM$`AMP-AD MSSM STG AD t-statistic`),]
-
-MAYO <- readr::read_tsv(synGet("syn8468023")$path) %>% 
-  dplyr::filter(Comparison %in% "AD-CONTROL",
-                Model %in% "Diagnosis") %>%
-  dplyr::select(ensembl_gene_id, `Tissue.ref`, t) %>%
-  dplyr::group_by(ensembl_gene_id, `Tissue.ref`) %>% 
-  dplyr::filter(row_number() == 1) %>% 
-  tidyr::pivot_wider(names_from = `Tissue.ref`,
-                     values_from = t) %>% 
-  dplyr::rename(`AMP-AD MAYO CBE AD t-statistic` = CBE,
-                `AMP-AD MAYO TCX AD t-statistic` = TCX,
-                gene_id = ensembl_gene_id) %>% 
-  dplyr::select(-`AMP-AD MAYO TCX AD t-statistic`)
-
-ROSMAP <- readr::read_tsv(synGet("syn8456721")$path) %>%
-    dplyr::filter(Comparison %in% "AD-CONTROL",
-                Model %in% "Diagnosis") %>%
-  dplyr::rename(gene_id = ensembl_gene_id, 
-                `AMP-AD ROSMAP DLPFC AD t-statistic` = t) %>% 
-  dplyr::select(gene_id, `AMP-AD ROSMAP DLPFC AD t-statistic`)
-
-out_group <- inner_join(MSSM, MAYO) %>% 
-  inner_join(., ROSMAP) %>% 
-  inner_join(., gandal)
-
-# merge with meta-analysis
-
-coef = 'DxSCZ'
-df = df_meta[[coef]]
-df_two = data.frame(gene_id = trim_ensembl_ids(rownames(df)),
-                `CMC - meta-analysis` = with(df, logFC/se), check.names=FALSE)
-
-df = tab.de.lst[[coef]][[1]]
-df.tmp = data.frame(gene_id = trim_ensembl_ids(rownames(df)),
-                `t` = df$t, check.names=FALSE)
-colnames(df.tmp)[2] = names(tab.de.lst[[coef]])[1]
-df_two = df_two %>% inner_join(., df.tmp)
-
-df = tab.de.lst[[coef]][[2]]
-df.tmp = data.frame(gene_id = trim_ensembl_ids(rownames(df)),
-                `t` = df$t, check.names=FALSE)
-colnames(df.tmp)[2] = names(tab.de.lst[[coef]])[2]
-df_two = df_two %>% inner_join(., df.tmp)
-
-# Remove transcript Ids
-df_two <- df_two %>% 
-  tidyr::separate(gene_id, c("gene_id", "transcript")) %>% 
-  dplyr::select(-one_of("transcript"))
-df_two <- inner_join(df_two, CMC1, by = c("gene_id")) 
-df_two <- inner_join(df_two, CMC_HBCC, by = c("gene_id"))
-df_two <- inner_join(df_two, out_group, by = c("gene_id")) %>% 
-  group_by(gene_id) %>% 
-  dplyr::filter(row_number()==1) %>% 
-  tibble::column_to_rownames(var = "gene_id")
-
-colnames(df_two) <- gsub(" t-statistic", "", colnames(df_two))
-
-col2 <- colorRampPalette(rev(c("#67001F", "#B2182B", "#D6604D", "#F4A582",
-                           "#FDDBC7", "#FFFFFF", "#D1E5F0", "#92C5DE",
-                           "#4393C3", "#2166AC", "#053061")))
-
-cor <- cor(df_two)
-# clustering based on distance
-hcl = hclust(as.dist(1-cor), method='ward.D2')
-
-results <- pvclust(cor, method.hclust='ward.D2')
-# reorder based on clustering
-reorder = cor[results$hclust$order, results$hclust$order]
-
-# plot clustering
-plot(results, hang=-1)
-# plot(hcl, hang=-1)
-corrplot(reorder, col = col2(200))
-```
-
-```{r test.multtest, eval=FALSE}
-
-library(irlba)
-library(pinnacle)
-library(poolR)
-library(matrixStats)
-
-i = with(res_meta_zenith, (Resource == 'GO') & (Coef == "DxSCZ"))
-df = res_meta_zenith[i,]
-df$qvalue = qvalue(df$PValue)$qvalue
-# df$p.adjust = p.adjust(df$PValue, method="bon" )
-df$p.adjust = pmin(1, 2805 * df$PValue)
-
-M = recodeToSparseMatrix(gs.GO)
-M = M[rownames(M) %in% trim_ensembl_ids(rownames(df_meta_SCZ)),]
-keep = (colSums(M) >= 10) & (colnames(M) %in% rownames(df) )
-M = M[,keep]
-
-m = colMeans(M)
-s = apply(M, 2, var) * nrow(M) / (nrow(M) + 1)
-
-dcmp = irlba( M, nv = 500, center=m, scale=sqrt(s), right_only=FALSE)
-ev = dcmp$d^2 / nrow(M)
-
-plot(cumsum(ev) / ncol(M))
-
-df_unsort = df[colnames(M),]
-edcmp = list(vectors = dcmp$v, values = dcmp$d^2 / nrow(M))
-
-res = p.corr.adjust( df_unsort$PValue, dcmp = edcmp, method.sim="eigen", B = nrow(df_unsort)*4)
-
-plot( cumsum(dcmp$d^2 / sum(dcmp$d^2)))
-
-C = cor(as.matrix(M[,1:1000]))
-evals = eigen(C)$values
-
-dcmp = svd(scale(M) / sqrt(nrow(M)))
-dcmp$d^2
-
-plot(cumsum(dcmp$d^2 / sum(dcmp$d^2)))
-
-meff(C, method='gao')
-
-dcmp = svd(scale(M[,1:100]))
-
-meff(eigen = dcmp$d^2 , method = "galwey")
-
-```
-
-```{r show.gene_set_test.marginal, eval=TRUE}
-show_zenith_results = function( res, coef, Resource, FDR = 0.05){
-
-  i = which((res$Resource == Resource) & (res$Coef == coef))
-  df = res[i,]
-
-	# filter by FDR
-	df = df[df$FDR <= FDR,]
-  df$Coef = c()
-  df$se = c()
-  df$Resource = c()
-
-	df %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption=paste(Resource, coef), row.names=FALSE) %>% kable_styling(full_width=FALSE)
-}
-
-show_zenith_results( res_meta_zenith, "DxSCZ", "GO", FDR=0.05)
-show_zenith_results( res_meta_zenith, "DxSCZ", "MSigDB", FDR=0.05)
-```
-
-# Sex Analysis
-```{r test.Sex, fig.height=5, fig.width=10}
-# Schizophrenia testing
-# extract from each chort
-coef = 'Reported_GenderMale'
-figList.Sex = lapply( names(df_DE_results), function(key){
-  df = df_DE_results[[key]][[coef]]
-  main = paste(key, coef, sep = ': ' )
-  plotVolcano( df, df$Symbol[1:20] ) + ggtitle( main )
-})
-
-# volcano plots
-plot_grid(plotlist=figList.Sex, nrow=1)
-```
-
-### Autosomes
-```{r test.Sex.autosome, fig.height=5, fig.width=10}
-# Schizophrenia testing
-# extract from each chort
-coef = 'Reported_GenderMale'
-figList.Sex = lapply( names(df_DE_results), function(key){
-  df = df_DE_results[[key]][[coef]]
-  main = paste(key, coef, sep = ': ' )
-  df = df[df$Chrom %in% 1:22,]
-  plotVolcano( df, df$Symbol[1:20] ) + ggtitle( main )
-})
-
-# volcano plots
-plot_grid(plotlist=figList.Sex, nrow=1)
-```
-
-## Meta-analysis across cohorts
-```{r concordance.Sex, fig.height=6, fig.width=10}
-# Concordance analysis
-resConcord.Sex = plot_concordance(tab.de.lst[[coef]], showGenes = df_meta[[coef]]$Symbol[1:20])
-
-main = "Concordance of Sex"
-plot_grid(  resConcord.Sex$fig_logFC + ggtitle(main),
-            resConcord.Sex$fig_tstat + ggtitle(main), nrow=1)
-
-data.frame(Cohort = names(tab.de.lst[[coef]]),
-          "# DE"  = resConcord.Sex$n.de,
-          pi1     = resConcord.Sex$pi_single,
-          check.names=FALSE) %>% 
-          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)
-
-resConcord.Sex$tab_pi  %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
-resConcord.Sex$tab_corr  %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
-```
-
-```{r concordance.Sex.autosome, fig.height=6, fig.width=10}
-
-tab.de.lst.autosome = lapply(tab.de.lst[[coef]], function(x) x[x$Chrom %in% 1:22,])
-
-# Concordance analysis
-resConcord.Sex = plot_concordance(tab.de.lst.autosome, showGenes = tab.de.lst.autosome$Symbol[1:20])
-
-main = "Concordance of Sex - autosomes"
-plot_grid(  resConcord.Sex$fig_logFC + ggtitle(main),
-            resConcord.Sex$fig_tstat + ggtitle(main), nrow=1)
-
-data.frame(Cohort = names(resConcord.Sex$pi_single),
-          "# DE"  = resConcord.Sex$n.de,
-          pi1     = resConcord.Sex$pi_single,
-          check.names=FALSE) %>% 
-          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)
-
-resConcord.Sex$tab_pi %>% mutate_if(is.numeric, format, digits=3)  %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
-resConcord.Sex$tab_corr  %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
-```
-
-```{r plot.meta.analysis.Sex, fig.height=5, fig.width=10}
-# volcano plot for meta-analysis
-main = 'Meta-analysis: Sex'
-fig1 = plotVolcano( df_meta[[coef]], df_meta[[coef]]$Symbol[1:20] ) + ggtitle( main )
-
-df = df_meta[[coef]][df_meta[[coef]]$Chrom %in% 1:22,]
-fig2 = plotVolcano( df, df$Symbol[1:20] ) + ggtitle( paste(main, "autosome") )
-plot_grid(fig1, fig2, ncol=2)
-
-data.frame(pi1 = 1 - qvalue(df_meta[[coef]]$P)$pi0,
-          `# DE` = sum(df_meta[[coef]]$adj.P.Val < 0.05),
-          check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for Sex meta-analysis") %>% kable_styling(full_width=FALSE)
-```
-
-## Sex circos
-```{r sex.circos}
-coef = 'Reported_GenderMale'
-sapply( names(df_DE_results), function(key){
-  plot_circos_logFC( df_DE_results[[key]][[coef]] ) + ggtitle(key)
-  })
-
-plot_circos_logFC( df_meta[[coef]] ) + ggtitle("Meta analysis")
-```
-
-## Plot top examples
-```{r plot.forrest.dlpfc.Sex, fig.width=8, fig.height=20}  
-figList = lapply(df_meta[[coef]]$Symbol[1:10], function(geneSymbol){
-
-  fig1 = plot_forrest( c(tab.de.lst[[coef]], list("Meta-analysis" = df_meta[[coef]])), geneSymbol)
-
-  fig2 = plot_gene( residuals.lst, df_meta[[coef]], geneSymbol, "Reported_Gender", fill=c("red", "blue")) + xlab("Sex")
-
-  plot_grid(fig1, fig2, nrow=1)
-})
-
-plot_grid(plotlist=figList, ncol=1)
-```
-
-## Meta-analysis of zenith results
-```{r show.gene_set_test.marginal.Sex, eval=TRUE}
-show_zenith_results( res_meta_zenith, "Reported_GenderMale", "GO", FDR=0.05)         
-show_zenith_results( res_meta_zenith, "Reported_GenderMale", "MSigDB", FDR=0.05)        
-```
-
-### Only autosomes
-```{r show.gene_set_test.marginal.Sex.autosome, eval=TRUE}
-show_zenith_results( res_meta_zenith.sex_autosome, "Reported_GenderMale", "GO", FDR=0.1) 
-show_zenith_results( res_meta_zenith.sex_autosome, "Reported_GenderMale", "MSigDB", FDR=0.1)   
-```
-
-# Age Analysis
-```{r test.Age, fig.height=5, fig.width=10}
-# Schizophrenia testing
-# extract from each chort
-coef = 'ageOfDeath'
-figList.Age = lapply( names(df_DE_results), function(key){
-  df = df_DE_results[[key]][[coef]]
-  main = paste(key, coef, sep = ': ' )
-  plotVolcano( df, df$Symbol[1:20] ) + ggtitle( main )
-})
-
-# volcano plots
-plot_grid(plotlist=figList.Age, nrow=1)
-```
-
-## Meta-analysis across cohorts
-```{r concordance.Age, fig.height=6, fig.width=10}
-# Concordance analysis
-resConcord.Age = plot_concordance(tab.de.lst[[coef]], showGenes = df_meta[[coef]]$Symbol[1:20])
-
-main = "Concordance of Age"
-plot_grid(  resConcord.Age$fig_logFC + ggtitle(main),
-            resConcord.Age$fig_tstat + ggtitle(main), nrow=1)
-
-data.frame(Cohort = names(tab.de.lst[[coef]]),
-          "# DE"  = resConcord.Age$n.de,
-          pi1     = resConcord.Age$pi_single,
-          check.names=FALSE) %>% 
-          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)
-
-resConcord.Age$tab_pi %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
-resConcord.Age$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
-```
-
-```{r plot.meta.analysis.Age, fig.height=5, fig.width=10}
-# volcano plot for meta-analysis
-main = 'Meta-analysis: Age'
-plotVolcano( df_meta[[coef]], df_meta[[coef]]$Symbol[1:20] ) + ggtitle( main )
-
-data.frame(pi1 = 1 - qvalue(df_meta[[coef]]$P)$pi0,
-          `# DE` = sum(df_meta[[coef]]$adj.P.Val < 0.05),
-          check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for Age meta-analysis") %>% kable_styling(full_width=FALSE)
-```
-
-## Plot top examples
-```{r plot.forrest.dlpfc.Age, fig.width=8, fig.height=20}  
-figList = lapply(df_meta[[coef]]$Symbol[1:10], function(geneSymbol){
-
-  fig1 = plot_forrest( c(tab.de.lst[[coef]], list("Meta-analysis" = df_meta[[coef]])), 
-    geneSymbol)
-
-  expr.list = lapply(vobj.lst, function(x)x$E)
-
-  fig2 = plot_gene( expr.list, df_meta[[coef]], geneSymbol, "ageOfDeath", fill=c("red", "blue")) + xlab("Age")
-
-  plot_grid(fig1, fig2, nrow=1)
-})
-
-plot_grid(plotlist=figList, ncol=1)
-```
-
-## Meta-analysis of zenith results
-```{r show.gene_set_test.marginal.Age, eval=TRUE}
-show_zenith_results( res_meta_zenith, "ageOfDeath", "GO", FDR=0.01)   
-show_zenith_results( res_meta_zenith, "ageOfDeath", "MSigDB", FDR=0.01)    
-```
-
-# Sex by Disease interaction
-```{r dream.sex.disease}
-form.lst = list(
-  `MSSM-Penn-Pitt`  = ~ Dx*Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
-  `NIMH-HBCC`       = ~ Dx*Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)
-
-# drop chrY genes
-df_genes  = data.frame(ENSEMBL = rownames(vobj.lst[[1]])) %>% getGeneSymbol('ENSEMBL')
-keep.ensID = df_genes$ENSEMBL[df_genes$Chrom %in% c(1:22, 'X', 'MT')]
-
-fit_inter_sex.disease.lst = lapply( vobj.lst, function(vobj){
-
-  i = match(colnames(vobj), rownames(METADATA))
-  info = METADATA[i,]
-
-  form_use = form.lst[[info$Cohort[1]]]
-
-  fit = dream(vobj[keep.ensID,], form_use, info, BPPARAM = SnowParam(18))  
-  eBayes( fit )
-  })
-names(fit_inter_sex.disease.lst) = names(vobj.lst)
-
-coefTest = c('DxSCZ:Reported_GenderMale')
-```
-
-```{r run.zenith.inter_sex.disease, eval=TRUE}
-res_zenith.inter_sex.disease = lapply( names(fit_inter_sex.disease.lst), function(key){
-
-  fit = fit_inter_sex.disease.lst[[key]]
-
-  res.GO = run_zenith(  fit,   
-                        coefTest,
-                        gs.GO.collection)
-  res.GO$Resource = "GO"   
-  res.GO$Cohort = key
-
-  res.MSigDB = run_zenith(fit, 
-                          coefTest,
-                          gs.MSigDB.collection) 
-  res.MSigDB$Resource = "MSigDB"
-  res.MSigDB$Cohort = key
-
-  rbind(res.GO, res.MSigDB)
-})
-names(res_zenith.inter_sex.disease) = names(fit_inter_sex.disease.lst)
-
-# only autosomes
-res_zenith.inter_sex.disease_autosome = lapply( names(fit_inter_sex.disease.lst), function(key){
-
-  fit = fit_inter_sex.disease.lst[[key]]
-
-  # only keep autosomes, for Reported_GenderMale
-  tab = coef(fit) %>% getGeneSymbol()
-  geneids = rownames(tab)[tab$Chrom %in% 1:22]
-  fit = fit[geneids,]
-
-  res.GO = run_zenith(  fit, 
-                        coefTest,
-                        gs.GO.collection)
-  res.GO$Resource = "GO"   
-  res.GO$Cohort = key
-
-  res.MSigDB = run_zenith(fit, 
-                          coefTest,
-                          gs.MSigDB.collection) 
-  res.MSigDB$Resource = "MSigDB"
-  res.MSigDB$Cohort = key
-
-  rbind(res.GO, res.MSigDB)
-})
-names(res_zenith.inter_sex.disease_autosome) = names(fit_inter_sex.disease.lst)
-```
-
-## Meta-analysis of zenith results
-```{r zenith.meta.inter_sex.disease, eval=TRUE}
-res_meta_zenith.inter_sex.disease = zenith_meta_analysis( res_zenith.inter_sex.disease )  
-res_meta_zenith.inter_sex.disease_autosome = zenith_meta_analysis( res_zenith.inter_sex.disease_autosome )  
-```
-
-```{r get.signatures.inter_sex.disease, fig.width=10, fig.height=5}
-# Get DE signatures
-df_DE_results.inter_sex.disease = lapply(fit_inter_sex.disease.lst, function(fit){
-  res = lapply(coefTest, function(coef){
-    res = topTable(fit, coef, number=Inf) 
-    res$Coef = coef
-    res %>% getGeneSymbol()  
-    })
-  names(res) = coefTest 
-  res
-  })
-names(df_DE_results.inter_sex.disease) = names(fit_inter_sex.disease.lst)
-```
-
-```{r test.scz.inter_sex.disease, fig.height=5, fig.width=10}
-coef = 'DxSCZ:Reported_GenderMale'
-
-figList.Scz = lapply( names(df_DE_results), function(key){
-  df = df_DE_results.inter_sex.disease[[key]][[coef]]
-
-  main = paste(key, coef, sep = ': ' )
-  plotVolcano( df, df$Symbol[1:10] ) + ggtitle( main )
-})
-
-# volcano plots
-plot_grid(plotlist=figList.Scz, nrow=1)
-```
-
-## Meta-analysis across cohorts
-```{r meta.analysis.DE.inter_sex.disease}
-# Store results for each coef and cohort
-tab.de.inter_sex.disease.lst = lapply(coefTest, function(coef){    
-  tab.de.inter_sex.disease.lst = lapply( df_DE_results.inter_sex.disease, function(lst){
-    lst[[coef]]
-  })
-  names(tab.de.inter_sex.disease.lst) = names(df_DE_results.inter_sex.disease)
-  tab.de.inter_sex.disease.lst
-  })
-names(tab.de.inter_sex.disease.lst) = coefTest
-
-# Create list preparing for meta-analysis for SCZ
-df_meta.inter_sex.disease = lapply(coefTest, function(coef){
-  # Run meta-analysis
-  run_meta_analysis( tab.de.inter_sex.disease.lst[[coef]], "FE") %>% getGeneSymbol
-}) 
-names(df_meta.inter_sex.disease) = coefTest  
-```
-
-```{r concordance.scz.inter_sex.disease, fig.height=6, fig.width=10}
-# Concordance analysis.  
-resConcord.inter_sex.disease = plot_concordance(tab.de.inter_sex.disease.lst[[coef]], showGenes = df_meta.inter_sex.disease[[coef]]$Symbol[1:20])
-
-main = "Concordance of DxSCZ:Reported_GenderMale"
-plot_grid(  resConcord.inter_sex.disease$fig_logFC + ggtitle(main),
-            resConcord.inter_sex.disease$fig_tstat + ggtitle(main), nrow=1)
-
-data.frame(Cohort = names(tab.de.inter_sex.disease.lst[[coef]]),
-          "# DE"  = resConcord.inter_sex.disease$n.de,
-          pi1     = resConcord.inter_sex.disease$pi_single,
-          check.names=FALSE) %>% 
-          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)
-
-resConcord.inter_sex.disease$tab_pi %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
-resConcord.inter_sex.disease$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
-```
-
-```{r plot.meta.analysis.SCZ.inter_sex.disease, fig.height=5, fig.width=10}
-# volcano plot for meta-analysis.  
-main = paste('Meta-analysis', coef, sep = ': ' )
-plotVolcano( df_meta.inter_sex.disease[[coef]], df_meta.inter_sex.disease[[coef]]$Symbol[1:20] ) + ggtitle( main )
-
-data.frame(pi1 = 1 - qvalue(df_meta.inter_sex.disease[[coef]]$P)$pi0,
-          `# DE` = sum(df_meta.inter_sex.disease[[coef]]$adj.P.Val < 0.05),
-          `# DE @ 10%` = sum(df_meta.inter_sex.disease[[coef]]$adj.P.Val < 0.1),
-          check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for SCZ:Sex meta-analysis") %>% kable_styling(full_width=FALSE)
-```
-
-## Plot top examples
-```{r plot.forrest.dlpfc.inter_sex.disease, fig.width=8, fig.height=40}
-
-figList = lapply(df_meta.inter_sex.disease[[coef]]$Symbol[1:20], function(geneSymbol){
-  fig1 = plot_forrest( c(tab.de.inter_sex.disease.lst[[coef]], list("Meta-analysis" = df_meta.inter_sex.disease[[coef]])), geneSymbol, plot2='p')
-
-  fig2 = plot_gene(residuals.lst, df_meta.inter_sex.disease[[coef]], geneSymbol, "Dx") + xlab("Diagnosis") + facet_wrap(~Reported_Gender + Cohort, nrow=1)
-
-  plot_grid(fig1, fig2, nrow=1)
-})
-
-plot_grid(plotlist=figList, ncol=1)
-```
-
-## Meta-analysis of zenith results
-```{r show.gene_set_test.marginal.inter_sex.disease, eval=TRUE}
-coef = 'DxSCZ:Reported_GenderMale'   
-show_zenith_results( res_meta_zenith.inter_sex.disease, coef, "GO", FDR=0.1)  
-show_zenith_results( res_meta_zenith.inter_sex.disease, coef, "MSigDB", FDR=0.1)
-
-show_zenith_results( res_meta_zenith.inter_sex.disease_autosome, coef, "GO", FDR=0.1) 
-show_zenith_results( res_meta_zenith.inter_sex.disease_autosome, coef, "MSigDB", FDR=0.1)    
-```
-
-# Dx test with in Males and then Females
-
-```{r store, eval=TRUE}
-if( params$upload ){    
-  # Code
-  # Make sure these match the folder name and parent Id in the first chunk.
-  CODE <- Folder(name = "Joint analysis of DLPFC count data", parentId = "syn22416298")
-  CODE <- synStore(CODE)
-
-  # Store data 
-  storeData = function( name, description){
-    file = paste0(name, ".RDS")
-    saveRDS( eval(parse(text=name)), file=file)
-
-    nEXP_OBJ = File(file, 
-                    name = description, 
-                    parentId = CODE$properties$id)
-    synStore(nEXP_OBJ, used = ALL_USED_IDs, executed = thisFile)
-  }
-
-  # Standard analysis
-  storeData( "geneCountsMerged", "Merged DLPFC counts for both cohorts")
-  storeData( "METADATA", "Merged metadata used here for both cohorts")
-  storeData( "isexpr", "Genes used in this analysis")
-  storeData( "vobj.lst", "Voom results for each cohort")
-  storeData( "resid.add", "Residuals correcting for covariates and add back - (Dx_Sex Dx Sex)")
-  storeData( "fit_marginal.lst", "dream fit for marginal analysis")
-  storeData( "res_zenith", "zenith results from fit_marginal.lst")
-  storeData( "res_meta_zenith", "Meta-analysis of zenith results from fit_marginal.lst")
-  storeData( "res_meta_zenith.sex_autosome", "Meta-analysis of zenith results from fit_marginal.lst - just autosomal genes")
-  storeData( "df_DE_results", "Differential expressed genes within each cohort")
-  storeData( "df_meta", "Meta-analysis of differential expression across cohorts")
-
-  # Sex by Disease interaction
-  storeData( "fit_inter_sex.disease.lst", "dream fit for interaction analysis")
-  storeData( "res_zenith.inter_sex.disease", "zenith results for interaction analysis")
-  storeData( "res_meta_zenith.inter_sex.disease", "Meta-analysis of zenith results for interaction analysis")
-  storeData( "res_meta_zenith.inter_sex.disease_autosome", "Meta-analysis of zenith results for interaction analysis -  autosomal genes only")
-  storeData( "df_DE_results.inter_sex.disease", "Differential expression result for each cohort for interaction analysis")
-  storeData( "df_meta.inter_sex.disease", "Meta-analysis for differential expression for interaction analysis")
-}
-
-#   rmarkdown::render("combined_analysis/combine_all_DLPFC.Rmd")
-
-```
-
-### Source
-[markdown](`r thisFile`)

### 2020-11-18 combined_analysis/peer_in_each.Rmd
Initial commit

@@
 knit2synapse::createAndKnitToFolderEntity(file = "combined_analysis/peer_in_each.Rmd",
-                                          parentId ="syn22231629",
+                                          parentId ="syn23565426",
                                           folderName = 'Run PEER in each dataset',

### 2020-11-18 combined_analysis/peer_in_each.Rmd
Initial commit

@@
 knit2synapse::createAndKnitToFolderEntity(file = "combined_analysis/peer_in_each.Rmd",
-                                          parentId ="syn23565426",
+                                          parentId ="syn22012942",
                                           folderName = 'Run PEER in each dataset',
@@
 library(plyr)
+library(compositions)
 library(BiocParallel)
@@
   cache = TRUE)
+```
+
+```{r synapse.parameters, include=FALSE, cache=TRUE}
+
+library(githubr)
+parentId = 'syn22012942';
+activityName = 'Run PEER in each dataset';
+activityDescription = 'Run PEER in each dataset';
+
+thisFileName <- 'peer_in_each.Rmd'
+
+# Github link
+thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
+thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('combined_analysis/',thisFileName))
 ```
@@
 ```{r summary}
-METADATA %>% group_by(Dx, Institution) %>% dplyr::summarise(count = n()) %>% 
-  kable(caption="Diagnosis, Institution") %>% kable_styling(full_width=FALSE)
+METADATA %>% group_by(Dx, Institution, Region) %>% dplyr::summarise(count = n()) %>% 
+  kable(caption="Diagnosis, Institution, Region") %>% kable_styling(full_width=FALSE)
 
-METADATA %>% group_by( Institution, Reported_Gender) %>% dplyr::summarise(count = n()) %>% 
-  kable(caption="Institution, Sex") %>% kable_styling(full_width=FALSE)
+METADATA %>% group_by( Institution, Reported_Gender, Region) %>% dplyr::summarise(count = n()) %>% 
+  kable(caption="Institution, Sex, Region") %>% kable_styling(full_width=FALSE)
 
-METADATA %>% group_by(Dx, Institution, Reported_Gender) %>% dplyr::summarise(count = n()) %>% 
-kable(caption="Diagnosis, Institution, Sex") %>% kable_styling(full_width=FALSE) 
+METADATA %>% group_by(Dx, Institution, Reported_Gender, Region) %>% dplyr::summarise(count = n()) %>% 
+kable(caption="Diagnosis, Institution, Sex, Region") %>% kable_styling(full_width=FALSE) 
 ```
+
+```{r}
+CODE <- Folder(name = "Run PEER in each dataset", parentId = "syn22012942")
+CODE <- synStore(CODE)
+
+# Store data 
+storeData = function( name, file, description){
+  saveRDS( eval(parse(text=name)), file=file)
+
+  nEXP_OBJ = File(file, 
+                  name = description, 
+                  parentId = CODE$properties$id)
+  synStore(nEXP_OBJ, used = ALL_USED_IDs, executed = thisFile)
+}
+```
+
+# Normalize and process
+Each region and chort separately
+```{r process.counts}
+# keep genes that are express in at least one combination of cohort and gender
+isexpr.lst = dlply(METADATA, .(Cohort, Region), .fun = function(mtd){
+  rowSums(cpm(geneCountsMerged[,mtd$SampleID])>.5) >= .3*ncol(geneCountsMerged[,mtd$SampleID])
+  })
+
+isexpr = do.call(rbind, isexpr.lst)
+isexpr = colSums(isexpr) > 0 # keep gene expressed in both cohorts 
+table(isexpr)
+
+grid = expand.grid(levels(METADATA$Cohort), unique(METADATA$Region))
+
+# Normalize each cohort separately
+dge.lst = lapply( 1:nrow(grid), function(i){
+
+  Cohort = grid$Var1[i]
+  Region = grid$Var2[i]
+
+  idx = (METADATA$Cohort==Cohort)&(METADATA$Region == Region)
+
+  cat(i, sum(idx), '\n')
+  if( sum(idx) == 0){
+    res = NULL
+  }else{
+    # normalize counts
+    dge = DGEList( geneCountsMerged[isexpr,idx] )
+    res = calcNormFactors(dge, "none")
+  }
+  res
+})
+names(dge.lst) = apply(grid, 1, paste, collapse=' - ')
+dge.lst = dge.lst[!is.null(sapply(dge.lst, nrow))]
+
+# form = ~ (1|Dx) + Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)
+
+form.lst = list(
+  `MSSM-Penn-Pitt`  = ~ Dx + Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
+  `NIMH-HBCC`       = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)
+
+form = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3
+
+residMat.lst = lapply( names(dge.lst), function(datID){
+
+  # get log2 CPM gene expression
+  dge = dge.lst[[datID]]
+
+  # get metadata in the same order
+  i = match(colnames(dge), rownames(METADATA))
+  info = METADATA[i,]
+
+  # get formula
+  j = unlist(sapply(names(form.lst), function(x) grep(x, datID)))
+  form_use = form.lst[[names(j)]]
+
+  # apply voom
+  vobj = voomWithDreamWeights(dge[1:100,], form_use, info, save.plot=TRUE, BPPARAM = SnowParam(18))
+
+  # Compute residuals, but add back in Dx*Reported_Gender
+  # for form = ~ Dx*Reported_Gender
+  f = function(fit){
+    residuals(fit) + get_prediction( fit, ~ Dx + Reported_Gender + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)
+  }
+
+  # Compute residuals while adding back biological variables
+  resid.lst = fitVarPartModel(vobj, form_use, info, fxn = f, BPPARAM = SnowParam(18))  
+  residMat = do.call(rbind, resid.lst)
+
+   # upload residuals
+  file = paste0('metadata_',gsub(" ", '', gsub("-", "_", datID)), ".RDS")
+  storeData( residMat, file, paste0(datID, " Residuals removing technical variables"))
+
+  # upload covariates
+  info.sub = info[,all.vars(form_use)]
+  file = paste0('metadata_',gsub(" ", '', gsub("-", "_", datID)), ".RDS")
+  storeData( info.sub, file, paste0(datID, " Meta data for variables relavant for QTL analysis"))
+
+  })
+names(residMat.lst) = names(dge.lst)
+```

### 2020-11-18 combined_analysis/peer_in_each.Rmd
Initial commit

@@
     dge = DGEList( geneCountsMerged[isexpr,idx] )
-    res = calcNormFactors(dge, "none")
+    res = calcNormFactors(dge, "TMM")
   }
@@
 names(dge.lst) = apply(grid, 1, paste, collapse=' - ')
-dge.lst = dge.lst[!is.null(sapply(dge.lst, nrow))]
-
-# form = ~ (1|Dx) + Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)
+dge.lst = dge.lst[!sapply(dge.lst, is.null)]
 
@@
   `NIMH-HBCC`       = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)
-
-form = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3
 
@@
   # apply voom
-  vobj = voomWithDreamWeights(dge[1:100,], form_use, info, save.plot=TRUE, BPPARAM = SnowParam(18))
+  vobj = voomWithDreamWeights(dge, form_use, info, save.plot=TRUE, BPPARAM = SnowParam(18))
 

### 2020-11-18 combined_analysis/eval_resid_in_each.Rmd
Initial commit

--- 2020-11-18 combined_analysis/peer_in_each.Rmd
@@
 ---
-title: "Run PEER in each dataset"
+title: "Evaluate residuals in each cohort and brain region"
 subtitle: ''
@@
 # rm -rf molec_profiling_cache/
-ml git pandoc
+ml git pandoc R/3.6.3 # to get peer
 git pull
@@
 
-rmarkdown::render("combined_analysis/peer_in_each.Rmd")
+rmarkdown::render("combined_analysis/eval_resid_in_each.Rmd")
 
@@
 synLogin()
-knit2synapse::createAndKnitToFolderEntity(file = "combined_analysis/peer_in_each.Rmd",
+knit2synapse::createAndKnitToFolderEntity(file = "combined_analysis/eval_resid_in_each.Rmd",
                                           parentId ="syn22012942",
-                                          folderName = 'Run PEER in each dataset',
+                                          folderName = 'Evaluate residuals in each cohort and brain region',
                                           overwrite=TRUE, 
@@
 parentId = 'syn22012942';
-activityName = 'Run PEER in each dataset';
-activityDescription = 'Run PEER in each dataset';
+activityName = 'Evaluate residuals in each cohort and brain region';
+activityDescription = 'Evaluate residuals in each cohort and brain region';
 
-thisFileName <- 'peer_in_each.Rmd'
+thisFileName <- 'eval_resid_in_each.Rmd'
 
@@
 ```{r}
-CODE <- Folder(name = "Run PEER in each dataset", parentId = "syn22012942")
+CODE <- Folder(name = "Evaluate residuals in each cohort and brain region", parentId = "syn22012942")
 CODE <- synStore(CODE)
@@
 
+# formula used to fit model
 form.lst = list(
@@
   `NIMH-HBCC`       = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)
+
+asdfas
 
@@
 
+  # add these terms back into the model
+  form_add = ifelse(names(j) == "MSSM-Penn-Pitt",
+        ~ Dx + Reported_Gender + Institution*(cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
+        ~ Dx + Reported_Gender + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)
+
   # apply voom
@@
   f = function(fit){
-    residuals(fit) + get_prediction( fit, ~ Dx + Reported_Gender + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)
+    residuals(fit) + get_prediction( fit, form_add)
   }
@@
 ```
+
+```{r run.peer}
+
+  model = PEER()
+
+  PEER_setPhenoMean(model,as.matrix(expr))
+
+  PEER_setCovariates(model, as.matrix(covs))
+
+  PEER_setNk(model,10)
+
+  PEER_update(model)
+
+  latentVariables = PEER_getX(model)
+
+```

### 2020-11-18 combined_analysis/eval_resid_in_each.Rmd
Initial commit

@@
 library(data.table)
+library(peer)
 library(cowplot)
@@
 
-asdfas
-
-residMat.lst = lapply( names(dge.lst), function(datID){
+for( datID in names(dge.lst) ){
 
@@
   form_use = form.lst[[names(j)]]
+  # apply voom
+  vobj = voomWithDreamWeights(dge[1:200,], form_use, info, save.plot=TRUE, BPPARAM = SnowParam(18))
+
+  # construct covariate matrix
+  cov_matrix = model.matrix(form_use, info)
+
+  # fit PEER
+  model = PEER()
+
+  PEER_setPhenoMean(model, t(vobj$E))
+
+  PEER_setCovariates(model, cov_matrix)
+
+  PEER_setNk(model,2)
+
+  PEER_update(model)
+
+  # extract PEER latent variables
+  # first ncol(cov_matrix) columns are the oringal covariates
+  latentVariables = PEER_getX(model)[,-c(1:ncol(cov_matrix))]
+  rownames(latentVariables) = rownames(cov_matrix)
+  colnames(latentVariables) = paste0("PEER_", 1:ncol(latentVariables))
+
+  # upload observed expression
+  file = paste0('geneExpr_',gsub(" ", '', gsub("-", "_", datID)), ".RDS")
+  storeData( vobj$E, file, paste0(datID, " Observed expression as log2 CPM"))
+
+  # upload covariate matrix
+  file = paste0('covariateMatrix_',gsub(" ", '', gsub("-", "_", datID)), ".RDS")
+  storeData( cov_matrix, file, paste0(datID, " Covariates for QTL analysis"))
+
+  # upload latent variables
+  file = paste0('metadata_',gsub(" ", '', gsub("-", "_", datID)), ".RDS")
+  storeData( latentVariables, file, paste0(datID, " Latent variables for QTL analysis"))
+}
+```
+
+For each cohort and each brain region, compute TMM normalization, log2 CPM with voomWithDreamWeights, and a set of covariates.
+Then compute PEER components using the observed expression and the covariates
+
+    `MSSM-Penn-Pitt`  = ~ Dx + Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)
+
+  `NIMH-HBCC`       = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3
+
+<!---
 
@@
   form_add = ifelse(names(j) == "MSSM-Penn-Pitt",
-        ~ Dx + Reported_Gender + Institution*(cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
-        ~ Dx + Reported_Gender + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)
-
-  # apply voom
-  vobj = voomWithDreamWeights(dge, form_use, info, save.plot=TRUE, BPPARAM = SnowParam(18))
+        '~ Dx + Reported_Gender + Institution*(cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)',
+       ' ~ Dx + Reported_Gender + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3')
+  form_add = as.formula(form_add)
 
@@
 names(residMat.lst) = names(dge.lst)
-```
-
-```{r run.peer}
-
-  model = PEER()
-
-  PEER_setPhenoMean(model,as.matrix(expr))
-
-  PEER_setCovariates(model, as.matrix(covs))
-
-  PEER_setNk(model,10)
-
-  PEER_update(model)
-
-  latentVariables = PEER_getX(model)
-
-```
+--->

### 2020-11-18 combined_analysis/eval_resid_in_each.Rmd
Initial commit

@@
 ---
-title: "Evaluate residuals in each cohort and brain region"
+title: "Evaluate PEER in each cohort and brain region"
 subtitle: ''
@@
 parentId = 'syn22012942';
-activityName = 'Evaluate residuals in each cohort and brain region';
-activityDescription = 'Evaluate residuals in each cohort and brain region';
+activityName = 'Evaluate PEER in each cohort and brain region';
+activityDescription = 'Evaluate PEER in each cohort and brain region';
 
@@
 ```{r}
-CODE <- Folder(name = "Evaluate residuals in each cohort and brain region", parentId = "syn22012942")
+CODE <- Folder(name = "Evaluate PEER in each cohort and brain region", parentId = "syn22012942")
 CODE <- synStore(CODE)
@@
   # apply voom
-  vobj = voomWithDreamWeights(dge[1:200,], form_use, info, save.plot=TRUE, BPPARAM = SnowParam(18))
+  vobj = voomWithDreamWeights(dge, form_use, info, save.plot=TRUE, BPPARAM = SnowParam(18))
 
@@
 
-  PEER_setNk(model,2)
+  PEER_setNk(model,20)
 

### 2020-11-18 combined_analysis/eval_resid_in_each.Rmd
Initial commit

@@
 # Store data 
-storeData = function( name, file, description){
-  saveRDS( eval(parse(text=name)), file=file)
+storeData = function( data, file, description){
+  saveRDS( data, file=file)
 
@@
 
-  PEER_setNk(model,20)
+  PEER_setNk(model,30)
 
@@
   # upload latent variables
-  file = paste0('metadata_',gsub(" ", '', gsub("-", "_", datID)), ".RDS")
+  file = paste0('PEER',gsub(" ", '', gsub("-", "_", datID)), ".RDS")
   storeData( latentVariables, file, paste0(datID, " Latent variables for QTL analysis"))

### 2020-11-19 combined_analysis/eval_resid_in_each.Rmd
Initial commit

@@
 # Set Control to baseline
-METADATA$Dx = factor(METADATA$Dx, c('Control', 'SCZ'))
+METADATA$Dx = factor(METADATA$Dx)#, c('Control', 'SCZ'))
 
@@
   # apply voom
-  vobj = voomWithDreamWeights(dge, form_use, info, save.plot=TRUE, BPPARAM = SnowParam(18))
+  vobj = voomWithDreamWeights(dge[1:100,], form_use, info, save.plot=TRUE, BPPARAM = SnowParam(18))
 
@@
 
-  PEER_setNk(model,30)
+  PEER_setNk(model,2)
 

### 2020-11-19 combined_analysis/eval_resid_in_each.Rmd
Initial commit

@@
   # apply voom
-  vobj = voomWithDreamWeights(dge[1:100,], form_use, info, save.plot=TRUE, BPPARAM = SnowParam(18))
+  vobj = voomWithDreamWeights(dge, form_use, info, save.plot=TRUE, BPPARAM = SerialParam())
 
@@
 
-  PEER_setNk(model,2)
+  PEER_setNk(model,30)
 

### 2020-11-23 combined_analysis/eval_resid_in_each.Rmd
add CTS

@@
 
+# Cell Type Specific expression from bMIND
+
+```{r}
+# create new parent folder for these results
+CODE <- Folder(name = "CTS - Evaluate PEER in each cohort and brain region", parentId = "syn23567522")
+CODE <- synStore(CODE)
+
+# Store data 
+storeData = function( data, file, description){
+  saveRDS( data, file=file)
+
+  nEXP_OBJ = File(file, 
+                  name = description, 
+                  parentId = CODE$properties$id)
+  synStore(nEXP_OBJ, used = ALL_USED_IDs, executed = thisFile)
+}
+```
+
+```{r CTS.download}
+geneExpr.CTS = readRDS( synGet('syn23234833')$path )
+
+# formula used to fit model
+form.lst = list(
+  `MSSM-Penn-Pitt`  = ~ Dx + Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*ageOfDeath,
+  `NIMH-HBCC`       = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath)
+
+# for each dataset
+for( datID in names(dge.lst) ){
+
+  # get log2 CPM gene expression
+  dge = dge.lst[[datID]]
+
+  # get metadata in the same order
+  i = match(colnames(dge), rownames(METADATA))
+  info = METADATA[i,]
+
+  # get formula
+  j = unlist(sapply(names(form.lst), function(x) grep(x, datID)))
+  form_use = form.lst[[names(j)]]
+
+  # extract CTS data corresponding to info
+  idx = match(info$SampleID, dimnames(geneExpr.CTS)[[3]])
+  geneExpr = geneExpr.CTS[,,idx]
+
+  # identical(dimnames(geneExpr)[[3]], info$SampleID)
+
+  # for each cell type
+  for( i in length(dimnames(geneExpr)[[2]]) ){
+
+    # construct covariate matrix
+    cov_matrix = model.matrix(form_use, info)
+
+    # fit PEER
+    model = PEER()
+
+    PEER_setPhenoMean(model, t(geneExpr[1:100,i,]))
+
+    PEER_setCovariates(model, cov_matrix)
+
+    PEER_setNk(model,2)
+
+    PEER_update(model)
+
+    # extract PEER latent variables
+    # first ncol(cov_matrix) columns are the oringal covariates
+    latentVariables = PEER_getX(model)[,-c(1:ncol(cov_matrix))]
+    rownames(latentVariables) = rownames(cov_matrix)
+    colnames(latentVariables) = paste0("PEER_", 1:ncol(latentVariables))
+
+    # upload observed expression
+    file = paste0('geneExpr_',gsub(" ", '', gsub("-", "_", datID)), ".RDS")
+    storeData( vobj$E, file, paste0(datID, " Observed expression"))
+
+    # upload covariate matrix
+    file = paste0('covariateMatrix_',gsub(" ", '', gsub("-", "_", datID)), ".RDS")
+    storeData( cov_matrix, file, paste0(datID, " Covariates for QTL analysis"))
+
+    # upload latent variables
+    file = paste0('PEER',gsub(" ", '', gsub("-", "_", datID)), ".RDS")
+    storeData( latentVariables, file, paste0(datID, " Latent variables for QTL analysis"))
+  }
+}
+```
+
 <!---

### 2020-11-23 combined_analysis/eval_resid_in_each.Rmd
add CTS

@@
 
+    CT = dimnames(geneExpr)[[2]][i]
+
     # construct covariate matrix
@@
     # upload observed expression
-    file = paste0('geneExpr_',gsub(" ", '', gsub("-", "_", datID)), ".RDS")
-    storeData( vobj$E, file, paste0(datID, " Observed expression"))
+    file = paste0('geneExpr_',gsub(" ", '', gsub("-", "_", datID)), '_', CT,".RDS")
+    storeData( vobj$E, file, paste0(datID, " Observed expression", '_', CT,))
 
     # upload covariate matrix
-    file = paste0('covariateMatrix_',gsub(" ", '', gsub("-", "_", datID)), ".RDS")
-    storeData( cov_matrix, file, paste0(datID, " Covariates for QTL analysis"))
+    file = paste0('covariateMatrix_',gsub(" ", '', gsub("-", "_", datID)), '_', CT, ".RDS")
+    storeData( cov_matrix, file, paste0(datID, " Covariates for QTL analysis", '_', CT,))
 
     # upload latent variables
-    file = paste0('PEER',gsub(" ", '', gsub("-", "_", datID)), ".RDS")
-    storeData( latentVariables, file, paste0(datID, " Latent variables for QTL analysis"))
+    file = paste0('PEER',gsub(" ", '', gsub("-", "_", datID)), '_', CT, ".RDS")
+    storeData( latentVariables, file, paste0(datID, " Latent variables for QTL analysis", '_', CT,))
   }

### 2020-11-23 combined_analysis/eval_resid_in_each.Rmd
add CTS

--- 2020-11-23 combined_analysis/eval_resid_in_each.Rmd
@@
     # upload observed expression
-    file = paste0('geneExpr_',gsub(" ", '', gsub("-", "_", datID)), ".RDS")
-    storeData( vobj$E, file, paste0(datID, " Observed expression"))
+    file = paste0('geneExpr_',gsub(" ", '', gsub("-", "_", datID)), '_', CT,".RDS")
+    storeData( geneExpr, file, paste0(datID, " Observed expression", ' - ', CT))
 
     # upload covariate matrix
-    file = paste0('covariateMatrix_',gsub(" ", '', gsub("-", "_", datID)), ".RDS")
-    storeData( cov_matrix, file, paste0(datID, " Covariates for QTL analysis"))
+    file = paste0('covariateMatrix_',gsub(" ", '', gsub("-", "_", datID)), '_', CT, ".RDS")
+    storeData( cov_matrix, file, paste0(datID, " Covariates for QTL analysis", ' - ' , CT))
 
     # upload latent variables
-    file = paste0('PEER',gsub(" ", '', gsub("-", "_", datID)), ".RDS")
-    storeData( latentVariables, file, paste0(datID, " Latent variables for QTL analysis"))
+    file = paste0('PEER',gsub(" ", '', gsub("-", "_", datID)), '_', CT, ".RDS")
+    storeData( latentVariables, file, paste0(datID, " Latent variables for QTL analysis", ' - ', CT))
   }

### 2020-11-23 combined_analysis/eval_resid_in_each.Rmd
add CTS

@@
 
-```{r}
+```{r create.folder.2}
 # create new parent folder for these results

### 2020-11-23 combined_analysis/eval_resid_in_each.Rmd
add CTS

@@
 geneExpr.CTS = readRDS( synGet('syn23234833')$path )
+```
 
+```{r CTS.eval_peer}
 # formula used to fit model

### 2020-11-23 combined_analysis/eval_resid_in_each.Rmd
add CTS

@@
 geneExpr.CTS = readRDS( synGet('syn23234833')$path )
+ALL_USED_IDs = c(ALL_USED_IDs, 'syn23234833')
 ```
@@
   # for each cell type
-  for( i in length(dimnames(geneExpr)[[2]]) ){
+  for( i in 1:length(dimnames(geneExpr)[[2]]) ){
 

### 2021-01-08 combined_analysis/combine_all_DLPFC.Rmd
add new plots

@@
 library(corrplot)
+library(fdrtool)
 library(pvclust)
+library(RColorBrewer)
 library(metafor)
@@
 library(ggrepel)
+library(corpcor)
 library(corrplot)
@@
 metadata = downloadFile_version(METADATA_QC_DLPFC_ID, version = 19) %>%
-  dplyr::select(`Individual_ID`, Institution, Cohort, `Reported_Gender`, Sex, Ethnicity, ageOfDeath, `PMI_(in_hours)`, Dx, 
+  dplyr::select(`Individual_ID`, Institution, Cohort, `Reported_Gender`, Sex, Ethnicity, ageOfDeath, `PMI_(in_hours)`, Dx, 'pH',
                 Sample_RNA_ID, one_of('rnaSeq_isolation:RIN',
@@
   dplyr::select(SampleID, Individual_ID, Institution, Cohort, Reported_Gender,Sex, Ethnicity, ageOfDeath, PMI, Dx, Region,
-                RIN, ReportExclude, GenotypeExclude, EV.1, EV.2, EV.3, EV.4, EV.5,
+                RIN, ReportExclude, GenotypeExclude, EV.1, EV.2, EV.3, EV.4, EV.5, pH,
                 LibraryBatch, FlowcellBatch, IsolationBatch, RibozeroBatch, MappedReads, TotalReads, GenesDetected, AlignmentRate,
@@
 
+```{r summarize.plots}
+
+METADATA %>% group_by(Dx, Institution, Reported_Gender) %>% dplyr::summarise(count = n()) %>% ggplot(aes(Institution, count, fill=Dx)) + geom_bar(stat="identity", position=position_dodge()) + facet_wrap(~ Reported_Gender) + theme_bw(16) + theme(aspect.ratio=1, axis.text.x = element_text(angle = 45, hjust=1)) + scale_fill_manual("Diagnosis", values=c("grey", "darkred"))
+```
+
+```{r summarize.plots2, fig.height=12}
+variables = c( 'ageOfDeath',"RIN", "PMI", 'rRNARate', 'IntronicRate', 'IntragenicRate', 'IntergenicRate', "pH")
+
+METADATA %>% group_by(Dx, Institution, Reported_Gender) %>% reshape2::melt(, id.vars=c('Dx', 'Institution', 'Reported_Gender'), measure.vars=variables ) %>% ggplot(aes(Institution, value, fill=Dx)) + geom_boxplot() + facet_grid( variable ~ Reported_Gender, scales='free_y') + theme_bw(16) + theme(aspect.ratio=1, axis.text.x = element_text(angle = 45, hjust=1), legend.position="none") + scale_fill_manual("Diagnosis", values=c("grey", "darkred"))
+```
+
+Wilcoxon test comparing Controls and SZ within each Institution and Sex
+```{r compare.case.control, fig.height=12}
+df = METADATA %>% group_by(Dx, Institution, Reported_Gender) %>% reshape2::melt(, id.vars=c('Dx', 'Institution', 'Reported_Gender'), measure.vars=variables ) %>% data.table
+
+# test difference bween CTL and SZ within each Institution, Sex
+df[,wilcox.test(value ~ Dx)$p.value, by=c("Institution", "Reported_Gender", "variable")] %>% ggplot(aes(Institution, -log10(V1))) + geom_bar(stat="identity", position=position_dodge()) + facet_grid(variable ~ Reported_Gender) + theme_bw(16) + theme(aspect.ratio=1, axis.text.x = element_text(angle = 45, hjust=1)) + scale_fill_manual("Diagnosis", values=c("grey", "darkred")) + ylab(bquote(-log[10]~P))
+```
+
+ANCOVA evaluating heterogenity in values across cohorts for each sex
+```{r ancova, fig.height=12}
+df2 = df[,anova(lm(value ~ Institution))$`Pr(>F)`[1], by=c("Reported_Gender", "variable")] 
+df2[,p.adj := p.adjust(V1, "bonferroni")]
+
+ggplot(df2, aes(Reported_Gender, -log10(V1), fill=p.adj < 0.05)) + geom_bar(stat="identity", position=position_dodge()) + facet_wrap(~variable, ncol=1) + theme_bw(16) + theme(aspect.ratio=1, axis.text.x = element_text(angle = 45, hjust=1)) + ylab(bquote(-log[10]~P)) + scale_fill_manual("Passes 5%\nBonferroni", values=c("grey", "navy"))
+
+df2[,p := format(V1, digits=2,scientific=TRUE)][,c('Reported_Gender', 'variable', 'p')] %>% kable(caption="ANCOVA p-values", scientific=TRUE) %>% kable_styling(full_width=FALSE)
+```
+
 ```{r joint.cca, fig.width=7, fig.height=7}
+
+# pH is not included because it is not available om many of the samples
+sum(!is.na(METADATA$pH))
+sum(is.na(METADATA$pH))
+
 variables = c('Reported_Gender', 'CellType', 'ageOfDeath', 'Institution', "Dx")
+variables2 = c("RIN", "PMI", 'rRNARate', 'IntronicRate', 'IntragenicRate', 'IntergenicRate')
+variables = c(variables, variables2)
 
@@
 
-    fit = cancor(M1, M2)
+    # If some rows are dropped for including NA's, 
+    # only analyze rows present in both variables
+    rowids = intersect(rownames(M1), rownames(M2))
+
+    fit = cancor(M1[rowids,], M2[rowids,])
     V_sqrt = sqrt(mean(fit$cor^2)) # Cramer's sqrt V
@@
 
+<!---
+Evaluate and perform hypothesis test for partial correlations
+```{r pcor}
+# Compute partial correlations
+pC = cor2pcor(C)
+colnames(pC) = colnames(C)
+rownames(pC) = rownames(C)
+
+res = lapply(colnames(C), function(v1){
+  res = lapply(colnames(C), function(v2){
+
+    r = pC[v1, v2]
+    n = nrow(METADATA)
+    k = ncol(C)
+    t = r * sqrt((n-2-k) / (1-r^2))
+
+    p = 2*pt( abs(t), n - 2 -k, lower.tail=FALSE)
+
+    a = sort(c(v1, v2))
+
+    data.frame(V1 = a[1], V2 = a[2], partial.corr = r, pvalue = p)
+  })
+  do.call(rbind, res)
+ })
+res = data.table(do.call(rbind, res))
+res = res[V1 != V2,]
+res[,title:=paste(V1, V2)]
+res = res[,.SD[1,], by='title']
+
+res$p.adj = p.adjust( res$pvalue, "BH")
+
+ggplot(res, aes(V1, V2, fill=partial.corr, label=ifelse(p.adj<0.05, '*', ''))) + geom_tile() + theme_bw() + scale_fill_gradient2(name = "Partial corr", low="blue", mid="white", high="red", limit=c(-1, 1)) + theme(aspect.ratio=1, panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(), legend.position="bottom") + geom_text() + coord_flip() + xlab("Variable 1") + ylab("Variable 2")
+```
+--->
+
 ```{r explain.CellType}
@@
 
-form = ~ ageOfDeath:Institution + Reported_Gender + Dx + Institution
+form = ~ ageOfDeath:Institution + Reported_Gender + Dx + Institution + RIN + rRNARate + PMI + IntronicRate + IntragenicRate + IntergenicRate
 
@@
 
-  fit = lm(Y[,id] ~ ageOfDeath:Institution + Reported_Gender + Dx + Institution, METADATA)
-
-  data.frame( Test     = rownames(anova(fit))[1:4],
+  form2 = as.formula(paste0('Y[,id] ~ ', as.character(form)[2]))
+
+  fit = lm(form2, METADATA)
+
+  vp = calcVarPart(fit)
+  n = length(vp) - 1 # omit residuals
+
+  data.frame( Test     = rownames(anova(fit))[1:n],
               CellType = id, 
-              p        = anova(fit)[,'Pr(>F)'][1:4], 
-              frac     =  calcVarPart(fit)[1:4])
+              p        = anova(fit)[,'Pr(>F)'][1:n], 
+              frac     =  calcVarPart(fit)[1:n])
 })
@@
 df$p.adj = p.adjust(df$p, "bonferroni")
+
+df2 = df[,sum(frac), by="Test"]
+setorder(df2, 'V1')
+
+df$Test = factor(df$Test, df2$Test)
 
@@
         legend.key = element_blank(), plot.margin = unit(c(0, 
-            0.3, 0, 0.8), "cm")) + scale_y_continuous(expand=c(0,0), limits=c(0, ymax)) + xlab('')
-
-ggplot(df, aes(CellType, Test, fill=-log10(pmax(1e-16,p)), label=ifelse(p.adj<0.05, '*', ''))) + geom_tile() + theme_bw() + scale_fill_gradient(name = bquote(-log[10]~P), low="white", high="red", limit=c(0, NA)) + theme(aspect.ratio=1, panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank()) + geom_text() + coord_flip()
+            0.3, 0, 0.8), "cm")) + scale_y_continuous(expand=c(0,0), limits=c(0, ymax)) + xlab('') + scale_fill_manual(values = rev(brewer.pal( length(table(df$Test)), "Set3")))
+
+ratio = length(CTs) / nlevels(df$Test)
+
+ggplot(df, aes(CellType, Test, fill=-log10(pmax(1e-16,p)), label=ifelse(p.adj<0.05, '*', ''))) + geom_tile() + theme_bw() + scale_fill_gradient(name = bquote(-log[10]~P), low="white", high="red", limit=c(0, NA)) + theme(aspect.ratio=ratio, panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(), legend.position="bottom") + geom_text() + coord_flip()
 ```
@@
 ```{r sex.graph, fig.width=7, fig.height=7, dev="png"}
+plotStarCorr = function(df_cca, cutoff=1e-3){
+
   df_cca2 = df_cca[with(df_cca, v1!=v2),]
@@
   df_cca2 = df_cca2[!duplicated(df_cca2$key),]
+
+  df_cca2 = df_cca2[df_cca2$V_sqrt^2 > cutoff,]
 
@@
     scale_edge_width(name=bquote(R^2)) +
-  scale_edge_color_gradient(name=bquote(R^2), low="white", high="red", lim=c(0, max(df_cca2$V_sqrt^2))) +  ggtitle("Pairwise squared correlation")
+    scale_edge_color_gradient(name=bquote(R^2), low="cornsilk", high="red", lim=c(0, max(df_cca2$V_sqrt^2))) +  ggtitle("Pairwise squared correlation")
+}
+
+plotStarCorr(df_cca, 0)
+```
+
+```{r sex.graph.sub, fig.width=7, fig.height=7, dev="png"}
+i1 = with(df_cca, v1 %in% setdiff(variables, variables2))
+i2 = with(df_cca, v2 %in% setdiff(variables, variables2))
+
+plotStarCorr(df_cca[i1 & i2,], 0)
 ```
@@
 ```{r sex.confounding, fig.width=5, fig.height=5}
-form = ~ Reported_Gender + GABA + GLU + Olig + MgAs + ageOfDeath 
+form = ~ Reported_Gender + GABA + GLU + Olig + MgAs + ageOfDeath + PMI + rRNARate + RIN + IntronicRate + IntragenicRate + IntergenicRate
 par(mfrow=c(2,2))
@@
 })
+```
+
+Despite the fact that the estimated GABA fraction is lower in females, this observation is not significant after accounting for other variables, notably age of death. 
+```{r sex.confounding.testing, fig.width=7, fig.height=4}
+
+res = lapply( unique(METADATA$Institution), function(INST){
+  res = lapply( c('GABA', 'GLU', 'Olig', 'MgAs'), function(CT){
+    i = METADATA$Institution == INST
+    y = factor(METADATA[i,]$Reported_Gender, c("Male", "Female"))
+
+    form = as.formula(paste0('y ~ ', CT, '+ ageOfDeath + PMI + rRNARate + RIN + IntronicRate + IntragenicRate + IntergenicRate'))
+
+    fit = glm(form,  METADATA[i,], family="binomial")
+
+    result = data.frame(coef(summary(fit))[-1,])
+    result$coef = rownames(result)
+    result$Institution = INST
+    result$CellType = CT
+    rownames(result) = NULL
+
+    result
+    })
+  do.call(rbind, res)
+})
+res = do.call(rbind, res)
+colnames(res)[colnames(res) == 'Pr...z..'] = 'pvalue'
+colnames(res)[colnames(res) == 'Std..Error'] = 'se'
+res$p.adj = p.adjust(res$pvalue, "BH")
+
+lim = max(abs(res$z.value))
+
+res2 = res[res$coef %in%  c('GABA', 'GLU', 'Olig', 'MgAs'),]
+res2$p.adj = p.adjust(res2$pvalue, "BH")
+
+ggplot(res2, aes(Institution, Estimate, color=coef)) + geom_point() + theme_bw() + ylab("Regression coefficient") + facet_wrap(~CellType, nrow=1) + theme(aspect.ratio=1) + geom_errorbar(aes(ymin = Estimate - se, ymax = Estimate + se), width=0.1) + geom_hline(yintercept=0, linetype="dashed")
+```
+
+```{r significance.CT.diff}
+colorCodes = c(
+  GABA = '#66A61E',
+  GLU = '#E6AB02',
+  Olig = '#E7298A',
+  MgAs = '#7570B3',
+  Bulk = "grey50")
+
+res2$color = colorCodes[match(res2$coef, names(colorCodes))]
+
+ggplot(res2, aes(Institution, -log10(pvalue))) + geom_point(color=res2$color) + theme_bw() + ylab("Regression coefficient") + facet_wrap(~CellType, nrow=1) + theme(aspect.ratio=1, axis.text.x = element_text(angle = 45, hjust=1)) + geom_hline(yintercept=-log10(0.05/nrow(res2)), linetype="dashed") + ylab(bquote(-log[10]~P)) + ylim(0, 3)
+```
+
+Sow results for hypothesis testing for: `ageOfDeath + PMI + rRNARate + RIN`
+
+```{r sex.confounding.testing2, fig.width=7, fig.height=4}
+
+res = lapply( unique(METADATA$Institution), function(INST){
+  i = METADATA$Institution == INST
+  y = factor(METADATA[i,]$Reported_Gender, c("Male", "Female"))
+
+  form = y ~ ageOfDeath + PMI + rRNARate + RIN + IntronicRate + IntragenicRate + IntergenicRate
+
+  fit = glm(form,  METADATA[i,], family="binomial")
+
+  result = data.frame(coef(summary(fit))[-1,])
+  result$coef = rownames(result)
+  result$Institution = INST
+  rownames(result) = NULL
+
+  result
+})
+res = do.call(rbind, res)
+colnames(res)[colnames(res) == 'Pr...z..'] = 'pvalue'
+res$p.adj = p.adjust(res$pvalue, "BH")
+
+ggplot(res, aes(Institution, z.value, fill=coef)) + geom_bar(stat="identity", position = position_dodge())  + theme_bw() + ylab("Z-score from logistic regression") + theme(aspect.ratio=1)
+```
+
+Evaluating each term in the model: `Reported_Gender ~ Institution*(ageOfDeath + PMI + rRNARate + RIN + IntronicRate + IntragenicRate + IntergenicRate)` using logistic regression
+```{r anova.test}
+# run ANOVA LRT omitting each variable
+y = factor(METADATA$Reported_Gender, c("Male", "Female"))
+
+form = y ~ Institution*(ageOfDeath + PMI + rRNARate + RIN + IntronicRate + IntragenicRate + IntergenicRate)
+fit_full = glm(form,  METADATA, family="binomial")
+
+res = lapply(c('ageOfDeath', 'PMI', 'rRNARate','RIN', 'IntronicRate', 'IntragenicRate', 'IntergenicRate'), function(V){
+
+  form2 = gsub(V, "1", form)
+  form2 = as.formula(paste(form2[2], '~', form2[3]))
+
+  fit_null = glm(form2,  METADATA, family="binomial")
+
+  res = anova(fit_full, fit_null, test="Chisq")
+
+  data.frame(Variable = V, pvalue = res$`Pr(>Chi)`[2])
+})
+res = do.call(rbind, res)
+res$p.adj = p.adjust(res$pvalue, "bonferroni")
+
+res %>% kable(caption="Hypothesis test ", digits=8) %>% kable_styling(full_width=FALSE)
 ```
@@
           check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for SCZ meta-analysis") %>% kable_styling(full_width=FALSE)
-
 ```
@@
 
-out_group <- inner_join(MSSM, MAYO) %>% 
-  inner_join(., ROSMAP) %>% 
-  inner_join(., gandal)
+# out_group <- inner_join(MSSM, MAYO) %>% 
+#   inner_join(., ROSMAP) %>% 
+#   inner_join(., gandal)
+
+out_group <- inner_join(gandal, ROSMAP)
 
@@
 
-cor <- cor(df_two)
+cor <- cor(df_two, method="spearman")
 # clustering based on distance
@@
 corrplot(reorder, col = col2(200))
+```
+
+## t-statistics for CASKIN1
+```{r CASKIN1}
+ df_two['ENSG00000167971',]  %>% kable(caption="t-statistics for CASKIN1", row.names=FALSE) %>% kable_styling(full_width=FALSE)
 ```
@@
 ```{r plot.gene_set_test.marginal, eval=TRUE}
- plot_zenith_results = function( res, coef, Resource, FDR = 0.05){
+
+  plot_zenith_results = function( res, coef, Resource, FDR = 0.05, compare=NULL){
 
@@
   # filter by FDR
+  df$FDR = pmax(1e-300, df$FDR)
   df = df[df$FDR <= FDR,]
@@
   ymax = max(-log10(df$FDR)) * 1.05
-  ggplot(df, aes(Set, -log10(FDR))) + geom_bar(stat="identity", fill="navy") + theme_bw() + theme(aspect.ratio=1,plot.title = element_text(hjust = 0.5)) + coord_flip() + ylab(bquote(-log[10]~FDR)) + xlab('') + ggtitle(coef) + scale_y_continuous(expand=c(0,0), limits=c(0, ymax)) + geom_hline(yintercept=-log10(0.05), color="red", linetype="dashed")
+
+  if( !is.null(compare) ){
+    df$Direction = compare[(sign(df$effect) + 1)/2 + 1]
+    fig = ggplot(df, aes(Set, -log10(FDR), fill = Direction )) + geom_bar(stat="identity") + scale_fill_manual(name = "Upregulated in", values = c('red', 'blue'))
+  }else{
+    fig = ggplot(df, aes(Set, -log10(FDR) )) + geom_bar(stat="identity", fill="navy")
+  }
+
+  fig + theme_bw() + theme(aspect.ratio=1,plot.title = element_text(hjust = 0.5)) + coord_flip() + ylab(bquote(-log[10]~FDR)) + xlab('') + ggtitle(coef) + scale_y_continuous(expand=c(0,0), limits=c(0, ymax)) + geom_hline(yintercept=-log10(0.05), color="black", linetype="dashed")
 }
@@
   main = paste(key, coef, sep = ': ' )
-  plotVolcano( df, df$Symbol[1:20] ) + ggtitle( main )
+  plotVolcanoSex( df, df$Symbol[1:20] ) + ggtitle( main )
 })
@@
 main = 'Meta-analysis: Sex'
-fig1 = plotVolcano( df_meta[[coef]], df_meta[[coef]]$Symbol[1:20] ) + ggtitle( main )
+fig1 = plotVolcanoSex( df_meta[[coef]], df_meta[[coef]]$Symbol[1:15] ) + ggtitle( main )
 
@@
 
-```{r plot.gene_set_test.marginal.Sex, eval=TRUE}
- plot_zenith_results( res_meta_zenith, "Reported_GenderMale", "GO", FDR=0.05)               
- plot_zenith_results( res_meta_zenith, "Reported_GenderMale", "MSigDB", FDR=0.05)        
+Evaluate if GABA pathways are different between sexes
+```{r zenith.gaba}
+df = data.table(res_meta_zenith)
+df = df[Coef == 'Reported_GenderMale',] 
+df2 = df[grep("GABA", Set),]
+
+# correct only for number of tests that have GABA in the name
+df2[,FDR := p.adjust(PValue, "fdr")] 
+df2 %>% kable(caption="Test only gene sets with GABA in the name") %>% kable_styling(full_width=FALSE)
+```
+
+```{r plot.gene_set_test.marginal.Sex, eval=TRUE, fig.width=12}
+  plot_zenith_results( res_meta_zenith, "Reported_GenderMale", "GO", FDR=0.05, compare=c('Female', "Male"))               
+  plot_zenith_results( res_meta_zenith, "Reported_GenderMale", "MSigDB", FDR=0.05, compare=c('Female', "Male"))         
+```
+
+Show genes in neurexin gene set
+```{r neurexin}
+geneNames = gs.GO.collection[[grep('GO0042043', names(gs.GO.collection))]]
+
+df = df_meta[['Reported_GenderMale']]
+
+i = match( geneNames, trim_ensembl_ids(rownames(df)))
+
+df = df[i[!is.na(i)],]
+df$ischrY = df$Chrom == "Y"
+
+figList = lapply( unique(df$ischrY), function(TF){
+
+  df2 = df[df$ischrY == TF,]
+  ggplot(df2, aes(Symbol, logFC, color=adj.P.Val < 0.05)) + geom_point() + geom_errorbar(aes(ymin = logFC - se, ymax = logFC + se, width=.1)) + theme_bw(16) + theme(aspect.ratio=1) + geom_hline(yintercept=0, linetype="dashed") + coord_flip() + scale_color_manual(name="FDR < 5%", values=c("black", "red"))
+  })
+plot_grid(plotlist = figList, ncol=1)
 ```
@@
 ```{r plot.gene_set_test.marginal.Sex.autosome, eval=TRUE}
- plot_zenith_results( res_meta_zenith.sex_autosome, "Reported_GenderMale", "GO", FDR=0.1) 
- plot_zenith_results( res_meta_zenith.sex_autosome, "Reported_GenderMale", "MSigDB", FDR=0.1)   
+  plot_zenith_results( res_meta_zenith.sex_autosome, "Reported_GenderMale", "GO", FDR=0.1, compare=c('Female', "Male"))  
+  plot_zenith_results( res_meta_zenith.sex_autosome, "Reported_GenderMale", "MSigDB", FDR=0.1, compare=c('Female', "Male"))    
 ```
@@
 plotVolcano( df_meta.inter_sex.disease[[coef]], df_meta.inter_sex.disease[[coef]]$Symbol[1:20] ) + ggtitle( main )  
-
-data.frame(pi1 = 1 - qvalue(df_meta.inter_sex.disease[[coef]]$P)$pi0,
-          `# DE` = sum(df_meta.inter_sex.disease[[coef]]$adj.P.Val < 0.05),
-          `# DE @ 10%` = sum(df_meta.inter_sex.disease[[coef]]$adj.P.Val < 0.1),
-          check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for SCZ:Sex meta-analysis") %>% kable_styling(full_width=FALSE)
+```
+
+Here we used multiple methods (BH, qvalue and FDRTool), to compute the local and global FDR and evalute the number of genes passing the 25% cutoff.
+```{r pv.hist.fdrtool}
+pv = df_meta.inter_sex.disease[[coef]]$P
+
+res_fdr = fdrtool( pv, "pvalue", plot=FALSE, verbose=FALSE)
+
+cutoff = .25
+df = data.frame(`BH` = sum(p.adjust(pv, "BH") < cutoff),
+          `qvalue (global)` = sum(qvalue(pv)$qvalue < cutoff),
+          `qvalue (local)` = sum(qvalue(pv)$lfdr < cutoff),
+          `FDRtool (global)` = sum(res_fdr$qval < cutoff),
+          `FDRtool (local)` = sum(res_fdr$lfdr < cutoff),
+          check.names=FALSE )  %>% t 
+colnames(df) = "Count DE genes"
+
+df %>% kable(caption="Number of DE genes by each multiple testing method at cutoff of 25%") %>% kable_styling(full_width=FALSE)
+```
+
+The pi1 value is `r 1 - qvalue(pv)$pi0`.
+
+Histogram of p-values from interaction testing is close to unifoirm
+
+```{r p.hist, fig.height=5, fig.width=5}
+hist( pv, xlab="p-value", main="Histogram of p-values")
 ```
@@
 
+1 - qvalue(df$p)$pi0
+
 head(df[df$FDR < 0.1,])
+
+coef(lm(t.x ~ t.y, df))
+coef(lm(t.y ~ t.x, df))
+
+with(df, hist(t.y - t.x))
+
+with(df, wilcox.test(t.y, t.x, paired=TRUE))
+
+with(df, hist(logFC.y - logFC.x))
 

### 2021-01-08 common_functions/common_functions.R

@@ -9,6 +9,7 @@ library(Rfast)
 library(viridis)
 library(data.table)
 library(ggraph)
+library(ggrepel)
 library(ggplot2)
 library(tidygraph)
 library(irlba)
@@ -71,6 +72,8 @@ plot_concordance = function( resList, showGenes = NULL, idx=c(1,2), size=15){
 
   df = merge( resList[[idx[1]]], resList[[idx[2]]], by="row.names")
 
+  # see link for simpler code to generate scatter plot with density
+  # https://gist.github.com/GabrielHoffman/0369ce8fd875b3ea5b7a364473add323
   lab = "log2 fold change"
   lim = with(df, max(abs(c(logFC.x, logFC.y))))
   df$density <- get_density(df$logFC.x, df$logFC.y, n = 100)
@@ -259,13 +262,41 @@ plotVolcano = function(df, showGenes = NULL, size=15, minp=1e-310 ){
 
   lim = -log10(min(df$P.Value))
   limx = c(-max(abs(df$logFC)), max(abs(df$logFC)))
-  fig = ggplot(df, aes( logFC, -log10(P.Value), color=isSignif)) + geom_point() + theme_bw(size) + theme(aspect.ratio=1, legend.position="none", plot.title = element_text(hjust = 0.5)) + xlab(bquote(log[2]~fold~change)) + ylab(bquote(-log[10]~P)) + scale_color_manual(values=c("grey", "red")) + scale_y_continuous(limits=c(0, lim*1.02), expand=c(0,0)) + xlim(limx)
+  fig = ggplot(df, aes( logFC, -log10(P.Value), color=isSignif)) + geom_point() + theme_bw(size) + theme(aspect.ratio=1, legend.position="none", plot.title = element_text(hjust = 0.5)) + xlab(bquote(log[2]~fold~change)) + ylab(bquote(-log[10]~P)) + scale_color_manual(values=c("grey", "darkred")) + scale_y_continuous(limits=c(0, lim*1.02), expand=c(0,0)) + xlim(limx)
 
   if( !is.null(showGenes) ){
     # top significant genes
     df2 = df[df$Symbol %in% showGenes,]
 
     fig = fig + geom_text_repel(data=df2, aes(logFC, -log10(P.Value), label=Symbol), segment.size=.5,  segment.color="black", color="black", force=1, nudge_x=.005, nudge_y=.5)
+  }
+
+  fig
+}
[…]

### 2021-01-13 createRelease/create_release_3.1.R
Initial commit

@@ -1,0 +1,60 @@
+# Gabriel Hoffman
+# January 13, 2021
+
+# Create a file with only DLPFC counts
+
+# source("createRelease/create_release_3.1.R")
+
+library(synapser)
+library(dplyr)
+
+synLogin()
+
+library(data.table)
+
+# Read raw counts from MPP and HBCC
+counts.MPP = fread(synGet('syn21867938')$path, data.table=FALSE)
+# Get metadata
+downloadFile <- function(id){
+  fread(synGet(id)$path, data.table = F)
+}
+downloadFile_version <- function(id , version){
+  fread(synGet(id, version = version)$path, data.table = F)
+}
+
+# Get RNASeq QCmetadata
+METADATA_QC_DLPFC_ID = 'syn16816488' 
+ALL_USED_IDs = METADATA_QC_DLPFC_ID
+metadata = downloadFile_version(METADATA_QC_DLPFC_ID, version = 19) %>%
+  dplyr::select(`Individual_ID`, Institution, Cohort, `Reported_Gender`, Sex, Ethnicity, ageOfDeath, `PMI_(in_hours)`, Dx, 'pH',
+                Sample_RNA_ID, one_of('rnaSeq_isolation:RIN',
+                                      'rnaSeq_dissection:Brain_Region')) 
+
+# Keep only DLPFC samples
+METADATA = metadata %>%
+  dplyr::left_join(GENOTYPE) %>% dplyr::rename(Region = `rnaSeq_dissection:Brain_Region`, 
+                SampleID = Sample_RNA_ID) %>% dplyr::select(SampleID, Region) %>% dplyr::filter(SampleID %in% colnames(counts.MPP)) %>% 
+  		 		 dplyr::filter(Region %in% c('DLPFC')) 
+
+# get counts from only DLPFC
+idx = which(colnames(counts.MPP) %in% METADATA$SampleID)
+countsFilter = counts.MPP[,c(1:6, idx)]
+
+# write result top file
+file = "MSSM.Penn.Pitt_DLPFC.featureCount.tsv"
+write.table(countsFilter, file = file,  sep="\t", row.names=FALSE, col.names=TRUE, quote=FALSE)
+system(paste("gzip", file))
+
+
+# 
+thisFileName <- 'create_release_3.1.R'
+
+# Github link
+thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
[…]

### 2021-01-13 createRelease/create_release_3.1.R
Initial commit

@@ -22,6 +22,27 @@ downloadFile_version <- function(id , version){
   fread(synGet(id, version = version)$path, data.table = F)
 }
 
+# Get ancestry vector calculated using gemtools
+ANCESTRY_ID = 'syn9922992'
+ALL_USED_IDs = c(ALL_USED_IDs, ANCESTRY_ID)
+ANCESTRY.HBCC = downloadFile(ANCESTRY_ID) %>% 
+  plyr::rename(c(ID = 'SNP_report:Genotyping_Sample_ID'))
+
+ANCESTRY_ID = 'syn2511399'
+ALL_USED_IDs[length(ALL_USED_IDs)+1] = ANCESTRY_ID
+ANCESTRY.MPP = downloadFile(ANCESTRY_ID) %>% 
+  plyr::rename(c('DNA_report..Genotyping.Sample_ID' = 'SNP_report:Genotyping_Sample_ID'))
+
+ANCESTRY = rbind(ANCESTRY.HBCC[,colnames(ANCESTRY.MPP)[-2]], ANCESTRY.MPP[,colnames(ANCESTRY.MPP)[-2]])
+
+# Get genotype ids from synapse. 
+GENOTYPE_ID = 'syn16816490'
+ALL_USED_IDs = c(ALL_USED_IDs, GENOTYPE_ID)
+GENOTYPE = downloadFile(GENOTYPE_ID) %>% 
+  dplyr::select(Individual_ID, `SNP_report:Genotyping_Sample_ID`, `SNP_report:Exclude?`) %>% 
+  dplyr::inner_join(ANCESTRY) %>% 
+  dplyr::filter(is.na(`SNP_report:Exclude?`))
+  
 # Get RNASeq QCmetadata
 METADATA_QC_DLPFC_ID = 'syn16816488' 
 ALL_USED_IDs = METADATA_QC_DLPFC_ID

### 2021-01-13 createRelease/create_release_3.1.R
Initial commit

@@ -22,6 +22,8 @@ downloadFile_version <- function(id , version){
   fread(synGet(id, version = version)$path, data.table = F)
 }
 
+ALL_USED_IDs = c()
+
 # Get ancestry vector calculated using gemtools
 ANCESTRY_ID = 'syn9922992'
 ALL_USED_IDs = c(ALL_USED_IDs, ANCESTRY_ID)
@@ -42,10 +44,10 @@ GENOTYPE = downloadFile(GENOTYPE_ID) %>%
   dplyr::select(Individual_ID, `SNP_report:Genotyping_Sample_ID`, `SNP_report:Exclude?`) %>% 
   dplyr::inner_join(ANCESTRY) %>% 
   dplyr::filter(is.na(`SNP_report:Exclude?`))
-  
+
 # Get RNASeq QCmetadata
 METADATA_QC_DLPFC_ID = 'syn16816488' 
-ALL_USED_IDs = METADATA_QC_DLPFC_ID
+ALL_USED_IDs = c(ALL_USED_IDs, METADATA_QC_DLPFC_ID)
 metadata = downloadFile_version(METADATA_QC_DLPFC_ID, version = 19) %>%
   dplyr::select(`Individual_ID`, Institution, Cohort, `Reported_Gender`, Sex, Ethnicity, ageOfDeath, `PMI_(in_hours)`, Dx, 'pH',
                 Sample_RNA_ID, one_of('rnaSeq_isolation:RIN',
@@ -78,4 +80,6 @@ thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('createRele
 nEXP_OBJ = File(paste0(file, ".gz"), 
                 name = "Count files for DLFPC from MSSM, Penn, Pitt", 
                 parentId = 'syn24172721')
-synStore(nEXP_OBJ, used = c('syn21867938', 'syn16816488'), executed = thisFile)
+synStore(nEXP_OBJ, used = ALL_USED_IDs, executed = thisFile)
+
+

### 2021-01-13 createRelease/create_release_3.1.R
Initial commit

@@ -7,6 +7,7 @@
 
 library(synapser)
 library(dplyr)
+library(githubr)
 
 synLogin()
 
@@ -66,7 +67,7 @@ countsFilter = counts.MPP[,c(1:6, idx)]
 # write result top file
 file = "MSSM.Penn.Pitt_DLPFC.featureCount.tsv"
 write.table(countsFilter, file = file,  sep="\t", row.names=FALSE, col.names=TRUE, quote=FALSE)
-system(paste("gzip", file))
+system(paste("gzip -f", file))
 
 
 # 

### 2021-01-14 combined_analysis/combine_all_DLPFC.Rmd
For public release

@@
 library(githubr)
-parentId = 'syn22416298';
+parentId = 'syn24172641';
 activityName = 'Joint analysis of DLPFC count data';
@@
 ```
-
-<!---
-
-TODO
-
-zenith 1 sided
-
-zenith plots
-replication or correlation as a function of expression magnitude
-
-difference networks
--------------------
-
-DONE
-
-Effect of ilr should depend on Instaition
-suffix 1.
-zenith for Sex with autosomes
-check zenith squared test statistics
-
---->
 
@@
 library(fdrtool)
+library(ashr)
 library(pvclust)
@@
 # Read raw counts from MPP and HBCC
-counts.MPP = fread(synGet('syn21867938')$path, data.table=FALSE)
+# counts.MPP = fread(synGet('syn21867938')$path, data.table=FALSE)
+
+ALL_USED_IDs = c()
+
+# Just DLPFC counts
+ALL_USED_IDs = c(ALL_USED_IDs, 'syn24172729')
+counts.MPP = fread(synGet('syn24172729')$path, data.table=FALSE)
 rownames(counts.MPP) = counts.MPP$Geneid
@@
 # HBCC - DLFPC
+ALL_USED_IDs = c(ALL_USED_IDs, 'syn21886235')
 counts.HBCC = fread(synGet('syn21886235')$path, data.table=FALSE)
@@
 # # HBCC - sgACC
-counts.HBCC4 = fread(synGet('syn21867866')$path, data.table=FALSE)
-rownames(counts.HBCC4) = counts.HBCC4$Geneid
-counts.HBCC4 = counts.HBCC4[,-c(1:6)]
-
-geneCountsMerged = cbind(counts.MPP, counts.HBCC, counts.HBCC4)
+# NOT INCLUDED in release
+# counts.HBCC4 = fread(synGet('syn21867866')$path, data.table=FALSE)
+# rownames(counts.HBCC4) = counts.HBCC4$Geneid
+# counts.HBCC4 = counts.HBCC4[,-c(1:6)]
+
+# geneCountsMerged = cbind(counts.MPP, counts.HBCC, counts.HBCC4)
+geneCountsMerged = cbind(counts.MPP, counts.HBCC)
 ```
@@
 ```{r download.data, cache=TRUE}
-downloadFile <- function(id){
-  fread(synGet(id)$path, data.table = F)
-}
-downloadFile_version <- function(id , version){
+downloadFile <- function(id, version=NULL){
   fread(synGet(id, version = version)$path, data.table = F)
 }
-
+```
+
+```{r get.metadata, eval=FALSE}
 # Get RNASeq QCmetadata
-METADATA_QC_DLPFC_ID = 'syn16816488' 
-ALL_USED_IDs = METADATA_QC_DLPFC_ID
-metadata = downloadFile_version(METADATA_QC_DLPFC_ID, version = 19) %>%
-  dplyr::select(`Individual_ID`, Institution, Cohort, `Reported_Gender`, Sex, Ethnicity, ageOfDeath, `PMI_(in_hours)`, Dx, 'pH',
-                Sample_RNA_ID, one_of('rnaSeq_isolation:RIN',
-                                      'rnaSeq_dissection:Brain_Region', 
-                                      'rnaSeq_isolation:RNA_Isolation_Batch',
-                                      'rnaSeq_report:Ribozero_Batch', 'rnaSeq_report:Library_Batch', 
-                                      'rnaSeq_report:Flowcell_Batch', 'rnaSeq_report:Exclude?',
-                                      'rnaSeq_report:Mapped_Reads',
-                                      'rnaSeq_report:Intragenic_Rate','rnaSeq_report:Intronic_Rate',
-                                      'rnaSeq_report:Intergenic_Rate','rnaSeq_report:Genes_Detected',
-                                      'rnaSeq_report:Expression_Profiling_Efficiency',
-                                      'rnaSeq_report:rRNA_Rate',
-                                      'rnaSeq_report:Total_Reads', 'rnaSeq_report:Percent_Aligned','rnaSeq_report:Transcripts_Detected'))
-
+# METADATA_QC_DLPFC_ID = 'syn16816488'; version = 19  # private version
+# # METADATA_QC_DLPFC_ID = 'syn18358379'; version = NULL  # public version
+# ALL_USED_IDs = METADATA_QC_DLPFC_ID
+# metadata = downloadFile_version(METADATA_QC_DLPFC_ID, version = version) %>%
+#   dplyr::select(`Individual_ID`, Institution, Cohort, `Reported_Gender`, Sex, Ethnicity, ageOfDeath, `PMI_(in_hours)`, Dx, 'pH',
+#                 Sample_RNA_ID, one_of('rnaSeq_isolation:RIN',
+#                                       'rnaSeq_dissection:Brain_Region', 
+#                                       'rnaSeq_isolation:RNA_Isolation_Batch',
+#                                       'rnaSeq_report:Ribozero_Batch', 'rnaSeq_report:Library_Batch', 
+#                                       'rnaSeq_report:Flowcell_Batch', 'rnaSeq_report:Exclude?',
+#                                       'rnaSeq_report:Mapped_Reads',
+#                                       'rnaSeq_report:Intragenic_Rate','rnaSeq_report:Intronic_Rate',
+#                                       'rnaSeq_report:Intergenic_Rate','rnaSeq_report:Genes_Detected',
+#                                       'rnaSeq_report:Expression_Profiling_Efficiency',
+#                                       'rnaSeq_report:rRNA_Rate',
+#                                       'rnaSeq_report:Total_Reads', 'rnaSeq_report:Percent_Aligned','rnaSeq_report:Transcripts_Detected'))
+
+# # Get ancestry vector calculated using gemtools
+# ANCESTRY_ID = 'syn9922992'
+# ALL_USED_IDs = c(ALL_USED_IDs, ANCESTRY_ID)
+# ANCESTRY.HBCC = downloadFile(ANCESTRY_ID) %>% 
+#   plyr::rename(c(ID = 'SNP_report:Genotyping_Sample_ID'))
+
+# ANCESTRY_ID = 'syn2511399'
+# ALL_USED_IDs[length(ALL_USED_IDs)+1] = ANCESTRY_ID
+# ANCESTRY.MPP = downloadFile(ANCESTRY_ID) %>% 
+#   plyr::rename(c('DNA_report..Genotyping.Sample_ID' = 'SNP_report:Genotyping_Sample_ID'))
+
+# ANCESTRY = rbind(ANCESTRY.HBCC[,colnames(ANCESTRY.MPP)[-2]], ANCESTRY.MPP[,colnames(ANCESTRY.MPP)[-2]])
+
+# # Get genotype ids from synapse. 
+# GENOTYPE_ID = 'syn16816490'
+# ALL_USED_IDs = c(ALL_USED_IDs, GENOTYPE_ID)
+# GENOTYPE = downloadFile(GENOTYPE_ID) %>% 
+#   dplyr::select(Individual_ID, `SNP_report:Genotyping_Sample_ID`, `SNP_report:Exclude?`) %>% 
+#   dplyr::inner_join(ANCESTRY) %>% 
+#   dplyr::filter(is.na(`SNP_report:Exclude?`))
+
+# # Merge metadata   
+# METADATA = metadata %>%
+#   dplyr::left_join(GENOTYPE) %>%      
+#   dplyr::rename(Region = `rnaSeq_dissection:Brain_Region`,
+#                 PMI = `PMI_(in_hours)`,
+#                 RIN = `rnaSeq_isolation:RIN`, 
+#                 ReportExclude = `rnaSeq_report:Exclude?`,
+#                 GenotypeExclude = `SNP_report:Exclude?`,
+#                 SampleID = Sample_RNA_ID,
+#                 LibraryBatch = `rnaSeq_report:Library_Batch`,
+#                 FlowcellBatch = `rnaSeq_report:Flowcell_Batch`,
+#                 IsolationBatch = `rnaSeq_isolation:RNA_Isolation_Batch`, #isolation batch to replace ribozero batch
+#                 RibozeroBatch = `rnaSeq_report:Ribozero_Batch`,
+#                 MappedReads = `rnaSeq_report:Mapped_Reads`,
+#                 IntragenicRate = `rnaSeq_report:Intragenic_Rate`, IntronicRate = `rnaSeq_report:Intronic_Rate`, 
+#                 IntergenicRate = `rnaSeq_report:Intergenic_Rate`, GenesDetected = `rnaSeq_report:Genes_Detected`,
+#                 ExpProfEfficiency = `rnaSeq_report:Expression_Profiling_Efficiency`, rRNARate = `rnaSeq_report:rRNA_Rate`,
+#                 TotalReads = `rnaSeq_report:Total_Reads`, AlignmentRate = `rnaSeq_report:Percent_Aligned`, TranscriptsDetected = `rnaSeq_report:Transcripts_Detected`) %>%
+#   dplyr::select(SampleID, Individual_ID, Institution, Cohort, Reported_Gender,Sex, Ethnicity, ageOfDeath, PMI, Dx, Region,
+#                 RIN, ReportExclude, GenotypeExclude, EV.1, EV.2, EV.3, EV.4, EV.5, pH,
+#                 LibraryBatch, FlowcellBatch, IsolationBatch, RibozeroBatch, MappedReads, TotalReads, GenesDetected, AlignmentRate,
+#                 IntragenicRate, IntergenicRate, IntronicRate, ExpProfEfficiency, rRNARate, TranscriptsDetected) 
+```
+
+```{r get.metadata.2}
 # Get ancestry vector calculated using gemtools
-ANCESTRY_ID = 'syn9922992'
+ANCESTRY_ID = 'syn17894713'
 ALL_USED_IDs = c(ALL_USED_IDs, ANCESTRY_ID)
 ANCESTRY.HBCC = downloadFile(ANCESTRY_ID) %>% 
-  plyr::rename(c(ID = 'SNP_report:Genotyping_Sample_ID'))
-
-ANCESTRY_ID = 'syn2511399'
-ALL_USED_IDs[length(ALL_USED_IDs)+1] = ANCESTRY_ID
+  dplyr::rename(ID = 'Genotyping_Sample_ID')
+ANCESTRY_ID = 'syn17346100'
+ALL_USED_IDs = c(ALL_USED_IDs, ANCESTRY_ID)
 ANCESTRY.MPP = downloadFile(ANCESTRY_ID) %>% 
-  plyr::rename(c('DNA_report..Genotyping.Sample_ID' = 'SNP_report:Genotyping_Sample_ID'))
-
-ANCESTRY = rbind(ANCESTRY.HBCC[,colnames(ANCESTRY.MPP)[-2]], ANCESTRY.MPP[,colnames(ANCESTRY.MPP)[-2]])
-
+  dplyr::rename('DNA_report..Genotyping.Sample_ID' = 'Genotyping_Sample_ID')
+ANCESTRY = rbind(
+  ANCESTRY.HBCC[,colnames(ANCESTRY.MPP)[-2]], 
+  ANCESTRY.MPP[,colnames(ANCESTRY.MPP)[-2]]
+  )
 # Get genotype ids from synapse. 
-GENOTYPE_ID = 'syn16816490'
+GENOTYPE_ID = "syn18358480"
 ALL_USED_IDs = c(ALL_USED_IDs, GENOTYPE_ID)
 GENOTYPE = downloadFile(GENOTYPE_ID) %>% 
-  dplyr::select(Individual_ID, `SNP_report:Genotyping_Sample_ID`, `SNP_report:Exclude?`) %>% 
+  dplyr::select(Individual_ID, `Genotyping_Sample_ID`, `Exclude`) %>% 
   dplyr::inner_join(ANCESTRY) %>% 
-  dplyr::filter(is.na(`SNP_report:Exclude?`))
+  dplyr::filter(is.na(`Exclude`))
+# Get RNASeq QCmetadata
+CLINICAL_ID = "syn3354385"
+clinical = downloadFile(CLINICAL_ID)
+ASSAY_ID = "syn24173489"
+rnaseq = downloadFile(ASSAY_ID)
+metadata <- right_join(
+  clinical, 
+  rnaseq,
+  )
+ALL_USED_IDs = c(ALL_USED_IDs,CLINICAL_ID, ASSAY_ID)
+metadata = metadata %>%
+  dplyr::select(
+    `Individual ID`, Institution, Cohort, `Reported Gender`, Sex, Ethnicity, 
+    `Age of Death`, `PMI (in hours)`, Dx, pH, SampleID, one_of(
+      'RIN','Brain_Region', 'Ribozero_Batch', 'Library_Batch', 
+      'Flowcell_Batch', 'Mapped_Reads', 'Intragenic_Rate','Intronic_Rate', 
+      'Intergenic_Rate','Genes_Detected', 'Expression_Profiling_Efficiency', 
+      'rRNA_Rate', 'Total_Reads', 'Percent_Aligned', 'Transcripts_Detected',
+      'Exclude?')
+    ) %>% 
+  dplyr::filter(`Brain_Region` %in% c("DLPFC"))
+metadata$`Age of Death`[metadata$`Age of Death`=="90+"] = 90
+metadata$`Age of Death` = as.numeric(metadata$`Age of Death`)
+
+# Merge metadata 
+METADATA = metadata %>%
+  dplyr::left_join(GENOTYPE, by = c("Individual ID" = "Individual_ID")) %>%  
+  dplyr::rename(
+    Region = `Brain_Region`,
+    PMI = `PMI (in hours)`,
+    RIN = `RIN`,
+    ReportExclude = `Exclude?`,
+    GenotypeExclude = Exclude,
+    LibraryBatch = `Library_Batch`,
+    FlowcellBatch = `Flowcell_Batch`,
+    RibozeroBatch = `Ribozero_Batch`,
+    MappedReads = `Mapped_Reads`,
+    IntragenicRate = `Intragenic_Rate`, 
+    IntronicRate = `Intronic_Rate`, 
+    IntergenicRate = `Intergenic_Rate`, 
+    GenesDetected = `Genes_Detected`,
+    ExpProfEfficiency = `Expression_Profiling_Efficiency`, 
+    rRNARate = `rRNA_Rate`,
+    TotalReads = `Total_Reads`, 
+    AlignmentRate = `Percent_Aligned`, 
+    TranscriptsDetected = `Transcripts_Detected`,    
+    TranscriptsDetected = `Transcripts_Detected`,    
+    Reported_Gender = `Reported Gender`,
+    ageOfDeath = `Age of Death`,
+    IndividualID = `Individual ID`) %>%
+  dplyr::select(
+    SampleID, IndividualID, Institution, Cohort,Reported_Gender, Sex, 
+    Ethnicity, ageOfDeath, PMI, Dx, RIN, EV.1, EV.2, EV.3, EV.4, EV.5, 
+    LibraryBatch, FlowcellBatch, RibozeroBatch, MappedReads, TotalReads, 
+    GenesDetected, AlignmentRate, IntragenicRate, IntergenicRate, IntronicRate, 
+    ExpProfEfficiency, rRNARate, TranscriptsDetected, ReportExclude, 
+    GenotypeExclude, pH)
 ```
@@
 ```{r preprocess.data}
-# Merge metadata 
-METADATA = metadata %>%
-  dplyr::left_join(GENOTYPE) %>%      
-  dplyr::rename(Region = `rnaSeq_dissection:Brain_Region`,
-                PMI = `PMI_(in_hours)`,
-                RIN = `rnaSeq_isolation:RIN`, 
-                ReportExclude = `rnaSeq_report:Exclude?`,
-                GenotypeExclude = `SNP_report:Exclude?`,
-                SampleID = Sample_RNA_ID,
-                LibraryBatch = `rnaSeq_report:Library_Batch`,
-                FlowcellBatch = `rnaSeq_report:Flowcell_Batch`,
-                IsolationBatch = `rnaSeq_isolation:RNA_Isolation_Batch`, #isolation batch to replace ribozero batch
-                RibozeroBatch = `rnaSeq_report:Ribozero_Batch`,
-                MappedReads = `rnaSeq_report:Mapped_Reads`,
-                IntragenicRate = `rnaSeq_report:Intragenic_Rate`, IntronicRate = `rnaSeq_report:Intronic_Rate`, 
-                IntergenicRate = `rnaSeq_report:Intergenic_Rate`, GenesDetected = `rnaSeq_report:Genes_Detected`,
-                ExpProfEfficiency = `rnaSeq_report:Expression_Profiling_Efficiency`, rRNARate = `rnaSeq_report:rRNA_Rate`,
-                TotalReads = `rnaSeq_report:Total_Reads`, AlignmentRate = `rnaSeq_report:Percent_Aligned`, TranscriptsDetected = `rnaSeq_report:Transcripts_Detected`) %>%
-  dplyr::select(SampleID, Individual_ID, Institution, Cohort, Reported_Gender,Sex, Ethnicity, ageOfDeath, PMI, Dx, Region,
-                RIN, ReportExclude, GenotypeExclude, EV.1, EV.2, EV.3, EV.4, EV.5, pH,
-                LibraryBatch, FlowcellBatch, IsolationBatch, RibozeroBatch, MappedReads, TotalReads, GenesDetected, AlignmentRate,
-                IntragenicRate, IntergenicRate, IntronicRate, ExpProfEfficiency, rRNARate, TranscriptsDetected) %>%
+
+METADATA = METADATA %>%
   		 		 dplyr::filter(SampleID %in% colnames(geneCountsMerged), !is.na(SampleID)) %>%
-  		 		 dplyr::filter(Region %in% c('DLPFC')) %>%
-			 	 dplyr::mutate(Cohort = forcats::fct_recode(Cohort, `MSSM-Penn-Pitt` = 'MSSM-Penn-Pitt', `NIMH-HBCC`='NIMH-HBCC', `NIMH-HBCC`='NIMH-HBCC-4')) %>%
+  		 		 # dplyr::filter(Region %in% c('DLPFC')) %>%
+			 	 dplyr::mutate(Cohort = forcats::fct_recode(Cohort, `MSSM-Penn-Pitt` = 'MSSM-Penn-Pitt', `NIMH-HBCC`='NIMH-HBCC')) %>% # , `NIMH-HBCC`='NIMH-HBCC-4'
 			 	 dplyr::filter(Dx %in% c('Control', "SCZ"))  %>%
@@
  df_two['ENSG00000167971',]  %>% kable(caption="t-statistics for CASKIN1", row.names=FALSE) %>% kable_styling(full_width=FALSE)
+ df_two['ENSG00000163634',]  %>% kable(caption="t-statistics for CASKIN1", row.names=FALSE) %>% kable_styling(full_width=FALSE)
 ```
@@
 
-```{r store, eval=TRUE}
+```{r store, eval=FALSE}
 if( params$upload ){    
@@
   # Make sure these match the folder name and parent Id in the first chunk.
-  CODE <- Folder(name = "Joint analysis of DLPFC count data", parentId = "syn22416298")
+  CODE <- Folder(name = "Joint analysis of DLPFC count data", parentId = "syn24172641")
   CODE <- synStore(CODE)
@@
 }
-
-#   rmarkdown::render("combined_analysis/combine_all_DLPFC.Rmd")
-
 ```

### 2021-01-14 combined_analysis/combine_all_DLPFC.Rmd
For public release

@@
 
-show_zenith_results( res_meta_zenith, "DxSCZ", "GO", FDR=0.05)
-show_zenith_results( res_meta_zenith, "DxSCZ", "MSigDB", FDR=0.05)
+show_zenith_results( res_meta_zenith, "DxSCZ", "GO", FDR=0.5)
+show_zenith_results( res_meta_zenith, "DxSCZ", "MSigDB", FDR=0.5)
 ```
@@
 
- plot_zenith_results( res_meta_zenith, "DxSCZ", "GO", FDR=0.05) 
+ plot_zenith_results( res_meta_zenith, "DxSCZ", "GO", FDR=0.5) 
 ```
@@
 ```{r, fig.width=9}
-  plot_zenith_results( res_meta_zenith, "DxSCZ", "MSigDB", FDR=0.05)  
+  plot_zenith_results( res_meta_zenith, "DxSCZ", "MSigDB", FDR=0.5)  
 ```

### 2021-01-15 combined_analysis/combine_all_DLPFC.Rmd
For public release

@@
  df_two['ENSG00000167971',]  %>% kable(caption="t-statistics for CASKIN1", row.names=FALSE) %>% kable_styling(full_width=FALSE)
- df_two['ENSG00000163634',]  %>% kable(caption="t-statistics for CASKIN1", row.names=FALSE) %>% kable_styling(full_width=FALSE)
 ```

### 2021-02-17 combined_analysis/eval_resid_in_each.Rmd
add CTS
add more summary info

@@
 
-METADATA %>% group_by(Dx, Institution, Reported_Gender, Region) %>% dplyr::summarise(count = n()) %>% 
-kable(caption="Diagnosis, Institution, Sex, Region") %>% kable_styling(full_width=FALSE) 
+METADATA %>% group_by(Dx, Institution, Reported_Gender, Region) %>% dplyr::summarise(count = n()) %>% kable(caption="Diagnosis, Institution, Sex, Region") %>% kable_styling(full_width=FALSE)
+
+METADATA %>% group_by(Ethnicity, Institution, Region) %>% dplyr::summarise(count = n()) %>% kable(caption="Ethnicity, Institution, Region") %>% kable_styling(full_width=FALSE)
+
+METADATA %>% group_by(Ethnicity, Cohort, Region) %>% dplyr::summarise(count = n()) %>%  kable(caption="Ethnicity, Cohort, Region") %>% kable_styling(full_width=FALSE) 
 ```
@@
                   parentId = CODE$properties$id)
-  synStore(nEXP_OBJ, used = ALL_USED_IDs, executed = thisFile)
+  # synStore(nEXP_OBJ, used = ALL_USED_IDs, executed = thisFile)
 }
@@
 
-    PEER_setPhenoMean(model, t(geneExpr[1:100,i,]))
+    PEER_setPhenoMean(model, t(geneExpr[,i,]))
 
@@
 
-    PEER_setNk(model,2)
+    PEER_setNk(model,30)
 
@@
     # upload latent variables
-    file = paste0('PEER',gsub(" ", '', gsub("-", "_", datID)), '_', CT, ".RDS")
+    file = paste0('PEER_',gsub(" ", '', gsub("-", "_", datID)), '_', CT, ".RDS")
     storeData( latentVariables, file, paste0(datID, " Latent variables for QTL analysis", ' - ', CT))
@@
 ```
+
+for FILE in $(ls PEER*)
+do
+  mv $FILE $(echo $FILE | sed 's/PEER/PEER_/')
+done
 

### 2021-02-18 combined_analysis/eval_resid_in_each.Rmd
add more summary info

@@
 
+```{r cache=FALSE}
+knitr::knit_exit()
+```
+
 ```{r preprocess.data1, results='asis'}
@@
 METADATA %>% group_by(Ethnicity, Cohort, Region) %>% dplyr::summarise(count = n()) %>%  kable(caption="Ethnicity, Cohort, Region") %>% kable_styling(full_width=FALSE)
+
+METADATA %>% group_by(Cohort, Region) %>% dplyr::summarise(count = n()) %>%  kable(caption="Cohort, Region") %>% kable_styling(full_width=FALSE) 
+```
+
+```{r cache=FALSE}
+knitr::knit_exit()
 ```

### 2021-02-18 combined_analysis/eval_resid_in_each.Rmd
add more summary info

@@
 
-# # HBCC - sgACC
+# HBCC - sgACC
 counts.HBCC4 = fread(synGet('syn21867866')$path, data.table=FALSE)
@@
 
-geneCountsMerged = cbind(counts.MPP, counts.HBCC, counts.HBCC4)
+# NIMH-HBCC-5
+counts.HBCC5 = fread(synGet('syn21867762')$path, data.table=FALSE)
+rownames(counts.HBCC5) = counts.HBCC5$Geneid
+counts.HBCC5 = counts.HBCC5[,-c(1:6)]
+
+geneCountsMerged = cbind(counts.MPP, counts.HBCC, counts.HBCC4, counts.HBCC5)
 ```
@@
   dplyr::filter(SampleID %in% colnames(geneCountsMerged), !is.na(SampleID)) %>%
-           # dplyr::filter(Region %in% c('DLPFC')) %>%
   dplyr::mutate(Cohort = forcats::fct_recode(Cohort, `MSSM-Penn-Pitt` = 'MSSM-Penn-Pitt', `NIMH-HBCC`='NIMH-HBCC', `NIMH-HBCC`='NIMH-HBCC-4')) %>%
-         # dplyr::filter(Dx %in% c('Control', "SCZ"))  %>%
   dplyr::filter(ageOfDeath > 17)
@@
 
-# drop individuals who where sequenced twice here
-tab = table(METADATA$Individual_ID)
-
-dropSamples = sapply( names(tab[tab > 1]), function(id){
-
-  idx = which(METADATA$Individual_ID == id)
-  df = METADATA[idx,]
-
-  # remove individuals that have less then max reads
-  i = which(df$MappedReads < max(df$MappedReads))
-  rownames(METADATA)[idx[i]]
-  } )
-METADATA = METADATA[!(rownames(METADATA) %in% dropSamples),]
+# When individuals were sequenced multiple times from the same brain region
+#   for technical replicates or QC failure,
+# Keep sample with most reads
+METADATA = METADATA %>%
+  group_by(Individual_ID, Region) %>% 
+  slice_max(MappedReads)
 ```

### 2021-02-18 combined_analysis/eval_resid_in_each.Rmd
add more summary info

@@
 cd /hpc/users/hoffmg01/work/covarr-de
-# rm -rf molec_profiling_cache/
+# rm -rf combined_analysis/eval_resid_in_each_cache/
 ml git pandoc R/3.6.3 # to get peer
@@
 METADATA = METADATA %>%
-  group_by(Individual_ID, Region) %>% 
+  group_by(Individual_ID, Region, Cohort) %>% 
   slice_max(MappedReads)

### 2021-02-18 combined_analysis/eval_resid_in_each.Rmd
add more summary info

@@
 
-```{r cache=FALSE}
-knitr::knit_exit()
-```
-
 ```{r preprocess.data1, results='asis'}
@@
 METADATA %>% group_by(Cohort, Region) %>% dplyr::summarise(count = n()) %>%  kable(caption="Cohort, Region") %>% kable_styling(full_width=FALSE)
+
+METADATA %>% group_by(Cohort, Region, Ethnicity) %>% dplyr::summarise(count = n()) %>%  kable(caption="Cohort, Region") %>% kable_styling(full_width=FALSE) 
 ```

### 2021-02-24 combined_analysis/eval_resid_in_each.Rmd
add more summary info

@@
 
+# Ethnicity
+# Genotype_Inferred_Ancestry
+# Race
+
 # Get ancestry vector calculated using gemtools
@@
   dplyr::filter(is.na(`SNP_report:Exclude?`))
+```
+
+```{r ancestry}
+fig1 = ggplot(ANCESTRY.MPP, aes(EV.1, EV.2, color=Cluster)) + geom_point() + ggtitle("MPP ancestry")  + theme(aspect.ratio=1)
+fig2 = ggplot(ANCESTRY.HBCC, aes(EV.1, EV.2, color=as.character(cluster))) + geom_point() + ggtitle("HBCC ancestry")  + theme(aspect.ratio=1)
+fig3 = ggplot(ANCESTRY.HBCC, aes(EV.1, EV.2, color=Population)) + geom_point() + ggtitle("HBCC ancestry") + theme(aspect.ratio=1)
+plot_grid(fig1, fig2, fig3, nrow=1)
 ```
@@
 
-```{r cache=FALSE}
-knitr::knit_exit()
-```
-
 ```{r}
@@
 form.lst = list(
-  `MSSM-Penn-Pitt`  = ~ Dx + Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
-  `NIMH-HBCC`       = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)
+  `MSSM-Penn-Pitt`  = ~ Dx + Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3 + EV.1 + EV.2 + EV.3 + EV.4 + EV.5),
+  `NIMH-HBCC`       = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3 + EV.1 + EV.2 + EV.3 + EV.4 + EV.5)
 
@@
 
-  `MSSM-Penn-Pitt`  = ~ Dx + Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)
+`MSSM-Penn-Pitt`  = ~ Dx + Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3) + EV.1 + EV.2 + EV.3 + EV.4 + EV.5
 
-  `NIMH-HBCC`       = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3
+`NIMH-HBCC`       = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3 + EV.1 + EV.2 + EV.3 + EV.4 + EV.5
 

### 2021-02-24 combined_analysis/eval_resid_in_each.Rmd
add more summary info

@@
 ## Summarize samples
+Note that these numbers are based on Ethnicity, rather then genetically determined ancestry used in the filtering for eQTL analysis.
 ```{r summary}

### 2021-02-24 combined_analysis/eval_resid_in_each.Rmd
add more summary info

@@
 
-```{r ancestry}
-fig1 = ggplot(ANCESTRY.MPP, aes(EV.1, EV.2, color=Cluster)) + geom_point() + ggtitle("MPP ancestry")  + theme(aspect.ratio=1)
-fig2 = ggplot(ANCESTRY.HBCC, aes(EV.1, EV.2, color=as.character(cluster))) + geom_point() + ggtitle("HBCC ancestry")  + theme(aspect.ratio=1)
-fig3 = ggplot(ANCESTRY.HBCC, aes(EV.1, EV.2, color=Population)) + geom_point() + ggtitle("HBCC ancestry") + theme(aspect.ratio=1)
-plot_grid(fig1, fig2, fig3, nrow=1)
+```{r ancestry, fig.height=12}
+ggplot(ANCESTRY.MPP, aes(EV.1, EV.2, color=Cluster)) + geom_point() + ggtitle("MPP ancestry")  + theme(aspect.ratio=1)
+ggplot(ANCESTRY.HBCC, aes(EV.1, EV.2, color=as.character(cluster))) + geom_point() + ggtitle("HBCC ancestry")  + theme(aspect.ratio=1)
+ggplot(ANCESTRY.HBCC, aes(EV.1, EV.2, color=Population)) + geom_point() + ggtitle("HBCC ancestry") + theme(aspect.ratio=1)
 ```

### 2021-02-24 combined_analysis/eval_resid_in_each.Rmd
add more summary info

@@
 
-```{r ancestry, fig.height=12}
+```{r ancestry, fig.height=8}
 ggplot(ANCESTRY.MPP, aes(EV.1, EV.2, color=Cluster)) + geom_point() + ggtitle("MPP ancestry")  + theme(aspect.ratio=1)

### 2021-02-24 combined_analysis/eval_resid_in_each.Rmd
add more summary info

@@
 
-```{r ancestry, fig.height=8}
+```{r ancestry, fig.height=6}
 ggplot(ANCESTRY.MPP, aes(EV.1, EV.2, color=Cluster)) + geom_point() + ggtitle("MPP ancestry")  + theme(aspect.ratio=1)

### 2021-03-03 combined_analysis/combine_all_EPHA1.Rmd
initial commit

--- 2021-01-15 combined_analysis/combine_all_DLPFC.Rmd
@@
 cd /hpc/users/hoffmg01/work/covarr-de
-# rm -rf combined_analysis/combine_all_DLPFC_cache
+# rm -rf combined_analysis/combine_all_EPHA1_cache
 ml git pandoc
@@
 
-rmarkdown::render("combined_analysis/combine_all_DLPFC.Rmd")
+rmarkdown::render("combined_analysis/combine_all_EPHA1.Rmd")
 
@@
 
+```{r cache=FALSE}
+knitr::knit_exit()
+```
+
 ```{r sex.confounding, fig.width=5, fig.height=5}

### 2021-03-04 combined_analysis/eval_resid_in_each.Rmd
PEER K = 60
update K = 50

@@
 
-cd /hpc/users/hoffmg01/work/covarr-de
+cd /hpc/users/hoffmg01/work/CMC/covarr-de
 # rm -rf combined_analysis/eval_resid_in_each_cache/
+export R_LIBS=/hpc/users/hoffmg01/.Rlib/R_403
+ml purge
 ml git pandoc R/3.6.3 # to get peer
@@
 
-  PEER_setNk(model,30)
+  PEER_setNk(model,50)
 

### 2021-03-10 combined_analysis/eval_resid_in_each.Rmd
add CTS

@@
 ```{r summary}
-METADATA %>% group_by(Dx, Institution, Region) %>% dplyr::summarise(count = n()) %>% 
+METADATA %>% group_by(Dx, Institution, Region) %>% dplyr::summarise(count = n()) %>% ©
   kable(caption="Diagnosis, Institution, Region") %>% kable_styling(full_width=FALSE)
@@
 
-    PEER_setNk(model,30)
+    PEER_setNk(model,2)
 

### 2021-03-11 combined_analysis/eval_resid_in_each.Rmd
add CTS

--- 2021-03-04 combined_analysis/eval_resid_in_each.Rmd
@@
                   parentId = CODE$properties$id)
-  # synStore(nEXP_OBJ, used = ALL_USED_IDs, executed = thisFile)
+  synStore(nEXP_OBJ, used = ALL_USED_IDs, executed = thisFile)
 }
@@
 form.lst = list(
-  `MSSM-Penn-Pitt`  = ~ Dx + Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*ageOfDeath,
-  `NIMH-HBCC`       = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath)
+  `MSSM-Penn-Pitt`  = ~ Dx + Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*ageOfDeath + EV.1 + EV.2 + EV.3 + EV.4 + EV.5,
+  `NIMH-HBCC`       = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + EV.1 + EV.2 + EV.3 + EV.4 + EV.5)
 
@@
 
-    PEER_setNk(model,30)
+    # PEER_setNmax_iterations(model, 2)
+    PEER_setNk(model,50)
 

### 2021-03-15 combined_analysis/eval_resid_in_each.Rmd
upload correct cell type

@@
     # fit PEER
-    model = PEER()
+    # model = PEER()
 
-    PEER_setPhenoMean(model, t(geneExpr[,i,]))
+    # PEER_setPhenoMean(model, t(geneExpr[,i,]))
 
-    PEER_setCovariates(model, cov_matrix)
+    # PEER_setCovariates(model, cov_matrix)
 
-    # PEER_setNmax_iterations(model, 2)
-    PEER_setNk(model,50)
+    # # PEER_setNmax_iterations(model, 2)
+    # PEER_setNk(model,50)
 
-    PEER_update(model)
+    # PEER_update(model)
 
@@
     # first ncol(cov_matrix) columns are the oringal covariates
-    latentVariables = PEER_getX(model)[,-c(1:ncol(cov_matrix))]
-    rownames(latentVariables) = rownames(cov_matrix)
-    colnames(latentVariables) = paste0("PEER_", 1:ncol(latentVariables))
+    # latentVariables = PEER_getX(model)[,-c(1:ncol(cov_matrix))]
+    # rownames(latentVariables) = rownames(cov_matrix)
+    # colnames(latentVariables) = paste0("PEER_", 1:ncol(latentVariables))
 
@@
     file = paste0('geneExpr_',gsub(" ", '', gsub("-", "_", datID)), '_', CT,".RDS")
-    storeData( geneExpr, file, paste0(datID, " Observed expression", ' - ', CT))
+    storeData( geneExpr[,i,], file, paste0(datID, " Observed expression", ' - ', CT))
 
-    # upload covariate matrix
-    file = paste0('covariateMatrix_',gsub(" ", '', gsub("-", "_", datID)), '_', CT, ".RDS")
-    storeData( cov_matrix, file, paste0(datID, " Covariates for QTL analysis", ' - ' , CT))
+    # # upload covariate matrix
+    # file = paste0('covariateMatrix_',gsub(" ", '', gsub("-", "_", datID)), '_', CT, ".RDS")
+    # storeData( cov_matrix, file, paste0(datID, " Covariates for QTL analysis", ' - ' , CT))
 
-    # upload latent variables
-    file = paste0('PEER_',gsub(" ", '', gsub("-", "_", datID)), '_', CT, ".RDS")
-    storeData( latentVariables, file, paste0(datID, " Latent variables for QTL analysis", ' - ', CT))
+    # # upload latent variables
+    # file = paste0('PEER_',gsub(" ", '', gsub("-", "_", datID)), '_', CT, ".RDS")
+    # storeData( latentVariables, file, paste0(datID, " Latent variables for QTL analysis", ' - ', CT))
   }

### 2021-03-15 combined_analysis/eval_resid_in_each.Rmd
upload correct cell type

--- 2021-03-11 combined_analysis/eval_resid_in_each.Rmd
@@
 
-    # PEER_setNmax_iterations(model, 2)
+    # # PEER_setNmax_iterations(model, 2)
     PEER_setNk(model,50)
@@
     file = paste0('geneExpr_',gsub(" ", '', gsub("-", "_", datID)), '_', CT,".RDS")
-    storeData( geneExpr, file, paste0(datID, " Observed expression", ' - ', CT))
+    storeData( geneExpr[,i,], file, paste0(datID, " Observed expression", ' - ', CT))
 
-    # upload covariate matrix
+    # # upload covariate matrix
     file = paste0('covariateMatrix_',gsub(" ", '', gsub("-", "_", datID)), '_', CT, ".RDS")
@@
 
-    # upload latent variables
+    # # upload latent variables
     file = paste0('PEER_',gsub(" ", '', gsub("-", "_", datID)), '_', CT, ".RDS")

### 2021-05-19 macaque/macaque_DE.Rmd
initial commit

---
title: "Rhesus macaque drug study"
author: "[Gabriel Hoffman](http://gabrielhoffman.github.io)"
date: "Run on `r Sys.time()`"
output: html_document
---

<!---

cd /Users/gabrielhoffman/workspace/repos/CMC/covarr-de/macaque
# rm -rf combined_analysis/combine_all_EPHA1_cache

R

rmarkdown::render("macaque_DE.Rmd")

--->

```{r setup, include=FALSE, echo=FALSE}
library(data.table)
library(synapser)
library(limma)
library(edgeR)
library(corrplot)
library(magrittr)
library(knitr)
library(kableExtra)
library(RColorBrewer)
library(biomaRt)

synLogin()

knitr::opts_chunk$set(
  echo = TRUE,
  warning=FALSE,
  message=TRUE,
  error = FALSE,
  tidy = FALSE,
  dev = c('png', 'pdf'),
  cache = TRUE)
```

## Get data
```{r get.data}
# Rhesus data: https://www.synapse.org/#!Synapse:syn11617745

# reads RNA-seq data
geneCounts = data.frame(fread( synGet('syn11615855')$path))
rownames(geneCounts) = geneCounts$GeneFeature
geneCounts = geneCounts[,-1]

# read sample metadata
info = data.frame(fread( synGet('syn11638462')$path))
rownames(info) = info$DLPFC_RNA_Sequencing_Sample_ID

# sort info based on geneCounts
idx = match(colnames(geneCounts), rownames(info))
info = info[idx,]

# Create drug and dosage variable
info$treatmentType = factor(info$treatmentType, c('placebo', 'clozapine', 'haloperidol'))
info$DrugExposure = with(info, paste0(treatmentType, ' (', dose..mg.kg.d., ' mg/kg)'))
info$DrugExposure = gsub(" \\(NA mg/kg\\)", '', info$DrugExposure)
info$DrugExposure = factor(info$DrugExposure, unique(info$DrugExposure))

# get mapping between Rh mac and human genes
#################

# genome annotation
rhmac = useMart("ensembl", dataset = "mmulatta_gene_ensembl")

geneInfo <- getBM(attributes=c("ensembl_gene_id", "hsapiens_homolog_ensembl_gene","hsapiens_homolog_associated_gene_name"), mart=rhmac, values = rownames(geneCounts))
```

## Differential expession analysis
```{r DE.analysis, fig.width=5, fig.height=5}
# Differential analysis
dsgn = model.matrix( ~ DrugExposure, info)

# voom
keep = filterByExpr(geneCounts, dsgn)
vobj = voom(geneCounts[keep,], dsgn, plot=TRUE)

# lmFit
fit = lmFit(vobj, dsgn)
fit = eBayes(fit)

# get and combine signatures
df_signatures = lapply( colnames(coef(fit))[-1], function(drug){
  tab = topTable(fit, coef=drug, sort.by="none", number=Inf)[,'logFC', drop=FALSE]
  colnames(tab) = gsub("^Drug", '', drug)
  tab
})
df_signatures = do.call(cbind, df_signatures)
df_signatures = merge(df_signatures, geneInfo, by.x="row.names", by.y="ensembl_gene_id")
```

## Compare with CommonMind
```{r cmc}
df_cmc = readRDS( synGet('syn22416508')$path)$DxSCZ

trim_ensembl_ids = function(x){
  gsub("(.*)\\.(.*)", "\\1", x) 
}
df_cmc$ENSEMBL = trim_ensembl_ids( rownames(df_cmc) )

df_signatures = merge(df_signatures, 
                      data.frame(CMCv4 = with(df_cmc, logFC), ENSEMBL=df_cmc$ENSEMBL),
                       by.x="hsapiens_homolog_ensembl_gene", 
                       by.y="ENSEMBL")
```

```{r compare, fig.width=5, fig.height=5}
cols = c(gsub("^Drug", '',colnames(coef(fit))[-1]), 'CMCv4')

C = cor(df_signatures[,cols], method="spearman")

corrplot(C, col = colorRampPalette(c("blue", "red"))(100), main = "Spearman correlation")
```

```{r table}
data.frame(`Fraction of variance` = C['CMCv4',-4]^2) %>%
  kbl() %>%
  kable_styling(full_width=FALSE)
```

```{r test, eval=FALSE}
lapply(names(C['CMCv4',-4]), function(drug){
  cor.test(df_signatures[,drug], df_signatures[,'CMCv4'], method="spearman")
})
```

### 2021-05-19 macaque/macaque_DE.Rmd
initial commit

@@
   tab = topTable(fit, coef=drug, sort.by="none", number=Inf)[,'logFC', drop=FALSE]
-  colnames(tab) = gsub("^Drug", '', drug)
+  colnames(tab) = gsub("^DrugExposure", '', drug)
   tab
@@
 ```{r compare, fig.width=5, fig.height=5}
-cols = c(gsub("^Drug", '',colnames(coef(fit))[-1]), 'CMCv4')
+cols = c(gsub("^DrugExposure", '',colnames(coef(fit))[-1]), 'CMCv4')
 

### 2021-07-26 individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_Diagnosis.Rmd

@@
 
-# Store contrasts
+# Store formula
 saveRDS(formula, file = 'CMC_DLPFC_Dx_formula.rds')

### 2021-07-26 individual-cohorts/split_by_analysis/HBCC_DLPFC_DE_DxSex.Rmd
update provenance - all cohorts

@@
 ```
-`r formula`
+
 ### Residual calculation

### 2021-07-28 deconvolution/fans4/molec_profiling.Rmd
Initial commit
freeze 2

@@
 # https://hoffmg01.u.hpc.mssm.edu/molec_profiling.html
+
+cd /Users/gabrielhoffman/workspace/repos/CMC/covarr-de/deconvolution/fans4
+R
+
+rmarkdown::render("molec_profiling.Rmd", output_dir='./', intermediates_dir='./')
 
@@
 library(synapser)
+library(githubr)
 synLogin()
@@
 Ids = c('syn22321012', 'syn22321013', 'syn22321014', 'syn22321016')
-# file = "/sc/arion/projects/CommonMind/roussp01a/MOLECULAR_PROFILING/rnaseq/step1-merged/featureCount/featureCount.gene.tsv.gz"
+
+# Read in RNA-seq featureCounts
+###############################
+
+# Phase 1
 file = synGet('syn22321012')$path
 df_featureCounts_gene = fread( file )
+
+# Phase 2+3
+# syn26000845 is really phase2 again
+ids = c('syn26000845','syn25976117', 'syn25976127')[-1]
+df_read = lapply(ids, function(id){
+  df = fread(synGet(id)$path)
+  colnames(df) = gsub("^Sample_", '', colnames(df))
+  colnames(df) = gsub("-", '.', colnames(df))
+  df
+  })
+
+# save batch info
+df_batch = data.frame(Name = colnames(df_featureCounts_gene)[-c(1:6)], Batch=1)
+df_batch = rbind(df_batch, data.frame(Name = colnames(df_read[[1]][,-c(1:6)]), Batch=2))
+df_batch = rbind(df_batch, data.frame(Name = colnames(df_read[[2]][,-c(1:6)]), Batch=3))
+df_batch$Batch = as.character(df_batch$Batch)
+
+# sapply(df_read, ncol)
+df_featureCounts_gene = cbind(df_featureCounts_gene, 
+                          df_read[[1]][,-c(1:6)], 
+                          df_read[[2]][,-c(1:6)])
 
@@
 ```{r QC, fig.width=18, fig.height=74}
+# Phase 1 
 # file = '/sc/arion/projects/CommonMind/roussp01a/MOLECULAR_PROFILING/rnaseq/step1-merged/QCmetrics/qc_metrics.RDS'
@@
 
+# Phase 2+3
+ids = c('syn26000850', 'syn25976122', 'syn25976136')[-1]
+df_read = lapply(ids, function(id){
+  qc.in = read.table(synGet(id)$path)
+  rownames(qc.in) = gsub('-', '.', rownames(qc.in))
+  qc.in
+  })
+
+# # sapply(df_read, ncol)
+# df_featureCounts_gene = cbind(df_featureCounts_gene, 
+#                           df_read[[1]][,-c(1:6)], 
+#                           df_read[[2]][,-c(1:6)])
+
 # extract information from sample names
@@
 # add information about each sample
-info = merge(info, metadata1, by.x="DissectionID", by.y='Dissection.ID')
-info = merge(info, qc_metrics, by.x="Name", by.y='row.names')
+info = merge(info, metadata1, by.x="DissectionID", by.y='Dissection.ID', all.x=TRUE)
+info = merge(info, qc_metrics, by.x="Name", by.y='row.names', all.x=TRUE)
 rownames(info) = info$Name
@@
 
+# Merge with batch info
+info = merge(info, df_batch, by="Name")
+rownames(info) = info$Name
+
 # set order to be the same as geneCounts
+# Must be last change to info
 info = info[match(colnames(geneCounts), info$Name),]
@@
 
-df = reshape2::melt(geneExpr)
-df = merge(df, info[,'CellType',drop=FALSE], by.x = "Var2", by.y="row.names")
+df = data.table(reshape2::melt(geneExpr))
+setkey(df, "Var2")
+df = merge(df, info[,c("Name", 'CellType')], by.x = "Var2", by.y="Name")
 
@@
 ### PCA
-```{r PCA}
-
-geneExpr = edgeR::cpm(dge, log=TRUE)
-
+```{r PCA, fig.width=10, fig.height=5}
 dcmp = prcomp( t(geneExpr), scale=TRUE)
@@
 ylab = paste0('PC2 (', format(varFrac[2]*100, digits=3), '%)')
-ggplot(df_pca, aes(PC1, PC2, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=colorCodes) + xlab(xlab) + ylab(ylab)
+fig1 = ggplot(df_pca, aes(PC1, PC2, color=CellType, shape=Batch)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=colorCodes) + xlab(xlab) + ylab(ylab)
 
@@
 ylab = paste0('PC3 (', format(varFrac[3]*100, digits=3), '%)')
-ggplot(df_pca, aes(PC2, PC3, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=colorCodes) + xlab(xlab) + ylab(ylab)
+fig2 = ggplot(df_pca, aes(PC2, PC3, color=CellType, shape=Batch)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=colorCodes) + xlab(xlab) + ylab(ylab)
+
+plot_grid(fig1, fig2 + theme(legend.position="none"), axis="tbr", align="v")
+```
+
+#### Batch corrected
+```{r PCA.batch, fig.width=10, fig.height=5}
+fit = dream(geneExpr, ~ Batch, info)
+
+dcmp = prcomp( t(residuals(fit)), scale=TRUE)
+
+df_pca = data.frame(dcmp$x)
+
+df_pca = merge(df_pca, info[,c("Name", "CellType", "Batch")], by.x="row.names", by.y="Name")
+
+varFrac = dcmp$sdev^2 / sum(dcmp$sdev^2)
+
+xlab = paste0('PC1 (', format(varFrac[1]*100, digits=3), '%)')
+ylab = paste0('PC2 (', format(varFrac[2]*100, digits=3), '%)')
+fig1 = ggplot(df_pca, aes(PC1, PC2, color=CellType, shape=Batch)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=colorCodes) + xlab(xlab) + ylab(ylab)
+
+xlab = paste0('PC2 (', format(varFrac[2]*100, digits=3), '%)')
+ylab = paste0('PC3 (', format(varFrac[3]*100, digits=3), '%)')
+fig2 = ggplot(df_pca, aes(PC2, PC3, color=CellType, shape=Batch)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=colorCodes) + xlab(xlab) + ylab(ylab)
+
+plot_grid(fig1, fig2 + theme(legend.position="none"), axis="tbr", align="v")
+```
+
+```{r predict.celltype}
+library(nnet)
+library(PRROC)
+
+hush=function(code){
+  sink("NUL") # use /dev/null in UNIX
+  tmp = code
+  sink()
+  return(tmp)
+}
+
+df_class = lapply(1:10, function(k){
+
+  fit = hush(multinom(df_pca$CellType ~ X[,1:k,drop=FALSE]))
+
+  data.table(k =k, n_mismatch = sum(predict(fit) != df_pca$CellType))
+  })
+df_class = do.call(rbind, df_class)
+
+fig1 = ggplot(df_class, aes(factor(k), n_mismatch)) + geom_bar(stat="identity") + theme_classic() + scale_y_continuous(expand=c(0,0)) + ggtitle("Ability to predict each cell type") + theme(aspect.ratio=1)
+
+df_class = lapply(1:10, function(k){
+
+  fit = hush(multinom(df_pca$CellType ~ X[,1:k,drop=FALSE]))
+
+  Y_pred = predict(fit, type="probs")
+
+  df_class = lapply( levels(df_pca$CellType), function(CT){
+
+    y0 = Y_pred[df_pca$CellType == CT,CT]
+    y1 = Y_pred[df_pca$CellType != CT,CT]
+
+    pr = pr.curve(scores.class0 = y0, scores.class1 = y1, curve=TRUE, rand.compute=TRUE)
+
+    data.table(k = k, CellType = CT, AUPR = pr$auc.integral, AUPR.rand = pr$rand$auc.integral)
+    })
+  do.call(rbind, df_class)
+  })
+df_class = do.call(rbind, df_class)
+
+fig2 = ggplot(df_class, aes(k, AUPR, color=CellType)) + geom_point() + geom_line() + theme_classic() + scale_color_manual(values = colorCodes) + scale_x_continuous(breaks = 1:10) + ggtitle("Ability to predict each cell type") + theme(aspect.ratio=1)
+
+plot_grid(fig1, fig2, axis="tbr", align="v")
 ```
@@
 ```{r vp}
-form = ~ (1|CellType) + (1|Dx) + (1|CellType:Dx) + (1|Individual.ID) + (1|Gender)
+form = ~ (1|CellType) + (1|Dx) + (1|CellType:Dx) + (1|Individual.ID) + (1|Gender) + (1|Batch)
 vp = fitExtractVarPartModel( geneExpr, form, info, BPPARAM=SnowParam(6, progressbar=TRUE))
@@
 ```{r DE_process}
-form = ~ 0 + CellType + (1|Individual.ID) + (1|Gender)
+form = ~ 0 + CellType + (1|Individual.ID) + (1|Gender) + (1|Batch)
 
@@
 ggplot(df_deCount, aes(CT1, CT2, fill=pi1)) + geom_tile() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + scale_fill_gradient(name = quote(pi[1]), low="white", high="red", limits=c(0, 1)) + geom_text(aes(label=format(pi1, digits=3))) + xlab('') + ylab('') + ggtitle(bquote(pi[1]))
+```
+
+```{r cache=FALSE}
+knitr::knit_exit()
 ```
@@
 # Differential expression
-file = Roussos_MSSM_MolProf_FANS_4_diffExpression.RDS
+file = "Roussos_MSSM_MolProf_FANS_4_diffExpression.RDS"
 

### 2021-07-28 deconvolution/fans4/molec_profiling.Rmd
freeze 2

@@
 form = ~ (1|CellType) + (1|Individual.ID) + (1|Gender)
-vobj = voomWithDreamWeights(dge, form, info, plot=TRUE, save.plot=TRUE, BPPARAM=SnowParam(6, progressbar=TRUE))
+vobj = voomWithDreamWeights(dge, form, info, plot=TRUE, save.plot=TRUE, BPPARAM=SnowParam(36, progressbar=TRUE))
 ```
@@
 form = ~ (1|CellType) + (1|Dx) + (1|CellType:Dx) + (1|Individual.ID) + (1|Gender) + (1|Batch)
-vp = fitExtractVarPartModel( geneExpr, form, info, BPPARAM=SnowParam(6, progressbar=TRUE))
+vp = fitExtractVarPartModel( geneExpr, form, info, BPPARAM=SnowParam(36, progressbar=TRUE))
 
@@
 
-fitBasic = dream( vobj, form, info, L=L, BPPARAM=SnowParam(6, progressbar=TRUE))
+fitBasic = dream( vobj, form, info, L=L, BPPARAM=SnowParam(36, progressbar=TRUE))
 ```

### 2021-07-28 common_functions/common_functions.R
add hush

@@ -18,6 +18,13 @@ library(Rfast)
 library(GSEABase)
 })
 
+hush = function(code){
+  sink("/dev/null") # use /dev/null in UNIX
+  tmp = code
+  sink()
+  return(tmp)
+}
+
 # For ENSEMBL id ENSG00000279457.4, return ENSG00000279457
 trim_ensembl_ids = function(x){
   gsub("(.*)\\.(.*)", "\\1", x) 

### 2021-07-28 deconvolution/fans4/molec_profiling.Rmd
fix dependency

@@
 
-cd /hpc/users/hoffmg01/work/cmc_analysis/src
+# cd /hpc/users/hoffmg01/work/cmc_analysis/src
+cd /hpc/users/hoffmg01/work/CMC/covarr-de/deconvolution/fans4
 # rm -rf molec_profiling_cache/
@@
 library(synapser)
+library(nnet)
+library(PRROC)
 library(githubr)
@@
 library(EnsDb.Hsapiens.v86)
-ensdb = EnsDb.Hsapiens.v86
-})
+})
+
+hush=function(code){
+  sink("NUL") # use /dev/null in UNIX
+  tmp = code
+  sink()
+  return(tmp)
+}
 
@@
 ```{r predict.celltype}
-library(nnet)
-library(PRROC)
-
-hush=function(code){
-  sink("NUL") # use /dev/null in UNIX
-  tmp = code
-  sink()
-  return(tmp)
-}
-
 df_class = lapply(1:10, function(k){

### 2021-07-28 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 
-hush=function(code){
-  sink("NUL") # use /dev/null in UNIX
-  tmp = code
-  sink()
-  return(tmp)
-}
-
 source("../../common_functions/common_functions.R")
@@
 ```{r predict.celltype}
+
+X = as.matrix(df_pca[,2:200])
+
 df_class = lapply(1:10, function(k){
@@
 ```{r voom}
-form = ~ (1|CellType) + (1|Individual.ID) + (1|Gender)
+form = ~ (1|CellType) + (1|Individual.ID) + (1|Gender) + (1|Batch)
 vobj = voomWithDreamWeights(dge, form, info, plot=TRUE, save.plot=TRUE, BPPARAM=SnowParam(36, progressbar=TRUE))

### 2021-07-28 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 git pull
+export OMP_NUM_THREADS=1
 R
@@
 
-fitBasic = dream( vobj, form, info, L=L, BPPARAM=SnowParam(36, progressbar=TRUE))
+fitBasic = dream( vobj, form, info, L=L, BPPARAM=SnowParam(12, progressbar=TRUE))
 ```

### 2021-07-28 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 form = ~ (1|CellType) + (1|Individual.ID) + (1|Gender) + (1|Batch)
-vobj = voomWithDreamWeights(dge, form, info, plot=TRUE, save.plot=TRUE, BPPARAM=SnowParam(36, progressbar=TRUE))
+vobj = voomWithDreamWeights(dge, form, info, plot=TRUE, save.plot=TRUE, BPPARAM=SnowParam(12, progressbar=TRUE))
 ```
@@
 form = ~ (1|CellType) + (1|Dx) + (1|CellType:Dx) + (1|Individual.ID) + (1|Gender) + (1|Batch)
-vp = fitExtractVarPartModel( geneExpr, form, info, BPPARAM=SnowParam(36, progressbar=TRUE))
+vp = fitExtractVarPartModel( geneExpr, form, info, BPPARAM=SnowParam(12, progressbar=TRUE))
 
@@
   form = ~ CellType + Gender
-  vobj = voomWithDreamWeights(dge, form, infoSub, plot=FALSE, save.plot=TRUE, BPPARAM=SnowParam(6, progressbar=TRUE))
+  vobj = voomWithDreamWeights(dge, form, infoSub, plot=FALSE, save.plot=TRUE, BPPARAM=SnowParam(12, progressbar=TRUE))
 
   form = ~ CellType + Gender
-  fit = dream( vobj, form, infoSub, BPPARAM=SnowParam(6, progressbar=TRUE))
+  fit = dream( vobj, form, infoSub, BPPARAM=SnowParam(12, progressbar=TRUE))
   fit = eBayes(fit)

### 2021-07-28 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 ### Cell type
-```{r DE_process}
+```{r DE_process, eval=FALSE}
 form = ~ 0 + CellType + (1|Individual.ID) + (1|Gender) + (1|Batch)
@@
 ```
+
+idx = as.integer(seq(1, 2000, length.out=100))
+lapply(2:length(idx), function(i){
+  message(i, '     ')
+fitBasic = dream( vobj[idx[i-1]:idx[i],], form, info, L=L, BPPARAM=SerialParam())
+
+})
+
+i = 173:173
+fitBasic = dream( vobj[i,,drop=FALSE], form, info, L=L, BPPARAM=SerialParam())
 

### 2021-07-28 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 rmarkdown::render("molec_profiling.Rmd", output_dir='./', intermediates_dir='./')
-
---->
 
@@
 
-idx = as.integer(seq(1, 2000, length.out=100))
-lapply(2:length(idx), function(i){
-  message(i, '     ')
-fitBasic = dream( vobj[idx[i-1]:idx[i],], form, info, L=L, BPPARAM=SerialParam())
-
-})
-
-i = 173:173
-fitBasic = dream( vobj[i,,drop=FALSE], form, info, L=L, BPPARAM=SerialParam())
+# idx = as.integer(seq(1, 2000, length.out=100))
+# lapply(2:length(idx), function(i){
+#   message(i, '     ')
+# fitBasic = dream( vobj[idx[i-1]:idx[i],], form, info, L=L, BPPARAM=SerialParam())
+
+# })
+
+# i = 445
+# form = ~ 0 + CellType + (1|Individual.ID) + (1|Gender) + (1|Batch)
+# fitBasic = dream( vobj[i,,drop=FALSE], form, info, BPPARAM=SerialParam())
 

### 2021-07-30 deconvolution/fans4/molec_profiling.Rmd
bug fix

--- 2021-07-28 deconvolution/fans4/molec_profiling.Rmd
@@
 
-# Prepare all pairwise contrasts
 index = combn(nlevels(info$CellType),2)
 
-L = lapply( 1:ncol(index), function(j){
+lvl = rbind(levels(info$CellType)[index[1,]], levels(info$CellType)[index[2,]])
+contrasts = apply(lvl, 2, function(x){
+  paste0('CellType', x[1], " - CellType", x[2])
+})
+names(contrasts) = apply(lvl, 2, function(x){
+  paste0(x[1], "_vs_", x[2])
+})
+
+contrasts['GABAGLU_vs_Other'] = "(CellTypeGABA + CellTypeGLU)/2 - (CellTypeMgAs + CellTypeOlig)/2"
+contrasts['MgAs_vs_Other'] = "CellTypeMgAs - (CellTypeGABA + CellTypeGLU + CellTypeOlig)/3"
+contrasts['Olig_vs_Other'] = "CellTypeOlig - (CellTypeGABA + CellTypeGLU + CellTypeMgAs)/3"
+
+# combined contrast matrix
+L = makeContrastsDream( form, info, contrasts = contrasts)
+
+fitBasic = dream( vobj, form, info, L=L, BPPARAM=SnowParam(36, progressbar=TRUE))
+```
+
+```{r plotContrasts}
+plotContrasts(round(L,2))   
+```
+
+```{r DE.analysis}
+
+index = combn(nlevels(info$CellType),2)
+
+df_deCount = lapply( 1:ncol(index), function(j){
 
@@
 
-  L = getContrast( vobj, form, info, paste0('CellType', c(CT1, CT2)))
-
-  L = as.matrix(L)
-  colnames(L) = paste0(CT1, '.', CT2)
-
-  L
-})
-L = do.call(cbind, L)
-
-L = cbind(L, data.frame(GABAGLU.Other = c(1, 1, -1, -1),
-                    MgAs.Other = c(-1/3, -1/3, -1/3, 1),
-                    Olig.Other = c(-1/3, -1/3, 1, -1/3) ))
-
-fitBasic = dream( vobj, form, info, L=L, BPPARAM=SnowParam(12, progressbar=TRUE))
-```
-
-```{r plotContrasts}
-plotContrasts(round(L,2)) + theme(axis.text.x=element_text(angle=45, hjust=1))   
-```
-
-```{r DE.analysis}
+    key = paste0(CT1, '_vs_', CT2)
+
+    tabl = topTable(fitBasic, coef=key, number=Inf)
+
+    data.frame(CT1, CT2, key, count = sum(tabl$adj.P.Val < 0.05), pi1 = 1 - qvalue(tabl$P.Value)$pi0)
+})
+df_deCount = do.call(rbind, df_deCount)
+
+ggplot(df_deCount, aes(CT1, CT2, fill=count)) + geom_tile() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + scale_fill_gradient(low="white", high="red", limits=c(0, 30000)) + geom_text(aes(label=format(count, big.mark=','))) + xlab('') + ylab('') + ggtitle("Count differentiall expression genes")
+
+ggplot(df_deCount, aes(CT1, CT2, fill=pi1)) + geom_tile() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + scale_fill_gradient(name = quote(pi[1]), low="white", high="red", limits=c(0, 1)) + geom_text(aes(label=format(pi1, digits=3))) + xlab('') + ylab('') + ggtitle(bquote(pi[1]))
+```
+
+```{r DE2}
 
@@
 
-df_deCount = lapply( 1:ncol(index), function(j){
+df = lapply( 1:ncol(index), function(j){
 
@@
 
-    key = paste0(CT1, '.', CT2)
+    key = paste0(CT1, '_vs_', CT2)
 
@@
 
-    data.frame(CT1, CT2, key, count = sum(tabl$adj.P.Val < 0.05), pi1 = 1 - qvalue(tabl$P.Value)$pi0)
-})
-df_deCount = do.call(rbind, df_deCount)
-
-ggplot(df_deCount, aes(CT1, CT2, fill=count)) + geom_tile() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + scale_fill_gradient(low="white", high="red", limits=c(0, 30000)) + geom_text(aes(label=format(count, big.mark=','))) + xlab('') + ylab('') + ggtitle("Count differentiall expression genes")
-
-ggplot(df_deCount, aes(CT1, CT2, fill=pi1)) + geom_tile() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + scale_fill_gradient(name = quote(pi[1]), low="white", high="red", limits=c(0, 1)) + geom_text(aes(label=format(pi1, digits=3))) + xlab('') + ylab('') + ggtitle(bquote(pi[1]))
-```
-
-```{r cache=FALSE}
-knitr::knit_exit()
-```
-
-```{r DE2}
+    data.frame(CT1, CT2, key, tabl)
+})
+df = do.call(rbind, df)
+
+ggplot(df, aes(logFC, ..count..)) + geom_density(alpha = .2) + facet_wrap(~key) + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) 
+```
+
+## Plots of top differentially expressed genes
+```{r DE.marker_genes, fig.width=15, fig.height=30}
 
@@
 
-df = lapply( 1:ncol(index), function(j){
+ensIDs = sapply( 1:ncol(index), function(j){
 
@@
 
-    key = paste0(CT1, '.', CT2)
-
-    tabl = topTable(fitBasic, coef=key, number=Inf)
-
-    data.frame(CT1, CT2, key, tabl)
-})
-df = do.call(rbind, df)
-
-ggplot(df, aes(logFC, ..count..)) + geom_density(alpha = .2) + facet_wrap(~key) + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) 
-```
-
-## Plots of top differentially expressed genes
-```{r DE.marker_genes, fig.width=15, fig.height=30}
-
-index = combn(nlevels(info$CellType),2)
-
-ensIDs = sapply( 1:ncol(index), function(j){
-
-    CT1 = levels(info$CellType)[index[1,j]]
-    CT2 = levels(info$CellType)[index[2,j]]
-
-    key = paste0(CT1, '.', CT2)
+    key = paste0(CT1, '_vs_', CT2)
 
@@
   CT2 = levels(info$CellType)[index[2,j]]
-  key = paste0(CT1, '.', CT2)
+  key = paste0(CT1, '_vs_', CT2)
 
@@
 # system("ml python; synapse add --parentid syn22243354 Roussos_MSSM_MolProf_FANS_4_diffExpression.RDS")
+```
+
+```{r cache=FALSE}
+knitr::knit_exit()
 ```

### 2021-07-30 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 
-df_hugo = select(org.Hs.eg.db, unique(as.character(df$Var1)), "SYMBOL", "ENSEMBL")
+df_hugo = AnnotationDbi::select(org.Hs.eg.db, unique(as.character(df$Var1)), "SYMBOL", "ENSEMBL")
 df = merge(df, df_hugo, by.x="Var1", by.y="ENSEMBL")
@@
 
-  df_hugo = select(org.Hs.eg.db, unique(as.character(df$Var1)), "SYMBOL", "ENSEMBL")
+  df_hugo = AnnotationDbi::select(org.Hs.eg.db, unique(as.character(df$Var1)), "SYMBOL", "ENSEMBL")
   df = merge(df, df_hugo, by.x="Var1", by.y="ENSEMBL")
@@
 
-  geneTab = select(ensdb, keys=names(isexpr[isexpr]), columns=c("GENEID", "GENEBIOTYPE"), keytype="GENEID")
+  geneTab = AnnotationDbi::select(ensdb, keys=names(isexpr[isexpr]), columns=c("GENEID", "GENEBIOTYPE"), keytype="GENEID")
 
@@
 
-    geneTab = select(ensdb, keys=rownames(tab), columns=c("GENEID", "GENEBIOTYPE"), keytype="GENEID")
+    geneTab = AnnotationDbi::select(ensdb, keys=rownames(tab), columns=c("GENEID", "GENEBIOTYPE"), keytype="GENEID")
 

### 2021-07-30 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 
-ggplot(df_deCount, aes(CT1, CT2, fill=count)) + geom_tile() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + scale_fill_gradient(low="white", high="red", limits=c(0, 30000)) + geom_text(aes(label=format(count, big.mark=','))) + xlab('') + ylab('') + ggtitle("Count differentiall expression genes")
+df_deCount$CT1 = factor(df_deCount$CT1, levels(info$CellType))
+df_deCount$CT2 = factor(df_deCount$CT2, levels(info$CellType))
+
+ggplot(df_deCount, aes(CT1, CT2, fill=count)) + geom_tile() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + scale_fill_gradient(low="white", high="red", limits=c(0, 30000)) + geom_text(aes(label=format(count, big.mark=','))) + xlab('') + ylab('') + ggtitle("Count differential expression genes")
 
@@
 
-ggplot(df, aes(CellType, value, fill=CellType)) + geom_boxplot() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position = "none") + scale_fill_manual("Cell Type", values=colorCodes) + facet_wrap(~SYMBOL, ncol=5) + ylab(bquote(log[2]~CPM))
+ggplot(df, aes(CellType, value, fill=CellType)) + geom_boxplot() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + scale_fill_manual("Cell Type", values=colorCodes) + facet_wrap(~SYMBOL, ncol=5) + ylab(bquote(log[2]~CPM))
 ```

### 2021-07-30 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 
-```{r cache=FALSE}
-knitr::knit_exit()
-```
-
 ## Plots of top differentially expressed genes
@@
 
-  geneTab = AnnotationDbi::select(ensdb, keys=names(isexpr[isexpr]), columns=c("GENEID", "GENEBIOTYPE"), keytype="GENEID")
+  geneTab = AnnotationDbi::select(EnsDb.Hsapiens.v86, keys=names(isexpr[isexpr]), columns=c("GENEID", "GENEBIOTYPE"), keytype="GENEID")
 
@@
 
-    geneTab = AnnotationDbi::select(ensdb, keys=rownames(tab), columns=c("GENEID", "GENEBIOTYPE"), keytype="GENEID")
+    geneTab = AnnotationDbi::select(EnsDb.Hsapiens.v86, keys=rownames(tab), columns=c("GENEID", "GENEBIOTYPE"), keytype="GENEID")
 

### 2021-07-30 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 
-  form = ~ CellType + Gender
+  form = ~ CellType + Gender + Batch
   vobj = voomWithDreamWeights(dge, form, infoSub, plot=FALSE, save.plot=TRUE, BPPARAM=SnowParam(12, progressbar=TRUE))
 
-  form = ~ CellType + Gender
+  form = ~ CellType + Gender + Batch
   fit = dream( vobj, form, infoSub, BPPARAM=SnowParam(12, progressbar=TRUE))
@@
 
-  form = ~ CellType + Gender
+  form = ~ CellType + Gender + Batch
   vobj = voomWithDreamWeights(dge, form, infoSub, plot=FALSE, save.plot=TRUE, BPPARAM=SnowParam(6, progressbar=TRUE))
@@
 
-    form = ~ CellType
+    form = ~ CellType + Batch
     vobj = voomWithDreamWeights(dge, form, infoSub2, plot=FALSE, save.plot=TRUE, BPPARAM=SnowParam(6, progressbar=TRUE), quiet=TRUE)
 
-    form = ~ CellType
+    form = ~ CellType + Batch
     fit = dream( vobj, form, infoSub2, BPPARAM=SnowParam(6, progressbar=TRUE), quiet=TRUE)
@@
 ```{r DE.disease}
+info$I = with(info, paste0(CellType, ".", Dx))
+
 # Disease
-fitDisease = dream( vobj, ~ 0 + CellType*Dx + (1|Individual.ID) + (1|Gender), info, BPPARAM=SnowParam(6, progressbar=TRUE))
+form = ~ 0 + I + (1|Individual.ID) + (1|Gender) + (1|Batch)
+
+# contrasts
+L = makeContrastsDream(form, info, contrasts=c(
+              GABA.Dx = "IGABA.SCZ - IGABA.Control",
+              GLU.Dx = "IGLU.SCZ - IGLU.Control",
+              MgAs.Dx = "IMgAs.SCZ - IMgAs.Control",
+              Olig.Dx = "IOlig.SCZ - IOlig.Control"))
+
+fitDisease = dream( vobj[1:10,], form, info, L=L, BPPARAM=SnowParam(12, progressbar=TRUE))
 ```
@@
 
-  key = paste0('CellType', CT, ':DxSCZ')
-
-  if( CT == levels(info$CellType)[1]){
-    key = 'DxSCZ'
-  }
+  key = paste0( CT, '.Dx')
 
@@
 
+```{r, eval=FALSE}
+knitr::knit_exit()
+```
+
 bMIND can incorporate the prior of cell type variance/covariance.
@@
 
-```{r store, eval=TRUE}
+```{r store, eval=FALSE}
 # Code

### 2021-07-30 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 rmarkdown::render("molec_profiling.Rmd", output_dir='./', intermediates_dir='./')
+
+# TODO 
+replace with FilterExpr()
 
@@
 
-    form = ~ CellType + Batch
+    form = ~ CellType #+ (1|Batch)
     vobj = voomWithDreamWeights(dge, form, infoSub2, plot=FALSE, save.plot=TRUE, BPPARAM=SnowParam(6, progressbar=TRUE), quiet=TRUE)
 
-    form = ~ CellType + Batch
+    form = ~ CellType #+ Batch
     fit = dream( vobj, form, infoSub2, BPPARAM=SnowParam(6, progressbar=TRUE), quiet=TRUE)

### 2021-08-02 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 replace with FilterExpr()
+add single cell
+Add CMC bulk RNA-seq
 
@@
 ```{r preprocess}
-isexpr = rowSums(edgeR::cpm(geneCounts)>1) >= 0.1*ncol(geneCounts)
+# isexpr = rowSums(edgeR::cpm(geneCounts)>1) >= 0.1*ncol(geneCounts)
+isexpr = filterByExpr(geneCounts, model.matrix(~CellType, info))
 table(isexpr)
@@
 ## Plots of top differentially expressed genes
-```{r DE.marker_genes, fig.width=15, fig.height=30}
+```{r DE.marker_genes, fig.width=15, fig.height=40}
 
@@
 
-  isexpr = rowSums(edgeR::cpm(geneCounts[,idx])>1) >= 0.3*ncol(geneCounts[,idx])
+  # isexpr = rowSums(edgeR::cpm(geneCounts[,idx])>1) >= 0.3*ncol(geneCounts[,idx])
+  isexpr = filterByExpr(geneCounts[,idx], model.matrix(~CellType, info[idx,]))
   table(isexpr)
@@
 # FANS4
-isexpr = rowSums(edgeR::cpm(geneCounts)>1) >= 0.3*ncol(geneCounts)
+# isexpr = rowSums(edgeR::cpm(geneCounts)>1) >= 0.3*ncol(geneCounts)
+isexpr = filterByExpr(geneCounts, model.matrix(~CellType, info))
 
@@
 
-isexpr = rowSums(edgeR::cpm(gtex_count[,idx])>1) >= 0.3*ncol(gtex_count[,idx])
+# isexpr = rowSums(edgeR::cpm(gtex_count[,idx])>1) >= 0.3*ncol(gtex_count[,idx])
+isexpr = filterByExpr(geneCounts[,idx], model.matrix(~CellType, info[idx,]))
 
@@
 
-  isexpr = rowSums(edgeR::cpm(geneCounts[,id])>1) >= 0.1*ncol(geneCounts[,id])
+  # isexpr = rowSums(edgeR::cpm(geneCounts[,id])>1) >= 0.1*ncol(geneCounts[,id])
+  isexpr = filterByExpr(geneCounts[,id], model.matrix(~CellType, info[id,]))
 
@@
 
-  isexpr = rowSums(edgeR::cpm(gtex_count[,id])>1) >= 0.1*ncol(gtex_count[,id])
+  # isexpr = rowSums(edgeR::cpm(gtex_count[,id])>1) >= 0.1*ncol(gtex_count[,id])
+  isexpr = filterByExpr(geneCounts[,id], model.matrix(~CellType, info[id,]))
 
@@
 
-isexpr = rowSums(edgeR::cpm(countsCombined)>1) >= 0.3*ncol(countsCombined)
+# isexpr = rowSums(edgeR::cpm(countsCombined)>1) >= 0.3*ncol(countsCombined)
+isexpr = filterByExpr(countsCombined, model.matrix(~CellType, countsCombined))
 
@@
 
-    isexpr = rowSums(edgeR::cpm(geneCountSub[,idx])>1) >= 0.3*ncol(geneCountSub[,idx])
+    # isexpr = rowSums(edgeR::cpm(geneCountSub[,idx])>1) >= 0.3*ncol(geneCountSub[,idx])
+    isexpr = filterByExpr(geneCountSub[,idx], model.matrix(~CellType, info[idx,]))
     table(isexpr)
@@
 
-fitDisease = dream( vobj[1:10,], form, info, L=L, BPPARAM=SnowParam(12, progressbar=TRUE))
+fitDisease = dream( vobj, form, info, L=L, BPPARAM=SnowParam(12, progressbar=TRUE))
 ```

### 2021-08-02 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 ```{r PCA.batch, fig.width=10, fig.height=5}
-fit = dream(geneExpr, ~ Batch, info)
+fit = dream(geneExpr, ~ (1|Batch), info)
 
@@
 ```{r predict.celltype}
-
 X = as.matrix(df_pca[,2:200])
@@
 # isexpr = rowSums(edgeR::cpm(gtex_count[,idx])>1) >= 0.3*ncol(gtex_count[,idx])
-isexpr = filterByExpr(geneCounts[,idx], model.matrix(~CellType, info[idx,]))
+isexpr = filterByExpr(gtex_count[,idx], model.matrix(~SMTSD, gtex_sample[idx,]))
 

### 2021-08-02 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 ```{r PCA.batch, fig.width=10, fig.height=5}
-fit = dream(geneExpr, ~ (1|Batch), info)
+fit = dream(geneExpr, ~ (1|Batch), info, BPPARAM=SnowParam(12, progressbar=TRUE))
 
@@
 # different because only genes expressed in given cell type are used
-res_cv_FANS = lapply(CTarray, function(CT){
+res_cv_FANS = lapply( unlist(CTarray), function(CT){
 

### 2021-08-02 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 ggplot(df_cv_FANS, aes(log2(sqrt(total)), color=key)) + geom_density(size=2) + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="right")  + scale_color_manual(values=colorCodes) + xlab(bquote(log[2]~CV))
+```
+
+```{r include.gtex}
 
@@
   # isexpr = rowSums(edgeR::cpm(gtex_count[,id])>1) >= 0.1*ncol(gtex_count[,id])
-  isexpr = filterByExpr(geneCounts[,id], model.matrix(~CellType, info[id,]))
+  isexpr = filterByExpr(gtex_count[,id], model.matrix(~SMTSD, gtex_sample[id,]))
 
@@
 ggplot( df_melt, aes(key, value, fill=key)) + geom_bar(stat='identity') + theme_bw(10) + theme(aspect.ratio=5, plot.title = element_text(hjust = 0.5), legend.position="none")  + scale_fill_manual(values=c(colsGTEx, colorCodes)[levels(df3$key)]) + coord_flip() + ylab("# genes expressed") + scale_y_continuous(expand=c(0,0)) + xlab('') + facet_wrap(~variable)
-
 ```

### 2021-08-02 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 ```{r include.gtex}
-
 #  GTEX
@@
   id = gtex_sample$SAMPID[which( gtex_sample$SMTSD == Tissue)]
+  rownames(gtex_sample) = gtex_sample$SAMPID
 
   # isexpr = rowSums(edgeR::cpm(gtex_count[,id])>1) >= 0.1*ncol(gtex_count[,id])
-  isexpr = filterByExpr(gtex_count[,id], model.matrix(~SMTSD, gtex_sample[id,]))
+  isexpr = filterByExpr( gtex_count[,id] )
 

### 2021-08-02 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 gtex_sample = gtex_sample[gtex_sample$SAMPID %in% colnames(rnaseq_sample_id),]
+rownames(gtex_sample) = gtex_sample$SAMPID
 
@@
 ```{r PCA.gtex}
-
 id = gtex_sample$SAMPID[which( gtex_sample$SMTSD == 'Brain - Frontal Cortex (BA9)')]
@@
 
+df_info = data.frame(CT = c(as.character(info$CellType), gtex_sample[id,]$SMTSD))
+
 # isexpr = rowSums(edgeR::cpm(countsCombined)>1) >= 0.3*ncol(countsCombined)
-isexpr = filterByExpr(countsCombined, model.matrix(~CellType, countsCombined))
+isexpr = filterByExpr(countsCombined, model.matrix(~CT, df_info))
 

### 2021-08-02 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 
-```{r Disease.analysis}
-df_disease = lapply(levels(info$CellType), function(CT){
+```{r Disease.analysis.corr}
+df_disease_de = lapply(levels(info$CellType), function(CT){
 
@@
   tabl = topTable( fitDisease, coef = key, number=Inf)
-
-  data.frame(CellType = CT, key, count = sum(tabl$adj.P.Val < 0.05), pi1 = 1 - qvalue(tabl$P.Value)$pi0)
-  })
-df_disease = do.call(rbind, df_disease)
-
-kable(df_disease, digits=2, row.names=FALSE) %>% kable_styling(bootstrap_options = "striped", full_width = FALSE)
+  tabl$CellType = CT
+  tabl$ensID = rownames(tabl)
+  tabl
+  })
+df_disease_de = data.table(do.call(rbind, df_disease_de))
+
+df_disease_de[,data.frame(count = sum(adj.P.Val < 0.05), pi1 = 1 - qvalue(P.Value)$pi0), by="CellType"] %>% kable(digits=2, row.names=FALSE) %>% kable_styling(bootstrap_options = "striped", full_width = FALSE)
+
+index = combn(nlevels(info$CellType),2)
+
+# for all pairs of cell types
+tabList = lapply( 1:ncol(index), function(j){
+
+  CT1 = levels(info$CellType)[index[1,j]]
+  CT2 = levels(info$CellType)[index[2,j]]
+
+  df_tstat = merge(df_disease_de[CellType==CT1,], df_disease_de[CellType==CT2,], by="ensID")
+
+  r = with(df_tstat, cor(z.std.x, z.std.y, method="spearman"))
+
+  data.frame(CT1, CT2, corr = r)
+})
+tabList = do.call(rbind, tabList)
+
+tabList %>% kable(digits=2, row.names=FALSE) %>% kable_styling(bootstrap_options = "striped", full_width = FALSE)
+
+ggplot(tabList, aes(CT1, CT2, fill=corr)) + geom_tile() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + scale_fill_gradient(name = quote(pi[1]), low="white", high="red", limits=c(0, 1)) + geom_text(aes(label=format(corr, digits=3))) + xlab('') + ylab('') + ggtitle("Correlation between t-statistics")
+
+C = matrix(1, 4,4)
+rownames(C) = levels(info$CellType)
+colnames(C) = levels(info$CellType)
+tabList$CT1 = factor(tabList$CT1, levels(info$CellType))
+tabList$CT2 = factor(tabList$CT2, levels(info$CellType))
+
+for(i in 1:nrow(tabList)){
+  C[tabList$CT1[i],tabList$CT2[i]] = tabList$corr[i]
+  C[tabList$CT2[i],tabList$CT1[i]] = tabList$corr[i]
+}
+
+plot(hclust(as.dist(1-C)))
 ```

### 2021-08-02 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 library(synapser)
+library(zenith)
 library(nnet)
@@
 
+# Gene set analysis
+```{r zenith.celltype, cache=FALSE}
+knitr::knit_exit()
+
+```
+
 ## Analysis using just pairs of samples
@@
 
+Gene-by-gene correlation
+fix PCA/MDS
+mvIC
+splicing
+add bulk, imputed, single cell
+zenith
+
 Volcano plot colored by cell type

### 2021-08-02 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 ```{r zenith.celltype, cache=FALSE}
+
+geneSets_GO = get_GeneOntology()
+
+# Map from Ensembl genes in geneSets_GO to 
+# Ensembl names from RNA-seq data 
+index = ids2indices( recodeToList(geneSets_GO), rownames(fitBasic))
+
+idxGO = index[sapply(index, length) > 20]
+
+df_CT_zenith = lapply( colnames(coef(fitBasic))[1:2], function(key){
+  res = zenith( fitBasic, coef = key, index = idxGO)
+  data.table(Geneset = rownames(res), res, coef = key)
+})
+df_CT_zenith = do.call(rbind, df_CT_zenith)
+
+df_CT_zenith[coef == 'GABA_vs_Olig',]
+
+```
+
+```{r exit, cache=FALSE}
 knitr::knit_exit()
-
 ```

### 2021-08-03 deconvolution/fans4/molec_profiling.Rmd
bug fix

--- 2021-08-02 deconvolution/fans4/molec_profiling.Rmd
@@
 git pull
-export OMP_NUM_THREADS=1
+# export OMP_NUM_THREADS=1
 R
@@
 ```{r zenith.celltype, cache=FALSE}
+
+geneSets_GO = get_GeneOntology("MF")
+
+# Map from Ensembl genes in geneSets_GO to 
+# Ensembl names from RNA-seq data 
+index = ids2indices( recodeToList(geneSets_GO), rownames(fitBasic))
+
+# filter by size of gene set
+n_genes_in = 30
+setSize = sapply(index, length)
+idxGO = index[(setSize > n_genes_in) & (setSize < 500)]
+
+df_CT_zenith = lapply( colnames(coef(fitBasic)), function(key){
+  res = zenith( fitBasic, coef = key, index = idxGO)
+  data.table(Geneset = rownames(res), res, coef = key)
+})
+df_CT_zenith = do.call(rbind, df_CT_zenith)
+
+df_gs = lapply( df_CT_zenith[,unique(coef)], function(key){
+
+  df_CT_zenith[(coef == key),][order(p.greater),][1:4,]
+  })
+df_gs = do.call(rbind, df_gs)
+```
+
+  res = zenith( fitBasic, coef = 'CellTypeGLU', index = idxGO)
+
+```{r eval=FALSE}
+# form = ~ CellType + Gender + Batch
+
+# idx = info$CellType %in% c("Olig", "GABA")
+# isexpr = filterByExpr(geneCounts[,idx], model.matrix(~CellType, info[idx,]))
+# dge = DGEList(geneCounts[isexpr,idx])
+# dge = calcNormFactors(dge)
+# vobj = voom(dge, model.matrix(~CellType, droplevels(info[idx,])))
+
+# fit = lmFit( vobj, model.matrix(form, droplevels(info[idx,])))
+# fit = eBayes(fit)
+# fit = dream(vobj, form, info[idx,])
+# fit = eBayes(fit)
+# res2 = zenith(fit, coef='CellTypeOlig',idxGO )
+
+# # Map from Ensembl genes in geneSets_GO to 
+# # Ensembl names from RNA-seq data 
+# index = ids2indices( recodeToList(geneSets_GO), rownames(fit))
+
+# n_genes_in = 20
+# idxGO = index[sapply(index, length) > n_genes_in]
+
+# res = camera( vobj, idxGO, model.matrix(form, droplevels(info[idx,])), inter.gene.cor=NA, contrast='CellTypeOlig')
+
+```
+
+```{r exit, cache=FALSE}
 knitr::knit_exit()
-
 ```

### 2021-08-03 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 git pull
-# export OMP_NUM_THREADS=1
+export OMP_NUM_THREADS=2
 R
@@
   Olig = '#E7298A',
-  MgAs = '#7570B3')
+  MgAs = '#7570B3',
+  GABAGLU_vs_Other = '#A6A910',
+  GABA_vs_GLU = "grey")
 
@@
 contrasts['GABAGLU_vs_Other'] = "(CellTypeGABA + CellTypeGLU)/2 - (CellTypeMgAs + CellTypeOlig)/2"
+contrasts['GLU_vs_Other'] = "CellTypeGLU - (CellTypeGABA + CellTypeMgAs + CellTypeOlig)/3"
+contrasts['GABA_vs_Other'] = "CellTypeGABA - (CellTypeGLU + CellTypeMgAs + CellTypeOlig)/3"
 contrasts['MgAs_vs_Other'] = "CellTypeMgAs - (CellTypeGABA + CellTypeGLU + CellTypeOlig)/3"
@@
 # Gene set analysis
-```{r zenith.celltype, cache=FALSE}
-
-geneSets_GO = get_GeneOntology("MF")
-
+```{r zenith.celltype, cache=FALSE, eval=FALSE}
+
+geneSets_GO = get_GeneOntology()
+
+filter_gs = function(fit, gs){
   # Map from Ensembl genes in geneSets_GO to 
   # Ensembl names from RNA-seq data 
-index = ids2indices( recodeToList(geneSets_GO), rownames(fitBasic))
+  index = ids2indices( recodeToList(gs), rownames(fit))
 
   # filter by size of gene set
-n_genes_in = 30
+  n_genes_in = 15
   setSize = sapply(index, length)
   idxGO = index[(setSize > n_genes_in) & (setSize < 500)]
+  idxGO
+}
+
+fit = fitBasic[fitBasic$Amean > -Inf,]
+idxGO = filter_gs( fit, geneSets_GO)
 
 df_CT_zenith = lapply( colnames(coef(fitBasic)), function(key){
-  res = zenith( fitBasic, coef = key, index = idxGO)
+  res = zenith( fit, coef = key, index = idxGO, inter.gene.cor=0.01)
   data.table(Geneset = rownames(res), res, coef = key)
@@
 
-  df_CT_zenith[(coef == key),][order(p.greater),][1:4,]
+  # results for this coef
+  df_sub = df_CT_zenith[(coef == key),][order(p.greater),]
+
+  # remove gene sets with small p-values from other tests
+  df_sub2 = df_CT_zenith[(coef != key),][order(p.greater),]
+  df_sub2 = df_sub2[,.SD[which.min(p.greater),], by="Geneset"]
+  df_sub2 = df_sub2[p.greater > 1e-3,]
+
+  df_sub[!(df_sub$Geneset %in% df_sub2$Geneset),][1:4,]
   })
 df_gs = do.call(rbind, df_gs)
-```
-
-  res = zenith( fitBasic, coef = 'CellTypeGLU', index = idxGO)
+
+df_gs %>% kable() %>% kable_styling(bootstrap_options = "striped", full_width = FALSE)
+
+df = df_gs[grep("CellType", coef),]
+# df = rbind(df, df_gs[grep('GABAGLU_vs_Other|GABA_vs_GLU', coef)])
+df[,coef := gsub("CellType", '',coef)]
+df$coef = factor(df$coef, names(colorCodes))
+
+ggplot(df, aes(Geneset,coef, -log10(p.greater))) + geom_tile() + coord_flip() + scale_y_continuous(expand=c(0,0)) + theme_classic() + ylab(bquote(-log[10]~p)) + scale_fill_manual(values = c(colorCodes, GABAGLU_vs_Other = "green"), name="Cell type")
+```
+
+```{r compare.celltypes}
+zenith_gsa = function (fit, coefs, geneSets, n_genes_min = 10, n_genes_max=1000,...){
+
+    # from GeneSetCollection to list 
+    geneSets = recodeToList(geneSets)
+
+    index = ids2indices(geneSets, rownames(fit))
+    index = index[sapply(index, length) >= n_genes_min]
+    index = index[sapply(index, length) < n_genes_max]
+
+    df_res = lapply(coefs, function(coef) {
+        df_res = zenith(fit, coef, index,...)
+        data.table(coef = coef, Geneset = rownames(df_res), df_res)
+    })
+    df_zenith = do.call(rbind, df_res)
+    rownames(df_zenith) = c()
+
+    df_zenith$FDR = p.adjust(df_zenith$PValue, "BH")
+    df_zenith
+}
+
+geneSets_GO = get_GeneOntology("BP")
+
+coefTests = colnames(coef(fitBasic))[c(1, 7:9)]
+df_zenith = zenith_gsa(fitBasic[fitBasic$Amean > 0,], coefTests, geneSets_GO, inter.gene.cor=0.01)
+
+df_zenith$FDR = p.adjust(df_zenith$p.greater, "BH")
+
+df = df_zenith[,.SD[order(p.greater)[1:3],],by=c("coef")]
+cls = unique(df$Geneset)
+df = df_zenith[Geneset %in% df$Geneset,]
+
+nrow = length(unique(df$coef))
+ncol = length(unique(df$Geneset))
+
+M = matrix(-log10(df$p.greater), nrow=nrow, ncol=ncol,dimnames=list(unique(df$coef), unique(df$Geneset)), byrow=TRUE)
+
+hcl_col = hclust(dist(M), method="ward.D")
+df$coef = factor(df$coef, rownames(M)[hcl_col$order])
+
+hcl_row = hclust(dist(t(M)), method="ward.D")
+# df$Geneset = factor(df$Geneset, colnames(M)[hcl_row$order])
+df$Geneset = factor(df$Geneset, cls)
+df$label = ifelse(df$FDR < 0.05, '*', '')
+
+ggplot(df, aes(coef, Geneset, fill=-log10(p.greater), label=label)) + geom_tile() + theme_classic() + scale_fill_gradient(limits=c(0, NA), low="white", high="red", name=bquote(-log[10]~P)) + theme(aspect.ratio = ncol/nrow, axis.text.x = element_text(angle = 90, vjust = 1, hjust=1)) + xlab('') + ylab('') + geom_text(vjust=1, hjust=0.5)
+```
 

### 2021-08-03 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
   Olig = '#E7298A',
-  MgAs = '#7570B3',
-  GABAGLU_vs_Other = '#A6A910',
-  GABA_vs_GLU = "grey")
+  MgAs = '#7570B3')
+  # GABAGLU_vs_Other = '#A6A910',
+  # GABA_vs_GLU = "grey")
 

### 2021-08-04 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 ```{r DE.analysis}
-
 index = combn(nlevels(info$CellType),2)
@@
 ```{r DE2}
-
 index = combn(nlevels(info$CellType),2)
@@
 ```{r DE.marker_genes, fig.width=15, fig.height=40}
-
 index = combn(nlevels(info$CellType),2)
@@
 
-```{r compare.celltypes}
+```{r compare.celltypes, fig.width=8, fig.height=12}
 zenith_gsa = function (fit, coefs, geneSets, n_genes_min = 10, n_genes_max=1000,...){
@@
 
-coefTests = colnames(coef(fitBasic))[c(1, 7:9)]
-df_zenith = zenith_gsa(fitBasic[fitBasic$Amean > 0,], coefTests, geneSets_GO, inter.gene.cor=0.01)
+coefTests = c('GABA_vs_GLU', 'GABAGLU_vs_Other', 'GLU_vs_Other', 'GABA_vs_Other', 'MgAs_vs_Other', 'Olig_vs_Other')
+df_zenith = zenith_gsa(fitBasic[fitBasic$Amean > 0,], coefTests, geneSets_GO)
+#, inter.gene.cor=0.01)
 
@@
 
-M = matrix(-log10(df$p.greater), nrow=nrow, ncol=ncol,dimnames=list(unique(df$coef), unique(df$Geneset)), byrow=TRUE)
-
-hcl_col = hclust(dist(M), method="ward.D")
-df$coef = factor(df$coef, rownames(M)[hcl_col$order])
-
-hcl_row = hclust(dist(t(M)), method="ward.D")
+# M = matrix(-log10(df$p.greater), nrow=nrow, ncol=ncol,dimnames=list(unique(df$coef), unique(df$Geneset)), byrow=TRUE)
+
+# M = matrix(-log10(df$p.greater), nrow=nrow, ncol=ncol, byrow=TRUE)
+# rownames(M) = unique(df$coef)
+# colnames(M) = unique(df$Geneset)
+
+# Heatmap(t(M), col = colorRampPalette(c("white", "red"))(100))
+
+# hcl_col = hclust(dist(M), method="ward.D2")
+# df$coef = factor(df$coef, rownames(M)[hcl_col$order])
+
+# hcl_row = hclust(dist(t(M)), method="ward.D2")
 # df$Geneset = factor(df$Geneset, colnames(M)[hcl_row$order])

### 2021-08-04 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 
-ggplot(df, aes(coef, Geneset, fill=-log10(p.greater), label=label)) + geom_tile() + theme_classic() + scale_fill_gradient(limits=c(0, NA), low="white", high="red", name=bquote(-log[10]~P)) + theme(aspect.ratio = ncol/nrow, axis.text.x = element_text(angle = 90, vjust = 1, hjust=1)) + xlab('') + ylab('') + geom_text(vjust=1, hjust=0.5)
+ggplot(df, aes(coef, Geneset, fill=-log10(p.greater), label=label)) + geom_tile() + theme_classic() + scale_fill_gradient(limits=c(0, NA), low="white", high="red", name=bquote(-log[10]~P)) + theme(aspect.ratio = ncol/nrow, axis.text.x = element_text(angle = 90, vjust = 1, hjust=1), plot.title = element_text(hjust = 0.5)) + xlab('') + ylab('') + geom_text(vjust=1, hjust=0.5) + ggtitle("Compare cell types")
 ```

### 2021-08-04 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 # Gene set analysis
-```{r zenith.celltype, cache=FALSE, eval=FALSE}
-
-geneSets_GO = get_GeneOntology()
-
-filter_gs = function(fit, gs){
-  # Map from Ensembl genes in geneSets_GO to 
-  # Ensembl names from RNA-seq data 
-  index = ids2indices( recodeToList(gs), rownames(fit))
-
-  # filter by size of gene set
-  n_genes_in = 15
-  setSize = sapply(index, length)
-  idxGO = index[(setSize > n_genes_in) & (setSize < 500)]
-  idxGO
-}
-
-fit = fitBasic[fitBasic$Amean > -Inf,]
-idxGO = filter_gs( fit, geneSets_GO)
-
-df_CT_zenith = lapply( colnames(coef(fitBasic)), function(key){
-  res = zenith( fit, coef = key, index = idxGO, inter.gene.cor=0.01)
-  data.table(Geneset = rownames(res), res, coef = key)
-})
-df_CT_zenith = do.call(rbind, df_CT_zenith)
-
-df_gs = lapply( df_CT_zenith[,unique(coef)], function(key){
-
-  # results for this coef
-  df_sub = df_CT_zenith[(coef == key),][order(p.greater),]
-
-  # remove gene sets with small p-values from other tests
-  df_sub2 = df_CT_zenith[(coef != key),][order(p.greater),]
-  df_sub2 = df_sub2[,.SD[which.min(p.greater),], by="Geneset"]
-  df_sub2 = df_sub2[p.greater > 1e-3,]
-
-  df_sub[!(df_sub$Geneset %in% df_sub2$Geneset),][1:4,]
-  })
-df_gs = do.call(rbind, df_gs)
-
-df_gs %>% kable() %>% kable_styling(bootstrap_options = "striped", full_width = FALSE)
-
-df = df_gs[grep("CellType", coef),]
-# df = rbind(df, df_gs[grep('GABAGLU_vs_Other|GABA_vs_GLU', coef)])
-df[,coef := gsub("CellType", '',coef)]
-df$coef = factor(df$coef, names(colorCodes))
-
-ggplot(df, aes(Geneset,coef, -log10(p.greater))) + geom_tile() + coord_flip() + scale_y_continuous(expand=c(0,0)) + theme_classic() + ylab(bquote(-log[10]~p)) + scale_fill_manual(values = c(colorCodes, GABAGLU_vs_Other = "green"), name="Cell type")
-```
-
 ```{r compare.celltypes, fig.width=8, fig.height=12}
@@
 coefTests = c('GABA_vs_GLU', 'GABAGLU_vs_Other', 'GLU_vs_Other', 'GABA_vs_Other', 'MgAs_vs_Other', 'Olig_vs_Other')
-df_zenith = zenith_gsa(fitBasic[fitBasic$Amean > 0,], coefTests, geneSets_GO)
-#, inter.gene.cor=0.01)
+df_zenith = zenith_gsa(fitBasic[fitBasic$Amean > 0,], coefTests, geneSets_GO, inter.gene.cor=0.01)
 
@@
 ggplot(df, aes(coef, Geneset, fill=-log10(p.greater), label=label)) + geom_tile() + theme_classic() + scale_fill_gradient(limits=c(0, NA), low="white", high="red", name=bquote(-log[10]~P)) + theme(aspect.ratio = ncol/nrow, axis.text.x = element_text(angle = 90, vjust = 1, hjust=1), plot.title = element_text(hjust = 0.5)) + xlab('') + ylab('') + geom_text(vjust=1, hjust=0.5) + ggtitle("Compare cell types")
-```
-
-```{r eval=FALSE}
-# form = ~ CellType + Gender + Batch
-
-# idx = info$CellType %in% c("Olig", "GABA")
-# isexpr = filterByExpr(geneCounts[,idx], model.matrix(~CellType, info[idx,]))
-# dge = DGEList(geneCounts[isexpr,idx])
-# dge = calcNormFactors(dge)
-# vobj = voom(dge, model.matrix(~CellType, droplevels(info[idx,])))
-
-# fit = lmFit( vobj, model.matrix(form, droplevels(info[idx,])))
-# fit = eBayes(fit)
-# fit = dream(vobj, form, info[idx,])
-# fit = eBayes(fit)
-# res2 = zenith(fit, coef='CellTypeOlig',idxGO )
-
-# # Map from Ensembl genes in geneSets_GO to 
-# # Ensembl names from RNA-seq data 
-# index = ids2indices( recodeToList(geneSets_GO), rownames(fit))
-
-# n_genes_in = 20
-# idxGO = index[sapply(index, length) > n_genes_in]
-
-# res = camera( vobj, idxGO, model.matrix(form, droplevels(info[idx,])), inter.gene.cor=NA, contrast='CellTypeOlig')
-
-```
-
-```{r exit, cache=FALSE}
-knitr::knit_exit()
 ```
@@
 boxplot_gene(ensIDs, tab) + ggtitle("FANS markers")
+```
+
+```{r exit, cache=FALSE}
+knitr::knit_exit()
 ```

### 2021-08-04 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 
+#### Correlation between individuals
+```{r corr.btw.indivs}
+form = ~ (1|CellType) + (1|Batch)
+vobj.simple = voomWithDreamWeights(dge, form, info, BPPARAM=SnowParam(12, progressbar=TRUE))
+
+fit.simple = dream(vobj.simple, form, info, BPPARAM=SnowParam(12, progressbar=TRUE))
+residCompare = residuals(fit.simple) 
+```
+
+```{r exit, cache=FALSE}
+knitr::knit_exit()
+```
+
 ```{r predict.celltype}
@@
 ## Plots of top differentially expressed genes
-```{r markerGenes, fig.width=15, fig.height=30}
+```{r markerGenes, fig.width=15, fig.height=40}
 
@@
 boxplot_gene(ensIDs, tab) + ggtitle("FANS markers")
-```
-
-```{r exit, cache=FALSE}
-knitr::knit_exit()
 ```
@@
 
+```{r exit, cache=FALSE}
+knitr::knit_exit()
+```
+
 # Disease

### 2021-08-04 deconvolution/fans4/molec_profiling.Rmd
bug fix

--- 2021-08-04 deconvolution/fans4/molec_profiling.Rmd
@@
 
+#### Correlation between individuals
+```{r corr.btw.indivs}
+form = ~ (1|CellType) + (1|Batch) + (1|Individual.ID) 
+vobj.simple = voomWithDreamWeights(dge, form, info, BPPARAM=SnowParam(12, progressbar=TRUE))
+
+# fit.simple = dream(vobj.simple, form, info, BPPARAM=SnowParam(12, progressbar=TRUE))
+# residCompare = residuals(fit.simple)
+
+residCompare = fitVarPartModel(vobj.simple, form, info, fxn=function(fit){
+  library(lme4)  
+  fit@resp[['y']] - variancePartition::get_prediction(fit, ~ (1|CellType) + (1|Batch))
+}, BPPARAM=SnowParam(12, progressbar=TRUE))
+
+residCompare = do.call(rbind, residCompare)
+```
+
+```{r plot_within_btw}
+plot_within_btw = function( simMat, design, col=c("lightblue", "lightgreen"),main='',ylim=NULL,...){
+  variation = list(type=c(), value=c(), donor1=c(), donor2=c())
+  for(i in 1:ncol(design)){
+    idx1 = which(design[,i] == 1)   
+    idx0 = which(design[,i] == 0)
+    C_sub = simMat[idx1, idx1]
+    variation$type = append( variation$type, rep("Same Donor", sum(lower.tri(C_sub))) )  
+    variation$value = append( variation$value, C_sub[lower.tri(C_sub)])   
+    variation$donor1 = append(variation$donor1, rep(colnames(design)[i], length(C_sub[lower.tri(C_sub)]) ))      
+    variation$donor2 = append(variation$donor2, rep(colnames(design)[i], length(C_sub[lower.tri(C_sub)]) ))
+    variation$type = append( variation$type, rep("Different Donor", length(simMat[idx1,idx0])) )  
+    variation$value = append( variation$value, simMat[idx1,idx0])    
+    variation$donor1 = append(variation$donor1, rep("", length(simMat[idx1,idx0])) )      
+    variation$donor2 = append(variation$donor2, rep("", length(simMat[idx1,idx0])))
+  }
+  variation = as.data.frame(variation)
+  fig = ggplot( variation, aes(type, value)) + geom_violin(scale = "width", aes(fill=type)) + 
+    ylab("Variance explained (%)") + xlab("") + geom_boxplot(width = 0.07, 
+        fill = "grey", outlier.color = "black") + theme_bw(15) + 
+        scale_fill_manual(values = col) + theme(legend.position = "none") + 
+        ylab("Correlation between samples") + scale_fill_manual(values=c("grey", "steelblue1"))
+    if( main !='' ){
+      fig = fig + ggtitle( main )
+    }
+    if( !is.null(ylim) ){
+      fig = fig + ylim( ylim )
+    }
+    fig = fig + theme(aspect.ratio=1)
+    return(list(fig=fig, variation=variation))
+}
+
+C = cor( residCompare, method="spearman" )
+
+res = plot_within_btw( C, model.matrix(~Individual.ID+0, info)) 
+res$fig
+```
+
 ```{r predict.celltype}

### 2021-08-04 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 
-infoExpand = rbind(data.frame(info[,c(1,3)], dataset="FANS4", stringsAsFactors=FALSE), 
-  data.frame(Name=colnames(gtex_count[,id]), CellType = "GTEx BA9", dataset="GTEx", stringsAsFactors=FALSE))
+infoExpand = rbind(data.frame(info[,c("Name", "CellType", "Batch")], 
+  dataset="FANS4", stringsAsFactors=FALSE), 
+  data.frame(Name=colnames(gtex_count[,id]), CellType = "GTEx BA9", dataset="GTEx", Batch = "GTEx", stringsAsFactors=FALSE))
 rownames(infoExpand) = infoExpand$Name
@@
 # the first PC separates FANS from bulk, so subtract that and use residuals
-fit = dream(geneExprCombined, ~ dataset, infoExpand)
+fit = dream(geneExprCombined, ~ Batch, infoExpand)
 resid = residuals(fit, geneExprCombined)
@@
 ylab = paste0('PC2 (', format(varFrac[2]*100, digits=3), '%)')
-ggplot(df_pca, aes(PC1, PC2, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=c(colorCodes, "grey")) + xlab(xlab) + ylab(ylab)
-
-# xlab = paste0('PC2 (', format(varFrac[2]*100, digits=3), '%)')
-# ylab = paste0('PC3 (', format(varFrac[3]*100, digits=3), '%)')
-# ggplot(df_pca, aes(PC2, PC3, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=c(colorCodes, "grey")) + xlab(xlab) + ylab(ylab)
-
+ggplot(df_pca, aes(PC1, PC2, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=c(colorCodes, 'GTEx BA9' = "grey")) + xlab(xlab) + ylab(ylab)
+
+xlab = paste0('PC2 (', format(varFrac[2]*100, digits=3), '%)')
+ylab = paste0('PC3 (', format(varFrac[3]*100, digits=3), '%)')
+ggplot(df_pca, aes(PC2, PC3, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=c(colorCodes, 'GTEx BA9' = "grey")) + xlab(xlab) + ylab(ylab)
 ```
@@
 ```{r GTEX.MDS}
-
-d = as.dist(1 - cor(geneExprCombined)^2)
+d = as.dist(1 - cor(resid, method="spearman")^2)
 # d = as.dist(1 - cor(apply(geneExprCombined, 2, RNOmni::rankNorm), method="sp")^2)
 
-df_mds = data.frame(cmdscale( d, k=3))
+df_mds = data.frame(cmdscale( d, k=2))
 
@@
 
-ggplot(df_mds, aes(X1, X2, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=c(colorCodes, "grey")) + xlab('Coordinate 1') + ylab('Coordinate 2')
-
-ggplot(df_mds, aes(X2, X3, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=c(colorCodes, "grey")) + xlab('Coordinate 2') + ylab('Coordinate 3')
+ggplot(df_mds, aes(X1, X2, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=c(colorCodes, 'GTEx BA9' = "grey")) + xlab('Coordinate 1') + ylab('Coordinate 2')
+
+# ggplot(df_mds, aes(X2, X3, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=c(colorCodes, , 'GTEx BA9' = "grey")) + xlab('Coordinate 2') + ylab('Coordinate 3')
 ```
@@
 
-```{r exit, cache=FALSE}
-knitr::knit_exit()
-```
-
 # Disease

### 2021-08-04 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 
-ggplot(tabList, aes(CT1, CT2, fill=corr)) + geom_tile() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + scale_fill_gradient(name = quote(pi[1]), low="white", high="red", limits=c(0, 1)) + geom_text(aes(label=format(corr, digits=3))) + xlab('') + ylab('') + ggtitle("Correlation between t-statistics")
-
-C = matrix(1, 4,4)
-rownames(C) = levels(info$CellType)
-colnames(C) = levels(info$CellType)
-tabList$CT1 = factor(tabList$CT1, levels(info$CellType))
-tabList$CT2 = factor(tabList$CT2, levels(info$CellType))
-
-for(i in 1:nrow(tabList)){
-  C[tabList$CT1[i],tabList$CT2[i]] = tabList$corr[i]
-  C[tabList$CT2[i],tabList$CT1[i]] = tabList$corr[i]
+tabList$CT1 = factor(tabList$CT1, names(colorCodes)[1:4])
+tabList$CT2 = factor(tabList$CT2, names(colorCodes)[1:4])
+
+ggplot(tabList, aes(CT1, CT2, fill=corr)) + geom_tile() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + scale_fill_gradient(name = "Correlation", low="white", high="red", limits=c(0, 1)) + geom_text(aes(label=format(corr, digits=3))) + xlab('') + ylab('') + ggtitle("Correlation between t-statistics")
+
+# C = matrix(1, 4,4)
+# rownames(C) = levels(info$CellType)
+# colnames(C) = levels(info$CellType)
+# tabList$CT1 = factor(tabList$CT1, levels(info$CellType))
+# tabList$CT2 = factor(tabList$CT2, levels(info$CellType))
+
+# for(i in 1:nrow(tabList)){
+#   C[tabList$CT1[i],tabList$CT2[i]] = tabList$corr[i]
+#   C[tabList$CT2[i],tabList$CT1[i]] = tabList$corr[i]
+# }
+
+# plot(hclust(as.dist(1-C)))
+```
+
+```{r scatterPlots, dev="png"}
+library(viridis)
+
+# Compute two dimensional density
+get_density <- function(x, y, n = 250) {
+  dens <- MASS::kde2d(x = x, y = y, n = n)
+  ix <- findInterval(x, dens$x)
+  iy <- findInterval(y, dens$y)
+  ii <- cbind(ix, iy)
+  return(dens$z[ii])
 }
 
-plot(hclust(as.dist(1-C)))
+# make scatter plot with 2D density using viridis colors
+plotScatterDensity = function(value1, value2){
+
+  # convert two vectors to a data frame
+  df = data.frame(cbind(value1, value2))
+
+  # determine limits of the plot
+  lim = with(df, max(abs(c(value1, value2))))
+
+  # Compute 2D density
+  df$density <- get_density(df$value1, df$value2, n = 100)
+
+  # Scatter plot colored by density
+  ggplot(df, aes(value1, value2, color=density)) + geom_point(size=.4) + theme_bw(16) + theme(aspect.ratio=1, legend.position="bottom", plot.title = element_text(hjust = 0.5)) + geom_abline(color="red") + geom_vline(xintercept=0, col="grey40", linetype="dashed") + geom_hline(yintercept=0, col="grey40", linetype="dashed") + xlim(-lim, lim) + ylim(-lim, lim) + scale_color_viridis() + geom_smooth(method="lm", se=FALSE, color="darkorange")
+}
+
+index = combn(nlevels(info$CellType),2)
+
+# for all pairs of cell types
+figList = lapply( 1:ncol(index), function(j){
+
+  CT1 = levels(info$CellType)[index[1,j]]
+  CT2 = levels(info$CellType)[index[2,j]]
+
+  df_tstat = merge(df_disease_de[CellType==CT1,], df_disease_de[CellType==CT2,], by="ensID")
+
+  fig = with(df_tstat, plotScatterDensity(t.x, t.y))
+  fig + theme(legend.position="none") + xlab(CT1) + ylab(CT2)
+})
+
+plot_grid(plotlist = figList, ncol=3, nrow=3)
+```
+
+```{r mashr}
+library(mashr)
+
+```
+
+```{r zenith.disease}
+plotZenith = function(df){
+
+  nrow = length(unique(df$coef))
+  ncol = length(unique(df$Geneset))
+
+  df$Geneset = factor(df$Geneset, cls)
+  df$label = ifelse(df$FDR < 0.05, '*', '')
+
+  M = reshape2::dcast(df, coef ~ Geneset, value.var="delta")
+  annot = M[,1:2]
+  M = as.matrix(M[,-1])
+  rownames(M) = annot$coef
+
+  hcl2 <- hclust(dist(t(M)))
+  df$Geneset = factor(df$Geneset, colnames(M)[hcl2$order])
+
+  statMax = with(df, max(abs(delta/se)))
+
+  ggplot(df, aes(coef, Geneset, fill=delta/se, label=label)) + geom_tile() + theme_classic() + scale_fill_gradient2(limits=c(-statMax, statMax), low="blue", mid="white", high="red", name="t-statistic") + theme(aspect.ratio = ncol/nrow, axis.text.x = element_text(angle = 90, vjust = 1, hjust=1), plot.title = element_text(hjust = 0.5)) + xlab('') + ylab('') + geom_text(vjust=1, hjust=0.5) + ggtitle("Compare cell types")
+}
+
+geneSets_GO = get_GeneOntology()
+
+df_zenith = zenith_gsa(fitDisease[fitDisease$Amean > 0,], colnames(coef(fitDisease))[1:4], geneSets_GO, inter.gene.cor=0.01)
+
+df = df_zenith[,.SD[order(PValue)[1:10],],by=c("coef")]
+cls = unique(df$Geneset)
+df = df_zenith[Geneset %in% df$Geneset,]
+
+pdf("test.pdf")
+plotZenith(df)
+dev.off()
+
+```
+
+```{r zenith.disease.mSigDB}
+geneSets_MSigDB = get_MSigDB(c('H', "C1", "C2", "C3", "C5", "C7"))
+
+df_zenith = zenith_gsa(fitDisease[fitDisease$Amean > 0,], colnames(coef(fitDisease))[1:4], geneSets_MSigDB, inter.gene.cor=0.01)
+
+df = df_zenith[,.SD[order(PValue)[1:10],],by=c("coef")]
+cls = unique(df$Geneset)
+df = df_zenith[Geneset %in% df$Geneset,]
+
+plotZenith(df)
 ```
@@
 zenith
+
+# Integrate CommonMind Consortium
+```{r load.cmc}
+
+geneCounts_CMC = readRDS(synGet('syn22416410')$path)
+info_CMC = readRDS(synGet('syn22416411')$path)
+
+```
 

### 2021-08-04 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 
-pdf("test.pdf")
 plotZenith(df)
-dev.off()
-
-```
-
-```{r zenith.disease.mSigDB}
+```
+
+```{r zenith.disease.mSigDB, fig.width=12}
 geneSets_MSigDB = get_MSigDB(c('H', "C1", "C2", "C3", "C5", "C7"))
@@
 ```{r load.cmc}
-
 geneCounts_CMC = readRDS(synGet('syn22416410')$path)
 info_CMC = readRDS(synGet('syn22416411')$path)
-
 ```

### 2021-08-04 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 library(zenith)
+library(UpSetR)
 library(nnet)
@@
 
+```{r disease.upsetr}
+de_genes = lapply(levels(info$CellType), function(CT){
+
+  key = paste0( CT, '.Dx')
+
+  tabl = topTable( fitDisease, coef = key, number=Inf)
+  tabl$ensID = rownames(tabl)
+  rownames(tabl[tabl$adj.P.Val < 0.05,])
+  })
+names(de_genes) = levels(info$CellType)
+
+lst = fromList(de_genes)
+a = upset(lst, keep.order = FALSE, order.by="freq" )
+
+upset(lst, keep.order = FALSE, order.by="freq", sets.bar.color=colorCodes[a$Set_names] )
+```
+
 ```{r scatterPlots, dev="png"}

### 2021-08-04 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 
+df_disease_de = lapply(levels(info$CellType), function(CT){
+
+  key = paste0( CT, '.Dx')
+
+  tab = topTable( fitDisease, coef = key, number=Inf)
+  tab$ensID = rownames(tab)
+  tab$CellType = CT
+  tab$se = with(tab, logFC / t)
+  tab
+  })
+names(df_disease_de) = levels(info$CellType)
+df_disease_de = do.call(rbind, df_disease_de)
+
+B.hat = reshape2::dcast( df_disease_de, ensID ~ CellType, value.var='logFC')
+rownames(B.hat) = B.hat$ensID
+B.hat = as.matrix(B.hat[,-1])
+
+se.hat = reshape2::dcast( df_disease_de, ensID ~ CellType, value.var='se')
+rownames(se.hat) = se.hat$ensID
+se.hat = as.matrix(se.hat[,-1])
+
+mashObj = mash_set_data( B.hat, se.hat)
+
+m.1by1 = mash_1by1(mashObj)
+strong = get_significant_results(m.1by1, 0.05)
+
+U.pca = cov_pca(mashObj,3,subset=strong)
+print(names(U.pca))
+
+U.ed = cov_ed(mashObj, U.pca, subset=strong)
+
+U.c = cov_canonical(mashObj)  
+resMash = mash(mashObj, c(U.c,U.ed))
+
+plotCorrMatrix(cor(resMash$result$PosteriorMean))
+```
+
+```{r upset.mashr}
+lst = fromList(apply(resMash$result$lfsr, 2, function(x) names(x[x < 0.05])))
+
+a = upset(lst, keep.order = FALSE, order.by="freq" )
+
+upset(lst, keep.order = FALSE, order.by="freq", sets.bar.color=colorCodes[a$Set_names] )
 ```

### 2021-08-04 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 
+Correlation between cell types
+```{r corr.celltype}
+C = get_pairwise_sharing( resMash )
+
+corrplot( C, addCoef.col = "black", title = "Pairwise sharing",
+  col=colorRampPalette(c("blue","white","red"))(200),  cl.lim=c(-1,1) )
+```
+
+```{r r heterogeneity}
+library(metafor)
+library(qqman)
+
+df_het = mclapply( 1:nrow(B.hat), function(j){
+
+  fit = rma( B.hat[j,], sei = se.hat[j,])
+
+  data.frame( EnsID = rownames(B.hat)[j], 
+    Isq = fit$I2, 
+    Hsq = fit$H2, 
+    p.het = fit$QEp)
+  },mc.cores = 12)
+df_het = do.call(rbind, df_het)
+
+# resMash$result$PosteriorMean[j,]
+
+par(mfrow=c(1,2))
+hist(df_het$p.het)
+qq(df_het$p.het)
+
+nDE = sum(p.adjust(df_het$p.het, "fdr") < 0.05)
+pi1 = 1 - qvalue(df_het$p.het)$pi0
+
+# Gene set analysis of heterogeneity
+stat = qchisq(df_het$p.het, 1, lower.tail=FALSE)
+names(stat) = df_het$EnsID
+
+geneSets_GO = get_GeneOntology()
+geneSets = recodeToList(geneSets_GO)
+
+index = ids2indices(geneSets, names(stat))
+index = index[sapply(index, length) >= 15]
+index = index[sapply(index, length) < 500]
+
+res_het = cameraPR( stat, index, use.ranks=TRUE)
+
+res_het[res_het$FDR < 0.05] %>% kable() %>% kable_styling(bootstrap_options = "striped", full_width = FALSE)
+```
+
+Number of DE genes: `r nDE`. pi1 = `r pi1`
+
 ```{r zenith.disease}

### 2021-08-04 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
 library(synapser)
+library(metafor)
+library(qqman)
 library(zenith)
@@
 
-```{r r heterogeneity}
-library(metafor)
-library(qqman)
-
+```{r heterogeneity}
 df_het = mclapply( 1:nrow(B.hat), function(j){

### 2021-08-04 deconvolution/fans4/molec_profiling.Rmd
bug fix

@@
   },mc.cores = 12)
-df_het = do.call(rbind, df_het)
+df_het = data.table(do.call(rbind, df_het))
+
+df_hugo = AnnotationDbi::select(org.Hs.eg.db, df_het$EnsID, "SYMBOL", "ENSEMBL")
+df_het = merge(df_het, df_hugo, by.x="EnsID", by.y = "ENSEMBL")
+df_het$SYMBOL[is.na(df_het$SYMBOL)] = df_het$EnsID[is.na(df_het$SYMBOL)]
 
@@
 
-res_het[res_het$FDR < 0.05] %>% kable() %>% kable_styling(bootstrap_options = "striped", full_width = FALSE)
+res_het[res_het$FDR < 0.05,] %>% kable() %>% kable_styling(bootstrap_options = "striped", full_width = FALSE)
 ```
@@
 Number of DE genes: `r nDE`. pi1 = `r pi1`
+
+```{r plot.hetero, fig.height=30}
+
+plotEffectSize = function(CellType, beta, se, main=''){
+
+  df = data.frame(CellType, beta, se)
+
+  ggplot(df, aes(CellType, beta, ymin = beta - 1.96*se, ymax = beta + 1.96*se, color=CellType)) + geom_point() + geom_errorbar(width=.1) + theme_classic() + theme(aspect.ratio=1, legend.position="none", plot.title = element_text(hjust = 0.5)) + coord_flip() + ylab(bquote(log[2]~FC)) + scale_color_manual(values = colorCodes[df$CellType]) + geom_hline(yintercept=0, linetype="dashed") + ggtitle(main)
+}
+
+figList = lapply( 1:10, function(k){
+
+  j = order(df_het$p.het)[k]
+  ensID = df_het$EnsID[j]
+  geneID = df_het$SYMBOL[j]
+
+  main = paste0(geneID, ' (', ensID, ")\nHeterogeneity (p=", format(df_het$p.het[j], digits=2), ')')
+
+  fig1 = plotEffectSize( colnames(B.hat),
+    B.hat[ensID,],
+    se.hat[ensID,],
+    main)
+
+  main = paste0(geneID, ' (', ensID, ")\nHeterogeneity (p=", format(df_het$p.het[j], digits=2), ')')
+
+  fig2 = plotEffectSize( colnames(resMash$result$PosteriorMean),
+    resMash$result$PosteriorMean[ensID,],
+    resMash$result$PosteriorSD[ensID,],
+    main)
+
+  plot_grid(fig1, fig2)
+})
+plot_grid(plotlist = figList, ncol=1)
+```
 

### 2021-11-01 deconvolution/fans4/molec_profiling.Rmd
per cell type

@@
 
+```{r excludeIDs}
+excludeIDs = c('CMC_MSSM_002')
+info = info[! info$Individual.ID %in% excludeIDs,]
+```
+
+```{r exit2, cache=FALSE}
+knitr::knit_exit()
+```
+
 ### Study Summary
@@
 ```
+
+# Per cell type analysus
+```{r preprocess}
+
+for( CT in unique(info$CellType) ){
+
+  idx = info$CellType == CT
+  isexpr = filterByExpr(geneCounts[,idx])
+
+  dge = DGEList( geneCounts[isexpr,] )
+  dge = calcNormFactors(dge)
+
+  form = ~ (1|Gender) + (1|Batch)
+  vobj = voomWithDreamWeights(dge, form, info, plot=TRUE, save.plot=TRUE, BPPARAM=SnowParam(12, progressbar=TRUE))
+
+  form = ~ (1|Gender) + (1|Batch) ~ Batch + Gender + Age.of.Death
+  fit = dream(vobj, form, info, BPPARAM=SnowParam(12, progressbar=TRUE)))
+
+# contrasts
+# L = makeContrastsDream(form, info, contrasts=c(
+#               GABA.Dx = "IGABA.SCZ - IGABA.Control",
+#               GLU.Dx = "IGLU.SCZ - IGLU.Control",
+#               MgAs.Dx = "IMgAs.SCZ - IMgAs.Control",
+#               Olig.Dx = "IOlig.SCZ - IOlig.Control"))
+
+}
+
+```
+
+# Jooint analysis
 
@@
 plot_grid(fig1, fig2 + theme(legend.position="none"), axis="tbr", align="v")
+```
+
+```{r cancor, eval=FALSE}
+
+pairwise_cca = function(fit){
+
+  exprResid = t(residuals(fit))
+
+  df = apply(combn(4, 2), 2, function(x){
+
+    CT1 = levels(info$CellType)[x[1]]
+    CT2 = levels(info$CellType)[x[2]]
+
+    X = exprResid[info$CellType == CT1,]
+    Y = exprResid[info$CellType == CT2,]
+
+    idx = match(rownames(X), rownames(info) )
+    rownames(X) = info$Individual.ID[idx]
+
+    idx = match(rownames(Y), rownames(info) )
+    rownames(Y) = info$Individual.ID[idx]
+
+    tab = table(c(rownames(X), rownames(Y)))
+    ids = names(tab[tab==2])
+
+    X = X[ids,]
+    Y = Y[ids,]
+
+    mx = rowMeans(geneExpr[,info$CellType == CT1])
+    my = rowMeans(geneExpr[,info$CellType == CT2])
+
+    # idx = sample.int(length(ids), length(ids), replace=FALSE)
+
+    X2 = scale(X[,mx>6])
+    Y2 = scale(Y[,my>6])
+
+    X2 = apply(X2, 2, RankNorm)
+    Y2 = apply(Y2, 2, RankNorm)
+
+    incl = seq(1, 0.8*nrow(X2))
+
+    fit = fastcca(X2[incl,],Y2[incl,])
+    X.pred = X2[-incl,] %*% fit$x.coef
+    Y.pred = Y2[-incl,] %*% fit$y.coef
+    rho.fastcca = cor(X.pred[,1], Y.pred[,1])^2
+    rho.fastcca
+
+    data.frame(CT1 = CT1, CT2 = CT2, rho = fit$cramer.V, rho.emp = rho.fastcca)
+  })
+  df = do.call(rbind, df)
+
+  df
+}
+
+dfToMatrix = function(df, V1, V2, value){
+
+  lvls = unique(c(df[,V1], df[,V2]))
+  C = diag(1, length(lvls))
+  C[C!=1] = NA
+  rownames(C) = lvls
+  colnames(C) = lvls
+
+  for(i in 1:nrow(df) ){
+    C[df[,V1][i], df[,V2][i]] = df[[value]][i]
+    C[df[,V2][i], df[,V1][i]] = df[[value]][i]
+  }
+
+  C
+}
+
+library(decorrelate)
+
+library(RNOmni)
+
+devtools::reload("/hpc/users/hoffmg01/build2/decorrelate")
+
+fit = dream(geneExpr, ~ Batch + Gender + Age.of.Death, info)
+fit = eBayes(fit)
+
+C = dfToMatrix(df, "CT1", "CT2", "rho.emp")
+
+plotCorrMatrix(C)
+
+cmd = cmdscale(dist(1-C^2))
+df_cmd = data.frame(CellType = rownames(cmd), data.frame(cmd))
+
+ggplot( df_cmd, aes(X1, X2, color=CellType)) + geom_point() + theme_classic() + theme(aspect.ratio=1)
+
 ```
@@
 
-### Esimate precision weights
+### Estimate precision weights
 ```{r voom}
@@
 
+```{r, eval=FALSE}
+knitr::knit_exit()
+```
+
 Volcano plot colored by cell type
@@
 
-```{r, eval=FALSE}
-knitr::knit_exit()
-```
-
 bMIND can incorporate the prior of cell type variance/covariance.

### 2021-11-01 deconvolution/fans4/molec_profiling.Rmd
per cell type
add per cell type analysis

@@
 
+```{r}
+CODE <- Folder(name = "results", parentId = "syn26433312")
+CODE <- synStore(CODE)
+
+# Store data 
+storeData = function( obj, file, description){
+  saveRDS( obj, file=file)
+
+  nEXP_OBJ = File(file, 
+                  name = description, 
+                  parentId = CODE$properties$id)
+  synStore(nEXP_OBJ, used = ALL_USED_IDs, executed = thisFile)
+}
+```
+
 ### Reading RNA-seq data
@@
 excludeIDs = c('CMC_MSSM_002')
+
+incl = !colnames(geneCounts) %in% rownames(info[info$Individual.ID %in% excludeIDs,])
+geneCounts = geneCounts[,incl]
 info = info[! info$Individual.ID %in% excludeIDs,]
+
+identical(rownames(info), colnames(geneCounts))
 ```
@@
 
+### Save processed data
+```{r save.processed}
+storeData( geneCounts, "geneCounts_merged.RDS", "Merged gene counts for Mol Prof data")
+storeData( info, "info_merged.RDS", "Merge metadata for Mol Prof data")
+```
+
 # Per cell type analysus
-```{r preprocess}
-
-for( CT in unique(info$CellType) ){
+```{r preprocess.CT}
+ALL_USED_IDs= unique(c('syn22321012', 'syn22321013', 'syn22321014', 'syn22321016', 'syn2232101', 'syn26000845','syn25976117', 'syn25976127', 'syn22321013', 'syn22321014', 'syn22321016', 'syn26000850', 'syn25976122', 'syn25976136'))
+
+resByCT = lapply(unique(info$CellType), function(CT){
 
@@
 
-  dge = DGEList( geneCounts[isexpr,] )
+  dge = DGEList( geneCounts[isexpr,idx] )
   dge = calcNormFactors(dge)
@@
   form = ~ (1|Gender) + (1|Batch)
-  vobj = voomWithDreamWeights(dge, form, info, plot=TRUE, save.plot=TRUE, BPPARAM=SnowParam(12, progressbar=TRUE))
-
-  form = ~ (1|Gender) + (1|Batch) ~ Batch + Gender + Age.of.Death
-  fit = dream(vobj, form, info, BPPARAM=SnowParam(12, progressbar=TRUE)))
-
-# contrasts
-# L = makeContrastsDream(form, info, contrasts=c(
-#               GABA.Dx = "IGABA.SCZ - IGABA.Control",
-#               GLU.Dx = "IGLU.SCZ - IGLU.Control",
-#               MgAs.Dx = "IMgAs.SCZ - IMgAs.Control",
-#               Olig.Dx = "IOlig.SCZ - IOlig.Control"))
-
-}
-
+  vobj = voomWithDreamWeights(dge, form, info[idx,], plot=TRUE, save.plot=TRUE, BPPARAM=SnowParam(12, progressbar=TRUE))
+
+  form = ~ Dx + (1|Gender) + (1|Batch) + Batch + Gender + Age.of.Death
+  fit = dream(vobj, form, info[idx,], BPPARAM=SnowParam(12, progressbar=TRUE))
+  fit = eBayes(fit)
+
+  resDE = topTable(fit, coef="DxSCZ", number=Inf)
+  resids = residuals(fit)
+
+  # save results for this cell type
+  storeData( vobj, paste0("voomObj_", CT, ".RDS"), paste0("voom fit for ", CT))
+  storeData( fit, paste0("fit_Dx_", CT, ".RDS"), paste0("dream fit for ", CT))
+  storeData( resDE, paste0("de_Dx_", CT, ".RDS"), paste0("Differential expression within ", CT))
+  storeData( resids, paste0("residuals_", CT, ".RDS"), paste0("Residuals for ", CT))
+})
 ```

### 2021-11-01 deconvolution/fans4/molec_profiling.Rmd
add per cell type analysis

@@
 
-```{r exit2, cache=FALSE}
-knitr::knit_exit()
-```
-
 ### Study Summary
@@
 plot_grid(fig1, fig2 + theme(legend.position="none"), axis="tbr", align="v")
-```
-
-```{r cancor, eval=FALSE}
-
-pairwise_cca = function(fit){
-
-  exprResid = t(residuals(fit))
-
-  df = apply(combn(4, 2), 2, function(x){
-
-    CT1 = levels(info$CellType)[x[1]]
-    CT2 = levels(info$CellType)[x[2]]
-
-    X = exprResid[info$CellType == CT1,]
-    Y = exprResid[info$CellType == CT2,]
-
-    idx = match(rownames(X), rownames(info) )
-    rownames(X) = info$Individual.ID[idx]
-
-    idx = match(rownames(Y), rownames(info) )
-    rownames(Y) = info$Individual.ID[idx]
-
-    tab = table(c(rownames(X), rownames(Y)))
-    ids = names(tab[tab==2])
-
-    X = X[ids,]
-    Y = Y[ids,]
-
-    mx = rowMeans(geneExpr[,info$CellType == CT1])
-    my = rowMeans(geneExpr[,info$CellType == CT2])
-
-    # idx = sample.int(length(ids), length(ids), replace=FALSE)
-
-    X2 = scale(X[,mx>6])
-    Y2 = scale(Y[,my>6])
-
-    X2 = apply(X2, 2, RankNorm)
-    Y2 = apply(Y2, 2, RankNorm)
-
-    incl = seq(1, 0.8*nrow(X2))
-
-    fit = fastcca(X2[incl,],Y2[incl,])
-    X.pred = X2[-incl,] %*% fit$x.coef
-    Y.pred = Y2[-incl,] %*% fit$y.coef
-    rho.fastcca = cor(X.pred[,1], Y.pred[,1])^2
-    rho.fastcca
-
-    data.frame(CT1 = CT1, CT2 = CT2, rho = fit$cramer.V, rho.emp = rho.fastcca)
-  })
-  df = do.call(rbind, df)
-
-  df
-}
-
-dfToMatrix = function(df, V1, V2, value){
-
-  lvls = unique(c(df[,V1], df[,V2]))
-  C = diag(1, length(lvls))
-  C[C!=1] = NA
-  rownames(C) = lvls
-  colnames(C) = lvls
-
-  for(i in 1:nrow(df) ){
-    C[df[,V1][i], df[,V2][i]] = df[[value]][i]
-    C[df[,V2][i], df[,V1][i]] = df[[value]][i]
-  }
-
-  C
-}
-
-library(decorrelate)
-
-library(RNOmni)
-
-devtools::reload("/hpc/users/hoffmg01/build2/decorrelate")
-
-fit = dream(geneExpr, ~ Batch + Gender + Age.of.Death, info)
-fit = eBayes(fit)
-
-C = dfToMatrix(df, "CT1", "CT2", "rho.emp")
-
-plotCorrMatrix(C)
-
-cmd = cmdscale(dist(1-C^2))
-df_cmd = data.frame(CellType = rownames(cmd), data.frame(cmd))
-
-ggplot( df_cmd, aes(X1, X2, color=CellType)) + geom_point() + theme_classic() + theme(aspect.ratio=1)
-
 ```

### 2021-11-01 deconvolution/fans4/molec_profiling.Rmd
add per cell type analysis

@@
 ```{r save.processed}
+ALL_USED_IDs= unique(c('syn22321012', 'syn22321013', 'syn22321014', 'syn22321016', 'syn2232101', 'syn26000845','syn25976117', 'syn25976127', 'syn22321013', 'syn22321014', 'syn22321016', 'syn26000850', 'syn25976122', 'syn25976136'))
+
 storeData( geneCounts, "geneCounts_merged.RDS", "Merged gene counts for Mol Prof data")
@@
 ```{r preprocess.CT}
-ALL_USED_IDs= unique(c('syn22321012', 'syn22321013', 'syn22321014', 'syn22321016', 'syn2232101', 'syn26000845','syn25976117', 'syn25976127', 'syn22321013', 'syn22321014', 'syn22321016', 'syn26000850', 'syn25976122', 'syn25976136'))
 

### 2021-11-01 deconvolution/fans4/molec_profiling.Rmd
add per cell type analysis

@@
 git pull
-export OMP_NUM_THREADS=2
 R

### 2021-11-01 deconvolution/fans4/molec_profiling.Rmd
add per cell type analysis

@@
 # rm -rf molec_profiling_cache/
-ml git pandoc
+ml git pandoc python
 git pull
@@
 
-```{r}
+```{r storeData}
 CODE <- Folder(name = "results", parentId = "syn26433312")
@@
 ```{r corr.celltype}
+library(corrplot)
 C = get_pairwise_sharing( resMash )

### 2022-03-30 deconvolution/fans4/molec_profiling.Rmd
update paths

@@
 
-# file_SampleAttributesDS = "/sc/arion/projects/CommonMind/data/resources/GTEx/GTEx_Analysis_v8_Annotations_SampleAttributesDS.txt"
-file_SampleAttributesDS = synGet('syn22321103')$path
-
-# file_SubjectPhenotypesDS= "/sc/arion/projects/CommonMind/data/resources/GTEx/GTEx_Analysis_v8_Annotations_SubjectPhenotypesDS.txt"
-file_SubjectPhenotypesDS = synGet('syn22321104')$path
-
-# file_counts = "/sc/arion/projects/CommonMind/data/resources/GTEx/GTEx_Analysis_2017-06-05_v8_RNASeQCv1.1.9_gene_reads.gct.gz"
-file_counts = synGet('syn22321107')$path
+file_SampleAttributesDS = "/sc/arion/projects/CommonMind/data/resources/GTEx/GTEx_Analysis_v8_Annotations_SampleAttributesDS.txt"
+# file_SampleAttributesDS = synGet('syn22321103')$path
+
+file_SubjectPhenotypesDS= "/sc/arion/projects/CommonMind/data/resources/GTEx/GTEx_Analysis_v8_Annotations_SubjectPhenotypesDS.txt"
+# file_SubjectPhenotypesDS = synGet('syn22321104')$path
+
+file_counts = "/sc/arion/projects/CommonMind/data/resources/GTEx/GTEx_Analysis_2017-06-05_v8_RNASeQCv1.1.9_gene_reads.gct.gz"
+# file_counts = synGet('syn22321107')$path
 

### 2022-03-30 deconvolution/fans4/molec_profiling.Rmd
update paths

@@
 
-# file= '/sc/arion/projects/CommonMind/data/resources/GTEx/GTEx_Tissue_colors.csv'
-file = synGet('syn22321106')$path
-gtexColors = read.csv( file, header=TRUE)
+file= '/sc/arion/projects/CommonMind/data/resources/GTEx/GTEx_Tissue_colors.tsv'
+# file = synGet('syn22321106')$path
+gtexColors = read.table( file, header=TRUE, sep="\t")
 ```
